### 2016-04-15 NatBiotech/CC-suppl_NatBiotech.tex
Include previous releases for different publications
On branch master
	new file:   MolSysBio/CC-suppl_molsysbio.docx
	new file:   MolSysBio/molsysbio_CC.docx
	new file:   NatBiotech/CC-suppl_NatBiotech.pdf
	new file:   NatBiotech/CC-suppl_NatBiotech.tex
	new file:   NatBiotech/CC_NatBiotech.pdf
	new file:   NatBiotech/CC_NatBiotech.tex
	new file:   NatBiotech/lettertoeditor_NatBiotech.docx
	new file:   Nature/CC-suppl_sub.pdf
	new file:   Nature/CC_sub.pdf
	new file:   Nature/lettertoeditor2.docx
	new file:   Science/lettertoeditor_sub.docx
	new file:   Science/science_CC.pdf
	new file:   Scie

% !TEX TS-program = pdflatex
% !TEX encoding = UTF-8 Unicode

%% Template for a preprint Letter or Article for submission
%% to the journal Nature.
%% Written by Peter Czoschke, 26 February 2004
%%
\documentclass{nature} % Commented includegraphics trick in /usr/local/texlive/2014/texmf-dist/tex/latex/nature/nature.cls
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage[usenames,dvipsnames,table]{xcolor}
\usepackage[nolists,tablesfirst,nomarkers]{endfloat}

\newcommand{\CC}[0]{\emph{cmplxcruncher}}
\newcommand{\task}[1]{\texttt{\bfseries\scshape\textcolor{MidnightBlue}{#1}}}
\newcommand{\un}[1]{\operatorname{#1}}
\newcommand{\unclassrate}[0]{\un{kpb/s/core}}

%% Roman footnote numbering style to not collide with bibliographic cites: 
\renewcommand{\thefootnote}{\Roman{footnote}}

%% Change floats name for supplemental material
\renewcommand{\figurename}{Supplementary Figure}
\renewcommand{\tablename}{Supplementary Table}

%% make sure you have the nature.cls and naturemag.bst files where
%% LaTeX can find them
\bibliographystyle{naturemag}

%%% HYPERREF
\usepackage[pdftex,
	pdftitle={Supplemental material of ``Microbiota: are you sick?''},
	pdfauthor={Jose Manuel Marti, Daniel M. Martinez, Manuel Peña, Cesar Gracia, Amparo Latorre, Andres Moya, Carlos P. Garay},
	pdfsubject={Nature Biotechnology},
	pdfkeywords={Supplemental material},
	bookmarksopen,
	colorlinks=true,
	urlcolor={red},
	bookmarks=true]{hyperref}

%% Title
\title{Supplemental material of ``Microbiota: are you sick?''}

%% Notice placement of commas and superscripts and use of &
%% in the author list
\author{Jose Manuel Mart\'i$^{1}$, Daniel M. Mart\'inez$^{1,2}$, Manuel Pe\~na$^{1}$, C\'esar Gracia$^{1}$, Amparo Latorre$^{2,3,4}$, Andr\'es Moya$^{2,3,4}$ \& Carlos P. Garay$^1$}

\begin{document}

\maketitle

\begin{affiliations}
 \item Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain.
 \item FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain.
 \item Cavanilles Institute of Biodiversity and Evolutionary Biology, Univ. de Valencia, 46980, Spain. 
 \item CIBER en Epidemiolog\'ia y Salud P\'ublica (CIBEResp), Madrid, Spain.
\end{affiliations}

\section{Model} \label{sec:model}

We model the microbial abundances across time along the lines of Blumm \textit{et al.}\cite{ranking}. The dynamics of taxon relative abundances is described by the Langevin equation: 
\begin{equation}
\dot{x_i} = F_i \cdot x_i^\alpha + V \cdot x_i^\beta \xi_i(t) - \phi(t) \cdot x_i,
\end{equation}
where F$_i$ captures the fitness of the taxon i, V corresponds to the noise amplitude and $\xi_i$(t) is a Gaussian random noise with zero mean  $<\xi_i(t)>$ = 0 and variance uncorrelated in time, $<\xi_i(t) \xi_i(t')>$ =  $\delta(t'-t)$, . The function $\phi(t)$ ensures the normalization at all times, $\sum x_i(t) = 1$, and corresponds to $\phi(t) = \sum F_i x_i^\alpha + \sum V x_i^\beta \xi_i(t)$ .
The temporal evolution of the probability that a taxon i has a relative abundance $x_i(t)$, P(x$_i$,t), is determined by the Fokker-Planck equation:
\begin{equation}
\frac{\partial P}{\partial t} = - \frac{\partial}{\partial x_i}  [(F_i \cdot x_i^\alpha - \phi(t) \cdot x_i ) \cdot P]+ \frac{1}{2} \frac{\partial^2}{\partial x_i^2} (V^2 \cdot x_i^{2\beta}\cdot P).
\end{equation}
The microbiota evolves towards a steady-state with a time-independent probability depending on the values of $\alpha$, $\beta$, F$_i$ and V. For $\alpha<1$ (otherwise, systems are always unstable), the steady-state probability may be localized in a region around a preferred value or broadly distributed over a wide range, depending on whether the fitness F$_i$  dominates or is overwhelmed by the noise amplitude V. The steady-state solution of the Fokker-Planck equation is given by:
\begin{eqnarray}
P_0 (x_i) &=& C_{ne}(\alpha,\beta,F_i,V)  \cdot x_i^{-2\beta}  \cdot \exp\Big[\frac{2F_i}{V^2}\frac{x_i^{1+\alpha-2\beta}}{1+\alpha-2\beta}-\frac{\phi_0}{V^2}\frac{x_i^{2-2\beta}}{1-\beta}\Big] \quad \textrm{if} \quad  2\beta \ne 1+\alpha, \notag \\
P_0 (x_i) &=& C_e(\alpha,\beta,F_i,V)  \cdot x_i^{\frac{2F_i}{V^2} -2\beta}  \cdot \exp\Big[\frac{\phi_0}{V^2}\frac{x_i^{2-2\beta}}{1-\beta}\Big] \quad \textrm{if} \quad  2\beta = 1+\alpha,
\notag
\end{eqnarray}
where $\phi_0 = (\sum_i F_i^{1/(1-\alpha)})^{1-\alpha}$ and C$_{ne}$ and C$_{e}$ are integrals that should be solved numerically for the parameters of interest. The ordered phase happens when the solution has a maximum in the physical interval ($0<x_i<1$). For larger V, the transition to a disordered phase happens when the maximum shifts to the unphysical region $x_i<0$, which sets the phase transition region V($\alpha,\beta,F_i$). The phase transition region can be calculated analytically in particular cases:
\begin{eqnarray}
F_i^2 &=& 4 \beta \phi_0 V^2\quad \textrm{if} \quad  \beta = \alpha \neq 1, \notag \\
F_i &=& \beta V^2\quad \textrm{if} \quad  2\beta = 1+\alpha,
\notag
\end{eqnarray}
where the first case, simplifies to $F= 3 V^2$ if $\beta = 0.75$ and the fitness of this taxon dominates in $\phi_0$. 
In many physical systems (Brownian motion is the classical example), the two terms of the Langevin equation are related.  The \emph{fluctuation--dissipation theorem} states a general relationship between the response to an external disturbance and the internal fluctuations of the system\cite{FD}. The theorem can be used as the basic formula to derive the fitness from the analysis of fluctuations 
of the microbiota, assuming that it is in equilibrium (the ordered phase).

\section{Selection and Methods}

\subsection{Sample selection}
We have chosen studies about relevant pathologies containing metagenomic sequencing time data series of bacterial populations from humans in different healthy and non-healthy states. We have selected only those individuals who had three or more time points of data available in databases. Metadata of each study is provided in Supplementary Tables \ref{tab:diet} to \ref{tab:LEA}. All used 16S rRNA gene sequencing except for the study of the discordant kwashiorkor twins\cite{kwashiorkor} (see Supplementary Tables \ref{tab:DH} and \ref{tab:DK}) where shotgun metagenomic sequencing (SMS) and 16S rRNA were used. In the latter case we selected to work with SMS data to show that our method is valid regardless of the source of taxonomic information. Each one of the datasets was treated as follows:

\subsection{16rRNA sequences processing}
Reads from the selected studies were first quality filtered using the FastX toolkit\cite{FASTX}, allowing only those reads which had more than 25 of quality along the 75\% of the complete sequence. 16S rRNA reads were then clustered at 97\% nucleotide sequence identity (97\% ID) into operational taxonomic units (OTUs) using QIIME package software\cite{QIIME} (version 1.8) We followed open reference OTU picking workflow in all cases. The clustering method used was uclust, and the OTUs were matched against Silva database\cite{SILVA} (version 111, July 2012) and were assigned to taxonomy with an uclust-based consensus taxonomy assigner. The parameters used in this step were: similarity 0.97, prefilter percent id 0.6, max accepts 20, max rejects 500.

\subsection{Metagenomic sequences processing}
Metagenomic shotgun (and 16S too) sequences were analyzed with LMAT (Livermore Metagenomics Analysis Toolkit) software package\cite{LMAT1, LMAT2} (version 1.2.4, with Feb'15 release of data base \emph{LMAT-Grand}). LMAT was run using a Bull shared-memory node belonging to the team's HPC (high performance computing) cluster. It is equipped with 32 cores (64 threads available using Intel Hyper-threading technology) as it has 2 Haswell-based Xeons, the E5-2698v3@2.3 GHz, sharing half a tebibyte (0.5 TiB, that is, 512 gibibytes) of DRAM memory. This node is also provided with a card PCIe SSD as NVRAM, the P420m HHHL, with 1.4 TB, and 750000 reading IOPS, 4 KB, achieving 3.3 GB/s, which Micron kindly issued free of charge, as a sample for testing purposes. The computing node was supplied with a RAID-0 (striping) scratch disk area. We used the ``Grand'' database\footnote{In this context, ``Grand''\cite{LMAT2} refers to a huge database that contains k-mers from all viral, prokaryote, fungal and protist genomes present in the NCBI database, plus Human reference genome (hg19), plus GenBank Human, plus the 1000 Human Genomes Project (HGP). This represent about 31.75 billion k-mers occupying 457.62 GB.}, release Feb'15, provided by the LMAT team\cite{LMAT2}. Previously to any calculation, the full database was loaded in the NVRAM. With this configuration the observed LMAT sustained sequence classification rate was $20\unclassrate$. Finally, it is worth mentioning that a complete set of Python scripts have been developed as back-end and front-end of the LMAT pipeline in order to manage the added complexity of time series analysis.

\subsection{Taxa level selection}
We selected genus as taxonomic level for the subsequent steps of our work. To ensure that, between adjacent taxonomic levels, there were not crucial differences which could still be of relevance after standardization (see Supplementary Section \ref{sec:stan}), we tested two different datasets. In the former, the antibiotics study\cite{antibiotic} with 16S data, we tested the differences between genus and family levels. The latter dataset tested was the kwashiorkor twins study\cite{kwashiorkor} for both genus and species taxonomic levels. The Supplementary Figures \ref{fig:taxlev1} (overview) and \ref{fig:taxlev2} (detail) plot the comparison between studies (and so, 16S and SMS) and between adjacent taxonomic levels.

\begin{figure} 
  \includegraphics[width=0.5\textwidth]{results/taxalevel/sum_raw_16S.pdf}
  \includegraphics[width=0.5\textwidth]{results/taxalevel/sum_sta_16S.pdf}
  \includegraphics[width=0.5\textwidth]{results/taxalevel/sum_raw_SMS.pdf} 
  \includegraphics[width=0.5\textwidth]{results/taxalevel/sum_sta_SMS.pdf}
\caption{Overview of comparison of different approaches based on adjacent taxonomic levels using plots in the Taylor-parameters space. For 16S (former row of subfigures), the levels are family vs. genus, whereas for SMS (latter row of subfigures) levels are genus vs. species. The left column shows the raw results and the right column plots the standardized results (see Section \ref{sec:stan})}
\label{fig:taxlev1}
\end{figure}

\begin{figure} 
  \includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_fam_16S.pdf}
  \includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_gen_16S.pdf}
  \includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_gen_SMS.pdf}
  \includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_spc_SMS.pdf}
\caption{Detail of comparison of different approaches based on adjacent taxonomic levels using plots of X-weighted power-law fits (see Section \ref{sec:X-w}). The former row of subfigures shows examples for 16S, whereas the latter row of subfigures plots examples for SMS. The left column shows results for the superior taxonomic level (family for 16S, genus for SMS), while the right column shows results for the inferior level (genus for 16S, specie for SMS).}
\label{fig:taxlev2}
\end{figure}

\begin{table} 
  \begin{center}
    \begin{tabular}{ccccccc}
	    \hline
		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
		\hline
		A&$0.26 \pm 0.05$&$0.826 \pm 0.025$&$0.918$&&$3.1 \pm 0.9$&$1.2 \pm 0.6$\\
		A&$0.32 \pm 0.06$&$0.857 \pm 0.025$&$0.924$&&$4.4 \pm 1.1$&$2.0 \pm 0.6$\\
		A&$0.194 \pm 0.033$&$0.813 \pm 0.024$&$0.918$&&$1.9 \pm 0.6$&$0.9 \pm 0.6$\\
		A&$0.24 \pm 0.04$&$0.824 \pm 0.020$&$0.924$&&$2.7 \pm 0.7$&$1.2 \pm 0.5$\\
		A&$0.34 \pm 0.06$&$0.855 \pm 0.024$&$0.931$&&$4.7 \pm 1.1$&$1.9 \pm 0.6$\\
		A&$0.30 \pm 0.05$&$0.847 \pm 0.022$&$0.921$&&$3.9 \pm 1.0$&$1.7 \pm 0.5$\\
		A&$0.133 \pm 0.021$&$0.784 \pm 0.023$&$0.916$&&$0.7 \pm 0.4$&$0.2 \pm 0.6$\\
		A&$0.25 \pm 0.04$&$0.831 \pm 0.024$&$0.929$&&$3.0 \pm 0.8$&$1.4 \pm 0.6$\\
		\hline
		P&$0.23 \pm 0.05$&$0.804 \pm 0.035$&$0.885$&&$2.6 \pm 0.9$&$0.7 \pm 0.8$\\
		P&$0.097 \pm 0.018$&$0.705 \pm 0.031$&$0.891$&&$0.03 \pm 0.34$&$-1.6 \pm 0.7$\\
		P&$0.037 \pm 0.006$&$0.642 \pm 0.025$&$0.881$&&$-1.12 \pm 0.11$&$-3.1 \pm 0.6$\\
		P&$0.118 \pm 0.019$&$0.723 \pm 0.025$&$0.895$&&$0.4 \pm 0.4$&$-1.2 \pm 0.6$\\
		P&$0.17 \pm 0.04$&$0.78 \pm 0.04$&$0.842$&&$1.5 \pm 0.7$&$0.1 \pm 0.9$\\
		P&$0.123 \pm 0.020$&$0.757 \pm 0.026$&$0.914$&&$0.5 \pm 0.4$&$-0.4 \pm 0.6$\\
		P&$0.19 \pm 0.05$&$0.77 \pm 0.04$&$0.871$&&$1.8 \pm 0.9$&$-0.0 \pm 0.9$\\
		P&$0.121 \pm 0.020$&$0.736 \pm 0.027$&$0.921$&&$0.5 \pm 0.4$&$-0.9 \pm 0.6$\\
		P&$0.187 \pm 0.034$&$0.771 \pm 0.030$&$0.908$&&$1.8 \pm 0.7$&$-0.1 \pm 0.7$\\
		P&$0.097 \pm 0.015$&$0.735 \pm 0.025$&$0.922$&&$0.05 \pm 0.28$&$-0.9 \pm 0.6$\\
	   \hline
	   \hline
    \end{tabular}
  \end{center}
  \caption{Taylor parameters. Individuals with either animal-based (A) or plant-based (P) diets\cite{diet}. Previous to diet,  the population sampled is described by $\bar{V} = 0.09 \pm 0.05, \bar{\beta} = 0.77 \pm 0.04$.}
  \label{tab:diet}
\end{table}

 \begin{table} 
  \begin{center}
    \begin{tabular}{ccccccc}
	    \hline
		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
		\hline
		Ab&$0.35 \pm 0.07$&$0.81 \pm 0.04$&$0.925$&&$4.3 \pm 1.4$&$1.3 \pm 0.9$\\
		Ab&$0.41 \pm 0.09$&$0.82 \pm 0.04$&$0.908$&&$5.6 \pm 1.8$&$1.6 \pm 0.9$\\
		Ab&$0.23 \pm 0.04$&$0.770 \pm 0.031$&$0.920$&&$2.1 \pm 0.8$&$0.5 \pm 0.7$\\
		Ab&$0.165 \pm 0.029$&$0.738 \pm 0.031$&$0.928$&&$0.9 \pm 0.6$&$-0.3 \pm 0.7$\\
		Ab&$0.34 \pm 0.06$&$0.812 \pm 0.032$&$0.936$&&$4.1 \pm 1.2$&$1.5 \pm 0.7$\\
		Ab&$0.26 \pm 0.05$&$0.798 \pm 0.033$&$0.931$&&$2.8 \pm 0.9$&$1.1 \pm 0.8$\\
	    \hline
	    \hline
    \end{tabular}
  \end{center}
  \caption{Taylor parameters for individuals taking antibiotics\cite{antibiotic}. Prior to antibiotics intake, the population sampled is described by $\bar{V} = 0.12 \pm 0.05, \bar{\beta} = 0.75 \pm 0.04$.}
  \label{tab:antibiotics}
\end{table}

\begin{table} 
  \begin{center}
    \begin{tabular}{ccccccc}
	    \hline
		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
		\hline
		IBS&$0.204 \pm 0.034$&$0.739 \pm 0.029$&$0.916$&&$7.6 \pm 3.7$&$1.9 \pm 1.2$\\
		IBS&$0.35 \pm 0.05$&$0.793 \pm 0.023$&$0.935$&&$23.1 \pm 5.9$&$4.0 \pm 0.9$\\
	     \hline
	     \hline
    \end{tabular}
  \end{center}
  \caption{Taylor parameters for persons diagnosed with irritable bowel syndrome (IBS)\cite{IBS}. Healthy individuals sampled in this study are characterized by $\bar{V} = 0.134 \pm 0.009, \bar{\beta} = 0.691 \pm 0.025$.}
  \label{tab:IBS}
\end{table}

 \begin{table} 
  \begin{center}
    \begin{tabular}{ccccccc}
	    \hline
		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
		\hline
		DH&$0.27 \pm 0.04$&$0.835 \pm 0.016$&$0.925$&&$0.2 \pm 0.4$&$-1.0 \pm 0.6$\\
		DH&$0.36 \pm 0.06$&$0.858 \pm 0.015$&$0.929$&&$1.1 \pm 0.6$&$-0.2 \pm 0.5$\\
		DH&$0.35 \pm 0.06$&$0.859 \pm 0.014$&$0.926$&&$1.0 \pm 0.5$&$-0.1 \pm 0.5$\\
		DH&$0.25 \pm 0.04$&$0.829 \pm 0.014$&$0.911$&&$0.0 \pm 0.4$&$-1.2 \pm 0.5$\\
		DH&$0.30 \pm 0.05$&$0.844 \pm 0.014$&$0.920$&&$0.5 \pm 0.4$&$-0.7 \pm 0.5$\\
		DH&$0.29 \pm 0.05$&$0.850 \pm 0.016$&$0.915$&&$0.4 \pm 0.5$&$-0.5 \pm 0.5$\\
		DH&$0.28 \pm 0.05$&$0.848 \pm 0.016$&$0.921$&&$0.3 \pm 0.5$&$-0.5 \pm 0.6$\\
		DH&$0.35 \pm 0.07$&$0.861 \pm 0.017$&$0.918$&&$0.9 \pm 0.6$&$-0.0 \pm 0.6$\\
		DH&$0.31 \pm 0.04$&$0.833 \pm 0.012$&$0.916$&&$0.6 \pm 0.4$&$-1.1 \pm 0.4$\\
		DH&$0.33 \pm 0.05$&$0.843 \pm 0.013$&$0.925$&&$0.8 \pm 0.5$&$-0.7 \pm 0.5$\\
		DH&$0.31 \pm 0.05$&$0.852 \pm 0.014$&$0.925$&&$0.6 \pm 0.5$&$-0.4 \pm 0.5$\\
		DH&$0.31 \pm 0.05$&$0.853 \pm 0.015$&$0.930$&&$0.6 \pm 0.5$&$-0.4 \pm 0.5$\\
		DH&$0.203 \pm 0.033$&$0.815 \pm 0.015$&$0.907$&&$-0.44 \pm 0.32$&$-1.7 \pm 0.5$\\
		\hline
    \end{tabular}
  \end{center}
  \caption{Taylor parameters for the healthy subject of the discordant twins\cite{kwashiorkor}. This table continues in Supplementary Table \ref{tab:DK}. The population of healthy twins is characterized by $\bar{V} = 0.25 \pm 0.10, \bar{\beta} = 0.863 \pm 0.028$.}
  \label{tab:DH}
\end{table}

\begin{table} 
  \begin{center}
    \begin{tabular}{ccccccc}
	    \hline
		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
		\hline
		DK&$0.40 \pm 0.07$&$0.859 \pm 0.017$&$0.926$&&$1.5 \pm 0.7$&$-0.1 \pm 0.6$\\
		DK&$0.44 \pm 0.08$&$0.868 \pm 0.016$&$0.919$&&$1.8 \pm 0.8$&$0.2 \pm 0.6$\\
		DK&$0.196 \pm 0.031$&$0.819 \pm 0.014$&$0.916$&&$-0.50 \pm 0.30$&$-1.5 \pm 0.5$\\
		DK&$0.160 \pm 0.026$&$0.798 \pm 0.015$&$0.904$&&$-0.85 \pm 0.25$&$-2.3 \pm 0.5$\\
		DK&$0.30 \pm 0.05$&$0.845 \pm 0.014$&$0.924$&&$0.5 \pm 0.4$&$-0.6 \pm 0.5$\\
		DK&$0.23 \pm 0.04$&$0.834 \pm 0.014$&$0.908$&&$-0.1 \pm 0.4$&$-1.0 \pm 0.5$\\
		DK&$0.27 \pm 0.05$&$0.848 \pm 0.015$&$0.930$&&$0.2 \pm 0.4$&$-0.5 \pm 0.5$\\
		DK&$0.35 \pm 0.07$&$0.860 \pm 0.019$&$0.916$&&$1.0 \pm 0.7$&$-0.1 \pm 0.7$\\
		DK&$0.34 \pm 0.05$&$0.835 \pm 0.012$&$0.917$&&$0.9 \pm 0.5$&$-1.0 \pm 0.4$\\
		DK&$0.25 \pm 0.04$&$0.831 \pm 0.012$&$0.912$&&$0.0 \pm 0.4$&$-1.1 \pm 0.4$\\
		DK&$0.36 \pm 0.06$&$0.858 \pm 0.013$&$0.918$&&$1.1 \pm 0.5$&$-0.2 \pm 0.5$\\
		DK&$0.31 \pm 0.06$&$0.851 \pm 0.016$&$0.924$&&$0.6 \pm 0.6$&$-0.4 \pm 0.6$\\
		DK&$0.149 \pm 0.022$&$0.799 \pm 0.013$&$0.905$&&$-0.96 \pm 0.22$&$-2.2 \pm 0.5$\\
	    \hline
	    \hline
    \end{tabular}
  \end{center}
  \caption{Taylor parameters for the kwashiorkor part of the discordant twins\cite{kwashiorkor}. This is a continuation of Supplementary Table \ref{tab:DH}. The population of healthy twins is characterized by $\bar{V} = 0.25 \pm 0.10, \bar{\beta} = 0.863 \pm 0.028$.}
  \label{tab:DK}
\end{table}

\begin{table} 
  \begin{center}
    \begin{tabular}{ccccccc}
	    \hline
		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
		\hline
		OW&$0.59 \pm 0.12$&$0.894 \pm 0.034$&$0.920$&&$6.6 \pm 2.0$&$2.6 \pm 1.0$\\
		OW&$0.22 \pm 0.04$&$0.830 \pm 0.030$&$0.904$&&$0.5 \pm 0.6$&$0.7 \pm 0.9$\\
		\hline
		OBI&$0.28 \pm 0.04$&$0.855 \pm 0.022$&$0.958$&&$1.5 \pm 0.6$&$1.4 \pm 0.6$\\
		OBI&$0.33 \pm 0.07$&$0.870 \pm 0.031$&$0.916$&&$2.4 \pm 1.1$&$1.9 \pm 0.9$\\
		\hline
		OBII&$0.223 \pm 0.032$&$0.823 \pm 0.023$&$0.938$&&$0.6 \pm 0.5$&$0.5 \pm 0.7$\\
		OBII&$0.208 \pm 0.029$&$0.844 \pm 0.022$&$0.935$&&$0.4 \pm 0.5$&$1.1 \pm 0.7$\\
		\hline
		OBIII&$0.34 \pm 0.05$&$0.855 \pm 0.025$&$0.943$&&$2.5 \pm 0.9$&$1.4 \pm 0.7$\\
		OBIII&$0.26 \pm 0.04$&$0.845 \pm 0.026$&$0.954$&&$1.1 \pm 0.7$&$1.2 \pm 0.8$\\
		OBIII&$0.33 \pm 0.06$&$0.870 \pm 0.027$&$0.908$&&$2.4 \pm 1.0$&$1.9 \pm 0.8$\\
		OBIII&$0.200 \pm 0.026$&$0.843 \pm 0.020$&$0.949$&&$0.2 \pm 0.4$&$1.1 \pm 0.6$\\
		OBIII&$0.30 \pm 0.05$&$0.846 \pm 0.026$&$0.929$&&$1.9 \pm 0.8$&$1.2 \pm 0.7$\\
		OBIII&$0.176 \pm 0.029$&$0.826 \pm 0.026$&$0.894$&&$-0.2 \pm 0.5$&$0.6 \pm 0.8$\\
		OBIII&$0.30 \pm 0.06$&$0.841 \pm 0.031$&$0.896$&&$1.8 \pm 0.9$&$1.0 \pm 0.9$\\
		OBIII&$0.28 \pm 0.04$&$0.857 \pm 0.025$&$0.941$&&$1.5 \pm 0.7$&$1.5 \pm 0.7$\\
		OBIII&$0.122 \pm 0.018$&$0.822 \pm 0.024$&$0.930$&&$-1.05 \pm 0.30$&$0.5 \pm 0.7$\\
		\hline
		OBIIId&$0.47 \pm 0.08$&$0.872 \pm 0.023$&$0.945$&&$4.7 \pm 1.3$&$1.9 \pm 0.7$\\
		OBIIId&$0.38 \pm 0.06$&$0.846 \pm 0.023$&$0.951$&&$3.2 \pm 1.0$&$1.2 \pm 0.7$\\
		OBIIId&$0.36 \pm 0.06$&$0.842 \pm 0.022$&$0.954$&&$2.9 \pm 0.9$&$1.1 \pm 0.6$\\
	    \hline
	    \hline
    \end{tabular}
  \end{center}
  \caption{Taylor parameters for individuals with different degrees of overweight and obesity\cite{LEA}. Healthy people in this study, whom were not obese, are characterized by $\bar{V} = 0.19 \pm 0.06, \bar{\beta} = 0.806 \pm 0.034$.}
  \label{tab:LEA}
\end{table}

\section{Temporal evolution of model parameters}

We have studied the time dependence of the variability $V$ and power law index $\beta$ (see Supplementary Section \ref{sec:model}) by using a sliding window approach. The total number of time points are divided in subsets of five points, where next subset is defined by adding next time sampling and by eliminating the earliest one. Both parameters were calculated for each subset against the average time lapse. Supplementary Figure \ref{fig:tempevo1} shows the variability  $V$ as a function of time for the largest sampling: two individuals in the Caporaso's study\cite{moving} corresponding to the gut microbiota of a male (upper plot) and a female (lower plot). Supplementary Figure \ref{fig:tempevo2} shows the time evolution of $V$ for patient P2 of the IBS study\cite{IBS} (upper plot) and patient D in the antibiotics study\cite{antibiotic} (lower plot).

\begin{figure}
	\includegraphics[width=1.0\textwidth]{results/sliwin/male_mov.pdf}
	\hspace*{3mm}\includegraphics[width=0.448\textwidth]{results/sliwin/female_mov.pdf}
\caption{$V$ as a function of time for the two individuals in the Caporaso's study\cite{moving}: samples of gut microbiome of a male (upper plot) and a female (lower plot). Both samples show changes in the variability V with quasi--periodic behavior peaked at about 10 days. Variability grows more for the gut microbiota of the male and share a minimal value around 0.1 with the gut microbiota of the female.}
\label{fig:tempevo1}
\end{figure}

\begin{figure}
	\centering 
 	\includegraphics[width=0.8\textwidth]{results/sliwin/patP2_IBS.pdf}
  	\includegraphics[width=1.0\textwidth]{results/sliwin/patD_antibio.pdf} 
\caption{$V$ as a function of time for patient P2 of the IBS study\cite{IBS} (upper plot) and patient D in the antibiotics study\cite{antibiotic} (lower plot). The variability 
of the gut microbiota of P2 decreases from above 0.3 to below 0.2, showing a slow tendency to increase the order of the system.  Antibiotic intake leaks to a quick increase of variability which lasts for a few days to recover ordering. The second antibiotic treatment shows some memory (lower increase of variability) with a slower recovery. NOTE: The blue vertical lines in the lower plot are showing the periods of antibiotic treatment.}
\label{fig:tempevo2}
\end{figure}

\section{Engineered Software}

A complete software framework, named 'ComplexCruncher', has been engineered to support the analysis of the dynamics of ranking processes in complex systems. Although the software was devised with a clear bias towards metagenomics, it is general enough to be able to cope with a ranking process in any complex system. Implemented in Python using well-known open-source community software, the software solution is composed of two parts that can be used together or apart: a web-based graphic front-end connected to a database, and a computing kernel. Used together, this software enables other users to reproduce our results easily and, furthermore, upload and analyse their own data or experiment with the preloaded metagenomics data sets. The next sections of this supplementary material deal with these two software pieces in detail.

\section{Web-based front-end}

`ComplexCruncher WebPortal' (CCWebPortal) is a web platform designed to allow the user to interact with a data repository of selected and well-documented metagenomics data sources. Through a few simple steps, the user can perform advanced searches on the complete set of records in the metagenomics repository.  The web application provides advanced filters that allow the user to reduce the search to a small set of interest. After this first step, the user can refine the search and discard those records that do not meet certain requirements.

The web application allows calculations to be done directly by the stable release of the \CC\ computing kernel. At the end of the calculations, the results are displayed to the user on the same browser which runs the web application. Then, the user can interact over the series of generated graphics thus allowing flexible comparison among them. In addition, CCWebPortal enables direct download of generated data (plots, spreadsheets, etc). The web application generates a report file summarizing all the results in PDF format. If the user has login permissions, CCWebPortal enables the option of insert new database records in addition to editing and deleting existing ones.

CCWebPortal is a web application that runs on current versions of many browsers. Additional software is not needed and only requires javaScript to be enabled on the browser to run applications. CCWebPortal is, at least, compatible with the following browsers: Internet Explorer 6+, Firefox 3.6+ (PC, Mac), Safari 4+, Chrome 10+ and Opera 11+ (PC, Mac).

CCWebPortal is implemented following the client$-$server distributed programming model. The user connects to the website and a set of javaScript files are downloaded and executed on the local machine. Subsequently, the client application connects to a remote server that enables the execution of calculations and transactions through a centralized database management system. The web application is deployed on the client side using ExtJs libraries. ExtJs is a javaScript development framework to implement interactive Web applications using AJAX, DHTML and DOM technologies. ExtJS has been widely adopted by large technology projects worldwide. Its success lies in its ability to design and implement web interfaces with a clear and flexible functionality for the end user. ExtJs is used under the GNU license GPL v3.

The CCWebPortal website communicates with a remote server that hosts a MySQL database. The database runs on a MyISAM engine that optimizes the speed of access to records in a transactional web-based environment. A set of relational tables allows the structuring of the metagenomics repository to establish relationships between records. Thus the search and information threshing is optimized for queries launched from the client interface. Access to the database on the server is implemented through Django framework. Django is an open-source framework written in Python using the model-view-controller (MVC) architectural pattern for implementing user interfaces, which eases the development and updating of code. This framework is used to implement web applications on the server side and allows flexible and modular developments and access to various types of database managers. Django runs as a web server, awaiting requests from the client CCWebPortal application. Once a user is authorized to perform a given operation, Django interacts with the database serving queries to the client application.

\section{Computing kernel}

The effective data analysis has been performed with a Python tool developed from scratch following the OOP\footnote{Object Oriented Programming, but OOP also stands for Object Oriented Python, which is applicable here too.} paradigm, called \CC. This software is the back-end of the website described in the previous subsection. However, it could be run as an independent piece of software since it is built as a Python package provided with a command-line front-end (\CC.py). Once installed, the tool can be run interactively (using the flag \texttt{-i}/\texttt{--interactive}) but also in automatic mode (flag \texttt{-a}/\texttt{--automatic} followed by the desired graphic format, e.g., \texttt{pdf}, \texttt{png}, \texttt{svg}, \texttt{ps}), which uses parallel computation to speed up the analysis of several data sources.

The external mandatory dependencies of \CC\ are with Python packages \emph{pandas}, \emph{matplotlib}, \emph{numpy}, \emph{scipy} and \emph{xlrd}, while the recommended ones are \emph{pylatex} and \emph{uncertainties}. The PEP8 compliant code of \CC\ could be executed under a Python 2.7 interpret, but a 3.X interpret is strongly advised. The complete distribution of \CC\ has the code signature shown in Supplementary Table \ref{tab:cloc}, obtained through the \emph{CLOC}\footnote{\emph{Count Lines of Code} is a perl script that counts blank lines, comment lines, and physical lines of source code.} tool release 1.62. Supplementary Figure \ref{fig:CC} shows the code layout of the complete distribution of \CC.

\begin{table}
  \begin{center}
    \begin{tabular}{lrrrrcrcr}
    \hline
	Language & files & blank & comment & code & x & scale & = & 3rd gen. equiv \\
	\hline
	Python & 13 & 384 & 666 & 4222 &x& 4.20 &=& 17732.40 \\
	R & 1 & 11 & 38 & 192 &x& 3.00 &=& 576.00 \\	
	\hline
	SUM: & 14 & 395 & 704 & 4414 &x& 4.15 &=& 18308.40 \\
	\hline
    \end{tabular}
  \end{center}
  \caption{Line counting for the complete distribution of \CC\ by \emph{CLOC}}
  \label{tab:cloc}
\end{table}

\begin{figure}
  \centering
	\[ \text{\task{ComplexCruncher}}
	\begin{cases}
		\text{\task{cmplxcruncher.py}} \\
		\\ 
		\text{\task{cmplxcruncher}}
		\begin{cases}
			\text{\task{\_\_init\_\_.py}} \\ 
			\\ 
			\text{\task{models.py}} \\
			\\ 
			\text{\task{plots.py}} \\
			\\ 
			\text{\task{sessions.py}} \\ 
		\end{cases}
		\\ \\
		\text{\task{tools}}
		\begin{cases}
			\text{\task{power\_analysis.R}} \\
			\\ 
			\text{\task{rawlmat2lmat.py}} \\
			\\ 
			\text{\task{lmat2cmplx.py}} \\
			\\ 
			\text{\task{pyLMAT\_rl.py}} \\
			\\ 
			\text{\task{pyLMAT\_cs.py}} \\
			\\ 
			\text{\task{pyLMAT\_gl.py}} \\
			\\
			\text{\task{LMAT}}
			\begin{cases}
	    		\text{\task{pyLMAT\_rescore.py}} \\ 
	    		\\ 
	    		\text{\task{HMPspiker.py}} 
			\end{cases}
		\end{cases}
	\end{cases} \]
	\caption{\CC\ layout}
	\label{fig:CC}
\end{figure}

\CC\ performs the power-law fit described in the \emph{Blumm, N. et al.} paper, but by fitting the best model, i.e. choosing between fitting a power-law using linear regression versus nonlinear regression\cite{ecology}. In the power-law fit plots we also show the generalized coefficient of determination computed for continuous models\cite{genR2,disR2}.

\section{Un-weighted power-law fit}\label{sec:unw}

\subsection{Fitting the best model}

As mentioned above, to choose between fitting power laws ($y=Vx^\beta$) using linear regression on log-transformed (LLR) data versus non-linear regression (NLR), we mainly follow \emph{General Guidelines for the Analysis of Biological Power Laws}\cite{ecology}. It consists of the following three steps:
\begin{enumerate}
    \item Determining the appropriate error structure by likelihood analysis.
    \begin{enumerate}
        \item Fit the Non-Linear Regression (NLR) model and obtain $V_\text{NLR}$, $\beta_\text{NLR}$ and $\sigma_\text{NLR}^2$.
        \item Calculate the loglikelihood that the data ($n$ is sample size) are generated from a normal distribution with additive error:
        \begin{itemize}
            \item The likelihood of a normal distribution is:  
            $$\mathcal{L}_\text{norm} = \prod_{i=1}^n\left[\frac{1}{\sqrt{2\pi\sigma^2_\text{NLR}}}\;\exp{\left(-\frac{\left(y_i-V_\text{NLR}x_i^{\beta_\text{NLR}}\right)^2}{2\sigma^2_\text{NLR}}\right)}\right]$$
            \item So, the loglikelihood of a normal distribution is:
            \begin{eqnarray*}
                \log\mathcal{L}_\text{norm} &=& -\frac{n}{2}\log\left|2\pi\sigma^2_\text{NLR}\right| - \frac{1}{2\sigma^2_\text{NLR}}\underbrace{\sum_{i=1}^n\left(y_i-V_\text{NLR}x_i^{\beta_\text{NLR}}\right)^2}_{\mathrm{RSS
                }_\text{NLR}}\\
                &=& -\frac{n}{2}\log\left|2\pi\sigma^2_\text{NLR}\right|-\frac{\mathrm{RSS}_\text{NLR}}{2\sigma^2_\text{NLR}}
            \end{eqnarray*}
        \end{itemize}
        \item Calculate the \emph{corrected Akaike's Information Criterion} for the NLR model:
        $$\mathrm{AIC_{c_{NLR}}} = 2k - 2\log\mathcal{L}_\text{norm} + \frac{2k(k+1)}{n-k-1}$$
        \item Fit the Log-transformed Linear Regression (LLR) model and obtain $V_\text{LLR}$, $\beta_\text{LLR}$ and $\sigma_\text{LLR}^2$.
        \item Calculate the loglikelihood that the data ($n$ is sample size) are generated from a lognormal distribution with multiplicative error:
        \begin{itemize}
            \item The likelihood of a lognormal distribution is: 
            $$\mathcal{L}_\text{logn} = \prod_{i=1}^n\left[\frac{1}{y_i\sqrt{2\pi\sigma^2_\text{LLR}}}\;\exp{\left(-\frac{\left(\log|y_i|-\log|V_\text{LLR}|-\beta_\text{LLR}\log|x_i|\right)^2}{2\sigma^2_\text{LLR}}\right)}\right]$$
            \item So, the loglikelihood of a lognormal distribution is: 
            \begin{eqnarray*}
                \log\mathcal{L}_\text{logn} &=& -\frac{n}{2}\log\left|2\pi\sigma^2_\text{LLR}\right| - \sum_{i=1}^n\log|y_i| -\\
                &&\qquad-\frac{1}{2\sigma^2_\text{LLR}}\underbrace{\sum_{i=1}^n\left(\log|y_i|-\log|V_\text{LLR}|-\beta_\text{LLR}\log|x_i|\right)^2}_{\mathrm{RSS
                }_\text{LLR}}\\
                &=& -\frac{n}{2}\log\left|2\pi\sigma^2_\text{LLR}\right| - \frac{\mathrm{RSS}_\text{LLR}}{2\sigma^2_\text{LLR}} - \sum_{i=1}^n\log|y_i|
            \end{eqnarray*}
        \end{itemize}
        \item Calculate the \emph{corrected Akaike's Information Criterion} for the LR model: 
        	$$\mathrm{AIC_{c_{LLR}}} = 2k - 2\log\mathcal{L}_\text{logn} + \frac{2k(k+1)}{n-k-1}$$
    \end{enumerate}
    \item Compare $\mathrm{AIC_{c_{NLR}}}$ with $\mathrm{AIC_{c_{LLR}}}$:
    \begin{itemize}
        \item If $\mathrm{AIC_{c_{NLR}}} - \mathrm{AIC_{c_{LLR}}} < -2$, the assumption of normal error is favoured compared to lognormal error, so proceed with the results obtained from the NLR fit.
        \item If $\mathrm{AIC_{c_{NLR}}} - \mathrm{AIC_{c_{LLR}}} > 2$, the assumption of lognormal error is favoured compared to normal error, so proceed with the results obtained from the LLR fit.
        \item If $\left|\mathrm{AIC_{c_{NLR}}} - \mathrm{AIC_{c_{LLR}}}\right| \leq 2$, no model is favoured, so proceed with model averaging:
        \begin{eqnarray*}
            B_\text{av} &=& w_\text{NLR}V_\text{NLR} + w_\text{LLR}V_\text{LLR} \\
            \beta_\text{av} &=& w_\text{NLR}\beta_\text{NLR} + w_\text{LLR}\beta_\text{LLR}
        \end{eqnarray*}
        where: 
        \begin{eqnarray*}
            w_\text{NLR} &=& \frac{1}{1+\mathrm{e}^{\frac{1}{2}\left(\mathrm{AIC_{c_{NLR}}}-\mathrm{AIC_{c_{LLR}}}\right)}} \\
            w_\text{LLR} &=& \frac{1}{1+\mathrm{e}^{\frac{1}{2}\left(\mathrm{AIC_{c_{LLR}}}-\mathrm{AIC_{c_{NLR}}}\right)}}
        \end{eqnarray*}
        which are obtained to fulfill the next condition: $w_\text{NLR} + w_\text{LLR} = 1$. The CIs for $B_\text{av}$ and $\beta_\text{av}$ are to be generated by ordinary bootstrapping\footnote{\CC\ has available the next bootstrapping alternatives\cite{boot}: ordinary, ``Resampling Residuals'' method, ``Wild'' method, and ``Monte-Carlo'' method.}.      
    \end{itemize}
    \item Assess the validity of the underlying statistical assumptions with diagnostic plots because while it is rare for all the assumptions to be fully satisfied by real-life data sets, major violations indicate the lack of appropriateness of the model and, thus, the potential invalidity of the results.
\end{enumerate}

\subsection{Calculating the coefficient of determination}

We think the best approach in this situation is to apply the generalized $R^2$ that, for continuous models, was defined as\cite{genR2}:
$$ R^2 = 1 - \left(\frac{\mathcal{L}(0)}{\mathcal{L}(\hat\theta)}\right)^{\!\!\frac{2}{n}} $$
where $\mathcal{L}(\hat\theta)$ and $\mathcal{L}(0)$  denote the likelihoods of the fitted and the ``null'' model, respectively, and $n$ is the sample size. In terms of the loglikelihoods, the generalized coefficient of determination would be:
$$ R^2 = 1 - \mathrm{e}^{-\frac{2}{n}\left(\log\mathcal{L}(\hat\theta)-\log\mathcal{L}(0)\right)} $$
We have the likelihoods calculated from the previous section, but what about the ``null'' models? We understand that they are the models with only the intercept. So for the Gaussian additive error model:
$$\mathcal{L}_\text{norm}(0) = \prod_{i=1}^n\left[\frac{1}{\sqrt{2\pi\sigma^2_\text{NLR0}}}\;\exp{\left(-\frac{\left(y_i-\bar y\right)^2}{2\sigma^2_\text{NLR0}}\right)}\right]$$
So:
\begin{eqnarray*}
\log\mathcal{L}_\text{norm}(0) &=& -\frac{n}{2}\log\left|2\pi\sigma^2_\text{NLR0}\right| - \frac{1}{2\sigma^2_\text{NLR0}}\sum_{i=1}^n\left(y_i-\bar y\right)^2\\
 &=& -\frac{n}{2}\left(\log\left|2\pi\sigma^2_\text{NLR0}\right| + 1 \right)
\end{eqnarray*}
since $\sigma^2_\text{NLR0}=\frac{1}{n}\sum\left(y_i-\bar y\right)^2=\frac{1}{n}\mathrm{TSS}_\text{NLR}$. Now, coming back to the coefficient of determination, we have:
\begin{eqnarray*}
R^2_\text{NLR} &=& 1 - \mathrm{e}^{\frac{2}{n}\left(\log\mathcal{L}_\text{NLR}(0)-\log\mathcal{L}_\text{NLR}(\hat\theta)\right)} = 1 - \exp{\left(\frac{\log(\mathrm{RSS}_\text{NLR})}{\log(\mathrm{TSS}_\text{NLR})}\right)}
 =\\ &=& 1 - \frac{\mathrm{RSS}_\text{NLR}}{\mathrm{TSS}_\text{NLR}} =
 1 - \frac{\sum_{i=1}^n\left(y_i-V_\text{NLR}x_i^{\beta_\text{NLR}}\right)^2}{\sum_{i=1}^n\left(y_i-\bar y\right)^2}
\end{eqnarray*}
recovering the traditional expression for $R^2$. Using the same approach for calculating $R^2_\text{LLR}$, then:
$$\mathcal{L}_\text{logn}(0) = \prod_{i=1}^n\left[\frac{1}{y_i\sqrt{2\pi\sigma^2_\text{LLR0}}}\;\exp{\left(-\frac{\left(\log|y_i|-\log|B_\text{LLR0}|\right)^2}{2\sigma^2_\text{LLR0}}\right)}\right]$$
So:
\begin{eqnarray*}
\log\mathcal{L}_\text{logn}(0) &=& -\frac{n}{2}\log\left|2\pi\sigma^2_\text{LLR0}\right| - \frac{1}{2\sigma^2_\text{LLR0}}\sum_{i=1}^n\left(\log|y_i|-\overline{\log|y|}\right)^2 - \sum_{i=1}^n\log|y_i|\\
 &=& -\frac{n}{2}\left(\log\left|2\pi\sigma^2_\text{LLR0}\right| + 1 \right) - \sum_{i=1}^n\log|y_i|
\end{eqnarray*}
since $\sigma^2_\text{LLR0}=\frac{1}{n}\sum\left(\log|y_i|-\overline{\log|y|}\right)^2=\frac{1}{n}\mathrm{TSS}_\text{logn}$. Again, recalling the expression for the generalized coefficient of determination, we have:
\begin{eqnarray*}
R^2_\text{LLR} &=& 1 - \mathrm{e}^{\frac{2}{n}\left(\log\mathcal{L}_\text{LLR}(0)-\log\mathcal{L}_\text{LLR}(\hat\theta)\right)} = 1 - \exp{\left(\frac{\log(\mathrm{RSS}_\text{LLR})}{\log(\mathrm{TSS}_\text{LLR})}\right)}
 =\\ &=& 1 - \frac{\mathrm{RSS}_\text{LLR}}{\mathrm{TSS}_\text{LLR}} =
 1 - \frac{\sum_{i=1}^n\left(\log|y_i|-\log|V_\text{LLR}|-\beta_\text{LLR}\log|x_i|\right)^2}{\sum_{i=1}^n\left(\log|y_i|-\overline{\log|y|}\right)^2}
\end{eqnarray*}

\section{X-weighted power-law fit}\label{sec:X-w}

When fitting the power-law of std vs. mean, we can take into account that every mean has uncertainty and estimate it for a sample size $n$ by the SEM (\emph{Standard Error of the Mean}):
$$ \mathrm{SEM} = \frac{s}{\sqrt{n}}$$
where $s$ is the sample standard deviation. So, the vector of weights is computed with:
$$ \mathbf{w} = \frac{1}{\overrightarrow{\mathrm{SEM}}} = \frac{\sqrt{\mathbf{n}}}{\mathbf{s}}$$

Here, the uncertainties affect the independent variable, so the fit is not so trivial as a Y-weighted fit, where the uncertainties affect the dependent variable. A standard approach to do this fit is: a) invert your variables before applying the weights, b) then perform the weighted fit, and finally, c) revert the inversion. This method is deterministic, but the approximate solution worsens with smaller $R^2$. For comparison, we develop a stochastic method by using a bootstrapping-like strategy that avoids the inversion and is applicable regardless of $R^2$. Both methods, detailed below, are implemented in \CC.

\subsection{Method 1: By inverting the data}

In the case of the log-LR model, we have:
$$\log y = \log V + \beta\log x \quad\rightarrow\quad \underbrace{\log x}_{\tilde y} = \overbrace{-\frac{1}{\beta}\log V}^{b} + \overbrace{\frac{1}{\beta}}^{m}\underbrace{\log y}_{\tilde x}$$ 
where $m$ determines the slope or gradient of the fitted line, and $b$ determines the point at which the line crosses the y-axis, otherwise known as the y-intercept. Once the model is fitted, the original parameters can be retrieved easily:
\begin{eqnarray*}
\beta &=& \frac{1}{m} \\
V &=& \mathrm{e}^{-\beta b} = \mathrm{e}^{-\frac{b}{m}}
\end{eqnarray*}
Their respective uncertainties are to be obtained using \emph{error propagation}:
\begin{eqnarray*}
    \sigma_\beta &=& \left|\frac{\mathrm{d}\beta}{\mathrm{d}m}\right|\sigma_m \quad=\quad \frac{1}{m^2}\;\sigma_m \\
    \sigma_V &=& \sqrt{\left(\frac{\partial V}{\partial b}\right)^{\!\!2}\sigma_b^2 +
      \left(\frac{\partial V}{\partial m}\right)^{\!\!2}\sigma_m^2} \quad=\quad
      \frac{1}{m}\;\mathrm{e}^{-\frac{b}{m}}\,\sqrt{\sigma_b^2 + \frac{b^2}{m^2}\;\sigma_m^2} 
\end{eqnarray*}

\subsection{Method 2: Bootstrapping-like strategy}

The basic idea of bootstrapping is that inference about a population from sample data (sample $\rightarrow$ population) can be modeled by resampling the sample data and performing inference on (resample $\rightarrow$ sample). To adapt this general idea to our problem, we resample the x-data array using its errors array. That is, for each replicate, a new x-data array is computed based on:
$$x^*_i = x_i + v_i$$
where $v_i$ is a Gaussian random variable with mean $\mu_i=0$ and standard deviation $\sigma_i=\mathrm{SEM}_i$, as defined previously in this supplementary material. For each replicate a complete un-weighted power-law fit is performed, as described in the previous section. It is worth mentioning that each replicate is filtered to avoid values of $x^*_i$ under \emph{eps} (obtained by \texttt{np.finfo(np.double).eps}) in order to keep away from the error of getting log of negatives or zero during the fit.

We devised and implemented a multi-step algorithm to estimate the fit parameters that finishes when a relative error of less than $10^{-4}$ is achieved. It also ends if the number of steps reaches $100$ to avoid too much time lapse, to prevent any pathologic numeric case which, in fact, we still have not detected in all the data sets analyzed.

In the previous version of the algorithm, for each step, the method generated $10$ replicates for each x-data point, in other words, it was computing the fit for $10$ times the length of the x-data array replicates, with a maximum of $10000$ fits per step. Nevertheless, we found that such an approach depending on the length of the x-data array did not perform better, so we decided to simplify the method and fix the number of fits per step in $100$. This latter approach improved the performance.

The parameters of the X-weighted fit are then estimated by averaging through all the replicate fits performed, and their errors are estimated by computing the standard deviation also for all the fits. At the end of each step, the relative error is calculated by comparing the fit parameters estimation in the last step with the previous one.

Finally, both the coefficient of determination of the fit and the coefficient of correlation between the fit parameters are estimated by averaging.

\section{Rank Stability Index (RSI)}\label{sec:RSI}

The Rank Stability Index is shown as a percentage in a separate bar on the right of the rank matrix plot provided by \CC. The RSI is strictly $1$ for an element whose range never changes over time, and is strictly $0$ for an element whose rank oscillates between the extremes from time to time. So, RSI is calculated, per element, as $1$ less the quotient of the number of true rank hops taken between the number of maximum possible rank hops, all powered to $p$:
$${\rm RSI} = \left(1-\frac{\text{true rank hops}}{\text{possible rank hops}}\right)^p = \left(1-\frac{D}{(N-1)(t-1)}\right)^p$$
where $D$ is the total of rank hops taken by the studied element, $N$ is the number of elements that have been ranked, and $t$ is the number of time samples. The power index $p$ is arbitrarily chosen to increase the resolution in the stable region; the value in the current version of the code is $p=4$.

As an example of this ``zooming'' effect in the stable region, to match a linear ($p=1$) RSI of 0.9 to a powered one of 0.1, we should select $p=21.8543$. An alternative way to obtain this effect and exactly map a linear RSI of 0.9 to a non-linear RSI (${\rm RSI'}$) of 0.1, is by applying the following function:
$${\rm RSI'} = \frac{10^{10\left(1-\frac{D}{(N-1)(t-1)}\right)}-1}{10^{10}-1} \approx 10^{-10\left(\frac{D}{(N-1)(t-1)}\right)} $$
where the approximation is valid because $10^{10}\gg 1$ but, the small price to pay for it is that, in the worst instability case, the ${\rm RSI'}$ would not be strictly $0$ but $10^{-10}$.

The colour code of the RSI percentage text in the rank plot of \CC\ is chosen following the first condition satisfied from those shown in Supplementary Table \ref{tab:RSI} (see page \pageref{tab:RSI}).

\begin{table}
  \begin{center}
    \begin{tabular}{ccc}
    \hline
    Case  &  Condition  &  Colour  \\
    \hline
    1  &  $1\ge{\rm RSI}>0.99$  & \textcolor{blue}{blue}  \\ 
    2  & ${\rm RSI}>0.90$  &  \textcolor{green}{green}  \\
    3  &  ${\rm RSI}>0.75$  &  \textcolor{orange}{orange} \\
    4  &  ${\rm RSI}>0.25$  &  \textcolor{red}{red} \\
    5  &  $0.25\ge{\rm RSI}\ge0$  &  \bfseries{black}  \\
    \hline
    \end{tabular}
  \end{center}
  \caption{Colour code of the RSI percentage text shown in rank plots, following the first condition satisfied.}
  \label{tab:RSI}
\end{table}

\section{\CC\ output}
In the previous sections, we have discussed details about the methods used in \CC. In this section we review the output from the package, which aims at summarizing the as yet undescribed functionality. Unless the flag \texttt{-n}/\texttt{--nodirs} is given to the \CC.py front-end script, the output will be spread into subdirectories (\texttt{corrank}, \texttt{fits}, \texttt{hist} and \texttt{other}) under the results directory. The path of the results directory is relative to the data directory, and is given by the optional argument \texttt{-r}/\texttt{--results}, which defaults to ``\texttt{results}''. A bunch of plots distributed in subdirectories of such path is generated by \CC\ (the flag that is reducing such number to the minimum is \texttt{-l}/\texttt{--less}). They are described in the next subsections. In addition, the code provides with output in CSV format (one file per input file), in Excel format (one summary spreadsheet with one sheet per input file, plus correlations and rank tables in the \texttt{corrank} subdirectory) and, optionally, in \LaTeX\ and PDF format (thanks to pdf\LaTeX).

\subsection{Fit Plots}
The subdirectory \texttt{fits} under the results directory contains the fit plots for each data set. Both for the unweighted fit (detailed in the Supplementary Section \ref{sec:unw}) and for the X-weighted fit (detailed in the Supplementary Section \ref{sec:X-w}), two plots are generated: the former with logarithmic scale, the latter with lineal scale. Supplementary Figures \ref{fig:unwFit} and \ref{fig:X-wFit} show examples of the unweighted and X-weighted fits, respectively. Additionally, for the unweighted fit, a complete residues analysis is performed, and a 4-in-1 figure is generated as shown in Supplementary Figure \ref{fig:unwRes} (corresponding to the fit of Supplementary Fig. \ref{fig:unwFit}). Among other tests, it allows to check for normality and homoscedasticity of the residues.

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_LOG.pdf}
	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_LIN.pdf}
	\caption{Example of unweighted fit, shown both in logarithmic and lineal scale plots}
	\label{fig:unwFit}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_xWboot_LOG.pdf}
	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_xWboot_LIN.pdf}
	\caption{The X-weighted fit log and lineal plots corresponding to the fit of Figure \ref{fig:unwFit}}
	\label{fig:X-wFit}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_RES.pdf}
	\caption{Residues analysis plot corresponding to the fit of Figure \ref{fig:unwFit}. The top-left subplot is a simple residues plot. The top-right subplot is a Normal quantiles plot with linear fitting (value of coefficient of determination is provided). The bottom-left subplot shows an accumulated ESS (Explained Sum of Squares) plot. Finally, the bottom-right subplot is a residues Normal histogram plot. This set of subplots allows to check for normality and homoscedasticity of the residues.}
	\label{fig:unwRes}
\end{figure}

\subsection{Histogram Plots} 
The subdirectory \texttt{hist} under the results directory contains the histogram plots. \CC\ generates three different histogram plots:
\begin{description}
	\item[$-$Absolute frequencies plot]: This plot is useful to visually assess the validity of the time points in terms of the accumulated absolute frequency of the elements (taxa), since absolute frequencies far (much higher or much lower) from those typically observed could mean a sampling problem. As an example, Supplementary Figure \ref{fig:histAFP} shows this histogram for the pre-treatment data (first 7 times) of patient ``D'' in the antibiotics study\cite{antibiotic}. 
	\item[$-$2D deviation plot]: The 2D semi-logarithmic histogram representing deviations from the mean versus the mean itself, is a useful tool in the analysis of the stability of ranking processes in complex systems\cite{ranking}. Supplementary Figure \ref{fig:hist2D} shows this plot for the data used in the fit shown in Supplementary Figure \ref{fig:unwFit}.
	\item[$-$Zero relative frequency plot]: We could define the ZRF (Zero Relative Frequency, thereby ranging from 0 to 1) of an element (taxon) as the portion of times where it is zero, i.e., it is not found. Attending to all the elements (taxa), we can plot the ZRF histogram, which then lies on the horizontal axis of the plot. The vertical axis shows the number of elements (taxa), so the height of a bar represents the number of elements that have determinate ZRF. In this respect, the bar over $0.0$ counts the number of elements (taxa) that are present at every time point of the data set (aka ``core''), while the bar over $1.0$ would count the number of elements (taxa) that are never found (this bar never appears because all these ``null'' elements are automatically filtered by the code). Supplementary Figure \ref{fig:histZRF} shows an example of this plot. There, we can see that $12$ taxa are present at all the time points of the time series while $9$ taxa basically appear only once. So, this plot is clearly useful to notice how the ``core'' is distributed.
\end{description}

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{results/hist/antibiotics_h_patD_pre_AbsFreqPlot}
	\caption{Histogram with the absolute frequencies of the pre-treatment data (7 first times) of patient ``D'' in the antibiotics study\cite{antibiotic}}
	\label{fig:histAFP}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{results/hist/IBS_h_A_amplicons_family_hist2D.pdf}
	\caption{2D histogram deviation plot of the data used in the fit shown in Figure \ref{fig:unwFit}}
	\label{fig:hist2D}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{results/hist/IBS_h_A_amplicons_family_ZRFhist.pdf}
	\caption{Histogram with the relative frequency of zero for the elements (taxa) in the data used in the fit shown in Figure \ref{fig:unwFit}}
	\label{fig:histZRF}
\end{figure}

\subsection{Correlation and Rank Plots} 
The subdirectory \texttt{corrank} under the results directory contains the correlation matrix and rank\&stability plots, as well as Excel files with the resulting matrices. \CC\ generates three different plots falling under this category:
\begin{description}
	\item[$-$Elements correlation matrix]: This plot shows a correlation matrix among the elements (taxa), calculated with the time as independent variable. For these calculations, the data set is not normalized to avoid entering an additional constraint. As an example, Supplementary Figure \ref{fig:corrElm} shows this matrix for the ``core'' elements (taxa) present in the pre-treatment data (first seven times) of patient ``D'' in the antibiotics study\cite{antibiotic}. 
	\item[$-$Times correlation matrix]: This plot presents a correlation matrix among the time points of the data set, calculated with the elements (taxa) as independent variable. Again, the data set is not normalized. Supplementary Figure \ref{fig:corrTim} shows this matrix for the ``core'' elements (taxa) present in pre-treatment data (first seven times) of patient ``D'' in the antibiotics study\cite{antibiotic}. 
	\item[$-$Rank dynamics and stability plot]: This plot shows the variation in the rank with time for the most dominant elements (taxa) and their calculated RSI, as discussed in Supplementary Section \ref{sec:RSI}. Supplementary Figure \ref{fig:corrank} shows this plot for the elements (taxa) in the data used in the fit shown in Supplementary Figure \ref{fig:unwFit}.
\end{description}

\begin{figure}
	\centering
	\includegraphics[width=0.9\textwidth]{results/corrank/antibiotics_h_patD_pre_CORE_ElementsCorr}
	\caption{Element correlation plot of the pre-treatment data (first seven times) of patient ``D'' in the antibiotics study\cite{antibiotic} for ``core'' taxa}
	\label{fig:corrElm}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.9\textwidth]{results/corrank/antibiotics_h_patD_pre_CORE_TimesCorr}
	\caption{Times correlation plot of the pre-treatment data (first seven times) of patient ``D'' in the antibiotics study\cite{antibiotic} for ``core'' taxa}
	\label{fig:corrTim}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{results/corrank/IBS_h_A_amplicons_family_Rank}
	\caption{Matrix showing the rank variation throughout time for the most dominant elements (taxa) and their calculated RSI (as discussed in Section \ref{sec:RSI}) in the data used in the fit shown in Figure \ref{fig:unwFit}}
	\label{fig:corrank}
\end{figure}

\subsection{Other Plots} 
The subdirectory \texttt{other} under the results directory contains other plots, like the elements (taxa) kurtosis and skewness against the mean. These plots are generated when the flag \texttt{-m}/\texttt{--more} is given to the \CC.py front-end script.

\subsection{\CC\ advanced options}
The advanced option flags are summarized here:
\vspace{-5mm} %%% CAUTION: FORCED LAYOUT
\begin{description}
	\item[$-$ZRF filter]: This is activated with \texttt{-c}/\texttt{--core} and allows global and detached results for ``core'', ``tailless'' and ``tail'', which allows comparative analysis depending on the presence/absence rate of the taxa throughout time, which is calculated in practice using the Zero Relative Frequency (ZRF) defined above. The criteria is as follows: 
	\begin{itemize}
		\item In any case, if $\operatorname{ZRF} < 0.2$ taxa are classified as ``core''. In other words, the taxa considered ``core'' are those appearing at over $80\%$ of the time points under analysis.
		\item Depending on the number of times:
		\begin{itemize}
			\item If there are only 3 time points, taxa with $0.2 \le\operatorname{ZRF} < 0.4$ are classified as ``tailless'' and, therefore, taxa with $\operatorname{ZRF} \ge 0.4$ are classified as ``tail''.
			\item If there are 4 time points, taxa with $0.2 \le\operatorname{ZRF} < 0.5$ are classified as ``tailless'' and, therefore, taxa with $\operatorname{ZRF} \ge 0.5$ are classified as ``tail''.
			\item If there are 5 or more time points, taxa with $0.2 \le\operatorname{ZRF} < 0.6$ are classified as ``tailless'' and, therefore, taxa with $\operatorname{ZRF} \ge 0.6$ are classified as ``tail''.			      
		\end{itemize}
    \end{itemize}
	\item[$-$Relative frequency filters]: The minimum frequency filter is activated with \texttt{-z}/\texttt{--fmin} and the maximum one with \texttt{-x}/\texttt{--fmax}, with values between $0$ and $1$. If both filters are activated simultaneously, the minimum frequency cannot be equal or higher than the maximum one. If no filter is activated, by default, \CC\ is filtering all the taxa that never appear throughout time, in the event there are such taxa among the input data. On the one hand, if a minimum frequency value is provided, all the taxa whose relative accumulated frequencies (the normalized sum of the values through time) are under such a value, are eliminated. But in addition, once the set of data is normalized (to work with relative frequencies), if there are data at any time points that are under the minimum frequency given value, they are automatically floored to zero (we call this \emph{item@time filtering}), and the set of data is renormalized. On the other hand, if a maximum frequency value is provided, all the taxa whose relative accumulated frequencies are over such value, are deleted. 
\end{description}

\section{\CC\ input}
The input data are provided by two simultaneously available ways:
\vspace{-5mm} %%% CAUTION: FORCED LAYOUT
\begin{enumerate}
	\item Text files (\texttt{txt} extension), to be processed in parallel\footnote{Currently, parallelization in \CC\ is only available under GNU/Linux}, with a shared-memory model.
	\item Excel files (\texttt{xls[x]} extension), whose sheets are related datasets to be processed in parallel\footnotemark[\value{footnote}]. Optionally, all the sheets of an Excel file are standardized using part of such Excel datasets, following the procedure described in the Supplementary Section \ref{sec:stan}, in which event, additional output is provided by \CC.
\end{enumerate}
Typically, the data sets are tables with the taxa lying in different rows and the time points ordered in different columns. So, the first column is considered the taxon labels, whereas the first row contains the time labels. It is worth mentioning that any Excel sheet or text file beginning with the underscore character is ignored by \CC, as they are supposed to hold metadata.

\section{Standardization} \label{sec:stan}
In order to show all the studies properly under common axes, we decided to standardize the Taylor parameters using the group of healthy individuals for each study. With this approach, all the studies can be visualized in a shared plot with units of Taylor-parameters standard-deviation on their axes.

For a Taylor parameter, e.g. $V$, the estimate of the mean ($\widehat{V}$) for the healthy subpopulation, composed of $h$ individuals, is:
$$\widehat{V} = \frac{1}{W_1}\sum_{i=1}^h V_i \omega_i=\sum_{i=1}^h V_i \omega_i$$
as $W_1=\sum_i^h \omega_i=1$, since $\omega_i$ are normalized weights calculated as:
$$\omega_i = \frac{\frac{1}{\sigma^2_{V_i}}}{\sum_i^h\frac{1}{\sigma^2_{V_i}}}$$
being $\sigma_{V_i}$ the estimation of the uncertainty in $V_i$ obtained together with $V_i$ from the X-weighted power-law fit described in Supplementary Section \ref{sec:X-w}, for healthy individuals.

Likewise, the estimation of the standard deviation for the healthy population ($\widehat{\sigma}_V$) is:
$$\widehat{\sigma}_V = \sqrt{\frac{1}{W_1-\frac{W_2}{W_1}}\sum_{i=1}^h\left[\omega_i\left(V_i-\hat{V}\right)^2\right]}$$
being $W_2=\sum_i^h \omega_i^2$, which finally yields to:
$$\widehat{\sigma}_V = \sqrt{\frac{1}{1-\sum_i^h \omega_i^2}\sum_{i=1}^h\left[\omega_i\left(V_i-\hat{V}\right)^2\right]}$$

\section{\CC\ interface with LMAT} \label{sec:lmat}
With the aim of adapting LMAT (Livermore Metagenomics Analysis Toolkit) software\cite{LMAT1, LMAT2} for its easy use in time series data, like the ones from the Kwashiorkor paper\cite{kwashiorkor}, we developed Python front and back-end scripts, which are supplied in the directory \texttt{tools} of the \CC\ package.

\subsection{LMAT front-ends} They are: \texttt{pyLMAT\_rl.py}, the launcher for LMAT \texttt{run\_rl.sh}, that is, the first step in the LMAT pipeline or \emph{read labeling}; \texttt{pyLMAT\_cs.py}, the launcher for LMAT \texttt{run\_cs.sh}, the \emph{content summarization} step in the LMAT pipeline; and, \texttt{pyLMAT\_gl.py}, the launcher for LMAT \texttt{run\_gl.sh}, the \emph{gene labeling/identification} step in the LMAT pipeline. Basically, all three scripts permit to manage the execution of parallelized LMAT against several fasta files coming from diferent time points in time series studies.

\subsection{LMAT back-ends} Currently, they are: \texttt{pyLMAT\_rescore.py} and the pipeline composed by \texttt{rawlmat2lmat.py} and \texttt{lmat2cmplx.py}. The first one allows not only to quick recalculate the LMAT results for different cutoff levels, but also to pull out reads with specific NCBI taxId (\emph{human} reads by default). For example, issuing "\texttt{pyLMAT\_rescore.py -p path/to/results -f file.fna -m -1 0 1 -t=32}" will regenerate the LMAT results for cutoffs $-1$, $0$ and $1$, and will pull human reads for each cutoff level, parallelizing in $32$ threads. On the other hand, the mentioned pipeline enables to post-process the several \emph{fastsummary} output files from LMAT first and second steps launched by the LMAT front-ends described above. The pipeline transforms the results so that they achieve a format and content ready for \CC\ computations. Such data arrangements are done independently for two different taxonomic levels: genus and specie.

\newpage
\section*{Supplemental material references}
\begin{thebibliography}{99}
	\bibitem{ranking} Blumm, N. \textit{et al.}, Dynamics of ranking processes in complex systems. {\it Phys. Rev. Lett.} {\bf 109,} 128701 (2012).
	\bibitem{FD} Weber, J. \textit{et al.}, Fluctuation dissipation theorem. {\it Phys. Rev.} {\bf 101}, 1620-6 (1956).
	\bibitem{kwashiorkor} Smith M.I.  \textit{et al.}, Gut microbiomes of Malawian twin pairs discordant for kwashiorkor. {\it Science} {\bf 339,} 548-54 (2013).
	\bibitem{FASTX} Gordon, A., Hannon, G.J., FASTX-Toolkit. FASTQ/A shortreads pre-processing tools (2010). http://hannonlab.cshl.edu/fastx\_toolkit/ (accessed 23 Feb 2015).
	\bibitem{QIIME} Caporaso, J.G. \textit{et al.}, QIIME allows analysis of high-throughput community sequencing data. {\it Nature Methods} {\bf 7,} 335-6 (2010).
	\bibitem{SILVA} Quast C.  \textit{et al.}, The SILVA ribosomal RNA gene database project: improved data processing and web-based tools (2013)
	\bibitem{LMAT1} Ames, S.K. Hysom, D.A., Gardner, S.N., Lloyd, G.S., Gokdale, M.B., Allen, J.E., Scalable metagenomic taxonomy classification using a reference genome database. {\it Bioinformatics} {\bf 29,} 2253-2260 (2013).
	\bibitem{LMAT2} Ames, S.K., Gardner, S.N., Marti, J.M., Slezak, T.R., Gokdale, M.B., Allen, J.E., Using populations of human and microbial genomes for organism detection in metagenomes {\it Genome Research} {doi:10.1101/gr.184879.114} (2015).
	\bibitem{antibiotic} Dethlefsen L., Relman D. A., Incomplete recovery and individualized responses of the human distal gut microbiota to repeated antibiotic perturbation. {\it Proc. Nat. Acad. Sci. USA} {\bf 108,} 4554-61 (2011).
	\bibitem{diet} David, L.A. \textit{et al.}, Diet rapidly and reproducibly alters the human gut microbiome. {\it Nature} {\bf 505,} 559-63 (2014).
	\bibitem{IBS} Durban, A. \textit{et al.}, Structural alterations of faecal and mucosa-associated bacterial communities in irritable bowel syndrome. {\it FEMS Microbiol. Ecol.} {\bf 86}, 581-9 (2013).
	\bibitem{LEA} Faith, J.J. \textit{et al.}, The long-term stability of the human gut microbiota. {\it Science} {\bf 341,} 1237439 (2013).
	\bibitem{moving} Caporaso, J.G. \textit{et al.}, Moving pictures of the human microbiome. {\it Genome Biol.} {\bf 12,} R50 (2011).
 	\bibitem{ecology} Xiao Xiao, Ethan P. White, Mevin B. Hooten, and Susan L. Durham., On the use of log-transformation vs. nonlinear regression for analyzing biological power laws. {\it Ecology} {\bf 92}, 10, 1887-1894 (2011).
 	\bibitem{genR2} Magee L., $R^2$ measures based on wald and likelihood ratio joint significance tests. {\it The American Statistician} {\bf 44}, 3, 250-253 (1990).
	\bibitem{disR2} Nagelkerke N.J.D., A note on a general definition of the coefficient of determination. {\it Biometrika} {\bf 78}, 3, 691-692 (1991).
	\bibitem{boot} Wu, C.F.J., Jackknife, bootstrap and other resampling methods in regression analysis. (with discussions) \textit{The Annals of Statistics} {\bf 14}: 1261-1350 (1986)
\end{thebibliography}

\end{document}

### 2016-04-15 NatBiotech/CC_NatBiotech.tex
Include previous releases for different publications
On branch master
	new file:   MolSysBio/CC-suppl_molsysbio.docx
	new file:   MolSysBio/molsysbio_CC.docx
	new file:   NatBiotech/CC-suppl_NatBiotech.pdf
	new file:   NatBiotech/CC-suppl_NatBiotech.tex
	new file:   NatBiotech/CC_NatBiotech.pdf
	new file:   NatBiotech/CC_NatBiotech.tex
	new file:   NatBiotech/lettertoeditor_NatBiotech.docx
	new file:   Nature/CC-suppl_sub.pdf
	new file:   Nature/CC_sub.pdf
	new file:   Nature/lettertoeditor2.docx
	new file:   Science/lettertoeditor_sub.docx
	new file:   Science/science_CC.pdf
	new file:   Scie

%% Template for a preprint Letter or Article for submission
%% to the journal Nature.
%% Written by Peter Czoschke, 26 February 2004
%%

\documentclass{nature}

%% make sure you have the nature.cls and naturemag.bst files where
%% LaTeX can find them

\bibliographystyle{naturemag}
\usepackage{graphicx}

%%% HYPERREF
\usepackage[pdftex,
	pdftitle={Microbiota: are you sick?},
	pdfauthor={Jose Manuel Marti, Daniel M. Martinez, Manuel Peña, Cesar Gracia, Amparo Latorre, Andres Moya, Carlos P. Garay},
	pdfsubject={Nature Biotechnology},
	pdfkeywords={brief communication},
	colorlinks=true,
	urlcolor={red},
	bookmarks=false]{hyperref}

% Title
\title{Microbiota: are you sick?}
% Alternative title
% \title{Health and disease imprinted in the temporal variability of the human microbiome.}

%% Notice placement of commas and superscripts and use of &
%% in the author list

\author{Jose Manuel Mart\' i$^{1}$, Daniel M. Mart\'inez$^{1,2}$, Manuel Pe\~na$^{1}$, C\'esar Gracia$^{1}$, Amparo Latorre$^{2,3,4}$, Andr\'es Moya$^{2,3,4}$ \& Carlos P. Garay$^1$}

\begin{document}

\maketitle

\begin{affiliations}
 \item Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain.
 \item FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain.
 \item Cavanilles Institute of Biodiversity and Evolutionary Biology, Univ. de Valencia, 46980, Spain. 
 \item CIBER en Epidemiolog\'ia y Salud P\'ublica (CIBEResp), Madrid, Spain.
\end{affiliations}

\begin{abstract}
Temporal fluctuations in the microbiome reveal significant differences due to factors that affect the microbiota. We show that a fluctuation scaling law describes the temporal variability of the system and that a noise-induced phase transition is central in the route to disease. The universal law distinguishes healthy from sick microbiota and quantitatively characterizes the path in the phase space, which opens up its potential clinical use and other technological applications.
\end{abstract}

High throughput methods for microbial 16S ribosomal RNA gene and shotgun metagenomic sequencing (SMS) have now begun to reveal the composition of archaeal, bacterial, fungal and viral communities located both, in and on the human body. Here we present the imprints of disease in macroscopic properties of the system, by studying the temporal variability in the microbiome.

To this end, we have analysed more than 35000 time series of taxa from the gut microbiomes of 97 individuals (sampling from three up to 332 time points): healthy individuals over a long term span\cite{moving}, people with various degrees of obesity\cite{lea}, twin pairs discordant for kwashiorkor\cite{kwashiorkor}, response to diet changes\cite{diet} or to antibiotic perturbation\cite{antibiotic}, and subjects diagnosed with irritable bowel syndrome (IBS)\cite{durban}. We engineered a complete software framework and a web platform, ComplexCruncher, ready to be implemented by other users.  We summarize our dataset in ST1-6 (Tables 1-6 in supplemental material). The bacteria and archaea taxonomic assignations were obtained by analysing 16S rRNA sequences, which were clustered into operational taxonomic units (OTUs) sharing 97 \% sequence identity using QIIME\cite{qiime}. SMS data\cite{kwashiorkor} were analysed and assigned at strain level by the Livermore Metagenomic Analysis Toolkit (LMAT)\cite{lmat1, lmat2}, according to their default quality threshold. Genus, with best balance between error assignment and number of taxa, was chosen as our reference taxonomic level.

\begin{figure}[!ht]
\includegraphics[width=1.05\textwidth]{comboplot1.pdf}
\caption{Taylor's law parameter space. X-weighted power-law fits of the standard deviations versus the mean values for each bacterial genus monitored in time. We show the fit for samples from a healthy subject (Fig. A) and from a subject diagnosed with irritable bowel syndrome (Fig. B) studied in our lab\cite{durban}. We have compiled all data studied in this work in Figure C. The coloured region (dashed line) corresponds to 68\% (95\%) CL region of healthy individuals in the Taylor parameter space. Points with errors place each individual gut microbiome in the Taylor space.}
\end{figure}

We have analysed the microbiome temporal variability to extract global properties of the system. As fluctuations in total counts are plagued by systematics errors, we worked with relative abundances. Our first finding was that, in all cases, changes in relative abundances of taxa follow a universal pattern known as the fluctuation scaling law\cite{fs} or Taylor's power law\cite{taylor}, i.e., microbiota of all detected taxa follows a power law dependence between mean relative abundance $x_i$ and dispersion $\sigma_i$, $\sigma_i  = V\cdot x_i^{\beta}$. While the law is universal, spanning six orders of magnitude in the observed relative abundances, the power law (or scaling) index $\beta$ and the variability $V$ (hereafter Taylor parameters) appear to be correlated with the stability of the community and the health status of the host (see Fig. 1). Our results hint at a universal behaviour. Firstly, the variability (which corresponds to the maximum amplitude of fluctuations) is large, which suggests resilient capacity of the microbiota, and the scaling index is always smaller than one, which means that, more abundant taxa are less volatile than less abundant ones. Secondly, Taylor parameters for the microbiome of healthy individuals in different studies are compatible within estimated errors. This enables us to define the health zone in the Taylor parameter space. We can better visualize the results of individuals from different studies by standardizing their Taylor parameters, where standardization means that each parameter is subtracted by the mean value and divided by the standard deviation of the group of healthy individuals in the study (see supplementary material). The zone of health and the standardized Taylor parameters for individuals whose gut microbiota is threatened (i.e., suffering from kwashiorkor, altered diet, antibiotics, IBS) is shown in Figure 1. Children developing kwashiorkor show smaller variability than their healthy twins. A meat/fish-based diet increases the variability significantly when compared to a plant-based diet. All other cases presented increased variability, which is particularly severe, and statistically significant at more than 95\% CL, for obese patients grade III on a diet, individuals taking antibiotics or IBS--diagnosed patients. A global property emerges from all worldwide data collected: Taylor parameters characterize the statistical behaviour of microbiome changes. Our conclusions are robust to systematic errors due to taxonomic assignment.

\begin{figure}[!ht]
\includegraphics[width=1.05\textwidth]{comboplot2.pdf}
\caption{Phase space of microbiota stability. We show the Variability parameter as a function of time for the male in Caporaso's study\cite{moving} (Fig. A) and for the patient D in the antibiotics study\cite{antibiotic} (Fig. B). Periods of antibiotic treatment are shown by blue vertical lines. Microbiota states can be placed in the phase space F-V (Fig. C). The light blue (grey) shaded region corresponds to the stable (unstable) phase (the phase transition line is calculated for  $\alpha$ = $\beta$ = 0.75). We place healthy individuals (green) and individuals whose gut microbiota is threatened (antibiotics, IBS) in the phase space fitness - variability. Healthy individuals (green) show quasi-periodical variability. Individuals whose gut microbiota was threatened (red) show a phase transition to the healthy phase.}
\end{figure}

Taylor's power law has been explained in terms of various effects, all without general consensus. It can be shown to have its origin in a mathematical convergence similar to the central limit theorem, so virtually any statistical model designed to produce a Taylor law converge to a Tweedie distribution\cite{stat}, providing a mechanistic explanation based on the statistical theory of errors\cite{convergence1,convergence3}. We model the system by assuming that taxon relative abundance follows a Langevin equation with a deterministic term that captures the fitness of each taxon and a randomness term with Gaussian random noise\cite{ranking}. Both terms are modelled by power laws, with coefficients that can be interpreted as the taxon fitness $F_i$ and the variability $V$ (see Section 1 in supplemental material). Differences in variability $V$ can induce a noise-induced phase transition in relative abundances of taxa. The temporal evolution of the probability of a taxon having abundance $x_i$ given its fitness is governed by the Fokker--Planck equation. The results of solving this equation show that  the stability is best captured by fitness F and amplitude of fluctuations V phase space (see Fig. 2). 
The model predicts two phases for the gut microbiome: a stable phase with large variability that permits some changes in the relative abundances of taxa and an unstable phase with larger variability, above the phase transition, where the order of abundant taxa varies significantly with time. The microbiome of all healthy individuals was found to be in the stable phase, while the microbiome of several other individuals was shown to be in the unstable phase. In particular, individuals taking antibiotics and IBS--diagnosed patient P2 had the most severe symptoms. In this phase diagram, each microbiota state is represented by a point at its measured variability V and inferred fitness F. The model predicts high average fitness for all taxa, i.e., taxa are narrowly distributed in F. The fitness parameter has been chosen with different values for demonstrative purposes. Fitness is larger for the healthiest subjects and smaller for the IBS--diagnosed patients.

In summary, we have quantitatively characterized whether the microbiota belongs to a healthy individual or a subject corresponding to an altered or pathological state (ie, altered diet, antibiotic treatment, early gut development, diagnosed IBS). Deciphering the mechanisms of disease requires in depth knowledge of the underlying biological mechanisms. We describe here the macroscopic behavior of disease by a noise-induced phase transition with a control parameter that can be measured by the temporal variability of the microbiome. The microbiota of healthy individuals and of individuals with pathologies represent different phases separated by this noise-induced phase transition. Improved high--throughput sequencing of samples from individuals monitored over time and taxonomic assigning methods will provide a better distinction among pathologies or altered states of the microbiota.

\begin{thebibliography}{1}
\bibitem{moving} Caporaso, J.G. \textit{et al.} {\it Genome Biol.} {\bf 12,} R50 (2011).
\bibitem{lea} Faith, J.J. \textit{et al.} {\it Science} {\bf 341,} 1237439 (2013).
\bibitem{kwashiorkor} Smith M.I. \textit{et al.} {\it Science} {\bf 339,} 548-54 (2013).
\bibitem{diet} David, L.A. \textit{et al.} {\it Nature} {\bf 505,} 559-63 (2014).
\bibitem{antibiotic} Dethlefsen L., Relman D.A. {\it Proc. Nat. Acad. Sci. USA} {\bf 108,} 4554-61 (2011).
\bibitem{durban} Durban, A. \textit{et al.} {\it FEMS Microbiol. Ecol.} {\bf 86}, 581-9 (2013).
\bibitem{qiime} Caporaso, J.G. \textit{et al.} {\it Nature Methods} {\bf 7,} 335-6 (2010).
\bibitem{lmat1} Ames, S.K. \textit{et al.} {\it Bioinformatics} {\bf 29,} 2253-2260 (2013).
\bibitem{lmat2} Ames, S.K. \textit{et al.} {\it Genome Research} {doi:10.1101/gr.184879.114} (2015).
\bibitem{fs} Eisler, Z., Bartos, I., Kertesz, J. {\it Adv. Phys.} {\bf 57,} 85 (2008).
\bibitem{taylor} Taylor, L.R. {\it Nature} {\bf 189,} 732-35 (1961).
\bibitem{stat} Jorgensen, B., Martinez, J.R. {\it Scand. J. Statist..} {\bf 21,} 223-243 (1994).
\bibitem{convergence1} Fronczak, A., Fronczak, P. {\it Phys. Rev. E} {\bf 81,} 066112 (2010).
\bibitem{convergence3} Kendal, W.S., Jorgensen, B. {\it Phys. Rev. E} {\bf 84,} 066120 (2011).
\bibitem{ranking} Blumm, N. \textit{et al.} {\it Phys. Rev. Lett.} {\bf 109,} 128701 (2012).
\end{thebibliography}

\end{document}

### 2016-04-15 Science/science_CC.tex
Include previous releases for different publications
On branch master
	new file:   MolSysBio/CC-suppl_molsysbio.docx
	new file:   MolSysBio/molsysbio_CC.docx
	new file:   NatBiotech/CC-suppl_NatBiotech.pdf
	new file:   NatBiotech/CC-suppl_NatBiotech.tex
	new file:   NatBiotech/CC_NatBiotech.pdf
	new file:   NatBiotech/CC_NatBiotech.tex
	new file:   NatBiotech/lettertoeditor_NatBiotech.docx
	new file:   Nature/CC-suppl_sub.pdf
	new file:   Nature/CC_sub.pdf
	new file:   Nature/lettertoeditor2.docx
	new file:   Science/lettertoeditor_sub.docx
	new file:   Science/science_CC.pdf
	new file:   Scie

% Use only LaTeX2e, calling the article.cls class and 12-point type.

\documentclass[12pt]{article}

% Users of the {thebibliography} environment or BibTeX should use the
% scicite.sty package, downloadable from *Science* at
% www.sciencemag.org/about/authors/prep/TeX_help/ .
% This package should properly format in-text
% reference calls and reference-list numbers.

\usepackage{scicite}
\usepackage{graphicx}

% Use times if you have the font installed; otherwise, comment out the
% following line.

\usepackage{times}

% The preamble here sets up a lot of new/revised commands and
% environments.  It's annoying, but please do *not* try to strip these
% out into a separate .sty file (which could lead to the loss of some
% information when we convert the file to other formats).  Instead, keep
% them in the preamble of your main LaTeX source file.

% The following parameters seem to provide a reasonable page setup.

\topmargin 0.0cm
\oddsidemargin 0.2cm
\textwidth 16cm 
\textheight 21cm
\footskip 1.0cm

%The next command sets up an environment for the abstract to your paper.

\newenvironment{sciabstract}{%
\begin{quote} \bf}
{\end{quote}}

% If your reference list includes text notes as well as references,
% include the following line; otherwise, comment it out.

\renewcommand\refname{References and Notes}

% The following lines set up an environment for the last note in the
% reference list, which commonly includes acknowledgments of funding,
% help, etc.  It's intended for users of BibTeX or the {thebibliography}
% environment.  Users who are hand-coding their references at the end
% using a list environment such as {enumerate} can simply add another
% item at the end, and it will be numbered automatically.

\newcounter{lastnote}
\newenvironment{scilastnote}{%
\setcounter{lastnote}{\value{enumiv}}%
\addtocounter{lastnote}{+1}%
\begin{list}%
{\arabic{lastnote}.}
{\setlength{\leftmargin}{.22in}}
{\setlength{\labelsep}{.5em}}}
{\end{list}}

% Include your paper's title here

\title{Microbiota: are you sick?}

% Place the author information here.  Please hand-code the contact
% information and notecalls; do *not* use \footnote commands.  Let the
% author contact information appear immediately below the author names
% as shown.  We would also prefer that you don't change the type-size
% settings shown here.

\author
{Jose Manuel Mart\'i,$^{1}$ Daniel M. Mart\'inez,$^{2}$ Manuel Pe\~na$^{1}$, C\'esar Gracia$^{1}$, \\
Amparo Latorre$^{2,3,4}$, Andr\'es Moya$^{2,3,4}$ \& Carlos P. Garay$^{1,\ast}$\\
\\
\normalsize{$^{1}$Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain.}\\
\normalsize{$^{2}$FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain.}\\
\normalsize{$^{3}$Cavanilles Institute of Biodiversity and Evolutionary Biology, Univ. de Valencia, 46980, Spain.}\\
\normalsize{$^{4}$CIBER en Epidemiolog\' ia y Salud P\'ublica (CIBEResp), Madrid, Spain.}\\
\\
\normalsize{$^\ast$To whom correspondence should be addressed; E-mail:  penya@ific.uv.es}
}

% Include the date command, but leave its argument blank.

\date{}

%%%%%%%%%%%%%%%%% END OF PREAMBLE %%%%%%%%%%%%%%%%

\begin{document}

% Double-space the manuscript.

\baselineskip24pt

% Make the title.

\maketitle

% Place your abstract within the special {sciabstract} environment.

\begin{sciabstract}
Human microbiota plays an important role in determining changes from health to disease. Increasing research activity is dedicated to understand its diversity and variability. 
We analyse 16S rRNA and whole genome sequencing (WGS) data from the gut microbiota of 97 individuals monitored in time. Temporal fluctuations in the microbiome reveal 
significant differences due to factors that affect the microbiota such as dietary changes, antibiotic intake, early gut development or disease. Here we show that a fluctuation 
scaling law describes the temporal variability of the system and that a noise-induced phase transition is central in the route to disease. The universal law distinguishes healthy 
from sick microbiota and quantitatively characterizes the path in the phase space, which opens up its potential clinical use and, more generally, other technological 
applications where microbiota plays an important role. 
\end{sciabstract}

% In setting up this template for *Science* papers, we've used both
% the \section* command and the \paragraph* command for topical
% divisions.  Which you use will of course depend on the type of paper
% you're writing.  Review Articles tend to have displayed headings, for
% which \section* is more appropriate; Research Articles, when they have
% formal topical divisions at all, tend to signal them with bold text
% that runs into the paragraph, for which \paragraph* is the right
% choice.  Either way, use the asterisk (*) modifier, as shown, to
% suppress numbering.

\section*{Introduction}

The desire to understand the factors that influence human health and cause diseases has always been one of the major driving forces of biological research. Modern high-throughput sequencing and bioinformatic tools provide a powerful means of understanding how the human microbiome contributes to health and its potential as a target for therapeutic interventions. High throughput methods for microbial 16S ribosomal RNA gene and WGS have now begun to reveal the composition of archaeal, bacterial, fungal and viral communities located both, in and on the human body. Biology has recently acquired new technological and conceptual tools to investigate, model and understand living organisms at the system level, thanks to the spectacular progress in quantitative techniques, large-scale measurement methods and the integration of experimental and computational approaches. Systems Biology has mostly been devoted to the study of well-characterized model organisms but, since the early days of the Human Genome Project it has become clear that applications of system-wide approaches to Human Biology would bring huge opportunities in Medicine. Here we present the imprints of disease in macroscopic properties of the system, by studying the temporal variability in the microbiome.

To this end, we have analysed more than 35000 time series of taxa from the gut microbiomes of 97 individuals (sampling from three up to 332 time points), obtained from publicly available high throughput sequencing data on: healthy individuals over a long term span\cite{moving}, people with various degrees of obesity\cite{lea}, twin pairs discordant for kwashiorkor\cite{kwashiorkor}, response to diet changes\cite{diet} or to antibiotic perturbation\cite{antibiotic}, and subjects diagnosed with irritable bowel syndrome (IBS)\cite{durban}. 
We engineered a complete software framework and a web platform, ComplexCruncher, ready to be implemented by other users.  
We summarize our dataset in ST1-6 (Tables 1-6 in suplemental material). The bacteria and archaea taxonomic assignations were obtained by analysing 16S rRNA sequences, which were clustered into operational taxonomic units (OTUs) sharing 97 \% sequence identity using QIIME\cite{qiime}. WGS data\cite{kwashiorkor} were analysed and assigned at strain level by the Livermore Metagenomic Analysis Toolkit (LMAT)\cite{lmat}, according to their default quality threshold. Genus, with best balance between error assignment and number of taxa, was chosen as our reference taxonomic level. We have verified that our conclusions are not significantly affected by selecting family or species as the reference taxonomic level (see Figure 1 in supplemental material).

\begin{figure}[!ht]
\includegraphics[width=1.05\textwidth]{comboplot1.pdf}
\caption{Taylor's law parameter space. X-weighted power-law fits of the standard deviations versus the mean values for each bacterial genus monitored in time. We show the fit for samples from a healthy subject (Figure A) and from a subject diagnosed with irritable bowel syndrome (Figure B) studied in our lab\cite{durban}. We have compiled all data studied in this work in Figure C. The coloured region (dashed line) corresponds to 68\% (95\%) CL region of healthy individuals in the Taylor parameter space. Points with errors place each individual gut microbiome in the Taylor space. Note that the parameters have been standardized (stdev units) to the healthy group in each study for demonstrative purposes.}
\end{figure}

We have analysed the microbiome temporal variability to extract global properties of the system. As fluctuations in total counts are plagued by systematic errors we worked on temporal variability of relative abundances for each taxon. Our first finding was that, in all cases, changes in relative abundances of taxa follow a universal pattern known as the fluctuation scaling law\cite{fs} or Taylor's  power law\cite{taylor}, i.e., microbiota of all detected taxa follows a power law dependence between mean relative abundance $x_i$ and dispersion $\sigma_i$, $\sigma_i  = V\cdot x_i^{\beta}$. While the law is universal, spanning six orders of magnitude in the observed relative abundances, the power law (or scaling) index $\beta$ and the variability $V$ (hereafter Taylor parameters) appear to be correlated with the stability of the community and the health status of the host, which is the main finding in this letter (see Figure 1). Taylor parameters describing the temporal variability of the gut microbiome in our sampled individuals are shown in ST1-6. Our results hint at a universal behaviour. Firstly, the variability (which corresponds to the maximum amplitude of fluctuations) is large, which suggests resilient capacity of the microbiota, and the scaling index is always smaller than one, which means that, more abundant taxa are less volatile than less abundant ones. Secondly, Taylor parameters for the microbiome of healthy individuals in different studies are compatible within estimated errors. This enables us to define the health zone in the Taylor parameter space. We can better visualize the results of individuals from different studies by standardizing their Taylor parameters, where standardization means that each parameter is subtracted by the mean value and divided by the standard deviation of the group of healthy individuals in the study (see Section 12 and ST1-6). The zone of health and the standardized Taylor parameters for individuals whose gut microbiota is threatened (i.e., suffering from kwashiorkor, altered diet, antibiotics, IBS) is shown in Figure 1. Children developing kwashiorkor show smaller variability than their healthy twins. A meat/fish-based diet increases the variability significantly when compared to a plant-based diet. All other cases presented increased variability, which is particularly severe, and statistically significant at more than 95\% CL, for obese patients grade III on a diet, individuals taking antibiotics or IBS--diagnosed patients. A global property emerges from all worldwide data collected: Taylor parameters characterize the statistical behaviour of microbiome changes. We have verified that our conclusions are robust to systematic errors due to taxonomic assignment.

\begin{figure}[!ht]
\includegraphics[width=1.05\textwidth]{comboplot2.pdf}
\caption{Phase space of microbiota stability. We show the Variability parameter as a function of time for the male in Caporaso's study\cite{moving} (Figure A) and for the patient D in the antibiotics study\cite{antibiotic} (Figure B). Periods of antibiotic treatment are shown by blue vertical lines. Microbiota states can be placed in the phase space F-V (Figure C). The light blue shaded region corresponds to the stable phase, while the grey shaded region is the unstable phase (the phase transition line is calculated for  $\alpha$ = $\beta$ = 0.75). We place healthy individuals (green) and individuals whose gut microbiota is threatened (antibiotics, IBS) in the phase space fitness - variability. Gut microbiota of healthy individuals over a long term span show a quasi--periodical variability (central period is ten days). We show that taking antibiotics (AB1 and AB2 correspond to first and second treatment respectively) induces a phase transition in the gut microbiota, which impacts its future changes. We also show an IBS--diagnosed patient transiting from the unstable to the stable phase.}
\end{figure}

Taylor's power law has been explained in terms of various effects, all without general consensus. It can be shown to have its origin in a mathematical convergence similar to the central limit theorem, so virtually any statistical model designed to produce a Taylor law converge to a Tweedie distribution\cite{stat}, providing a mechanistic explanation based on the statistical theory of errors\cite{convergence1,convergence2,convergence3}. To unveil the generic mechanisms that drive different scenarios in the $\beta$--V space, we model the system by assuming that taxon relative abundance follows a Langevin equation with a deterministic term that captures the fitness of each taxon and a randomness term with Gaussian random noise\cite{ranking}. Both terms are modelled by power laws, with coefficients that can be interpreted as the taxon fitness $F_i$ and the variability $V$ (see Section 1 in supplemental material). When $V$ is sufficiently low, abundances are stable in time.  Differences in variability $V$ can induce a noise-induced phase transition in relative abundances of taxa. The temporal evolution of the probability of a taxon having abundance $x_i$ given its fitness is governed by the Fokker--Planck equation. The results of solving this equation show that  the stability is best captured by fitness F and amplitude of fluctuations V phase space (see Figure 2). 
The model predicts two phases for the gut microbiome: a stable phase with large variability that permits some changes in the relative abundances of taxa and an unstable phase with larger variability, above the phase transition, where the order of abundant taxa varies significantly with time. The microbiome of all healthy individuals was found to be in the stable phase, while the microbiome of several other individuals was shown to be in the unstable phase. In particular, individuals taking antibiotics and IBS--diagnosed patient P2 had the most severe symptoms. In this phase diagram, each microbiota state is represented by a point at its measured variability V and inferred fitness F. The model predicts high average fitness for all taxa, i.e., taxa are narrowly distributed in F. The fitness parameter has been chosen with different values for demonstrative purposes. Fitness is larger for the healthiest subjects and smaller for the IBS--diagnosed patients.

In summary, we have quantitatively characterized whether the microbiota belongs to a healthy individual or a subject corresponding to an altered or pathological state (ie, altered diet, antibiotic treatment, early gut development, diagnosed IBS). Deciphering the mechanisms of disease requires in depth knowledge of the underlying biological mechanisms. We describe here the macroscopic behavior of disease by a noise-induced phase transition with a control parameter that can be measured by the temporal variability of the microbiome. The microbiota of healthy individuals and of individuals with pathologies represent different phases separated by this noise-induced phase transition. Improved high--throughput sequencing of samples from individuals monitored over time and taxonomic assigning methods will provide a better distinction among pathologies or altered states of the microbiota.

\begin{thebibliography}{1}
\bibitem{moving} Caporaso, J.G. et al. Moving pictures of the human microbiome. {\it Genome Biol.} {\bf 12,} R50 (2011).
\bibitem{lea} Faith, J.J. et al. The long-term stability of the human gut microbiota. {\it Science} {\bf 341,} 1237439 (2013).
\bibitem{kwashiorkor} Smith M.I. et al. Gut microbiomes of Malawian twin pairs discordant for kwashiorkor. {\it Science} {\bf 339,} 548-54 (2013).
\bibitem{diet} David, L.A. et al. Diet rapidly and reproducibly alters the human gut microbiome. {\it Nature} {\bf 505,} 559-63 (2014).
\bibitem{antibiotic} Dethlefsen L., Relman D. A. Incomplete recovery and individualized responses of the human distal gut microbiota to repeated antibiotic perturbation. {\it Proc. Nat. Acad. Sci. USA} {\bf 108,} 4554-61 (2011).
\bibitem{durban} Durban, A. \textit{et al.} Structural alterations of faecal and mucosa-associated bacterial communities in irritable bowel syndrome. {\it FEMS Microbiol. Ecol.} {\bf 86}, 581-9 (2013).
\bibitem{qiime} Caporaso, J.G. et al. QIIME allows analysis of high-throughput community sequencing data. {\it Nature Methods} {\bf 7,} 335-6 (2010).
\bibitem{lmat} Ames, S.K. et al. Scalable metagenomic taxonomy classification usng a reference genome database. {\it Bioinformatics} {\bf 29,} 2253-2260 (2013).
\bibitem{fs} Eisler,Z., Bartos,I., Kertesz,J. Fluctuation scaling in complex systems: Taylor's law and beyond. {\it Adv. Phys.} {\bf 57,} 85 (2008).
\bibitem{taylor} Taylor, L.R. Aggregation, Variance and the mean. {\it Nature} {\bf 189,} 732-35 (1961).
\bibitem{stat} Jorgensen,B., Martinez,J.R., Tsao,M. Asymptotic behaviour of the variance function. {\it Scand. J. Statist..} {\bf 21,} 223-243 (1994).
\bibitem{convergence1} Fronczak,A., Fronczak,P. Origins of Taylor's power law for fluctuation scaling in complex systems. {\it Phys. Rev. E} {\bf 81,} 066112 (2010).
\bibitem{convergence2} Kendal, W.S., Jorgensen,B. Taylor's power law and fluctuation scaling explained by a central-limit-like convergence. {\it Phys. Rev. E} {\bf 83,} 066115 (2011).
\bibitem{convergence3} Kendal, W.S., Jorgensen,B. Tweedie convergence: A mathematical basis for Taylor's power law. {\it Phys. Rev. E} {\bf 84,} 066120 (2011).
\bibitem{ranking} Blumm, N. et al. Dynamics of ranking processes in complex systems. {\it Phys. Rev. Lett.} {\bf 109,} 128701 (2012).
\end{thebibliography}

\end{document}

### 2016-04-15 Science/science_CC_suppl.tex
Include previous releases for different publications
On branch master
	new file:   MolSysBio/CC-suppl_molsysbio.docx
	new file:   MolSysBio/molsysbio_CC.docx
	new file:   NatBiotech/CC-suppl_NatBiotech.pdf
	new file:   NatBiotech/CC-suppl_NatBiotech.tex
	new file:   NatBiotech/CC_NatBiotech.pdf
	new file:   NatBiotech/CC_NatBiotech.tex
	new file:   NatBiotech/lettertoeditor_NatBiotech.docx
	new file:   Nature/CC-suppl_sub.pdf
	new file:   Nature/CC_sub.pdf
	new file:   Nature/lettertoeditor2.docx
	new file:   Science/lettertoeditor_sub.docx
	new file:   Science/science_CC.pdf
	new file:   Scie

--- 2016-04-15 NatBiotech/CC-suppl_NatBiotech.tex
@@
-% !TEX TS-program = pdflatex
-% !TEX encoding = UTF-8 Unicode
+% Use only LaTeX2e, calling the article.cls class and 12-point type.
 
-%% Template for a preprint Letter or Article for submission
-%% to the journal Nature.
-%% Written by Peter Czoschke, 26 February 2004
-%%
-\documentclass{nature} % Commented includegraphics trick in /usr/local/texlive/2014/texmf-dist/tex/latex/nature/nature.cls
+\documentclass[12pt]{article}
+
+% Users of the {thebibliography} environment or BibTeX should use the
+% scicite.sty package, downloadable from *Science* at
+% www.sciencemag.org/about/authors/prep/TeX_help/ .
+% This package should properly format in-text
+% reference calls and reference-list numbers.
+
+\usepackage{scicite}
 \usepackage{graphicx}
@@
 \usepackage[usenames,dvipsnames,table]{xcolor}
-\usepackage[nolists,tablesfirst,nomarkers]{endfloat}
+\usepackage[nolists,tablesfirst, nomarkers]{endfloat}
+
+% Use times if you have the font installed; otherwise, comment out the
+% following line.
+
+\usepackage{times}
+
+% The preamble here sets up a lot of new/revised commands and
+% environments.  It's annoying, but please do *not* try to strip these
+% out into a separate .sty file (which could lead to the loss of some
+% information when we convert the file to other formats).  Instead, keep
+% them in the preamble of your main LaTeX source file.
+
+% The following parameters seem to provide a reasonable page setup.
+
+\topmargin 0.0cm
+\oddsidemargin 0.2cm
+\textwidth 16cm 
+\textheight 21cm
+\footskip 1.0cm
+
+%The next command sets up an environment for the abstract to your paper.
+
+\newenvironment{sciabstract}{%
+\begin{quote} \bf}
+{\end{quote}}
+
+% If your reference list includes text notes as well as references,
+% include the following line; otherwise, comment it out.
 
@@
 \newcommand{\unclassrate}[0]{\un{kpb/s/core}}
-
-%% Roman footnote numbering style to not collide with bibliographic cites: 
+\renewcommand\refname{References and Notes}
 \renewcommand{\thefootnote}{\Roman{footnote}}
 
-%% Change floats name for supplemental material
-\renewcommand{\figurename}{Supplementary Figure}
-\renewcommand{\tablename}{Supplementary Table}
+% The following lines set up an environment for the last note in the
+% reference list, which commonly includes acknowledgments of funding,
+% help, etc.  It's intended for users of BibTeX or the {thebibliography}
+% environment.  Users who are hand-coding their references at the end
+% using a list environment such as {enumerate} can simply add another
+% item at the end, and it will be numbered automatically.
 
-%% make sure you have the nature.cls and naturemag.bst files where
-%% LaTeX can find them
-\bibliographystyle{naturemag}
+\newcounter{lastnote}
+\newenvironment{scilastnote}{%
+\setcounter{lastnote}{\value{enumiv}}%
+\addtocounter{lastnote}{+1}%
+\begin{list}%
+{\arabic{lastnote}.}
+{\setlength{\leftmargin}{.22in}}
+{\setlength{\labelsep}{.5em}}}
+{\end{list}}
 
-%%% HYPERREF
-\usepackage[pdftex,
-	pdftitle={Supplemental material of ``Microbiota: are you sick?''},
-	pdfauthor={Jose Manuel Marti, Daniel M. Martinez, Manuel Peña, Cesar Gracia, Amparo Latorre, Andres Moya, Carlos P. Garay},
-	pdfsubject={Nature Biotechnology},
-	pdfkeywords={Supplemental material},
-	bookmarksopen,
-	colorlinks=true,
-	urlcolor={red},
-	bookmarks=true]{hyperref}
+% Include your paper's title here
 
-%% Title
 \title{Supplemental material of ``Microbiota: are you sick?''}
 
-%% Notice placement of commas and superscripts and use of &
-%% in the author list
-\author{Jose Manuel Mart\'i$^{1}$, Daniel M. Mart\'inez$^{1,2}$, Manuel Pe\~na$^{1}$, C\'esar Gracia$^{1}$, Amparo Latorre$^{2,3,4}$, Andr\'es Moya$^{2,3,4}$ \& Carlos P. Garay$^1$}
+% Place the author information here.  Please hand-code the contact
+% information and notecalls; do *not* use \footnote commands.  Let the
+% author contact information appear immediately below the author names
+% as shown.  We would also prefer that you don't change the type-size
+% settings shown here.
+
+\author
+{Jose Manuel Mart\'i,$^{1}$ Daniel M. Mart\'inez,$^{2}$ Manuel Pe\~na$^{1}$, C\'esar Gracia$^{1}$, \\
+Amparo Latorre$^{2,3,4}$, Andr\'es Moya$^{2,3,4}$ \& Carlos P. Garay$^{1,\ast}$\\
+\\
+\normalsize{$^{1}$Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain.}\\
+\normalsize{$^{2}$FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain.}\\
+\normalsize{$^{3}$Cavanilles Institute of Biodiversity and Evolutionary Biology, Univ. de Valencia, 46980, Spain.}\\
+\normalsize{$^{4}$CIBER en Epidemiolog\' ia y Salud P\'ublica (CIBEResp), Madrid, Spain.}\\
+\\
+\normalsize{$^\ast$To whom correspondence should be addressed; E-mail:  penya@ific.uv.es}
+}
+
+% Include the date command, but leave its argument blank.
+
+\date{}
+
+%%%%%%%%%%%%%%%%% END OF PREAMBLE %%%%%%%%%%%%%%%%
 
@@
 
+% Double-space the manuscript.
+
+\baselineskip24pt
+
+% Make the title.
+
 \maketitle
 
-\begin{affiliations}
- \item Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain.
- \item FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain.
- \item Cavanilles Institute of Biodiversity and Evolutionary Biology, Univ. de Valencia, 46980, Spain. 
- \item CIBER en Epidemiolog\'ia y Salud P\'ublica (CIBEResp), Madrid, Spain.
-\end{affiliations}
+% In setting up this template for *Science* papers, we've used both
+% the \section* command and the \paragraph* command for topical
+% divisions.  Which you use will of course depend on the type of paper
+% you're writing.  Review Articles tend to have displayed headings, for
+% which \section* is more appropriate; Research Articles, when they have
+% formal topical divisions at all, tend to signal them with bold text
+% that runs into the paragraph, for which \paragraph* is the right
+% choice.  Either way, use the asterisk (*) modifier, as shown, to
+% suppress numbering.
 
-\section{Model} \label{sec:model}
+\section*{Model} \label{sec:model}
 
@@
 \subsection{Metagenomic sequences processing}
-Metagenomic shotgun (and 16S too) sequences were analyzed with LMAT (Livermore Metagenomics Analysis Toolkit) software package\cite{LMAT1, LMAT2} (version 1.2.4, with Feb'15 release of data base \emph{LMAT-Grand}).
+Metagenomic shotgun (and 16S too) sequences were analyzed with LMAT (Livermore Metagenomics Analysis Toolkit) software package\cite{LMAT} (version 1.2.4, with Feb'15 release of data base \emph{LMAT-Grand}).
 LMAT was run using a Bull shared-memory node belonging to the team's HPC (high performance computing) cluster.
@@
 The computing node was supplied with a RAID-0 (striping) scratch disk area.
-We used the ``Grand'' database\footnote{In this context, ``Grand''\cite{LMAT2} refers to a huge database that contains k-mers from all viral, prokaryote, fungal and protist genomes present in the NCBI database, plus Human reference genome (hg19), plus GenBank Human, plus the 1000 Human Genomes Project (HGP).
-This represent about 31.75 billion k-mers occupying 457.62 GB.}, release Feb'15, provided by the LMAT team\cite{LMAT2}.
+We used the ``Grand'' database\footnote{In this context, ``Grand'' refers to a huge database that contains k-mers from all viral, prokaryote, fungal and protist genomes present in the NCBI database, plus Human reference genome (hg19), plus GenBank Human, plus the 1000 Human Genomes Project (HGP).
+This represent about 31.75 billion k-mers occupying 457.62 GB.}, release Feb'15, provided by the LMAT team.
 Previously to any calculation, the full database was loaded in the NVRAM.
@@
 We selected genus as taxonomic level for the subsequent steps of our work.
-To ensure that, between adjacent taxonomic levels, there were not crucial differences which could still be of relevance after standardization (see Supplementary Section \ref{sec:stan}), we tested two different datasets.
+In order to ensure that, between adjacent taxonomic levels, there were not crucial differences which could still be of relevance after standardization (see Section \ref{sec:stan}), we tested two different data sets.
 In the former, the antibiotics study\cite{antibiotic} with 16S data, we tested the differences between genus and family levels.
-The latter dataset tested was the kwashiorkor twins study\cite{kwashiorkor} for both genus and species taxonomic levels.
+The latter dataset tested was the kwashiorkor discordant twins study\cite{kwashiorkor} for both genus and species taxonomic levels.
 The Supplementary Figures \ref{fig:taxlev1} (overview) and \ref{fig:taxlev2} (detail) plot the comparison between studies (and so, 16S and SMS) and between adjacent taxonomic levels.
@@
 
-We have studied the time dependence of the variability $V$ and power law index $\beta$ (see Supplementary Section \ref{sec:model}) by using a sliding window approach.
+We have studied the time dependence of the variability $V$ and power law index $\beta$ (see Section \ref{sec:model}) by using a sliding window approach.
 The total number of time points are divided in subsets of five points, where next subset is defined by adding next time sampling and by eliminating the earliest one.
 Both parameters were calculated for each subset against the average time lapse.
-Supplementary Figure \ref{fig:tempevo1} shows the variability  $V$ as a function of time for the largest sampling: two individuals in the Caporaso's study\cite{moving} corresponding to the gut microbiota of a male (upper plot) and a female (lower plot).
-Supplementary Figure \ref{fig:tempevo2} shows the time evolution of $V$ for patient P2 of the IBS study\cite{IBS} (upper plot) and patient D in the antibiotics study\cite{antibiotic} (lower plot).
+Figure \ref{fig:tempevo1} shows the variability  $V$ as a function of time for the largest sampling: two individuals in the Caporaso's study\cite{moving} corresponding to the gut microbiota of a male (upper plot) and a female (lower plot).
+Figure \ref{fig:tempevo2} shows the time evolution of $V$ for patient P2 of the IBS study\cite{IBS} (upper plot) and patient D in the antibiotics study\cite{antibiotic} (lower plot).
 
@@
 The PEP8 compliant code of \CC\ could be executed under a Python 2.7 interpret, but a 3.X interpret is strongly advised.
-The complete distribution of \CC\ has the code signature shown in Supplementary Table \ref{tab:cloc}, obtained through the \emph{CLOC}\footnote{\emph{Count Lines of Code} is a perl script that counts blank lines, comment lines, and physical lines of source code.} tool release 1.62.
-Supplementary Figure \ref{fig:CC} shows the code layout of the complete distribution of \CC.
+The complete distribution of \CC\ has the code signature shown in the Table \ref{tab:cloc}, obtained through the \emph{CLOC}\footnote{\emph{Count Lines of Code} is a perl script that counts blank lines, comment lines, and physical lines of source code.} tool release 1.62.
+Figure \ref{fig:CC} shows the code layout of the complete distribution of \CC.
 
@@
         \item If $\mathrm{AIC_{c_{NLR}}} - \mathrm{AIC_{c_{LLR}}} > 2$, the assumption of lognormal error is favoured compared to normal error, so proceed with the results obtained from the LLR fit.
-        \item If $\left|\mathrm{AIC_{c_{NLR}}} - \mathrm{AIC_{c_{LLR}}}\right| \leq 2$, no model is favoured, so proceed with model averaging:
+        \item If $\left|\mathrm{AIC_{c_{NLR}}} - \mathrm{AIC_{c_{LLR}}}\right| \leq 2$, no model is favoured, so proceed with averaging:
         \begin{eqnarray*}
@@
 
-The colour code of the RSI percentage text in the rank plot of \CC\ is chosen following the first condition satisfied from those shown in Supplementary Table \ref{tab:RSI} (see page \pageref{tab:RSI}).
+The colour code of the RSI percentage text in the rank plot of \CC\ is chosen following the first condition satisfied from those shown in Table \ref{tab:RSI} (see page \pageref{tab:RSI}).
 
@@
 The subdirectory \texttt{fits} under the results directory contains the fit plots for each data set.
-Both for the unweighted fit (detailed in the Supplementary Section \ref{sec:unw}) and for the X-weighted fit (detailed in the Supplementary Section \ref{sec:X-w}), two plots are generated: the former with logarithmic scale, the latter with lineal scale.
-Supplementary Figures \ref{fig:unwFit} and \ref{fig:X-wFit} show examples of the unweighted and X-weighted fits, respectively.
-Additionally, for the unweighted fit, a complete residues analysis is performed, and a 4-in-1 figure is generated as shown in Supplementary Figure \ref{fig:unwRes} (corresponding to the fit of Supplementary Fig.
-\ref{fig:unwFit}).
+Both for the unweighted fit (detailed in the section \ref{sec:unw}) and for the X-weighted fit (detailed in the section \ref{sec:X-w}), two plots are generated: the former with logarithmic scale, the latter with lineal scale.
+Figure \ref{fig:unwFit} shows an example of the unweighted fit plots and Figure \ref{fig:X-wFit} does the same with the X-weighted fit plots.
+Additionally, for the unweighted fit, a complete residues analysis is performed, and a 4-in-1 figure is generated as shown in Figure \ref{fig:unwRes}, corresponding to the fit of Figure \ref{fig:unwFit}.
 Among other tests, it allows to check for normality and homoscedasticity of the residues.
@@
 	\item[$-$Absolute frequencies plot]: This plot is useful to visually assess the validity of the time points in terms of the accumulated absolute frequency of the elements (taxa), since absolute frequencies far (much higher or much lower) from those typically observed could mean a sampling problem.
-As an example, Supplementary Figure \ref{fig:histAFP} shows this histogram for the pre-treatment data (first 7 times) of patient ``D'' in the antibiotics study\cite{antibiotic}.
+As an example, Figure \ref{fig:histAFP} shows this histogram for the pre-treatment data (first 7 times) of patient ``D'' in the antibiotics study\cite{antibiotic}.
 	\item[$-$2D deviation plot]: The 2D semi-logarithmic histogram representing deviations from the mean versus the mean itself, is a useful tool in the analysis of the stability of ranking processes in complex systems\cite{ranking}.
-Supplementary Figure \ref{fig:hist2D} shows this plot for the data used in the fit shown in Supplementary Figure \ref{fig:unwFit}.
+Figure \ref{fig:hist2D} shows this plot for the data used in the fit shown in Figure \ref{fig:unwFit}.
 	\item[$-$Zero relative frequency plot]: We could define the ZRF (Zero Relative Frequency, thereby ranging from 0 to 1) of an element (taxon) as the portion of times where it is zero, i.e., it is not found.
@@
 In this respect, the bar over $0.0$ counts the number of elements (taxa) that are present at every time point of the data set (aka ``core''), while the bar over $1.0$ would count the number of elements (taxa) that are never found (this bar never appears because all these ``null'' elements are automatically filtered by the code).
-Supplementary Figure \ref{fig:histZRF} shows an example of this plot.
+Figure \ref{fig:histZRF} shows an example of this plot.
 There, we can see that $12$ taxa are present at all the time points of the time series while $9$ taxa basically appear only once.
@@
 For these calculations, the data set is not normalized to avoid entering an additional constraint.
-As an example, Supplementary Figure \ref{fig:corrElm} shows this matrix for the ``core'' elements (taxa) present in the pre-treatment data (first seven times) of patient ``D'' in the antibiotics study\cite{antibiotic}.
+As an example, Figure \ref{fig:corrElm} shows this matrix for the ``core'' elements (taxa) present in the pre-treatment data (first seven times) of patient ``D'' in the antibiotics study\cite{antibiotic}.
 	\item[$-$Times correlation matrix]: This plot presents a correlation matrix among the time points of the data set, calculated with the elements (taxa) as independent variable.
 Again, the data set is not normalized.
-Supplementary Figure \ref{fig:corrTim} shows this matrix for the ``core'' elements (taxa) present in pre-treatment data (first seven times) of patient ``D'' in the antibiotics study\cite{antibiotic}.
-	\item[$-$Rank dynamics and stability plot]: This plot shows the variation in the rank with time for the most dominant elements (taxa) and their calculated RSI, as discussed in Supplementary Section \ref{sec:RSI}.
-Supplementary Figure \ref{fig:corrank} shows this plot for the elements (taxa) in the data used in the fit shown in Supplementary Figure \ref{fig:unwFit}.
+Figure \ref{fig:corrTim} shows this matrix for the ``core'' elements (taxa) present in pre-treatment data (first seven times) of patient ``D'' in the antibiotics study\cite{antibiotic}.
+	\item[$-$Rank dynamics and stability plot]: This plot shows the variation in the rank with time for the most dominant elements (taxa) and their calculated RSI, as discussed in Section \ref{sec:RSI}.
+Figure \ref{fig:corrank} shows this plot for the elements (taxa) in the data used in the fit shown in Figure \ref{fig:unwFit}.
 \end{description}
@@
 The advanced option flags are summarized here:
-\vspace{-5mm} %%% CAUTION: FORCED LAYOUT
 \begin{description}
@@
 The input data are provided by two simultaneously available ways:
-\vspace{-5mm} %%% CAUTION: FORCED LAYOUT
 \begin{enumerate}
@@
 $$\omega_i = \frac{\frac{1}{\sigma^2_{V_i}}}{\sum_i^h\frac{1}{\sigma^2_{V_i}}}$$
-being $\sigma_{V_i}$ the estimation of the uncertainty in $V_i$ obtained together with $V_i$ from the X-weighted power-law fit described in Supplementary Section \ref{sec:X-w}, for healthy individuals.
+being $\sigma_{V_i}$ the estimation of the uncertainty in $V_i$ obtained together with $V_i$ from the X-weighted power-law fit described in Section \ref{sec:X-w}, for healthy individuals.
 
@@
 \section{\CC\ interface with LMAT} \label{sec:lmat}
-With the aim of adapting LMAT (Livermore Metagenomics Analysis Toolkit) software\cite{LMAT1, LMAT2} for its easy use in time series data, like the ones from the Kwashiorkor paper\cite{kwashiorkor}, we developed Python front and back-end scripts, which are supplied in the directory \texttt{tools} of the \CC\ package.
+With the aim of adapting LMAT (Livermore Metagenomics Analysis Toolkit) software\cite{LMAT} for its easy use in time series data, like the ones we have analyzed in this paper\cite{kwashiorkor}, we developed Python front and back-end scripts.
+They are supplied in the directory \texttt{tools} of the \CC\ package.
 
@@
 \newpage
-\section*{Supplemental material references}
-\begin{thebibliography}{99}
-	\bibitem{ranking} Blumm, N. \textit{et al.}, Dynamics of ranking processes in complex systems. {\it Phys. Rev. Lett.} {\bf 109,} 128701 (2012).
-	\bibitem{FD} Weber, J. \textit{et al.}, Fluctuation dissipation theorem. {\it Phys. Rev.} {\bf 101}, 1620-6 (1956).
-	\bibitem{kwashiorkor} Smith M.I.  \textit{et al.}, Gut microbiomes of Malawian twin pairs discordant for kwashiorkor. {\it Science} {\bf 339,} 548-54 (2013).
-	\bibitem{FASTX} Gordon, A., Hannon, G.J., FASTX-Toolkit. FASTQ/A shortreads pre-processing tools (2010). http://hannonlab.cshl.edu/fastx\_toolkit/ (accessed 23 Feb 2015).
-	\bibitem{QIIME} Caporaso, J.G. \textit{et al.}, QIIME allows analysis of high-throughput community sequencing data. {\it Nature Methods} {\bf 7,} 335-6 (2010).
-	\bibitem{SILVA} Quast C.  \textit{et al.}, The SILVA ribosomal RNA gene database project: improved data processing and web-based tools (2013)
-	\bibitem{LMAT1} Ames, S.K.
-Hysom, D.A., Gardner, S.N., Lloyd, G.S., Gokdale, M.B., Allen, J.E., Scalable metagenomic taxonomy classification using a reference genome database.
-{\it Bioinformatics} {\bf 29,} 2253-2260 (2013).
-	\bibitem{LMAT2} Ames, S.K., Gardner, S.N., Marti, J.M., Slezak, T.R., Gokdale, M.B., Allen, J.E., Using populations of human and microbial genomes for organism detection in metagenomes {\it Genome Research} {doi:10.1101/gr.184879.114} (2015).
+\begin{thebibliography}{1}
+	\bibitem{ranking} Blumm, N. \textit{et al.} Dynamics of ranking processes in complex systems. {\it Phys. Rev. Lett.} {\bf 109,} 128701 (2012).
+	\bibitem{FD} Weber, J. \textit{et al.} Fluctuation dissipation theorem. {\it Phys. Rev.} {\bf 101}, 1620-6 (1956).
+	\bibitem{kwashiorkor} Smith M.I.  \textit{et al.} Gut microbiomes of Malawian twin pairs discordant for kwashiorkor. {\it Science} {\bf 339,} 548-54 (2013).
+	\bibitem{FASTX} Gordon, A., Hannon, G.J. FASTX-Toolkit. FASTQ/A shortreads pre-processing tools (2010). http://hannonlab.cshl.edu/fastx\_toolkit/ (accessed 23 Feb 2015).
+	\bibitem{QIIME} Caporaso, J.G. \textit{et al.} QIIME allows analysis of high-throughput community sequencing data. {\it Nature Methods} {\bf 7,} 335-6 (2010).
+	\bibitem{SILVA} Quast C.  \textit{et al.} The SILVA ribosomal RNA gene database project: improved data processing and web-based tools (2013)
+	\bibitem{LMAT} Ames, S.K.  \textit{et al.} Scalable metagenomic taxonomy classification usng a reference genome database. {\it Bioinformatics} {\bf 29,} 2253-2260 (2013).
 	\bibitem{antibiotic} Dethlefsen L., Relman D.
-A., Incomplete recovery and individualized responses of the human distal gut microbiota to repeated antibiotic perturbation.
+A.
+Incomplete recovery and individualized responses of the human distal gut microbiota to repeated antibiotic perturbation.
 {\it Proc.
@@
 USA} {\bf 108,} 4554-61 (2011).
-	\bibitem{diet} David, L.A. \textit{et al.}, Diet rapidly and reproducibly alters the human gut microbiome. {\it Nature} {\bf 505,} 559-63 (2014).
-	\bibitem{IBS} Durban, A. \textit{et al.}, Structural alterations of faecal and mucosa-associated bacterial communities in irritable bowel syndrome. {\it FEMS Microbiol. Ecol.} {\bf 86}, 581-9 (2013).
-	\bibitem{LEA} Faith, J.J. \textit{et al.}, The long-term stability of the human gut microbiota. {\it Science} {\bf 341,} 1237439 (2013).
-	\bibitem{moving} Caporaso, J.G. \textit{et al.}, Moving pictures of the human microbiome. {\it Genome Biol.} {\bf 12,} R50 (2011).
+	\bibitem{diet} David, L.A. \textit{et al.} Diet rapidly and reproducibly alters the human gut microbiome. {\it Nature} {\bf 505,} 559-63 (2014).
+	\bibitem{IBS} Durban, A. \textit{et al.} Structural alterations of faecal and mucosa-associated bacterial communities in irritable bowel syndrome. {\it FEMS Microbiol. Ecol.} {\bf 86}, 581-9 (2013).
+	\bibitem{LEA} Faith, J.J.  \textit{et al.} The long-term stability of the human gut microbiota. {\it Science} {\bf 341,} 1237439 (2013).
+	\bibitem{moving} Caporaso, J.G.  \textit{et al.} Moving pictures of the human microbiome. {\it Genome Biol.} {\bf 12,} R50 (2011).
  	\bibitem{ecology} Xiao Xiao, Ethan P.
@@
 Hooten, and Susan L.
-Durham., On the use of log-transformation vs.
+Durham.
+On the use of log-transformation vs.
 nonlinear regression for analyzing biological power laws.
@@
 	\bibitem{disR2} Nagelkerke N.J.D., A note on a general definition of the coefficient of determination. {\it Biometrika} {\bf 78}, 3, 691-692 (1991).
-	\bibitem{boot} Wu, C.F.J., Jackknife, bootstrap and other resampling methods in regression analysis. (with discussions) \textit{The Annals of Statistics} {\bf 14}: 1261-1350 (1986)
+	\bibitem{boot} Wu, C.F.J. Jackknife, bootstrap and other resampling methods in regression analysis. (with discussions) \textit{The Annals of Statistics} {\bf 14}: 1261?1350 (1986)
 \end{thebibliography}

### 2016-04-15 eLife/elife_sub.tex
Add eLife docs and ISMEJ letter to editor
On branch master
	new file:   ISMEJ/lettertoeditor_sub.docx
	deleted:    eLife/To_be_populated.txt
	new file:   eLife/elife_sub.tex
	new file:   eLife/elife_suppl_sub.tex

--- 2016-04-15 Science/science_CC.tex
@@
-% Use only LaTeX2e, calling the article.cls class and 12-point type.
+\documentclass[11pt,oneside,letterpaper]{article}
 
-\documentclass[12pt]{article}
+% graphicx package, useful for including eps and pdf graphics
+\usepackage{graphicx}
+\DeclareGraphicsExtensions{.pdf,.png,.jpg}
 
-% Users of the {thebibliography} environment or BibTeX should use the
-% scicite.sty package, downloadable from *Science* at
-% www.sciencemag.org/about/authors/prep/TeX_help/ .
-% This package should properly format in-text
-% reference calls and reference-list numbers.
+% basic packages
+\usepackage{color} 
+\usepackage{parskip}
+\usepackage{float}
 
-\usepackage{scicite}
-\usepackage{graphicx}
+% text layout
+\usepackage{geometry}
+\geometry{textwidth=15cm} % 15.25cm for single-space, 16.25cm for double-space
+\geometry{textheight=22cm} % 22cm for single-space, 22.5cm for double-space
 
-% Use times if you have the font installed; otherwise, comment out the
-% following line.
+% helps to keep figures from being orphaned on a page by themselves
+\renewcommand{\topfraction}{0.85}
+\renewcommand{\textfraction}{0.1}
 
-\usepackage{times}
+% bold the 'Figure #' in the caption and separate it with a period
+% Captions will be left justified
+\usepackage[labelfont=bf,labelsep=period,font=small]{caption}
 
-% The preamble here sets up a lot of new/revised commands and
-% environments.  It's annoying, but please do *not* try to strip these
-% out into a separate .sty file (which could lead to the loss of some
-% information when we convert the file to other formats).  Instead, keep
-% them in the preamble of your main LaTeX source file.
+% review layout with double-spacing
+\usepackage{setspace} 
+\doublespacing
+\captionsetup{labelfont=bf,labelsep=period,font=doublespacing}
 
-% The following parameters seem to provide a reasonable page setup.
+% cite package, to clean up citations in the main text. Do not remove.
+\usepackage{cite}
+\renewcommand\citeleft{(}
+\renewcommand\citeright{)}
+\renewcommand\citeform[1]{\textsl{#1}}
 
-\topmargin 0.0cm
-\oddsidemargin 0.2cm
-\textwidth 16cm 
-\textheight 21cm
-\footskip 1.0cm
+% Remove brackets from numbering in list of References
+\renewcommand\refname{\large References}
+\makeatletter
+\renewcommand{\@biblabel}[1]{\quad#1.}
+\makeatother
 
-%The next command sets up an environment for the abstract to your paper.
+\usepackage{authblk}
+\renewcommand\Authands{ \& }
+\renewcommand\Authfont{\normalsize \bf}
+\renewcommand\Affilfont{\small \normalfont}
+\makeatletter
+\renewcommand\AB@affilsepx{, \protect\Affilfont}
+\makeatother
 
-\newenvironment{sciabstract}{%
-\begin{quote} \bf}
-{\end{quote}}
+% notation
+\usepackage{amsmath}
+\usepackage{amssymb}
 
-% If your reference list includes text notes as well as references,
-% include the following line; otherwise, comment it out.
-
-\renewcommand\refname{References and Notes}
-
-% The following lines set up an environment for the last note in the
-% reference list, which commonly includes acknowledgments of funding,
-% help, etc.  It's intended for users of BibTeX or the {thebibliography}
-% environment.  Users who are hand-coding their references at the end
-% using a list environment such as {enumerate} can simply add another
-% item at the end, and it will be numbered automatically.
-
-\newcounter{lastnote}
-\newenvironment{scilastnote}{%
-\setcounter{lastnote}{\value{enumiv}}%
-\addtocounter{lastnote}{+1}%
-\begin{list}%
-{\arabic{lastnote}.}
-{\setlength{\leftmargin}{.22in}}
-{\setlength{\labelsep}{.5em}}}
-{\end{list}}
-
-% Include your paper's title here
-
-\title{Microbiota: are you sick?}
-
-% Place the author information here.  Please hand-code the contact
-% information and notecalls; do *not* use \footnote commands.  Let the
-% author contact information appear immediately below the author names
-% as shown.  We would also prefer that you don't change the type-size
-% settings shown here.
-
-\author
-{Jose Manuel Mart\'i,$^{1}$ Daniel M. Mart\'inez,$^{2}$ Manuel Pe\~na$^{1}$, C\'esar Gracia$^{1}$, \\
-Amparo Latorre$^{2,3,4}$, Andr\'es Moya$^{2,3,4}$ \& Carlos P. Garay$^{1,\ast}$\\
-\\
-\normalsize{$^{1}$Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain.}\\
-\normalsize{$^{2}$FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain.}\\
-\normalsize{$^{3}$Cavanilles Institute of Biodiversity and Evolutionary Biology, Univ. de Valencia, 46980, Spain.}\\
-\normalsize{$^{4}$CIBER en Epidemiolog\' ia y Salud P\'ublica (CIBEResp), Madrid, Spain.}\\
-\\
-\normalsize{$^\ast$To whom correspondence should be addressed; E-mail:  penya@ific.uv.es}
+%%% TITLE %%%
+\title{\vspace{1.0cm} \Large \bf 
+Health and disease imprinted in the time variability of the human microbiome
 }
 
-% Include the date command, but leave its argument blank.
+\author[1]{Jose Manuel Mart\' i}
+\author[1,2]{Daniel M. Mart\' inez}
+\author[1]{Manuel Pe\~na}
+\author[1]{C\'esar Gracia}
+\author[2,3,4]{Amparo Latorre}
+\author[2,3,4]{Andr\'es Moya}
+\author[1]{Carlos P. Garay}
+
+\affil[1]{Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain.}
+\affil[2]{FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain.}
+\affil[3]{Cavanilles Institute of Biodiversity and Evolutionary Biology, Univ. de Valencia, 46980, Spain. }
+\affil[4]{CIBER en Epidemiolog\' ia y Salud P\'ublica (CIBEResp), Madrid, Spain}
 
@@
 
-%%%%%%%%%%%%%%%%% END OF PREAMBLE %%%%%%%%%%%%%%%%
-
 \begin{document}
-
-% Double-space the manuscript.
-
-\baselineskip24pt
-
-% Make the title.
 
@@
 
-% Place your abstract within the special {sciabstract} environment.
+%%% ABSTRACT %%%
+\begin{abstract}
 
-\begin{sciabstract}
 Human microbiota plays an important role in determining changes from health to disease. Increasing research activity is dedicated to understand its diversity and variability. 
@@
 applications where microbiota plays an important role.
-\end{sciabstract}
 
-% In setting up this template for *Science* papers, we've used both
-% the \section* command and the \paragraph* command for topical
-% divisions.  Which you use will of course depend on the type of paper
-% you're writing.  Review Articles tend to have displayed headings, for
-% which \section* is more appropriate; Research Articles, when they have
-% formal topical divisions at all, tend to signal them with bold text
-% that runs into the paragraph, for which \paragraph* is the right
-% choice.  Either way, use the asterisk (*) modifier, as shown, to
-% suppress numbering.
+\end{abstract}
 
+\pagebreak
+
+%%% INTRODUCTION %%%
 \section*{Introduction}
@@
 
-\begin{figure}[!ht]
-\includegraphics[width=1.05\textwidth]{comboplot1.pdf}
-\caption{Taylor's law parameter space.
-X-weighted power-law fits of the standard deviations versus the mean values for each bacterial genus monitored in time.
-We show the fit for samples from a healthy subject (Figure A) and from a subject diagnosed with irritable bowel syndrome (Figure B) studied in our lab\cite{durban}.
-We have compiled all data studied in this work in Figure C.
-The coloured region (dashed line) corresponds to 68\% (95\%) CL region of healthy individuals in the Taylor parameter space.
-Points with errors place each individual gut microbiome in the Taylor space.
-Note that the parameters have been standardized (stdev units) to the healthy group in each study for demonstrative purposes.}
-\end{figure}
+%%% RESULTS %%%
+\section*{Results and discussion}
 
@@
 \begin{figure}[!ht]
+\includegraphics[width=1.05\textwidth]{comboplot1.pdf}
+\caption{Taylor's law parameter space.
+X-weighted power-law fits of the standard deviations versus the mean values for each bacterial genus monitored in time.
+We show the fit for samples from a healthy subject (Figure A) and from a subject diagnosed with irritable bowel syndrome (Figure B) studied in our lab\cite{durban}.
+We have compiled all data studied in this work in Figure C.
+The coloured region (dashed line) corresponds to 68\% (95\%) CL region of healthy individuals in the Taylor parameter space.
+Points with errors place each individual gut microbiome in the Taylor space.
+Note that the parameters have been standardized (stdev units) to the healthy group in each study for demonstrative purposes.}
+\end{figure}
+
+Taylor's power law has been explained in terms of various effects, all without general consensus.
+It can be shown to have its origin in a mathematical convergence similar to the central limit theorem, so virtually any statistical model designed to produce a Taylor law converge to a Tweedie distribution\cite{stat}, providing a mechanistic explanation based on the statistical theory of errors\cite{convergence1,convergence2,convergence3}.
+To unveil the generic mechanisms that drive different scenarios in the $\beta$--V space, we model the system by assuming that taxon relative abundance follows a Langevin equation with a deterministic term that captures the fitness of each taxon and a randomness term with Gaussian random noise\cite{ranking}.
+Both terms are modelled by power laws, with coefficients that can be interpreted as the taxon fitness $F_i$ and the variability $V$ (see Section 1 in supplemental material).
+When $V$ is sufficiently low, abundances are stable in time.
+Differences in variability $V$ can induce a noise-induced phase transition in relative abundances of taxa.
+The temporal evolution of the probability of a taxon having abundance $x_i$ given its fitness is governed by the Fokker--Planck equation.
+The results of solving this equation show that  the stability is best captured by fitness F and amplitude of fluctuations V phase space (see Figure 2).
+
+\begin{figure}[!ht]
 \includegraphics[width=1.05\textwidth]{comboplot2.pdf}
@@
 
-Taylor's power law has been explained in terms of various effects, all without general consensus.
-It can be shown to have its origin in a mathematical convergence similar to the central limit theorem, so virtually any statistical model designed to produce a Taylor law converge to a Tweedie distribution\cite{stat}, providing a mechanistic explanation based on the statistical theory of errors\cite{convergence1,convergence2,convergence3}.
-To unveil the generic mechanisms that drive different scenarios in the $\beta$--V space, we model the system by assuming that taxon relative abundance follows a Langevin equation with a deterministic term that captures the fitness of each taxon and a randomness term with Gaussian random noise\cite{ranking}.
-Both terms are modelled by power laws, with coefficients that can be interpreted as the taxon fitness $F_i$ and the variability $V$ (see Section 1 in supplemental material).
-When $V$ is sufficiently low, abundances are stable in time.
-Differences in variability $V$ can induce a noise-induced phase transition in relative abundances of taxa.
-The temporal evolution of the probability of a taxon having abundance $x_i$ given its fitness is governed by the Fokker--Planck equation.
-The results of solving this equation show that  the stability is best captured by fitness F and amplitude of fluctuations V phase space (see Figure 2).
 The model predicts two phases for the gut microbiome: a stable phase with large variability that permits some changes in the relative abundances of taxa and an unstable phase with larger variability, above the phase transition, where the order of abundant taxa varies significantly with time.
@@
 
-In summary, we have quantitatively characterized whether the microbiota belongs to a healthy individual or a subject corresponding to an altered or pathological state (ie, altered diet, antibiotic treatment, early gut development, diagnosed IBS).
+\section*{Conclusions}
+
+We have quantitatively characterized whether the microbiota belongs to a healthy individual or a subject corresponding to an altered or pathological state (ie, altered diet, antibiotic treatment, early gut development, diagnosed IBS).
 Deciphering the mechanisms of disease requires in depth knowledge of the underlying biological mechanisms.

### 2016-04-15 eLife/elife_suppl_sub.tex
Add eLife docs and ISMEJ letter to editor
On branch master
	new file:   ISMEJ/lettertoeditor_sub.docx
	deleted:    eLife/To_be_populated.txt
	new file:   eLife/elife_sub.tex
	new file:   eLife/elife_suppl_sub.tex

--- 2016-04-15 Science/science_CC_suppl.tex
@@
-% Use only LaTeX2e, calling the article.cls class and 12-point type.
+\documentclass[11pt,oneside,letterpaper]{article}
 
-\documentclass[12pt]{article}
+% graphicx package, useful for including eps and pdf graphics
+\usepackage{graphicx}
+\DeclareGraphicsExtensions{.pdf,.png,.jpg}
 
-% Users of the {thebibliography} environment or BibTeX should use the
-% scicite.sty package, downloadable from *Science* at
-% www.sciencemag.org/about/authors/prep/TeX_help/ .
-% This package should properly format in-text
-% reference calls and reference-list numbers.
+% basic packages
+\usepackage{color} 
+\usepackage{parskip}
+\usepackage{float}
 
-\usepackage{scicite}
-\usepackage{graphicx}
-\usepackage{amsmath}
-\usepackage[usenames,dvipsnames,table]{xcolor}
-\usepackage[nolists,tablesfirst, nomarkers]{endfloat}
+% text layout
+\usepackage{geometry}
+\geometry{textwidth=15cm} % 15.25cm for single-space, 16.25cm for double-space
+\geometry{textheight=22cm} % 22cm for single-space, 22.5cm for double-space
 
-% Use times if you have the font installed; otherwise, comment out the
-% following line.
-
-\usepackage{times}
-
-% The preamble here sets up a lot of new/revised commands and
-% environments.  It's annoying, but please do *not* try to strip these
-% out into a separate .sty file (which could lead to the loss of some
-% information when we convert the file to other formats).  Instead, keep
-% them in the preamble of your main LaTeX source file.
-
-% The following parameters seem to provide a reasonable page setup.
-
-\topmargin 0.0cm
-\oddsidemargin 0.2cm
-\textwidth 16cm 
-\textheight 21cm
-\footskip 1.0cm
-
-%The next command sets up an environment for the abstract to your paper.
-
-\newenvironment{sciabstract}{%
-\begin{quote} \bf}
-{\end{quote}}
-
-% If your reference list includes text notes as well as references,
-% include the following line; otherwise, comment it out.
+% helps to keep figures from being orphaned on a page by themselves
+\renewcommand{\topfraction}{0.85}
+\renewcommand{\textfraction}{0.1}
+%% Roman footnote numbering style to not collide with bibliographic cites: 
+\renewcommand{\thefootnote}{\Roman{footnote}}
 
@@
 \newcommand{\unclassrate}[0]{\un{kpb/s/core}}
-\renewcommand\refname{References and Notes}
-\renewcommand{\thefootnote}{\Roman{footnote}}
 
-% The following lines set up an environment for the last note in the
-% reference list, which commonly includes acknowledgments of funding,
-% help, etc.  It's intended for users of BibTeX or the {thebibliography}
-% environment.  Users who are hand-coding their references at the end
-% using a list environment such as {enumerate} can simply add another
-% item at the end, and it will be numbered automatically.
+% bold the 'Figure #' in the caption and separate it with a period
+% Captions will be left justified
+\usepackage[labelfont=bf,labelsep=period,font=small]{caption}
 
-\newcounter{lastnote}
-\newenvironment{scilastnote}{%
-\setcounter{lastnote}{\value{enumiv}}%
-\addtocounter{lastnote}{+1}%
-\begin{list}%
-{\arabic{lastnote}.}
-{\setlength{\leftmargin}{.22in}}
-{\setlength{\labelsep}{.5em}}}
-{\end{list}}
+% review layout with double-spacing
+\usepackage{setspace} 
+\doublespacing
+\captionsetup{labelfont=bf,labelsep=period,font=doublespacing}
 
-% Include your paper's title here
+% cite package, to clean up citations in the main text. Do not remove.
+\usepackage{cite}
+\renewcommand\citeleft{(}
+\renewcommand\citeright{)}
+\renewcommand\citeform[1]{\textsl{#1}}
 
-\title{Supplemental material of ``Microbiota: are you sick?''}
+% Remove brackets from numbering in list of References
+\renewcommand\refname{\large References}
+\makeatletter
+\renewcommand{\@biblabel}[1]{\quad#1.}
+\makeatother
 
-% Place the author information here.  Please hand-code the contact
-% information and notecalls; do *not* use \footnote commands.  Let the
-% author contact information appear immediately below the author names
-% as shown.  We would also prefer that you don't change the type-size
-% settings shown here.
+\usepackage{authblk}
+\renewcommand\Authands{ \& }
+\renewcommand\Authfont{\normalsize \bf}
+\renewcommand\Affilfont{\small \normalfont}
+\makeatletter
+\renewcommand\AB@affilsepx{, \protect\Affilfont}
+\makeatother
 
-\author
-{Jose Manuel Mart\'i,$^{1}$ Daniel M. Mart\'inez,$^{2}$ Manuel Pe\~na$^{1}$, C\'esar Gracia$^{1}$, \\
-Amparo Latorre$^{2,3,4}$, Andr\'es Moya$^{2,3,4}$ \& Carlos P. Garay$^{1,\ast}$\\
-\\
-\normalsize{$^{1}$Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain.}\\
-\normalsize{$^{2}$FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain.}\\
-\normalsize{$^{3}$Cavanilles Institute of Biodiversity and Evolutionary Biology, Univ. de Valencia, 46980, Spain.}\\
-\normalsize{$^{4}$CIBER en Epidemiolog\' ia y Salud P\'ublica (CIBEResp), Madrid, Spain.}\\
-\\
-\normalsize{$^\ast$To whom correspondence should be addressed; E-mail:  penya@ific.uv.es}
+% notation
+\usepackage{amsmath}
+\usepackage{amssymb}
+
+%%% TITLE %%%
+\title{\vspace{1.0cm} \Large \bf 
+Supplemental material of ``Health and disease imprinted in the time variability of the human microbiome''
 }
 
-% Include the date command, but leave its argument blank.
+\author[1]{Jose Manuel Mart\' i}
+\author[1,2]{Daniel M. Mart\' inez}
+\author[1]{Manuel Pe\~na}
+\author[1]{C\'esar Gracia}
+\author[2,3,4]{Amparo Latorre}
+\author[2,3,4]{Andr\'es Moya}
+\author[1]{Carlos P. Garay}
+
+\affil[1]{Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain.}
+\affil[2]{FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain.}
+\affil[3]{Cavanilles Institute of Biodiversity and Evolutionary Biology, Univ. de Valencia, 46980, Spain. }
+\affil[4]{CIBER en Epidemiolog\' ia y Salud P\'ublica (CIBEResp), Madrid, Spain}
 
@@
 
-% In setting up this template for *Science* papers, we've used both
-% the \section* command and the \paragraph* command for topical
-% divisions.  Which you use will of course depend on the type of paper
-% you're writing.  Review Articles tend to have displayed headings, for
-% which \section* is more appropriate; Research Articles, when they have
-% formal topical divisions at all, tend to signal them with bold text
-% that runs into the paragraph, for which \paragraph* is the right
-% choice.  Either way, use the asterisk (*) modifier, as shown, to
-% suppress numbering.
-
-\section*{Model} \label{sec:model}
+\section{Model} \label{sec:model}
 
@@
         \item If $\mathrm{AIC_{c_{NLR}}} - \mathrm{AIC_{c_{LLR}}} > 2$, the assumption of lognormal error is favoured compared to normal error, so proceed with the results obtained from the LLR fit.
-        \item If $\left|\mathrm{AIC_{c_{NLR}}} - \mathrm{AIC_{c_{LLR}}}\right| \leq 2$, no model is favoured, so proceed with averaging:
+        \item If $\left|\mathrm{AIC_{c_{NLR}}} - \mathrm{AIC_{c_{LLR}}}\right| \leq 2$, no model is favoured, so proceed with model averaging:
         \begin{eqnarray*}
@@
 \newpage
-\begin{thebibliography}{1}
+\begin{thebibliography}{99}
 	\bibitem{ranking} Blumm, N. \textit{et al.} Dynamics of ranking processes in complex systems. {\it Phys. Rev. Lett.} {\bf 109,} 128701 (2012).

### 2016-04-15 biorxiv/arxiv_sub.tex
Move docx to MSWord folder in mSystems and add biorxiv files
On branch master
	new file:   biorxiv/arxiv_sub.tex
	new file:   biorxiv/arxiv_suppl_sub.tex
	renamed:    mSystems/WHOLE_draft_0.1.docx -> mSystems/MSWord/WHOLE_draft_0.1.docx
	renamed:    mSystems/discussion_draft_0.1.docx -> mSystems/MSWord/discussion_draft_0.1.docx
	renamed:    mSystems/intro_draft_0.1.docx -> mSystems/MSWord/intro_draft_0.1.docx
	renamed:    mSystems/mat&met_draft_0.1.docx -> mSystems/MSWord/mat&met_draft_0.1.docx
	renamed:    mSystems/results_draft_0.1.docx -> mSystems/MSWord/results_draft_0.1.docx
	renamed:

--- 2016-04-15 Science/science_CC.tex
@@
-% Use only LaTeX2e, calling the article.cls class and 12-point type.
-
-\documentclass[12pt]{article}
-
-% Users of the {thebibliography} environment or BibTeX should use the
-% scicite.sty package, downloadable from *Science* at
-% www.sciencemag.org/about/authors/prep/TeX_help/ .
-% This package should properly format in-text
-% reference calls and reference-list numbers.
-
-\usepackage{scicite}
+\documentclass[aps,prl,twocolumn,superscriptaddress,showpacs]{revtex4}
 \usepackage{graphicx}
-
-% Use times if you have the font installed; otherwise, comment out the
-% following line.
-
-\usepackage{times}
-
-% The preamble here sets up a lot of new/revised commands and
-% environments.  It's annoying, but please do *not* try to strip these
-% out into a separate .sty file (which could lead to the loss of some
-% information when we convert the file to other formats).  Instead, keep
-% them in the preamble of your main LaTeX source file.
-
-% The following parameters seem to provide a reasonable page setup.
-
-\topmargin 0.0cm
-\oddsidemargin 0.2cm
-\textwidth 16cm 
-\textheight 21cm
-\footskip 1.0cm
-
-%The next command sets up an environment for the abstract to your paper.
-
-\newenvironment{sciabstract}{%
-\begin{quote} \bf}
-{\end{quote}}
-
-% If your reference list includes text notes as well as references,
-% include the following line; otherwise, comment it out.
-
-\renewcommand\refname{References and Notes}
-
-% The following lines set up an environment for the last note in the
-% reference list, which commonly includes acknowledgments of funding,
-% help, etc.  It's intended for users of BibTeX or the {thebibliography}
-% environment.  Users who are hand-coding their references at the end
-% using a list environment such as {enumerate} can simply add another
-% item at the end, and it will be numbered automatically.
-
-\newcounter{lastnote}
-\newenvironment{scilastnote}{%
-\setcounter{lastnote}{\value{enumiv}}%
-\addtocounter{lastnote}{+1}%
-\begin{list}%
-{\arabic{lastnote}.}
-{\setlength{\leftmargin}{.22in}}
-{\setlength{\labelsep}{.5em}}}
-{\end{list}}
-
-% Include your paper's title here
-
-\title{Microbiota: are you sick?}
-
-% Place the author information here.  Please hand-code the contact
-% information and notecalls; do *not* use \footnote commands.  Let the
-% author contact information appear immediately below the author names
-% as shown.  We would also prefer that you don't change the type-size
-% settings shown here.
-
-\author
-{Jose Manuel Mart\'i,$^{1}$ Daniel M. Mart\'inez,$^{2}$ Manuel Pe\~na$^{1}$, C\'esar Gracia$^{1}$, \\
-Amparo Latorre$^{2,3,4}$, Andr\'es Moya$^{2,3,4}$ \& Carlos P. Garay$^{1,\ast}$\\
-\\
-\normalsize{$^{1}$Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain.}\\
-\normalsize{$^{2}$FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain.}\\
-\normalsize{$^{3}$Cavanilles Institute of Biodiversity and Evolutionary Biology, Univ. de Valencia, 46980, Spain.}\\
-\normalsize{$^{4}$CIBER en Epidemiolog\' ia y Salud P\'ublica (CIBEResp), Madrid, Spain.}\\
-\\
-\normalsize{$^\ast$To whom correspondence should be addressed; E-mail:  penya@ific.uv.es}
-}
-
-% Include the date command, but leave its argument blank.
-
-\date{}
-
-%%%%%%%%%%%%%%%%% END OF PREAMBLE %%%%%%%%%%%%%%%%
+\usepackage{mathrsfs}
+\usepackage{bm}
 
@@
 
-% Double-space the manuscript.
+\title{Health and disease imprinted in the time variability of the human microbiome}
+\author{Jose Manuel Mart\' i}
+\affiliation{Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain}
+\author{Daniel M. Mart\' inez}
+\affiliation{Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain}
+\affiliation{FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain}
+\author{C\'esar Gracia}
+\affiliation{Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain}
+\author{Amparo Latorre}
+\affiliation{FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain}
+\affiliation{Cavanilles Institute of Biodiversity and Evolutionary Biology, Univ. de Valencia, 46980, Spain}
+\affiliation{CIBER en Epidemiolog\' ia y Salud P\'ublica (CIBEResp), Madrid, Spain}
+\author{Andr\'es Moya}
+\affiliation{FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain}
+\affiliation{Cavanilles Institute of Biodiversity and Evolutionary Biology, Univ. de Valencia, 46980, Spain}
+\affiliation{CIBER en Epidemiolog\' ia y Salud P\'ublica (CIBEResp), Madrid, Spain}
+\author{Carlos P. Garay}
+\affiliation{Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain}
 
-\baselineskip24pt
-
-% Make the title.
-
-\maketitle
-
-% Place your abstract within the special {sciabstract} environment.
-
-\begin{sciabstract}
+ \begin{abstract}
 Human microbiota plays an important role in determining changes from health to disease. Increasing research activity is dedicated to understand its diversity and variability. 
@@
 applications where microbiota plays an important role. 
-\end{sciabstract}
+ \end{abstract}
 
-% In setting up this template for *Science* papers, we've used both
-% the \section* command and the \paragraph* command for topical
-% divisions.  Which you use will of course depend on the type of paper
-% you're writing.  Review Articles tend to have displayed headings, for
-% which \section* is more appropriate; Research Articles, when they have
-% formal topical divisions at all, tend to signal them with bold text
-% that runs into the paragraph, for which \paragraph* is the right
-% choice.  Either way, use the asterisk (*) modifier, as shown, to
-% suppress numbering.
-
-\section*{Introduction}
+\maketitle
 
@@
 
-\begin{figure}[!ht]
-\includegraphics[width=1.05\textwidth]{comboplot1.pdf}
+\begin{figure*}
+\includegraphics[width=14cm,angle=0]{comboplot1.pdf}
 \caption{Taylor's law parameter space.
@@
 Note that the parameters have been standardized (stdev units) to the healthy group in each study for demonstrative purposes.}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}[!ht]
-\includegraphics[width=1.05\textwidth]{comboplot2.pdf}
+\begin{figure*}
+\includegraphics[width=14cm,angle=0]{comboplot2.pdf}
 \caption{Phase space of microbiota stability.
@@
 We also show an IBS--diagnosed patient transiting from the unstable to the stable phase.}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{thebibliography}{1}
+\textit{Acknowledgements.---} This work is supported by Generalitat Valencia Prometeo Grants II/2014/050, II/2014/065, by the Spanish Grants FPA2011-29678 , BFU2012-39816-C02-01 of MINECO and by PITN- GA-2011-289442-INVISIBLES.
+JMM and DMM acknowledge FPI and FISABIO fellowships.
+
+\begin{thebibliography}{99}
+
 \bibitem{moving} Caporaso, J.G. et al. Moving pictures of the human microbiome. {\it Genome Biol.} {\bf 12,} R50 (2011).

### 2016-04-17 mSystems/mySys_MAIN.tex
Create 1st mySystems LaTeX release, which makes MSWord files obsolete!
On branch master
	renamed:    MSWord/WHOLE_draft_0.1.docx -> MSWord_OBSOLETE/WHOLE_draft_0.1.docx
	renamed:    MSWord/discussion_draft_0.1.docx -> MSWord_OBSOLETE/discussion_draft_0.1.docx
	renamed:    MSWord/intro_draft_0.1.docx -> MSWord_OBSOLETE/intro_draft_0.1.docx
	renamed:    MSWord/mat_met_draft_0.3.docx -> MSWord_OBSOLETE/mat_met_draft_0.3.docx
	renamed:    MSWord/results_draft_0.1.docx -> MSWord_OBSOLETE/results_draft_0.1.docx
	renamed:    MSWord/skeleton_draft_0.3.docx -> MSWord_OBSOLETE/skeleton_draft_0.3.docx
	n

% !TEX TS-program = pdflatex
% !TEX encoding = UTF-8 Unicode

\documentclass[12pt,oneside,letterpaper]{article}

%%% PACKAGES
\usepackage{lmodern} % Activa fuentes y macros de Latin Modern
\usepackage[T1]{fontenc} % set font encoding
\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)
\usepackage[english]{babel} 
\usepackage[expert, charter]{mathdesign} % Fuentes... disponibles conjuntos {charter|utopia}
%%\usepackage{calligra}
\usepackage[printwatermark]{xwatermark}
\usepackage{graphicx}
\usepackage[usenames,dvipsnames,table]{xcolor}
\usepackage{parskip}
%%\usepackage{float}
%\usepackage[nolists,tablesfirst, nomarkers]{endfloat}
\usepackage{moreverb} % for verbatim output

% text layout
\usepackage[letterpaper]{geometry}
\geometry{textwidth=16.25cm} % 15.25cm for single-space, 16.25cm for double-space
\geometry{textheight=22.5cm} % 22cm for single-space, 22.5cm for double-space

% helps to keep figures from being orphaned on a page by themselves
\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
%% Roman footnote numbering style to not collide with bibliographic cites: 
\renewcommand{\thefootnote}{\Roman{footnote}}

% line numbering
\usepackage[running,mathlines]{lineno}
\renewcommand\thelinenumber{\color{red}\arabic{linenumber}}
\linenumbers

% bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,font=small]{caption}

% review layout with double-spacing
\usepackage{setspace} 
\captionsetup{labelfont=bf,labelsep=period,font=doublespacing}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}
\renewcommand\citeleft{(}
\renewcommand\citeright{)}
\renewcommand\citeform[1]{\textsl{#1}}

% Remove brackets from numbering in list of References
\renewcommand\refname{\large References}
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Importance environment and keywords command
\newcommand\importancename{Importance}
\newcommand{\importance}[1]{
    \small
    \begin{center}
        {\bfseries \importancename\vspace{-.5em}\vspace{0pt}}
    \end{center}
    \begin{quote}
    #1
    \end{quote}
}
\providecommand{\keywords}[1]{\textbf{\textit{Keywords---}} #1}

% Package authblk
\usepackage{authblk}
\renewcommand\Authands{ \& }
\renewcommand\Authfont{\normalsize \bf}
\renewcommand\Affilfont{\small \normalfont}

% notation
\usepackage{amsmath}

%% Roman footnote numbering style to not collide with bibliographic cites: 
\renewcommand{\thefootnote}{\Roman{footnote}}

% aux commands
\newcommand{\CC}[0]{\emph{cmplxcruncher}}
\newcommand{\task}[1]{\texttt{\bfseries\scshape\textcolor{MidnightBlue}{#1}}}
\newcommand{\un}[1]{\operatorname{#1}}
\newcommand{\unclassrate}[0]{\un{kpb/s/core}}
\newcommand{\todo}[1]{\texttt{\bfseries\textcolor{Orange}{#1}}}

% watermark (time consuming!)
\newwatermark[allpages,color=red!4,angle=45,scale=9,xpos=-40,ypos=0]{DRAFT}

% word count (Need --enable-write18 or --shell-escape)
\immediate\write18{texcount -opt=option.tc \jobname.tex > wordcount.aux} 
\newcommand\wordcount{\verbatiminput{wordcount.aux}}

%%% DOCUMENT %%%

\begin{document}

\title{
	\vspace*{0mm} % Helps to include the corresponding author footnote in the title page
	\singlespacing
	\begin{flushleft}
		\texttt{\large Title:} \\
	\end{flushleft}
	\vspace*{2mm}
	Health and disease imprinted in the time variability of the human microbiome\\
	\vspace*{6mm}
	\begin{flushleft}
		\texttt{\large Running title:} \\
	\end{flushleft}
	\vspace*{0mm}
	Microbiota, are you sick?
	\vspace*{4mm}
	}

\doublespacing

\author[1,2]{Jose Manuel Martí}
\author[1,2,3]{Daniel M. Martínez}
\author[2]{Manuel Peña}
\author[1,2]{César Gracia}
\author[1,3,4,5]{Amparo Latorre}
\author[1,3,4,5]{Andrés Moya}
\author[1,2,\#]{Carlos P. Garay}

\affil[1]{Institute for Integrative Systems Biology (I2SysBio), 46980, Spain.}
\affil[2]{Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain.}
\affil[3]{FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain.}
\affil[4]{Cavanilles Institute of Biodiversity and Evolutionary Biology, Univ. de Valencia, 46980, Spain.}
\affil[5]{CIBER en Epidemiología y Salud Pública (CIBEResp), Madrid, Spain}

\date{}

\maketitle
\wordcount
\footnote[0]{$^\#$ Corresponding author: penagaray@gmail.com}

\clearpage

%TC:newcounter abswords Words in abstract
%TC:macro \abstract [abswords]
\abstract{Human microbiota plays an important role in determining changes from health to disease. Increasing research activity is dedicated to understand its diversity and variability. 
We analyse 16S rRNA and whole genome sequencing (WGS) data from the gut microbiota of 97 individuals monitored in time. Temporal fluctuations in the microbiome reveal 
significant differences due to factors that affect the microbiota such as dietary changes, antibiotic intake, early gut development or disease. Here we show that a fluctuation 
scaling law describes the temporal variability of the system and that a noise-induced phase transition is central in the route to disease. The universal law distinguishes healthy 
from sick microbiota and quantitatively characterizes the path in the phase space, which opens up its potential clinical use and, more generally, other technological 
applications where microbiota plays an important role.}

%TC:newcounter impwords Words in importance section
%TC:macro \importance [impwords]
\importance{Human microbiota is tightly associated to the health status of a person. Here we analyse the microbial composition of several subjects under different conditions, over a time span that ranges from days to months. Using the Langevin equation as the basis of our mathematical framework in order to evaluate microbial temporal stability, we prove that we are capable to distinguish stable from unstable microbiotas. This first step will help us to determine how microbiota temporal stability is related to the healthiness of the people, and it will allow the development of a more complete framework in order to deepen the knowledge of this complex system. }

\vspace{4mm}
\begin{keywords}
microbiome, systems biology, ecological modelling, community composition, stability	
\end{keywords}

%===================================== SECTIONS =======================================
%%% Introduction
\section*{Introduction}
\input{mySys_intro.tex}   
%%% Result
\section*{Results}
\input{mySys_results.tex}
%%% Discussion
\section*{Discussion}
\input{mySys_discussion.tex}    
%%% Material & Methods
\section*{Materials and Methods}
\input{mySys_MandM.tex}

\section*{Acknowledgments}
Authors declare that there are no competing financial interests in relation to the work described here.

\section*{Funding Information}
This work is supported by Generalitat Valenciana Prometeo Grants II/2014/050, II/2014/065, by the Spanish Grants FPA2011-29678, BFU2012-39816-C02-01 of MINECO and by PITN-GA-2011-289442-INVISIBLES. JMM \& DMM acknowledge FPI and FISABIO fellowships. \todo{Modificar becas de JMM y DMM ¿poner algún grant más?}

\newpage
\begin{thebibliography}{99}
\bibitem{moving} Caporaso, J.G. et al. Moving pictures of the human microbiome. {\it Genome Biol.} {\bf 12,} R50 (2011).
\bibitem{LEA} Faith, J.J. et al. The long-term stability of the human gut microbiota. {\it Science} {\bf 341,} 1237439 (2013).
\bibitem{kwashiorkor} Smith M.I. et al. Gut microbiomes of Malawian twin pairs discordant for kwashiorkor. {\it Science} {\bf 339,} 548-54 (2013).
\bibitem{diet} David, L.A. et al. Diet rapidly and reproducibly alters the human gut microbiome. {\it Nature} {\bf 505,} 559-63 (2014).
\bibitem{antibiotic} Dethlefsen L., Relman D. A. Incomplete recovery and individualized responses of the human distal gut microbiota to repeated antibiotic perturbation. {\it Proc. Nat. Acad. Sci. USA} {\bf 108,} 4554-61 (2011).
\bibitem{IBS} Durban, A. \textit{et al.} Structural alterations of faecal and mucosa-associated bacterial communities in irritable bowel syndrome. {\it FEMS Microbiol. Ecol.} {\bf 86}, 581-9 (2013).
\bibitem{QIIME} Caporaso, J.G. et al. QIIME allows analysis of high-throughput community sequencing data. {\it Nature Methods} {\bf 7,} 335-6 (2010).
\bibitem{LMAT} Ames, S.K. et al. Scalable metagenomic taxonomy classification usng a reference genome database. {\it Bioinformatics} {\bf 29,} 2253-2260 (2013).
\bibitem{fs} Eisler,Z., Bartos,I., Kertesz,J. Fluctuation scaling in complex systems: Taylor's law and beyond. {\it Adv. Phys.} {\bf 57,} 85 (2008).
\bibitem{taylor} Taylor, L.R. Aggregation, Variance and the mean. {\it Nature} {\bf 189,} 732-35 (1961).
\bibitem{stat} Jorgensen,B., Martinez,J.R., Tsao,M. Asymptotic behaviour of the variance function. {\it Scand. J. Statist..} {\bf 21,} 223-243 (1994).
\bibitem{convergence1} Fronczak,A., Fronczak,P. Origins of Taylor's power law for fluctuation scaling in complex systems. {\it Phys. Rev. E} {\bf 81,} 066112 (2010).
\bibitem{convergence2} Kendal, W.S., Jorgensen,B. Taylor's power law and fluctuation scaling explained by a central-limit-like convergence. {\it Phys. Rev. E} {\bf 83,} 066115 (2011).
\bibitem{convergence3} Kendal, W.S., Jorgensen,B. Tweedie convergence: A mathematical basis for Taylor's power law. {\it Phys. Rev. E} {\bf 84,} 066120 (2011).
\bibitem{ranking} Blumm, N. et al. Dynamics of ranking processes in complex systems. {\it Phys. Rev. Lett.} {\bf 109,} 128701 (2012).
	\bibitem{FD} Weber, J. \textit{et al.} Fluctuation dissipation theorem. {\it Phys. Rev.} {\bf 101}, 1620-6 (1956).
	\bibitem{FASTX} Gordon, A., Hannon, G.J. FASTX-Toolkit. FASTQ/A shortreads pre-processing tools (2010). http://hannonlab.cshl.edu/fastx\_toolkit/ (accessed 23 Feb 2015).
	\bibitem{SILVA} Quast C.  \textit{et al.} The SILVA ribosomal RNA gene database project: improved data processing and web-based tools (2013)
 	\bibitem{ecology} Xiao Xiao, Ethan P. White, Mevin B. Hooten, and Susan L. Durham. On the use of log-transformation vs. nonlinear regression for analyzing biological power laws. {\it Ecology} {\bf 92}, 10, 1887-1894 (2011).
 	\bibitem{genR2} Magee L., $R^2$ measures based on wald and likelihood ratio joint significance tests. {\it The American Statistician} {\bf 44}, 3, 250-253 (1990).
	\bibitem{disR2} Nagelkerke N.J.D., A note on a general definition of the coefficient of determination. {\it Biometrika} {\bf 78}, 3, 691-692 (1991).
	\bibitem{boot} Wu, C.F.J. Jackknife, bootstrap and other resampling methods in regression analysis. (with discussions) \textit{The Annals of Statistics} {\bf 14}: 1261?1350 (1986)
\end{thebibliography}
\todo{Eliminar \textit{et al.} y poner la referencia completa como exige la guía de estilo de la revista...}

\end{document}

### 2016-04-17 mSystems/mSys_MandM.tex
MINOR: rename mySys_ file prefix to mSys_ prefix
On branch master
	renamed:    mySys_MAIN.pdf -> mSys_MAIN.pdf
	renamed:    mySys_MAIN.tex -> mSys_MAIN.tex
	renamed:    mySys_MandM.tex -> mSys_MandM.tex
	renamed:    mySys_discussion.tex -> mSys_discussion.tex
	renamed:    mySys_intro.tex -> mSys_intro.tex
	renamed:    mySys_results.tex -> mSys_results.tex
	renamed:    mySys_suppl_0.1.tex -> mSys_suppl_0.1.tex

--- 2016-04-15 eLife/elife_suppl_sub.tex
@@
-\documentclass[11pt,oneside,letterpaper]{article}
+We have analysed more than 35000 time series of taxa from the gut microbiomes of 97 individuals (sampling from three up to 332 time points), obtained from publicly available high throughput sequencing data on: healthy individuals over a long term span\cite{moving}, people with various degrees of obesity\cite{LEA}, twin pairs discordant for kwashiorkor\cite{kwashiorkor}, response to diet changes\cite{diet} or to antibiotic perturbation\cite{antibiotic}, and subjects diagnosed with irritable bowel syndrome (IBS)\cite{IBS}.
+We engineered a complete software framework and a web platform, ComplexCruncher, ready to be implemented by other users.
 
-% graphicx package, useful for including eps and pdf graphics
-\usepackage{graphicx}
-\DeclareGraphicsExtensions{.pdf,.png,.jpg}
+We summarize our dataset in ST1-6 (Tables 1-6 in suplemental material).
+The bacteria and archaea taxonomic assignations were obtained by analysing 16S rRNA sequences, which were clustered into operational taxonomic units (OTUs) sharing 97 \% sequence identity using QIIME\cite{QIIME}.
+WGS data\cite{kwashiorkor} were analysed and assigned at strain level by the Livermore Metagenomic Analysis Toolkit (LMAT)\cite{LMAT}, according to their default quality threshold.
+Genus, with best balance between error assignment and number of taxa, was chosen as our reference taxonomic level.
+We have verified that our conclusions are not significantly affected by selecting family or species as the reference taxonomic level (see Figure 1 in supplemental material).
 
-% basic packages
-\usepackage{color} 
-\usepackage{parskip}
-\usepackage{float}
+\todo{Specify, in each study treated, the nature of the samples (conditions, timespan between timepoints, subjects).
+Specify, and it is very important, what we consider ?healthy?
+in each study (for example: pre-antibiotics is healthy)}
 
-% text layout
-\usepackage{geometry}
-\geometry{textwidth=15cm} % 15.25cm for single-space, 16.25cm for double-space
-\geometry{textheight=22cm} % 22cm for single-space, 22.5cm for double-space
-
-% helps to keep figures from being orphaned on a page by themselves
-\renewcommand{\topfraction}{0.85}
-\renewcommand{\textfraction}{0.1}
-%% Roman footnote numbering style to not collide with bibliographic cites: 
-\renewcommand{\thefootnote}{\Roman{footnote}}
-
-\newcommand{\CC}[0]{\emph{cmplxcruncher}}
-\newcommand{\task}[1]{\texttt{\bfseries\scshape\textcolor{MidnightBlue}{#1}}}
-\newcommand{\un}[1]{\operatorname{#1}}
-\newcommand{\unclassrate}[0]{\un{kpb/s/core}}
-
-% bold the 'Figure #' in the caption and separate it with a period
-% Captions will be left justified
-\usepackage[labelfont=bf,labelsep=period,font=small]{caption}
-
-% review layout with double-spacing
-\usepackage{setspace} 
-\doublespacing
-\captionsetup{labelfont=bf,labelsep=period,font=doublespacing}
-
-% cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
-\renewcommand\citeleft{(}
-\renewcommand\citeright{)}
-\renewcommand\citeform[1]{\textsl{#1}}
-
-% Remove brackets from numbering in list of References
-\renewcommand\refname{\large References}
-\makeatletter
-\renewcommand{\@biblabel}[1]{\quad#1.}
-\makeatother
-
-\usepackage{authblk}
-\renewcommand\Authands{ \& }
-\renewcommand\Authfont{\normalsize \bf}
-\renewcommand\Affilfont{\small \normalfont}
-\makeatletter
-\renewcommand\AB@affilsepx{, \protect\Affilfont}
-\makeatother
-
-% notation
-\usepackage{amsmath}
-\usepackage{amssymb}
-
-%%% TITLE %%%
-\title{\vspace{1.0cm} \Large \bf 
-Supplemental material of ``Health and disease imprinted in the time variability of the human microbiome''
-}
-
-\author[1]{Jose Manuel Mart\' i}
-\author[1,2]{Daniel M. Mart\' inez}
-\author[1]{Manuel Pe\~na}
-\author[1]{C\'esar Gracia}
-\author[2,3,4]{Amparo Latorre}
-\author[2,3,4]{Andr\'es Moya}
-\author[1]{Carlos P. Garay}
-
-\affil[1]{Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain.}
-\affil[2]{FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain.}
-\affil[3]{Cavanilles Institute of Biodiversity and Evolutionary Biology, Univ. de Valencia, 46980, Spain. }
-\affil[4]{CIBER en Epidemiolog\' ia y Salud P\'ublica (CIBEResp), Madrid, Spain}
-
-\date{}
-
-\begin{document}
-
-\maketitle
-
-\section{Model} \label{sec:model}
+%%%
+\subsection*{Model} \label{sec:model}
 
 We model the microbial abundances across time along the lines of Blumm \textit{et al.}\cite{ranking}. The dynamics of taxon relative abundances is described by the Langevin equation:
+\begin{linenomath}
 \begin{equation}
@@
 \end{equation}
+\end{linenomath}
 where F$_i$ captures the fitness of the taxon i, V corresponds to the noise amplitude and $\xi_i$(t) is a Gaussian random noise with zero mean  $<\xi_i(t)>$ = 0 and variance uncorrelated in time, $<\xi_i(t) \xi_i(t')>$ =  $\delta(t'-t)$, .
-The function $\phi(t)$ ensures the normalization at all times, $\sum x_i(t) = 1$, and corresponds to $\phi(t) = \sum F_i x_i^\alpha + \sum V x_i^\beta \xi_i(t)$ .
+The function $\phi(t)$ ensures the normalization at all times, $\sum x_i(t) = 1$, and corresponds to $\phi(t) = \sum F_i x_i^\alpha + \sum V x_i^\beta \xi_i(t)$.
 The temporal evolution of the probability that a taxon i has a relative abundance $x_i(t)$, P(x$_i$,t), is determined by the Fokker-Planck equation:
+\begin{linenomath}
 \begin{equation}
@@
 \end{equation}
+\end{linenomath}
 The microbiota evolves towards a steady-state with a time-independent probability depending on the values of $\alpha$, $\beta$, F$_i$ and V.
@@
 
-\section{Selection and Methods}
+\todo{Explain better the fluctiation-dissipation theorem}
 
-\subsection{Sample selection}
+%%%
+\subsection*{Selection and Methods}
+
+\subsubsection*{Sample selection}
 We have chosen studies about relevant pathologies containing metagenomic sequencing time data series of bacterial populations from humans in different healthy and non-healthy states.
@@
 
-\subsection{16rRNA sequences processing}
+\subsubsection*{16rRNA sequences processing}
 Reads from the selected studies were first quality filtered using the FastX toolkit\cite{FASTX}, allowing only those reads which had more than 25 of quality along the 75\% of the complete sequence.
@@
 
-\subsection{Metagenomic sequences processing}
+\subsubsection*{Metagenomic sequences processing}
 Metagenomic shotgun (and 16S too) sequences were analyzed with LMAT (Livermore Metagenomics Analysis Toolkit) software package\cite{LMAT} (version 1.2.4, with Feb'15 release of data base \emph{LMAT-Grand}).
@@
 
-\subsection{Taxa level selection}
+\subsubsection*{Taxa level selection}
 We selected genus as taxonomic level for the subsequent steps of our work.
@@
 
-\section{Temporal evolution of model parameters}
-
-We have studied the time dependence of the variability $V$ and power law index $\beta$ (see Section \ref{sec:model}) by using a sliding window approach.
-The total number of time points are divided in subsets of five points, where next subset is defined by adding next time sampling and by eliminating the earliest one.
-Both parameters were calculated for each subset against the average time lapse.
-Figure \ref{fig:tempevo1} shows the variability  $V$ as a function of time for the largest sampling: two individuals in the Caporaso's study\cite{moving} corresponding to the gut microbiota of a male (upper plot) and a female (lower plot).
-Figure \ref{fig:tempevo2} shows the time evolution of $V$ for patient P2 of the IBS study\cite{IBS} (upper plot) and patient D in the antibiotics study\cite{antibiotic} (lower plot).
-
-\begin{figure}
-	\includegraphics[width=1.0\textwidth]{results/sliwin/male_mov.pdf}
-	\hspace*{3mm}\includegraphics[width=0.448\textwidth]{results/sliwin/female_mov.pdf}
-\caption{$V$ as a function of time for the two individuals in the Caporaso's study\cite{moving}: samples of gut microbiome of a male (upper plot) and a female (lower plot).
-Both samples show changes in the variability V with quasi--periodic behavior peaked at about 10 days.
-Variability grows more for the gut microbiota of the male and share a minimal value around 0.1 with the gut microbiota of the female.}
-\label{fig:tempevo1}
-\end{figure}
-
-\begin{figure}
-	\centering 
- 	\includegraphics[width=0.8\textwidth]{results/sliwin/patP2_IBS.pdf}
-  	\includegraphics[width=1.0\textwidth]{results/sliwin/patD_antibio.pdf} 
-\caption{$V$ as a function of time for patient P2 of the IBS study\cite{IBS} (upper plot) and patient D in the antibiotics study\cite{antibiotic} (lower plot). The variability 
-of the gut microbiota of P2 decreases from above 0.3 to below 0.2, showing a slow tendency to increase the order of the system.
-Antibiotic intake leaks to a quick increase of variability which lasts for a few days to recover ordering.
-The second antibiotic treatment shows some memory (lower increase of variability) with a slower recovery.
-NOTE: The blue vertical lines in the lower plot are showing the periods of antibiotic treatment.}
-\label{fig:tempevo2}
-\end{figure}
-
-\section{Engineered Software}
+\subsection*{ComplexCruncher}
 
@@
 Used together, this software enables other users to reproduce our results easily and, furthermore, upload and analyse their own data or experiment with the preloaded metagenomics data sets.
-The next sections of this supplementary material deal with these two software pieces in detail.
-
-\section{Web-based front-end}
 
@@
 
-CCWebPortal is a web application that runs on current versions of many browsers.
-Additional software is not needed and only requires javaScript to be enabled on the browser to run applications.
-CCWebPortal is, at least, compatible with the following browsers: Internet Explorer 6+, Firefox 3.6+ (PC, Mac), Safari 4+, Chrome 10+ and Opera 11+ (PC, Mac).
-
-CCWebPortal is implemented following the client$-$server distributed programming model.
-The user connects to the website and a set of javaScript files are downloaded and executed on the local machine.
-Subsequently, the client application connects to a remote server that enables the execution of calculations and transactions through a centralized database management system.
-The web application is deployed on the client side using ExtJs libraries.
-ExtJs is a javaScript development framework to implement interactive Web applications using AJAX, DHTML and DOM technologies.
-ExtJS has been widely adopted by large technology projects worldwide.
-Its success lies in its ability to design and implement web interfaces with a clear and flexible functionality for the end user.
-ExtJs is used under the GNU license GPL v3.
-
-The CCWebPortal website communicates with a remote server that hosts a MySQL database.
-The database runs on a MyISAM engine that optimizes the speed of access to records in a transactional web-based environment.
+CCWebPortal is a web application that runs on current versions of many browsers. Additional software is not needed and only requires javaScript to be enabled on the browser to run applications. 
+CCWebPortal is implemented following the client$-$server distributed programming model, where the javaScript client application connects to a remote server that enables the execution of calculations and transactions through a centralized database management system.
 A set of relational tables allows the structuring of the metagenomics repository to establish relationships between records.
 Thus the search and information threshing is optimized for queries launched from the client interface.
-Access to the database on the server is implemented through Django framework.
-Django is an open-source framework written in Python using the model-view-controller (MVC) architectural pattern for implementing user interfaces, which eases the development and updating of code.
-This framework is used to implement web applications on the server side and allows flexible and modular developments and access to various types of database managers.
-Django runs as a web server, awaiting requests from the client CCWebPortal application.
-Once a user is authorized to perform a given operation, Django interacts with the database serving queries to the client application.
+Access to the database on the server is implemented through Django framework, an open-source framework written in Python using the model-view-controller (MVC) architectural pattern for implementing user interfaces.
 
-\section{Computing kernel}
-
-The effective data analysis has been performed with a Python tool developed from scratch following the OOP\footnote{Object Oriented Programming, but OOP also stands for Object Oriented Python, which is applicable here too.} paradigm, called \CC.
-This software is the back-end of the website described in the previous subsection.
+The effective data analysis has been performed with a Python tool developed from scratch to more than 4200 lines of code.
+Implemented following the Object Oriented Programming, paradigm, this software is the back-end of the website described above.
 However, it could be run as an independent piece of software since it is built as a Python package provided with a command-line front-end (\CC.py).
-Once installed, the tool can be run interactively (using the flag \texttt{-i}/\texttt{--interactive}) but also in automatic mode (flag \texttt{-a}/\texttt{--automatic} followed by the desired graphic format, e.g., \texttt{pdf}, \texttt{png}, \texttt{svg}, \texttt{ps}), which uses parallel computation to speed up the analysis of several data sources.
-
-The external mandatory dependencies of \CC\ are with Python packages \emph{pandas}, \emph{matplotlib}, \emph{numpy}, \emph{scipy} and \emph{xlrd}, while the recommended ones are \emph{pylatex} and \emph{uncertainties}.
-The PEP8 compliant code of \CC\ could be executed under a Python 2.7 interpret, but a 3.X interpret is strongly advised.
-The complete distribution of \CC\ has the code signature shown in the Table \ref{tab:cloc}, obtained through the \emph{CLOC}\footnote{\emph{Count Lines of Code} is a perl script that counts blank lines, comment lines, and physical lines of source code.} tool release 1.62.
-Figure \ref{fig:CC} shows the code layout of the complete distribution of \CC.
-
-\begin{table}
-  \begin{center}
-    \begin{tabular}{lrrrrcrcr}
-    \hline
-	Language & files & blank & comment & code & x & scale & = & 3rd gen. equiv \\
-	\hline
-	Python & 13 & 384 & 666 & 4222 &x& 4.20 &=& 17732.40 \\
-	R & 1 & 11 & 38 & 192 &x& 3.00 &=& 576.00 \\	
-	\hline
-	SUM: & 14 & 395 & 704 & 4414 &x& 4.15 &=& 18308.40 \\
-	\hline
-    \end{tabular}
-  \end{center}
-  \caption{Line counting for the complete distribution of \CC\ by \emph{CLOC}}
-  \label{tab:cloc}
-\end{table}
-
-\begin{figure}
-  \centering
-	\[ \text{\task{ComplexCruncher}}
-	\begin{cases}
-		\text{\task{cmplxcruncher.py}} \\
-		\\ 
-		\text{\task{cmplxcruncher}}
-		\begin{cases}
-			\text{\task{\_\_init\_\_.py}} \\ 
-			\\ 
-			\text{\task{models.py}} \\
-			\\ 
-			\text{\task{plots.py}} \\
-			\\ 
-			\text{\task{sessions.py}} \\ 
-		\end{cases}
-		\\ \\
-		\text{\task{tools}}
-		\begin{cases}
-			\text{\task{power\_analysis.R}} \\
-			\\ 
-			\text{\task{rawlmat2lmat.py}} \\
-			\\ 
-			\text{\task{lmat2cmplx.py}} \\
-			\\ 
-			\text{\task{pyLMAT\_rl.py}} \\
-			\\ 
-			\text{\task{pyLMAT\_cs.py}} \\
-			\\ 
-			\text{\task{pyLMAT\_gl.py}} \\
-			\\
-			\text{\task{LMAT}}
-			\begin{cases}
-	    		\text{\task{pyLMAT\_rescore.py}} \\ 
-	    		\\ 
-	    		\text{\task{HMPspiker.py}} 
-			\end{cases}
-		\end{cases}
-	\end{cases} \]
-	\caption{\CC\ layout}
-	\label{fig:CC}
-\end{figure}
+Once installed, the tool can be run interactively but also in automatic mode, which uses parallel computation to speed up the analysis of several data sources.
 
@@
 
-\section{Un-weighted power-law fit}\label{sec:unw}
+\subsection*{Un-weighted power-law fit}\label{sec:unw}
 
-\subsection{Fitting the best model}
+\subsubsection*{Fitting the best model}
 
-As mentioned above, to choose between fitting power laws ($y=Vx^\beta$) using linear regression on log-transformed (LLR) data versus non-linear regression (NLR), we mainly follow \emph{General Guidelines for the Analysis of Biological Power Laws}\cite{ecology}.
+As already mentioned, to choose between fitting power laws ($y=Vx^\beta$) using linear regression on log-transformed (LLR) data versus non-linear regression (NLR), we mainly follow \emph{General Guidelines for the Analysis of Biological Power Laws}\cite{ecology}.
 It consists of the following three steps:
@@
             \item The likelihood of a normal distribution is: 
+            \begin{linenomath}
             $$\mathcal{L}_\text{norm} = \prod_{i=1}^n\left[\frac{1}{\sqrt{2\pi\sigma^2_\text{NLR}}}\;\exp{\left(-\frac{\left(y_i-V_\text{NLR}x_i^{\beta_\text{NLR}}\right)^2}{2\sigma^2_\text{NLR}}\right)}\right]$$
+            \end{linenomath}
             \item So, the loglikelihood of a normal distribution is:
@@
         \item Calculate the \emph{corrected Akaike's Information Criterion} for the NLR model:
+		\begin{linenomath}
         $$\mathrm{AIC_{c_{NLR}}} = 2k - 2\log\mathcal{L}_\text{norm} + \frac{2k(k+1)}{n-k-1}$$
+		\end{linenomath}
         \item Fit the Log-transformed Linear Regression (LLR) model and obtain $V_\text{LLR}$, $\beta_\text{LLR}$ and $\sigma_\text{LLR}^2$.
@@
             \item The likelihood of a lognormal distribution is: 
+            \begin{linenomath}
             $$\mathcal{L}_\text{logn} = \prod_{i=1}^n\left[\frac{1}{y_i\sqrt{2\pi\sigma^2_\text{LLR}}}\;\exp{\left(-\frac{\left(\log|y_i|-\log|V_\text{LLR}|-\beta_\text{LLR}\log|x_i|\right)^2}{2\sigma^2_\text{LLR}}\right)}\right]$$
+            \end{linenomath}
             \item So, the loglikelihood of a lognormal distribution is: 
@@
         \item Calculate the \emph{corrected Akaike's Information Criterion} for the LR model: 
+        	\begin{linenomath}
         	$$\mathrm{AIC_{c_{LLR}}} = 2k - 2\log\mathcal{L}_\text{logn} + \frac{2k(k+1)}{n-k-1}$$
+			\end{linenomath}
     \end{enumerate}
@@
 
-\subsection{Calculating the coefficient of determination}
+\subsubsection*{Calculating the coefficient of determination}
 
 We think the best approach in this situation is to apply the generalized $R^2$ that, for continuous models, was defined as\cite{genR2}:
+\begin{linenomath}
 $$ R^2 = 1 - \left(\frac{\mathcal{L}(0)}{\mathcal{L}(\hat\theta)}\right)^{\!\!\frac{2}{n}} $$
+\end{linenomath}
 where $\mathcal{L}(\hat\theta)$ and $\mathcal{L}(0)$  denote the likelihoods of the fitted and the ``null'' model, respectively, and $n$ is the sample size.
 In terms of the loglikelihoods, the generalized coefficient of determination would be:
+\begin{linenomath}
 $$ R^2 = 1 - \mathrm{e}^{-\frac{2}{n}\left(\log\mathcal{L}(\hat\theta)-\log\mathcal{L}(0)\right)} $$
+\end{linenomath}
 We have the likelihoods calculated from the previous section, but what about the ``null'' models?
@@
 So for the Gaussian additive error model:
+\begin{linenomath}
 $$\mathcal{L}_\text{norm}(0) = \prod_{i=1}^n\left[\frac{1}{\sqrt{2\pi\sigma^2_\text{NLR0}}}\;\exp{\left(-\frac{\left(y_i-\bar y\right)^2}{2\sigma^2_\text{NLR0}}\right)}\right]$$
+\end{linenomath}
 So:
@@
 recovering the traditional expression for $R^2$. Using the same approach for calculating $R^2_\text{LLR}$, then:
+\begin{linenomath}
 $$\mathcal{L}_\text{logn}(0) = \prod_{i=1}^n\left[\frac{1}{y_i\sqrt{2\pi\sigma^2_\text{LLR0}}}\;\exp{\left(-\frac{\left(\log|y_i|-\log|B_\text{LLR0}|\right)^2}{2\sigma^2_\text{LLR0}}\right)}\right]$$
+\end{linenomath}
 So:
@@
 
-\section{X-weighted power-law fit}\label{sec:X-w}
+\subsection*{X-weighted power-law fit}\label{sec:X-w}
 
 When fitting the power-law of std vs. mean, we can take into account that every mean has uncertainty and estimate it for a sample size $n$ by the SEM (\emph{Standard Error of the Mean}):
+\begin{linenomath}
 $$ \mathrm{SEM} = \frac{s}{\sqrt{n}}$$
+\end{linenomath}
 where $s$ is the sample standard deviation. So, the vector of weights is computed with:
+\begin{linenomath}
 $$ \mathbf{w} = \frac{1}{\overrightarrow{\mathrm{SEM}}} = \frac{\sqrt{\mathbf{n}}}{\mathbf{s}}$$
+\end{linenomath}
 
@@
 
-\subsection{Method 1: By inverting the data}
+\subsubsection*{Method 1: By inverting the data}
 
 In the case of the log-LR model, we have:
+\begin{linenomath}
 $$\log y = \log V + \beta\log x \quad\rightarrow\quad \underbrace{\log x}_{\tilde y} = \overbrace{-\frac{1}{\beta}\log V}^{b} + \overbrace{\frac{1}{\beta}}^{m}\underbrace{\log y}_{\tilde x}$$ 
+\end{linenomath}
 where $m$ determines the slope or gradient of the fitted line, and $b$ determines the point at which the line crosses the y-axis, otherwise known as the y-intercept.
@@
 
-\subsection{Method 2: Bootstrapping-like strategy}
+\subsubsection*{Method 2: Bootstrapping-like strategy}
 
@@
 That is, for each replicate, a new x-data array is computed based on:
+\begin{linenomath}
 $$x^*_i = x_i + v_i$$
+\end{linenomath}
 where $v_i$ is a Gaussian random variable with mean $\mu_i=0$ and standard deviation $\sigma_i=\mathrm{SEM}_i$, as defined previously in this supplementary material.
@@
 
-\section{Rank Stability Index (RSI)}\label{sec:RSI}
+\subsection*{Rank Stability Index (RSI)}\label{sec:RSI}
 
@@
 So, RSI is calculated, per element, as $1$ less the quotient of the number of true rank hops taken between the number of maximum possible rank hops, all powered to $p$:
+\begin{linenomath}
 $${\rm RSI} = \left(1-\frac{\text{true rank hops}}{\text{possible rank hops}}\right)^p = \left(1-\frac{D}{(N-1)(t-1)}\right)^p$$
+\end{linenomath}
 where $D$ is the total of rank hops taken by the studied element, $N$ is the number of elements that have been ranked, and $t$ is the number of time samples.
@@
 An alternative way to obtain this effect and exactly map a linear RSI of 0.9 to a non-linear RSI (${\rm RSI'}$) of 0.1, is by applying the following function:
-$${\rm RSI'} = \frac{10^{10\left(1-\frac{D}{(N-1)(t-1)}\right)}-1}{10^{10}-1} \approx 10^{-10\left(\frac{D}{(N-1)(t-1)}\right)} $$
+\begin{linenomath}
+$${\rm RSI'} = \frac{10^{10\left(1-\frac{D}{(N-1)(t-1)}\right)}-1}{10^{10}-1} \approx 10^{-10\left(\frac{D}{(N-1)(t-1)}\right)}$$
+\end{linenomath}
 where the approximation is valid because $10^{10}\gg 1$ but, the small price to pay for it is that, in the worst instability case, the ${\rm RSI'}$ would not be strictly $0$ but $10^{-10}$.
@@
 
-\section{\CC\ output}
+\subsection*{\CC\ output}
 In the previous sections, we have discussed details about the methods used in \CC.
 In this section we review the output from the package, which aims at summarizing the as yet undescribed functionality.
-Unless the flag \texttt{-n}/\texttt{--nodirs} is given to the \CC.py front-end script, the output will be spread into subdirectories (\texttt{corrank}, \texttt{fits}, \texttt{hist} and \texttt{other}) under the results directory.
-The path of the results directory is relative to the data directory, and is given by the optional argument \texttt{-r}/\texttt{--results}, which defaults to ``\texttt{results}''.
-A bunch of plots distributed in subdirectories of such path is generated by \CC\ (the flag that is reducing such number to the minimum is \texttt{-l}/\texttt{--less}).
-They are described in the next subsections.
-In addition, the code provides with output in CSV format (one file per input file), in Excel format (one summary spreadsheet with one sheet per input file, plus correlations and rank tables in the \texttt{corrank} subdirectory) and, optionally, in \LaTeX\ and PDF format (thanks to pdf\LaTeX).
 
-\subsection{Fit Plots}
-The subdirectory \texttt{fits} under the results directory contains the fit plots for each data set.
+\subsubsection*{Fit Plots}
 Both for the unweighted fit (detailed in the section \ref{sec:unw}) and for the X-weighted fit (detailed in the section \ref{sec:X-w}), two plots are generated: the former with logarithmic scale, the latter with lineal scale.
@@
 
-\subsection{Histogram Plots} 
-The subdirectory \texttt{hist} under the results directory contains the histogram plots. \CC\ generates three different histogram plots:
+\subsubsection*{Histogram Plots} 
+\CC\ generates three different histogram plots:
 \begin{description}
@@
 
-\subsection{Correlation and Rank Plots} 
-The subdirectory \texttt{corrank} under the results directory contains the correlation matrix and rank\&stability plots, as well as Excel files with the resulting matrices.
-\CC\ generates three different plots falling under this category:
+\subsubsection*{Correlation and Rank Plots} 
+\CC\ generates three different plots falling under this category, as well as Excel files with the resulting matrices:
 \begin{description}
@@
 
-\subsection{Other Plots} 
-The subdirectory \texttt{other} under the results directory contains other plots, like the elements (taxa) kurtosis and skewness against the mean.
-These plots are generated when the flag \texttt{-m}/\texttt{--more} is given to the \CC.py front-end script.
-
-\subsection{\CC\ advanced options}
-The advanced option flags are summarized here:
-\begin{description}
-	\item[$-$ZRF filter]: This is activated with \texttt{-c}/\texttt{--core} and allows global and detached results for ``core'', ``tailless'' and ``tail'', which allows comparative analysis depending on the presence/absence rate of the taxa throughout time, which is calculated in practice using the Zero Relative Frequency (ZRF) defined above.
-The criteria is as follows:
-	\begin{itemize}
-		\item In any case, if $\operatorname{ZRF} < 0.2$ taxa are classified as ``core''. In other words, the taxa considered ``core'' are those appearing at over $80\%$ of the time points under analysis.
-		\item Depending on the number of times:
-		\begin{itemize}
-			\item If there are only 3 time points, taxa with $0.2 \le\operatorname{ZRF} < 0.4$ are classified as ``tailless'' and, therefore, taxa with $\operatorname{ZRF} \ge 0.4$ are classified as ``tail''.
-			\item If there are 4 time points, taxa with $0.2 \le\operatorname{ZRF} < 0.5$ are classified as ``tailless'' and, therefore, taxa with $\operatorname{ZRF} \ge 0.5$ are classified as ``tail''.
-			\item If there are 5 or more time points, taxa with $0.2 \le\operatorname{ZRF} < 0.6$ are classified as ``tailless'' and, therefore, taxa with $\operatorname{ZRF} \ge 0.6$ are classified as ``tail''.
-		\end{itemize}
-    \end{itemize}
-	\item[$-$Relative frequency filters]: The minimum frequency filter is activated with \texttt{-z}/\texttt{--fmin} and the maximum one with \texttt{-x}/\texttt{--fmax}, with values between $0$ and $1$.
-If both filters are activated simultaneously, the minimum frequency cannot be equal or higher than the maximum one.
-If no filter is activated, by default, \CC\ is filtering all the taxa that never appear throughout time, in the event there are such taxa among the input data.
-On the one hand, if a minimum frequency value is provided, all the taxa whose relative accumulated frequencies (the normalized sum of the values through time) are under such a value, are eliminated.
-But in addition, once the set of data is normalized (to work with relative frequencies), if there are data at any time points that are under the minimum frequency given value, they are automatically floored to zero (we call this \emph{item@time filtering}), and the set of data is renormalized.
-On the other hand, if a maximum frequency value is provided, all the taxa whose relative accumulated frequencies are over such value, are deleted.
-\end{description}
-
-\section{\CC\ input}
-The input data are provided by two simultaneously available ways:
-\begin{enumerate}
-	\item Text files (\texttt{txt} extension), to be processed in parallel\footnote{Currently, parallelization in \CC\ is only available under GNU/Linux}, with a shared-memory model.
-	\item Excel files (\texttt{xls[x]} extension), whose sheets are related datasets to be processed in parallel\footnotemark[\value{footnote}].
-Optionally, all the sheets of an Excel file are standardized using part of such Excel datasets, following the procedure described in the Supplementary Section \ref{sec:stan}, in which event, additional output is provided by \CC.
-\end{enumerate}
-Typically, the data sets are tables with the taxa lying in different rows and the time points ordered in different columns.
-So, the first column is considered the taxon labels, whereas the first row contains the time labels.
-It is worth mentioning that any Excel sheet or text file beginning with the underscore character is ignored by \CC, as they are supposed to hold metadata.
-
-\section{Standardization} \label{sec:stan}
+\subsection*{Standardization} \label{sec:stan}
 In order to show all the studies properly under common axes, we decided to standardize the Taylor parameters using the group of healthy individuals for each study.
@@
 For a Taylor parameter, e.g. $V$, the estimate of the mean ($\widehat{V}$) for the healthy subpopulation, composed of $h$ individuals, is:
+\begin{linenomath}
 $$\widehat{V} = \frac{1}{W_1}\sum_{i=1}^h V_i \omega_i=\sum_{i=1}^h V_i \omega_i$$
+\end{linenomath}
 as $W_1=\sum_i^h \omega_i=1$, since $\omega_i$ are normalized weights calculated as:
+\begin{linenomath}
 $$\omega_i = \frac{\frac{1}{\sigma^2_{V_i}}}{\sum_i^h\frac{1}{\sigma^2_{V_i}}}$$
+\end{linenomath}
 being $\sigma_{V_i}$ the estimation of the uncertainty in $V_i$ obtained together with $V_i$ from the X-weighted power-law fit described in Section \ref{sec:X-w}, for healthy individuals.
@@
 Likewise, the estimation of the standard deviation for the healthy population ($\widehat{\sigma}_V$) is:
+\begin{linenomath}
 $$\widehat{\sigma}_V = \sqrt{\frac{1}{W_1-\frac{W_2}{W_1}}\sum_{i=1}^h\left[\omega_i\left(V_i-\hat{V}\right)^2\right]}$$
+\end{linenomath}
 being $W_2=\sum_i^h \omega_i^2$, which finally yields to:
+\begin{linenomath}
 $$\widehat{\sigma}_V = \sqrt{\frac{1}{1-\sum_i^h \omega_i^2}\sum_{i=1}^h\left[\omega_i\left(V_i-\hat{V}\right)^2\right]}$$
-
-\section{\CC\ interface with LMAT} \label{sec:lmat}
-With the aim of adapting LMAT (Livermore Metagenomics Analysis Toolkit) software\cite{LMAT} for its easy use in time series data, like the ones we have analyzed in this paper\cite{kwashiorkor}, we developed Python front and back-end scripts.
-They are supplied in the directory \texttt{tools} of the \CC\ package.
-
-\subsection{LMAT front-ends} They are: \texttt{pyLMAT\_rl.py}, the launcher for LMAT \texttt{run\_rl.sh}, that is, the first step in the LMAT pipeline or \emph{read labeling}; \texttt{pyLMAT\_cs.py}, the launcher for LMAT \texttt{run\_cs.sh}, the \emph{content summarization} step in the LMAT pipeline; and, \texttt{pyLMAT\_gl.py}, the launcher for LMAT \texttt{run\_gl.sh}, the \emph{gene labeling/identification} step in the LMAT pipeline.
-Basically, all three scripts permit to manage the execution of parallelized LMAT against several fasta files coming from diferent time points in time series studies.
-
-\subsection{LMAT back-ends} Currently, they are: \texttt{pyLMAT\_rescore.py} and the pipeline composed by \texttt{rawlmat2lmat.py} and \texttt{lmat2cmplx.py}.
-The first one allows not only to quick recalculate the LMAT results for different cutoff levels, but also to pull out reads with specific NCBI taxId (\emph{human} reads by default).
-For example, issuing "\texttt{pyLMAT\_rescore.py -p path/to/results -f file.fna -m -1 0 1 -t=32}" will regenerate the LMAT results for cutoffs $-1$, $0$ and $1$, and will pull human reads for each cutoff level, parallelizing in $32$ threads.
-On the other hand, the mentioned pipeline enables to post-process the several \emph{fastsummary} output files from LMAT first and second steps launched by the LMAT front-ends described above.
-The pipeline transforms the results so that they achieve a format and content ready for \CC\ computations.
-Such data arrangements are done independently for two different taxonomic levels: genus and specie.
-
-\newpage
-\begin{thebibliography}{99}
-	\bibitem{ranking} Blumm, N. \textit{et al.} Dynamics of ranking processes in complex systems. {\it Phys. Rev. Lett.} {\bf 109,} 128701 (2012).
-	\bibitem{FD} Weber, J. \textit{et al.} Fluctuation dissipation theorem. {\it Phys. Rev.} {\bf 101}, 1620-6 (1956).
-	\bibitem{kwashiorkor} Smith M.I.  \textit{et al.} Gut microbiomes of Malawian twin pairs discordant for kwashiorkor. {\it Science} {\bf 339,} 548-54 (2013).
-	\bibitem{FASTX} Gordon, A., Hannon, G.J. FASTX-Toolkit. FASTQ/A shortreads pre-processing tools (2010). http://hannonlab.cshl.edu/fastx\_toolkit/ (accessed 23 Feb 2015).
-	\bibitem{QIIME} Caporaso, J.G. \textit{et al.} QIIME allows analysis of high-throughput community sequencing data. {\it Nature Methods} {\bf 7,} 335-6 (2010).
-	\bibitem{SILVA} Quast C.  \textit{et al.} The SILVA ribosomal RNA gene database project: improved data processing and web-based tools (2013)
-	\bibitem{LMAT} Ames, S.K.  \textit{et al.} Scalable metagenomic taxonomy classification usng a reference genome database. {\it Bioinformatics} {\bf 29,} 2253-2260 (2013).
-	\bibitem{antibiotic} Dethlefsen L., Relman D.
-A.
-Incomplete recovery and individualized responses of the human distal gut microbiota to repeated antibiotic perturbation.
-{\it Proc.
-Nat.
-Acad.
-Sci.
-USA} {\bf 108,} 4554-61 (2011).
-	\bibitem{diet} David, L.A. \textit{et al.} Diet rapidly and reproducibly alters the human gut microbiome. {\it Nature} {\bf 505,} 559-63 (2014).
-	\bibitem{IBS} Durban, A. \textit{et al.} Structural alterations of faecal and mucosa-associated bacterial communities in irritable bowel syndrome. {\it FEMS Microbiol. Ecol.} {\bf 86}, 581-9 (2013).
-	\bibitem{LEA} Faith, J.J.  \textit{et al.} The long-term stability of the human gut microbiota. {\it Science} {\bf 341,} 1237439 (2013).
-	\bibitem{moving} Caporaso, J.G.  \textit{et al.} Moving pictures of the human microbiome. {\it Genome Biol.} {\bf 12,} R50 (2011).
- 	\bibitem{ecology} Xiao Xiao, Ethan P.
-White, Mevin B.
-Hooten, and Susan L.
-Durham.
-On the use of log-transformation vs.
-nonlinear regression for analyzing biological power laws.
-{\it Ecology} {\bf 92}, 10, 1887-1894 (2011).
- 	\bibitem{genR2} Magee L., $R^2$ measures based on wald and likelihood ratio joint significance tests. {\it The American Statistician} {\bf 44}, 3, 250-253 (1990).
-	\bibitem{disR2} Nagelkerke N.J.D., A note on a general definition of the coefficient of determination. {\it Biometrika} {\bf 78}, 3, 691-692 (1991).
-	\bibitem{boot} Wu, C.F.J. Jackknife, bootstrap and other resampling methods in regression analysis. (with discussions) \textit{The Annals of Statistics} {\bf 14}: 1261?1350 (1986)
-\end{thebibliography}
-
-\end{document}
+\end{linenomath}

### 2016-04-17 mSystems/mSys_discussion.tex
MINOR: rename mySys_ file prefix to mSys_ prefix
On branch master
	renamed:    mySys_MAIN.pdf -> mSys_MAIN.pdf
	renamed:    mySys_MAIN.tex -> mSys_MAIN.tex
	renamed:    mySys_MandM.tex -> mSys_MandM.tex
	renamed:    mySys_discussion.tex -> mSys_discussion.tex
	renamed:    mySys_intro.tex -> mSys_intro.tex
	renamed:    mySys_results.tex -> mSys_results.tex
	renamed:    mySys_suppl_0.1.tex -> mSys_suppl_0.1.tex

--- 2016-04-15 biorxiv/arxiv_sub.tex
@@
-\documentclass[aps,prl,twocolumn,superscriptaddress,showpacs]{revtex4}
-\usepackage{graphicx}
-\usepackage{mathrsfs}
-\usepackage{bm}
-
-\begin{document}
-
-\title{Health and disease imprinted in the time variability of the human microbiome}
-\author{Jose Manuel Mart\' i}
-\affiliation{Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain}
-\author{Daniel M. Mart\' inez}
-\affiliation{Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain}
-\affiliation{FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain}
-\author{C\'esar Gracia}
-\affiliation{Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain}
-\author{Amparo Latorre}
-\affiliation{FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain}
-\affiliation{Cavanilles Institute of Biodiversity and Evolutionary Biology, Univ. de Valencia, 46980, Spain}
-\affiliation{CIBER en Epidemiolog\' ia y Salud P\'ublica (CIBEResp), Madrid, Spain}
-\author{Andr\'es Moya}
-\affiliation{FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain}
-\affiliation{Cavanilles Institute of Biodiversity and Evolutionary Biology, Univ. de Valencia, 46980, Spain}
-\affiliation{CIBER en Epidemiolog\' ia y Salud P\'ublica (CIBEResp), Madrid, Spain}
-\author{Carlos P. Garay}
-\affiliation{Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain}
-
- \begin{abstract}
-Human microbiota plays an important role in determining changes from health to disease. Increasing research activity is dedicated to understand its diversity and variability. 
-We analyse 16S rRNA and whole genome sequencing (WGS) data from the gut microbiota of 97 individuals monitored in time. Temporal fluctuations in the microbiome reveal 
-significant differences due to factors that affect the microbiota such as dietary changes, antibiotic intake, early gut development or disease. Here we show that a fluctuation 
-scaling law describes the temporal variability of the system and that a noise-induced phase transition is central in the route to disease. The universal law distinguishes healthy 
-from sick microbiota and quantitatively characterizes the path in the phase space, which opens up its potential clinical use and, more generally, other technological 
-applications where microbiota plays an important role. 
- \end{abstract}
-
-\maketitle
-
-The desire to understand the factors that influence human health and cause diseases has always been one of the major driving forces of biological research.
-Modern high-throughput sequencing and bioinformatic tools provide a powerful means of understanding how the human microbiome contributes to health and its potential as a target for therapeutic interventions.
-High throughput methods for microbial 16S ribosomal RNA gene and WGS have now begun to reveal the composition of archaeal, bacterial, fungal and viral communities located both, in and on the human body.
-Biology has recently acquired new technological and conceptual tools to investigate, model and understand living organisms at the system level, thanks to the spectacular progress in quantitative techniques, large-scale measurement methods and the integration of experimental and computational approaches.
-Systems Biology has mostly been devoted to the study of well-characterized model organisms but, since the early days of the Human Genome Project it has become clear that applications of system-wide approaches to Human Biology would bring huge opportunities in Medicine.
-Here we present the imprints of disease in macroscopic properties of the system, by studying the temporal variability in the microbiome.
-
-To this end, we have analysed more than 35000 time series of taxa from the gut microbiomes of 97 individuals (sampling from three up to 332 time points), obtained from publicly available high throughput sequencing data on: healthy individuals over a long term span\cite{moving}, people with various degrees of obesity\cite{lea}, twin pairs discordant for kwashiorkor\cite{kwashiorkor}, response to diet changes\cite{diet} or to antibiotic perturbation\cite{antibiotic}, and subjects diagnosed with irritable bowel syndrome (IBS)\cite{durban}.
-We engineered a complete software framework and a web platform, ComplexCruncher, ready to be implemented by other users.  
-We summarize our dataset in ST1-6 (Tables 1-6 in suplemental material).
-The bacteria and archaea taxonomic assignations were obtained by analysing 16S rRNA sequences, which were clustered into operational taxonomic units (OTUs) sharing 97 \% sequence identity using QIIME\cite{qiime}.
-WGS data\cite{kwashiorkor} were analysed and assigned at strain level by the Livermore Metagenomic Analysis Toolkit (LMAT)\cite{lmat}, according to their default quality threshold.
-Genus, with best balance between error assignment and number of taxa, was chosen as our reference taxonomic level.
-We have verified that our conclusions are not significantly affected by selecting family or species as the reference taxonomic level (see Figure 1 in supplemental material).
+We have quantitatively characterized whether the microbiota belongs to a healthy individual or a subject corresponding to an altered or pathological state (i.e., altered diet, antibiotic treatment, early gut development, diagnosed IBS).
+Deciphering the mechanisms of disease requires in depth knowledge of the underlying biological mechanisms.
+We describe here the macroscopic behavior of disease by a noise-induced phase transition with a control parameter that can be measured by the temporal variability of the microbiome.
+The microbiota of healthy individuals and of individuals with pathologies represent different phases separated by this noise-induced phase transition.
+Improved high?throughput sequencing of samples from individuals monitored over time and taxonomic assigning methods will provide a better distinction among pathologies or altered states of the microbiota.
 
 \begin{figure*}
-\includegraphics[width=14cm,angle=0]{comboplot1.pdf}
+\includegraphics[width=0.9\textwidth]{results/comboplot1.pdf}
 \caption{Taylor's law parameter space.
@@
 
-We have analysed the microbiome temporal variability to extract global properties of the system.
-As fluctuations in total counts are plagued by systematic errors we worked on temporal variability of relative abundances for each taxon.
-Our first finding was that, in all cases, changes in relative abundances of taxa follow a universal pattern known as the fluctuation scaling law\cite{fs} or Taylor's  power law\cite{taylor}, i.e., microbiota of all detected taxa follows a power law dependence between mean relative abundance $x_i$ and dispersion $\sigma_i$, $\sigma_i  = V\cdot x_i^{\beta}$.
-While the law is universal, spanning six orders of magnitude in the observed relative abundances, the power law (or scaling) index $\beta$ and the variability $V$ (hereafter Taylor parameters) appear to be correlated with the stability of the community and the health status of the host, which is the main finding in this letter (see Figure 1).
-Taylor parameters describing the temporal variability of the gut microbiome in our sampled individuals are shown in ST1-6.
-Our results hint at a universal behaviour.
-Firstly, the variability (which corresponds to the maximum amplitude of fluctuations) is large, which suggests resilient capacity of the microbiota, and the scaling index is always smaller than one, which means that, more abundant taxa are less volatile than less abundant ones.
-Secondly, Taylor parameters for the microbiome of healthy individuals in different studies are compatible within estimated errors.
-This enables us to define the health zone in the Taylor parameter space.
-We can better visualize the results of individuals from different studies by standardizing their Taylor parameters, where standardization means that each parameter is subtracted by the mean value and divided by the standard deviation of the group of healthy individuals in the study (see Section 12 and ST1-6).
-The zone of health and the standardized Taylor parameters for individuals whose gut microbiota is threatened (i.e., suffering from kwashiorkor, altered diet, antibiotics, IBS) is shown in Figure 1.
-Children developing kwashiorkor show smaller variability than their healthy twins.
-A meat/fish-based diet increases the variability significantly when compared to a plant-based diet.
-All other cases presented increased variability, which is particularly severe, and statistically significant at more than 95\% CL, for obese patients grade III on a diet, individuals taking antibiotics or IBS--diagnosed patients.
-A global property emerges from all worldwide data collected: Taylor parameters characterize the statistical behaviour of microbiome changes.
-We have verified that our conclusions are robust to systematic errors due to taxonomic assignment.
-
 \begin{figure*}
-\includegraphics[width=14cm,angle=0]{comboplot2.pdf}
+\includegraphics[width=0.9\textwidth]{results/comboplot2.pdf}
 \caption{Phase space of microbiota stability.
@@
 
-Taylor's power law has been explained in terms of various effects, all without general consensus.
-It can be shown to have its origin in a mathematical convergence similar to the central limit theorem, so virtually any statistical model designed to produce a Taylor law converge to a Tweedie distribution\cite{stat}, providing a mechanistic explanation based on the statistical theory of errors\cite{convergence1,convergence2,convergence3}.
-To unveil the generic mechanisms that drive different scenarios in the $\beta$--V space, we model the system by assuming that taxon relative abundance follows a Langevin equation with a deterministic term that captures the fitness of each taxon and a randomness term with Gaussian random noise\cite{ranking}.
-Both terms are modelled by power laws, with coefficients that can be interpreted as the taxon fitness $F_i$ and the variability $V$ (see Section 1 in supplemental material).
-When $V$ is sufficiently low, abundances are stable in time.
-Differences in variability $V$ can induce a noise-induced phase transition in relative abundances of taxa.
-The temporal evolution of the probability of a taxon having abundance $x_i$ given its fitness is governed by the Fokker--Planck equation.
-The results of solving this equation show that  the stability is best captured by fitness F and amplitude of fluctuations V phase space (see Figure 2).
-The model predicts two phases for the gut microbiome: a stable phase with large variability that permits some changes in the relative abundances of taxa and an unstable phase with larger variability, above the phase transition, where the order of abundant taxa varies significantly with time.
-The microbiome of all healthy individuals was found to be in the stable phase, while the microbiome of several other individuals was shown to be in the unstable phase.
-In particular, individuals taking antibiotics and IBS--diagnosed patient P2 had the most severe symptoms.
-In this phase diagram, each microbiota state is represented by a point at its measured variability V and inferred fitness F.
-The model predicts high average fitness for all taxa, i.e., taxa are narrowly distributed in F.
-The fitness parameter has been chosen with different values for demonstrative purposes.
-Fitness is larger for the healthiest subjects and smaller for the IBS--diagnosed patients.
+\subsection*{Temporal evolution of model parameters}
 
-In summary, we have quantitatively characterized whether the microbiota belongs to a healthy individual or a subject corresponding to an altered or pathological state (ie, altered diet, antibiotic treatment, early gut development, diagnosed IBS).
-Deciphering the mechanisms of disease requires in depth knowledge of the underlying biological mechanisms.
-We describe here the macroscopic behavior of disease by a noise-induced phase transition with a control parameter that can be measured by the temporal variability of the microbiome.
-The microbiota of healthy individuals and of individuals with pathologies represent different phases separated by this noise-induced phase transition.
-Improved high--throughput sequencing of samples from individuals monitored over time and taxonomic assigning methods will provide a better distinction among pathologies or altered states of the microbiota.
+We have studied the time dependence of the variability $V$ and power law index $\beta$ (see Section \ref{sec:model}) by using a sliding window approach.
+The total number of time points are divided in subsets of five points, where next subset is defined by adding next time sampling and by eliminating the earliest one.
+Both parameters were calculated for each subset against the average time lapse.
+Figure \ref{fig:tempevo1} shows the variability  $V$ as a function of time for the largest sampling: two individuals in the Caporaso's study\cite{moving} corresponding to the gut microbiota of a male (upper plot) and a female (lower plot).
+Figure \ref{fig:tempevo2} shows the time evolution of $V$ for patient P2 of the IBS study\cite{IBS} (upper plot) and patient D in the antibiotics study\cite{antibiotic} (lower plot).
 
-\textit{Acknowledgements.---} This work is supported by Generalitat Valencia Prometeo Grants II/2014/050, II/2014/065, by the Spanish Grants FPA2011-29678 , BFU2012-39816-C02-01 of MINECO and by PITN- GA-2011-289442-INVISIBLES.
-JMM and DMM acknowledge FPI and FISABIO fellowships.
+\begin{figure}
+	\includegraphics[width=1.0\textwidth]{results/sliwin/male_mov.pdf}
+	\hspace*{3mm}\includegraphics[width=0.448\textwidth]{results/sliwin/female_mov.pdf}
+\caption{$V$ as a function of time for the two individuals in the Caporaso's study\cite{moving}: samples of gut microbiome of a male (upper plot) and a female (lower plot).
+Both samples show changes in the variability V with quasi--periodic behavior peaked at about 10 days.
+Variability grows more for the gut microbiota of the male and share a minimal value around 0.1 with the gut microbiota of the female.}
+\label{fig:tempevo1}
+\end{figure}
 
-\begin{thebibliography}{99}
-
-\bibitem{moving} Caporaso, J.G. et al. Moving pictures of the human microbiome. {\it Genome Biol.} {\bf 12,} R50 (2011).
-\bibitem{lea} Faith, J.J. et al. The long-term stability of the human gut microbiota. {\it Science} {\bf 341,} 1237439 (2013).
-\bibitem{kwashiorkor} Smith M.I. et al. Gut microbiomes of Malawian twin pairs discordant for kwashiorkor. {\it Science} {\bf 339,} 548-54 (2013).
-\bibitem{diet} David, L.A. et al. Diet rapidly and reproducibly alters the human gut microbiome. {\it Nature} {\bf 505,} 559-63 (2014).
-\bibitem{antibiotic} Dethlefsen L., Relman D.
-A.
-Incomplete recovery and individualized responses of the human distal gut microbiota to repeated antibiotic perturbation.
-{\it Proc.
-Nat.
-Acad.
-Sci.
-USA} {\bf 108,} 4554-61 (2011).
-\bibitem{durban} Durban, A.
-\textit{et al.} Structural alterations of faecal and mucosa-associated bacterial communities in irritable bowel syndrome.
-{\it FEMS Microbiol.
-Ecol.} {\bf 86}, 581-9 (2013).
-\bibitem{qiime} Caporaso, J.G. et al. QIIME allows analysis of high-throughput community sequencing data. {\it Nature Methods} {\bf 7,} 335-6 (2010).
-\bibitem{lmat} Ames, S.K. et al. Scalable metagenomic taxonomy classification usng a reference genome database. {\it Bioinformatics} {\bf 29,} 2253-2260 (2013).
-\bibitem{fs} Eisler,Z., Bartos,I., Kertesz,J. Fluctuation scaling in complex systems: Taylor's law and beyond. {\it Adv. Phys.} {\bf 57,} 85 (2008).
-\bibitem{taylor} Taylor, L.R. Aggregation, Variance and the mean. {\it Nature} {\bf 189,} 732-35 (1961).
-\bibitem{stat} Jorgensen,B., Martinez,J.R., Tsao,M. Asymptotic behaviour of the variance function. {\it Scand. J. Statist..} {\bf 21,} 223-243 (1994).
-\bibitem{convergence1} Fronczak,A., Fronczak,P. Origins of Taylor's power law for fluctuation scaling in complex systems. {\it Phys. Rev. E} {\bf 81,} 066112 (2010).
-\bibitem{convergence2} Kendal, W.S., Jorgensen,B. Taylor's power law and fluctuation scaling explained by a central-limit-like convergence. {\it Phys. Rev. E} {\bf 83,} 066115 (2011).
-\bibitem{convergence3} Kendal, W.S., Jorgensen,B. Tweedie convergence: A mathematical basis for Taylor's power law. {\it Phys. Rev. E} {\bf 84,} 066120 (2011).
-\bibitem{ranking} Blumm, N. et al. Dynamics of ranking processes in complex systems. {\it Phys. Rev. Lett.} {\bf 109,} 128701 (2012).
-\end{thebibliography}
-
-\end{document}
+\begin{figure}
+	\centering 
+ 	\includegraphics[width=0.8\textwidth]{results/sliwin/patP2_IBS.pdf}
+  	\includegraphics[width=1.0\textwidth]{results/sliwin/patD_antibio.pdf} 
+\caption{$V$ as a function of time for patient P2 of the IBS study\cite{IBS} (upper plot) and patient D in the antibiotics study\cite{antibiotic} (lower plot). The variability 
+of the gut microbiota of P2 decreases from above 0.3 to below 0.2, showing a slow tendency to increase the order of the system.
+Antibiotic intake leaks to a quick increase of variability which lasts for a few days to recover ordering.
+The second antibiotic treatment shows some memory (lower increase of variability) with a slower recovery.
+NOTE: The blue vertical lines in the lower plot are showing the periods of antibiotic treatment.}
+\label{fig:tempevo2}
+\end{figure}

### 2016-04-17 mSystems/mSys_intro.tex
MINOR: rename mySys_ file prefix to mSys_ prefix
On branch master
	renamed:    mySys_MAIN.pdf -> mSys_MAIN.pdf
	renamed:    mySys_MAIN.tex -> mSys_MAIN.tex
	renamed:    mySys_MandM.tex -> mSys_MandM.tex
	renamed:    mySys_discussion.tex -> mSys_discussion.tex
	renamed:    mySys_intro.tex -> mSys_intro.tex
	renamed:    mySys_results.tex -> mSys_results.tex
	renamed:    mySys_suppl_0.1.tex -> mSys_suppl_0.1.tex

--- 2016-04-15 biorxiv/arxiv_sub.tex
@@
-\documentclass[aps,prl,twocolumn,superscriptaddress,showpacs]{revtex4}
-\usepackage{graphicx}
-\usepackage{mathrsfs}
-\usepackage{bm}
-
-\begin{document}
-
-\title{Health and disease imprinted in the time variability of the human microbiome}
-\author{Jose Manuel Mart\' i}
-\affiliation{Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain}
-\author{Daniel M. Mart\' inez}
-\affiliation{Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain}
-\affiliation{FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain}
-\author{C\'esar Gracia}
-\affiliation{Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain}
-\author{Amparo Latorre}
-\affiliation{FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain}
-\affiliation{Cavanilles Institute of Biodiversity and Evolutionary Biology, Univ. de Valencia, 46980, Spain}
-\affiliation{CIBER en Epidemiolog\' ia y Salud P\'ublica (CIBEResp), Madrid, Spain}
-\author{Andr\'es Moya}
-\affiliation{FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain}
-\affiliation{Cavanilles Institute of Biodiversity and Evolutionary Biology, Univ. de Valencia, 46980, Spain}
-\affiliation{CIBER en Epidemiolog\' ia y Salud P\'ublica (CIBEResp), Madrid, Spain}
-\author{Carlos P. Garay}
-\affiliation{Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain}
-
- \begin{abstract}
-Human microbiota plays an important role in determining changes from health to disease. Increasing research activity is dedicated to understand its diversity and variability. 
-We analyse 16S rRNA and whole genome sequencing (WGS) data from the gut microbiota of 97 individuals monitored in time. Temporal fluctuations in the microbiome reveal 
-significant differences due to factors that affect the microbiota such as dietary changes, antibiotic intake, early gut development or disease. Here we show that a fluctuation 
-scaling law describes the temporal variability of the system and that a noise-induced phase transition is central in the route to disease. The universal law distinguishes healthy 
-from sick microbiota and quantitatively characterizes the path in the phase space, which opens up its potential clinical use and, more generally, other technological 
-applications where microbiota plays an important role. 
- \end{abstract}
-
-\maketitle
-
 The desire to understand the factors that influence human health and cause diseases has always been one of the major driving forces of biological research.
-Modern high-throughput sequencing and bioinformatic tools provide a powerful means of understanding how the human microbiome contributes to health and its potential as a target for therapeutic interventions.
+Modern high-throughput sequencing and bioinformatics tools provide a powerful means of understanding how the human microbiome contributes to health and its potential as a target for therapeutic interventions.
 High throughput methods for microbial 16S ribosomal RNA gene and WGS have now begun to reveal the composition of archaeal, bacterial, fungal and viral communities located both, in and on the human body.
@@
 Here we present the imprints of disease in macroscopic properties of the system, by studying the temporal variability in the microbiome.
-
-To this end, we have analysed more than 35000 time series of taxa from the gut microbiomes of 97 individuals (sampling from three up to 332 time points), obtained from publicly available high throughput sequencing data on: healthy individuals over a long term span\cite{moving}, people with various degrees of obesity\cite{lea}, twin pairs discordant for kwashiorkor\cite{kwashiorkor}, response to diet changes\cite{diet} or to antibiotic perturbation\cite{antibiotic}, and subjects diagnosed with irritable bowel syndrome (IBS)\cite{durban}.
-We engineered a complete software framework and a web platform, ComplexCruncher, ready to be implemented by other users.  
-We summarize our dataset in ST1-6 (Tables 1-6 in suplemental material).
-The bacteria and archaea taxonomic assignations were obtained by analysing 16S rRNA sequences, which were clustered into operational taxonomic units (OTUs) sharing 97 \% sequence identity using QIIME\cite{qiime}.
-WGS data\cite{kwashiorkor} were analysed and assigned at strain level by the Livermore Metagenomic Analysis Toolkit (LMAT)\cite{lmat}, according to their default quality threshold.
-Genus, with best balance between error assignment and number of taxa, was chosen as our reference taxonomic level.
-We have verified that our conclusions are not significantly affected by selecting family or species as the reference taxonomic level (see Figure 1 in supplemental material).
-
-\begin{figure*}
-\includegraphics[width=14cm,angle=0]{comboplot1.pdf}
-\caption{Taylor's law parameter space.
-X-weighted power-law fits of the standard deviations versus the mean values for each bacterial genus monitored in time.
-We show the fit for samples from a healthy subject (Figure A) and from a subject diagnosed with irritable bowel syndrome (Figure B) studied in our lab\cite{durban}.
-We have compiled all data studied in this work in Figure C.
-The coloured region (dashed line) corresponds to 68\% (95\%) CL region of healthy individuals in the Taylor parameter space.
-Points with errors place each individual gut microbiome in the Taylor space.
-Note that the parameters have been standardized (stdev units) to the healthy group in each study for demonstrative purposes.}
-\end{figure*}
-
-We have analysed the microbiome temporal variability to extract global properties of the system.
-As fluctuations in total counts are plagued by systematic errors we worked on temporal variability of relative abundances for each taxon.
-Our first finding was that, in all cases, changes in relative abundances of taxa follow a universal pattern known as the fluctuation scaling law\cite{fs} or Taylor's  power law\cite{taylor}, i.e., microbiota of all detected taxa follows a power law dependence between mean relative abundance $x_i$ and dispersion $\sigma_i$, $\sigma_i  = V\cdot x_i^{\beta}$.
-While the law is universal, spanning six orders of magnitude in the observed relative abundances, the power law (or scaling) index $\beta$ and the variability $V$ (hereafter Taylor parameters) appear to be correlated with the stability of the community and the health status of the host, which is the main finding in this letter (see Figure 1).
-Taylor parameters describing the temporal variability of the gut microbiome in our sampled individuals are shown in ST1-6.
-Our results hint at a universal behaviour.
-Firstly, the variability (which corresponds to the maximum amplitude of fluctuations) is large, which suggests resilient capacity of the microbiota, and the scaling index is always smaller than one, which means that, more abundant taxa are less volatile than less abundant ones.
-Secondly, Taylor parameters for the microbiome of healthy individuals in different studies are compatible within estimated errors.
-This enables us to define the health zone in the Taylor parameter space.
-We can better visualize the results of individuals from different studies by standardizing their Taylor parameters, where standardization means that each parameter is subtracted by the mean value and divided by the standard deviation of the group of healthy individuals in the study (see Section 12 and ST1-6).
-The zone of health and the standardized Taylor parameters for individuals whose gut microbiota is threatened (i.e., suffering from kwashiorkor, altered diet, antibiotics, IBS) is shown in Figure 1.
-Children developing kwashiorkor show smaller variability than their healthy twins.
-A meat/fish-based diet increases the variability significantly when compared to a plant-based diet.
-All other cases presented increased variability, which is particularly severe, and statistically significant at more than 95\% CL, for obese patients grade III on a diet, individuals taking antibiotics or IBS--diagnosed patients.
-A global property emerges from all worldwide data collected: Taylor parameters characterize the statistical behaviour of microbiome changes.
-We have verified that our conclusions are robust to systematic errors due to taxonomic assignment.
-
-\begin{figure*}
-\includegraphics[width=14cm,angle=0]{comboplot2.pdf}
-\caption{Phase space of microbiota stability.
-We show the Variability parameter as a function of time for the male in Caporaso's study\cite{moving} (Figure A) and for the patient D in the antibiotics study\cite{antibiotic} (Figure B).
-Periods of antibiotic treatment are shown by blue vertical lines.
-Microbiota states can be placed in the phase space F-V (Figure C).
-The light blue shaded region corresponds to the stable phase, while the grey shaded region is the unstable phase (the phase transition line is calculated for  $\alpha$ = $\beta$ = 0.75).
-We place healthy individuals (green) and individuals whose gut microbiota is threatened (antibiotics, IBS) in the phase space fitness - variability.
-Gut microbiota of healthy individuals over a long term span show a quasi--periodical variability (central period is ten days).
-We show that taking antibiotics (AB1 and AB2 correspond to first and second treatment respectively) induces a phase transition in the gut microbiota, which impacts its future changes.
-We also show an IBS--diagnosed patient transiting from the unstable to the stable phase.}
-\end{figure*}
-
-Taylor's power law has been explained in terms of various effects, all without general consensus.
-It can be shown to have its origin in a mathematical convergence similar to the central limit theorem, so virtually any statistical model designed to produce a Taylor law converge to a Tweedie distribution\cite{stat}, providing a mechanistic explanation based on the statistical theory of errors\cite{convergence1,convergence2,convergence3}.
-To unveil the generic mechanisms that drive different scenarios in the $\beta$--V space, we model the system by assuming that taxon relative abundance follows a Langevin equation with a deterministic term that captures the fitness of each taxon and a randomness term with Gaussian random noise\cite{ranking}.
-Both terms are modelled by power laws, with coefficients that can be interpreted as the taxon fitness $F_i$ and the variability $V$ (see Section 1 in supplemental material).
-When $V$ is sufficiently low, abundances are stable in time.
-Differences in variability $V$ can induce a noise-induced phase transition in relative abundances of taxa.
-The temporal evolution of the probability of a taxon having abundance $x_i$ given its fitness is governed by the Fokker--Planck equation.
-The results of solving this equation show that  the stability is best captured by fitness F and amplitude of fluctuations V phase space (see Figure 2).
-The model predicts two phases for the gut microbiome: a stable phase with large variability that permits some changes in the relative abundances of taxa and an unstable phase with larger variability, above the phase transition, where the order of abundant taxa varies significantly with time.
-The microbiome of all healthy individuals was found to be in the stable phase, while the microbiome of several other individuals was shown to be in the unstable phase.
-In particular, individuals taking antibiotics and IBS--diagnosed patient P2 had the most severe symptoms.
-In this phase diagram, each microbiota state is represented by a point at its measured variability V and inferred fitness F.
-The model predicts high average fitness for all taxa, i.e., taxa are narrowly distributed in F.
-The fitness parameter has been chosen with different values for demonstrative purposes.
-Fitness is larger for the healthiest subjects and smaller for the IBS--diagnosed patients.
-
-In summary, we have quantitatively characterized whether the microbiota belongs to a healthy individual or a subject corresponding to an altered or pathological state (ie, altered diet, antibiotic treatment, early gut development, diagnosed IBS).
-Deciphering the mechanisms of disease requires in depth knowledge of the underlying biological mechanisms.
-We describe here the macroscopic behavior of disease by a noise-induced phase transition with a control parameter that can be measured by the temporal variability of the microbiome.
-The microbiota of healthy individuals and of individuals with pathologies represent different phases separated by this noise-induced phase transition.
-Improved high--throughput sequencing of samples from individuals monitored over time and taxonomic assigning methods will provide a better distinction among pathologies or altered states of the microbiota.
-
-\textit{Acknowledgements.---} This work is supported by Generalitat Valencia Prometeo Grants II/2014/050, II/2014/065, by the Spanish Grants FPA2011-29678 , BFU2012-39816-C02-01 of MINECO and by PITN- GA-2011-289442-INVISIBLES.
-JMM and DMM acknowledge FPI and FISABIO fellowships.
-
-\begin{thebibliography}{99}
-
-\bibitem{moving} Caporaso, J.G. et al. Moving pictures of the human microbiome. {\it Genome Biol.} {\bf 12,} R50 (2011).
-\bibitem{lea} Faith, J.J. et al. The long-term stability of the human gut microbiota. {\it Science} {\bf 341,} 1237439 (2013).
-\bibitem{kwashiorkor} Smith M.I. et al. Gut microbiomes of Malawian twin pairs discordant for kwashiorkor. {\it Science} {\bf 339,} 548-54 (2013).
-\bibitem{diet} David, L.A. et al. Diet rapidly and reproducibly alters the human gut microbiome. {\it Nature} {\bf 505,} 559-63 (2014).
-\bibitem{antibiotic} Dethlefsen L., Relman D.
-A.
-Incomplete recovery and individualized responses of the human distal gut microbiota to repeated antibiotic perturbation.
-{\it Proc.
-Nat.
-Acad.
-Sci.
-USA} {\bf 108,} 4554-61 (2011).
-\bibitem{durban} Durban, A.
-\textit{et al.} Structural alterations of faecal and mucosa-associated bacterial communities in irritable bowel syndrome.
-{\it FEMS Microbiol.
-Ecol.} {\bf 86}, 581-9 (2013).
-\bibitem{qiime} Caporaso, J.G. et al. QIIME allows analysis of high-throughput community sequencing data. {\it Nature Methods} {\bf 7,} 335-6 (2010).
-\bibitem{lmat} Ames, S.K. et al. Scalable metagenomic taxonomy classification usng a reference genome database. {\it Bioinformatics} {\bf 29,} 2253-2260 (2013).
-\bibitem{fs} Eisler,Z., Bartos,I., Kertesz,J. Fluctuation scaling in complex systems: Taylor's law and beyond. {\it Adv. Phys.} {\bf 57,} 85 (2008).
-\bibitem{taylor} Taylor, L.R. Aggregation, Variance and the mean. {\it Nature} {\bf 189,} 732-35 (1961).
-\bibitem{stat} Jorgensen,B., Martinez,J.R., Tsao,M. Asymptotic behaviour of the variance function. {\it Scand. J. Statist..} {\bf 21,} 223-243 (1994).
-\bibitem{convergence1} Fronczak,A., Fronczak,P. Origins of Taylor's power law for fluctuation scaling in complex systems. {\it Phys. Rev. E} {\bf 81,} 066112 (2010).
-\bibitem{convergence2} Kendal, W.S., Jorgensen,B. Taylor's power law and fluctuation scaling explained by a central-limit-like convergence. {\it Phys. Rev. E} {\bf 83,} 066115 (2011).
-\bibitem{convergence3} Kendal, W.S., Jorgensen,B. Tweedie convergence: A mathematical basis for Taylor's power law. {\it Phys. Rev. E} {\bf 84,} 066120 (2011).
-\bibitem{ranking} Blumm, N. et al. Dynamics of ranking processes in complex systems. {\it Phys. Rev. Lett.} {\bf 109,} 128701 (2012).
-\end{thebibliography}
-
-\end{document}

### 2016-04-17 mSystems/mSys_results.tex
MINOR: rename mySys_ file prefix to mSys_ prefix
On branch master
	renamed:    mySys_MAIN.pdf -> mSys_MAIN.pdf
	renamed:    mySys_MAIN.tex -> mSys_MAIN.tex
	renamed:    mySys_MandM.tex -> mSys_MandM.tex
	renamed:    mySys_discussion.tex -> mSys_discussion.tex
	renamed:    mySys_intro.tex -> mSys_intro.tex
	renamed:    mySys_results.tex -> mSys_results.tex
	renamed:    mySys_suppl_0.1.tex -> mSys_suppl_0.1.tex

--- 2016-04-15 biorxiv/arxiv_sub.tex
@@
-\documentclass[aps,prl,twocolumn,superscriptaddress,showpacs]{revtex4}
-\usepackage{graphicx}
-\usepackage{mathrsfs}
-\usepackage{bm}
-
-\begin{document}
-
-\title{Health and disease imprinted in the time variability of the human microbiome}
-\author{Jose Manuel Mart\' i}
-\affiliation{Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain}
-\author{Daniel M. Mart\' inez}
-\affiliation{Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain}
-\affiliation{FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain}
-\author{C\'esar Gracia}
-\affiliation{Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain}
-\author{Amparo Latorre}
-\affiliation{FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain}
-\affiliation{Cavanilles Institute of Biodiversity and Evolutionary Biology, Univ. de Valencia, 46980, Spain}
-\affiliation{CIBER en Epidemiolog\' ia y Salud P\'ublica (CIBEResp), Madrid, Spain}
-\author{Andr\'es Moya}
-\affiliation{FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain}
-\affiliation{Cavanilles Institute of Biodiversity and Evolutionary Biology, Univ. de Valencia, 46980, Spain}
-\affiliation{CIBER en Epidemiolog\' ia y Salud P\'ublica (CIBEResp), Madrid, Spain}
-\author{Carlos P. Garay}
-\affiliation{Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain}
-
- \begin{abstract}
-Human microbiota plays an important role in determining changes from health to disease. Increasing research activity is dedicated to understand its diversity and variability. 
-We analyse 16S rRNA and whole genome sequencing (WGS) data from the gut microbiota of 97 individuals monitored in time. Temporal fluctuations in the microbiome reveal 
-significant differences due to factors that affect the microbiota such as dietary changes, antibiotic intake, early gut development or disease. Here we show that a fluctuation 
-scaling law describes the temporal variability of the system and that a noise-induced phase transition is central in the route to disease. The universal law distinguishes healthy 
-from sick microbiota and quantitatively characterizes the path in the phase space, which opens up its potential clinical use and, more generally, other technological 
-applications where microbiota plays an important role. 
- \end{abstract}
-
-\maketitle
-
-The desire to understand the factors that influence human health and cause diseases has always been one of the major driving forces of biological research.
-Modern high-throughput sequencing and bioinformatic tools provide a powerful means of understanding how the human microbiome contributes to health and its potential as a target for therapeutic interventions.
-High throughput methods for microbial 16S ribosomal RNA gene and WGS have now begun to reveal the composition of archaeal, bacterial, fungal and viral communities located both, in and on the human body.
-Biology has recently acquired new technological and conceptual tools to investigate, model and understand living organisms at the system level, thanks to the spectacular progress in quantitative techniques, large-scale measurement methods and the integration of experimental and computational approaches.
-Systems Biology has mostly been devoted to the study of well-characterized model organisms but, since the early days of the Human Genome Project it has become clear that applications of system-wide approaches to Human Biology would bring huge opportunities in Medicine.
-Here we present the imprints of disease in macroscopic properties of the system, by studying the temporal variability in the microbiome.
-
-To this end, we have analysed more than 35000 time series of taxa from the gut microbiomes of 97 individuals (sampling from three up to 332 time points), obtained from publicly available high throughput sequencing data on: healthy individuals over a long term span\cite{moving}, people with various degrees of obesity\cite{lea}, twin pairs discordant for kwashiorkor\cite{kwashiorkor}, response to diet changes\cite{diet} or to antibiotic perturbation\cite{antibiotic}, and subjects diagnosed with irritable bowel syndrome (IBS)\cite{durban}.
-We engineered a complete software framework and a web platform, ComplexCruncher, ready to be implemented by other users.  
-We summarize our dataset in ST1-6 (Tables 1-6 in suplemental material).
-The bacteria and archaea taxonomic assignations were obtained by analysing 16S rRNA sequences, which were clustered into operational taxonomic units (OTUs) sharing 97 \% sequence identity using QIIME\cite{qiime}.
-WGS data\cite{kwashiorkor} were analysed and assigned at strain level by the Livermore Metagenomic Analysis Toolkit (LMAT)\cite{lmat}, according to their default quality threshold.
-Genus, with best balance between error assignment and number of taxa, was chosen as our reference taxonomic level.
-We have verified that our conclusions are not significantly affected by selecting family or species as the reference taxonomic level (see Figure 1 in supplemental material).
-
-\begin{figure*}
-\includegraphics[width=14cm,angle=0]{comboplot1.pdf}
-\caption{Taylor's law parameter space.
-X-weighted power-law fits of the standard deviations versus the mean values for each bacterial genus monitored in time.
-We show the fit for samples from a healthy subject (Figure A) and from a subject diagnosed with irritable bowel syndrome (Figure B) studied in our lab\cite{durban}.
-We have compiled all data studied in this work in Figure C.
-The coloured region (dashed line) corresponds to 68\% (95\%) CL region of healthy individuals in the Taylor parameter space.
-Points with errors place each individual gut microbiome in the Taylor space.
-Note that the parameters have been standardized (stdev units) to the healthy group in each study for demonstrative purposes.}
-\end{figure*}
-
 We have analysed the microbiome temporal variability to extract global properties of the system.
@@
 This enables us to define the health zone in the Taylor parameter space.
-We can better visualize the results of individuals from different studies by standardizing their Taylor parameters, where standardization means that each parameter is subtracted by the mean value and divided by the standard deviation of the group of healthy individuals in the study (see Section 12 and ST1-6).
+We can better visualize the results of individuals from different studies by standardizing their Taylor parameters, where standardization means that each parameter is subtracted by the mean value and divided by the standard deviation of the group of healthy individuals in the study (see Supplementary Section 12 and ST1-6).
 The zone of health and the standardized Taylor parameters for individuals whose gut microbiota is threatened (i.e., suffering from kwashiorkor, altered diet, antibiotics, IBS) is shown in Figure 1.
@@
 We have verified that our conclusions are robust to systematic errors due to taxonomic assignment.
-
-\begin{figure*}
-\includegraphics[width=14cm,angle=0]{comboplot2.pdf}
-\caption{Phase space of microbiota stability.
-We show the Variability parameter as a function of time for the male in Caporaso's study\cite{moving} (Figure A) and for the patient D in the antibiotics study\cite{antibiotic} (Figure B).
-Periods of antibiotic treatment are shown by blue vertical lines.
-Microbiota states can be placed in the phase space F-V (Figure C).
-The light blue shaded region corresponds to the stable phase, while the grey shaded region is the unstable phase (the phase transition line is calculated for  $\alpha$ = $\beta$ = 0.75).
-We place healthy individuals (green) and individuals whose gut microbiota is threatened (antibiotics, IBS) in the phase space fitness - variability.
-Gut microbiota of healthy individuals over a long term span show a quasi--periodical variability (central period is ten days).
-We show that taking antibiotics (AB1 and AB2 correspond to first and second treatment respectively) induces a phase transition in the gut microbiota, which impacts its future changes.
-We also show an IBS--diagnosed patient transiting from the unstable to the stable phase.}
-\end{figure*}
 
@@
 The results of solving this equation show that  the stability is best captured by fitness F and amplitude of fluctuations V phase space (see Figure 2).
+
 The model predicts two phases for the gut microbiome: a stable phase with large variability that permits some changes in the relative abundances of taxa and an unstable phase with larger variability, above the phase transition, where the order of abundant taxa varies significantly with time.
@@
 Fitness is larger for the healthiest subjects and smaller for the IBS--diagnosed patients.
-
-In summary, we have quantitatively characterized whether the microbiota belongs to a healthy individual or a subject corresponding to an altered or pathological state (ie, altered diet, antibiotic treatment, early gut development, diagnosed IBS).
-Deciphering the mechanisms of disease requires in depth knowledge of the underlying biological mechanisms.
-We describe here the macroscopic behavior of disease by a noise-induced phase transition with a control parameter that can be measured by the temporal variability of the microbiome.
-The microbiota of healthy individuals and of individuals with pathologies represent different phases separated by this noise-induced phase transition.
-Improved high--throughput sequencing of samples from individuals monitored over time and taxonomic assigning methods will provide a better distinction among pathologies or altered states of the microbiota.
-
-\textit{Acknowledgements.---} This work is supported by Generalitat Valencia Prometeo Grants II/2014/050, II/2014/065, by the Spanish Grants FPA2011-29678 , BFU2012-39816-C02-01 of MINECO and by PITN- GA-2011-289442-INVISIBLES.
-JMM and DMM acknowledge FPI and FISABIO fellowships.
-
-\begin{thebibliography}{99}
-
-\bibitem{moving} Caporaso, J.G. et al. Moving pictures of the human microbiome. {\it Genome Biol.} {\bf 12,} R50 (2011).
-\bibitem{lea} Faith, J.J. et al. The long-term stability of the human gut microbiota. {\it Science} {\bf 341,} 1237439 (2013).
-\bibitem{kwashiorkor} Smith M.I. et al. Gut microbiomes of Malawian twin pairs discordant for kwashiorkor. {\it Science} {\bf 339,} 548-54 (2013).
-\bibitem{diet} David, L.A. et al. Diet rapidly and reproducibly alters the human gut microbiome. {\it Nature} {\bf 505,} 559-63 (2014).
-\bibitem{antibiotic} Dethlefsen L., Relman D.
-A.
-Incomplete recovery and individualized responses of the human distal gut microbiota to repeated antibiotic perturbation.
-{\it Proc.
-Nat.
-Acad.
-Sci.
-USA} {\bf 108,} 4554-61 (2011).
-\bibitem{durban} Durban, A.
-\textit{et al.} Structural alterations of faecal and mucosa-associated bacterial communities in irritable bowel syndrome.
-{\it FEMS Microbiol.
-Ecol.} {\bf 86}, 581-9 (2013).
-\bibitem{qiime} Caporaso, J.G. et al. QIIME allows analysis of high-throughput community sequencing data. {\it Nature Methods} {\bf 7,} 335-6 (2010).
-\bibitem{lmat} Ames, S.K. et al. Scalable metagenomic taxonomy classification usng a reference genome database. {\it Bioinformatics} {\bf 29,} 2253-2260 (2013).
-\bibitem{fs} Eisler,Z., Bartos,I., Kertesz,J. Fluctuation scaling in complex systems: Taylor's law and beyond. {\it Adv. Phys.} {\bf 57,} 85 (2008).
-\bibitem{taylor} Taylor, L.R. Aggregation, Variance and the mean. {\it Nature} {\bf 189,} 732-35 (1961).
-\bibitem{stat} Jorgensen,B., Martinez,J.R., Tsao,M. Asymptotic behaviour of the variance function. {\it Scand. J. Statist..} {\bf 21,} 223-243 (1994).
-\bibitem{convergence1} Fronczak,A., Fronczak,P. Origins of Taylor's power law for fluctuation scaling in complex systems. {\it Phys. Rev. E} {\bf 81,} 066112 (2010).
-\bibitem{convergence2} Kendal, W.S., Jorgensen,B. Taylor's power law and fluctuation scaling explained by a central-limit-like convergence. {\it Phys. Rev. E} {\bf 83,} 066115 (2011).
-\bibitem{convergence3} Kendal, W.S., Jorgensen,B. Tweedie convergence: A mathematical basis for Taylor's power law. {\it Phys. Rev. E} {\bf 84,} 066120 (2011).
-\bibitem{ranking} Blumm, N. et al. Dynamics of ranking processes in complex systems. {\it Phys. Rev. Lett.} {\bf 109,} 128701 (2012).
-\end{thebibliography}
-
-\end{document}

### 2016-04-18 mSystems/mSys_MandM.tex
Move text from mSys_MandM.tex to intro, and other minor changes
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_discussion.tex
	modified:   mSys_intro.tex
	new file:   results/finalPlot11.pdf
	new file:   results/finalPlot33_new.pdf

@@
-We have analysed more than 35000 time series of taxa from the gut microbiomes of 97 individuals (sampling from three up to 332 time points), obtained from publicly available high throughput sequencing data on: healthy individuals over a long term span\cite{moving}, people with various degrees of obesity\cite{LEA}, twin pairs discordant for kwashiorkor\cite{kwashiorkor}, response to diet changes\cite{diet} or to antibiotic perturbation\cite{antibiotic}, and subjects diagnosed with irritable bowel syndrome (IBS)\cite{IBS}.
-We engineered a complete software framework and a web platform, ComplexCruncher, ready to be implemented by other users.
-
 We summarize our dataset in ST1-6 (Tables 1-6 in suplemental material).
@@
 
-\todo{Explain better the fluctiation-dissipation theorem}
+\todo{Explain better the fluctuation-dissipation theorem}
 

### 2016-04-18 mSystems/mSys_discussion.tex
Move text from mSys_MandM.tex to intro, and other minor changes
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_discussion.tex
	modified:   mSys_intro.tex
	new file:   results/finalPlot11.pdf
	new file:   results/finalPlot33_new.pdf

@@
 The microbiota of healthy individuals and of individuals with pathologies represent different phases separated by this noise-induced phase transition.
-Improved high?throughput sequencing of samples from individuals monitored over time and taxonomic assigning methods will provide a better distinction among pathologies or altered states of the microbiota.
+Improved high-throughput sequencing of samples from individuals monitored over time and taxonomic assigning methods will provide a better distinction among pathologies or altered states of the microbiota.
 

### 2016-04-18 mSystems/mSys_intro.tex
Move text from mSys_MandM.tex to intro, and other minor changes
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_discussion.tex
	modified:   mSys_intro.tex
	new file:   results/finalPlot11.pdf
	new file:   results/finalPlot33_new.pdf

@@
 Here we present the imprints of disease in macroscopic properties of the system, by studying the temporal variability in the microbiome.
+
+Here we present the imprints of disease in macroscopic properties of the system, by studying the temporal variability in the microbiome.
+We have analyzed more than 35000 time series of taxa from the gut microbiome of 97 individuals (sampling from three up to 332 time points), obtained from publicly available high throughput sequencing data on different conditions: diseases, diets, obese status, antibiotic perturbation and healthy individuals.
+We engineered a complete software framework, ComplexCruncher, to support the analysis of the dynamics of ranking processes in complex systems, which is ready to be utilized by other users.?

### 2016-04-18 mSystems/mSys_MandM.tex
Initial major review to the mSystems paper discussion section
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_discussion.tex
	modified:   mSys_results.tex
	modified:   results/comboplot2.pdf
	deleted:    results/fits/IBS_P2 Metatranscriptores_stdVSmean_xWboot_LOG.pdf
	new file:   results/fits/IBS_P2_Metatranscriptores_stdVSmean_xWboot_LOG.pdf
	modified:   results/fits/IBS_h_A_amplicons_family_stdVSmean_xWboot_LOG.pdf

@@
 \end{figure}
-
-\begin{table} 
-  \begin{center}
-    \begin{tabular}{ccccccc}
-	    \hline
-		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
-		\hline
-		A&$0.26 \pm 0.05$&$0.826 \pm 0.025$&$0.918$&&$3.1 \pm 0.9$&$1.2 \pm 0.6$\\
-		A&$0.32 \pm 0.06$&$0.857 \pm 0.025$&$0.924$&&$4.4 \pm 1.1$&$2.0 \pm 0.6$\\
-		A&$0.194 \pm 0.033$&$0.813 \pm 0.024$&$0.918$&&$1.9 \pm 0.6$&$0.9 \pm 0.6$\\
-		A&$0.24 \pm 0.04$&$0.824 \pm 0.020$&$0.924$&&$2.7 \pm 0.7$&$1.2 \pm 0.5$\\
-		A&$0.34 \pm 0.06$&$0.855 \pm 0.024$&$0.931$&&$4.7 \pm 1.1$&$1.9 \pm 0.6$\\
-		A&$0.30 \pm 0.05$&$0.847 \pm 0.022$&$0.921$&&$3.9 \pm 1.0$&$1.7 \pm 0.5$\\
-		A&$0.133 \pm 0.021$&$0.784 \pm 0.023$&$0.916$&&$0.7 \pm 0.4$&$0.2 \pm 0.6$\\
-		A&$0.25 \pm 0.04$&$0.831 \pm 0.024$&$0.929$&&$3.0 \pm 0.8$&$1.4 \pm 0.6$\\
-		\hline
-		P&$0.23 \pm 0.05$&$0.804 \pm 0.035$&$0.885$&&$2.6 \pm 0.9$&$0.7 \pm 0.8$\\
-		P&$0.097 \pm 0.018$&$0.705 \pm 0.031$&$0.891$&&$0.03 \pm 0.34$&$-1.6 \pm 0.7$\\
-		P&$0.037 \pm 0.006$&$0.642 \pm 0.025$&$0.881$&&$-1.12 \pm 0.11$&$-3.1 \pm 0.6$\\
-		P&$0.118 \pm 0.019$&$0.723 \pm 0.025$&$0.895$&&$0.4 \pm 0.4$&$-1.2 \pm 0.6$\\
-		P&$0.17 \pm 0.04$&$0.78 \pm 0.04$&$0.842$&&$1.5 \pm 0.7$&$0.1 \pm 0.9$\\
-		P&$0.123 \pm 0.020$&$0.757 \pm 0.026$&$0.914$&&$0.5 \pm 0.4$&$-0.4 \pm 0.6$\\
-		P&$0.19 \pm 0.05$&$0.77 \pm 0.04$&$0.871$&&$1.8 \pm 0.9$&$-0.0 \pm 0.9$\\
-		P&$0.121 \pm 0.020$&$0.736 \pm 0.027$&$0.921$&&$0.5 \pm 0.4$&$-0.9 \pm 0.6$\\
-		P&$0.187 \pm 0.034$&$0.771 \pm 0.030$&$0.908$&&$1.8 \pm 0.7$&$-0.1 \pm 0.7$\\
-		P&$0.097 \pm 0.015$&$0.735 \pm 0.025$&$0.922$&&$0.05 \pm 0.28$&$-0.9 \pm 0.6$\\
-	   \hline
-	   \hline
-    \end{tabular}
-  \end{center}
-  \caption{Taylor parameters.
-Individuals with either animal-based (A) or plant-based (P) diets\cite{diet}.
-Previous to diet,  the population sampled is described by $\bar{V} = 0.09 \pm 0.05, \bar{\beta} = 0.77 \pm 0.04$.}
-  \label{tab:diet}
-\end{table}
-
- \begin{table} 
-  \begin{center}
-    \begin{tabular}{ccccccc}
-	    \hline
-		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
-		\hline
-		Ab&$0.35 \pm 0.07$&$0.81 \pm 0.04$&$0.925$&&$4.3 \pm 1.4$&$1.3 \pm 0.9$\\
-		Ab&$0.41 \pm 0.09$&$0.82 \pm 0.04$&$0.908$&&$5.6 \pm 1.8$&$1.6 \pm 0.9$\\
-		Ab&$0.23 \pm 0.04$&$0.770 \pm 0.031$&$0.920$&&$2.1 \pm 0.8$&$0.5 \pm 0.7$\\
-		Ab&$0.165 \pm 0.029$&$0.738 \pm 0.031$&$0.928$&&$0.9 \pm 0.6$&$-0.3 \pm 0.7$\\
-		Ab&$0.34 \pm 0.06$&$0.812 \pm 0.032$&$0.936$&&$4.1 \pm 1.2$&$1.5 \pm 0.7$\\
-		Ab&$0.26 \pm 0.05$&$0.798 \pm 0.033$&$0.931$&&$2.8 \pm 0.9$&$1.1 \pm 0.8$\\
-	    \hline
-	    \hline
-    \end{tabular}
-  \end{center}
-  \caption{Taylor parameters for individuals taking antibiotics\cite{antibiotic}.
-Prior to antibiotics intake, the population sampled is described by $\bar{V} = 0.12 \pm 0.05, \bar{\beta} = 0.75 \pm 0.04$.}
-  \label{tab:antibiotics}
-\end{table}
-
-\begin{table} 
-  \begin{center}
-    \begin{tabular}{ccccccc}
-	    \hline
-		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
-		\hline
-		IBS&$0.204 \pm 0.034$&$0.739 \pm 0.029$&$0.916$&&$7.6 \pm 3.7$&$1.9 \pm 1.2$\\
-		IBS&$0.35 \pm 0.05$&$0.793 \pm 0.023$&$0.935$&&$23.1 \pm 5.9$&$4.0 \pm 0.9$\\
-	     \hline
-	     \hline
-    \end{tabular}
-  \end{center}
-  \caption{Taylor parameters for persons diagnosed with irritable bowel syndrome (IBS)\cite{IBS}.
-Healthy individuals sampled in this study are characterized by $\bar{V} = 0.134 \pm 0.009, \bar{\beta} = 0.691 \pm 0.025$.}
-  \label{tab:IBS}
-\end{table}
-
- \begin{table} 
-  \begin{center}
-    \begin{tabular}{ccccccc}
-	    \hline
-		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
-		\hline
-		DH&$0.27 \pm 0.04$&$0.835 \pm 0.016$&$0.925$&&$0.2 \pm 0.4$&$-1.0 \pm 0.6$\\
-		DH&$0.36 \pm 0.06$&$0.858 \pm 0.015$&$0.929$&&$1.1 \pm 0.6$&$-0.2 \pm 0.5$\\
-		DH&$0.35 \pm 0.06$&$0.859 \pm 0.014$&$0.926$&&$1.0 \pm 0.5$&$-0.1 \pm 0.5$\\
-		DH&$0.25 \pm 0.04$&$0.829 \pm 0.014$&$0.911$&&$0.0 \pm 0.4$&$-1.2 \pm 0.5$\\
-		DH&$0.30 \pm 0.05$&$0.844 \pm 0.014$&$0.920$&&$0.5 \pm 0.4$&$-0.7 \pm 0.5$\\
-		DH&$0.29 \pm 0.05$&$0.850 \pm 0.016$&$0.915$&&$0.4 \pm 0.5$&$-0.5 \pm 0.5$\\
-		DH&$0.28 \pm 0.05$&$0.848 \pm 0.016$&$0.921$&&$0.3 \pm 0.5$&$-0.5 \pm 0.6$\\
-		DH&$0.35 \pm 0.07$&$0.861 \pm 0.017$&$0.918$&&$0.9 \pm 0.6$&$-0.0 \pm 0.6$\\
-		DH&$0.31 \pm 0.04$&$0.833 \pm 0.012$&$0.916$&&$0.6 \pm 0.4$&$-1.1 \pm 0.4$\\
-		DH&$0.33 \pm 0.05$&$0.843 \pm 0.013$&$0.925$&&$0.8 \pm 0.5$&$-0.7 \pm 0.5$\\
-		DH&$0.31 \pm 0.05$&$0.852 \pm 0.014$&$0.925$&&$0.6 \pm 0.5$&$-0.4 \pm 0.5$\\
-		DH&$0.31 \pm 0.05$&$0.853 \pm 0.015$&$0.930$&&$0.6 \pm 0.5$&$-0.4 \pm 0.5$\\
-		DH&$0.203 \pm 0.033$&$0.815 \pm 0.015$&$0.907$&&$-0.44 \pm 0.32$&$-1.7 \pm 0.5$\\
-		\hline
-    \end{tabular}
-  \end{center}
-  \caption{Taylor parameters for the healthy subject of the discordant twins\cite{kwashiorkor}.
-This table continues in Supplementary Table \ref{tab:DK}.
-The population of healthy twins is characterized by $\bar{V} = 0.25 \pm 0.10, \bar{\beta} = 0.863 \pm 0.028$.}
-  \label{tab:DH}
-\end{table}
-
-\begin{table} 
-  \begin{center}
-    \begin{tabular}{ccccccc}
-	    \hline
-		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
-		\hline
-		DK&$0.40 \pm 0.07$&$0.859 \pm 0.017$&$0.926$&&$1.5 \pm 0.7$&$-0.1 \pm 0.6$\\
-		DK&$0.44 \pm 0.08$&$0.868 \pm 0.016$&$0.919$&&$1.8 \pm 0.8$&$0.2 \pm 0.6$\\
-		DK&$0.196 \pm 0.031$&$0.819 \pm 0.014$&$0.916$&&$-0.50 \pm 0.30$&$-1.5 \pm 0.5$\\
-		DK&$0.160 \pm 0.026$&$0.798 \pm 0.015$&$0.904$&&$-0.85 \pm 0.25$&$-2.3 \pm 0.5$\\
-		DK&$0.30 \pm 0.05$&$0.845 \pm 0.014$&$0.924$&&$0.5 \pm 0.4$&$-0.6 \pm 0.5$\\
-		DK&$0.23 \pm 0.04$&$0.834 \pm 0.014$&$0.908$&&$-0.1 \pm 0.4$&$-1.0 \pm 0.5$\\
-		DK&$0.27 \pm 0.05$&$0.848 \pm 0.015$&$0.930$&&$0.2 \pm 0.4$&$-0.5 \pm 0.5$\\
-		DK&$0.35 \pm 0.07$&$0.860 \pm 0.019$&$0.916$&&$1.0 \pm 0.7$&$-0.1 \pm 0.7$\\
-		DK&$0.34 \pm 0.05$&$0.835 \pm 0.012$&$0.917$&&$0.9 \pm 0.5$&$-1.0 \pm 0.4$\\
-		DK&$0.25 \pm 0.04$&$0.831 \pm 0.012$&$0.912$&&$0.0 \pm 0.4$&$-1.1 \pm 0.4$\\
-		DK&$0.36 \pm 0.06$&$0.858 \pm 0.013$&$0.918$&&$1.1 \pm 0.5$&$-0.2 \pm 0.5$\\
-		DK&$0.31 \pm 0.06$&$0.851 \pm 0.016$&$0.924$&&$0.6 \pm 0.6$&$-0.4 \pm 0.6$\\
-		DK&$0.149 \pm 0.022$&$0.799 \pm 0.013$&$0.905$&&$-0.96 \pm 0.22$&$-2.2 \pm 0.5$\\
-	    \hline
-	    \hline
-    \end{tabular}
-  \end{center}
-  \caption{Taylor parameters for the kwashiorkor part of the discordant twins\cite{kwashiorkor}.
-This is a continuation of Supplementary Table \ref{tab:DH}.
-The population of healthy twins is characterized by $\bar{V} = 0.25 \pm 0.10, \bar{\beta} = 0.863 \pm 0.028$.}
-  \label{tab:DK}
-\end{table}
-
-\begin{table} 
-  \begin{center}
-    \begin{tabular}{ccccccc}
-	    \hline
-		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
-		\hline
-		OW&$0.59 \pm 0.12$&$0.894 \pm 0.034$&$0.920$&&$6.6 \pm 2.0$&$2.6 \pm 1.0$\\
-		OW&$0.22 \pm 0.04$&$0.830 \pm 0.030$&$0.904$&&$0.5 \pm 0.6$&$0.7 \pm 0.9$\\
-		\hline
-		OBI&$0.28 \pm 0.04$&$0.855 \pm 0.022$&$0.958$&&$1.5 \pm 0.6$&$1.4 \pm 0.6$\\
-		OBI&$0.33 \pm 0.07$&$0.870 \pm 0.031$&$0.916$&&$2.4 \pm 1.1$&$1.9 \pm 0.9$\\
-		\hline
-		OBII&$0.223 \pm 0.032$&$0.823 \pm 0.023$&$0.938$&&$0.6 \pm 0.5$&$0.5 \pm 0.7$\\
-		OBII&$0.208 \pm 0.029$&$0.844 \pm 0.022$&$0.935$&&$0.4 \pm 0.5$&$1.1 \pm 0.7$\\
-		\hline
-		OBIII&$0.34 \pm 0.05$&$0.855 \pm 0.025$&$0.943$&&$2.5 \pm 0.9$&$1.4 \pm 0.7$\\
-		OBIII&$0.26 \pm 0.04$&$0.845 \pm 0.026$&$0.954$&&$1.1 \pm 0.7$&$1.2 \pm 0.8$\\
-		OBIII&$0.33 \pm 0.06$&$0.870 \pm 0.027$&$0.908$&&$2.4 \pm 1.0$&$1.9 \pm 0.8$\\
-		OBIII&$0.200 \pm 0.026$&$0.843 \pm 0.020$&$0.949$&&$0.2 \pm 0.4$&$1.1 \pm 0.6$\\
-		OBIII&$0.30 \pm 0.05$&$0.846 \pm 0.026$&$0.929$&&$1.9 \pm 0.8$&$1.2 \pm 0.7$\\
-		OBIII&$0.176 \pm 0.029$&$0.826 \pm 0.026$&$0.894$&&$-0.2 \pm 0.5$&$0.6 \pm 0.8$\\
-		OBIII&$0.30 \pm 0.06$&$0.841 \pm 0.031$&$0.896$&&$1.8 \pm 0.9$&$1.0 \pm 0.9$\\
-		OBIII&$0.28 \pm 0.04$&$0.857 \pm 0.025$&$0.941$&&$1.5 \pm 0.7$&$1.5 \pm 0.7$\\
-		OBIII&$0.122 \pm 0.018$&$0.822 \pm 0.024$&$0.930$&&$-1.05 \pm 0.30$&$0.5 \pm 0.7$\\
-		\hline
-		OBIIId&$0.47 \pm 0.08$&$0.872 \pm 0.023$&$0.945$&&$4.7 \pm 1.3$&$1.9 \pm 0.7$\\
-		OBIIId&$0.38 \pm 0.06$&$0.846 \pm 0.023$&$0.951$&&$3.2 \pm 1.0$&$1.2 \pm 0.7$\\
-		OBIIId&$0.36 \pm 0.06$&$0.842 \pm 0.022$&$0.954$&&$2.9 \pm 0.9$&$1.1 \pm 0.6$\\
-	    \hline
-	    \hline
-    \end{tabular}
-  \end{center}
-  \caption{Taylor parameters for individuals with different degrees of overweight and obesity\cite{LEA}.
-Healthy people in this study, whom were not obese, are characterized by $\bar{V} = 0.19 \pm 0.06, \bar{\beta} = 0.806 \pm 0.034$.}
-  \label{tab:LEA}
-\end{table}
 

### 2016-04-18 mSystems/mSys_discussion.tex
Initial major review to the mSystems paper discussion section
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_discussion.tex
	modified:   mSys_results.tex
	modified:   results/comboplot2.pdf
	deleted:    results/fits/IBS_P2 Metatranscriptores_stdVSmean_xWboot_LOG.pdf
	new file:   results/fits/IBS_P2_Metatranscriptores_stdVSmean_xWboot_LOG.pdf
	modified:   results/fits/IBS_h_A_amplicons_family_stdVSmean_xWboot_LOG.pdf

@@
 Improved high-throughput sequencing of samples from individuals monitored over time and taxonomic assigning methods will provide a better distinction among pathologies or altered states of the microbiota.
-
-\begin{figure*}
-\includegraphics[width=0.9\textwidth]{results/comboplot1.pdf}
-\caption{Taylor's law parameter space.
-X-weighted power-law fits of the standard deviations versus the mean values for each bacterial genus monitored in time.
-We show the fit for samples from a healthy subject (Figure A) and from a subject diagnosed with irritable bowel syndrome (Figure B) studied in our lab\cite{durban}.
-We have compiled all data studied in this work in Figure C.
-The coloured region (dashed line) corresponds to 68\% (95\%) CL region of healthy individuals in the Taylor parameter space.
-Points with errors place each individual gut microbiome in the Taylor space.
-Note that the parameters have been standardized (stdev units) to the healthy group in each study for demonstrative purposes.}
-\end{figure*}
-
-\begin{figure*}
-\includegraphics[width=0.9\textwidth]{results/comboplot2.pdf}
-\caption{Phase space of microbiota stability.
-We show the Variability parameter as a function of time for the male in Caporaso's study\cite{moving} (Figure A) and for the patient D in the antibiotics study\cite{antibiotic} (Figure B).
-Periods of antibiotic treatment are shown by blue vertical lines.
-Microbiota states can be placed in the phase space F-V (Figure C).
-The light blue shaded region corresponds to the stable phase, while the grey shaded region is the unstable phase (the phase transition line is calculated for  $\alpha$ = $\beta$ = 0.75).
-We place healthy individuals (green) and individuals whose gut microbiota is threatened (antibiotics, IBS) in the phase space fitness - variability.
-Gut microbiota of healthy individuals over a long term span show a quasi--periodical variability (central period is ten days).
-We show that taking antibiotics (AB1 and AB2 correspond to first and second treatment respectively) induces a phase transition in the gut microbiota, which impacts its future changes.
-We also show an IBS--diagnosed patient transiting from the unstable to the stable phase.}
-\end{figure*}
 

### 2016-04-18 mSystems/mSys_results.tex
Initial major review to the mSystems paper discussion section
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_discussion.tex
	modified:   mSys_results.tex
	modified:   results/comboplot2.pdf
	deleted:    results/fits/IBS_P2 Metatranscriptores_stdVSmean_xWboot_LOG.pdf
	new file:   results/fits/IBS_P2_Metatranscriptores_stdVSmean_xWboot_LOG.pdf
	modified:   results/fits/IBS_h_A_amplicons_family_stdVSmean_xWboot_LOG.pdf

@@
 As fluctuations in total counts are plagued by systematic errors we worked on temporal variability of relative abundances for each taxon.
-Our first finding was that, in all cases, changes in relative abundances of taxa follow a universal pattern known as the fluctuation scaling law\cite{fs} or Taylor's  power law\cite{taylor}, i.e., microbiota of all detected taxa follows a power law dependence between mean relative abundance $x_i$ and dispersion $\sigma_i$, $\sigma_i  = V\cdot x_i^{\beta}$.
-While the law is universal, spanning six orders of magnitude in the observed relative abundances, the power law (or scaling) index $\beta$ and the variability $V$ (hereafter Taylor parameters) appear to be correlated with the stability of the community and the health status of the host, which is the main finding in this letter (see Figure 1).
-Taylor parameters describing the temporal variability of the gut microbiome in our sampled individuals are shown in ST1-6.
-Our results hint at a universal behaviour.
+Our first finding was that, in all cases, changes in relative abundances of taxa follow a ubiquitous pattern known as the fluctuation scaling law\cite{fs} or Taylor's  power law\cite{taylor}, i.e., microbiota of all detected taxa follows a power law dependence between mean relative abundance $x_i$ and dispersion $\sigma_i$, $\sigma_i  = V\cdot x_i^{\beta}$.
+The law seem to be ubiquitous, spanning to six orders of magnitude in the observed relative abundances (see Figure \ref{fig:main1}).
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_xWboot_LOG.pdf}
+	\includegraphics[width=0.8\textwidth]{results/fits/IBS_P2_Metatranscriptores_stdVSmean_xWboot_LOG.pdf}
+	\caption{Taylor's law parameter space.
+X-weighted power-law fits of the standard deviations versus the mean values for each bacterial genus monitored in time.
+We show the fit for samples from a healthy subject (top) and from a subject diagnosed with irritable bowel syndrome (bottom), studied in our lab \cite{IBS}.}
+	\label{fig:main1}
+\end{figure}
+
+The power law (or scaling) index $\beta$ and the variability $V$ (hereafter Taylor parameters) appear to be correlated with the stability of the community and the health status of the host, which we consider the main finding exposed in this article (see Figure \ref{fig:main2}).
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\textwidth]{results/finalplot11.pdf}
+	\caption{We have compiled here all the data studied in this work.
+The coloured circle corresponds to 68\% confidence level region of healthy individuals in the Taylor parameter space, while dashed line delimites the 98\% CL region.
+Points with errors place each individual gut microbiome in the Taylor space.
+Note that the parameters have been standardized (stdev units) to the healthy group in each study for demonstrative and comparative purposes.}
+	\label{fig:main2}
+\end{figure}
+
+Taylor parameters describing the temporal variability of the gut microbiome in our sampled individuals are shown in Tables \label{tab:diet} to \label{tab:LEA}.
+Our results hint at an ubiquotous behaviour.
 Firstly, the variability (which corresponds to the maximum amplitude of fluctuations) is large, which suggests resilient capacity of the microbiota, and the scaling index is always smaller than one, which means that, more abundant taxa are less volatile than less abundant ones.
 Secondly, Taylor parameters for the microbiome of healthy individuals in different studies are compatible within estimated errors.
-This enables us to define the health zone in the Taylor parameter space.
-We can better visualize the results of individuals from different studies by standardizing their Taylor parameters, where standardization means that each parameter is subtracted by the mean value and divided by the standard deviation of the group of healthy individuals in the study (see Supplementary Section 12 and ST1-6).
-The zone of health and the standardized Taylor parameters for individuals whose gut microbiota is threatened (i.e., suffering from kwashiorkor, altered diet, antibiotics, IBS) is shown in Figure 1.
+This enables us to define the healthy zone in the Taylor parameter space.
+
+In order to jointly visualize and compare the results of individuals from different studies, their Taylor parameters have been standardized, where standardization means that each parameter is subtracted by the mean value and divided by the standard deviation of the group of healthy individuals for each study (for details of the procedure, please see Standardization subsection in Material and Methods).
+The zone of health and the standardized Taylor parameters for individuals whose gut microbiota is threatened (i.e., suffering from kwashiorkor, altered diet, antibiotics or IBS) is shown in Figure \ref{fig:main2}.
 Children developing kwashiorkor show smaller variability than their healthy twins.
@@
 A global property emerges from all worldwide data collected: Taylor parameters characterize the statistical behaviour of microbiome changes.
-We have verified that our conclusions are robust to systematic errors due to taxonomic assignment.
+In addition, we have verified that our conclusions are robust to systematic errors due to taxonomic assignment.
+
+\begin{table} 
+  \begin{center}
+    \begin{tabular}{ccccccc}
+	    \hline
+		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
+		\hline
+		A&$0.26 \pm 0.05$&$0.826 \pm 0.025$&$0.918$&&$3.1 \pm 0.9$&$1.2 \pm 0.6$\\
+		A&$0.32 \pm 0.06$&$0.857 \pm 0.025$&$0.924$&&$4.4 \pm 1.1$&$2.0 \pm 0.6$\\
+		A&$0.194 \pm 0.033$&$0.813 \pm 0.024$&$0.918$&&$1.9 \pm 0.6$&$0.9 \pm 0.6$\\
+		A&$0.24 \pm 0.04$&$0.824 \pm 0.020$&$0.924$&&$2.7 \pm 0.7$&$1.2 \pm 0.5$\\
+		A&$0.34 \pm 0.06$&$0.855 \pm 0.024$&$0.931$&&$4.7 \pm 1.1$&$1.9 \pm 0.6$\\
+		A&$0.30 \pm 0.05$&$0.847 \pm 0.022$&$0.921$&&$3.9 \pm 1.0$&$1.7 \pm 0.5$\\
+		A&$0.133 \pm 0.021$&$0.784 \pm 0.023$&$0.916$&&$0.7 \pm 0.4$&$0.2 \pm 0.6$\\
+		A&$0.25 \pm 0.04$&$0.831 \pm 0.024$&$0.929$&&$3.0 \pm 0.8$&$1.4 \pm 0.6$\\
+		\hline
+		P&$0.23 \pm 0.05$&$0.804 \pm 0.035$&$0.885$&&$2.6 \pm 0.9$&$0.7 \pm 0.8$\\
+		P&$0.097 \pm 0.018$&$0.705 \pm 0.031$&$0.891$&&$0.03 \pm 0.34$&$-1.6 \pm 0.7$\\
+		P&$0.037 \pm 0.006$&$0.642 \pm 0.025$&$0.881$&&$-1.12 \pm 0.11$&$-3.1 \pm 0.6$\\
+		P&$0.118 \pm 0.019$&$0.723 \pm 0.025$&$0.895$&&$0.4 \pm 0.4$&$-1.2 \pm 0.6$\\
+		P&$0.17 \pm 0.04$&$0.78 \pm 0.04$&$0.842$&&$1.5 \pm 0.7$&$0.1 \pm 0.9$\\
+		P&$0.123 \pm 0.020$&$0.757 \pm 0.026$&$0.914$&&$0.5 \pm 0.4$&$-0.4 \pm 0.6$\\
+		P&$0.19 \pm 0.05$&$0.77 \pm 0.04$&$0.871$&&$1.8 \pm 0.9$&$-0.0 \pm 0.9$\\
+		P&$0.121 \pm 0.020$&$0.736 \pm 0.027$&$0.921$&&$0.5 \pm 0.4$&$-0.9 \pm 0.6$\\
+		P&$0.187 \pm 0.034$&$0.771 \pm 0.030$&$0.908$&&$1.8 \pm 0.7$&$-0.1 \pm 0.7$\\
+		P&$0.097 \pm 0.015$&$0.735 \pm 0.025$&$0.922$&&$0.05 \pm 0.28$&$-0.9 \pm 0.6$\\
+	   \hline
+	   \hline
+    \end{tabular}
+  \end{center}
+  \caption{Taylor parameters.
+Individuals with either animal-based (A) or plant-based (P) diets\cite{diet}.
+Previous to diet,  the population sampled is described by $\bar{V} = 0.09 \pm 0.05, \bar{\beta} = 0.77 \pm 0.04$.}
+  \label{tab:diet}
+\end{table}
+
+ \begin{table} 
+  \begin{center}
+    \begin{tabular}{ccccccc}
+	    \hline
+		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
+		\hline
+		Ab&$0.35 \pm 0.07$&$0.81 \pm 0.04$&$0.925$&&$4.3 \pm 1.4$&$1.3 \pm 0.9$\\
+		Ab&$0.41 \pm 0.09$&$0.82 \pm 0.04$&$0.908$&&$5.6 \pm 1.8$&$1.6 \pm 0.9$\\
+		Ab&$0.23 \pm 0.04$&$0.770 \pm 0.031$&$0.920$&&$2.1 \pm 0.8$&$0.5 \pm 0.7$\\
+		Ab&$0.165 \pm 0.029$&$0.738 \pm 0.031$&$0.928$&&$0.9 \pm 0.6$&$-0.3 \pm 0.7$\\
+		Ab&$0.34 \pm 0.06$&$0.812 \pm 0.032$&$0.936$&&$4.1 \pm 1.2$&$1.5 \pm 0.7$\\
+		Ab&$0.26 \pm 0.05$&$0.798 \pm 0.033$&$0.931$&&$2.8 \pm 0.9$&$1.1 \pm 0.8$\\
+	    \hline
+	    \hline
+    \end{tabular}
+  \end{center}
+  \caption{Taylor parameters for individuals taking antibiotics\cite{antibiotic}.
+Prior to antibiotics intake, the population sampled is described by $\bar{V} = 0.12 \pm 0.05, \bar{\beta} = 0.75 \pm 0.04$.}
+  \label{tab:antibiotics}
+\end{table}
+
+\begin{table} 
+  \begin{center}
+    \begin{tabular}{ccccccc}
+	    \hline
+		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
+		\hline
+		IBS&$0.204 \pm 0.034$&$0.739 \pm 0.029$&$0.916$&&$7.6 \pm 3.7$&$1.9 \pm 1.2$\\
+		IBS&$0.35 \pm 0.05$&$0.793 \pm 0.023$&$0.935$&&$23.1 \pm 5.9$&$4.0 \pm 0.9$\\
+	     \hline
+	     \hline
+    \end{tabular}
+  \end{center}
+  \caption{Taylor parameters for persons diagnosed with irritable bowel syndrome (IBS)\cite{IBS}.
+Healthy individuals sampled in this study are characterized by $\bar{V} = 0.134 \pm 0.009, \bar{\beta} = 0.691 \pm 0.025$.}
+  \label{tab:IBS}
+\end{table}
+
+ \begin{table} 
+  \begin{center}
+    \begin{tabular}{ccccccc}
+	    \hline
+		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
+		\hline
+		DH&$0.27 \pm 0.04$&$0.835 \pm 0.016$&$0.925$&&$0.2 \pm 0.4$&$-1.0 \pm 0.6$\\
+		DH&$0.36 \pm 0.06$&$0.858 \pm 0.015$&$0.929$&&$1.1 \pm 0.6$&$-0.2 \pm 0.5$\\
+		DH&$0.35 \pm 0.06$&$0.859 \pm 0.014$&$0.926$&&$1.0 \pm 0.5$&$-0.1 \pm 0.5$\\
+		DH&$0.25 \pm 0.04$&$0.829 \pm 0.014$&$0.911$&&$0.0 \pm 0.4$&$-1.2 \pm 0.5$\\
+		DH&$0.30 \pm 0.05$&$0.844 \pm 0.014$&$0.920$&&$0.5 \pm 0.4$&$-0.7 \pm 0.5$\\
+		DH&$0.29 \pm 0.05$&$0.850 \pm 0.016$&$0.915$&&$0.4 \pm 0.5$&$-0.5 \pm 0.5$\\
+		DH&$0.28 \pm 0.05$&$0.848 \pm 0.016$&$0.921$&&$0.3 \pm 0.5$&$-0.5 \pm 0.6$\\
+		DH&$0.35 \pm 0.07$&$0.861 \pm 0.017$&$0.918$&&$0.9 \pm 0.6$&$-0.0 \pm 0.6$\\
+		DH&$0.31 \pm 0.04$&$0.833 \pm 0.012$&$0.916$&&$0.6 \pm 0.4$&$-1.1 \pm 0.4$\\
+		DH&$0.33 \pm 0.05$&$0.843 \pm 0.013$&$0.925$&&$0.8 \pm 0.5$&$-0.7 \pm 0.5$\\
+		DH&$0.31 \pm 0.05$&$0.852 \pm 0.014$&$0.925$&&$0.6 \pm 0.5$&$-0.4 \pm 0.5$\\
+		DH&$0.31 \pm 0.05$&$0.853 \pm 0.015$&$0.930$&&$0.6 \pm 0.5$&$-0.4 \pm 0.5$\\
+		DH&$0.203 \pm 0.033$&$0.815 \pm 0.015$&$0.907$&&$-0.44 \pm 0.32$&$-1.7 \pm 0.5$\\
+		\hline
+    \end{tabular}
+  \end{center}
+  \caption{Taylor parameters for the healthy subject of the discordant twins\cite{kwashiorkor}.
+This table continues in Table \ref{tab:DK}.
+The population of healthy twins is characterized by $\bar{V} = 0.25 \pm 0.10, \bar{\beta} = 0.863 \pm 0.028$.}
+  \label{tab:DH}
+\end{table}
+
+\begin{table} 
+  \begin{center}
+    \begin{tabular}{ccccccc}
+	    \hline
+		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
+		\hline
+		DK&$0.40 \pm 0.07$&$0.859 \pm 0.017$&$0.926$&&$1.5 \pm 0.7$&$-0.1 \pm 0.6$\\
+		DK&$0.44 \pm 0.08$&$0.868 \pm 0.016$&$0.919$&&$1.8 \pm 0.8$&$0.2 \pm 0.6$\\
+		DK&$0.196 \pm 0.031$&$0.819 \pm 0.014$&$0.916$&&$-0.50 \pm 0.30$&$-1.5 \pm 0.5$\\
+		DK&$0.160 \pm 0.026$&$0.798 \pm 0.015$&$0.904$&&$-0.85 \pm 0.25$&$-2.3 \pm 0.5$\\
+		DK&$0.30 \pm 0.05$&$0.845 \pm 0.014$&$0.924$&&$0.5 \pm 0.4$&$-0.6 \pm 0.5$\\
+		DK&$0.23 \pm 0.04$&$0.834 \pm 0.014$&$0.908$&&$-0.1 \pm 0.4$&$-1.0 \pm 0.5$\\
+		DK&$0.27 \pm 0.05$&$0.848 \pm 0.015$&$0.930$&&$0.2 \pm 0.4$&$-0.5 \pm 0.5$\\
+		DK&$0.35 \pm 0.07$&$0.860 \pm 0.019$&$0.916$&&$1.0 \pm 0.7$&$-0.1 \pm 0.7$\\
+		DK&$0.34 \pm 0.05$&$0.835 \pm 0.012$&$0.917$&&$0.9 \pm 0.5$&$-1.0 \pm 0.4$\\
+		DK&$0.25 \pm 0.04$&$0.831 \pm 0.012$&$0.912$&&$0.0 \pm 0.4$&$-1.1 \pm 0.4$\\
+		DK&$0.36 \pm 0.06$&$0.858 \pm 0.013$&$0.918$&&$1.1 \pm 0.5$&$-0.2 \pm 0.5$\\
+		DK&$0.31 \pm 0.06$&$0.851 \pm 0.016$&$0.924$&&$0.6 \pm 0.6$&$-0.4 \pm 0.6$\\
+		DK&$0.149 \pm 0.022$&$0.799 \pm 0.013$&$0.905$&&$-0.96 \pm 0.22$&$-2.2 \pm 0.5$\\
+	    \hline
+	    \hline
+    \end{tabular}
+  \end{center}
+  \caption{Taylor parameters for the kwashiorkor part of the discordant twins\cite{kwashiorkor}.
+This is a continuation of Table \ref{tab:DH}.
+The population of healthy twins is characterized by $\bar{V} = 0.25 \pm 0.10, \bar{\beta} = 0.863 \pm 0.028$.}
+  \label{tab:DK}
+\end{table}
+
+\begin{table} 
+  \begin{center}
+    \begin{tabular}{ccccccc}
+	    \hline
+		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
+		\hline
+		OW&$0.59 \pm 0.12$&$0.894 \pm 0.034$&$0.920$&&$6.6 \pm 2.0$&$2.6 \pm 1.0$\\
+		OW&$0.22 \pm 0.04$&$0.830 \pm 0.030$&$0.904$&&$0.5 \pm 0.6$&$0.7 \pm 0.9$\\
+		\hline
+		OBI&$0.28 \pm 0.04$&$0.855 \pm 0.022$&$0.958$&&$1.5 \pm 0.6$&$1.4 \pm 0.6$\\
+		OBI&$0.33 \pm 0.07$&$0.870 \pm 0.031$&$0.916$&&$2.4 \pm 1.1$&$1.9 \pm 0.9$\\
+		\hline
+		OBII&$0.223 \pm 0.032$&$0.823 \pm 0.023$&$0.938$&&$0.6 \pm 0.5$&$0.5 \pm 0.7$\\
+		OBII&$0.208 \pm 0.029$&$0.844 \pm 0.022$&$0.935$&&$0.4 \pm 0.5$&$1.1 \pm 0.7$\\
+		\hline
+		OBIII&$0.34 \pm 0.05$&$0.855 \pm 0.025$&$0.943$&&$2.5 \pm 0.9$&$1.4 \pm 0.7$\\
+		OBIII&$0.26 \pm 0.04$&$0.845 \pm 0.026$&$0.954$&&$1.1 \pm 0.7$&$1.2 \pm 0.8$\\
+		OBIII&$0.33 \pm 0.06$&$0.870 \pm 0.027$&$0.908$&&$2.4 \pm 1.0$&$1.9 \pm 0.8$\\
+		OBIII&$0.200 \pm 0.026$&$0.843 \pm 0.020$&$0.949$&&$0.2 \pm 0.4$&$1.1 \pm 0.6$\\
+		OBIII&$0.30 \pm 0.05$&$0.846 \pm 0.026$&$0.929$&&$1.9 \pm 0.8$&$1.2 \pm 0.7$\\
+		OBIII&$0.176 \pm 0.029$&$0.826 \pm 0.026$&$0.894$&&$-0.2 \pm 0.5$&$0.6 \pm 0.8$\\
+		OBIII&$0.30 \pm 0.06$&$0.841 \pm 0.031$&$0.896$&&$1.8 \pm 0.9$&$1.0 \pm 0.9$\\
+		OBIII&$0.28 \pm 0.04$&$0.857 \pm 0.025$&$0.941$&&$1.5 \pm 0.7$&$1.5 \pm 0.7$\\
+		OBIII&$0.122 \pm 0.018$&$0.822 \pm 0.024$&$0.930$&&$-1.05 \pm 0.30$&$0.5 \pm 0.7$\\
+		\hline
+		OBIIId&$0.47 \pm 0.08$&$0.872 \pm 0.023$&$0.945$&&$4.7 \pm 1.3$&$1.9 \pm 0.7$\\
+		OBIIId&$0.38 \pm 0.06$&$0.846 \pm 0.023$&$0.951$&&$3.2 \pm 1.0$&$1.2 \pm 0.7$\\
+		OBIIId&$0.36 \pm 0.06$&$0.842 \pm 0.022$&$0.954$&&$2.9 \pm 0.9$&$1.1 \pm 0.6$\\
+	    \hline
+	    \hline
+    \end{tabular}
+  \end{center}
+  \caption{Taylor parameters for individuals with different degrees of overweight and obesity\cite{LEA}.
+Healthy people in this study, whom were not obese, are characterized by $\bar{V} = 0.19 \pm 0.06, \bar{\beta} = 0.806 \pm 0.034$.}
+  \label{tab:LEA}
+\end{table}
 
@@
 It can be shown to have its origin in a mathematical convergence similar to the central limit theorem, so virtually any statistical model designed to produce a Taylor law converge to a Tweedie distribution\cite{stat}, providing a mechanistic explanation based on the statistical theory of errors\cite{convergence1,convergence2,convergence3}.
-To unveil the generic mechanisms that drive different scenarios in the $\beta$--V space, we model the system by assuming that taxon relative abundance follows a Langevin equation with a deterministic term that captures the fitness of each taxon and a randomness term with Gaussian random noise\cite{ranking}.
-Both terms are modelled by power laws, with coefficients that can be interpreted as the taxon fitness $F_i$ and the variability $V$ (see Section 1 in supplemental material).
+To unveil the generic mechanisms that drive different scenarios in the $\beta$--$V$ space, we model the system by assuming that taxon relative abundance follows a Langevin equation with a deterministic term that captures the fitness of each taxon and a randomness term with Gaussian random noise\cite{ranking}.
+Both terms are modeled by power laws, with coefficients that can be interpreted as the taxon fitness $F_i$ and the variability $V$ (see Model in Material and Methods).
 When $V$ is sufficiently low, abundances are stable in time.
@@
 The temporal evolution of the probability of a taxon having abundance $x_i$ given its fitness is governed by the Fokker--Planck equation.
-The results of solving this equation show that  the stability is best captured by fitness F and amplitude of fluctuations V phase space (see Figure 2).
+The results of solving this equation show that the stability is best captured by fitness $F$ and amplitude of fluctuations $V$ phase space (see Figure \ref{fig:main3}).
+
+\begin{figure}
+	\centering
+	\includegraphics[width=1.0\textwidth]{results/comboplot2.pdf}
+	\caption{Phase space of microbiota stability.
+We show the Variability parameter as a function of time for the male in Caporaso's study\cite{moving} (Figure A) and for the patient D in the antibiotics study\cite{antibiotic} (Figure B).
+Periods of antibiotic treatment are shown by blue vertical lines.
+Microbiota states can be placed in the phase space $F$--$V$ (Figure C).
+The light blue shaded region corresponds to the stable phase, while the grey shaded region is the unstable phase (the phase transition line is calculated for  $\alpha$ = $\beta$ = 0.75).
+We place healthy individuals (green) and individuals whose gut microbiota is threatened (antibiotics, IBS) in the phase space fitness--variability.
+Gut microbiota of healthy individuals over a long term span show a quasi--periodical variability (central period is ten days).
+We show that taking antibiotics (AB1 and AB2 correspond to first and second treatment respectively) induces a phase transition in the gut microbiota, which impacts its future changes.
+We also show an IBS--diagnosed patient transiting from the unstable to the stable phase.}
+	\label{fig:main3}
+\end{figure}
 
@@
 In particular, individuals taking antibiotics and IBS--diagnosed patient P2 had the most severe symptoms.
-In this phase diagram, each microbiota state is represented by a point at its measured variability V and inferred fitness F.
+In this phase diagram, each microbiota state is represented by a point at its measured variability $V$ and inferred fitness $F$.
 The model predicts high average fitness for all taxa, i.e., taxa are narrowly distributed in F.

### 2016-04-21 mSystems/mSys_intro.tex
first version of intro done, refs no complete

--- 2016-04-17 mSystems/mSys_intro.tex
@@
 The desire to understand the factors that influence human health and cause diseases has always been one of the major driving forces of biological research.
-Modern high-throughput sequencing and bioinformatics tools provide a powerful means of understanding how the human microbiome contributes to health and its potential as a target for therapeutic interventions.
+We are populated by a myriad of microorganisms that are interacting with us in several physiological processes such as metabolism regulation or maturation of the immune system.
+Human microbiota has been suggested to be closely related to diseases like type 2 diabetes [ref], cardiovascular disease (CVD) [ref], irritable bowel syndrome [ref], Crohn?s disease [ref] or some affections as obesity [ref] or malnutrition [ref].
 High throughput methods for microbial 16S ribosomal RNA gene and WGS have now begun to reveal the composition of archaeal, bacterial, fungal and viral communities located both, in and on the human body.
+Modern high-throughput sequencing and bioinformatics tools provide a powerful means of understanding how the human microbiome contributes to health and its potential as a target for therapeutic interventions [ref?].
+
 Biology has recently acquired new technological and conceptual tools to investigate, model and understand living organisms at the system level, thanks to the spectacular progress in quantitative techniques, large-scale measurement methods and the integration of experimental and computational approaches.
-Systems Biology has mostly been devoted to the study of well-characterized model organisms but, since the early days of the Human Genome Project it has become clear that applications of system-wide approaches to Human Biology would bring huge opportunities in Medicine.
+Systems Biology has mostly been devoted to the study of well-characterized model organisms but, since the early days of the Human Genome Project [ref] it has become clear that applications of system-wide approaches to Human Biology would bring huge opportunities in Medicine.
+Great effort has been placed to unveil the general laws governing the behaviour of this complex system [ref].
+Due to his nature, microbiota can be studied under the light of the ecology, where we can find general principles as the Taylor?s law, which relates spatial or temporal variability of the population with its mean.
+This law, also known as fluctuation scale law, is ubiquitous in the natural world and can be found in several systems as cosmic rays [ref1], stock markets [ref2,3], animal populations [ref4, 5, 6], gene expression [ref7], or in the human genome [ref8].
+Taylor?s law has been applied to microbiota in a spatial way in the work of Zhang et al., (2014), where they show that this population tend to be in an aggregated way rather than in a random distribution.
+
 Here we present the imprints of disease in macroscopic properties of the system, by studying the temporal variability in the microbiome.
+We have analyzed more than 35000 time series of taxa from the gut microbiome of 97 individuals obtained from publicly available high throughput sequencing data on different conditions: diseases, diets, obese status, antibiotic perturbation and healthy individuals.
+Having seen that all cases follows Taylor?s law, we use this empirical fact to model how the relative abundances of taxa evolves toward time thanks to the Langevin equation, in a similar way as Blumm et al., did in their (2012).
+We use this mathematical framework to explore the temporal stability of the microbiota in different conditions in order to understand how this affects the healthy status of the subjects.
+Finally, we have engineered a complete software framework, ComplexCruncher, to support the analysis of the dynamics of ranking processes in complex systems, which is ready to be implemented by other users.

### 2016-04-21 mSystems/mSys_MAIN.tex
MINOR: rename mySys_ file prefix to mSys_ prefix
On branch master
	renamed:    mySys_MAIN.pdf -> mSys_MAIN.pdf
	renamed:    mySys_MAIN.tex -> mSys_MAIN.tex
	renamed:    mySys_MandM.tex -> mSys_MandM.tex
	renamed:    mySys_discussion.tex -> mSys_discussion.tex
	renamed:    mySys_intro.tex -> mSys_intro.tex
	renamed:    mySys_results.tex -> mSys_results.tex
	renamed:    mySys_suppl_0.1.tex -> mSys_suppl_0.1.tex
For mSystems paper, include new watermark with the release number
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
Move text from mSys_MandM.tex to intro, and other

--- 2016-04-17 mSystems/mySys_MAIN.tex
@@
 \documentclass[12pt,oneside,letterpaper]{article}
+
+% release codenumber (update it here!)
+\newcommand\releasecode{0.0.3}
 
@@
 \geometry{textheight=22.5cm} % 22cm for single-space, 22.5cm for double-space
+
+% watermarks (time consuming!)
+%\newwatermark[allpages,color=red!4,angle=45,scale=9,xpos=-40,ypos=0]{DRAFT}
+%\newwatermark[allpages,color=red!12,angle=0,scale=1,xpos=0,ypos=137]{
+%\emph{mSystems} \LaTeX\ manuscript release \releasecode}
 
@@
 \newcommand{\todo}[1]{\texttt{\bfseries\textcolor{Orange}{#1}}}
-
-% watermark (time consuming!)
-\newwatermark[allpages,color=red!4,angle=45,scale=9,xpos=-40,ypos=0]{DRAFT}
 
@@
 \section*{Introduction}
-\input{mySys_intro.tex}   
+\input{mSys_intro.tex}   
 %%% Result
 \section*{Results}
-\input{mySys_results.tex}
+\input{mSys_results.tex}
 %%% Discussion
 \section*{Discussion}
-\input{mySys_discussion.tex}    
+\input{mSys_discussion.tex}    
 %%% Material & Methods
 \section*{Materials and Methods}
-\input{mySys_MandM.tex}
+\input{mSys_MandM.tex}
 
@@
 \begin{thebibliography}{99}
+\bibitem{diabetes2} {\bf Qin J, Li Y, Cai Z, Li S, Zhu J, Zhang F, Liang S, Zhang W, Guan Y, Shen D, Peng Y, Zhang D, Jie Z, Wu W, Qin Y, Xue W, Li J, Han L, Lu D, Wu P, Dai Y, Sun X, Li Z, Tang A, Zhong S, Li X, Chen W, Xu R, Wang M, Feng Q, Gong M, Yu J, Zhang Y, Zhang M, Hansen T, Sanchez G, Raes J, Falony G, Okuda S, Almeida M, LeChatelier E, Renault P, Pons N, Batto J-M, Zhang Z, Chen H, Yang R, Zheng W, Li S, Yang H, Wang J, Ehrlich SD, Nielsen R, Pedersen O, Kristiansen K, Wang J.} 2012.
+A metagenome-wide association study of gut microbiota in type 2 diabetes.
+Nature {\bf 490}:55–60.
+\bibitem{CVD} {\bf Brown JM, Hazen SL.} 2015. The Gut Microbial Endocrine Organ: Bacterially Derived Signals Driving Cardiometabolic Diseases. Annu Rev Med {\bf 66}:343–359.
+\bibitem{IBS} {\bf Durbán A, Abellán JJ, Jiménez-Hernández N, Artacho A, Garrigues V, Ortiz V, Ponce J, Latorre A, Moya A.} 2013.
+Instability of the faecal microbiota in diarrhoea-predominant irritable bowel syndrome.
+FEMS Microbiol Ecol {\bf 86}:581–589.
+\bibitem{CD} {\bf Gevers D, Kugathasan S, Denson LA, Vázquez-Baeza Y, Van Treuren W, Ren B, Schwager E, Knights D, Song SJ, Yassour M, Morgan XC, Kostic AD, Luo C, González A, McDonald D, Haberman Y, Walters T, Baker S, Rosh J, Stephens M, Heyman M, Markowitz J, Baldassano R, Griffiths A, Sylvester F, Mack D, Kim S, Crandall W, Hyams J, Huttenhower C, Knight R, Xavier RJ.} 2014.
+The treatment-naive microbiome in new-onset Crohn’s disease.
+Cell Host Microbe {\bf 15}:382–392.
+\bibitem{ob1} {\bf Ridaura VK, Faith JJ, Rey FE, Cheng J, Duncan AE, Kau L, Griffi NW, Lombard V, Henrissat B, Bain JR, Michael J, Ilkayeva O, Semenkovich CF, Funai K, Hayashi DK, Lyle J, Martini MC, Ursell LK, Clemente JC, Treuren W Van, William A, Knight R, Newgard CB, Heath AC, Gordon JI, Kau AL, Griffin NW, Muehlbauer MJ.} 2013.
+Gut Microbiota from Twins Discordant for Obesity Modulate Metabolism in Mice Gut Microbiota from Twins Metabolism in Mice.
+Science {\bf 341}:1241214.
+\bibitem{ob2} {\bf Turnbaugh PJ, Hamady M, Yatsunenko T, Cantarel BL, Duncan A, Ley RE, Sogin ML, Jones WJ, Roe BA, Affourtit JP, Egholm M, Henrissat B, Heath AC, Knight R, Gordon JI.} 2009.
+LETTERS A core gut microbiome in obese and lean twins.
+Nature {\bf 457}:480–484.
+\bibitem{nutr} {\bf Subramanian S, Huq S, Yatsunenko T, Haque R, Mahfuz M, Alam MA, Benezra A, DeStefano J, Meier MF, Muegge BD, Barratt MJ, VanArendonk LG, Zhang Q, Province MA, Petri WA, Ahmed T, Gordon JI.} 2014.
+Persistent gut microbiota immaturity in malnourished Bangladeshi children.
+Nature {\bf 510}:417–21.
 \bibitem{moving} Caporaso, J.G. et al. Moving pictures of the human microbiome. {\it Genome Biol.} {\bf 12,} R50 (2011).
@@
 USA} {\bf 108,} 4554-61 (2011).
-\bibitem{IBS} Durban, A. \textit{et al.} Structural alterations of faecal and mucosa-associated bacterial communities in irritable bowel syndrome. {\it FEMS Microbiol. Ecol.} {\bf 86}, 581-9 (2013).
 \bibitem{QIIME} Caporaso, J.G. et al. QIIME allows analysis of high-throughput community sequencing data. {\it Nature Methods} {\bf 7,} 335-6 (2010).

### 2016-04-21 mSystems/mSys_intro.tex
intro added to MAIN; new refs added to intro

@@
 We are populated by a myriad of microorganisms that are interacting with us in several physiological processes such as metabolism regulation or maturation of the immune system.
-Human microbiota has been suggested to be closely related to diseases like type 2 diabetes [ref], cardiovascular disease (CVD) [ref], irritable bowel syndrome [ref], Crohn?s disease [ref] or some affections as obesity [ref] or malnutrition [ref].
+Human microbiota has been suggested to be closely related to diseases like type 2 diabetes
+\cite{diabetes2}, cardiovascular disease (CVD) \cite{CVD}, irritable bowel syndrome \cite{IBS}, Crohn's disease \cite{CD} or some affections as obesity \cite{ob1, ob2} or malnutrition \cite{nutr}.
 High throughput methods for microbial 16S ribosomal RNA gene and WGS have now begun to reveal the composition of archaeal, bacterial, fungal and viral communities located both, in and on the human body.

### 2016-04-21 mSystems/mSys_MandM.tex
mSystems release 0.0.4
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_discussion.tex
	modified:   mSys_results.tex
	modified:   results/finalPlot33_new.pdf

@@
-We summarize our dataset in ST1-6 (Tables 1-6 in suplemental material).
-The bacteria and archaea taxonomic assignations were obtained by analysing 16S rRNA sequences, which were clustered into operational taxonomic units (OTUs) sharing 97 \% sequence identity using QIIME\cite{QIIME}.
-WGS data\cite{kwashiorkor} were analysed and assigned at strain level by the Livermore Metagenomic Analysis Toolkit (LMAT)\cite{LMAT}, according to their default quality threshold.
-Genus, with best balance between error assignment and number of taxa, was chosen as our reference taxonomic level.
-We have verified that our conclusions are not significantly affected by selecting family or species as the reference taxonomic level (see Figure 1 in supplemental material).
-
-\todo{Specify, in each study treated, the nature of the samples (conditions, timespan between timepoints, subjects).
-Specify, and it is very important, what we consider ?healthy?
-in each study (for example: pre-antibiotics is healthy)}
-
 %%%
@@
 \subsection*{Selection and Methods}
+
+The bacteria and archaea taxonomic assignations were obtained by analysing 16S rRNA sequences, which were clustered into operational taxonomic units (OTUs) sharing 97 \% sequence identity using QIIME\cite{QIIME}.
+WGS data\cite{kwashiorkor} were analysed and assigned at strain level by the Livermore Metagenomic Analysis Toolkit (LMAT)\cite{LMAT}, according to their default quality threshold.
+Genus, with best balance between error assignment and number of taxa, was chosen as our reference taxonomic level.
+We have verified that our conclusions are not significantly affected by selecting family or species as the reference taxonomic level (see Figure \ref{fig:taxlev1}).
+
+\todo{Specify, in each study treated, the nature of the samples (conditions, timespan between timepoints, subjects).
+Specify, and it is very important, what we consider \emph{healthy} in each study (for example: pre-antibiotics is healthy)}
 
@@
 The effective data analysis has been performed with a Python tool developed from scratch to more than 4200 lines of code.
-Implemented following the Object Oriented Programming, paradigm, this software is the back-end of the website described above.
+Implemented following the Object Oriented Programming paradigm, this software is the back-end of the website described above.
 However, it could be run as an independent piece of software since it is built as a Python package provided with a command-line front-end (\CC.py).
@@
 \end{linenomath}
-where $v_i$ is a Gaussian random variable with mean $\mu_i=0$ and standard deviation $\sigma_i=\mathrm{SEM}_i$, as defined previously in this supplementary material.
+where $v_i$ is a Gaussian random variable with mean $\mu_i=0$ and standard deviation $\sigma_i=\mathrm{SEM}_i$, as defined previously.
 For each replicate a complete un-weighted power-law fit is performed, as described in the previous section.

### 2016-04-21 mSystems/mSys_discussion.tex
mSystems release 0.0.4
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_discussion.tex
	modified:   mSys_results.tex
	modified:   results/finalPlot33_new.pdf

@@
 Improved high-throughput sequencing of samples from individuals monitored over time and taxonomic assigning methods will provide a better distinction among pathologies or altered states of the microbiota.
-
-\subsection*{Temporal evolution of model parameters}
-
-We have studied the time dependence of the variability $V$ and power law index $\beta$ (see Section \ref{sec:model}) by using a sliding window approach.
-The total number of time points are divided in subsets of five points, where next subset is defined by adding next time sampling and by eliminating the earliest one.
-Both parameters were calculated for each subset against the average time lapse.
-Figure \ref{fig:tempevo1} shows the variability  $V$ as a function of time for the largest sampling: two individuals in the Caporaso's study\cite{moving} corresponding to the gut microbiota of a male (upper plot) and a female (lower plot).
-Figure \ref{fig:tempevo2} shows the time evolution of $V$ for patient P2 of the IBS study\cite{IBS} (upper plot) and patient D in the antibiotics study\cite{antibiotic} (lower plot).
-
-\begin{figure}
-	\includegraphics[width=1.0\textwidth]{results/sliwin/male_mov.pdf}
-	\hspace*{3mm}\includegraphics[width=0.448\textwidth]{results/sliwin/female_mov.pdf}
-\caption{$V$ as a function of time for the two individuals in the Caporaso's study\cite{moving}: samples of gut microbiome of a male (upper plot) and a female (lower plot).
-Both samples show changes in the variability V with quasi--periodic behavior peaked at about 10 days.
-Variability grows more for the gut microbiota of the male and share a minimal value around 0.1 with the gut microbiota of the female.}
-\label{fig:tempevo1}
-\end{figure}
-
-\begin{figure}
-	\centering 
- 	\includegraphics[width=0.8\textwidth]{results/sliwin/patP2_IBS.pdf}
-  	\includegraphics[width=1.0\textwidth]{results/sliwin/patD_antibio.pdf} 
-\caption{$V$ as a function of time for patient P2 of the IBS study\cite{IBS} (upper plot) and patient D in the antibiotics study\cite{antibiotic} (lower plot). The variability 
-of the gut microbiota of P2 decreases from above 0.3 to below 0.2, showing a slow tendency to increase the order of the system.
-Antibiotic intake leaks to a quick increase of variability which lasts for a few days to recover ordering.
-The second antibiotic treatment shows some memory (lower increase of variability) with a slower recovery.
-NOTE: The blue vertical lines in the lower plot are showing the periods of antibiotic treatment.}
-\label{fig:tempevo2}
-\end{figure}

### 2016-04-21 mSystems/mSys_results.tex
mSystems release 0.0.4
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_discussion.tex
	modified:   mSys_results.tex
	modified:   results/finalPlot33_new.pdf

@@
 
-Taylor parameters describing the temporal variability of the gut microbiome in our sampled individuals are shown in Tables \label{tab:diet} to \label{tab:LEA}.
+Taylor parameters describing the temporal variability of the gut microbiome in our sampled individuals are shown in Tables \ref{tab:diet} to \ref{tab:LEA}.
 Our results hint at an ubiquotous behaviour.
@@
 To unveil the generic mechanisms that drive different scenarios in the $\beta$--$V$ space, we model the system by assuming that taxon relative abundance follows a Langevin equation with a deterministic term that captures the fitness of each taxon and a randomness term with Gaussian random noise\cite{ranking}.
-Both terms are modeled by power laws, with coefficients that can be interpreted as the taxon fitness $F_i$ and the variability $V$ (see Model in Material and Methods).
+Both terms are modeled by power laws, with coefficients that can be interpreted as the taxon fitness $F_i$ and the variability $V$ (see Model under Material and Methods).
 When $V$ is sufficiently low, abundances are stable in time.
@@
 	\centering
-	\includegraphics[width=1.0\textwidth]{results/comboplot2.pdf}
-	\caption{Phase space of microbiota stability.
-We show the Variability parameter as a function of time for the male in Caporaso's study\cite{moving} (Figure A) and for the patient D in the antibiotics study\cite{antibiotic} (Figure B).
-Periods of antibiotic treatment are shown by blue vertical lines.
-Microbiota states can be placed in the phase space $F$--$V$ (Figure C).
+	\includegraphics[width=0.8\textwidth]{results/finalPlot33_new.pdf}
+	\caption{Microbiota states can be placed in the phase space $F$--$V$.
 The light blue shaded region corresponds to the stable phase, while the grey shaded region is the unstable phase (the phase transition line is calculated for  $\alpha$ = $\beta$ = 0.75).
@@
 Fitness is larger for the healthiest subjects and smaller for the IBS--diagnosed patients.
+
+Finally, we have studied the time dependence of the variability $V$ and power law index $\beta$ (see Model under Material and Methods) by using a sliding window approach.
+The total number of time points are divided in subsets of five points, where next subset is defined by adding next time sampling and by eliminating the earliest one.
+Both parameters were calculated for each subset against the average time lapse.
+Figure \ref{fig:tempevo1} shows the variability  $V$ as a function of time for the largest sampling: two individuals in the Caporaso's study\cite{moving} corresponding to the gut microbiota of a male (upper plot) and a female (lower plot).
+Figure \ref{fig:tempevo2} shows the time evolution of $V$ for patient P2 of the IBS study\cite{IBS} (upper plot) and patient D in the antibiotics study\cite{antibiotic} (lower plot).
+
+\begin{figure}
+	\includegraphics[width=1.0\textwidth]{results/sliwin/male_mov.pdf}
+	\hspace*{3mm}\includegraphics[width=0.448\textwidth]{results/sliwin/female_mov.pdf}
+\caption{$V$ as a function of time for the two individuals in the Caporaso's study\cite{moving}: samples of gut microbiome of a male (upper plot) and a female (lower plot).
+Both samples show changes in the variability V with quasi--periodic behavior peaked at about 10 days.
+Variability grows more for the gut microbiota of the male and share a minimal value around 0.1 with the gut microbiota of the female.}
+\label{fig:tempevo1}
+\end{figure}
+
+\begin{figure}
+	\centering 
+ 	\includegraphics[width=0.8\textwidth]{results/sliwin/patP2_IBS.pdf}
+  	\includegraphics[width=1.0\textwidth]{results/sliwin/patD_antibio.pdf} 
+\caption{$V$ as a function of time for patient P2 of the IBS study\cite{IBS} (upper plot) and patient D in the antibiotics study\cite{antibiotic} (lower plot). The variability 
+of the gut microbiota of P2 decreases from above 0.3 to below 0.2, showing a slow tendency to increase the order of the system.
+Antibiotic intake leaks to a quick increase of variability which lasts for a few days to recover ordering.
+The second antibiotic treatment shows some memory (lower increase of variability) with a slower recovery.
+NOTE: The blue vertical lines in the lower plot are showing the periods of antibiotic treatment.}
+\label{fig:tempevo2}
+\end{figure}

### 2016-04-22 mSystems/mSys_results.tex
Changes to mSystems Results section (rel0.0.5)
On branch master
	modified:   ../.gitignore
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_results.tex

@@
 As fluctuations in total counts are plagued by systematic errors we worked on temporal variability of relative abundances for each taxon.
-Our first finding was that, in all cases, changes in relative abundances of taxa follow a ubiquitous pattern known as the fluctuation scaling law\cite{fs} or Taylor's  power law\cite{taylor}, i.e., microbiota of all detected taxa follows a power law dependence between mean relative abundance $x_i$ and dispersion $\sigma_i$, $\sigma_i  = V\cdot x_i^{\beta}$.
-The law seem to be ubiquitous, spanning to six orders of magnitude in the observed relative abundances (see Figure \ref{fig:main1}).
+Our first finding was that, in all cases, changes in relative abundances of taxa follow a ubiquitous pattern known as the fluctuation scaling law\cite{fs} or Taylor's  power law\cite{taylor}, i.e., microbiota of all detected taxa follows $\sigma_i  = V\cdot x_i^{\beta}$, a power law dependence between mean relative abundance $x_i$ and dispersion $\sigma_i$.
+The law seem to be ubiquitous, spanning even to six orders of magnitude in the observed relative abundances (see Figure \ref{fig:main1}).
 
@@
 	\includegraphics[width=0.8\textwidth]{results/fits/IBS_P2_Metatranscriptores_stdVSmean_xWboot_LOG.pdf}
-	\caption{Taylor's law parameter space.
-X-weighted power-law fits of the standard deviations versus the mean values for each bacterial genus monitored in time.
-We show the fit for samples from a healthy subject (top) and from a subject diagnosed with irritable bowel syndrome (bottom), studied in our lab \cite{IBS}.}
+	\caption{X-weighted power-law fits of the standard deviations versus the mean values for each bacterial genus monitored in time.
+We show the fit for samples from a healthy subject (top) and from a subject diagnosed with irritable bowel syndrome (bottom), studied in our lab \cite{IBS}.
+Taylor's power law seems to be ubiquitous, spanning to six orders of magnitude.}
 	\label{fig:main1}
@@
 
-The power law (or scaling) index $\beta$ and the variability $V$ (hereafter Taylor parameters) appear to be correlated with the stability of the community and the health status of the host, which we consider the main finding exposed in this article (see Figure \ref{fig:main2}).
+The power law (or scaling) index $\beta$ and the variability $V$ (hereafter Taylor parameters) appear to be correlated with the stability of the community and related with the health status of the host, which we consider the main finding exposed in this article (see Figure \ref{fig:main2}).
 
@@
 	\includegraphics[width=0.9\textwidth]{results/finalplot11.pdf}
-	\caption{We have compiled here all the data studied in this work.
-The coloured circle corresponds to 68\% confidence level region of healthy individuals in the Taylor parameter space, while dashed line delimites the 98\% CL region.
+	\caption{Taylor's law parameter space.
+We have compiled here all the data studied in this work.
+The coloured circle corresponds to 68\% confidence level (CL) region of healthy individuals in the Taylor parameter space, while dashed line delimites the 98\% CL region.
 Points with errors place each individual gut microbiome in the Taylor space.
@@
 Taylor parameters describing the temporal variability of the gut microbiome in our sampled individuals are shown in Tables \ref{tab:diet} to \ref{tab:LEA}.
-Our results hint at an ubiquotous behaviour.
-Firstly, the variability (which corresponds to the maximum amplitude of fluctuations) is large, which suggests resilient capacity of the microbiota, and the scaling index is always smaller than one, which means that, more abundant taxa are less volatile than less abundant ones.
-Secondly, Taylor parameters for the microbiome of healthy individuals in different studies are compatible within estimated errors.
-This enables us to define the healthy zone in the Taylor parameter space.
-
-In order to jointly visualize and compare the results of individuals from different studies, their Taylor parameters have been standardized, where standardization means that each parameter is subtracted by the mean value and divided by the standard deviation of the group of healthy individuals for each study (for details of the procedure, please see Standardization subsection in Material and Methods).
-The zone of health and the standardized Taylor parameters for individuals whose gut microbiota is threatened (i.e., suffering from kwashiorkor, altered diet, antibiotics or IBS) is shown in Figure \ref{fig:main2}.
-Children developing kwashiorkor show smaller variability than their healthy twins.
-A meat/fish-based diet increases the variability significantly when compared to a plant-based diet.
-All other cases presented increased variability, which is particularly severe, and statistically significant at more than 95\% CL, for obese patients grade III on a diet, individuals taking antibiotics or IBS--diagnosed patients.
-A global property emerges from all worldwide data collected: Taylor parameters characterize the statistical behaviour of microbiome changes.
-In addition, we have verified that our conclusions are robust to systematic errors due to taxonomic assignment.
+Our results hint at an ubiquitous behaviour.
+On the first hand, the variability (which corresponds to the maximum amplitude of fluctuations) is large, which suggests resilient capacity of the microbiota.
+On the other hand, the scaling index is always smaller than one, which means that more abundant taxa are less volatile than less abundant ones.
+In addition, Taylor parameters for the microbiome of healthy individuals in different studies are compatible within estimated errors.
+This enables us to define an area in the Taylor parameter space that we called the \emph{healthy zone}.
 
@@
   \end{center}
-  \caption{Taylor parameters.
-Individuals with either animal-based (A) or plant-based (P) diets\cite{diet}.
-Previous to diet,  the population sampled is described by $\bar{V} = 0.09 \pm 0.05, \bar{\beta} = 0.77 \pm 0.04$.}
+  \caption{Taylor parameters for individuals with either animal-based (A) or plant-based (P) diets\cite{diet}.
+Previous to diet, the population sampled is described by $\bar{V} = 0.09 \pm 0.05, \bar{\beta} = 0.77 \pm 0.04$, which we used to describe the \emph{healthy zone} for this study.}
   \label{tab:diet}
@@
   \caption{Taylor parameters for individuals taking antibiotics\cite{antibiotic}.
-Prior to antibiotics intake, the population sampled is described by $\bar{V} = 0.12 \pm 0.05, \bar{\beta} = 0.75 \pm 0.04$.}
+Prior to antibiotics intake, the population sampled is described by $\bar{V} = 0.12 \pm 0.05, \bar{\beta} = 0.75 \pm 0.04$, which characterize the \emph{healthy zone} for this study.}
   \label{tab:antibiotics}
@@
   \caption{Taylor parameters for persons diagnosed with irritable bowel syndrome (IBS)\cite{IBS}.
-Healthy individuals sampled in this study are characterized by $\bar{V} = 0.134 \pm 0.009, \bar{\beta} = 0.691 \pm 0.025$.}
+Healthy individuals sampled in this study are characterized by $\bar{V} = 0.134 \pm 0.009, \bar{\beta} = 0.691 \pm 0.025$, which we used to define the correspondent \emph{healthy zone}.}
   \label{tab:IBS}
@@
 This table continues in Table \ref{tab:DK}.
-The population of healthy twins is characterized by $\bar{V} = 0.25 \pm 0.10, \bar{\beta} = 0.863 \pm 0.028$.}
+The population of healthy twins is characterized by $\bar{V} = 0.25 \pm 0.10, \bar{\beta} = 0.863 \pm 0.028$, values which we used to describe the \emph{healthy zone} for this study.}
   \label{tab:DH}
@@
   \caption{Taylor parameters for the kwashiorkor part of the discordant twins\cite{kwashiorkor}.
-This is a continuation of Table \ref{tab:DH}.
-The population of healthy twins is characterized by $\bar{V} = 0.25 \pm 0.10, \bar{\beta} = 0.863 \pm 0.028$.}
+This is a continuation of Table \ref{tab:DH}, so that the population of healthy twins is also characterized by $\bar{V} = 0.25 \pm 0.10 and \bar{\beta} = 0.863 \pm 0.028$.}
   \label{tab:DK}
@@
   \caption{Taylor parameters for individuals with different degrees of overweight and obesity\cite{LEA}.
-Healthy people in this study, whom were not obese, are characterized by $\bar{V} = 0.19 \pm 0.06, \bar{\beta} = 0.806 \pm 0.034$.}
+Healthy people in this study, whom were not obese, are characterized by $\bar{V} = 0.19 \pm 0.06, \bar{\beta} = 0.806 \pm 0.034$, which we used to determine the correspondent \emph{healthy zone} for this study.}
   \label{tab:LEA}
@@
 
+In order to jointly visualize and compare the results of individuals from different studies, their Taylor parameters have been standardized, where standardization means that each parameter is subtracted by the mean value and divided by the standard deviation of the group of healthy individuals for each study (for details of the procedure, please see Standardization subsection in Material and Methods).
+The healthy zone and the standardized Taylor parameters for individuals whose gut microbiota is threatened (i.e., suffering from kwashiorkor, altered diet, antibiotics or IBS) is shown in Figure \ref{fig:main2}.
+Children developing kwashiorkor show smaller variability than their healthy twins.
+A meat/fish-based diet increases the variability significantly when compared to a plant-based diet.
+All other cases presented increased variability, which is particularly severe, and statistically significant at more than 95\% CL, for obese patients grade III on a diet, individuals taking antibiotics or IBS--diagnosed patients.
+A global property emerges from all worldwide data collected: Taylor parameters characterize the statistical behaviour of microbiome changes.
+Furthermore, we have verified that our conclusions are robust to systematic errors due to taxonomic assignment.
+
 Taylor's power law has been explained in terms of various effects, all without general consensus.
 It can be shown to have its origin in a mathematical convergence similar to the central limit theorem, so virtually any statistical model designed to produce a Taylor law converge to a Tweedie distribution\cite{stat}, providing a mechanistic explanation based on the statistical theory of errors\cite{convergence1,convergence2,convergence3}.
-To unveil the generic mechanisms that drive different scenarios in the $\beta$--$V$ space, we model the system by assuming that taxon relative abundance follows a Langevin equation with a deterministic term that captures the fitness of each taxon and a randomness term with Gaussian random noise\cite{ranking}.
+To unveil the generic mechanisms that drive different scenarios in the $\beta$--$V$ space, we model the system by assuming that taxon relative abundance follows a Langevin equation with, on the one hand, a deterministic term that captures the fitness of each taxon and, on the other hand, a randomness term associated with Gaussian random noise\cite{ranking}.
 Both terms are modeled by power laws, with coefficients that can be interpreted as the taxon fitness $F_i$ and the variability $V$ (see Model under Material and Methods).
-When $V$ is sufficiently low, abundances are stable in time.
+In this model, when $V$ is sufficiently low, abundances are stable in time.
 Differences in variability $V$ can induce a noise-induced phase transition in relative abundances of taxa.
 The temporal evolution of the probability of a taxon having abundance $x_i$ given its fitness is governed by the Fokker--Planck equation.
-The results of solving this equation show that the stability is best captured by fitness $F$ and amplitude of fluctuations $V$ phase space (see Figure \ref{fig:main3}).
+The results of solving this equation show that the stability is best captured by a phase space determined by fitness $F$ and amplitude of fluctuations $V$ (see Figure \ref{fig:main3}).
 

### 2016-04-25 mSystems/mSys_MandM.tex
In mSystems move section from MandM to Results (v0.0.6)
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_results.tex

@@
 We have selected only those individuals who had three or more time points of data available in databases.
-Metadata of each study is provided in Supplementary Tables \ref{tab:diet} to \ref{tab:LEA}.
-All used 16S rRNA gene sequencing except for the study of the discordant kwashiorkor twins\cite{kwashiorkor} (see Supplementary Tables \ref{tab:DH} and \ref{tab:DK}) where shotgun metagenomic sequencing (SMS) and 16S rRNA were used.
+Metadata of each study is provided in Tables \ref{tab:diet} to \ref{tab:LEA}.
+All used 16S rRNA gene sequencing except for the study of the discordant kwashiorkor twins\cite{kwashiorkor} (see Tables \ref{tab:DH} and \ref{tab:DK}) where shotgun metagenomic sequencing (SMS) and 16S rRNA were used.
 In the latter case we selected to work with SMS data to show that our method is valid regardless of the source of taxonomic information.
@@
 The latter dataset tested was the kwashiorkor discordant twins study\cite{kwashiorkor} for both genus and species taxonomic levels.
-The Supplementary Figures \ref{fig:taxlev1} (overview) and \ref{fig:taxlev2} (detail) plot the comparison between studies (and so, 16S and SMS) and between adjacent taxonomic levels.
+The Figures \ref{fig:taxlev1} (overview) and \ref{fig:taxlev2} (detail) plot the comparison between studies (and so, 16S and SMS) and between adjacent taxonomic levels.
 
@@
 
-\subsection*{\CC\ output}
-In the previous sections, we have discussed details about the methods used in \CC.
-In this section we review the output from the package, which aims at summarizing the as yet undescribed functionality.
-
-\subsubsection*{Fit Plots}
-Both for the unweighted fit (detailed in the section \ref{sec:unw}) and for the X-weighted fit (detailed in the section \ref{sec:X-w}), two plots are generated: the former with logarithmic scale, the latter with lineal scale.
-Figure \ref{fig:unwFit} shows an example of the unweighted fit plots and Figure \ref{fig:X-wFit} does the same with the X-weighted fit plots.
-Additionally, for the unweighted fit, a complete residues analysis is performed, and a 4-in-1 figure is generated as shown in Figure \ref{fig:unwRes}, corresponding to the fit of Figure \ref{fig:unwFit}.
-Among other tests, it allows to check for normality and homoscedasticity of the residues.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_LOG.pdf}
-	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_LIN.pdf}
-	\caption{Example of unweighted fit, shown both in logarithmic and lineal scale plots}
-	\label{fig:unwFit}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_xWboot_LOG.pdf}
-	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_xWboot_LIN.pdf}
-	\caption{The X-weighted fit log and lineal plots corresponding to the fit of Figure \ref{fig:unwFit}}
-	\label{fig:X-wFit}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_RES.pdf}
-	\caption{Residues analysis plot corresponding to the fit of Figure \ref{fig:unwFit}.
-The top-left subplot is a simple residues plot.
-The top-right subplot is a Normal quantiles plot with linear fitting (value of coefficient of determination is provided).
-The bottom-left subplot shows an accumulated ESS (Explained Sum of Squares) plot.
-Finally, the bottom-right subplot is a residues Normal histogram plot.
-This set of subplots allows to check for normality and homoscedasticity of the residues.}
-	\label{fig:unwRes}
-\end{figure}
-
-\subsubsection*{Histogram Plots} 
-\CC\ generates three different histogram plots:
-\begin{description}
-	\item[$-$Absolute frequencies plot]: This plot is useful to visually assess the validity of the time points in terms of the accumulated absolute frequency of the elements (taxa), since absolute frequencies far (much higher or much lower) from those typically observed could mean a sampling problem.
-As an example, Figure \ref{fig:histAFP} shows this histogram for the pre-treatment data (first 7 times) of patient ``D'' in the antibiotics study\cite{antibiotic}.
-	\item[$-$2D deviation plot]: The 2D semi-logarithmic histogram representing deviations from the mean versus the mean itself, is a useful tool in the analysis of the stability of ranking processes in complex systems\cite{ranking}.
-Figure \ref{fig:hist2D} shows this plot for the data used in the fit shown in Figure \ref{fig:unwFit}.
-	\item[$-$Zero relative frequency plot]: We could define the ZRF (Zero Relative Frequency, thereby ranging from 0 to 1) of an element (taxon) as the portion of times where it is zero, i.e., it is not found.
-Attending to all the elements (taxa), we can plot the ZRF histogram, which then lies on the horizontal axis of the plot.
-The vertical axis shows the number of elements (taxa), so the height of a bar represents the number of elements that have determinate ZRF.
-In this respect, the bar over $0.0$ counts the number of elements (taxa) that are present at every time point of the data set (aka ``core''), while the bar over $1.0$ would count the number of elements (taxa) that are never found (this bar never appears because all these ``null'' elements are automatically filtered by the code).
-Figure \ref{fig:histZRF} shows an example of this plot.
-There, we can see that $12$ taxa are present at all the time points of the time series while $9$ taxa basically appear only once.
-So, this plot is clearly useful to notice how the ``core'' is distributed.
-\end{description}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.8\textwidth]{results/hist/antibiotics_h_patD_pre_AbsFreqPlot}
-	\caption{Histogram with the absolute frequencies of the pre-treatment data (7 first times) of patient ``D'' in the antibiotics study\cite{antibiotic}}
-	\label{fig:histAFP}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.8\textwidth]{results/hist/IBS_h_A_amplicons_family_hist2D.pdf}
-	\caption{2D histogram deviation plot of the data used in the fit shown in Figure \ref{fig:unwFit}}
-	\label{fig:hist2D}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.8\textwidth]{results/hist/IBS_h_A_amplicons_family_ZRFhist.pdf}
-	\caption{Histogram with the relative frequency of zero for the elements (taxa) in the data used in the fit shown in Figure \ref{fig:unwFit}}
-	\label{fig:histZRF}
-\end{figure}
-
-\subsubsection*{Correlation and Rank Plots} 
-\CC\ generates three different plots falling under this category, as well as Excel files with the resulting matrices:
-\begin{description}
-	\item[$-$Elements correlation matrix]: This plot shows a correlation matrix among the elements (taxa), calculated with the time as independent variable.
-For these calculations, the data set is not normalized to avoid entering an additional constraint.
-As an example, Figure \ref{fig:corrElm} shows this matrix for the ``core'' elements (taxa) present in the pre-treatment data (first seven times) of patient ``D'' in the antibiotics study\cite{antibiotic}.
-	\item[$-$Times correlation matrix]: This plot presents a correlation matrix among the time points of the data set, calculated with the elements (taxa) as independent variable.
-Again, the data set is not normalized.
-Figure \ref{fig:corrTim} shows this matrix for the ``core'' elements (taxa) present in pre-treatment data (first seven times) of patient ``D'' in the antibiotics study\cite{antibiotic}.
-	\item[$-$Rank dynamics and stability plot]: This plot shows the variation in the rank with time for the most dominant elements (taxa) and their calculated RSI, as discussed in Section \ref{sec:RSI}.
-Figure \ref{fig:corrank} shows this plot for the elements (taxa) in the data used in the fit shown in Figure \ref{fig:unwFit}.
-\end{description}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\textwidth]{results/corrank/antibiotics_h_patD_pre_CORE_ElementsCorr}
-	\caption{Element correlation plot of the pre-treatment data (first seven times) of patient ``D'' in the antibiotics study\cite{antibiotic} for ``core'' taxa}
-	\label{fig:corrElm}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\textwidth]{results/corrank/antibiotics_h_patD_pre_CORE_TimesCorr}
-	\caption{Times correlation plot of the pre-treatment data (first seven times) of patient ``D'' in the antibiotics study\cite{antibiotic} for ``core'' taxa}
-	\label{fig:corrTim}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\textwidth]{results/corrank/IBS_h_A_amplicons_family_Rank}
-	\caption{Matrix showing the rank variation throughout time for the most dominant elements (taxa) and their calculated RSI (as discussed in Section \ref{sec:RSI}) in the data used in the fit shown in Figure \ref{fig:unwFit}}
-	\label{fig:corrank}
-\end{figure}
-
 \subsection*{Standardization} \label{sec:stan}

### 2016-04-25 mSystems/mSys_results.tex
In mSystems move section from MandM to Results (v0.0.6)
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_results.tex

@@
+%%%
+\subsection*{Global results}
+
 We have analysed the microbiome temporal variability to extract global properties of the system.
@@
 
+%%%
+\subsection*{DRAFT: Specific results}
+
+\subsubsection*{Fit Plots}
+Both for the unweighted fit (see Material and Methods section for details) and for the X-weighted fit (detailed in Material and Methods subsection), two plots are generated: the former with logarithmic scale, the latter with lineal scale.
+Figure \ref{fig:unwFit} shows an example of the unweighted fit plots and Figure \ref{fig:X-wFit} does the same with the X-weighted fit plots.
+Additionally, for the unweighted fit, a complete residues analysis is performed, and a 4-in-1 figure is generated as shown in Figure \ref{fig:unwRes}, corresponding to the fit of Figure \ref{fig:unwFit}.
+Among other tests, it allows to check for normality and homoscedasticity of the residues.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_LOG.pdf}
+	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_LIN.pdf}
+	\caption{Example of unweighted fit, shown both in logarithmic and lineal scale plots}
+	\label{fig:unwFit}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_xWboot_LOG.pdf}
+	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_xWboot_LIN.pdf}
+	\caption{The X-weighted fit log and lineal plots corresponding to the fit of Figure \ref{fig:unwFit}}
+	\label{fig:X-wFit}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_RES.pdf}
+	\caption{Residues analysis plot corresponding to the fit of Figure \ref{fig:unwFit}.
+The top-left subplot is a simple residues plot.
+The top-right subplot is a Normal quantiles plot with linear fitting (value of coefficient of determination is provided).
+The bottom-left subplot shows an accumulated ESS (Explained Sum of Squares) plot.
+Finally, the bottom-right subplot is a residues Normal histogram plot.
+This set of subplots allows to check for normality and homoscedasticity of the residues.}
+	\label{fig:unwRes}
+\end{figure}
+
+\subsubsection*{Histogram Plots} 
+\CC\ generates three different histogram plots:
+\begin{description}
+	\item[$-$Absolute frequencies plot]: This plot is useful to visually assess the validity of the time points in terms of the accumulated absolute frequency of the elements (taxa), since absolute frequencies far (much higher or much lower) from those typically observed could mean a sampling problem.
+As an example, Figure \ref{fig:histAFP} shows this histogram for the pre-treatment data (first 7 times) of patient ``D'' in the antibiotics study\cite{antibiotic}.
+	\item[$-$2D deviation plot]: The 2D semi-logarithmic histogram representing deviations from the mean versus the mean itself, is a useful tool in the analysis of the stability of ranking processes in complex systems\cite{ranking}.
+Figure \ref{fig:hist2D} shows this plot for the data used in the fit shown in Figure \ref{fig:unwFit}.
+	\item[$-$Zero relative frequency plot]: We could define the ZRF (Zero Relative Frequency, thereby ranging from 0 to 1) of an element (taxon) as the portion of times where it is zero, i.e., it is not found.
+Attending to all the elements (taxa), we can plot the ZRF histogram, which then lies on the horizontal axis of the plot.
+The vertical axis shows the number of elements (taxa), so the height of a bar represents the number of elements that have determinate ZRF.
+In this respect, the bar over $0.0$ counts the number of elements (taxa) that are present at every time point of the data set (aka ``core''), while the bar over $1.0$ would count the number of elements (taxa) that are never found (this bar never appears because all these ``null'' elements are automatically filtered by the code).
+Figure \ref{fig:histZRF} shows an example of this plot.
+There, we can see that $12$ taxa are present at all the time points of the time series while $9$ taxa basically appear only once.
+So, this plot is clearly useful to notice how the ``core'' is distributed.
+\end{description}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.8\textwidth]{results/hist/antibiotics_h_patD_pre_AbsFreqPlot}
+	\caption{Histogram with the absolute frequencies of the pre-treatment data (7 first times) of patient ``D'' in the antibiotics study\cite{antibiotic}}
+	\label{fig:histAFP}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.8\textwidth]{results/hist/IBS_h_A_amplicons_family_hist2D.pdf}
+	\caption{2D histogram deviation plot of the data used in the fit shown in Figure \ref{fig:unwFit}}
+	\label{fig:hist2D}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.8\textwidth]{results/hist/IBS_h_A_amplicons_family_ZRFhist.pdf}
+	\caption{Histogram with the relative frequency of zero for the elements (taxa) in the data used in the fit shown in Figure \ref{fig:unwFit}}
+	\label{fig:histZRF}
+\end{figure}
+
+\subsubsection*{Correlation and Rank Plots} 
+\CC\ generates three different plots falling under this category, as well as Excel files with the resulting matrices:
+\begin{description}
+	\item[$-$Elements correlation matrix]: This plot shows a correlation matrix among the elements (taxa), calculated with the time as independent variable.
+For these calculations, the data set is not normalized to avoid entering an additional constraint.
+As an example, Figure \ref{fig:corrElm} shows this matrix for the ``core'' elements (taxa) present in the pre-treatment data (first seven times) of patient ``D'' in the antibiotics study\cite{antibiotic}.
+	\item[$-$Times correlation matrix]: This plot presents a correlation matrix among the time points of the data set, calculated with the elements (taxa) as independent variable.
+Again, the data set is not normalized.
+Figure \ref{fig:corrTim} shows this matrix for the ``core'' elements (taxa) present in pre-treatment data (first seven times) of patient ``D'' in the antibiotics study\cite{antibiotic}.
+	\item[$-$Rank dynamics and stability plot]: This plot shows the variation in the rank with time for the most dominant elements (taxa) and their calculated RSI, as discussed in Section \ref{sec:RSI}.
+Figure \ref{fig:corrank} shows this plot for the elements (taxa) in the data used in the fit shown in Figure \ref{fig:unwFit}.
+\end{description}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\textwidth]{results/corrank/antibiotics_h_patD_pre_CORE_ElementsCorr}
+	\caption{Element correlation plot of the pre-treatment data (first seven times) of patient ``D'' in the antibiotics study\cite{antibiotic} for ``core'' taxa}
+	\label{fig:corrElm}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\textwidth]{results/corrank/antibiotics_h_patD_pre_CORE_TimesCorr}
+	\caption{Times correlation plot of the pre-treatment data (first seven times) of patient ``D'' in the antibiotics study\cite{antibiotic} for ``core'' taxa}
+	\label{fig:corrTim}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{results/corrank/IBS_h_A_amplicons_family_Rank}
+	\caption{Matrix showing the rank variation throughout time for the most dominant elements (taxa) and their calculated RSI (as discussed in Section \ref{sec:RSI}) in the data used in the fit shown in Figure \ref{fig:unwFit}}
+	\label{fig:corrank}
+\end{figure}
+
+%%%
+\subsection*{Time dependence of model parameters}
+
 Finally, we have studied the time dependence of the variability $V$ and power law index $\beta$ (see Model under Material and Methods) by using a sliding window approach.

### 2016-04-25 mSystems/mSys_MAIN.tex
mSystems release 0.0.4
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_discussion.tex
	modified:   mSys_results.tex
	modified:   results/finalPlot33_new.pdf
Changes to mSystems Results section (rel0.0.5)
On branch master
	modified:   ../.gitignore
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_results.tex
Daniel M. Martinez -> Daniel Martínez-Martínez
In mSystems move section from MandM to Results (v0.0.6)
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.t

@@
 % release codenumber (update it here!)
-\newcommand\releasecode{0.0.3}
+\newcommand\releasecode{0.0.7}
 
@@
 % text layout
-\usepackage[letterpaper]{geometry}
-\geometry{textwidth=16.25cm} % 15.25cm for single-space, 16.25cm for double-space
-\geometry{textheight=22.5cm} % 22cm for single-space, 22.5cm for double-space
+%\usepackage[letterpaper]{geometry}
+%\geometry{textwidth=16.25cm} % 15.25cm for single-space, 16.25cm for double-space
+%\geometry{textheight=22.5cm} % 22cm for single-space, 22.5cm for double-space
 
 % watermarks (time consuming!)
-%\newwatermark[allpages,color=red!4,angle=45,scale=9,xpos=-40,ypos=0]{DRAFT}
-%\newwatermark[allpages,color=red!12,angle=0,scale=1,xpos=0,ypos=137]{
-%\emph{mSystems} \LaTeX\ manuscript release \releasecode}
+\newwatermark[allpages,color=red!4,angle=45,scale=9,xpos=-40,ypos=0]{DRAFT}
+\newwatermark[allpages,color=red!12,angle=0,scale=1,xpos=0,ypos=137]{
+\emph{mSystems} \LaTeX\ manuscript release \releasecode}
 
@@
 \newcommand{\importance}[1]{
+	{
     \small
@@
     }
-\providecommand{\keywords}[1]{\textbf{\textit{Keywords---}} #1}
+}
+\providecommand{\keywords}[1]{\footnotesize\textbf{\textit{Keywords---}} #1}
 
@@
 \title{
-	\vspace*{0mm} % Helps to include the corresponding author footnote in the title page
+	\vspace*{-16mm} % Helps to include the corresponding author footnote in the title page
 	\singlespacing
@@
 \author[1,2]{Jose Manuel Martí}
-\author[1,2,3]{Daniel M. Martínez}
+\author[1,2,3]{Daniel Martínez-Martínez}
 \author[2]{Manuel Peña}
@@
 \affil[3]{FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain.}
-\affil[4]{Cavanilles Institute of Biodiversity and Evolutionary Biology, Univ. de Valencia, 46980, Spain.}
+\affil[4]{Cavanilles Institute of Biodiversity and Evolutionary Biology, UVEG, 46980, Spain.}
 \affil[5]{CIBER en Epidemiología y Salud Pública (CIBEResp), Madrid, Spain}
@@
 %TC:macro \abstract [abswords]
-\abstract{Human microbiota plays an important role in determining changes from health to disease. Increasing research activity is dedicated to understand its diversity and variability. 
+{\abstract{Human microbiota plays an important role in determining changes from health to disease. Increasing research activity is dedicated to understand its diversity and variability. 
 We analyse 16S rRNA and whole genome sequencing (WGS) data from the gut microbiota of 97 individuals monitored in time. Temporal fluctuations in the microbiome reveal 
@@
 from sick microbiota and quantitatively characterizes the path in the phase space, which opens up its potential clinical use and, more generally, other technological 
-applications where microbiota plays an important role.}
+applications where microbiota plays an important role.}}
 
@@
 Using the Langevin equation as the basis of our mathematical framework in order to evaluate microbial temporal stability, we prove that we are capable to distinguish stable from unstable microbiotas.
-This first step will help us to determine how microbiota temporal stability is related to the healthiness of the people, and it will allow the development of a more complete framework in order to deepen the knowledge of this complex system.
-}
+This first step will help us to determine how microbiota temporal stability is related to the healthiness of the people, and it will allow the development of a more complete framework in order to deepen the knowledge of this complex system.}
 
@@
 \begin{keywords}
-microbiome, systems biology, ecological modelling, community composition, stability	
+microbiome, systems biology, ecological modeling, metagenomics, stability	
 \end{keywords}

### 2016-04-25 mSystems/mSys_MandM.tex
Some work in mSys MAIN (fix font size) and Results (v0.0.7)
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_results.tex
	modified:   results/corrank/IBS_h_A_amplicons_family_ElmCorrelation.pdf
	modified:   results/corrank/IBS_h_A_amplicons_family_Rank.pdf
	modified:   results/corrank/IBS_h_A_amplicons_family_TimesCorr.pdf

@@
   \includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_spc_SMS.pdf}
-\caption{Detail of comparison of different approaches based on adjacent taxonomic levels using plots of X-weighted power-law fits (see Section \ref{sec:X-w}).
+\caption{Detail of comparison of different approaches based on adjacent taxonomic levels using plots of X-weighted power-law fits (see Material and Methods).
 The former row of subfigures shows examples for 16S, whereas the latter row of subfigures plots examples for SMS.

### 2016-04-25 mSystems/mSys_results.tex
Some work in mSys MAIN (fix font size) and Results (v0.0.7)
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_results.tex
	modified:   results/corrank/IBS_h_A_amplicons_family_ElmCorrelation.pdf
	modified:   results/corrank/IBS_h_A_amplicons_family_Rank.pdf
	modified:   results/corrank/IBS_h_A_amplicons_family_TimesCorr.pdf

@@
 	\centering
+	\vspace*{-15mm} % Corrects overbox of the figures
 	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_xWboot_LOG.pdf}
@@
   \caption{Taylor parameters for the kwashiorkor part of the discordant twins\cite{kwashiorkor}.
-This is a continuation of Table \ref{tab:DH}, so that the population of healthy twins is also characterized by $\bar{V} = 0.25 \pm 0.10 and \bar{\beta} = 0.863 \pm 0.028$.}
+This is a continuation of Table \ref{tab:DH}, so that the population of healthy twins is also characterized by $\bar{V} = 0.25 \pm 0.10$ and $\bar{\beta} = 0.863 \pm 0.028$.}
   \label{tab:DK}
@@
 	\centering
+	\vspace*{-5mm} % Corrects overbox of the figures
 	\includegraphics[width=0.8\textwidth]{results/finalPlot33_new.pdf}
@@
 %%%
-\subsection*{DRAFT: Specific results}
+\subsection*{Specific results}
 
 \subsubsection*{Fit Plots}
-Both for the unweighted fit (see Material and Methods section for details) and for the X-weighted fit (detailed in Material and Methods subsection), two plots are generated: the former with logarithmic scale, the latter with lineal scale.
-Figure \ref{fig:unwFit} shows an example of the unweighted fit plots and Figure \ref{fig:X-wFit} does the same with the X-weighted fit plots.
-Additionally, for the unweighted fit, a complete residues analysis is performed, and a 4-in-1 figure is generated as shown in Figure \ref{fig:unwRes}, corresponding to the fit of Figure \ref{fig:unwFit}.
+For each and every dataset included in the study, an unweighted fit (see Material and Methods section for details) and a X-weighted fit (detailed in Material and Methods subsection) have been calculated for standard deviations versus the mean values for each bacterial genus monitored in time.
+Figure \ref{fig:main1} showed the X-weighted fit for samples from a healthy subject (top) and from a subject diagnosed with irritable bowel syndrome (bottom) studied in our lab \cite{IBS}, while Figure \ref{fig:fits} shows the corresponding unweighted fits.
+Additionally, for the unweighted fit, a complete residues analysis is performed, and a 4-in-1 figure is generated as shown in Figure \ref{fig:unwRes}, corresponding to patient A (top plot in Figure \ref{fig:main1} and Figure \ref{fig:fits}).
 Among other tests, it allows to check for normality and homoscedasticity of the residues.
@@
 	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_LOG.pdf}
-	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_LIN.pdf}
-	\caption{Example of unweighted fit, shown both in logarithmic and lineal scale plots}
-	\label{fig:unwFit}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_xWboot_LOG.pdf}
-	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_xWboot_LIN.pdf}
-	\caption{The X-weighted fit log and lineal plots corresponding to the fit of Figure \ref{fig:unwFit}}
-	\label{fig:X-wFit}
+	\includegraphics[width=0.8\textwidth]{results/fits/IBS_P1_metatranscriptomes_family_stdVSmean_LLR_LOG.pdf}
+	\caption{Log plots of unweighted fits corresponding to the datasets shown in Figure \ref{fig:main1}}
+	\label{fig:fits}
 \end{figure}
@@
 	\includegraphics[width=\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_RES.pdf}
-	\caption{Residues analysis plot corresponding to the fit of Figure \ref{fig:unwFit}.
+	\caption{Residues analysis plot corresponding to the unweighted fit for patient A of the IBS study.
 The top-left subplot is a simple residues plot.
@@
 
-\subsubsection*{Histogram Plots} 
+\subsubsection*{Histogram Plots (DRAFT)} 
 \CC\ generates three different histogram plots:
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{results/hist/antibiotics_h_patD_pre_AbsFreqPlot}
-	\caption{Histogram with the absolute frequencies of the pre-treatment data (7 first times) of patient ``D'' in the antibiotics study\cite{antibiotic}}
-	\label{fig:histAFP}
-\end{figure}
-
-\begin{figure}
-	\centering
 	\includegraphics[width=0.8\textwidth]{results/hist/IBS_h_A_amplicons_family_hist2D.pdf}
-	\caption{2D histogram deviation plot of the data used in the fit shown in Figure \ref{fig:unwFit}}
+	\caption{2D histogram deviation plot of the data for patient A of the IBS study}
 	\label{fig:hist2D}
@@
 	\includegraphics[width=0.8\textwidth]{results/hist/IBS_h_A_amplicons_family_ZRFhist.pdf}
-	\caption{Histogram with the relative frequency of zero for the elements (taxa) in the data used in the fit shown in Figure \ref{fig:unwFit}}
+	\caption{Histogram with the relative frequency of zero for the elements (taxa) in the data for patient A of the IBS study}
 	\label{fig:histZRF}
@@
 	\centering
-	\includegraphics[width=0.9\textwidth]{results/corrank/antibiotics_h_patD_pre_CORE_ElementsCorr}
-	\caption{Element correlation plot of the pre-treatment data (first seven times) of patient ``D'' in the antibiotics study\cite{antibiotic} for ``core'' taxa}
+	\includegraphics[width=0.9\textwidth]{results/corrank/IBS_h_A_amplicons_family_ElmCorrelation.pdf}
+	\caption{Element correlation plot of the for the most dominant taxa in the data for patient A of the IBS study}
 	\label{fig:corrElm}
@@
 	\centering
-	\includegraphics[width=0.9\textwidth]{results/corrank/antibiotics_h_patD_pre_CORE_TimesCorr}
-	\caption{Times correlation plot of the pre-treatment data (first seven times) of patient ``D'' in the antibiotics study\cite{antibiotic} for ``core'' taxa}
-	\label{fig:corrTim}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\textwidth]{results/corrank/IBS_h_A_amplicons_family_Rank}
-	\caption{Matrix showing the rank variation throughout time for the most dominant elements (taxa) and their calculated RSI (as discussed in Section \ref{sec:RSI}) in the data used in the fit shown in Figure \ref{fig:unwFit}}
+	\includegraphics[width=\textwidth]{results/corrank/IBS_h_A_amplicons_family_Rank.pdf}
+	\caption{Matrix showing the rank variation throughout time for the most dominant elements (taxa) and their calculated Rank Stability Index (as discussed in Material and Methods) in the data used in the fit shown in Figure \ref{fig:unwFit}}
 	\label{fig:corrank}

### 2016-04-27 mSystems/mSys_MAIN.tex
Some work in mSys MAIN (fix paper geometry) and Results (v0.0.8)
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_results.tex
	modified:   results/hist/antibiotics_h_patE_pre_AbsFreqPlot.pdf

@@
 % release codenumber (update it here!)
-\newcommand\releasecode{0.0.7}
+\newcommand\releasecode{0.0.8}
 
@@
 % text layout
-%\usepackage[letterpaper]{geometry}
-%\geometry{textwidth=16.25cm} % 15.25cm for single-space, 16.25cm for double-space
-%\geometry{textheight=22.5cm} % 22cm for single-space, 22.5cm for double-space
+\usepackage[letterpaper]{geometry}
+\geometry{textwidth=16.25cm} % 15.25cm for single-space, 16.25cm for double-space
+\geometry{textheight=22.5cm} % 22cm for single-space, 22.5cm for double-space
 
@@
 \title{
-	\vspace*{-16mm} % Helps to include the corresponding author footnote in the title page
+	\vspace*{0mm} % May help to include the corresponding author footnote in the title page
 	\singlespacing

### 2016-04-27 mSystems/mSys_results.tex
Some work in mSys MAIN (fix paper geometry) and Results (v0.0.8)
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_results.tex
	modified:   results/hist/antibiotics_h_patE_pre_AbsFreqPlot.pdf

--- 2016-04-22 mSystems/mSys_results.tex
@@
 Fitness is larger for the healthiest subjects and smaller for the IBS--diagnosed patients.
+
+%%%
+\subsection*{Specific results}
+
+\subsubsection*{Fit Plots}
+For each and every dataset included in the study, an unweighted fit (see Material and Methods section for details) and a X-weighted fit (detailed in Material and Methods subsection) have been calculated for standard deviations versus the mean values for each bacterial genus monitored in time.
+Figure \ref{fig:main1} showed the X-weighted fit for samples from a healthy subject (patient A, top) and from a subject diagnosed with irritable bowel syndrome (patient P2, bottom) studied in our lab\cite{IBS}, while Figure \ref{fig:fits} shows the corresponding unweighted fits.
+Additionally, for the unweighted fit, a complete residues analysis was performed, and a 4-in-1 figure was generated as shown in Figure \ref{fig:unwRes}, corresponding to patient A (top plot in Figure \ref{fig:main1} and Figure \ref{fig:fits}).
+Among other tests, it allows to check for normality and homoscedasticity of the residues and, therefore, to further evaluate the goodness of the fit.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_LOG.pdf}
+	\includegraphics[width=0.8\textwidth]{results/fits/IBS_P1_metatranscriptomes_family_stdVSmean_LLR_LOG.pdf}
+	\caption{Log plots of unweighted fits corresponding to the datasets shown in Figure \ref{fig:main1}}
+	\label{fig:fits}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_RES.pdf}
+	\caption{Residues analysis plot corresponding to the unweighted fit for patient A of the IBS study\cite{IBS}.
+The top-left subplot is a simple residues plot.
+The top-right subplot is a Normal quantiles plot with linear fitting (value of coefficient of determination is provided).
+The bottom-left subplot shows an accumulated ESS (Explained Sum of Squares) plot.
+Finally, the bottom-right subplot is a residues Normal histogram plot.
+This set of subplots allows to check for normality and homoscedasticity of the residues.}
+	\label{fig:unwRes}
+\end{figure}
+
+\subsubsection*{Histogram Plots} 
+\CC\ generates two related histogram plots, absolute frequencies plot and zero relative frequency plot.
+The former is useful to visually assess the validity of the time points in terms of the accumulated absolute frequency of the elements (taxa), since absolute frequencies far (much higher or much lower) from those typically observed could mean a sampling problem.
+In Figure \ref{fig:histAFP} shows this histogram for the pre-treatment data (first 7 times) of patient ``D'' in the antibiotics study\cite{antibiotic}.
+On the other hand, we could define the ZRF (Zero Relative Frequency, thereby ranging from 0 to 1) of an element (taxon) as the portion of times where it is zero, i.e., it is not found.
+Attending to all taxa, we can plot the ZRF histogram, which then lies on the horizontal axis of the plot.
+The vertical axis shows the number of taxa, so the height of a bar represents the amount of taxa that have determinate ZRF.
+In this respect, the bar over $0.0$ counts the quantity of taxa that are present at every time point of the data set (aka ``core''), while the bar over $1.0$ would count the total of taxa that are never found (this bar never appears because all these ``null'' elements are automatically filtered by the software.
+Figure \ref{fig:histZRF} shows this plot for the healthy patient A of the IBS study performed in our lab\cite{IBS}.
+There, we can see that $12$ taxa are present at all the time points of the time series while $9$ taxa basically appear only once.
+So, this plot is clearly useful to notice how the ``core'' is distributed.
+
+\begin{figure}		
+	\centering		
+ 	\includegraphics[width=0.8\textwidth]{results/hist/antibiotics_h_patD_pre_AbsFreqPlot}		
+ 	\caption{Histogram with the absolute frequencies of the pre-treatment data (7 first times) of patient ``D'' in the antibiotics study\cite{antibiotic}.
+We can see that there are no \emph{outlayers} among the total taxa sum for the time series points, so all of them were considered for further analysis by \CC\ software}
+ 	\label{fig:histAFP}		
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.8\textwidth]{results/hist/IBS_h_A_amplicons_family_ZRFhist.pdf}
+	\caption{Histogram with the relative frequency of zero for the elements (taxa) in the data for patient A of the IBS study\cite{IBS}.
+In this particular case, the bar over $0.0$ counts the quantity of taxa that are present at every time point of the data set (``core''), while the bar over $1.0$ sums the total of taxa that are never found (this bar never appears because all these ``null'' elements are automatically filtered by \CC): here, $12$ taxa are present at all the time points of the time series while $9$ taxa basically appear only once.}
+	\label{fig:histZRF}
+\end{figure}
+
+A 2D semi-logarithmic histogram representing deviations from the mean versus the mean itself is a useful tool in the analysis of the stability of ranking processes in complex systems\cite{ranking}.
+Figure \ref{fig:hist2D} shows this 2D deviation plot for the patient A of the IBS study\cite{IBS}.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.8\textwidth]{results/hist/IBS_h_A_amplicons_family_hist2D.pdf}
+	\caption{2D histogram deviation plot of the data for patient A of the IBS study\cite{IBS}}
+	\label{fig:hist2D}
+\end{figure}
+
+\subsubsection*{Correlation and Rank Plots} 
+\CC\ generates two different plots falling under this category, as well as Excel files with the resulting matrices.
+On the one hand, the elements correlation matrix plot shows a correlation matrix among the taxa, calculated with the time as independent variable.
+For these calculations, the data set is not normalized to avoid entering an additional constraint.
+Figure \ref{fig:corrElm} shows this matrix for the most dominant taxa present in the data of the patient ``A'' of the IBS study\cite{IBS}.
+On the other hand, the rank dynamics and stability plot shows the variation in the rank with time for the most dominant taxa and their calculated RSI, as discussed in Material and Methods.
+Figure \ref{fig:corrank} shows this plot for the taxa of the aforementioned patient A.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\textwidth]{results/corrank/IBS_h_A_amplicons_family_ElmCorrelation.pdf}
+	\caption{Element correlation plot of the for the most dominant taxa in the data for patient A of the IBS study\cite{IBS}}
+	\label{fig:corrElm}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{results/corrank/IBS_h_A_amplicons_family_Rank.pdf}
+	\caption{Matrix showing the rank variation throughout time for the most dominant elements (taxa) and their calculated Rank Stability Index (as discussed in Material and Methods) in the data  for patient A of the IBS study\cite{IBS}}
+	\label{fig:corrank}
+\end{figure}
+
+%%%
+\subsection*{Time dependence of model parameters}
 

### 2016-04-28 mSystems/mSys_results.tex
Some work in mSys Results section (v0.0.9)
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_results.tex
	modified:   results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_RES.pdf
	modified:   results/hist/IBS_h_A_amplicons_family_ZRFhist.pdf
	modified:   results/hist/IBS_h_A_amplicons_family_hist2D.pdf
	modified:   results/hist/antibiotics_h_patD_pre_AbsFreqPlot.pdf

@@
 \subsubsection*{Fit Plots}
-For each and every dataset included in the study, an unweighted fit (see Material and Methods section for details) and a X-weighted fit (detailed in Material and Methods subsection) have been calculated for standard deviations versus the mean values for each bacterial genus monitored in time.
+For each and every dataset included in the study, an unweighted fit and a X-weighted fit (see Material and Methods section for details on both fits) have been calculated for standard deviations versus the mean values for each bacterial genus monitored in time.
 Figure \ref{fig:main1} showed the X-weighted fit for samples from a healthy subject (patient A, top) and from a subject diagnosed with irritable bowel syndrome (patient P2, bottom) studied in our lab\cite{IBS}, while Figure \ref{fig:fits} shows the corresponding unweighted fits.
-Additionally, for the unweighted fit, a complete residues analysis was performed, and a 4-in-1 figure was generated as shown in Figure \ref{fig:unwRes}, corresponding to patient A (top plot in Figure \ref{fig:main1} and Figure \ref{fig:fits}).
-Among other tests, it allows to check for normality and homoscedasticity of the residues and, therefore, to further evaluate the goodness of the fit.
+Both for the X-weighted and the unweighted fit, the IBS patient shows significative higher values of $V$ and $\beta$ than the healthy subject.
 
@@
 	\includegraphics[width=0.8\textwidth]{results/fits/IBS_P1_metatranscriptomes_family_stdVSmean_LLR_LOG.pdf}
-	\caption{Log plots of unweighted fits corresponding to the datasets shown in Figure \ref{fig:main1}}
+	\caption{Log plots of unweighted fits corresponding to the datasets shown in Figure \ref{fig:main1}, a healthy subject (top) and from a subject diagnosed with irritable bowel syndrome (bottom) studied in our lab \cite{IBS}}
 	\label{fig:fits}
 \end{figure}
+
+Additionally, for the unweighted fit, a complete residues analysis was performed, and a 4-in-1 figure was generated as shown in Figure \ref{fig:unwRes}, corresponding to patient A (top plot in Figure \ref{fig:main1} and Figure \ref{fig:fits}).
+This allows to check for normality and homoscedasticity of the residues and, therefore, to further assess the goodness of the model's fit to the data.
 
@@
 Finally, the bottom-right subplot is a residues Normal histogram plot.
-This set of subplots allows to check for normality and homoscedasticity of the residues.}
+This set of subplots allows to check for normality and homoscedasticity of the residues and, therefore, to further assess the goodness of the model's fit to the real data.}
 	\label{fig:unwRes}
@@
 \subsubsection*{Histogram Plots} 
-\CC\ generates two related histogram plots, absolute frequencies plot and zero relative frequency plot.
+The \CC\ software generates two related histogram plots: absolute frequencies plot and zero relative frequency plot.
 The former is useful to visually assess the validity of the time points in terms of the accumulated absolute frequency of the elements (taxa), since absolute frequencies far (much higher or much lower) from those typically observed could mean a sampling problem.
 In Figure \ref{fig:histAFP} shows this histogram for the pre-treatment data (first 7 times) of patient ``D'' in the antibiotics study\cite{antibiotic}.
+We can see that there are no \emph{outlayers} among the total taxa sum for the sample points of this time series, so all of them were considered for further analysis by \CC.
+
 On the other hand, we could define the ZRF (Zero Relative Frequency, thereby ranging from 0 to 1) of an element (taxon) as the portion of times where it is zero, i.e., it is not found.
@@
  	\caption{Histogram with the absolute frequencies of the pre-treatment data (7 first times) of patient ``D'' in the antibiotics study\cite{antibiotic}.
-We can see that there are no \emph{outlayers} among the total taxa sum for the time series points, so all of them were considered for further analysis by \CC\ software}
+There are no \emph{outlayers} among the total taxa sum for the time series sample points, so all of them were considered for further analysis by \CC\ software}
  	\label{fig:histAFP}		

### 2016-04-28 mSystems/mSys_MAIN.tex
Some work in mSys Results section (v0.0.9)
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_results.tex
	modified:   results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_RES.pdf
	modified:   results/hist/IBS_h_A_amplicons_family_ZRFhist.pdf
	modified:   results/hist/IBS_h_A_amplicons_family_hist2D.pdf
	modified:   results/hist/antibiotics_h_patD_pre_AbsFreqPlot.pdf
intro citations updated

@@
 % release codenumber (update it here!)
-\newcommand\releasecode{0.0.8}
+\newcommand\releasecode{0.0.9}
 
@@
 Nature {\bf 510}:417–21.
+\bibitem{microb&health} {\bf Marchesi JR, Adams DH, Fava F, Hermes GD a, Hirschfield GM, Hold G, Quraishi MN, Kinross J, Smidt H, Tuohy KM, Thomas L V, Zoetendal EG, Hart A.} 2015.
+The gut microbiota and host health: a new clinical frontier.
+Gut 1–10.
+\bibitem{sysbio&microb} {\bf Wu H, Tremaroli V, Bäckhed F.} 2015. Linking Microbiota to Human Diseases: A Systems Biology Perspective. Trends Endocrinol Metab {\bf 26}:758–770.
+\bibitem{humangenome} {\bf Lander ES, Linton LM, Birren B, Nusbaum C, Zody MC, Baldwin J, Devon K, Dewar K, Doyle M, FitzHugh W, Funke R, Gage D, Harris K, Heaford A, Howland J, Kann L, Lehoczky J, LeVine R, McEwan P, McKernan K, Meldrim J, Mesirov JP, Miranda C, Morris W, Naylor J, Raymond C, Rosetti M, Santos R, Sheridan A, Sougnez C, Stange-Thomann N, Stojanovic N, Subramanian A, Wyman D, Rogers J, Sulston J, Ainscough R, Beck S, Bentley D, Burton J, Clee C, Carter N, Coulson A, Deadman R, Deloukas P, Dunham A, Dunham I, Durbin R, French L, Grafham D, Gregory S, Hubbard T, Humphray S, Hunt A, Jones M, Lloyd C, McMurray A, Matthews L, Mercer S, Milne S, Mullikin JC, Mungall A, Plumb R, Ross M, Shownkeen R, Sims S, Waterston RH, Wilson RK, Hillier LW, McPherson JD, Marra MA, Mardis ER, Fulton LA, Chinwalla AT, Pepin KH, Gish WR, Chissoe SL, Wendl MC, Delehaunty KD, Miner TL, Delehaunty A, Kramer JB, Cook LL, Fulton RS, Johnson DL, Minx PJ, Clifton SW, Hawkins T, Branscomb E, Predki P, Richardson P, Wenning S, Slezak T, Doggett N, Cheng JF, Olsen A, Lucas S, Elkin C, Uberbacher E, Frazier M, Gibbs RA, Muzny DM, Scherer SE, Bouck JB, Sodergren EJ, Worley KC, Rives CM, Gorrell JH, Metzker ML, Naylor SL, Kucherlapati RS, Nelson DL, Weinstock GM, Sakaki Y, Fujiyama A, Hattori M, Yada T, Toyoda A, Itoh T, Kawagoe C, Watanabe H, Totoki Y, Taylor T, Weissenbach J, Heilig R, Saurin W, Artiguenave F, Brottier P, Bruls T, Pelletier E, Robert C, Wincker P, Smith DR, Doucette-Stamm L, Rubenfield M, Weinstock K, Lee HM, Dubois J, Rosenthal A, Platzer M, Nyakatura G, Taudien S, Rump A, Yang H, Yu J, Wang J, Huang G, Gu J, Hood L, Rowen L, Madan A, Qin S, Davis RW, Federspiel NA, Abola AP, Proctor MJ, Myers RM, Schmutz J, Dickson M, Grimwood J, Cox DR, Olson M V, Kaul R, Raymond C, Shimizu N, Kawasaki K, Minoshima S, Evans GA, Athanasiou M, Schultz R, Roe BA, Chen F, Pan H, Ramser J, Lehrach H, Reinhardt R, McCombie WR, de la Bastide M, Dedhia N, Blocker H, Hornischer K, Nordsiek G, Agarwala R, Aravind L, Bailey JA, Bateman A, Batzoglou S, Birney E, Bork P, Brown DG, Burge CB, Cerutti L, Chen HC, Church D, Clamp M, Copley RR, Doerks T, Eddy SR, Eichler EE, Furey TS, Galagan J, Gilbert JG, Harmon C, Hayashizaki Y, Haussler D, Hermjakob H, Hokamp K, Jang W, Johnson LS, Jones TA, Kasif S, Kaspryzk A, Kennedy S, Kent WJ, Kitts P, Koonin E V, Korf I, Kulp D, Lancet D, Lowe TM, McLysaght A, Mikkelsen T, Moran J V, Mulder N, Pollara VJ, Ponting CP, Schuler G, Schultz J, Slater G, Smit AF, Stupka E, Szustakowski J, Thierry-Mieg D, Thierry-Mieg J, Wagner L, Wallis J, Wheeler R, Williams A, Wolf YI, Wolfe KH, Yang SP, Yeh RF, Collins F, Guyer MS, Peterson J, Felsenfeld A, Wetterstrand KA, Patrinos A, Morgan MJ, de Jong P, Catanese JJ, Osoegawa K, Shizuya H, Choi S, Chen YJ, International Human Genome Sequencing C.} 2001.
+Initial sequencing and analysis of the human genome.
+Nature {\bf 409}:860–921.
+
+\bibitem{taylor} {\bf Taylor, L.R.} 1961. Aggregation, Variance and the mean. Nature {\bf 189,} 732-35.
+\bibitem{randomwalks} {\bf de Menezes MA, Barabási A-L.} 2004. Fluctuations in network dynamics. Phys Rev Lett {\bf 92}:1–4.
+\bibitem{economics1} {\bf Mantegna RN, Stanley HE.} 1995. Scaling behaviour in the dynamics of an economic index. Nature {\bf 376}:46–49.
+\bibitem{economics2} {\bf Eisler Z, Kertesz J, Yook SH, Barabasi AL.} 2005. Multiscaling and non-universality in fluctuations of driven complex systems. Europhys Lett {\bf 69}:664–670.
+\bibitem{animal1} {\bf Reed DH, Hobbs GR.} 2004. The relationship between population size and temporal variability in population size. Anim Conserv {\bf 7}:1–8.
+\bibitem{animal2} {\bf Anderson RM, Gordon DM, Crawley MJ, Hassell MP.} 1982. Variability in the abundance of animal and plant species. Nature {\bf 18}: 245–248
+\bibitem{genexpress} {\bf Živković J, Tadić B, Wick N, Thurner S.} 2006. Statistical indicators of collective behavior and functional clusters in gene networks of yeast. Eur Phys J B {\bf 50}:255–258.
+\bibitem{genome} {\bf Kendal WS.} 2003. An Exponential Dispersion Model for the Distribution of Human Single Nucleotide Polymorphisms. Mol Biol Evol {\bf 20}:579–590.
+\bibitem{isme1} {\bf Zhang Z, Geng J, Tang X, Fan H, Xu J, Wen X, Ma ZS, Shi P.} 2014.
+Spatial heterogeneity and co-occurrence patterns of human mucosal-associated intestinal microbiota.
+ISME J {\bf 8}:881–93.
+\bibitem{ranking} {\bf Blumm N, Ghoshal G, Forró Z, Schich M, Bianconi G, Bouchaud J-P, Barabási A-L.} 2012. Dynamics of Ranking Processes in Complex Systems. Phys Rev Lett {\bf 109}:128701.
+
 \bibitem{moving} Caporaso, J.G. et al. Moving pictures of the human microbiome. {\it Genome Biol.} {\bf 12,} R50 (2011).
@@
 \bibitem{fs} Eisler,Z., Bartos,I., Kertesz,J. Fluctuation scaling in complex systems: Taylor's law and beyond. {\it Adv. Phys.} {\bf 57,} 85 (2008).
-\bibitem{taylor} Taylor, L.R. Aggregation, Variance and the mean. {\it Nature} {\bf 189,} 732-35 (1961).
 \bibitem{stat} Jorgensen,B., Martinez,J.R., Tsao,M. Asymptotic behaviour of the variance function. {\it Scand. J. Statist..} {\bf 21,} 223-243 (1994).
@@
 \bibitem{convergence3} Kendal, W.S., Jorgensen,B. Tweedie convergence: A mathematical basis for Taylor's power law. {\it Phys. Rev. E} {\bf 84,} 066120 (2011).
-\bibitem{ranking} Blumm, N. et al. Dynamics of ranking processes in complex systems. {\it Phys. Rev. Lett.} {\bf 109,} 128701 (2012).
+
 	\bibitem{FD} Weber, J. \textit{et al.} Fluctuation dissipation theorem. {\it Phys. Rev.} {\bf 101}, 1620-6 (1956).

### 2016-04-28 mSystems/mSys_intro.tex
intro citations updated

@@
 High throughput methods for microbial 16S ribosomal RNA gene and WGS have now begun to reveal the composition of archaeal, bacterial, fungal and viral communities located both, in and on the human body.
-Modern high-throughput sequencing and bioinformatics tools provide a powerful means of understanding how the human microbiome contributes to health and its potential as a target for therapeutic interventions [ref?].
+Modern high-throughput sequencing and bioinformatics tools provide a powerful means of understanding how the human microbiome contributes to health and its potential as a target for therapeutic interventions \cite{microb&health, sysbio&microb}.
 
 Biology has recently acquired new technological and conceptual tools to investigate, model and understand living organisms at the system level, thanks to the spectacular progress in quantitative techniques, large-scale measurement methods and the integration of experimental and computational approaches.
-Systems Biology has mostly been devoted to the study of well-characterized model organisms but, since the early days of the Human Genome Project [ref] it has become clear that applications of system-wide approaches to Human Biology would bring huge opportunities in Medicine.
+Systems Biology has mostly been devoted to the study of well-characterized model organisms but, since the early days of the Human Genome Project \cite{humangenome} it has become clear that applications of system-wide approaches to Human Biology would bring huge opportunities in Medicine.
 Great effort has been placed to unveil the general laws governing the behaviour of this complex system [ref].
-Due to his nature, microbiota can be studied under the light of the ecology, where we can find general principles as the Taylor?s law, which relates spatial or temporal variability of the population with its mean.
-This law, also known as fluctuation scale law, is ubiquitous in the natural world and can be found in several systems as cosmic rays [ref1], stock markets [ref2,3], animal populations [ref4, 5, 6], gene expression [ref7], or in the human genome [ref8].
-Taylor?s law has been applied to microbiota in a spatial way in the work of Zhang et al., (2014), where they show that this population tend to be in an aggregated way rather than in a random distribution.
+Due to his nature, microbiota can be studied under the light of the ecology, where we can find general principles as the Taylor's law \cite{taylor}, which relates spatial or temporal variability of the population with its mean.
+This law, also known as fluctuation scale law, is ubiquitous in the natural world and can be found in several systems as random walks \cite{randomwalks}, stock markets \cite{economics1, economics2}, animal populations \cite{taylor, animal1, animal2}, gene expression \cite{genexpress}, or in the human genome \cite{genome}.
+Taylor's law has been applied to microbiota in a spatial way in the work of Zhang {\it et al.}, (2014) \cite{isme1}, where they show that this population tend to be in an aggregated way rather than in a random distribution.
 
@@
 We have analyzed more than 35000 time series of taxa from the gut microbiome of 97 individuals obtained from publicly available high throughput sequencing data on different conditions: diseases, diets, obese status, antibiotic perturbation and healthy individuals.
-Having seen that all cases follows Taylor?s law, we use this empirical fact to model how the relative abundances of taxa evolves toward time thanks to the Langevin equation, in a similar way as Blumm et al., did in their (2012).
+Having seen that all cases follows Taylor's law, we use this empirical fact to model how the relative abundances of taxa evolves toward time thanks to the Langevin equation, in a similar way as Blumm et al., did in their (2012) \cite{ranking}.
 We use this mathematical framework to explore the temporal stability of the microbiota in different conditions in order to understand how this affects the healthy status of the subjects.

### 2016-04-28 mSystems/mSys_results.tex
Work in mSys Results section (v0.0.10)
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_results.tex
	modified:   results/fits/IBS_P1_metatranscriptomes_family_stdVSmean_LLR_LOG.pdf
	modified:   results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_LOG.pdf

@@
 
-On the other hand, we could define the ZRF (Zero Relative Frequency, thereby ranging from 0 to 1) of an element (taxon) as the portion of times where it is zero, i.e., it is not found.
-Attending to all taxa, we can plot the ZRF histogram, which then lies on the horizontal axis of the plot.
-The vertical axis shows the number of taxa, so the height of a bar represents the amount of taxa that have determinate ZRF.
-In this respect, the bar over $0.0$ counts the quantity of taxa that are present at every time point of the data set (aka ``core''), while the bar over $1.0$ would count the total of taxa that are never found (this bar never appears because all these ``null'' elements are automatically filtered by the software.
+On the other hand, we could define the ZRF (Zero Relative Frequency, thereby ranging from 0 to 1) of an element (taxon) as the portion of times of the time series where the corresponding frequency is zero, i.e., such taxon is not found.
+Attending to all taxa, we can plot the ZRF histogram, which then lies on the horizontal axis of the plot, with the vertical axis showing the number of taxa.
+So, the height of a bar represents the amount of taxa that have determinate ZRF.
+In this respect, the bar over $0.0$ counts the quantity of taxa that are present at every time point of the data set (aka ``core'').
+In the same way, in the opposite case, the bar over $1.0$ would count the total of taxa that are never found, though this bar never appears in our plots since all these ``null'' elements are automatically filtered by the software.
 Figure \ref{fig:histZRF} shows this plot for the healthy patient A of the IBS study performed in our lab\cite{IBS}.
 There, we can see that $12$ taxa are present at all the time points of the time series while $9$ taxa basically appear only once.
-So, this plot is clearly useful to notice how the ``core'' is distributed.
+All in all, this plot is clearly useful to notice how the ``core'' is distributed.
 
@@
 \subsubsection*{Correlation and Rank Plots} 
-\CC\ generates two different plots falling under this category, as well as Excel files with the resulting matrices.
+\CC\ generates two different plots falling under this category in addition to the Excel files with the resulting matrices.
 On the one hand, the elements correlation matrix plot shows a correlation matrix among the taxa, calculated with the time as independent variable.

### 2016-05-01 mSystems/mSys_results.tex
Improvements to mSys ZRF results from cmplxCruncher 1.1rc2, which leads to 0.1.0
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_results.tex
	modified:   results/hist/IBS_h_A_amplicons_family_ZRFhist.pdf

@@
 On the other hand, we could define the ZRF (Zero Relative Frequency, thereby ranging from 0 to 1) of an element (taxon) as the portion of times of the time series where the corresponding frequency is zero, i.e., such taxon is not found.
-Attending to all taxa, we can plot the ZRF histogram, which then lies on the horizontal axis of the plot, with the vertical axis showing the number of taxa.
+The ZRF is useful for assessing how regularly is appearing the microbiota in the samples.
+In that sense, we define the ``core'' as the taxa that has a ZRF of less than 0.2, i.e., is present in more than the 80\% of the samples.
+
+Attending to all taxa in a dataset, we can plot the ZRF histogram, which then lies on the horizontal axis of the plot, with the vertical axis showing the number of taxa.
 So, the height of a bar represents the amount of taxa that have determinate ZRF.
-In this respect, the bar over $0.0$ counts the quantity of taxa that are present at every time point of the data set (aka ``core'').
-In the same way, in the opposite case, the bar over $1.0$ would count the total of taxa that are never found, though this bar never appears in our plots since all these ``null'' elements are automatically filtered by the software.
+In this respect, the bar near $0$ counts the quantity of taxa that are present at every (or almost every) time point of the data set.
+In the same way, in the opposite case, the bar near $1$ counts the total of taxa that are scarcely found but present at least in one time (the software filters any taxon not appearing at any time).
 Figure \ref{fig:histZRF} shows this plot for the healthy patient A of the IBS study performed in our lab\cite{IBS}.
 There, we can see that $12$ taxa are present at all the time points of the time series while $9$ taxa basically appear only once.
-All in all, this plot is clearly useful to notice how the ``core'' is distributed.
+Ranging between those extremes, there are a significative amount of elements, $23$ taxa, appearing from often to seldom, but not belonging to the ``core'' region.
 
@@
 	\caption{Histogram with the relative frequency of zero for the elements (taxa) in the data for patient A of the IBS study\cite{IBS}.
-In this particular case, the bar over $0.0$ counts the quantity of taxa that are present at every time point of the data set (``core''), while the bar over $1.0$ sums the total of taxa that are never found (this bar never appears because all these ``null'' elements are automatically filtered by \CC): here, $12$ taxa are present at all the time points of the time series while $9$ taxa basically appear only once.}
+In this particular case, the bar near $0$ counts the quantity of taxa that are present at almost every time point of the data set, while the bar near $1$ sums the total of taxa that are scarcely found: here, $12$ taxa are present at all the time points of the time series while $9$ taxa basically appear only once.}
 	\label{fig:histZRF}

### 2016-05-13 mSystems/mSys_MAIN.tex
Work in mSys Results section (v0.0.10)
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_results.tex
	modified:   results/fits/IBS_P1_metatranscriptomes_family_stdVSmean_LLR_LOG.pdf
	modified:   results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_LOG.pdf
Improvements to mSys ZRF results from cmplxCruncher 1.1rc2, which leads to 0.1.0
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_results.tex
	modified:   results/hist/IBS_h_A_amplicons_family_ZRFhist.pdf
abstract modified; intro greatly modified, with several new c

--- 2016-04-27 mSystems/mSys_MAIN.tex
@@
 % release codenumber (update it here!)
-\newcommand\releasecode{0.0.8}
+\newcommand\releasecode{0.1.0}
 
@@
 %TC:macro \abstract [abswords]
-{\abstract{Human microbiota plays an important role in determining changes from health to disease. Increasing research activity is dedicated to understand its diversity and variability. 
-We analyse 16S rRNA and whole genome sequencing (WGS) data from the gut microbiota of 97 individuals monitored in time. Temporal fluctuations in the microbiome reveal 
-significant differences due to factors that affect the microbiota such as dietary changes, antibiotic intake, early gut development or disease. Here we show that a fluctuation 
-scaling law describes the temporal variability of the system and that a noise-induced phase transition is central in the route to disease. The universal law distinguishes healthy 
-from sick microbiota and quantitatively characterizes the path in the phase space, which opens up its potential clinical use and, more generally, other technological 
-applications where microbiota plays an important role.}}
+{\abstract{Animal microbiota (human included) plays an important role keeping healthy the physiological status of the host.
+Increasing research activity is dedicated to understand how changes in composition and function of the microbiota are associated to disease or not.
+We analyze 16S rRNA and whole genome sequencing (WGS) published data from the gut microbiota of 97 individuals monitored in time.
+Temporal fluctuations in the microbial composition reveal significant differences due to factors such us dietary changes, antibiotic intake, age or disease.
+Here we show that a fluctuation scaling law describes the temporal changes in the gut microbiota.
+This law allows to estimate the temporal variability of the microbial population and quantitatively characterizes the path toward disease by a noise-induced phase transition.
+The estimation of the systemic parameters for follow-up studies may have clinical use and, more generally, applications in other fields where it is important to know if a given community is stable or not.}}
 
@@
 \begin{thebibliography}{99}
+\bibitem{bileacids} {\bf Swann JR, Want EJ, Geier FM, Spagou K, Wilson ID, Sidaway JE, Nicholson JK, Holmes E.} 2011.
+Systemic gut microbial modulation of bile acid metabolism in host tissue compartments.
+Proc Natl Acad Sci {\bf 108}:4523–4530.
+\bibitem{choline} {\bf nSpencer MD, Hamp TJ, Reid RW, Fischer LM, Zeisel SH, Fodor AA.} 2011.
+Association between composition of the human gastrointestinal microbiome and development of fatty liver with choline deficiency.
+Gastroenterology {\bf 140}:976–986.
+\bibitem{scfa1} {\bf Samuel BS, Shaito A, Motoike T, Rey FE, Backhed F, Manchester JK, Hammer RE, Williams SC, Crowley J, Yanagisawa M, Gordon JI}.
+2008.
+Effects of the gut microbiota on host adiposity are modulated by the short-chain fatty-acid binding G protein-coupled receptor, Gpr41.
+Proc Natl Acad Sci {\bf 105}:16767–16772.
+\bibitem{scfa2} {\bf Smith PM, Howitt MR, Panikov N, Michaud M, Gallini CA, Bohlooly-Y M, Glickman JN, Garrett WS.} 2013.
+The Microbial Metabolites, Short-Chain Fatty Acids, Regulate Colonic Treg Cell Homeostasis.
+Science (80- ) {\bf 341}:569–573.
+\bibitem {scfa3} {\bf Kimura I, Ozawa K, Inoue D, Imamura T, Kimura K, Maeda T, Terasawa K, Kashihara D, Hirano K, Tani T, Takahashi T, Miyauchi S, Shioi G, Inoue H, Tsujimoto G.} 2013.
+The gut microbiota suppresses insulin-mediated fat accumulation via the short-chain fatty acid receptor GPR43.
+Nat Commun {\bf 4}:1829.
+\bibitem {scfa4} {\bf Maslowski KM, Vieira AT, Ng A, Kranich J, Sierro F, Di Yu, Schilter HC, Rolph MS, Mackay F, Artis D, Xavier RJ, Teixeira MM, Mackay CR.} 2009.
+Regulation of inflammatory responses by gut microbiota and chemoattractant receptor GPR43.
+Nature {\bf 461}:1282–1286.
 \bibitem{diabetes2} {\bf Qin J, Li Y, Cai Z, Li S, Zhu J, Zhang F, Liang S, Zhang W, Guan Y, Shen D, Peng Y, Zhang D, Jie Z, Wu W, Qin Y, Xue W, Li J, Han L, Lu D, Wu P, Dai Y, Sun X, Li Z, Tang A, Zhong S, Li X, Chen W, Xu R, Wang M, Feng Q, Gong M, Yu J, Zhang Y, Zhang M, Hansen T, Sanchez G, Raes J, Falony G, Okuda S, Almeida M, LeChatelier E, Renault P, Pons N, Batto J-M, Zhang Z, Chen H, Yang R, Zheng W, Li S, Yang H, Wang J, Ehrlich SD, Nielsen R, Pedersen O, Kristiansen K, Wang J.} 2012.
@@
 Nature {\bf 510}:417–21.
+\bibitem{Moya_trends} {\bf Moya A, Ferrer M.} 2016. Functional Redundancy-Induced Stability of Gut Microbiota Subjected to Disturbance. Trends Microbiol {\bf 24}:402–413.
+\bibitem{microb&health} {\bf Marchesi JR, Adams DH, Fava F, Hermes GD a, Hirschfield GM, Hold G, Quraishi MN, Kinross J, Smidt H, Tuohy KM, Thomas L V, Zoetendal EG, Hart A.} 2015.
+The gut microbiota and host health: a new clinical frontier.
+Gut 1–10.
+\bibitem{normal1} {\bf Falony G, Joossens M, Vieira-Silva S, Wang J, Darzi Y, Faust K, Kurilshikov A, Bonder MJ, Valles-Colomer M, Vandeputte D, Tito RY, Chaffron S, Rymenans L, Verspecht C, De Sutter L, Lima-Mendez G, Dhoe K, Jonckheere K, Homola D, Garcia R, Tigchelaar EF, Eeckhaudt L, Fu J, Henckaerts L, Zhernakova A, Wijmenga C, Raes J.} 2016.
+Population-level analysis of gut microbiome variation.
+Science (80- ) {\bf 352}:560–564.
+\bibitem{normal2} {\bf Zhernakova A, Kurilshikov A, Bonder MJ, Tigchelaar EF, Schirmer M, Vatanen T, Mujagic Z, Vila AV, Falony G, Vieira-Silva S, Wang J, Imhann F, Brandsma E, Jankipersadsing SA, Joossens M, Cenit MC, Deelen P, Swertz MA, Weersma RK, Feskens EJM, Netea MG, Gevers D, Jonkers D, Franke L, Aulchenko YS, Huttenhower C, Raes J, Hofker MH, Xavier RJ, Wijmenga C, Fu J.} 2016.
+Population-based metagenomics analysis reveals markers for gut microbiome composition and diversity.
+Science (80- ) {\bf 352}:565–569.
+\bibitem{sysbio&microb} {\bf Wu H, Tremaroli V, Bäckhed F.} 2015. Linking Microbiota to Human Diseases: A Systems Biology Perspective. Trends Endocrinol Metab {\bf 26}:758–770.
+\bibitem{msys1} {\bf Noecker C, Eng A, Srinivasan S, Theriot CM, Young VB, Jansson JK, Fredricks DN, Borenstein E.} 2016.
+Metabolic Model-Based Integration of Microbiome Taxonomic and Metabolomic Profiles Elucidates Mechanistic Links between Ecological and Metabolic Variation.
+mSystems {\bf 1}:e00013–15.
+\bibitem {metasysbio} {\bf Greenblum S, Turnbaugh PJ, Borenstein E.} 2012.
+Metagenomic systems biology of the human gut microbiome reveals topological shifts associated with obesity and inflammatory bowel disease.
+Proc Natl Acad Sci {\bf 109}:594–599.
+\bibitem{taylor} {\bf Taylor, L.R.} 1961. Aggregation, Variance and the mean. Nature {\bf 189,} 732-35.
+\bibitem{randomwalks} {\bf de Menezes MA, Barabási A-L.} 2004. Fluctuations in network dynamics. Phys Rev Lett {\bf 92}:1–4.
+\bibitem{economics1} {\bf Mantegna RN, Stanley HE.} 1995. Scaling behaviour in the dynamics of an economic index. Nature {\bf 376}:46–49.
+\bibitem{economics2} {\bf Eisler Z, Kertesz J, Yook SH, Barabasi AL.} 2005. Multiscaling and non-universality in fluctuations of driven complex systems. Europhys Lett {\bf 69}:664–670.
+\bibitem{animal1} {\bf Reed DH, Hobbs GR.} 2004. The relationship between population size and temporal variability in population size. Anim Conserv {\bf 7}:1–8.
+\bibitem{animal2} {\bf Anderson RM, Gordon DM, Crawley MJ, Hassell MP.} 1982. Variability in the abundance of animal and plant species. Nature {\bf 18}: 245–248
+\bibitem{genexpress} {\bf Živković J, Tadić B, Wick N, Thurner S.} 2006. Statistical indicators of collective behavior and functional clusters in gene networks of yeast. Eur Phys J B {\bf 50}:255–258.
+\bibitem{genome} {\bf Kendal WS.} 2003. An Exponential Dispersion Model for the Distribution of Human Single Nucleotide Polymorphisms. Mol Biol Evol {\bf 20}:579–590.
+\bibitem{isme1} {\bf Zhang Z, Geng J, Tang X, Fan H, Xu J, Wen X, Ma ZS, Shi P.} 2014.
+Spatial heterogeneity and co-occurrence patterns of human mucosal-associated intestinal microbiota.
+ISME J {\bf 8}:881–93.
+\bibitem{cobas} {\bf Pérez-Cobas AE, Artacho A, Ott SJ, Moya A, Gosalbes MJ, Latorre A.} 2014.
+Structural and functional changes in the gut microbiota associated to Clostridium difficile infection.
+Front Microbiol {\bf 5}:1–15.
+\bibitem{schloss} {\bf Ding T, Schloss PD.} 2014. Dynamics and associations of microbial community types across the human body. Nature {\bf 509}:357–360.
+\bibitem{ravel} {\bf Gajer P, Brotman RM, Bai G, Sakamoto J, Schütte UME, Zhong X, Koenig SSK, Fu L, Ma ZS, Zhou X, Abdo Z, Forney LJ, Ravel J.} 2012.
+Temporal dynamics of the human vaginal microbiota.
+Sci Transl Med {\bf 4}:132ra52.
+\bibitem{ranking} {\bf Blumm N, Ghoshal G, Forró Z, Schich M, Bianconi G, Bouchaud J-P, Barabási A-L.} 2012. Dynamics of Ranking Processes in Complex Systems. Phys Rev Lett {\bf 109}:128701.
 \bibitem{moving} Caporaso, J.G. et al. Moving pictures of the human microbiome. {\it Genome Biol.} {\bf 12,} R50 (2011).

### 2016-05-13 mSystems/mSys_intro.tex
abstract modified; intro greatly modified, with several new citations; discussion modified with an idea for last paragraph

@@
 The desire to understand the factors that influence human health and cause diseases has always been one of the major driving forces of biological research.
-We are populated by a myriad of microorganisms that are interacting with us in several physiological processes such as metabolism regulation or maturation of the immune system.
-Human microbiota has been suggested to be closely related to diseases like type 2 diabetes
-\cite{diabetes2}, cardiovascular disease (CVD) \cite{CVD}, irritable bowel syndrome \cite{IBS}, Crohn's disease \cite{CD} or some affections as obesity \cite{ob1, ob2} or malnutrition \cite{nutr}.
+We are populated by a myriad of microorganisms that are interacting with us in several physiological processes such as metabolism of the bile acids \cite{bileacids}, of the choline \cite{choline} or key-route metabolites as short-chain fatty acids \cite{scfa1, scfa2} which is also involved in immune system maturation \cite{scfa3, scfa4}.
+Human microbiota has been suggested to be closely related to diseases like type 2 diabetes \cite{diabetes2}, cardiovascular disease (CVD) \cite{CVD}, irritable bowel syndrome \cite{IBS}, Crohn's disease \cite{CD}, some affections as obesity \cite{ob1, ob2}, malnutrition \cite{nutr} among other multiple diseases \cite{Moya_trends}.
 High throughput methods for microbial 16S ribosomal RNA gene and WGS have now begun to reveal the composition of archaeal, bacterial, fungal and viral communities located both, in and on the human body.
-Modern high-throughput sequencing and bioinformatics tools provide a powerful means of understanding how the human microbiome contributes to health and its potential as a target for therapeutic interventions \cite{microb&health, sysbio&microb}.
+Modern high-throughput sequencing and bioinformatics tools provide a powerful means of understanding how the human microbiome contributes to health and its potential as a target for therapeutic interventions \cite{microb&health}.
+To define normal microbiota and how it's compositional changes can origin some diseases are important issues still in need for scientific answers \cite{normal1, normal2}.
 
 Biology has recently acquired new technological and conceptual tools to investigate, model and understand living organisms at the system level, thanks to the spectacular progress in quantitative techniques, large-scale measurement methods and the integration of experimental and computational approaches.
-Systems Biology has mostly been devoted to the study of well-characterized model organisms but, since the early days of the Human Genome Project \cite{humangenome} it has become clear that applications of system-wide approaches to Human Biology would bring huge opportunities in Medicine.
-Great effort has been placed to unveil the general laws governing the behaviour of this complex system [ref].
-Due to his nature, microbiota can be studied under the light of the ecology, where we can find general principles as the Taylor's law \cite{taylor}, which relates spatial or temporal variability of the population with its mean.
+In particular, Systems Biology has placed a great effort to unveil the general laws governing the complex behaviour of microbial communities \cite{sysbio&microb, msys1, metasysbio}.
+Microbiota can be approached under the light of ecological theory where we can find, for instance, general principles as the Taylor's law \cite{taylor}, which relates spatial or temporal variability of the population with its mean.
 This law, also known as fluctuation scale law, is ubiquitous in the natural world and can be found in several systems as random walks \cite{randomwalks}, stock markets \cite{economics1, economics2}, animal populations \cite{taylor, animal1, animal2}, gene expression \cite{genexpress}, or in the human genome \cite{genome}.
 Taylor's law has been applied to microbiota in a spatial way in the work of Zhang {\it et al.}, (2014) \cite{isme1}, where they show that this population tend to be in an aggregated way rather than in a random distribution.
+Despite its ubiquity, it has never been applied in follow-up studies from microbiota even that a great effort has been made to infer the community structure from a dynamical point of view \cite{cobas, schloss, ravel}
 
-Here we present the imprints of disease in macroscopic properties of the system, by studying the temporal variability in the microbiome.
-We have analyzed more than 35000 time series of taxa from the gut microbiome of 97 individuals obtained from publicly available high throughput sequencing data on different conditions: diseases, diets, obese status, antibiotic perturbation and healthy individuals.
-Having seen that all cases follows Taylor's law, we use this empirical fact to model how the relative abundances of taxa evolves toward time thanks to the Langevin equation, in a similar way as Blumm et al., did in their (2012) \cite{ranking}.
+Here we present the imprints of health status (healthy or disease) in macroscopic properties of microbiota, by studying the temporal variability.
+We have analyzed more than 35000 time series of taxa from the gut microbiome of 97 individuals obtained from publicly available high throughput sequencing data on different conditions: diseases, diets, obese status, antibiotic therapy and healthy individuals.
+Having seen that all cases follows Taylor's law, we use this empirical fact to model how the relative abundances of taxa evolves toward time thanks to the Langevin equation, in a similar way as it was applied recently by Blumm {\it et al.} \cite{ranking}.
 We use this mathematical framework to explore the temporal stability of the microbiota in different conditions in order to understand how this affects the healthy status of the subjects.
-Finally, we have engineered a complete software framework, ComplexCruncher, to support the analysis of the dynamics of ranking processes in complex systems, which is ready to be implemented by other users.
+Finally, we have engineered a complete software framework, {\it ComplexCruncher}, to carry out the analysis of the temporal dynamics of microbiotas, which is ready to be implemented by other users.

### 2016-05-18 mSystems/2CCpaper.tex
Detach CC from mSys and improve rank stability results, which leads to 0.2.0
On branch master
	copied:     mSys_MandM.tex -> 2CCpaper.tex
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_discussion.tex
	modified:   mSys_results.tex
	renamed:    results/corrank/IBS_P2 Metatranscriptores_Rank.pdf -> results/corrank/IBS_P2_Metatranscriptores_Rank.pdf

--- 2016-04-25 mSystems/mSys_MandM.tex
@@
-%%%
-\subsection*{Model} \label{sec:model}
+% !TEX TS-program = pdflatex
+% !TEX encoding = UTF-8 Unicode
+%
+% Aux .tex file to store some of the content of the mSystems paper to be moved
+% to the complexCrucher specific paper that should be prepared once decided to
+% detach CC in a different specific paper (probably targeting a bioinformatics
+% journal) 
+%
 
-We model the microbial abundances across time along the lines of Blumm \textit{et al.}\cite{ranking}. The dynamics of taxon relative abundances is described by the Langevin equation:
-\begin{linenomath}
-\begin{equation}
-\dot{x_i} = F_i \cdot x_i^\alpha + V \cdot x_i^\beta \xi_i(t) - \phi(t) \cdot x_i,
-\end{equation}
-\end{linenomath}
-where F$_i$ captures the fitness of the taxon i, V corresponds to the noise amplitude and $\xi_i$(t) is a Gaussian random noise with zero mean  $<\xi_i(t)>$ = 0 and variance uncorrelated in time, $<\xi_i(t) \xi_i(t')>$ =  $\delta(t'-t)$, .
-The function $\phi(t)$ ensures the normalization at all times, $\sum x_i(t) = 1$, and corresponds to $\phi(t) = \sum F_i x_i^\alpha + \sum V x_i^\beta \xi_i(t)$.
-The temporal evolution of the probability that a taxon i has a relative abundance $x_i(t)$, P(x$_i$,t), is determined by the Fokker-Planck equation:
-\begin{linenomath}
-\begin{equation}
-\frac{\partial P}{\partial t} = - \frac{\partial}{\partial x_i}  [(F_i \cdot x_i^\alpha - \phi(t) \cdot x_i ) \cdot P]+ \frac{1}{2} \frac{\partial^2}{\partial x_i^2} (V^2 \cdot x_i^{2\beta}\cdot P).
-\end{equation}
-\end{linenomath}
-The microbiota evolves towards a steady-state with a time-independent probability depending on the values of $\alpha$, $\beta$, F$_i$ and V.
-For $\alpha<1$ (otherwise, systems are always unstable), the steady-state probability may be localized in a region around a preferred value or broadly distributed over a wide range, depending on whether the fitness F$_i$  dominates or is overwhelmed by the noise amplitude V.
-The steady-state solution of the Fokker-Planck equation is given by:
-\begin{eqnarray}
-P_0 (x_i) &=& C_{ne}(\alpha,\beta,F_i,V)  \cdot x_i^{-2\beta}  \cdot \exp\Big[\frac{2F_i}{V^2}\frac{x_i^{1+\alpha-2\beta}}{1+\alpha-2\beta}-\frac{\phi_0}{V^2}\frac{x_i^{2-2\beta}}{1-\beta}\Big] \quad \textrm{if} \quad  2\beta \ne 1+\alpha, \notag \\
-P_0 (x_i) &=& C_e(\alpha,\beta,F_i,V)  \cdot x_i^{\frac{2F_i}{V^2} -2\beta}  \cdot \exp\Big[\frac{\phi_0}{V^2}\frac{x_i^{2-2\beta}}{1-\beta}\Big] \quad \textrm{if} \quad  2\beta = 1+\alpha,
-\notag
-\end{eqnarray}
-where $\phi_0 = (\sum_i F_i^{1/(1-\alpha)})^{1-\alpha}$ and C$_{ne}$ and C$_{e}$ are integrals that should be solved numerically for the parameters of interest.
-The ordered phase happens when the solution has a maximum in the physical interval ($0<x_i<1$).
-For larger V, the transition to a disordered phase happens when the maximum shifts to the unphysical region $x_i<0$, which sets the phase transition region V($\alpha,\beta,F_i$).
-The phase transition region can be calculated analytically in particular cases:
-\begin{eqnarray}
-F_i^2 &=& 4 \beta \phi_0 V^2\quad \textrm{if} \quad  \beta = \alpha \neq 1, \notag \\
-F_i &=& \beta V^2\quad \textrm{if} \quad  2\beta = 1+\alpha,
-\notag
-\end{eqnarray}
-where the first case, simplifies to $F= 3 V^2$ if $\beta = 0.75$ and the fitness of this taxon dominates in $\phi_0$. 
-In many physical systems (Brownian motion is the classical example), the two terms of the Langevin equation are related.
-The \emph{fluctuation--dissipation theorem} states a general relationship between the response to an external disturbance and the internal fluctuations of the system\cite{FD}.
-The theorem can be used as the basic formula to derive the fitness from the analysis of fluctuations
-of the microbiota, assuming that it is in equilibrium (the ordered phase).
+\subsection*{Specific results}
 
-\todo{Explain better the fluctuation-dissipation theorem}
-
-%%%
-\subsection*{Selection and Methods}
-
-The bacteria and archaea taxonomic assignations were obtained by analysing 16S rRNA sequences, which were clustered into operational taxonomic units (OTUs) sharing 97 \% sequence identity using QIIME\cite{QIIME}.
-WGS data\cite{kwashiorkor} were analysed and assigned at strain level by the Livermore Metagenomic Analysis Toolkit (LMAT)\cite{LMAT}, according to their default quality threshold.
-Genus, with best balance between error assignment and number of taxa, was chosen as our reference taxonomic level.
-We have verified that our conclusions are not significantly affected by selecting family or species as the reference taxonomic level (see Figure \ref{fig:taxlev1}).
-
-\todo{Specify, in each study treated, the nature of the samples (conditions, timespan between timepoints, subjects).
-Specify, and it is very important, what we consider \emph{healthy} in each study (for example: pre-antibiotics is healthy)}
-
-\subsubsection*{Sample selection}
-We have chosen studies about relevant pathologies containing metagenomic sequencing time data series of bacterial populations from humans in different healthy and non-healthy states.
-We have selected only those individuals who had three or more time points of data available in databases.
-Metadata of each study is provided in Tables \ref{tab:diet} to \ref{tab:LEA}.
-All used 16S rRNA gene sequencing except for the study of the discordant kwashiorkor twins\cite{kwashiorkor} (see Tables \ref{tab:DH} and \ref{tab:DK}) where shotgun metagenomic sequencing (SMS) and 16S rRNA were used.
-In the latter case we selected to work with SMS data to show that our method is valid regardless of the source of taxonomic information.
-Each one of the datasets was treated as follows:
-
-\subsubsection*{16rRNA sequences processing}
-Reads from the selected studies were first quality filtered using the FastX toolkit\cite{FASTX}, allowing only those reads which had more than 25 of quality along the 75\% of the complete sequence.
-16S rRNA reads were then clustered at 97\% nucleotide sequence identity (97\% ID) into operational taxonomic units (OTUs) using QIIME package software\cite{QIIME} (version 1.8) We followed open reference OTU picking workflow in all cases.
-The clustering method used was uclust, and the OTUs were matched against Silva database\cite{SILVA} (version 111, July 2012) and were assigned to taxonomy with an uclust-based consensus taxonomy assigner.
-The parameters used in this step were: similarity 0.97, prefilter percent id 0.6, max accepts 20, max rejects 500.
-
-\subsubsection*{Metagenomic sequences processing}
-Metagenomic shotgun (and 16S too) sequences were analyzed with LMAT (Livermore Metagenomics Analysis Toolkit) software package\cite{LMAT} (version 1.2.4, with Feb'15 release of data base \emph{LMAT-Grand}).
-LMAT was run using a Bull shared-memory node belonging to the team's HPC (high performance computing) cluster.
-It is equipped with 32 cores (64 threads available using Intel Hyper-threading technology) as it has 2 Haswell-based Xeons, the E5-2698v3@2.3 GHz, sharing half a tebibyte (0.5 TiB, that is, 512 gibibytes) of DRAM memory.
-This node is also provided with a card PCIe SSD as NVRAM, the P420m HHHL, with 1.4 TB, and 750000 reading IOPS, 4 KB, achieving 3.3 GB/s, which Micron kindly issued free of charge, as a sample for testing purposes.
-The computing node was supplied with a RAID-0 (striping) scratch disk area.
-We used the ``Grand'' database\footnote{In this context, ``Grand'' refers to a huge database that contains k-mers from all viral, prokaryote, fungal and protist genomes present in the NCBI database, plus Human reference genome (hg19), plus GenBank Human, plus the 1000 Human Genomes Project (HGP).
-This represent about 31.75 billion k-mers occupying 457.62 GB.}, release Feb'15, provided by the LMAT team.
-Previously to any calculation, the full database was loaded in the NVRAM.
-With this configuration the observed LMAT sustained sequence classification rate was $20\unclassrate$.
-Finally, it is worth mentioning that a complete set of Python scripts have been developed as back-end and front-end of the LMAT pipeline in order to manage the added complexity of time series analysis.
-
-\subsubsection*{Taxa level selection}
-We selected genus as taxonomic level for the subsequent steps of our work.
-In order to ensure that, between adjacent taxonomic levels, there were not crucial differences which could still be of relevance after standardization (see Section \ref{sec:stan}), we tested two different data sets.
-In the former, the antibiotics study\cite{antibiotic} with 16S data, we tested the differences between genus and family levels.
-The latter dataset tested was the kwashiorkor discordant twins study\cite{kwashiorkor} for both genus and species taxonomic levels.
-The Figures \ref{fig:taxlev1} (overview) and \ref{fig:taxlev2} (detail) plot the comparison between studies (and so, 16S and SMS) and between adjacent taxonomic levels.
+%%
+\subsubsection*{Fit Plots}
+For each and every dataset included in the study, an unweighted fit and a X-weighted fit (see Material and Methods section for details on both fits) have been calculated for standard deviations versus the mean values for each bacterial genus monitored in time.
+Figure \ref{fig:main1} showed the X-weighted fit for samples from a healthy subject (patient A, top) and from a subject diagnosed with irritable bowel syndrome (patient P2, bottom) studied in our lab\cite{IBS}, while Figure \ref{fig:fits} shows the corresponding unweighted fits.
+Both for the X-weighted and the unweighted fit, the IBS patient shows significative higher values of $V$ and $\beta$ than the healthy subject.
 
 \begin{figure}
-  \includegraphics[width=0.5\textwidth]{results/taxalevel/sum_raw_16S.pdf}
-  \includegraphics[width=0.5\textwidth]{results/taxalevel/sum_sta_16S.pdf}
-  \includegraphics[width=0.5\textwidth]{results/taxalevel/sum_raw_SMS.pdf} 
-  \includegraphics[width=0.5\textwidth]{results/taxalevel/sum_sta_SMS.pdf}
-\caption{Overview of comparison of different approaches based on adjacent taxonomic levels using plots in the Taylor-parameters space.
-For 16S (former row of subfigures), the levels are family vs.
-genus, whereas for SMS (latter row of subfigures) levels are genus vs.
-species.
-The left column shows the raw results and the right column plots the standardized results (see Section \ref{sec:stan})}
-\label{fig:taxlev1}
+	\centering
+	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_LOG.pdf}
+	\includegraphics[width=0.8\textwidth]{results/fits/IBS_P1_metatranscriptomes_family_stdVSmean_LLR_LOG.pdf}
+	\caption{Log plots of unweighted fits corresponding to the datasets shown in Figure \ref{fig:main1}, a healthy subject (top) and from a subject diagnosed with irritable bowel syndrome (bottom) studied in our lab \cite{IBS}}
+	\label{fig:fits}
+\end{figure}
+
+Additionally, for the unweighted fit, a complete residues analysis was performed, and a 4-in-1 figure was generated as shown in Figure \ref{fig:unwRes}, corresponding to patient A (top plot in Figure \ref{fig:main1} and Figure \ref{fig:fits}).
+This allows to check for normality and homoscedasticity of the residues and, therefore, to further assess the goodness of the model's fit to the data.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_RES.pdf}
+	\caption{Residues analysis plot corresponding to the unweighted fit for patient A of the IBS study\cite{IBS}.
+The top-left subplot is a simple residues plot.
+The top-right subplot is a Normal quantiles plot with linear fitting (value of coefficient of determination is provided).
+The bottom-left subplot shows an accumulated ESS (Explained Sum of Squares) plot.
+Finally, the bottom-right subplot is a residues Normal histogram plot.
+This set of subplots allows to check for normality and homoscedasticity of the residues and, therefore, to further assess the goodness of the model's fit to the real data.}
+	\label{fig:unwRes}
+\end{figure}
+
+\subsubsection*{Histogram Plots} 
+The \CC\ software generates two related histogram plots: absolute frequencies plot and zero relative frequency plot.
+The former is useful to visually assess the validity of the time points in terms of the accumulated absolute frequency of the elements (taxa), since absolute frequencies far (much higher or much lower) from those typically observed could mean a sampling problem.
+In Figure \ref{fig:histAFP} shows this histogram for the pre-treatment data (first 7 times) of patient ``D'' in the antibiotics study\cite{antibiotic}.
+We can see that there are no \emph{outlayers} among the total taxa sum for the sample points of this time series, so all of them were considered for further analysis by \CC.
+
+On the other hand, we could define the ZRF (Zero Relative Frequency, thereby ranging from 0 to 1) of an element (taxon) as the portion of times of the time series where the corresponding frequency is zero, i.e., such taxon is not found.
+The ZRF is useful for assessing how regularly is appearing the microbiota in the samples.
+In that sense, we define the ``core'' as the taxa that has a ZRF of less than 0.2, i.e., is present in more than the 80\% of the samples.
+
+Attending to all taxa in a dataset, we can plot the ZRF histogram, which then lies on the horizontal axis of the plot, with the vertical axis showing the number of taxa.
+So, the height of a bar represents the amount of taxa that have determinate ZRF.
+In this respect, the bar near $0$ counts the quantity of taxa that are present at every (or almost every) time point of the data set.
+In the same way, in the opposite case, the bar near $1$ counts the total of taxa that are scarcely found but present at least in one time (the software filters any taxon not appearing at any time).
+Figure \ref{fig:histZRF} shows this plot for the healthy patient A of the IBS study performed in our lab\cite{IBS}.
+There, we can see that $12$ taxa are present at all the time points of the time series while $9$ taxa basically appear only once.
+Ranging between those extremes, there are a significative amount of elements, $23$ taxa, appearing from often to seldom, but not belonging to the ``core'' region.
+
+\begin{figure}		
+	\centering		
+ 	\includegraphics[width=0.8\textwidth]{results/hist/antibiotics_h_patD_pre_AbsFreqPlot}		
+ 	\caption{Histogram with the absolute frequencies of the pre-treatment data (7 first times) of patient ``D'' in the antibiotics study\cite{antibiotic}.
+There are no \emph{outlayers} among the total taxa sum for the time series sample points, so all of them were considered for further analysis by \CC\ software}
+ 	\label{fig:histAFP}		
 \end{figure}
@@
 \begin{figure}
-  \includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_fam_16S.pdf}
-  \includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_gen_16S.pdf}
-  \includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_gen_SMS.pdf}
-  \includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_spc_SMS.pdf}
-\caption{Detail of comparison of different approaches based on adjacent taxonomic levels using plots of X-weighted power-law fits (see Material and Methods).
-The former row of subfigures shows examples for 16S, whereas the latter row of subfigures plots examples for SMS.
-The left column shows results for the superior taxonomic level (family for 16S, genus for SMS), while the right column shows results for the inferior level (genus for 16S, specie for SMS).}
-\label{fig:taxlev2}
+	\centering
+	\includegraphics[width=0.8\textwidth]{results/hist/IBS_h_A_amplicons_family_ZRFhist.pdf}
+	\caption{Histogram with the relative frequency of zero for the elements (taxa) in the data for patient A of the IBS study\cite{IBS}.
+In this particular case, the bar near $0$ counts the quantity of taxa that are present at almost every time point of the data set, while the bar near $1$ sums the total of taxa that are scarcely found: here, $12$ taxa are present at all the time points of the time series while $9$ taxa basically appear only once.}
+	\label{fig:histZRF}
 \end{figure}
 
+A 2D semi-logarithmic histogram representing deviations from the mean versus the mean itself is a useful tool in the analysis of the stability of ranking processes in complex systems\cite{ranking}.
+Figure \ref{fig:hist2D} shows this 2D deviation plot for the patient A of the IBS study\cite{IBS}.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.8\textwidth]{results/hist/IBS_h_A_amplicons_family_hist2D.pdf}
+	\caption{2D histogram deviation plot of the data for patient A of the IBS study\cite{IBS}}
+	\label{fig:hist2D}
+\end{figure}
+
+%% Partial inclusion
+\subsubsection*{Correlation and Rank Plots} 
+\CC\ generates two different plots falling under this category in addition to the Excel files with the resulting matrices.
+On the one hand, the elements correlation matrix plot shows a correlation matrix among the taxa, calculated with the time as independent variable.
+For these calculations, the data set is not normalized to avoid entering an additional constraint.
+Figure \ref{fig:corrElm} shows this matrix for the most dominant taxa present in the data of the patient ``A'' of the IBS study\cite{IBS}.
+On the other hand, the rank dynamics and stability plot shows the variation in the rank with time for the most dominant taxa and their calculated RSI, as discussed in Material and Methods.
+Figure \ref{fig:corrank} shows this plot for the taxa of the aforementioned patient A.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\textwidth]{results/corrank/IBS_h_A_amplicons_family_ElmCorrelation.pdf}
+	\caption{Element correlation plot of the for the most dominant taxa in the data for patient A of the IBS study\cite{IBS}}
+	\label{fig:corrElm}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{results/corrank/IBS_h_A_amplicons_family_Rank.pdf}
+	\caption{Matrix showing the rank variation throughout time for the most dominant elements (taxa) and their calculated Rank Stability Index (as discussed in Material and Methods) in the data  for patient A of the IBS study\cite{IBS}}
+	\label{fig:corrank}
+\end{figure}
+
+%%
 \subsection*{ComplexCruncher}
@@
 
+%%
 \subsection*{Un-weighted power-law fit}\label{sec:unw}
@@
 
+%% Partially included
 \subsection*{X-weighted power-law fit}\label{sec:X-w}
@@
 
+%% Omitted
 \subsubsection*{Method 1: By inverting the data}
@@
 
+%% Partially included
 \subsubsection*{Method 2: Bootstrapping-like strategy}
@@
 
+%% Mostly included
 \subsection*{Rank Stability Index (RSI)}\label{sec:RSI}
@@
   \begin{center}
-    \begin{tabular}{ccc}
+    \begin{tabular}{cccr}
     \hline
-    Case  &  Condition  &  Colour  \\
+    Case  &  Condition  &  Colour  &  Description \\
     \hline
-    1  &  $1\ge{\rm RSI}>0.99$  & \textcolor{blue}{blue}  \\ 
-    2  & ${\rm RSI}>0.90$  &  \textcolor{green}{green}  \\
-    3  &  ${\rm RSI}>0.75$  &  \textcolor{orange}{orange} \\
-    4  &  ${\rm RSI}>0.25$  &  \textcolor{red}{red} \\
-    5  &  $0.25\ge{\rm RSI}\ge0$  &  \bfseries{black}  \\
+    1  &  $1\ge{\rm RSI}>0.99$  & \textcolor{blue}{blue} & constant rank \\ 
+    2  & ${\rm RSI}>0.90$  &  \textcolor{green}{green}  & highly stable rank \\
+    3  &  ${\rm RSI}>0.75$  &  \textcolor{orange}{orange} & moderately stable rank \\
+    4  &  ${\rm RSI}>0.25$  &  \textcolor{red}{red} & unstable rank \\
+    5  &  $0.25\ge{\rm RSI}\ge0$  &  \bfseries{black} & very unstable rank \\
     \hline
@@
   \end{center} 
-  \caption{Colour code of the RSI percentage text shown in rank plots, following the first condition satisfied.}
+  \caption{Colour code of the RSI percentage text shown in Figure \ref{fig:corrank}, following the first condition satisfied.}
   \label{tab:RSI}
 \end{table}
-
-\subsection*{Standardization} \label{sec:stan}
-In order to show all the studies properly under common axes, we decided to standardize the Taylor parameters using the group of healthy individuals for each study.
-With this approach, all the studies can be visualized in a shared plot with units of Taylor-parameters standard-deviation on their axes.
-
-For a Taylor parameter, e.g. $V$, the estimate of the mean ($\widehat{V}$) for the healthy subpopulation, composed of $h$ individuals, is:
-\begin{linenomath}
-$$\widehat{V} = \frac{1}{W_1}\sum_{i=1}^h V_i \omega_i=\sum_{i=1}^h V_i \omega_i$$
-\end{linenomath}
-as $W_1=\sum_i^h \omega_i=1$, since $\omega_i$ are normalized weights calculated as:
-\begin{linenomath}
-$$\omega_i = \frac{\frac{1}{\sigma^2_{V_i}}}{\sum_i^h\frac{1}{\sigma^2_{V_i}}}$$
-\end{linenomath}
-being $\sigma_{V_i}$ the estimation of the uncertainty in $V_i$ obtained together with $V_i$ from the X-weighted power-law fit described in Section \ref{sec:X-w}, for healthy individuals.
-
-Likewise, the estimation of the standard deviation for the healthy population ($\widehat{\sigma}_V$) is:
-\begin{linenomath}
-$$\widehat{\sigma}_V = \sqrt{\frac{1}{W_1-\frac{W_2}{W_1}}\sum_{i=1}^h\left[\omega_i\left(V_i-\hat{V}\right)^2\right]}$$
-\end{linenomath}
-being $W_2=\sum_i^h \omega_i^2$, which finally yields to:
-\begin{linenomath}
-$$\widehat{\sigma}_V = \sqrt{\frac{1}{1-\sum_i^h \omega_i^2}\sum_{i=1}^h\left[\omega_i\left(V_i-\hat{V}\right)^2\right]}$$
-\end{linenomath}

### 2016-05-18 mSystems/mSys_MAIN.tex
Detach CC from mSys and improve rank stability results, which leads to 0.2.0
On branch master
	copied:     mSys_MandM.tex -> 2CCpaper.tex
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_discussion.tex
	modified:   mSys_results.tex
	renamed:    results/corrank/IBS_P2 Metatranscriptores_Rank.pdf -> results/corrank/IBS_P2_Metatranscriptores_Rank.pdf

@@
 % release codenumber (update it here!)
-\newcommand\releasecode{0.1.0}
+\newcommand\releasecode{0.2.0}
 
@@
 % watermarks (time consuming!)
-\newwatermark[allpages,color=red!4,angle=45,scale=9,xpos=-40,ypos=0]{DRAFT}
-\newwatermark[allpages,color=red!12,angle=0,scale=1,xpos=0,ypos=137]{
+\newwatermark[allpages,color=blue!4,angle=45,scale=9,xpos=-40,ypos=0]{DRAFT}
+\newwatermark[allpages,color=blue!12,angle=0,scale=1,xpos=0,ypos=137]{
 \emph{mSystems} \LaTeX\ manuscript release \releasecode}
@@
 
-\author[1,2]{Jose Manuel Martí}
-\author[1,2,3]{Daniel Martínez-Martínez}
+\author[1,2,$*$]{Jose Manuel Martí}
+\author[1,2,3,$*$]{Daniel Martínez-Martínez}
 \author[2]{Manuel Peña}
@@
 \wordcount
+\footnote[0]{$^*$ Equally contributed}
 \footnote[0]{$^\#$ Corresponding author: penagaray@gmail.com}
@@
 \input{mSys_intro.tex}
+\clearpage
 %%% Result
@@
 \input{mSys_results.tex}
+\clearpage
 %%% Discussion
@@
 \input{mSys_discussion.tex}
+\clearpage    
 %%% Material & Methods
@@
 \input{mSys_MandM.tex}   
+\clearpage
 
@@
 
-\newpage
+\clearpage
 \begin{thebibliography}{99}
@@
 \bibitem{QIIME} Caporaso, J.G. et al. QIIME allows analysis of high-throughput community sequencing data. {\it Nature Methods} {\bf 7,} 335-6 (2010).
-\bibitem{LMAT} Ames, S.K. et al. Scalable metagenomic taxonomy classification usng a reference genome database. {\it Bioinformatics} {\bf 29,} 2253-2260 (2013).
+\bibitem{LMAT} {\bf Ames SK, Hysom DA, Gardner SN, Lloyd GS, Gokhale MB, Allen JE}.
+2013.
+Scalable metagenomic taxonomy classification using a reference genome database.
+Bioinformatics {\bf 29}:2253-2260.
+\bibitem{LMAT2} {\bf Ames SK, Gardner SN, Marti JM, Slezak TR, Gokhale MB, Allen JE}.
+2015.
+Using populations of human and microbial genomes for organism detection in metagenomes.
+Genome Res.
+{\bf 25}:1056-67.
 \bibitem{fs} Eisler,Z., Bartos,I., Kertesz,J. Fluctuation scaling in complex systems: Taylor's law and beyond. {\it Adv. Phys.} {\bf 57,} 85 (2008).

### 2016-05-18 mSystems/mSys_MandM.tex
Detach CC from mSys and improve rank stability results, which leads to 0.2.0
On branch master
	copied:     mSys_MandM.tex -> 2CCpaper.tex
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_discussion.tex
	modified:   mSys_results.tex
	renamed:    results/corrank/IBS_P2 Metatranscriptores_Rank.pdf -> results/corrank/IBS_P2_Metatranscriptores_Rank.pdf

@@
-%%%
+%%
 \subsection*{Model} \label{sec:model}
@@
 
-%%%
+%%
 \subsection*{Selection and Methods}
@@
 
+%%
 \subsubsection*{Sample selection}
@@
 
+%%
 \subsubsection*{16rRNA sequences processing}
@@
 
+%%
 \subsubsection*{Metagenomic sequences processing}
@@
 The computing node was supplied with a RAID-0 (striping) scratch disk area.
-We used the ``Grand'' database\footnote{In this context, ``Grand'' refers to a huge database that contains k-mers from all viral, prokaryote, fungal and protist genomes present in the NCBI database, plus Human reference genome (hg19), plus GenBank Human, plus the 1000 Human Genomes Project (HGP).
-This represent about 31.75 billion k-mers occupying 457.62 GB.}, release Feb'15, provided by the LMAT team.
+We used the ``Grand'' database\cite{LMAT2}, release Feb'15, provided by the LMAT team, where ``Grand'' refers to a huge database that contains k-mers from all viral, prokaryote, fungal and protist genomes present in the NCBI database, plus Human reference genome (hg19), plus GenBank Human, plus the 1000 Human Genomes Project (HGP) (this represent about 31.75 billion k-mers occupying 457.62 GB)\cite{LMAT2}.
 Previously to any calculation, the full database was loaded in the NVRAM.
@@
 
+%%
 \subsubsection*{Taxa level selection}
 We selected genus as taxonomic level for the subsequent steps of our work.
-In order to ensure that, between adjacent taxonomic levels, there were not crucial differences which could still be of relevance after standardization (see Section \ref{sec:stan}), we tested two different data sets.
+In order to ensure that, between adjacent taxonomic levels, there were not crucial differences which could still be of relevance after standardization (see last subsection of Material and Methods), we tested two different data sets.
 In the former, the antibiotics study\cite{antibiotic} with 16S data, we tested the differences between genus and family levels.
@@
 
-\subsection*{ComplexCruncher}
-
-A complete software framework, named 'ComplexCruncher', has been engineered to support the analysis of the dynamics of ranking processes in complex systems.
-Although the software was devised with a clear bias towards metagenomics, it is general enough to be able to cope with a ranking process in any complex system.
-Implemented in Python using well-known open-source community software, the software solution is composed of two parts that can be used together or apart: a web-based graphic front-end connected to a database, and a computing kernel.
-Used together, this software enables other users to reproduce our results easily and, furthermore, upload and analyse their own data or experiment with the preloaded metagenomics data sets.
-
-`ComplexCruncher WebPortal' (CCWebPortal) is a web platform designed to allow the user to interact with a data repository of selected and well-documented metagenomics data sources.
-Through a few simple steps, the user can perform advanced searches on the complete set of records in the metagenomics repository.
-The web application provides advanced filters that allow the user to reduce the search to a small set of interest.
-After this first step, the user can refine the search and discard those records that do not meet certain requirements.
-
-The web application allows calculations to be done directly by the stable release of the \CC\ computing kernel.
-At the end of the calculations, the results are displayed to the user on the same browser which runs the web application.
-Then, the user can interact over the series of generated graphics thus allowing flexible comparison among them.
-In addition, CCWebPortal enables direct download of generated data (plots, spreadsheets, etc).
-The web application generates a report file summarizing all the results in PDF format.
-If the user has login permissions, CCWebPortal enables the option of insert new database records in addition to editing and deleting existing ones.
-
-CCWebPortal is a web application that runs on current versions of many browsers. Additional software is not needed and only requires javaScript to be enabled on the browser to run applications. 
-CCWebPortal is implemented following the client$-$server distributed programming model, where the javaScript client application connects to a remote server that enables the execution of calculations and transactions through a centralized database management system.
-A set of relational tables allows the structuring of the metagenomics repository to establish relationships between records.
-Thus the search and information threshing is optimized for queries launched from the client interface.
-Access to the database on the server is implemented through Django framework, an open-source framework written in Python using the model-view-controller (MVC) architectural pattern for implementing user interfaces.
-
-The effective data analysis has been performed with a Python tool developed from scratch to more than 4200 lines of code.
-Implemented following the Object Oriented Programming paradigm, this software is the back-end of the website described above.
-However, it could be run as an independent piece of software since it is built as a Python package provided with a command-line front-end (\CC.py).
-Once installed, the tool can be run interactively but also in automatic mode, which uses parallel computation to speed up the analysis of several data sources.
-
-\CC\ performs the power-law fit described in the \emph{Blumm, N.
-et al.} paper, but by fitting the best model, i.e.
-choosing between fitting a power-law using linear regression versus nonlinear regression\cite{ecology}.
-In the power-law fit plots we also show the generalized coefficient of determination computed for continuous models\cite{genR2,disR2}.
-
-\subsection*{Un-weighted power-law fit}\label{sec:unw}
-
-\subsubsection*{Fitting the best model}
-
-As already mentioned, to choose between fitting power laws ($y=Vx^\beta$) using linear regression on log-transformed (LLR) data versus non-linear regression (NLR), we mainly follow \emph{General Guidelines for the Analysis of Biological Power Laws}\cite{ecology}.
-It consists of the following three steps:
-\begin{enumerate}
-    \item Determining the appropriate error structure by likelihood analysis.
-    \begin{enumerate}
-        \item Fit the Non-Linear Regression (NLR) model and obtain $V_\text{NLR}$, $\beta_\text{NLR}$ and $\sigma_\text{NLR}^2$.
-        \item Calculate the loglikelihood that the data ($n$ is sample size) are generated from a normal distribution with additive error:
-        \begin{itemize}
-            \item The likelihood of a normal distribution is: 
-            \begin{linenomath}
-            $$\mathcal{L}_\text{norm} = \prod_{i=1}^n\left[\frac{1}{\sqrt{2\pi\sigma^2_\text{NLR}}}\;\exp{\left(-\frac{\left(y_i-V_\text{NLR}x_i^{\beta_\text{NLR}}\right)^2}{2\sigma^2_\text{NLR}}\right)}\right]$$
-            \end{linenomath}
-            \item So, the loglikelihood of a normal distribution is:
-            \begin{eqnarray*}
-                \log\mathcal{L}_\text{norm} &=& -\frac{n}{2}\log\left|2\pi\sigma^2_\text{NLR}\right| - \frac{1}{2\sigma^2_\text{NLR}}\underbrace{\sum_{i=1}^n\left(y_i-V_\text{NLR}x_i^{\beta_\text{NLR}}\right)^2}_{\mathrm{RSS
-                }_\text{NLR}}\\
-                &=& -\frac{n}{2}\log\left|2\pi\sigma^2_\text{NLR}\right|-\frac{\mathrm{RSS}_\text{NLR}}{2\sigma^2_\text{NLR}}
-            \end{eqnarray*}
-        \end{itemize}
-        \item Calculate the \emph{corrected Akaike's Information Criterion} for the NLR model:
-		\begin{linenomath}
-        $$\mathrm{AIC_{c_{NLR}}} = 2k - 2\log\mathcal{L}_\text{norm} + \frac{2k(k+1)}{n-k-1}$$
-		\end{linenomath}
-        \item Fit the Log-transformed Linear Regression (LLR) model and obtain $V_\text{LLR}$, $\beta_\text{LLR}$ and $\sigma_\text{LLR}^2$.
-        \item Calculate the loglikelihood that the data ($n$ is sample size) are generated from a lognormal distribution with multiplicative error:
-        \begin{itemize}
-            \item The likelihood of a lognormal distribution is: 
-            \begin{linenomath}
-            $$\mathcal{L}_\text{logn} = \prod_{i=1}^n\left[\frac{1}{y_i\sqrt{2\pi\sigma^2_\text{LLR}}}\;\exp{\left(-\frac{\left(\log|y_i|-\log|V_\text{LLR}|-\beta_\text{LLR}\log|x_i|\right)^2}{2\sigma^2_\text{LLR}}\right)}\right]$$
-            \end{linenomath}
-            \item So, the loglikelihood of a lognormal distribution is: 
-            \begin{eqnarray*}
-                \log\mathcal{L}_\text{logn} &=& -\frac{n}{2}\log\left|2\pi\sigma^2_\text{LLR}\right| - \sum_{i=1}^n\log|y_i| -\\
-                &&\qquad-\frac{1}{2\sigma^2_\text{LLR}}\underbrace{\sum_{i=1}^n\left(\log|y_i|-\log|V_\text{LLR}|-\beta_\text{LLR}\log|x_i|\right)^2}_{\mathrm{RSS
-                }_\text{LLR}}\\
-                &=& -\frac{n}{2}\log\left|2\pi\sigma^2_\text{LLR}\right| - \frac{\mathrm{RSS}_\text{LLR}}{2\sigma^2_\text{LLR}} - \sum_{i=1}^n\log|y_i|
-            \end{eqnarray*}
-        \end{itemize}
-        \item Calculate the \emph{corrected Akaike's Information Criterion} for the LR model: 
-        	\begin{linenomath}
-        	$$\mathrm{AIC_{c_{LLR}}} = 2k - 2\log\mathcal{L}_\text{logn} + \frac{2k(k+1)}{n-k-1}$$
-			\end{linenomath}
-    \end{enumerate}
-    \item Compare $\mathrm{AIC_{c_{NLR}}}$ with $\mathrm{AIC_{c_{LLR}}}$:
-    \begin{itemize}
-        \item If $\mathrm{AIC_{c_{NLR}}} - \mathrm{AIC_{c_{LLR}}} < -2$, the assumption of normal error is favoured compared to lognormal error, so proceed with the results obtained from the NLR fit.
-        \item If $\mathrm{AIC_{c_{NLR}}} - \mathrm{AIC_{c_{LLR}}} > 2$, the assumption of lognormal error is favoured compared to normal error, so proceed with the results obtained from the LLR fit.
-        \item If $\left|\mathrm{AIC_{c_{NLR}}} - \mathrm{AIC_{c_{LLR}}}\right| \leq 2$, no model is favoured, so proceed with model averaging:
-        \begin{eqnarray*}
-            B_\text{av} &=& w_\text{NLR}V_\text{NLR} + w_\text{LLR}V_\text{LLR} \\
-            \beta_\text{av} &=& w_\text{NLR}\beta_\text{NLR} + w_\text{LLR}\beta_\text{LLR}
-        \end{eqnarray*}
-        where: 
-        \begin{eqnarray*}
-            w_\text{NLR} &=& \frac{1}{1+\mathrm{e}^{\frac{1}{2}\left(\mathrm{AIC_{c_{NLR}}}-\mathrm{AIC_{c_{LLR}}}\right)}} \\
-            w_\text{LLR} &=& \frac{1}{1+\mathrm{e}^{\frac{1}{2}\left(\mathrm{AIC_{c_{LLR}}}-\mathrm{AIC_{c_{NLR}}}\right)}}
-        \end{eqnarray*}
-        which are obtained to fulfill the next condition: $w_\text{NLR} + w_\text{LLR} = 1$.
-The CIs for $B_\text{av}$ and $\beta_\text{av}$ are to be generated by ordinary bootstrapping\footnote{\CC\ has available the next bootstrapping alternatives\cite{boot}: ordinary, ``Resampling Residuals'' method, ``Wild'' method, and ``Monte-Carlo'' method.}.
-    \end{itemize}
-    \item Assess the validity of the underlying statistical assumptions with diagnostic plots because while it is rare for all the assumptions to be fully satisfied by real-life data sets, major violations indicate the lack of appropriateness of the model and, thus, the potential invalidity of the results.
-\end{enumerate}
-
-\subsubsection*{Calculating the coefficient of determination}
-
-We think the best approach in this situation is to apply the generalized $R^2$ that, for continuous models, was defined as\cite{genR2}:
-\begin{linenomath}
-$$ R^2 = 1 - \left(\frac{\mathcal{L}(0)}{\mathcal{L}(\hat\theta)}\right)^{\!\!\frac{2}{n}} $$
-\end{linenomath}
-where $\mathcal{L}(\hat\theta)$ and $\mathcal{L}(0)$  denote the likelihoods of the fitted and the ``null'' model, respectively, and $n$ is the sample size.
-In terms of the loglikelihoods, the generalized coefficient of determination would be:
-\begin{linenomath}
-$$ R^2 = 1 - \mathrm{e}^{-\frac{2}{n}\left(\log\mathcal{L}(\hat\theta)-\log\mathcal{L}(0)\right)} $$
-\end{linenomath}
-We have the likelihoods calculated from the previous section, but what about the ``null'' models?
-We understand that they are the models with only the intercept.
-So for the Gaussian additive error model:
-\begin{linenomath}
-$$\mathcal{L}_\text{norm}(0) = \prod_{i=1}^n\left[\frac{1}{\sqrt{2\pi\sigma^2_\text{NLR0}}}\;\exp{\left(-\frac{\left(y_i-\bar y\right)^2}{2\sigma^2_\text{NLR0}}\right)}\right]$$
-\end{linenomath}
-So:
-\begin{eqnarray*}
-\log\mathcal{L}_\text{norm}(0) &=& -\frac{n}{2}\log\left|2\pi\sigma^2_\text{NLR0}\right| - \frac{1}{2\sigma^2_\text{NLR0}}\sum_{i=1}^n\left(y_i-\bar y\right)^2\\
- &=& -\frac{n}{2}\left(\log\left|2\pi\sigma^2_\text{NLR0}\right| + 1 \right)
-\end{eqnarray*}
-since $\sigma^2_\text{NLR0}=\frac{1}{n}\sum\left(y_i-\bar y\right)^2=\frac{1}{n}\mathrm{TSS}_\text{NLR}$. Now, coming back to the coefficient of determination, we have:
-\begin{eqnarray*}
-R^2_\text{NLR} &=& 1 - \mathrm{e}^{\frac{2}{n}\left(\log\mathcal{L}_\text{NLR}(0)-\log\mathcal{L}_\text{NLR}(\hat\theta)\right)} = 1 - \exp{\left(\frac{\log(\mathrm{RSS}_\text{NLR})}{\log(\mathrm{TSS}_\text{NLR})}\right)}
- =\\ &=& 1 - \frac{\mathrm{RSS}_\text{NLR}}{\mathrm{TSS}_\text{NLR}} =
- 1 - \frac{\sum_{i=1}^n\left(y_i-V_\text{NLR}x_i^{\beta_\text{NLR}}\right)^2}{\sum_{i=1}^n\left(y_i-\bar y\right)^2}
-\end{eqnarray*}
-recovering the traditional expression for $R^2$. Using the same approach for calculating $R^2_\text{LLR}$, then:
-\begin{linenomath}
-$$\mathcal{L}_\text{logn}(0) = \prod_{i=1}^n\left[\frac{1}{y_i\sqrt{2\pi\sigma^2_\text{LLR0}}}\;\exp{\left(-\frac{\left(\log|y_i|-\log|B_\text{LLR0}|\right)^2}{2\sigma^2_\text{LLR0}}\right)}\right]$$
-\end{linenomath}
-So:
-\begin{eqnarray*}
-\log\mathcal{L}_\text{logn}(0) &=& -\frac{n}{2}\log\left|2\pi\sigma^2_\text{LLR0}\right| - \frac{1}{2\sigma^2_\text{LLR0}}\sum_{i=1}^n\left(\log|y_i|-\overline{\log|y|}\right)^2 - \sum_{i=1}^n\log|y_i|\\
- &=& -\frac{n}{2}\left(\log\left|2\pi\sigma^2_\text{LLR0}\right| + 1 \right) - \sum_{i=1}^n\log|y_i|
-\end{eqnarray*}
-since $\sigma^2_\text{LLR0}=\frac{1}{n}\sum\left(\log|y_i|-\overline{\log|y|}\right)^2=\frac{1}{n}\mathrm{TSS}_\text{logn}$.
-Again, recalling the expression for the generalized coefficient of determination, we have:
-\begin{eqnarray*}
-R^2_\text{LLR} &=& 1 - \mathrm{e}^{\frac{2}{n}\left(\log\mathcal{L}_\text{LLR}(0)-\log\mathcal{L}_\text{LLR}(\hat\theta)\right)} = 1 - \exp{\left(\frac{\log(\mathrm{RSS}_\text{LLR})}{\log(\mathrm{TSS}_\text{LLR})}\right)}
- =\\ &=& 1 - \frac{\mathrm{RSS}_\text{LLR}}{\mathrm{TSS}_\text{LLR}} =
- 1 - \frac{\sum_{i=1}^n\left(\log|y_i|-\log|V_\text{LLR}|-\beta_\text{LLR}\log|x_i|\right)^2}{\sum_{i=1}^n\left(\log|y_i|-\overline{\log|y|}\right)^2}
-\end{eqnarray*}
-
+%%
 \subsection*{X-weighted power-law fit}\label{sec:X-w}
 
-When fitting the power-law of std vs. mean, we can take into account that every mean has uncertainty and estimate it for a sample size $n$ by the SEM (\emph{Standard Error of the Mean}):
-\begin{linenomath}
-$$ \mathrm{SEM} = \frac{s}{\sqrt{n}}$$
-\end{linenomath}
-where $s$ is the sample standard deviation. So, the vector of weights is computed with:
-\begin{linenomath}
-$$ \mathbf{w} = \frac{1}{\overrightarrow{\mathrm{SEM}}} = \frac{\sqrt{\mathbf{n}}}{\mathbf{s}}$$
-\end{linenomath}
-
+When fitting the power-law of std vs.
+mean, we can take into account that every mean has uncertainty and estimate it for a sample size $n$ by the SEM (\emph{Standard Error of the Mean}).
 Here, the uncertainties affect the independent variable, so the fit is not so trivial as a Y-weighted fit, where the uncertainties affect the dependent variable.
 A standard approach to do this fit is: a) invert your variables before applying the weights, b) then perform the weighted fit, and finally, c) revert the inversion.
-This method is deterministic, but the approximate solution worsens with smaller $R^2$.
-For comparison, we develop a stochastic method by using a bootstrapping-like strategy that avoids the inversion and is applicable regardless of $R^2$.
-Both methods, detailed below, are implemented in \CC.
-
-\subsubsection*{Method 1: By inverting the data}
-
-In the case of the log-LR model, we have:
-\begin{linenomath}
-$$\log y = \log V + \beta\log x \quad\rightarrow\quad \underbrace{\log x}_{\tilde y} = \overbrace{-\frac{1}{\beta}\log V}^{b} + \overbrace{\frac{1}{\beta}}^{m}\underbrace{\log y}_{\tilde x}$$ 
-\end{linenomath}
-where $m$ determines the slope or gradient of the fitted line, and $b$ determines the point at which the line crosses the y-axis, otherwise known as the y-intercept.
-Once the model is fitted, the original parameters can be retrieved easily:
-\begin{eqnarray*}
-\beta &=& \frac{1}{m} \\
-V &=& \mathrm{e}^{-\beta b} = \mathrm{e}^{-\frac{b}{m}}
-\end{eqnarray*}
-Their respective uncertainties are to be obtained using \emph{error propagation}:
-\begin{eqnarray*}
-    \sigma_\beta &=& \left|\frac{\mathrm{d}\beta}{\mathrm{d}m}\right|\sigma_m \quad=\quad \frac{1}{m^2}\;\sigma_m \\
-    \sigma_V &=& \sqrt{\left(\frac{\partial V}{\partial b}\right)^{\!\!2}\sigma_b^2 +
-      \left(\frac{\partial V}{\partial m}\right)^{\!\!2}\sigma_m^2} \quad=\quad
-      \frac{1}{m}\;\mathrm{e}^{-\frac{b}{m}}\,\sqrt{\sigma_b^2 + \frac{b^2}{m^2}\;\sigma_m^2} 
-\end{eqnarray*}
-
-\subsubsection*{Method 2: Bootstrapping-like strategy}
+This method is deterministic, but the approximate solution worsens with smaller coefficients of determination.
+To overcome this limitation, we developed a stochastic method by using a bootstrapping-like strategy that avoids the inversion and is applicable regardless of the coefficient of determination.
 
@@
 where $v_i$ is a Gaussian random variable with mean $\mu_i=0$ and standard deviation $\sigma_i=\mathrm{SEM}_i$, as defined previously.
-For each replicate a complete un-weighted power-law fit is performed, as described in the previous section.
-It is worth mentioning that each replicate is filtered to avoid values of $x^*_i$ under \emph{eps} (obtained by \texttt{np.finfo(np.double).eps}) in order to keep away from the error of getting log of negatives or zero during the fit.
-
-We devised and implemented a multi-step algorithm to estimate the fit parameters that finishes when a relative error of less than $10^{-4}$ is achieved.
-It also ends if the number of steps reaches $100$ to avoid too much time lapse, to prevent any pathologic numeric case which, in fact, we still have not detected in all the data sets analyzed.
-
-In the previous version of the algorithm, for each step, the method generated $10$ replicates for each x-data point, in other words, it was computing the fit for $10$ times the length of the x-data array replicates, with a maximum of $10000$ fits per step.
-Nevertheless, we found that such an approach depending on the length of the x-data array did not perform better, so we decided to simplify the method and fix the number of fits per step in $100$.
-This latter approach improved the performance.
-
+For each replicate a complete un-weighted power-law fit is performed, where to choose between fitting power laws ($y=Vx^\beta$) using linear regression on log-transformed (LLR) data versus non-linear regression (NLR) we mainly follow \emph{General Guidelines for the Analysis of Biological Power Laws}\cite{ecology}.
 The parameters of the X-weighted fit are then estimated by averaging through all the replicate fits performed, and their errors are estimated by computing the standard deviation also for all the fits.
 At the end of each step, the relative error is calculated by comparing the fit parameters estimation in the last step with the previous one.
-
 Finally, both the coefficient of determination of the fit and the coefficient of correlation between the fit parameters are estimated by averaging.
 
-\subsection*{Rank Stability Index (RSI)}\label{sec:RSI}
+%%
+\subsection*{Rank Stability Index}\label{sec:RSI}
 
-The Rank Stability Index is shown as a percentage in a separate bar on the right of the rank matrix plot provided by \CC.
+The Rank Stability Index (RSI) is shown as a percentage in a separate bar on the right of the rank matrix plot shown in Figure \ref{fig:corrank}.
 The RSI is strictly $1$ for an element whose range never changes over time, and is strictly $0$ for an element whose rank oscillates between the extremes from time to time.
@@
 where $D$ is the total of rank hops taken by the studied element, $N$ is the number of elements that have been ranked, and $t$ is the number of time samples.
-The power index $p$ is arbitrarily chosen to increase the resolution in the stable region; the value in the current version of the code is $p=4$.
+The power index $p=4$ is arbitrarily chosen to increase the resolution in the stable region.
 
-As an example of this ``zooming'' effect in the stable region, to match a linear ($p=1$) RSI of 0.9 to a powered one of 0.1, we should select $p=21.8543$.
-An alternative way to obtain this effect and exactly map a linear RSI of 0.9 to a non-linear RSI (${\rm RSI'}$) of 0.1, is by applying the following function:
-\begin{linenomath}
-$${\rm RSI'} = \frac{10^{10\left(1-\frac{D}{(N-1)(t-1)}\right)}-1}{10^{10}-1} \approx 10^{-10\left(\frac{D}{(N-1)(t-1)}\right)}$$
-\end{linenomath}
-where the approximation is valid because $10^{10}\gg 1$ but, the small price to pay for it is that, in the worst instability case, the ${\rm RSI'}$ would not be strictly $0$ but $10^{-10}$.
-
-The colour code of the RSI percentage text in the rank plot of \CC\ is chosen following the first condition satisfied from those shown in Table \ref{tab:RSI} (see page \pageref{tab:RSI}).
+The colour code of the RSI percentage text in the rank plot shown in Figure \ref{fig:corrank} is chosen following the first condition satisfied from those shown in Table \ref{tab:RSI}.
 
@@
   \end{center} 
-  \caption{Colour code of the RSI percentage text shown in rank plots, following the first condition satisfied.}
+  \caption{Colour code of the RSI percentage text shown in Figure \ref{fig:corrank}, following the first condition satisfied.}
   \label{tab:RSI}
@@
 
+%%
 \subsection*{Standardization} \label{sec:stan}
-In order to show all the studies properly under common axes, we decided to standardize the Taylor parameters using the group of healthy individuals for each study.
+In order to properly show all the studies under common axes, we decided to standardize the Taylor parameters using the group of healthy individuals for each study.
 With this approach, all the studies can be visualized in a shared plot with units of Taylor-parameters standard-deviation on their axes.

### 2016-05-18 mSystems/mSys_discussion.tex
abstract modified; intro greatly modified, with several new citations; discussion modified with an idea for last paragraph
Detach CC from mSys and improve rank stability results, which leads to 0.2.0
On branch master
	copied:     mSys_MandM.tex -> 2CCpaper.tex
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_discussion.tex
	modified:   mSys_results.tex
	renamed:    results/corrank/IBS_P2 Metatranscriptores_Rank.pdf -> results/corrank/IBS_P2_Metatranscriptores_Rank.pdf

@@
 Improved high-throughput sequencing of samples from individuals monitored over time and taxonomic assigning methods will provide a better distinction among pathologies or altered states of the microbiota.
+
+Specifically, the analysis of the rank stability of the samples of healthy and IBS diagnosed patients studied in our lab\cite{IBS}, suggests that the presence of a \emph{rank stability island} among medium-ranked taxa could be an indicator of a healthy microbiota.
+
+Final paragraph to talk about limitations and future perspectives: can we model stability in the functional landscape?
+Community assembly for itself doesn't explain everything, we need to move forward and think about what can be happening in complex ecosystems as the human microbiota.

### 2016-05-18 mSystems/mSys_results.tex
Detach CC from mSys and improve rank stability results, which leads to 0.2.0
On branch master
	copied:     mSys_MandM.tex -> 2CCpaper.tex
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_discussion.tex
	modified:   mSys_results.tex
	renamed:    results/corrank/IBS_P2 Metatranscriptores_Rank.pdf -> results/corrank/IBS_P2_Metatranscriptores_Rank.pdf

--- 2016-04-28 mSystems/mSys_results.tex
@@
-%%%
-\subsection*{Global results}
-
 We have analysed the microbiome temporal variability to extract global properties of the system.
@@
 %%%
-\subsection*{Specific results}
-
-\subsubsection*{Fit Plots}
-For each and every dataset included in the study, an unweighted fit and a X-weighted fit (see Material and Methods section for details on both fits) have been calculated for standard deviations versus the mean values for each bacterial genus monitored in time.
-Figure \ref{fig:main1} showed the X-weighted fit for samples from a healthy subject (patient A, top) and from a subject diagnosed with irritable bowel syndrome (patient P2, bottom) studied in our lab\cite{IBS}, while Figure \ref{fig:fits} shows the corresponding unweighted fits.
-Both for the X-weighted and the unweighted fit, the IBS patient shows significative higher values of $V$ and $\beta$ than the healthy subject.
+\subsection*{Rank stability of the taxa} 
+The rank dynamics and stability plot in Figure \ref{fig:corrank} shows the variation in the rank with time for the most dominant taxa and their calculated Rank Stability Index (RSI, as discussed in Material and Methods) for the taxa of a healthy subject (individual \emph{A}, top) and from a subject diagnosed with IBS (patient \emph{P2}, bottom) of the IBS study\cite{IBS}.
+The taxa are listed ordered by the accumulated frequency along the time series, so y-axis is an overall dominance axis for each sample set.
+Generally speaking, we observe that the most dominant taxa are the most rank stable.
 
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_LOG.pdf}
-	\includegraphics[width=0.8\textwidth]{results/fits/IBS_P1_metatranscriptomes_family_stdVSmean_LLR_LOG.pdf}
-	\caption{Log plots of unweighted fits corresponding to the datasets shown in Figure \ref{fig:main1}, a healthy subject (top) and from a subject diagnosed with irritable bowel syndrome (bottom) studied in our lab \cite{IBS}}
-	\label{fig:fits}
+	\vspace*{-10mm} % Corrects overbox of the figures
+	\includegraphics[width=0.8\textwidth]{results/corrank/IBS_h_A_amplicons_family_Rank.pdf}\vspace*{-3mm} % Corrects overbox of the figures
+	\hspace*{-4.5mm}\includegraphics[width=0.83\textwidth]{results/corrank/IBS_P2_Metatranscriptores_Rank.pdf}
+	\caption{Matrixes showing the rank variation throughout time for the most dominant elements (taxa) and their calculated Rank Stability Index (as discussed in Material and Methods).
+We show the matrix for samples from a healthy subject (top) and from a subject diagnosed with irritable bowel syndrome (bottom), studied in our lab\cite{IBS}.}
+	\label{fig:corrank}
 \end{figure}
 
-Additionally, for the unweighted fit, a complete residues analysis was performed, and a 4-in-1 figure was generated as shown in Figure \ref{fig:unwRes}, corresponding to patient A (top plot in Figure \ref{fig:main1} and Figure \ref{fig:fits}).
-This allows to check for normality and homoscedasticity of the residues and, therefore, to further assess the goodness of the model's fit to the data.
+Nevertheless, in the particular case of the healthy individual, \emph{Burkholderiales} and \emph{Betaproteobacteria} (taxa ordered as 18th and 25th in the dominance axis) show comparatively very low rank stability regarding similar dominant taxa while, on the other hand, \emph{Comamonadaceae}, \emph{Lactobacillaceae}, \emph{Fusobacteriaceae}, \emph{Aerococcaceae} and \emph{Carnobacteriaceae} show higher stability than other more dominant taxa, forming a kind of \emph{rank stability island} for medium-ranked taxa around position 40 in the dominance axis, and thus colored in orange following Table \ref{tab:RSI} criteria, since they show a moderately stable RSI.
 
-\begin{figure}
-	\centering
-	\includegraphics[width=\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_RES.pdf}
-	\caption{Residues analysis plot corresponding to the unweighted fit for patient A of the IBS study\cite{IBS}.
-The top-left subplot is a simple residues plot.
-The top-right subplot is a Normal quantiles plot with linear fitting (value of coefficient of determination is provided).
-The bottom-left subplot shows an accumulated ESS (Explained Sum of Squares) plot.
-Finally, the bottom-right subplot is a residues Normal histogram plot.
-This set of subplots allows to check for normality and homoscedasticity of the residues and, therefore, to further assess the goodness of the model's fit to the real data.}
-	\label{fig:unwRes}
-\end{figure}
+In the IBS diagnosed patient, beyond the differences in dominance for the particular taxa, we still observe that the most dominant are the most rank stable.
+However, as opposed to the healthy individual results, far from presenting a \emph{rank stability island}, the medium-ranked taxa are very rank unstable, mostly due to transient (often one or two consecutive samples) but deep drops in their relative abundance, which are usually happening more than twice along the time series.
+That is, for instance, the case of \emph{Sphingobacteriales} with two non-consecutive samples dropping to 111th rank position.
+In other cases, the high rank instability comes from a rank fluctuation over all the time series, as for \emph{Streptococcaceae} and \emph{Burkholderiales}, which are ranking 26th and 29th respectively in the overall dominance axis but show very low RSI, and thus colored in black attending to Table \ref{tab:RSI}.
 
-\subsubsection*{Histogram Plots} 
-The \CC\ software generates two related histogram plots: absolute frequencies plot and zero relative frequency plot.
-The former is useful to visually assess the validity of the time points in terms of the accumulated absolute frequency of the elements (taxa), since absolute frequencies far (much higher or much lower) from those typically observed could mean a sampling problem.
-In Figure \ref{fig:histAFP} shows this histogram for the pre-treatment data (first 7 times) of patient ``D'' in the antibiotics study\cite{antibiotic}.
-We can see that there are no \emph{outlayers} among the total taxa sum for the sample points of this time series, so all of them were considered for further analysis by \CC.
-
-On the other hand, we could define the ZRF (Zero Relative Frequency, thereby ranging from 0 to 1) of an element (taxon) as the portion of times where it is zero, i.e., it is not found.
-Attending to all taxa, we can plot the ZRF histogram, which then lies on the horizontal axis of the plot.
-The vertical axis shows the number of taxa, so the height of a bar represents the amount of taxa that have determinate ZRF.
-In this respect, the bar over $0.0$ counts the quantity of taxa that are present at every time point of the data set (aka ``core''), while the bar over $1.0$ would count the total of taxa that are never found (this bar never appears because all these ``null'' elements are automatically filtered by the software.
-Figure \ref{fig:histZRF} shows this plot for the healthy patient A of the IBS study performed in our lab\cite{IBS}.
-There, we can see that $12$ taxa are present at all the time points of the time series while $9$ taxa basically appear only once.
-So, this plot is clearly useful to notice how the ``core'' is distributed.
-
-\begin{figure}		
-	\centering		
- 	\includegraphics[width=0.8\textwidth]{results/hist/antibiotics_h_patD_pre_AbsFreqPlot}		
- 	\caption{Histogram with the absolute frequencies of the pre-treatment data (7 first times) of patient ``D'' in the antibiotics study\cite{antibiotic}.
-There are no \emph{outlayers} among the total taxa sum for the time series sample points, so all of them were considered for further analysis by \CC\ software}
- 	\label{fig:histAFP}		
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.8\textwidth]{results/hist/IBS_h_A_amplicons_family_ZRFhist.pdf}
-	\caption{Histogram with the relative frequency of zero for the elements (taxa) in the data for patient A of the IBS study\cite{IBS}.
-In this particular case, the bar over $0.0$ counts the quantity of taxa that are present at every time point of the data set (``core''), while the bar over $1.0$ sums the total of taxa that are never found (this bar never appears because all these ``null'' elements are automatically filtered by \CC): here, $12$ taxa are present at all the time points of the time series while $9$ taxa basically appear only once.}
-	\label{fig:histZRF}
-\end{figure}
-
-A 2D semi-logarithmic histogram representing deviations from the mean versus the mean itself is a useful tool in the analysis of the stability of ranking processes in complex systems\cite{ranking}.
-Figure \ref{fig:hist2D} shows this 2D deviation plot for the patient A of the IBS study\cite{IBS}.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.8\textwidth]{results/hist/IBS_h_A_amplicons_family_hist2D.pdf}
-	\caption{2D histogram deviation plot of the data for patient A of the IBS study\cite{IBS}}
-	\label{fig:hist2D}
-\end{figure}
-
-\subsubsection*{Correlation and Rank Plots} 
-\CC\ generates two different plots falling under this category, as well as Excel files with the resulting matrices.
-On the one hand, the elements correlation matrix plot shows a correlation matrix among the taxa, calculated with the time as independent variable.
-For these calculations, the data set is not normalized to avoid entering an additional constraint.
-Figure \ref{fig:corrElm} shows this matrix for the most dominant taxa present in the data of the patient ``A'' of the IBS study\cite{IBS}.
-On the other hand, the rank dynamics and stability plot shows the variation in the rank with time for the most dominant taxa and their calculated RSI, as discussed in Material and Methods.
-Figure \ref{fig:corrank} shows this plot for the taxa of the aforementioned patient A.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\textwidth]{results/corrank/IBS_h_A_amplicons_family_ElmCorrelation.pdf}
-	\caption{Element correlation plot of the for the most dominant taxa in the data for patient A of the IBS study\cite{IBS}}
-	\label{fig:corrElm}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\textwidth]{results/corrank/IBS_h_A_amplicons_family_Rank.pdf}
-	\caption{Matrix showing the rank variation throughout time for the most dominant elements (taxa) and their calculated Rank Stability Index (as discussed in Material and Methods) in the data  for patient A of the IBS study\cite{IBS}}
-	\label{fig:corrank}
-\end{figure}
+We found the presence of such of \emph{rank stability island} for medium-ranked taxa in the other healthy subjects (\emph{B} and \emph{C}) of the IBS study\cite{IBS} together with its total absence for the other IBS diagnosed patient (patient \emph{P1}), which also presents very high rank instability in its medium-ranked taxa.
 

### 2016-05-23 mSystems/mSys_intro.tex
two more cites in introduction, minor changes in structure

@@
 The desire to understand the factors that influence human health and cause diseases has always been one of the major driving forces of biological research.
-We are populated by a myriad of microorganisms that are interacting with us in several physiological processes such as metabolism of the bile acids \cite{bileacids}, of the choline \cite{choline} or key-route metabolites as short-chain fatty acids \cite{scfa1, scfa2} which is also involved in immune system maturation \cite{scfa3, scfa4}.
+As evidence of new concepts 'holobiont' and 'hologenome' is increasing each day \cite{holo1, holo2}, research not only focus on the human physiology but also on the microbial population that surround ourselves.
+We are populated by a myriad of microorganisms that are interacting with us in several physiological processes such as metabolism of the bile acids \cite{bileacids}, of the choline \cite{choline} or key-route metabolites as short-chain fatty acids \cite{scfa1, scfa2} which are also involved in immune system maturation \cite{scfa3, scfa4}.
 Human microbiota has been suggested to be closely related to diseases like type 2 diabetes \cite{diabetes2}, cardiovascular disease (CVD) \cite{CVD}, irritable bowel syndrome \cite{IBS}, Crohn's disease \cite{CD}, some affections as obesity \cite{ob1, ob2}, malnutrition \cite{nutr} among other multiple diseases \cite{Moya_trends}.
@@
 
-Here we present the imprints of health status (healthy or disease) in macroscopic properties of microbiota, by studying the temporal variability.
+Here we present the imprints of health status (healthy or disease) in macroscopic properties of microbiota, by studying its temporal variability.
 We have analyzed more than 35000 time series of taxa from the gut microbiome of 97 individuals obtained from publicly available high throughput sequencing data on different conditions: diseases, diets, obese status, antibiotic therapy and healthy individuals.

### 2016-06-07 mSystems/mSys_results.tex
tables removed from results, minor changes

@@
 
-Taylor parameters describing the temporal variability of the gut microbiome in our sampled individuals are shown in Tables \ref{tab:diet} to \ref{tab:LEA}.
+Taylor parameters describing the temporal variability of the gut microbiome in our sampled individuals are shown in Tables S1 to S6.
 Our results hint at an ubiquitous behaviour.
@@
 
-\begin{table} 
-  \begin{center}
-    \begin{tabular}{ccccccc}
-	    \hline
-		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
-		\hline
-		A&$0.26 \pm 0.05$&$0.826 \pm 0.025$&$0.918$&&$3.1 \pm 0.9$&$1.2 \pm 0.6$\\
-		A&$0.32 \pm 0.06$&$0.857 \pm 0.025$&$0.924$&&$4.4 \pm 1.1$&$2.0 \pm 0.6$\\
-		A&$0.194 \pm 0.033$&$0.813 \pm 0.024$&$0.918$&&$1.9 \pm 0.6$&$0.9 \pm 0.6$\\
-		A&$0.24 \pm 0.04$&$0.824 \pm 0.020$&$0.924$&&$2.7 \pm 0.7$&$1.2 \pm 0.5$\\
-		A&$0.34 \pm 0.06$&$0.855 \pm 0.024$&$0.931$&&$4.7 \pm 1.1$&$1.9 \pm 0.6$\\
-		A&$0.30 \pm 0.05$&$0.847 \pm 0.022$&$0.921$&&$3.9 \pm 1.0$&$1.7 \pm 0.5$\\
-		A&$0.133 \pm 0.021$&$0.784 \pm 0.023$&$0.916$&&$0.7 \pm 0.4$&$0.2 \pm 0.6$\\
-		A&$0.25 \pm 0.04$&$0.831 \pm 0.024$&$0.929$&&$3.0 \pm 0.8$&$1.4 \pm 0.6$\\
-		\hline
-		P&$0.23 \pm 0.05$&$0.804 \pm 0.035$&$0.885$&&$2.6 \pm 0.9$&$0.7 \pm 0.8$\\
-		P&$0.097 \pm 0.018$&$0.705 \pm 0.031$&$0.891$&&$0.03 \pm 0.34$&$-1.6 \pm 0.7$\\
-		P&$0.037 \pm 0.006$&$0.642 \pm 0.025$&$0.881$&&$-1.12 \pm 0.11$&$-3.1 \pm 0.6$\\
-		P&$0.118 \pm 0.019$&$0.723 \pm 0.025$&$0.895$&&$0.4 \pm 0.4$&$-1.2 \pm 0.6$\\
-		P&$0.17 \pm 0.04$&$0.78 \pm 0.04$&$0.842$&&$1.5 \pm 0.7$&$0.1 \pm 0.9$\\
-		P&$0.123 \pm 0.020$&$0.757 \pm 0.026$&$0.914$&&$0.5 \pm 0.4$&$-0.4 \pm 0.6$\\
-		P&$0.19 \pm 0.05$&$0.77 \pm 0.04$&$0.871$&&$1.8 \pm 0.9$&$-0.0 \pm 0.9$\\
-		P&$0.121 \pm 0.020$&$0.736 \pm 0.027$&$0.921$&&$0.5 \pm 0.4$&$-0.9 \pm 0.6$\\
-		P&$0.187 \pm 0.034$&$0.771 \pm 0.030$&$0.908$&&$1.8 \pm 0.7$&$-0.1 \pm 0.7$\\
-		P&$0.097 \pm 0.015$&$0.735 \pm 0.025$&$0.922$&&$0.05 \pm 0.28$&$-0.9 \pm 0.6$\\
-	   \hline
-	   \hline
-    \end{tabular}
-  \end{center}
-  \caption{Taylor parameters for individuals with either animal-based (A) or plant-based (P) diets\cite{diet}.
-Previous to diet, the population sampled is described by $\bar{V} = 0.09 \pm 0.05, \bar{\beta} = 0.77 \pm 0.04$, which we used to describe the \emph{healthy zone} for this study.}
-  \label{tab:diet}
-\end{table}
-
- \begin{table} 
-  \begin{center}
-    \begin{tabular}{ccccccc}
-	    \hline
-		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
-		\hline
-		Ab&$0.35 \pm 0.07$&$0.81 \pm 0.04$&$0.925$&&$4.3 \pm 1.4$&$1.3 \pm 0.9$\\
-		Ab&$0.41 \pm 0.09$&$0.82 \pm 0.04$&$0.908$&&$5.6 \pm 1.8$&$1.6 \pm 0.9$\\
-		Ab&$0.23 \pm 0.04$&$0.770 \pm 0.031$&$0.920$&&$2.1 \pm 0.8$&$0.5 \pm 0.7$\\
-		Ab&$0.165 \pm 0.029$&$0.738 \pm 0.031$&$0.928$&&$0.9 \pm 0.6$&$-0.3 \pm 0.7$\\
-		Ab&$0.34 \pm 0.06$&$0.812 \pm 0.032$&$0.936$&&$4.1 \pm 1.2$&$1.5 \pm 0.7$\\
-		Ab&$0.26 \pm 0.05$&$0.798 \pm 0.033$&$0.931$&&$2.8 \pm 0.9$&$1.1 \pm 0.8$\\
-	    \hline
-	    \hline
-    \end{tabular}
-  \end{center}
-  \caption{Taylor parameters for individuals taking antibiotics\cite{antibiotic}.
-Prior to antibiotics intake, the population sampled is described by $\bar{V} = 0.12 \pm 0.05, \bar{\beta} = 0.75 \pm 0.04$, which characterize the \emph{healthy zone} for this study.}
-  \label{tab:antibiotics}
-\end{table}
-
-\begin{table} 
-  \begin{center}
-    \begin{tabular}{ccccccc}
-	    \hline
-		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
-		\hline
-		IBS&$0.204 \pm 0.034$&$0.739 \pm 0.029$&$0.916$&&$7.6 \pm 3.7$&$1.9 \pm 1.2$\\
-		IBS&$0.35 \pm 0.05$&$0.793 \pm 0.023$&$0.935$&&$23.1 \pm 5.9$&$4.0 \pm 0.9$\\
-	     \hline
-	     \hline
-    \end{tabular}
-  \end{center}
-  \caption{Taylor parameters for persons diagnosed with irritable bowel syndrome (IBS)\cite{IBS}.
-Healthy individuals sampled in this study are characterized by $\bar{V} = 0.134 \pm 0.009, \bar{\beta} = 0.691 \pm 0.025$, which we used to define the correspondent \emph{healthy zone}.}
-  \label{tab:IBS}
-\end{table}
-
- \begin{table} 
-  \begin{center}
-    \begin{tabular}{ccccccc}
-	    \hline
-		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
-		\hline
-		DH&$0.27 \pm 0.04$&$0.835 \pm 0.016$&$0.925$&&$0.2 \pm 0.4$&$-1.0 \pm 0.6$\\
-		DH&$0.36 \pm 0.06$&$0.858 \pm 0.015$&$0.929$&&$1.1 \pm 0.6$&$-0.2 \pm 0.5$\\
-		DH&$0.35 \pm 0.06$&$0.859 \pm 0.014$&$0.926$&&$1.0 \pm 0.5$&$-0.1 \pm 0.5$\\
-		DH&$0.25 \pm 0.04$&$0.829 \pm 0.014$&$0.911$&&$0.0 \pm 0.4$&$-1.2 \pm 0.5$\\
-		DH&$0.30 \pm 0.05$&$0.844 \pm 0.014$&$0.920$&&$0.5 \pm 0.4$&$-0.7 \pm 0.5$\\
-		DH&$0.29 \pm 0.05$&$0.850 \pm 0.016$&$0.915$&&$0.4 \pm 0.5$&$-0.5 \pm 0.5$\\
-		DH&$0.28 \pm 0.05$&$0.848 \pm 0.016$&$0.921$&&$0.3 \pm 0.5$&$-0.5 \pm 0.6$\\
-		DH&$0.35 \pm 0.07$&$0.861 \pm 0.017$&$0.918$&&$0.9 \pm 0.6$&$-0.0 \pm 0.6$\\
-		DH&$0.31 \pm 0.04$&$0.833 \pm 0.012$&$0.916$&&$0.6 \pm 0.4$&$-1.1 \pm 0.4$\\
-		DH&$0.33 \pm 0.05$&$0.843 \pm 0.013$&$0.925$&&$0.8 \pm 0.5$&$-0.7 \pm 0.5$\\
-		DH&$0.31 \pm 0.05$&$0.852 \pm 0.014$&$0.925$&&$0.6 \pm 0.5$&$-0.4 \pm 0.5$\\
-		DH&$0.31 \pm 0.05$&$0.853 \pm 0.015$&$0.930$&&$0.6 \pm 0.5$&$-0.4 \pm 0.5$\\
-		DH&$0.203 \pm 0.033$&$0.815 \pm 0.015$&$0.907$&&$-0.44 \pm 0.32$&$-1.7 \pm 0.5$\\
-		\hline
-    \end{tabular}
-  \end{center}
-  \caption{Taylor parameters for the healthy subject of the discordant twins\cite{kwashiorkor}.
-This table continues in Table \ref{tab:DK}.
-The population of healthy twins is characterized by $\bar{V} = 0.25 \pm 0.10, \bar{\beta} = 0.863 \pm 0.028$, values which we used to describe the \emph{healthy zone} for this study.}
-  \label{tab:DH}
-\end{table}
-
-\begin{table} 
-  \begin{center}
-    \begin{tabular}{ccccccc}
-	    \hline
-		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
-		\hline
-		DK&$0.40 \pm 0.07$&$0.859 \pm 0.017$&$0.926$&&$1.5 \pm 0.7$&$-0.1 \pm 0.6$\\
-		DK&$0.44 \pm 0.08$&$0.868 \pm 0.016$&$0.919$&&$1.8 \pm 0.8$&$0.2 \pm 0.6$\\
-		DK&$0.196 \pm 0.031$&$0.819 \pm 0.014$&$0.916$&&$-0.50 \pm 0.30$&$-1.5 \pm 0.5$\\
-		DK&$0.160 \pm 0.026$&$0.798 \pm 0.015$&$0.904$&&$-0.85 \pm 0.25$&$-2.3 \pm 0.5$\\
-		DK&$0.30 \pm 0.05$&$0.845 \pm 0.014$&$0.924$&&$0.5 \pm 0.4$&$-0.6 \pm 0.5$\\
-		DK&$0.23 \pm 0.04$&$0.834 \pm 0.014$&$0.908$&&$-0.1 \pm 0.4$&$-1.0 \pm 0.5$\\
-		DK&$0.27 \pm 0.05$&$0.848 \pm 0.015$&$0.930$&&$0.2 \pm 0.4$&$-0.5 \pm 0.5$\\
-		DK&$0.35 \pm 0.07$&$0.860 \pm 0.019$&$0.916$&&$1.0 \pm 0.7$&$-0.1 \pm 0.7$\\
-		DK&$0.34 \pm 0.05$&$0.835 \pm 0.012$&$0.917$&&$0.9 \pm 0.5$&$-1.0 \pm 0.4$\\
-		DK&$0.25 \pm 0.04$&$0.831 \pm 0.012$&$0.912$&&$0.0 \pm 0.4$&$-1.1 \pm 0.4$\\
-		DK&$0.36 \pm 0.06$&$0.858 \pm 0.013$&$0.918$&&$1.1 \pm 0.5$&$-0.2 \pm 0.5$\\
-		DK&$0.31 \pm 0.06$&$0.851 \pm 0.016$&$0.924$&&$0.6 \pm 0.6$&$-0.4 \pm 0.6$\\
-		DK&$0.149 \pm 0.022$&$0.799 \pm 0.013$&$0.905$&&$-0.96 \pm 0.22$&$-2.2 \pm 0.5$\\
-	    \hline
-	    \hline
-    \end{tabular}
-  \end{center}
-  \caption{Taylor parameters for the kwashiorkor part of the discordant twins\cite{kwashiorkor}.
-This is a continuation of Table \ref{tab:DH}, so that the population of healthy twins is also characterized by $\bar{V} = 0.25 \pm 0.10$ and $\bar{\beta} = 0.863 \pm 0.028$.}
-  \label{tab:DK}
-\end{table}
-
-\begin{table} 
-  \begin{center}
-    \begin{tabular}{ccccccc}
-	    \hline
-		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
-		\hline
-		OW&$0.59 \pm 0.12$&$0.894 \pm 0.034$&$0.920$&&$6.6 \pm 2.0$&$2.6 \pm 1.0$\\
-		OW&$0.22 \pm 0.04$&$0.830 \pm 0.030$&$0.904$&&$0.5 \pm 0.6$&$0.7 \pm 0.9$\\
-		\hline
-		OBI&$0.28 \pm 0.04$&$0.855 \pm 0.022$&$0.958$&&$1.5 \pm 0.6$&$1.4 \pm 0.6$\\
-		OBI&$0.33 \pm 0.07$&$0.870 \pm 0.031$&$0.916$&&$2.4 \pm 1.1$&$1.9 \pm 0.9$\\
-		\hline
-		OBII&$0.223 \pm 0.032$&$0.823 \pm 0.023$&$0.938$&&$0.6 \pm 0.5$&$0.5 \pm 0.7$\\
-		OBII&$0.208 \pm 0.029$&$0.844 \pm 0.022$&$0.935$&&$0.4 \pm 0.5$&$1.1 \pm 0.7$\\
-		\hline
-		OBIII&$0.34 \pm 0.05$&$0.855 \pm 0.025$&$0.943$&&$2.5 \pm 0.9$&$1.4 \pm 0.7$\\
-		OBIII&$0.26 \pm 0.04$&$0.845 \pm 0.026$&$0.954$&&$1.1 \pm 0.7$&$1.2 \pm 0.8$\\
-		OBIII&$0.33 \pm 0.06$&$0.870 \pm 0.027$&$0.908$&&$2.4 \pm 1.0$&$1.9 \pm 0.8$\\
-		OBIII&$0.200 \pm 0.026$&$0.843 \pm 0.020$&$0.949$&&$0.2 \pm 0.4$&$1.1 \pm 0.6$\\
-		OBIII&$0.30 \pm 0.05$&$0.846 \pm 0.026$&$0.929$&&$1.9 \pm 0.8$&$1.2 \pm 0.7$\\
-		OBIII&$0.176 \pm 0.029$&$0.826 \pm 0.026$&$0.894$&&$-0.2 \pm 0.5$&$0.6 \pm 0.8$\\
-		OBIII&$0.30 \pm 0.06$&$0.841 \pm 0.031$&$0.896$&&$1.8 \pm 0.9$&$1.0 \pm 0.9$\\
-		OBIII&$0.28 \pm 0.04$&$0.857 \pm 0.025$&$0.941$&&$1.5 \pm 0.7$&$1.5 \pm 0.7$\\
-		OBIII&$0.122 \pm 0.018$&$0.822 \pm 0.024$&$0.930$&&$-1.05 \pm 0.30$&$0.5 \pm 0.7$\\
-		\hline
-		OBIIId&$0.47 \pm 0.08$&$0.872 \pm 0.023$&$0.945$&&$4.7 \pm 1.3$&$1.9 \pm 0.7$\\
-		OBIIId&$0.38 \pm 0.06$&$0.846 \pm 0.023$&$0.951$&&$3.2 \pm 1.0$&$1.2 \pm 0.7$\\
-		OBIIId&$0.36 \pm 0.06$&$0.842 \pm 0.022$&$0.954$&&$2.9 \pm 0.9$&$1.1 \pm 0.6$\\
-	    \hline
-	    \hline
-    \end{tabular}
-  \end{center}
-  \caption{Taylor parameters for individuals with different degrees of overweight and obesity\cite{LEA}.
-Healthy people in this study, whom were not obese, are characterized by $\bar{V} = 0.19 \pm 0.06, \bar{\beta} = 0.806 \pm 0.034$, which we used to determine the correspondent \emph{healthy zone} for this study.}
-  \label{tab:LEA}
-\end{table}
-
 In order to jointly visualize and compare the results of individuals from different studies, their Taylor parameters have been standardized, where standardization means that each parameter is subtracted by the mean value and divided by the standard deviation of the group of healthy individuals for each study (for details of the procedure, please see Standardization subsection in Material and Methods).
-The healthy zone and the standardized Taylor parameters for individuals whose gut microbiota is threatened (i.e., suffering from kwashiorkor, altered diet, antibiotics or IBS) is shown in Figure \ref{fig:main2}.
+The healthy zone and the standardized Taylor parameters for individuals whose gut microbiota is altered (i.e., suffering from kwashiorkor, altered diet, antibiotics or IBS) is shown in Figure \ref{fig:main2}.
 Children developing kwashiorkor show smaller variability than their healthy twins.
@@
 All other cases presented increased variability, which is particularly severe, and statistically significant at more than 95\% CL, for obese patients grade III on a diet, individuals taking antibiotics or IBS--diagnosed patients.
-A global property emerges from all worldwide data collected: Taylor parameters characterize the statistical behaviour of microbiome changes.
-Furthermore, we have verified that our conclusions are robust to systematic errors due to taxonomic assignment.
+A global property emerges from all worldwide data collected: Taylor parameters characterize the statistical behavior of microbiome changes.
+Furthermore, we have verified that our conclusions are robust to systematic errors due to taxonomic assignment (see Figure Sx in Supplemental Material).
 

### 2016-06-16 mSystems/mSys_MAIN.tex
two more cites in introduction, minor changes in structure
discussion started, more citations added

@@
 \begin{thebibliography}{99}
+\bibitem{holo1} {\bf Rosenberg E, Zilber-Rosenberg I.} 2016. Microbes Drive Evolution of Animals and Plants: the Hologenome Concept. MBio {\bf 7}:e01395–15–.
+\bibitem{holo2} {\bf Bordenstein SR, Theis KR.} 2015. Host Biology in Light of the Microbiome: Ten Principles of Holobionts and Hologenomes. PLOS Biol {\bf 13}:e1002226.
 \bibitem{bileacids} {\bf Swann JR, Want EJ, Geier FM, Spagou K, Wilson ID, Sidaway JE, Nicholson JK, Holmes E.} 2011.
@@
 
+%discussion
+\bibitem {kilpatrick} {\bf Kilpatrick a M, Ives a R.} 2003. Species interactions can explain Taylor’s power law for ecological time series. Nature {\bf 422}:65–68.
+\bibitem {ballantyne} {\bf Ballantyne IV F, J. Kerkhoff A.} 2007. The observed range for temporal mean-variance scaling exponents can be explained by reproductive correlation. Oikos {\bf 116}:174–180.
+\bibitem {joao} {\bf Stein RR, Bucci V, Toussaint NC, Buffie CG, Rätsch G, Pamer EG, Sander C, Xavier JB.}  2013.
+Ecological modeling from time-series inference: insight into dynamics and stability of intestinal microbiota.
+PLoS Comput Biol {\bf 9}:e1003388.
+\bibitem {mehta} {\bf Fisher CK, Mehta P.} 2014. Identifying keystone species in the human gut microbiome from metagenomic timeseries using sparse linear regression. PLoS One {\bf 9}:e102451.
+\bibitem {cohen_taylor} {\bf Cohen JE, Xu M, Schuster WSF}.
+2013.
+Stochastic multiplicative population growth predicts and interprets Taylor ’ s power law of fluctuation scaling Stochastic multiplicative population growth predicts and interprets Taylor ’ s power law of fluctuation scaling.
+Proc R Soc B Biol Sci {\bf 280}:20122955.
+
 	\bibitem{FD} Weber, J. \textit{et al.} Fluctuation dissipation theorem. {\it Phys. Rev.} {\bf 101}, 1620-6 (1956).

### 2016-06-16 mSystems/mSys_discussion.tex
discussion started, more citations added

@@
-We have quantitatively characterized whether the microbiota belongs to a healthy individual or a subject corresponding to an altered or pathological state (i.e., altered diet, antibiotic treatment, early gut development, diagnosed IBS).
-Deciphering the mechanisms of disease requires in depth knowledge of the underlying biological mechanisms.
-We describe here the macroscopic behavior of disease by a noise-induced phase transition with a control parameter that can be measured by the temporal variability of the microbiome.
-The microbiota of healthy individuals and of individuals with pathologies represent different phases separated by this noise-induced phase transition.
-Improved high-throughput sequencing of samples from individuals monitored over time and taxonomic assigning methods will provide a better distinction among pathologies or altered states of the microbiota.
+The main feature of this work is to have shown that the microbiota, independently of its condition, follows the Taylor's law.
+We have seen that the value of beta in each case is always less than the unity (using standard deviation in the law), which is informing us about the community structure.
+This means that the most abundant elements in the population are less volatile to perturbations in relative terms than the less abundant.
+The biological explanation for this universal pattern is not clear.
+There are studies that prove that having this kind of exponents may be due to the negative interactions in the population \cite{kilpatrick}, while others have demonstrated that it may depend on reproductive correlation \cite{ballantyne}.
+Nevertheless, none of these explanations are enough when we are talking about microbiota as the reproduction term is diffuse, and the interactions between its components are numerous and complex \cite{joao, mehta}.
+
+Model and temporal variation of V.
+
+From the biological perspective, too many questions arise with this work.
+Knowing that microbiota has universal properties, one of the main relevant questions is which mechanisms are involved in maintaining the population structure in order to have always values of beta below 1...
 

### 2016-06-21 mSystems/mSys_MAIN.tex
discussion in alpha version

@@
 Proc Natl Acad Sci {\bf 108}:4523–4530.
-\bibitem{choline} {\bf nSpencer MD, Hamp TJ, Reid RW, Fischer LM, Zeisel SH, Fodor AA.} 2011.
+\bibitem{choline} {\bf Spencer MD, Hamp TJ, Reid RW, Fischer LM, Zeisel SH, Fodor AA.} 2011.
 Association between composition of the human gastrointestinal microbiome and development of fatty liver with choline deficiency.
@@
 ISME J {\bf 8}:881–93.
+\bibitem{cohen_bac} {\bf Kaltz O, Escobar-Paramo P, Hochberg M, Cohen} JE.
+2012.
+Bacterial microcosms obey Taylor’s law: Effects of abiotic and biotic stress and genetics on mean and variance of population density.
+Ecol Process {\bf 1}:5.
+\bibitem {ramslayer} {\bf Ramsayer J, Fellous S, Cohen JE, Hochberg ME.} 2012.
+Taylor’s Law holds in experimental bacterial populations but competition does not influence the slope.
+Biol Lett {\bf 8}:316–319.
 \bibitem{cobas} {\bf Pérez-Cobas AE, Artacho A, Ott SJ, Moya A, Gosalbes MJ, Latorre A.} 2014.
@@
 \bibitem {mehta} {\bf Fisher CK, Mehta P.} 2014. Identifying keystone species in the human gut microbiome from metagenomic timeseries using sparse linear regression. PLoS One {\bf 9}:e102451.
+\bibitem {bucci}  {\bf Bucci V, Tzen B, Li N, Simmons M, Tanoue T, Bogart E, Deng L, Yeliseyev V, Delaney ML, Liu Q, Olle B, Stein RR, Honda K, Bry L, Gerber GK.} 2016.
+MDSINE: Microbial Dynamical Systems INference Engine for microbiome time-series analyses.
+Genome Biol {\bf 17}:121.
 \bibitem {cohen_taylor} {\bf Cohen JE, Xu M, Schuster WSF}.
@@
 Proc R Soc B Biol Sci {\bf 280}:20122955.
+\bibitem {koenig} {\bf Koenig JE, Spor A, Scalfone N, Fricker AD, Stombaugh J, Knight R, Angenent LT, Ley RE.} 2011.
+Succession of microbial consortia in the developing infant gut microbiome.
+Proc Natl Acad Sci {\bf 108} :4578–4585.
+\bibitem {tikhonov} {\bf  Tikhonov M.} 2016. Community-level cohesion without cooperation. Elife {\bf 5}.
 

### 2016-06-21 mSystems/mSys_discussion.tex
discussion in alpha version

--- 2016-04-21 mSystems/mSys_discussion.tex
@@
-We have quantitatively characterized whether the microbiota belongs to a healthy individual or a subject corresponding to an altered or pathological state (i.e., altered diet, antibiotic treatment, early gut development, diagnosed IBS).
-Deciphering the mechanisms of disease requires in depth knowledge of the underlying biological mechanisms.
-We describe here the macroscopic behavior of disease by a noise-induced phase transition with a control parameter that can be measured by the temporal variability of the microbiome.
-The microbiota of healthy individuals and of individuals with pathologies represent different phases separated by this noise-induced phase transition.
-Improved high-throughput sequencing of samples from individuals monitored over time and taxonomic assigning methods will provide a better distinction among pathologies or altered states of the microbiota.
+One of the mains features of this work is to have shown that the microbiota, independently of its condition, follows the Taylor's law.
+We have seen that the value of the scaling index in each case is always less than the unity (using standard deviation in the law), which is informing us about the community structure.
+This means that the most abundant elements in the population are less volatile to perturbations in relative terms than the less abundant.
+The explanation for this universal pattern is not clear although some hypothesis have been tested in other studies as the presence of negative interactions in the population \cite{kilpatrick}, while others have demonstrated that it may depend on reproductive correlation \cite{ballantyne}.
+Nevertheless, none of these explanations are enough when we are talking about microbiota as the reproduction term is diffuse, the interactions between its components are not only based on competition \cite{joao, mehta, bucci} and that kind of negative interaction may not effectively yield in values less than the unity when referring to a bacterial species \cite{ramslayer}.
+In any case, the values obtained in all cases are very similar between them, hence these theoretical questions may be addressed in future works.
+
+The second parameter is informing about the noise and can be directly related with the variability or the fluctuation amplitude of the population over time, and it is a direct estimator of the stability of the system under study.
+As we showed in above, the healthy part of each study have lower variability than the non-healthy part when dealing with adult individuals.
+Interestingly, the variability parameter was higher in the healthy part of the study of the discordant twins suffering from kwashiorkor disease \cite{kwashiorkor}.
+Taking into account that the infant microbiota is evolving toward a definite, adult state \cite{koenig}, it means that the temporal variability would be greater than in a adult who has reached a stability in his gut microbiota, while our results could be directing in the possibility that this variability is necessary in order to reach that adult state.
+Furthermore, as we wanted to see how this variability was over time, we calculated the evolution of this parameter for the samples which had enough time sampling.
+As can be seen in Figure \ref{fig:tempevo1}, the variability of the microbiota has some fluctuations over time.
+It is interesting to note in Figure \ref{fig:tempevo2} how this parameter can capture the two antibiotic intakes in one of the patients from the study of Dethlefsen and Relman \cite{antibiotic}, especially that it seems to be some resiliance process in the microbiota due to the lower variability increase in the second antibiotic intake.
+
+The primary hypothesis of this work is that having a healthy microbiota means, in adult individuals, that population is stable in time and does not have huge flips or jumps into another states.
+In order to use the valuable information which gives us the empirical law of Taylor's work, we propose the use of Langevin equation to model how the ranking stability evolves in time.
+While we can measure directly the component of the noise of the system as their variability, the other main term needs to be inferred from the model.
+This term, which we have named as 'fitness', is the one that gives the ability to the system to be stable to a possible perturbations.
+In ecological terms, this could mean the nature of interactions that are present between the bacteria, between bacteria and other minority populations as fungi or archaea, between bacteria and the viral component in the microbiota, and the interactions between host and the whole microbiota.
+Being this a first step to model the temporal stability of the microbiota and due to its complicated nature, we have calculated the fitness term using the fluctuation-dissipation theorem as a first approximation.
+Thus, the fitness of the microbiota still remains to be modeled in future works in order to make the model more accurate and with a higher predictive power.
+
+By solving the Langevin differential equation, we can obtain a phase diagram where each microbiota sample can be placed according to its fitness and variability in one of the two phases according to the ranking stability of the system.
+As we can see in the phase space in Figure \ref{fig:main3}, we are showing three different conditions that could happen.
+First, we can have a healthy microbiota which could have some fluctuations as showed by one of the subjects of Caporasso et al study \cite{moving}.
+Because the fitness of this cases will be high enough, the temporal variability will not place the microbiota in the unstable phase of the diagram.
+Second, we have a subject from the study of Dethlefsen and Relman \cite{antibiotic} which is perturbed twice by an antibiotic intake.
+His microbiota is altered enough to loose its stability and then be placed in the unstable part, being more sensible to a possible perturbations as, for example, opportunist infections.
+Third, the subject is already in the unstable phase due to some healthy issue as IBS, as can be observed in one of the patients from Durban et al study \cite{IBS}.
+It is shown also that this subject improved its healthy status in the time when the experiment was done, implying that his microbiota also recovered the lost stability.
+
+Specifically, the analysis of the rank stability of the samples of healthy and IBS diagnosed patients studied in our lab\cite{IBS}, suggests that the presence of a \emph{rank stability island} among medium-ranked taxa could be an indicator of a healthy microbiota.
+(to complete)
+
+But we have to be aware that the hypothesis above is too simplistic to be related with the reality.
+It has been demonstrated that the situation is more complex than to separate healthy people from non-healthy people by compositional terms only as Moya and Ferrer underlines in their review \cite{Moya_trends}.
+There are several different scenarios that can be possible in which we can consider the microbiota as stable independently of their compositional evolution over time, as for example in their ability to recover the initial composition (resiliance), or if it can recover the original function despite the composition (functional redundancy).
+What we have showed in this work could be explained as the transitions of a stable microbiota into a dysbiosis state.
+
+As a first step toward understanding the microbiota stability, the model present some limitations and there is still work to do.
+From the biological perspective, many questions arise from this work as for example which mechanisms are involved in maintaining the population structure.
+The nature of the interactions between the elements of the community is surely of great importance in this matter, and it is related to the fitness of the community as has been commented above.
+To capture these interactions, we could try to use the generalized Lotka-Volterra system of differential equations or other methods which are mathematically similar.
+There are also other ways in which we could infer the community fitness without taking into account the interactions as has been exposed by Tikhonov \cite{tikhonov}.
+There is also the possibility to work with the fluctuation-dissipation theorem and try to understand what could mean the 'temperature' in ecological terms.

### 2016-06-21 mSystems/mSys_intro.tex
deleted last sentence in intro
discussion in alpha version

@@
 Taylor's law has been applied to microbiota in a spatial way in the work of Zhang {\it et al.}, (2014) \cite{isme1}, where they show that this population tend to be in an aggregated way rather than in a random distribution.
-Despite its ubiquity, it has never been applied in follow-up studies from microbiota even that a great effort has been made to infer the community structure from a dynamical point of view \cite{cobas, schloss, ravel}
+Despite its ubiquity, it has been studied only in experimental settings \cite{cohen_bac, ramslayer} but never been applied in follow-up studies from microbiota even that a great effort has been made to infer the community structure from a dynamical point of view \cite{cobas, schloss, ravel}
 
@@
 We use this mathematical framework to explore the temporal stability of the microbiota in different conditions in order to understand how this affects the healthy status of the subjects.
-Finally, we have engineered a complete software framework, {\it ComplexCruncher}, to carry out the analysis of the temporal dynamics of microbiotas, which is ready to be implemented by other users.

### 2016-06-23 mSystems/mSys_discussion.tex
minor changes in intro (reference against holobiont), results (more text to time evolution of Taylors parameters), and discussion

@@
 One of the mains features of this work is to have shown that the microbiota, independently of its condition, follows the Taylor's law.
-We have seen that the value of the scaling index in each case is always less than the unity (using standard deviation in the law), which is informing us about the community structure.
+We have seen that the value of the scaling index in each case is always less than the unity (using standard deviation), which is informing us about the community structure.
 This means that the most abundant elements in the population are less volatile to perturbations in relative terms than the less abundant.
@@
 Nevertheless, none of these explanations are enough when we are talking about microbiota as the reproduction term is diffuse, the interactions between its components are not only based on competition \cite{joao, mehta, bucci} and that kind of negative interaction may not effectively yield in values less than the unity when referring to a bacterial species \cite{ramslayer}.
-In any case, the values obtained in all cases are very similar between them, hence these theoretical questions may be addressed in future works.
+In any case, the values obtained in all cases are very similar between them, what could be suggesting that the community structure is preserved in all the different scenarios that we have studied.
 
@@
 
-As a first step toward understanding the microbiota stability, the model present some limitations and there is still work to do.
-From the biological perspective, many questions arise from this work as for example which mechanisms are involved in maintaining the population structure.
+As a first step toward understanding the microbiota stability, the model presents some limitations and there is still work to do.
+From the biological perspective, many questions arise from this work.
+We have observed the same pattern in Taylor's parameters in all the different conditions we have studied, but a pertinent question is if it is really a universal feature in the huge diversity of microbial niches.
+Also, another relevant question is which mechanisms are involved in maintaining the population structure.
 The nature of the interactions between the elements of the community is surely of great importance in this matter, and it is related to the fitness of the community as has been commented above.
-To capture these interactions, we could try to use the generalized Lotka-Volterra system of differential equations or other methods which are mathematically similar.
-There are also other ways in which we could infer the community fitness without taking into account the interactions as has been exposed by Tikhonov \cite{tikhonov}.
-There is also the possibility to work with the fluctuation-dissipation theorem and try to understand what could mean the 'temperature' in ecological terms.
+How we should address the community fitness is not clear, but works as Tikhonov's \cite{tikhonov} could help us to aim in the correct direction toward unraveling the complexity of the microbiota.

### 2016-06-23 mSystems/mSys_intro.tex
minor changes in intro (reference against holobiont), results (more text to time evolution of Taylors parameters), and discussion

@@
 As evidence of new concepts 'holobiont' and 'hologenome' is increasing each day \cite{holo1, holo2}, research not only focus on the human physiology but also on the microbial population that surround ourselves.
+However, these concepts are still in debate \cite{holo3}.
 We are populated by a myriad of microorganisms that are interacting with us in several physiological processes such as metabolism of the bile acids \cite{bileacids}, of the choline \cite{choline} or key-route metabolites as short-chain fatty acids \cite{scfa1, scfa2} which are also involved in immune system maturation \cite{scfa3, scfa4}.
@@
 Biology has recently acquired new technological and conceptual tools to investigate, model and understand living organisms at the system level, thanks to the spectacular progress in quantitative techniques, large-scale measurement methods and the integration of experimental and computational approaches.
-In particular, Systems Biology has placed a great effort to unveil the general laws governing the complex behaviour of microbial communities \cite{sysbio&microb, msys1, metasysbio}.
+In particular, Systems Biology has placed a great effort to unveil the general laws governing the complex behaviour of microbial communities \cite{sysbio&microb, msys1, metasysbio}, even proposing that they have universal dynamics \cite{uni_dynam}.
 Microbiota can be approached under the light of ecological theory where we can find, for instance, general principles as the Taylor's law \cite{taylor}, which relates spatial or temporal variability of the population with its mean.

### 2016-06-23 mSystems/mSys_results.tex
minor changes in intro (reference against holobiont), results (more text to time evolution of Taylors parameters), and discussion

@@
 	\hspace*{-4.5mm}\includegraphics[width=0.83\textwidth]{results/corrank/IBS_P2_Metatranscriptores_Rank.pdf}
-	\caption{Matrixes showing the rank variation throughout time for the most dominant elements (taxa) and their calculated Rank Stability Index (as discussed in Material and Methods).
+	\caption{Matrixes showing the rank variation throughout time for the most dominant elements (taxa) and their calculated Rank Stability Index (as shown in Material and Methods).
 We show the matrix for samples from a healthy subject (top) and from a subject diagnosed with irritable bowel syndrome (bottom), studied in our lab\cite{IBS}.}
@@
 Figure \ref{fig:tempevo2} shows the time evolution of $V$ for patient P2 of the IBS study\cite{IBS} (upper plot) and patient D in the antibiotics study\cite{antibiotic} (lower plot).
+Both samples show changes in the variability V with quasi--periodic behavior peaked at about 10 days.
+Variability grows more for the gut microbiota of the male and share a minimal value around 0.1 with the gut microbiota of the female.
+The variability of the gut microbiota of P2 decreases from above 0.3 to below 0.2, showing a slow tendency to increase the order of the system.
+Antibiotic intake leaks to a quick increase of variability which lasts for a few days to recover ordering.
+The second antibiotic treatment shows some memory (lower increase of variability) with a slower recovery.
 
@@
 	\hspace*{3mm}\includegraphics[width=0.448\textwidth]{results/sliwin/female_mov.pdf}
-\caption{$V$ as a function of time for the two individuals in the Caporaso's study\cite{moving}: samples of gut microbiome of a male (upper plot) and a female (lower plot).
-Both samples show changes in the variability V with quasi--periodic behavior peaked at about 10 days.
-Variability grows more for the gut microbiota of the male and share a minimal value around 0.1 with the gut microbiota of the female.}
+\caption{$V$ as a function of time for the two individuals in the Caporaso's study\cite{moving}: samples of gut microbiome of a male (upper plot) and a female (lower plot).}
 \label{fig:tempevo1}
@@
   	\includegraphics[width=1.0\textwidth]{results/sliwin/patD_antibio.pdf} 
-\caption{$V$ as a function of time for patient P2 of the IBS study\cite{IBS} (upper plot) and patient D in the antibiotics study\cite{antibiotic} (lower plot). The variability 
-of the gut microbiota of P2 decreases from above 0.3 to below 0.2, showing a slow tendency to increase the order of the system.
-Antibiotic intake leaks to a quick increase of variability which lasts for a few days to recover ordering.
-The second antibiotic treatment shows some memory (lower increase of variability) with a slower recovery.
-NOTE: The blue vertical lines in the lower plot are showing the periods of antibiotic treatment.}
+\caption{$V$ as a function of time for patient P2 of the IBS study\cite{IBS} (upper plot) and patient D in the antibiotics study\cite{antibiotic} (lower plot).
+The blue vertical lines in the lower plot are showing the periods of antibiotic treatment.}
 \label{fig:tempevo2}

### 2016-06-23 mSystems/mSys_MAIN.tex
minor changes in intro (reference against holobiont), results (more text to time evolution of Taylors parameters), and discussion
bibliography fixed (finally!)

@@
 \author[1,3,4,5]{Amparo Latorre}
-\author[1,3,4,5]{Andrés Moya}
+\author[1,3,4,5, \#]{Andrés Moya}
 \author[1,2,\#]{Carlos P. Garay}
@@
 \footnote[0]{$^*$ Equally contributed}
-\footnote[0]{$^\#$ Corresponding author: penagaray@gmail.com}
+\footnote[0]{$^\#$ Corresponding authors: andres.moya@uv.es, penagaray@gmail.com}
 
@@
 \bibitem{holo2} {\bf Bordenstein SR, Theis KR.} 2015. Host Biology in Light of the Microbiome: Ten Principles of Holobionts and Hologenomes. PLOS Biol {\bf 13}:e1002226.
+\bibitem{holo3} {\bf Moran NA, Sloan DB.} 2015. The Hologenome Concept: Helpful or Hollow? PLoS Biol {\bf 13}:1–10.
 \bibitem{bileacids} {\bf Swann JR, Want EJ, Geier FM, Spagou K, Wilson ID, Sidaway JE, Nicholson JK, Holmes E.} 2011.
@@
 Proc Natl Acad Sci {\bf 109}:594–599.
+\bibitem {uni_dynam} {\bf Bashan A, Gibson TE, Friedman J, Carey VJ, Weiss ST, Hohmann EL, Liu Y-Y.} 2016. Universality of human microbial dynamics. Nature {\bf 534}:259–262.
 \bibitem{taylor} {\bf Taylor, L.R.} 1961. Aggregation, Variance and the mean. Nature {\bf 189,} 732-35.
@@
 \bibitem{ranking} {\bf Blumm N, Ghoshal G, Forró Z, Schich M, Bianconi G, Bouchaud J-P, Barabási A-L.} 2012. Dynamics of Ranking Processes in Complex Systems. Phys Rev Lett {\bf 109}:128701.
-\bibitem{moving} Caporaso, J.G. et al. Moving pictures of the human microbiome. {\it Genome Biol.} {\bf 12,} R50 (2011).
-\bibitem{LEA} Faith, J.J. et al. The long-term stability of the human gut microbiota. {\it Science} {\bf 341,} 1237439 (2013).
-\bibitem{kwashiorkor} Smith M.I. et al. Gut microbiomes of Malawian twin pairs discordant for kwashiorkor. {\it Science} {\bf 339,} 548-54 (2013).
-\bibitem{diet} David, L.A. et al. Diet rapidly and reproducibly alters the human gut microbiome. {\it Nature} {\bf 505,} 559-63 (2014).
-\bibitem{antibiotic} Dethlefsen L., Relman D.
-A.
+
+\bibitem{moving} {\bf Caporaso JG, Lauber CL, Costello EK, Berg-Lyons D, Gonzalez A, Stombaugh J, Knights D, Gajer P, Ravel J, Fierer N, Gordon JI, Knight R.} 2011.
+Moving pictures of the human microbiome.
+Genome Biol {\bf 12}:R50.
+\bibitem{LEA} {\bf Faith JJ, Guruge JL, Charbonneau M, Subramanian S, Seedorf H, Goodman AL, Clemente JC, Knight R, Heath AC, Leibel RL, Rosenbaum M, Gordon JI.} 2013.
+The long-term stability of the human gut microbiota.
+Science (80- ) {\bf 341}:1237439.
+\bibitem{kwashiorkor} {\bf Smith MI, Yatsunenko T, Manary MJ, Trehan I, Mkakosya R, Cheng J, Kau AL, Rich SS, Concannon P, Mychaleckyj JC, Liu J, Houpt E, Li J V, Holmes E, Nicholson J, Knights D, Ursell LK, Knight R, Gordon JI.} 2013.
+Gut microbiomes of Malawian twin pairs discordant for kwashiorkor.
+Science {\bf 339}:548–54.
+\bibitem{diet} {\bf David LA, Maurice CF, Carmody RN, Gootenberg DB, Button JE, Wolfe BE, Ling A V, Devlin AS, Varma Y, Fischbach MA, Biddinger SB, Dutton RJ, Turnbaugh PJ.} 2014.
+Diet rapidly and reproducibly alters the human gut microbiome.
+Nature {\bf 505}:559–63.
+\bibitem{antibiotic} {\bf Dethlefsen L, Relman DA.} 2011.
 Incomplete recovery and individualized responses of the human distal gut microbiota to repeated antibiotic perturbation.
-{\it Proc.
-Nat.
-Acad.
-Sci.
-USA} {\bf 108,} 4554-61 (2011).
-\bibitem{QIIME} Caporaso, J.G. et al. QIIME allows analysis of high-throughput community sequencing data. {\it Nature Methods} {\bf 7,} 335-6 (2010).
+Proc Natl Acad Sci {\bf108}:4554–61.
+\bibitem{QIIME} {\bf Caporaso JG, Kuczynski J, Stombaugh J, Bittinger K, Bushman FD, Costello EK, Fierer N, Peña AG, Goodrich JK, Gordon JI, Huttley G a, Kelley ST, Knights D, Koenig JE, Ley RE, Lozupone C a, Mcdonald D, Muegge BD, Pirrung M, Reeder J, Sevinsky JR, Turnbaugh PJ, Walters W a, Widmann J, Yatsunenko T, Zaneveld J, Knight R.} 2010.
+correspondence QIIME allows analysis of high- throughput community sequencing data Intensity normalization improves color calling in SOLiD sequencing.
+Nat Publ Gr {\bf 7}:335–336.
 \bibitem{LMAT} {\bf Ames SK, Hysom DA, Gardner SN, Lloyd GS, Gokhale MB, Allen JE}.
@@
 {\bf 25}:1056-67.
-\bibitem{fs} Eisler,Z., Bartos,I., Kertesz,J. Fluctuation scaling in complex systems: Taylor's law and beyond. {\it Adv. Phys.} {\bf 57,} 85 (2008).
-\bibitem{stat} Jorgensen,B., Martinez,J.R., Tsao,M. Asymptotic behaviour of the variance function. {\it Scand. J. Statist..} {\bf 21,} 223-243 (1994).
-\bibitem{convergence1} Fronczak,A., Fronczak,P. Origins of Taylor's power law for fluctuation scaling in complex systems. {\it Phys. Rev. E} {\bf 81,} 066112 (2010).
-\bibitem{convergence2} Kendal, W.S., Jorgensen,B. Taylor's power law and fluctuation scaling explained by a central-limit-like convergence. {\it Phys. Rev. E} {\bf 83,} 066115 (2011).
-\bibitem{convergence3} Kendal, W.S., Jorgensen,B. Tweedie convergence: A mathematical basis for Taylor's power law. {\it Phys. Rev. E} {\bf 84,} 066120 (2011).
+\bibitem{fs} {\bf Eisler Z, Bartos I, Kertész J.} 2008. Fluctuation scaling in complex systems: Taylor’s law and beyond1. Adv Phys {\bf 57}:89–142.
+\bibitem{stat} {\bf Jørgensen B, Martinez JR, Tsao M.} 1994. Asymptotic behaviour of the variance function. Scand J Stat {\bf 21}:223–243.
+
+\bibitem{convergence1} {\bf Fronczak A, Fronczak P.} 2010. Origins of Taylor’s power law for fluctuation scaling in complex systems. Phys Rev E {\bf 81}:066112.
+\bibitem{convergence2} {\bf Kendal, W.S., Jorgensen,B.} Taylor's power law and fluctuation scaling explained by a central-limit-like convergence.  Phys. Rev. E {\bf 83} 066115.
+\bibitem{convergence3} {\bf Kendal WS, Jørgensen B.} 2011. Tweedie convergence: A mathematical basis for Taylor’s power law, 1/f noise, and multifractality. Phys Rev E {\bf 84}:066120.
 
@@
 \end{thebibliography}
-\todo{Eliminar \textit{et al.} y poner la referencia completa como exige la guía de estilo de la revista...}
 

### 2016-06-23 mSystems/mSys_MAIN.tex
Several minor changes that leads to 0.3.0, last version before pre-releasing
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_discussion.tex

@@
 % release codenumber (update it here!)
-\newcommand\releasecode{0.2.0}
+\newcommand\releasecode{0.3.0}
 
@@
 % watermarks (time consuming!)
-\newwatermark[allpages,color=blue!4,angle=45,scale=9,xpos=-40,ypos=0]{DRAFT}
-\newwatermark[allpages,color=blue!12,angle=0,scale=1,xpos=0,ypos=137]{
+\newwatermark[allpages,color=PineGreen!4,angle=45,scale=9,xpos=-40,ypos=0]{DRAFT}
+\newwatermark[allpages,color=PineGreen!12,angle=0,scale=1,xpos=0,ypos=137]{
 \emph{mSystems} \LaTeX\ manuscript release \releasecode}
@@
 \author[1,3,4,5]{Amparo Latorre}
-\author[1,3,4,5, \#]{Andrés Moya}
+\author[1,3,4,5,\#]{Andrés Moya}
 \author[1,2,\#]{Carlos P. Garay}
@@
 Authors declare that there are no competing financial interests in relation to the work described here.
+We thereby express our acknowledgement to Bull/Atos and Micron Technology for providing us with the PCIe SSD card Micron P420m HHHL as a free-of-charge sample for high performance throughput database testing purposes.
 
@@
 Proc Natl Acad Sci {\bf 108} :4578–4585.
-\bibitem {tikhonov} {\bf  Tikhonov M.} 2016. Community-level cohesion without cooperation. Elife {\bf 5}.
-
+\bibitem {tikhonov} {\bf  Tikhonov M.} 2016.
+Community-level cohesion without cooperation.
+Elife {\bf 5}.
+\todo{Revisar referencias a partir de aquí: si proceden una vez que CC no se incluye y, en caso aformativo, adaptar el formato}
 	\bibitem{FD} Weber, J. \textit{et al.} Fluctuation dissipation theorem. {\it Phys. Rev.} {\bf 101}, 1620-6 (1956).

### 2016-06-23 mSystems/mSys_MandM.tex
Several minor changes that leads to 0.3.0, last version before pre-releasing
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_discussion.tex

@@
 \todo{Specify, in each study treated, the nature of the samples (conditions, timespan between timepoints, subjects).
-Specify, and it is very important, what we consider \emph{healthy} in each study (for example: pre-antibiotics is healthy)}
+Specify, and it is very important, what (and why) we consider \emph{healthy} in each study (for example: pre-antibiotics is healthy)}
 
@@
 LMAT was run using a Bull shared-memory node belonging to the team's HPC (high performance computing) cluster.
-It is equipped with 32 cores (64 threads available using Intel Hyper-threading technology) as it has 2 Haswell-based Xeons, the E5-2698v3@2.3 GHz, sharing half a tebibyte (0.5 TiB, that is, 512 gibibytes) of DRAM memory.
-This node is also provided with a card PCIe SSD as NVRAM, the P420m HHHL, with 1.4 TB, and 750000 reading IOPS, 4 KB, achieving 3.3 GB/s, which Micron kindly issued free of charge, as a sample for testing purposes.
+It is equipped with 32 cores (64 threads available using Intel Hyper-threading technology) as it has two Haswell-based Xeons ($22\un{nm}$ technology), the E5-2698v3@2.3 GHz, sharing half a tebibyte (0.5 TiB, that is, 512 gibibytes) of DRAM memory.
+This node is also provided with a PCIe SSD card as NVRAM, the Micron P420m HHHL, with 1.4 TB, and 750000 reading IOPS, 4 KB, achieving 3.3 GB/s.
 The computing node was supplied with a RAID-0 (striping) scratch disk area.

### 2016-06-23 mSystems/mSys_discussion.tex
Several minor changes that leads to 0.3.0, last version before pre-releasing
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_discussion.tex

--- 2016-06-16 mSystems/mSys_discussion.tex
@@
-The main feature of this work is to have shown that the microbiota, independently of its condition, follows the Taylor's law.
-We have seen that the value of beta in each case is always less than the unity (using standard deviation in the law), which is informing us about the community structure.
-This means that the most abundant elements in the population are less volatile to perturbations in relative terms than the less abundant.
-The biological explanation for this universal pattern is not clear.
-There are studies that prove that having this kind of exponents may be due to the negative interactions in the population \cite{kilpatrick}, while others have demonstrated that it may depend on reproductive correlation \cite{ballantyne}.
-Nevertheless, none of these explanations are enough when we are talking about microbiota as the reproduction term is diffuse, and the interactions between its components are numerous and complex \cite{joao, mehta}.
+One of the mains features of this work is to have shown that, independently of its condition, the microbiota follows the Taylor's law.
+We have seen that the value of the scaling index in each case is always less than the unity (using standard deviation as the measurement for dispersion), which is informing us about the community structure.
+This means that, in relative terms, the most abundant elements in the population are less volatile to perturbations than the less abundant ones.
+The explanation for this universal pattern is not clear although some hypothesis have been tested in other studies, as the presence of negative interactions in the population \cite{kilpatrick}, or the demonstration that it may depend on reproductive correlation \cite{ballantyne}.
+Nevertheless, none of these explanations are enough when we are talking about microbiota as the reproduction term is diffuse, the interactions between its components are not only based on competition \cite{joao, mehta, bucci} and that even that kind of negative interaction may not effectively yield in values less than the unity when referring to a bacterial species \cite{ramslayer}.
+Anyhow, the values obtained in all cases are very similar among them, which could be suggesting that the community structure is preserved throughout the different scenarios that we have studied.
 
-Model and temporal variation of V.
+The second parameter is informing about the noise and can be directly related with the variability or the fluctuation amplitude of the population over time.
+It is a direct estimator of the stability of the system under study.
+As we showed above, the healthy subset of each study have lower variability than the non-healthy subset when dealing with adult individuals.
+Interestingly, the variability parameter was higher in the healthy subset for the study of the discordant twins suffering from kwashiorkor disease \cite{kwashiorkor}.
+Taking into account that the infant microbiota is evolving toward a definite, adult state \cite{koenig}, it means that the temporal variability would be greater than in a adult who has reached a stability in his gut microbiota, while our results could be directing in the possibility that this variability is necessary in order to reach that adult state  \todo{<-review the wording of the whole sentence}.
+Furthermore, as we wanted to see how this variability was over time, we calculated the evolution of this parameter for the samples which had enough time sampling \todo{<-we should clarify which is and why we take that number as the minimum number of time samples}.
+As can be seen in Figure \ref{fig:tempevo1}, the variability of the microbiota has some fluctuations over time.
+It is interesting to note in Figure \ref{fig:tempevo2} how this parameter can capture the two antibiotic intakes in one of the patients from the study of Dethlefsen and Relman \cite{antibiotic}, especially that it seems to be some resilience process in the microbiota due to the lower variability increase in the second antibiotic intake.
 
-From the biological perspective, too many questions arise with this work.
-Knowing that microbiota has universal properties, one of the main relevant questions is which mechanisms are involved in maintaining the population structure in order to have always values of beta below 1...
+The primary hypothesis of this work is that, in adults, having a healthy microbiota means that population is stable in time and does not have huge flips or jumps into another states.
+In order to use the valuable information which gives us the empirical law of Taylor's work, we propose the use of Langevin equation to model how the ranking stability evolves in time.
+While we can measure directly the component of the noise of the system as their variability, the other main term needs to be inferred from the model.
+This term, which we have named as 'fitness', is the one that gives the ability to the system to be stable to potential perturbations.
+In ecological terms, this could mean the nature of interactions that are present among the bacteria, between bacteria and other minority populations as fungi or archaea, between bacteria and the viral component in the microbiota, and the interactions between host and the whole microbiota.
+Being this a first step to model the temporal stability of the microbiota and due to its complicated nature, we have calculated the fitness term using the fluctuation-dissipation theorem as a first approximation.
+Thus, the fitness of the microbiota still remains to be modelled in future works in order to make the model more accurate and with a higher predictive power.
 
-Specifically, the analysis of the rank stability of the samples of healthy and IBS diagnosed patients studied in our lab\cite{IBS}, suggests that the presence of a \emph{rank stability island} among medium-ranked taxa could be an indicator of a healthy microbiota.
+By solving the Langevin differential equation, we can obtain a phase diagram where each microbiota sample can be placed according to its fitness and variability in one of two phases according to the ranking stability of the system.
+As we can see in the phase-space in Figure \ref{fig:main3}, we are showing three different conditions that could happen.
+First, we can have a healthy microbiota which could have some fluctuations as showed by one of the subjects of Caporasso et al study \cite{moving}.
+Because the fitness of this cases will be high enough, the temporal variability will not place the microbiota in the unstable phase of the diagram.
+Second, we have a subject from the study of Dethlefsen and Relman \cite{antibiotic} which is perturbed twice by an antibiotic intake.
+His microbiota is altered enough to lose its stability, and hence be placed in the unstable part.
+So located, it is more sensible to potential perturbations as, for example, opportunist infections.
+Third and last condition, the subject is already in the unstable phase due to some healthy issue as IBS.
+This can be observed in one of the patients from Durban et al study \cite{IBS}.
+In addition, it was shown that this subject improved its healthy status in the time when the experiment was done, implying that his microbiota also recovered the lost stability.
 
-Final paragraph to talk about limitations and future perspectives: can we model stability in the functional landscape?
-Community assembly for itself doesn't explain everything, we need to move forward and think about what can be happening in complex ecosystems as the human microbiota.
+Specifically, the analysis of the rank stability of the samples of healthy and IBS diagnosed patients studied in our lab\cite{IBS}, suggests that the presence of what we called a \emph{rank stability island} among medium-ranked taxa could be an indicator of a healthy microbiota.
+\todo{<-review}
+
+However, we have to be aware that the hypothesis above is too simplistic to be directly related with reality.
+It has been demonstrated that the situation is more complex than the outlook provided separating healthy people from non-healthy people just by compositional terms, as Moya and Ferrer underline in their recent review \cite{Moya_trends}.
+There are several different feasible scenarios in which we can consider the microbiota as stable independently of their compositional evolution over time.
+For example, depending on their ability to recover the initial composition (resilience), or whether it can recover the original function despite the composition (functional redundancy).
+What we have shown in this work could be explained as the transitions of a stable microbiota into a dysbiosis state.
+
+As a first step toward understanding the microbiota stability, the model presents some limitations and there is still work to do.
+From the biological perspective, many questions arise from this work.
+We have observed the same pattern in Taylor's parameters in all the different conditions we have studied, but a pertinent question is whether it is really a universal feature in the huge diversity of microbial niches.
+Furthermore, another relevant question is which mechanisms are involved in maintaining the population structure.
+The nature of the interactions among the elements of the community is surely of great importance in this matter, and it is related to the fitness of the community as has been commented above.
+How we should address the community fitness is not clear, but works as Tikhonov's \cite{tikhonov} could help us to aim at the correct direction toward unraveling the complexity of the microbiota.

### 2016-07-06 mSystems/mSys_MAIN.tex
minor changes in supplementary: sample selection updated, grants added

@@
 \section*{Funding Information}
-This work is supported by Generalitat Valenciana Prometeo Grants II/2014/050, II/2014/065, by the Spanish Grants FPA2011-29678, BFU2012-39816-C02-01 of MINECO and by PITN-GA-2011-289442-INVISIBLES.
-JMM \& DMM acknowledge FPI and FISABIO fellowships.
+This work was supported by grants to AM from the Spanish Ministry of Science and Competitivity (projects SAF 2012-31187, SAF2013-49788-EXP, SAF2015-65878-R), Carlos III Institute of Health (projects PIE14/00045 and AC15/00022), Generalitat Valenciana (project PrometeoII/2014/065) and co-financed by FEDER, and grants to CPG from the Generalitat Valenciana Prometeo Grants II/2014/050, II/2014/065, 419 by the Spanish Grants FPA2011-29678, BFU2012-39816-C02-01 of MINECO and by PITN-GA-420 2011-289442-INVISIBLES.
+From the Ministry of Economy and Competitiveness (grant FPI BES-2013-062767).
+
  \todo{Modificar becas de JMM y DMM ¿poner algún grant más?}
@@
 
+% main studies
 \bibitem{moving} {\bf Caporaso JG, Lauber CL, Costello EK, Berg-Lyons D, Gonzalez A, Stombaugh J, Knights D, Gajer P, Ravel J, Fierer N, Gordon JI, Knight R.} 2011.
@@
 Proc Natl Acad Sci {\bf108}:4554–61.
+
 \bibitem{QIIME} {\bf Caporaso JG, Kuczynski J, Stombaugh J, Bittinger K, Bushman FD, Costello EK, Fierer N, Peña AG, Goodrich JK, Gordon JI, Huttley G a, Kelley ST, Knights D, Koenig JE, Ley RE, Lozupone C a, Mcdonald D, Muegge BD, Pirrung M, Reeder J, Sevinsky JR, Turnbaugh PJ, Walters W a, Widmann J, Yatsunenko T, Zaneveld J, Knight R.} 2010.

### 2016-07-06 mSystems/mSys_MandM.tex
minor changes in supplementary: sample selection updated, grants added

@@
 
-\todo{Specify, in each study treated, the nature of the samples (conditions, timespan between timepoints, subjects).
-Specify, and it is very important, what (and why) we consider \emph{healthy} in each study (for example: pre-antibiotics is healthy)}
-
 %%
@@
 We have selected only those individuals who had three or more time points of data available in databases.
+Caporaso et al.
+study \cite{moving} was selected as it has two healthy individuals measured over a very large timespan, with almost daily sampling.
+Faith et al.
+study \cite{LEA} was selected due to the BMI differences between subjects, moreover some of them had diets which can be treated as system perturbations.
+We considered healthy only those individuals who had normal or overweight BMI.
+Smith et al., study \cite{kwashiorkor} was selected for both the age of the patients and the rare disease.
+We only worked with the discordant twins, and considered healthy those who were not affected by kwashiorkor in each pair of patients.
+David et al., study \cite{diet} was selected for its differential diets.
+We considered as the healthy part the first time samples of each individual before the diet, and the rest of time points as perturbations.
+Dethlefsen and Relman work \cite{antibiotics} was selected due to the interesting treatment of two antibiotic intakes of the same antibiotic by three different subjects.
+We considered as the healthy part only those times before any antibiotic treatment, and as perturbations the time of antibiotic intakes and the period after that.
+And finally, we also considered a study made in our group carried by Durbán et al., \cite{IBS} in which we considered as healthy subjects those who didn't suffer from irritable bowel disease, and as perturbation the patients who had this disease.
+
 Metadata of each study is provided in Tables \ref{tab:diet} to \ref{tab:LEA}.

### 2016-07-21 mSystems/mSys_MandM.tex
release version 0.4! minor changes in almost every section
Andrés, estás ahora como otro corresponding author.
Intro: cambios menores
Resultados: reestructuración de algunas partes. Hemos eliminado los "as discussed in supplementary..." de varias de las partes. Subida la Tabla 1 a su sección. Ampliada la sección de análisis temporal de V y beta.
Discusión: inglés algo más corregido (gracias sobre todo a Jose Manuel). Final menos abrupto y sin tantos detalles innecesarios. Ampliada el párrafo que habla sobre estas "islas de estabilidad" que tan hábilmente encontró JM (a mí me parecen un resulta

@@
 We considered as the healthy part only those times before any antibiotic treatment, and as perturbations the time of antibiotic intakes and the period after that.
-And finally, we also considered a study made in our group carried by Durbán et al., \cite{IBS} in which we considered as healthy subjects those who didn't suffer from irritable bowel disease, and as perturbation the patients who had this disease.
+And finally, we also considered a study made in our group carried by Durban et al., \cite{IBS} in which we considered as healthy subjects those who didn't suffer from irritable bowel disease, and as perturbation the patients who had this disease.
 
@@
 
-The colour code of the RSI percentage text in the rank plot shown in Figure \ref{fig:corrank} is chosen following the first condition satisfied from those shown in Table \ref{tab:RSI}.
-
-\begin{table}
-  \begin{center}
-    \begin{tabular}{cccr}
-    \hline
-    Case  &  Condition  &  Colour  &  Description \\
-    \hline
-    1  &  $1\ge{\rm RSI}>0.99$  & \textcolor{blue}{blue} & constant rank \\ 
-    2  & ${\rm RSI}>0.90$  &  \textcolor{green}{green}  & highly stable rank \\
-    3  &  ${\rm RSI}>0.75$  &  \textcolor{orange}{orange} & moderately stable rank \\
-    4  &  ${\rm RSI}>0.25$  &  \textcolor{red}{red} & unstable rank \\
-    5  &  $0.25\ge{\rm RSI}\ge0$  &  \bfseries{black} & very unstable rank \\
-    \hline
-    \end{tabular}
-  \end{center} 
-  \caption{Colour code of the RSI percentage text shown in Figure \ref{fig:corrank}, following the first condition satisfied.}
-  \label{tab:RSI}
-\end{table}
-
 %%

### 2016-07-21 mSystems/mSys_discussion.tex
release version 0.4! minor changes in almost every section
Andrés, estás ahora como otro corresponding author.
Intro: cambios menores
Resultados: reestructuración de algunas partes. Hemos eliminado los "as discussed in supplementary..." de varias de las partes. Subida la Tabla 1 a su sección. Ampliada la sección de análisis temporal de V y beta.
Discusión: inglés algo más corregido (gracias sobre todo a Jose Manuel). Final menos abrupto y sin tantos detalles innecesarios. Ampliada el párrafo que habla sobre estas "islas de estabilidad" que tan hábilmente encontró JM (a mí me parecen un resulta

@@
 Taking into account that the infant microbiota is evolving toward a definite, adult state \cite{koenig}, it means that the temporal variability would be greater than in a adult who has reached a stability in his gut microbiota, while our results could be directing in the possibility that this variability is necessary in order to reach that adult state  \todo{<-review the wording of the whole sentence}.
-Furthermore, as we wanted to see how this variability was over time, we calculated the evolution of this parameter for the samples which had enough time sampling \todo{<-we should clarify which is and why we take that number as the minimum number of time samples}.
+Furthermore, as we wanted to see how this variability behaved over time, we calculated the evolution of this parameter for the samples which had enough time sampling.
 As can be seen in Figure \ref{fig:tempevo1}, the variability of the microbiota has some fluctuations over time.
@@
 
-Specifically, the analysis of the rank stability of the samples of healthy and IBS diagnosed patients studied in our lab\cite{IBS}, suggests that the presence of what we called a \emph{rank stability island} among medium-ranked taxa could be an indicator of a healthy microbiota.
-\todo{<-review}
+Specifically, the analysis of the rank stability of the samples of healthy and IBS diagnosed patients studied in our lab\cite{IBS}, suggests that the presence of \emph{rank stability islands} among medium-ranked taxa is a interesting feature.
+The higher stability of these taxa goes against the global meaning of the scaling index.
+Interestingly, that stability disappears when we look at the IBS patients.
+One could ask if these taxa are key players in the phase transition of the microbiota, of if they are more susceptible to perturbations than the most abundant.
+The types of interactions that could be sustaining this particular behavior are not clear, as these non-abundant taxa are not usually included in dynamical studies in order to get the community matrix.
+Further experiments and data analysis is needed to clarify if this is not a unique event, or it is a widespread feature of stable microbiotas.
 

### 2016-07-21 mSystems/mSys_results.tex
release version 0.4! minor changes in almost every section
Andrés, estás ahora como otro corresponding author.
Intro: cambios menores
Resultados: reestructuración de algunas partes. Hemos eliminado los "as discussed in supplementary..." de varias de las partes. Subida la Tabla 1 a su sección. Ampliada la sección de análisis temporal de V y beta.
Discusión: inglés algo más corregido (gracias sobre todo a Jose Manuel). Final menos abrupto y sin tantos detalles innecesarios. Ampliada el párrafo que habla sobre estas "islas de estabilidad" que tan hábilmente encontró JM (a mí me parecen un resulta

@@
-We have analysed the microbiome temporal variability to extract global properties of the system.
+We have analyzed the microbiome temporal variability to extract global properties of the system.
 As fluctuations in total counts are plagued by systematic errors we worked on temporal variability of relative abundances for each taxon.
 Our first finding was that, in all cases, changes in relative abundances of taxa follow a ubiquitous pattern known as the fluctuation scaling law\cite{fs} or Taylor's power law\cite{taylor}, i.e., microbiota of all detected taxa follows $\sigma_i  = V\cdot x_i^{\beta}$, a power law dependence between mean relative abundance $x_i$ and dispersion $\sigma_i$.
-The law seem to be ubiquitous, spanning even to six orders of magnitude in the observed relative abundances (see Figure \ref{fig:main1}).
+The law seem to be ubiquitous, spanning even to six orders of magnitude in the observed relative abundances.
+As can be seen in Figure \ref{fig:main1}, the most abundant species are less volatile in relative terms than the less abundant.
+The fitting to the power law is always robust ($R^{2}$ > 0.88) and does not depend on the microbiome condition.
+The power law (or scaling) index $\beta$ and the variability $V$ (hereafter Taylor parameters) appear to be correlated with the stability of the community and related with the health status of the host, which we consider the main finding exposed in this article.
 
@@
 
-The power law (or scaling) index $\beta$ and the variability $V$ (hereafter Taylor parameters) appear to be correlated with the stability of the community and related with the health status of the host, which we consider the main finding exposed in this article (see Figure \ref{fig:main2}).
+Taylor parameters describing the temporal variability of the gut microbiome in our sampled individuals are shown in Tables S1 to S6.
+Our results hint at an ubiquitous behavior.
+On the first hand, the variability (which corresponds to the maximum amplitude of fluctuations) is large, which suggests resilient capacity of the microbiota.
+On the other hand, the scaling index is always smaller than one, which means that more abundant taxa are less volatile than less abundant ones.
+In addition, Taylor parameters for the microbiome of healthy individuals in different studies are compatible within estimated errors.
+This enables us to define an area in the Taylor parameter space that we called the \emph{healthy zone}.
+% should be changed for any other name: stable zone?
+
+In order to jointly visualize and compare the results of individuals from different studies, their Taylor parameters have been standardized, where standardization means that each parameter is subtracted by the mean value and divided by the standard deviation of the group of healthy individuals for each study (for details of the procedure, please see Standardization subsection in Material and Methods).
+The healthy zone and the standardized Taylor parameters for individuals whose gut microbiota is altered (i.e., suffering from kwashiorkor, altered diet, antibiotics or IBS) is shown in Figure \ref{fig:main2}.
+Children developing kwashiorkor show smaller variability than their healthy twins.
+A meat/fish-based diet increases the variability significantly when compared to a plant-based diet.
+All other cases presented increased variability, which is particularly severe, and statistically significant at more than 95\% CL, for obese patients grade III on a diet, individuals taking antibiotics or IBS--diagnosed patients.
+A global property emerges from all worldwide data collected: Taylor parameters characterize the statistical behavior of microbiome changes.
+Furthermore, we have verified that our conclusions are robust to systematic errors due to taxonomic assignment (see Figure Sx in Supplemental Material).
 
@@
 \end{figure}
-
-Taylor parameters describing the temporal variability of the gut microbiome in our sampled individuals are shown in Tables S1 to S6.
-Our results hint at an ubiquitous behaviour.
-On the first hand, the variability (which corresponds to the maximum amplitude of fluctuations) is large, which suggests resilient capacity of the microbiota.
-On the other hand, the scaling index is always smaller than one, which means that more abundant taxa are less volatile than less abundant ones.
-In addition, Taylor parameters for the microbiome of healthy individuals in different studies are compatible within estimated errors.
-This enables us to define an area in the Taylor parameter space that we called the \emph{healthy zone}.
-
-In order to jointly visualize and compare the results of individuals from different studies, their Taylor parameters have been standardized, where standardization means that each parameter is subtracted by the mean value and divided by the standard deviation of the group of healthy individuals for each study (for details of the procedure, please see Standardization subsection in Material and Methods).
-The healthy zone and the standardized Taylor parameters for individuals whose gut microbiota is altered (i.e., suffering from kwashiorkor, altered diet, antibiotics or IBS) is shown in Figure \ref{fig:main2}.
-Children developing kwashiorkor show smaller variability than their healthy twins.
-A meat/fish-based diet increases the variability significantly when compared to a plant-based diet.
-All other cases presented increased variability, which is particularly severe, and statistically significant at more than 95\% CL, for obese patients grade III on a diet, individuals taking antibiotics or IBS--diagnosed patients.
-A global property emerges from all worldwide data collected: Taylor parameters characterize the statistical behavior of microbiome changes.
-Furthermore, we have verified that our conclusions are robust to systematic errors due to taxonomic assignment (see Figure Sx in Supplemental Material).
 
@@
 
-The model predicts two phases for the gut microbiome: a stable phase with large variability that permits some changes in the relative abundances of taxa and an unstable phase with larger variability, above the phase transition, where the order of abundant taxa varies significantly with time.
+The model predicts two phases for the gut microbiome: a stable phase with large variability that permits some changes in the relative abundances of taxa; and an unstable phase with larger variability, above the phase transition, where the order of abundant taxa varies significantly with time.
 The microbiome of all healthy individuals was found to be in the stable phase, while the microbiome of several other individuals was shown to be in the unstable phase.
@@
 
+\begin{table}
+  \begin{center}
+    \begin{tabular}{cccr}
+    \hline
+    Case  &  Condition  &  Colour  &  Description \\
+    \hline
+    1  &  $1\ge{\rm RSI}>0.99$  & \textcolor{blue}{blue} & constant rank \\ 
+    2  & ${\rm RSI}>0.90$  &  \textcolor{green}{green}  & highly stable rank \\
+    3  &  ${\rm RSI}>0.75$  &  \textcolor{orange}{orange} & moderately stable rank \\
+    4  &  ${\rm RSI}>0.25$  &  \textcolor{red}{red} & unstable rank \\
+    5  &  $0.25\ge{\rm RSI}\ge0$  &  \bfseries{black} & very unstable rank \\
+    \hline
+    \end{tabular}
+  \end{center} 
+  \caption{Colour code of the RSI percentage text following the first condition satisfied.}
+  \label{tab:RSI}
+\end{table}
+
 In the IBS diagnosed patient, beyond the differences in dominance for the particular taxa, we still observe that the most dominant are the most rank stable.
@@
 Finally, we have studied the time dependence of the variability $V$ and power law index $\beta$ (see Model under Material and Methods) by using a sliding window approach.
-The total number of time points are divided in subsets of five points, where next subset is defined by adding next time sampling and by eliminating the earliest one.
+The total number of time points are divided in subsets of five points, where the next subset is defined by adding the next time sampling and by eliminating the earliest one.
 Both parameters were calculated for each subset against the average time lapse.
@@
 Figure \ref{fig:tempevo2} shows the time evolution of $V$ for patient P2 of the IBS study\cite{IBS} (upper plot) and patient D in the antibiotics study\cite{antibiotic} (lower plot).
-Both samples show changes in the variability V with quasi--periodic behavior peaked at about 10 days.
+Both samples shows changes in the variability V with quasi--periodic behavior peaked at about 10 days.
 Variability grows more for the gut microbiota of the male and share a minimal value around 0.1 with the gut microbiota of the female.

### 2016-07-26 mSystems/mSys_MAIN.tex
release version 0.4! minor changes in almost every section
Andrés, estás ahora como otro corresponding author.
Intro: cambios menores
Resultados: reestructuración de algunas partes. Hemos eliminado los "as discussed in supplementary..." de varias de las partes. Subida la Tabla 1 a su sección. Ampliada la sección de análisis temporal de V y beta.
Discusión: inglés algo más corregido (gracias sobre todo a Jose Manuel). Final menos abrupto y sin tantos detalles innecesarios. Ampliada el párrafo que habla sobre estas "islas de estabilidad" que tan hábilmente encontró JM (a mí me parecen un resulta

@@
 % release codenumber (update it here!)
-\newcommand\releasecode{0.3.0}
+\newcommand\releasecode{0.5.0}
 
@@
 % watermarks (time consuming!)
-\newwatermark[allpages,color=PineGreen!4,angle=45,scale=9,xpos=-40,ypos=0]{DRAFT}
-\newwatermark[allpages,color=PineGreen!12,angle=0,scale=1,xpos=0,ypos=137]{
+\newwatermark[allpages,color=Fuchsia!4,angle=45,scale=9,xpos=-40,ypos=0]{DRAFT}
+\newwatermark[allpages,color=Fuchsia!12,angle=0,scale=1,xpos=0,ypos=137]{
 \emph{mSystems} \LaTeX\ manuscript release \releasecode}
@@
 Increasing research activity is dedicated to understand how changes in composition and function of the microbiota are associated to disease or not.
-We analyze 16S rRNA and whole genome sequencing (WGS) published data from the gut microbiota of 97 individuals monitored in time.
+We analyze 16S rRNA and shotgun metagenomic sequencing (SMS) published data from the gut microbiota of 97 individuals monitored in time.
 Temporal fluctuations in the microbial composition reveal significant differences due to factors such us dietary changes, antibiotic intake, age or disease.
@@
 \input{mSys_intro.tex}
-\clearpage
+%\clearpage
 %%% Result
@@
 \input{mSys_results.tex}
-\clearpage
+%\clearpage
 %%% Discussion
@@
 \input{mSys_discussion.tex}
-\clearpage    
+%\clearpage    
 %%% Material & Methods
@@
 \input{mSys_MandM.tex}   
-\clearpage
+%\clearpage
 
@@
 \bibitem{Moya_trends} {\bf Moya A, Ferrer M.} 2016. Functional Redundancy-Induced Stability of Gut Microbiota Subjected to Disturbance. Trends Microbiol {\bf 24}:402–413.
+\bibitem{CFS} {\bf Giloteaux L, Goodrich JK, Walters WA, Levine SM, Ley RE, Hanson MR}. 2016.
+Reduced diversity and altered composition of the gut microbiome in individuals with myalgic encephalomyelitis/chronic fatigue syndrome. Microbiome. {\bf 4}:30
+
 \bibitem{microb&health} {\bf Marchesi JR, Adams DH, Fava F, Hermes GD a, Hirschfield GM, Hold G, Quraishi MN, Kinross J, Smidt H, Tuohy KM, Thomas L V, Zoetendal EG, Hart A.} 2015.
@@
 2012.
-Bacterial microcosms obey Taylor’s law: Effects of abiotic and biotic stress and genetics on mean and variance of population density.
+Bacterial microcosmos obey Taylor’s law: Effects of abiotic and biotic stress and genetics on mean and variance of population density.
 Ecol Process {\bf 1}:5.
@@
 
+\bibitem{FASTX} {\bf Gordon, A., Hannon, G.J.} 2010. FASTX-Toolkit. FASTQ/A shortreads pre-processing tools v0.0.13. http://hannonlab.cshl.edu/fastx\_toolkit/ (last accessed 26 Jul 2016).
 \bibitem{QIIME} {\bf Caporaso JG, Kuczynski J, Stombaugh J, Bittinger K, Bushman FD, Costello EK, Fierer N, Peña AG, Goodrich JK, Gordon JI, Huttley G a, Kelley ST, Knights D, Koenig JE, Ley RE, Lozupone C a, Mcdonald D, Muegge BD, Pirrung M, Reeder J, Sevinsky JR, Turnbaugh PJ, Walters W a, Widmann J, Yatsunenko T, Zaneveld J, Knight R.} 2010.
@@
 Nat Publ Gr {\bf 7}:335–336.
+\bibitem{SILVA} {\bf Quast C, Pruesse E, Yilmaz P, Gerken J, Schweer T, Yarza P, Peplies J, Glöckner FO}.
+2013.
+The SILVA ribosomal RNA gene database project: improved data processing and web-based tools.
+Acids Res.
+{\bf 41}:D590-D596
 \bibitem{LMAT} {\bf Ames SK, Hysom DA, Gardner SN, Lloyd GS, Gokhale MB, Allen JE}.
@@
 Succession of microbial consortia in the developing infant gut microbiome.
-Proc Natl Acad Sci {\bf 108} :4578–4585.
-\bibitem {tikhonov} {\bf  Tikhonov M.} 2016.
-Community-level cohesion without cooperation.
-Elife {\bf 5}.
-\todo{Revisar referencias a partir de aquí: si proceden una vez que CC no se incluye y, en caso aformativo, adaptar el formato}
-	\bibitem{FD} Weber, J. \textit{et al.} Fluctuation dissipation theorem. {\it Phys. Rev.} {\bf 101}, 1620-6 (1956).
-	\bibitem{FASTX} Gordon, A., Hannon, G.J. FASTX-Toolkit. FASTQ/A shortreads pre-processing tools (2010). http://hannonlab.cshl.edu/fastx\_toolkit/ (accessed 23 Feb 2015).
-	\bibitem{SILVA} Quast C.  \textit{et al.} The SILVA ribosomal RNA gene database project: improved data processing and web-based tools (2013)
- 	\bibitem{ecology} Xiao Xiao, Ethan P.
-White, Mevin B.
-Hooten, and Susan L.
-Durham.
-On the use of log-transformation vs.
-nonlinear regression for analyzing biological power laws.
-{\it Ecology} {\bf 92}, 10, 1887-1894 (2011).
- 	\bibitem{genR2} Magee L., $R^2$ measures based on wald and likelihood ratio joint significance tests. {\it The American Statistician} {\bf 44}, 3, 250-253 (1990).
-	\bibitem{disR2} Nagelkerke N.J.D., A note on a general definition of the coefficient of determination. {\it Biometrika} {\bf 78}, 3, 691-692 (1991).
-	\bibitem{boot} Wu, C.F.J. Jackknife, bootstrap and other resampling methods in regression analysis. (with discussions) \textit{The Annals of Statistics} {\bf 14}: 1261?1350 (1986)
+Proc Natl Acad Sci {\bf 108}:4578–4585.
+\bibitem {FD} {\bf Weber, J.} 1956. Fluctuation Dissipation Theorem. Phys. Rev. {\bf 101}:1620-6
+\bibitem {tikhonov} {\bf  Tikhonov M.} 2016. Community-level cohesion without cooperation. Elife {\bf 5}.
+
+%M&M
+\bibitem{biopwrlaw} {\bf Xiao X, White EP, Hooten MB, Durham SL}. 2011. On the use of log-transformation vs. nonlinear regression for analyzing biological power laws. Ecology {\bf 92}:1887-1894
+\bibitem{boot} {\bf Wu, C.F.J.} 1986. Jackknife, bootstrap and other resampling methods in regression analysis. (with discussions) \textit{The Annals of Statistics} {\bf 14}:1261-1350 
 \end{thebibliography}

### 2016-07-26 mSystems/mSys_MandM.tex
Several minor changes leading to 0.5.0, another step towards pre-releasing
On branch master changes to be committed:
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_discussion.tex
	modified:   mSys_intro.tex
	modified:   mSys_results.tex

@@
 \begin{linenomath}
-\begin{equation}
+\begin{equation*}
 \dot{x_i} = F_i \cdot x_i^\alpha + V \cdot x_i^\beta \xi_i(t) - \phi(t) \cdot x_i,
-\end{equation}
+\end{equation*}
 \end{linenomath}
@@
 \begin{linenomath}
-\begin{equation}
+\begin{equation*}
 \frac{\partial P}{\partial t} = - \frac{\partial}{\partial x_i}  [(F_i \cdot x_i^\alpha - \phi(t) \cdot x_i ) \cdot P]+ \frac{1}{2} \frac{\partial^2}{\partial x_i^2} (V^2 \cdot x_i^{2\beta}\cdot P).
-\end{equation}
+\end{equation*}
 \end{linenomath}
@@
 In many physical systems (Brownian motion is the classical example), the two terms of the Langevin equation are related.
-The \emph{fluctuation--dissipation theorem} states a general relationship between the response to an external disturbance and the internal fluctuations of the system\cite{FD}.
+The \emph{Fluctuation Dissipation Theorem} states a general relationship between the response to an external disturbance and the internal fluctuations of the system\cite{FD}.
 The theorem can be used as the basic formula to derive the fitness from the analysis of fluctuations
@@
 
-\todo{Explain better the fluctuation-dissipation theorem}
+\todo{Explain better the Fluctuation Dissipation Theorem}
 
@@
 We considered as the healthy part the first time samples of each individual before the diet, and the rest of time points as perturbations.
-Dethlefsen and Relman work \cite{antibiotics} was selected due to the interesting treatment of two antibiotic intakes of the same antibiotic by three different subjects.
+Dethlefsen and Relman work \cite{antibiotic} was selected due to the interesting treatment of two antibiotic intakes of the same antibiotic by three different subjects.
 We considered as the healthy part only those times before any antibiotic treatment, and as perturbations the time of antibiotic intakes and the period after that.
@@
 
-Metadata of each study is provided in Tables \ref{tab:diet} to \ref{tab:LEA}.
-All used 16S rRNA gene sequencing except for the study of the discordant kwashiorkor twins\cite{kwashiorkor} (see Tables \ref{tab:DH} and \ref{tab:DK}) where shotgun metagenomic sequencing (SMS) and 16S rRNA were used.
+Metadata of each study is provided in Supplementary Tables S1 to S4.
+All used 16S rRNA gene sequencing except for the study of the discordant kwashiorkor twins\cite{kwashiorkor}, where both shotgun metagenomic sequencing (SMS) and 16S rRNA were used.
 In the latter case we selected to work with SMS data to show that our method is valid regardless of the source of taxonomic information.
@@
 \subsubsection*{Metagenomic sequences processing}
-Metagenomic shotgun (and 16S too) sequences were analyzed with LMAT (Livermore Metagenomics Analysis Toolkit) software package\cite{LMAT} (version 1.2.4, with Feb'15 release of data base \emph{LMAT-Grand}).
+Shotgun metagenomic (and 16S too) sequences were analyzed with LMAT (Livermore Metagenomics Analysis Toolkit) software package\cite{LMAT} (version 1.2.4, with Feb'15 release of data base \emph{LMAT-Grand}).
 LMAT was run using a Bull shared-memory node belonging to the team's HPC (high performance computing) cluster.
@@
 
-The basic idea of bootstrapping is that inference about a population from sample data (sample $\rightarrow$ population) can be modeled by resampling the sample data and performing inference on (resample $\rightarrow$ sample).
+The basic idea of bootstrapping is that inference about a population from sample data (sample $\rightarrow$ population) can be modeled by resampling the sample data and performing inference on (resample $\rightarrow$ sample)\cite{boot}.
 To adapt this general idea to our problem, we resample the x-data array using its errors array.
@@
 where $v_i$ is a Gaussian random variable with mean $\mu_i=0$ and standard deviation $\sigma_i=\mathrm{SEM}_i$, as defined previously.
-For each replicate a complete un-weighted power-law fit is performed, where to choose between fitting power laws ($y=Vx^\beta$) using linear regression on log-transformed (LLR) data versus non-linear regression (NLR) we mainly follow \emph{General Guidelines for the Analysis of Biological Power Laws}\cite{ecology}.
+For each replicate a complete un-weighted power-law fit is performed, where to choose between fitting power laws ($y=Vx^\beta$) using linear regression on log-transformed (LLR) data versus non-linear regression (NLR) we mainly follow \emph{General Guidelines for the Analysis of Biological Power Laws}\cite{biopwrlaw}.
 The parameters of the X-weighted fit are then estimated by averaging through all the replicate fits performed, and their errors are estimated by computing the standard deviation also for all the fits.
@@
 
-The Rank Stability Index (RSI) is shown as a percentage in a separate bar on the right of the rank matrix plot shown in Figure \ref{fig:corrank}.
+The Rank Stability Index (RSI) is shown as a percentage in a separate bar on the right of the rank matrix plot shown in Figures \ref{fig:corrank_IBS_A} and \ref{fig:corrank_IBS_P2}.
 The RSI is strictly $1$ for an element whose range never changes over time, and is strictly $0$ for an element whose rank oscillates between the extremes from time to time.

### 2016-07-26 mSystems/mSys_discussion.tex
Several minor changes leading to 0.5.0, another step towards pre-releasing
On branch master changes to be committed:
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_discussion.tex
	modified:   mSys_intro.tex
	modified:   mSys_results.tex

@@
 In ecological terms, this could mean the nature of interactions that are present among the bacteria, between bacteria and other minority populations as fungi or archaea, between bacteria and the viral component in the microbiota, and the interactions between host and the whole microbiota.
-Being this a first step to model the temporal stability of the microbiota and due to its complicated nature, we have calculated the fitness term using the fluctuation-dissipation theorem as a first approximation.
+Being this a first step to model the temporal stability of the microbiota and due to its complicated nature, we have calculated the fitness term using the Fluctuation Dissipation Theorem as a first approximation\cite{FD}.
 Thus, the fitness of the microbiota still remains to be modelled in future works in order to make the model more accurate and with a higher predictive power.

### 2016-07-26 mSystems/mSys_results.tex
Several minor changes leading to 0.5.0, another step towards pre-releasing
On branch master changes to be committed:
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_discussion.tex
	modified:   mSys_intro.tex
	modified:   mSys_results.tex

@@
 \subsection*{Rank stability of the taxa}
-The rank dynamics and stability plot in Figure \ref{fig:corrank} shows the variation in the rank with time for the most dominant taxa and their calculated Rank Stability Index (RSI, as discussed in Material and Methods) for the taxa of a healthy subject (individual \emph{A}, top) and from a subject diagnosed with IBS (patient \emph{P2}, bottom) of the IBS study\cite{IBS}.
+
+\def \RSItable {
+\scriptsize
+	\begin{tabular}{cccr}
+	\multicolumn{4}{c}{Colour code for the RSI percentage column} \\
+    \hline
+    Case  &  Condition  &  Colour  &  Description \\
+    \hline
+    1  &  $1\ge{\rm RSI}>0.99$  & \textcolor{blue}{blue} & constant rank \\ 
+    2  &  $0.99\ge{\rm RSI}>0.90$  &  \textcolor{green}{green}  & highly stable rank \\
+    3  &  $0.90\ge{\rm RSI}>0.75$  &  \textcolor{orange}{orange} & moderately stable rank \\
+    4  &  $0.75\ge{\rm RSI}>0.25$  &  \textcolor{red}{red} & unstable rank \\
+    5  &  $0.25\ge{\rm RSI}\ge0$  &  \bfseries{black} & very unstable rank \\
+    \hline
+    \end{tabular}	
+}
+
+The rank dynamics and stability plots in Figure \ref{fig:corrank_IBS_A} and \ref{fig:corrank_IBS_P2} show the variation in the rank with time for the most dominant taxa and their calculated Rank Stability Index (RSI, as discussed in Material and Methods) for the taxa of a healthy subject (individual \emph{A}, Figure \ref{fig:corrank_IBS_A}) and from a subject diagnosed with IBS (patient \emph{P2}, Figure \ref{fig:corrank_IBS_P2}) of the IBS study\cite{IBS}.
 The taxa are listed ordered by the accumulated frequency along the time series, so y-axis is an overall dominance axis for each sample set.
@@
 	\centering
-	\vspace*{-10mm} % Corrects overbox of the figures
-	\includegraphics[width=0.8\textwidth]{results/corrank/IBS_h_A_amplicons_family_Rank.pdf}\vspace*{-3mm} % Corrects overbox of the figures
-	\hspace*{-4.5mm}\includegraphics[width=0.83\textwidth]{results/corrank/IBS_P2_Metatranscriptores_Rank.pdf}
-	\caption{Matrixes showing the rank variation throughout time for the most dominant elements (taxa) and their calculated Rank Stability Index (as shown in Material and Methods).
-We show the matrix for samples from a healthy subject (top) and from a subject diagnosed with irritable bowel syndrome (bottom), studied in our lab\cite{IBS}.}
-	\label{fig:corrank}
+	%\vspace*{-10mm} % Corrects overbox of the figures
+	\includegraphics[width=0.99\textwidth]{results/corrank/IBS_h_A_amplicons_family_Rank.pdf}
+	\RSItable	
+	\caption{Rank variation throughout time for the most dominant elements (taxa) and their calculated Rank Stability Index (as shown in Material and Methods) for samples from a healthy subject studied in our lab\cite{IBS}.}
+	\label{fig:corrank_IBS_A}
 \end{figure}
 
-Nevertheless, in the particular case of the healthy individual, \emph{Burkholderiales} and \emph{Betaproteobacteria} (taxa ordered as 18th and 25th in the dominance axis) show comparatively very low rank stability regarding similar dominant taxa while, on the other hand, \emph{Comamonadaceae}, \emph{Lactobacillaceae}, \emph{Fusobacteriaceae}, \emph{Aerococcaceae} and \emph{Carnobacteriaceae} show higher stability than other more dominant taxa, forming a kind of \emph{rank stability island} for medium-ranked taxa around position 40 in the dominance axis, and thus colored in orange following Table \ref{tab:RSI} criteria, since they show a moderately stable RSI.
+\begin{figure}
+	\centering
+	%\vspace*{-10mm} % Corrects overbox of the figures
+	\hspace*{-4.5mm}\includegraphics[width=0.99\textwidth]{results/corrank/IBS_P2_Metatranscriptores_Rank.pdf}
+	\RSItable
+	\caption{Rank variation throughout time for the most dominant elements (taxa) and their calculated Rank Stability Index for samples from a subject diagnosed with irritable bowel syndrome studied in our lab\cite{IBS}.}
+	\label{fig:corrank_IBS_P2}
+\end{figure}
 
-\begin{table}
-  \begin{center}
-    \begin{tabular}{cccr}
-    \hline
-    Case  &  Condition  &  Colour  &  Description \\
-    \hline
-    1  &  $1\ge{\rm RSI}>0.99$  & \textcolor{blue}{blue} & constant rank \\ 
-    2  & ${\rm RSI}>0.90$  &  \textcolor{green}{green}  & highly stable rank \\
-    3  &  ${\rm RSI}>0.75$  &  \textcolor{orange}{orange} & moderately stable rank \\
-    4  &  ${\rm RSI}>0.25$  &  \textcolor{red}{red} & unstable rank \\
-    5  &  $0.25\ge{\rm RSI}\ge0$  &  \bfseries{black} & very unstable rank \\
-    \hline
-    \end{tabular}
-  \end{center} 
-  \caption{Colour code of the RSI percentage text following the first condition satisfied.}
-  \label{tab:RSI}
-\end{table}
+Nevertheless, in the particular case of the healthy individual in Figure \ref{fig:corrank_IBS_A}, \emph{Burkholderiales} and \emph{Betaproteobacteria} (taxa ordered as 18th and 25th in the dominance axis) show comparatively very low rank stability regarding similar dominant taxa while, on the other hand, \emph{Comamonadaceae}, \emph{Lactobacillaceae}, \emph{Fusobacteriaceae}, \emph{Aerococcaceae} and \emph{Carnobacteriaceae} show higher stability than other more dominant taxa, forming a kind of \emph{rank stability island} for medium-ranked taxa around position 40 in the dominance axis, and thus colored in orange, following the color criteria shown in the table included in Figure \ref{fig:corrank_IBS_A}, since they show a moderately stable RSI.
 
-In the IBS diagnosed patient, beyond the differences in dominance for the particular taxa, we still observe that the most dominant are the most rank stable.
+In the IBS diagnosed patient of Figure \ref{fig:corrank_IBS_P2}, beyond the differences in dominance for the particular taxa, we still observe that the most dominant are the most rank stable.
 However, as opposed to the healthy individual results, far from presenting a \emph{rank stability island}, the medium-ranked taxa are very rank unstable, mostly due to transient (often one or two consecutive samples) but deep drops in their relative abundance, which are usually happening more than twice along the time series.
 That is, for instance, the case of \emph{Sphingobacteriales} with two non-consecutive samples dropping to 111th rank position.
-In other cases, the high rank instability comes from a rank fluctuation over all the time series, as for \emph{Streptococcaceae} and \emph{Burkholderiales}, which are ranking 26th and 29th respectively in the overall dominance axis but show very low RSI, and thus colored in black attending to Table \ref{tab:RSI}.
+In other cases, the high rank instability comes from a rank fluctuation over all the time series, as for \emph{Streptococcaceae} and \emph{Burkholderiales}, which are ranking 26th and 29th respectively in the overall dominance axis but show very low RSI, and thus colored in black attending to the color criteria shown in the table included in Figure \ref{fig:corrank_IBS_P2}.
 

### 2016-09-22 mSystems/mSys_MAIN.tex
Some changes in main text and supplementary update leads to RC version
On branch master, changes to be committed:
	deleted:    doc/665380-revised-april.pdf
	new file:   doc/mSys-ITA-8-1-2016.pdf
	new file:   doc/mSys-sub-checklistAugust2016.pdf
	renamed:    mSys_suppl_0.1.pdf -> mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_discussion.tex
	modified:   mSys_intro.tex
	modified:   mSys_results.tex
	new file:   mSys_suppl.pdf
	new file:   mSys_suppl.tex
	deleted:    mSys_suppl_0.1.tex
	modified:   option.tc
	new file:   results/finalPlot11.bkp.pdf
	modifie

@@
 % release codenumber (update it here!)
-\newcommand\releasecode{0.5.0}
+\newcommand\releasecode{$\beta$}
 
@@
 %%\usepackage{float}
-%\usepackage[nolists,tablesfirst, nomarkers]{endfloat}
+\usepackage[nolists,tablesfirst, nomarkers]{endfloat}
 \usepackage{moreverb} % for verbatim output
@@
 % watermarks (time consuming!)
-\newwatermark[allpages,color=Fuchsia!4,angle=45,scale=9,xpos=-40,ypos=0]{DRAFT}
-\newwatermark[allpages,color=Fuchsia!12,angle=0,scale=1,xpos=0,ypos=137]{
-\emph{mSystems} \LaTeX\ manuscript release \releasecode}
+\newwatermark[allpages,color=Green!4,angle=45,scale=12,xpos=-60,ypos=0]{RC\releasecode}
+\newwatermark[allpages,color=Green!12,angle=0,scale=1,xpos=0,ypos=137]{
+\emph{mSystems} \LaTeX\ manuscript RC\releasecode}
 
@@
 \section*{Funding Information}
-This work was supported by grants to AM from the Spanish Ministry of Science and Competitivity (projects SAF 2012-31187, SAF2013-49788-EXP, SAF2015-65878-R), Carlos III Institute of Health (projects PIE14/00045 and AC15/00022), Generalitat Valenciana (project PrometeoII/2014/065) and co-financed by FEDER, and grants to CPG from the Generalitat Valenciana Prometeo Grants II/2014/050, II/2014/065, 419 by the Spanish Grants FPA2011-29678, BFU2012-39816-C02-01 of MINECO and by PITN-GA-420 2011-289442-INVISIBLES.
-From the Ministry of Economy and Competitiveness (grant FPI BES-2013-062767).
+This work was supported by grants to AM from the Spanish Ministry of Science and Competitivity (projects SAF2012-31187, SAF2013-49788-EXP, SAF2015-65878-R), Carlos III Institute of Health (projects PIE14/00045 and AC15/00022), Generalitat Valenciana (project PrometeoII/2014/065) and co-financed by FEDER, and grants to CPG from the Generalitat Valenciana Prometeo Grants II/2014/050, II/2014/065, 419 by the Spanish Grants FPA2011-29678, BFU2012-39816-C02-01 of MINECO and by PITN-GA-420 2011-289442-INVISIBLES.
+From the Ministry of Economy and Competitiveness (grants FPI BES-2012-052900 and FPI BES-2013-062767).
 
- \todo{Modificar becas de JMM y DMM ¿poner algún grant más?}
-
-\clearpage
+%\clearpage
 \begin{thebibliography}{99}
+%% Introduction
 \bibitem{holo1} {\bf Rosenberg E, Zilber-Rosenberg I.} 2016. Microbes Drive Evolution of Animals and Plants: the Hologenome Concept. MBio {\bf 7}:e01395–15–.
@@
 \bibitem{Moya_trends} {\bf Moya A, Ferrer M.} 2016. Functional Redundancy-Induced Stability of Gut Microbiota Subjected to Disturbance. Trends Microbiol {\bf 24}:402–413.
+\bibitem{mind} {\bf Cryan JF, Dinan TG}. 2012. Mind-altering microorganisms: the impact of the gut microbiota on brain and behaviour. Nat Rev Neurosci {\bf 13}:701–712
+\bibitem{AD} {\bf Xu R, Wang Q}. 2016. Towards understanding brain-gut-microbiome connections in Alzheimer's disease. BMC Systems Biology {\bf 10}:277-285
 \bibitem{CFS} {\bf Giloteaux L, Goodrich JK, Walters WA, Levine SM, Ley RE, Hanson MR}. 2016.
 Reduced diversity and altered composition of the gut microbiome in individuals with myalgic encephalomyelitis/chronic fatigue syndrome. Microbiome. {\bf 4}:30
-
+%
 \bibitem{microb&health} {\bf Marchesi JR, Adams DH, Fava F, Hermes GD a, Hirschfield GM, Hold G, Quraishi MN, Kinross J, Smidt H, Tuohy KM, Thomas L V, Zoetendal EG, Hart A.} 2015.
@@
 Science (80- ) {\bf 352}:565–569.
+\bibitem{panthropology} {\bf Amato KR} 2016. Incorporating the Gut Microbiota Into Models of Human and Non-Human Primate Ecology and Evolution. Yearbook Of Physical Anthropology {\bf 157}:S196–S215.
 \bibitem{sysbio&microb} {\bf Wu H, Tremaroli V, Bäckhed F.} 2015. Linking Microbiota to Human Diseases: A Systems Biology Perspective. Trends Endocrinol Metab {\bf 26}:758–770.
@@
 \bibitem{economics2} {\bf Eisler Z, Kertesz J, Yook SH, Barabasi AL.} 2005. Multiscaling and non-universality in fluctuations of driven complex systems. Europhys Lett {\bf 69}:664–670.
+\bibitem {cohen_taylor} {\bf Cohen JE, Xu M, Schuster WSF}.
+2013.
+Stochastic multiplicative population growth predicts and interprets Taylor’s power law of fluctuation scaling.
+Proc R Soc B Biol Sci {\bf 280}:20122955.
 \bibitem{animal1} {\bf Reed DH, Hobbs GR.} 2004. The relationship between population size and temporal variability in population size. Anim Conserv {\bf 7}:1–8.
@@
 
-% main studies
+%% Results
+\bibitem{fs} {\bf Eisler Z, Bartos I, Kertész J.} 2008. Fluctuation scaling in complex systems: Taylor’s law and beyond1. Adv Phys {\bf 57}:89–142.
+% -> main studies
 \bibitem{moving} {\bf Caporaso JG, Lauber CL, Costello EK, Berg-Lyons D, Gonzalez A, Stombaugh J, Knights D, Gajer P, Ravel J, Fierer N, Gordon JI, Knight R.} 2011.
@@
 Genome Biol {\bf 12}:R50.
+\bibitem{antibiotic} {\bf Dethlefsen L, Relman DA.} 2011.
+Incomplete recovery and individualized responses of the human distal gut microbiota to repeated antibiotic perturbation.
+Proc Natl Acad Sci {\bf108}:4554–61.
 \bibitem{LEA} {\bf Faith JJ, Guruge JL, Charbonneau M, Subramanian S, Seedorf H, Goodman AL, Clemente JC, Knight R, Heath AC, Leibel RL, Rosenbaum M, Gordon JI.} 2013.
 The long-term stability of the human gut microbiota.
-Science (80- ) {\bf 341}:1237439.
+Science {\bf 341}:1237439.
 \bibitem{kwashiorkor} {\bf Smith MI, Yatsunenko T, Manary MJ, Trehan I, Mkakosya R, Cheng J, Kau AL, Rich SS, Concannon P, Mychaleckyj JC, Liu J, Houpt E, Li J V, Holmes E, Nicholson J, Knights D, Ursell LK, Knight R, Gordon JI.} 2013.
@@
 Nature {\bf 505}:559–63.
-\bibitem{antibiotic} {\bf Dethlefsen L, Relman DA.} 2011.
-Incomplete recovery and individualized responses of the human distal gut microbiota to repeated antibiotic perturbation.
-Proc Natl Acad Sci {\bf108}:4554–61.
-
-\bibitem{FASTX} {\bf Gordon, A., Hannon, G.J.} 2010. FASTX-Toolkit. FASTQ/A shortreads pre-processing tools v0.0.13. http://hannonlab.cshl.edu/fastx\_toolkit/ (last accessed 26 Jul 2016).
-\bibitem{QIIME} {\bf Caporaso JG, Kuczynski J, Stombaugh J, Bittinger K, Bushman FD, Costello EK, Fierer N, Peña AG, Goodrich JK, Gordon JI, Huttley G a, Kelley ST, Knights D, Koenig JE, Ley RE, Lozupone C a, Mcdonald D, Muegge BD, Pirrung M, Reeder J, Sevinsky JR, Turnbaugh PJ, Walters W a, Widmann J, Yatsunenko T, Zaneveld J, Knight R.} 2010.
-correspondence QIIME allows analysis of high- throughput community sequencing data Intensity normalization improves color calling in SOLiD sequencing.
-Nat Publ Gr {\bf 7}:335–336.
-\bibitem{SILVA} {\bf Quast C, Pruesse E, Yilmaz P, Gerken J, Schweer T, Yarza P, Peplies J, Glöckner FO}.
-2013.
-The SILVA ribosomal RNA gene database project: improved data processing and web-based tools.
-Acids Res.
-{\bf 41}:D590-D596
-\bibitem{LMAT} {\bf Ames SK, Hysom DA, Gardner SN, Lloyd GS, Gokhale MB, Allen JE}.
-2013.
-Scalable metagenomic taxonomy classification using a reference genome database.
-Bioinformatics {\bf 29}:2253-2260.
-\bibitem{LMAT2} {\bf Ames SK, Gardner SN, Marti JM, Slezak TR, Gokhale MB, Allen JE}.
-2015.
-Using populations of human and microbial genomes for organism detection in metagenomes.
-Genome Res.
-{\bf 25}:1056-67.
-\bibitem{fs} {\bf Eisler Z, Bartos I, Kertész J.} 2008. Fluctuation scaling in complex systems: Taylor’s law and beyond1. Adv Phys {\bf 57}:89–142.
+% <- END main studies
 \bibitem{stat} {\bf Jørgensen B, Martinez JR, Tsao M.} 1994. Asymptotic behaviour of the variance function. Scand J Stat {\bf 21}:223–243.
-
 \bibitem{convergence1} {\bf Fronczak A, Fronczak P.} 2010. Origins of Taylor’s power law for fluctuation scaling in complex systems. Phys Rev E {\bf 81}:066112.
@@
 
-%discussion
+%% Discussion
 \bibitem {kilpatrick} {\bf Kilpatrick a M, Ives a R.} 2003. Species interactions can explain Taylor’s power law for ecological time series. Nature {\bf 422}:65–68.
@@
 Genome Biol {\bf 17}:121.
-\bibitem {cohen_taylor} {\bf Cohen JE, Xu M, Schuster WSF}.
-2013.
-Stochastic multiplicative population growth predicts and interprets Taylor ’ s power law of fluctuation scaling Stochastic multiplicative population growth predicts and interprets Taylor ’ s power law of fluctuation scaling.
-Proc R Soc B Biol Sci {\bf 280}:20122955.
 \bibitem {koenig} {\bf Koenig JE, Spor A, Scalfone N, Fricker AD, Stombaugh J, Knight R, Angenent LT, Ley RE.} 2011.
@@
 
-%M&M
+%% M&M
+% Selection and Methods
+\bibitem{Einstein} {\bf Einstein A.} 1905.
+Über die von der molekularkinetischen Theorie der Wärme geforderte Bewegung von in ruhenden Flüssigkeiten suspendierten Teilchen.
+Annalen der Physik, {\bf 322}:549-560.
+\bibitem{QIIME} {\bf Caporaso JG, Kuczynski J, Stombaugh J, Bittinger K, Bushman FD, Costello EK, Fierer N, Peña AG, Goodrich JK, Gordon JI, Huttley G a, Kelley ST, Knights D, Koenig JE, Ley RE, Lozupone C a, Mcdonald D, Muegge BD, Pirrung M, Reeder J, Sevinsky JR, Turnbaugh PJ, Walters W a, Widmann J, Yatsunenko T, Zaneveld J, Knight R.} 2010.
+correspondence QIIME allows analysis of high- throughput community sequencing data Intensity normalization improves color calling in SOLiD sequencing.
+Nat Publ Gr {\bf 7}:335–336.
+\bibitem{LMAT} {\bf Ames SK, Hysom DA, Gardner SN, Lloyd GS, Gokhale MB, Allen JE}.
+2013.
+Scalable metagenomic taxonomy classification using a reference genome database.
+Bioinformatics {\bf 29}:2253-2260.
+\bibitem{FASTX} {\bf Gordon, A, Hannon, GJ.} 2010. FASTX-Toolkit. FASTQ/A shortreads pre-processing tools v0.0.13. http://hannonlab.cshl.edu/fastx\_toolkit/ (last accessed 26 Jul 2016).
+\bibitem{SILVA} {\bf Quast C, Pruesse E, Yilmaz P, Gerken J, Schweer T, Yarza P, Peplies J, Glöckner FO}.
+2013.
+The SILVA ribosomal RNA gene database project: improved data processing and web-based tools.
+Acids Res.
+{\bf 41}:D590-D596
+\bibitem{LMAT2} {\bf Ames SK, Gardner SN, Marti JM, Slezak TR, Gokhale MB, Allen JE}.
+2015.
+Using populations of human and microbial genomes for organism detection in metagenomes.
+Genome Res.
+{\bf 25}:1056-67.
+\bibitem{boot} {\bf Wu, C.F.J.} 1986. Jackknife, bootstrap and other resampling methods in regression analysis. (with discussions) \textit{The Annals of Statistics} {\bf 14}:1261-1350 
 \bibitem{biopwrlaw} {\bf Xiao X, White EP, Hooten MB, Durham SL}. 2011. On the use of log-transformation vs. nonlinear regression for analyzing biological power laws. Ecology {\bf 92}:1887-1894
-\bibitem{boot} {\bf Wu, C.F.J.} 1986. Jackknife, bootstrap and other resampling methods in regression analysis. (with discussions) \textit{The Annals of Statistics} {\bf 14}:1261-1350 
 \end{thebibliography}
-
 \end{document}

### 2016-09-22 mSystems/mSys_MandM.tex
Some changes in main text and supplementary update leads to RC version
On branch master, changes to be committed:
	deleted:    doc/665380-revised-april.pdf
	new file:   doc/mSys-ITA-8-1-2016.pdf
	new file:   doc/mSys-sub-checklistAugust2016.pdf
	renamed:    mSys_suppl_0.1.pdf -> mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_discussion.tex
	modified:   mSys_intro.tex
	modified:   mSys_results.tex
	new file:   mSys_suppl.pdf
	new file:   mSys_suppl.tex
	deleted:    mSys_suppl_0.1.tex
	modified:   option.tc
	new file:   results/finalPlot11.bkp.pdf
	modifie

@@
 %%
+%\small
 \subsection*{Model} \label{sec:model}
@@
 where the first case, simplifies to $F= 3 V^2$ if $\beta = 0.75$ and the fitness of this taxon dominates in $\phi_0$. 
-In many physical systems (Brownian motion is the classical example), the two terms of the Langevin equation are related.
+In many physical systems (Brownian motion is the classical example \cite{Einstein}), the two terms of the Langevin equation are related.
 The \emph{Fluctuation Dissipation Theorem} states a general relationship between the response to an external disturbance and the internal fluctuations of the system\cite{FD}.
@@
 
-\todo{Explain better the Fluctuation Dissipation Theorem}
+%%
+\subsection*{Standardization} \label{sec:stan}
+In order to properly show all the studies under common axes, we decided to standardize the Taylor parameters using the group of healthy individuals for each study.
+With this approach, all the studies can be visualized in a shared plot with units of Taylor-parameters standard-deviation on their axes.
+
+For a Taylor parameter, e.g. $V$, the estimate of the mean ($\widehat{V}$) for the healthy subpopulation, composed of $h$ individuals, is:
+\begin{linenomath}
+$$\widehat{V} = \frac{1}{W_1}\sum_{i=1}^h V_i \omega_i=\sum_{i=1}^h V_i \omega_i$$
+\end{linenomath}
+as $W_1=\sum_i^h \omega_i=1$, since $\omega_i$ are normalized weights calculated as:
+\begin{linenomath}
+$$\omega_i = \frac{\frac{1}{\sigma^2_{V_i}}}{\sum_i^h\frac{1}{\sigma^2_{V_i}}}$$
+\end{linenomath}
+being $\sigma_{V_i}$ the estimation of the uncertainty in $V_i$ obtained together with $V_i$ from the X-weighted power-law fit described in Section \ref{sec:X-w}, for healthy individuals.
+
+Likewise, the estimation of the standard deviation for the healthy population ($\widehat{\sigma}_V$) is:
+\begin{linenomath}
+$$\widehat{\sigma}_V = \sqrt{\frac{1}{W_1-\frac{W_2}{W_1}}\sum_{i=1}^h\left[\omega_i\left(V_i-\hat{V}\right)^2\right]}$$
+\end{linenomath}
+being $W_2=\sum_i^h \omega_i^2$, which finally yields to:
+\begin{linenomath}
+$$\widehat{\sigma}_V = \sqrt{\frac{1}{1-\sum_i^h \omega_i^2}\sum_{i=1}^h\left[\omega_i\left(V_i-\hat{V}\right)^2\right]}$$
+\end{linenomath}
 
@@
 The bacteria and archaea taxonomic assignations were obtained by analysing 16S rRNA sequences, which were clustered into operational taxonomic units (OTUs) sharing 97 \% sequence identity using QIIME\cite{QIIME}.
-WGS data\cite{kwashiorkor} were analysed and assigned at strain level by the Livermore Metagenomic Analysis Toolkit (LMAT)\cite{LMAT}, according to their default quality threshold.
+Shotgun metagenomic sequencing (SMS) data\cite{kwashiorkor} were analysed and assigned at strain level by the Livermore Metagenomic Analysis Toolkit (LMAT)\cite{LMAT}, according to their default quality threshold.
 Genus, with best balance between error assignment and number of taxa, was chosen as our reference taxonomic level.
@@
 Metadata of each study is provided in Supplementary Tables S1 to S4.
-All used 16S rRNA gene sequencing except for the study of the discordant kwashiorkor twins\cite{kwashiorkor}, where both shotgun metagenomic sequencing (SMS) and 16S rRNA were used.
+All used 16S rRNA gene sequencing except for the study of the discordant kwashiorkor twins\cite{kwashiorkor}, where both SMS and 16S rRNA data were used.
 In the latter case we selected to work with SMS data to show that our method is valid regardless of the source of taxonomic information.
@@
 %%
-\subsubsection*{Taxa level selection}
+\subsubsection*{Taxa level robustness}
 We selected genus as taxonomic level for the subsequent steps of our work.
@@
 \caption{Overview of comparison of different approaches based on adjacent taxonomic levels using plots in the Taylor-parameters space.
-For 16S (former row of subfigures), the levels are family vs.
-genus, whereas for SMS (latter row of subfigures) levels are genus vs.
-species.
-The left column shows the raw results and the right column plots the standardized results (see Section \ref{sec:stan})}
+The former row of subfigures is for 16S, where levels are family (blue circles) vs.
+genus (purple triangles).
+Whereas the latter row of subfigures is for SMS, where levels are genus (blue circles) vs.
+species (purple triangles).
+The left column shows the raw results and the right column plots the standardized results (see Standardization under Material and Methods).}
 \label{fig:taxlev1}
@@
 
-%%
-\subsection*{Standardization} \label{sec:stan}
-In order to properly show all the studies under common axes, we decided to standardize the Taylor parameters using the group of healthy individuals for each study.
-With this approach, all the studies can be visualized in a shared plot with units of Taylor-parameters standard-deviation on their axes.
-
-For a Taylor parameter, e.g. $V$, the estimate of the mean ($\widehat{V}$) for the healthy subpopulation, composed of $h$ individuals, is:
-\begin{linenomath}
-$$\widehat{V} = \frac{1}{W_1}\sum_{i=1}^h V_i \omega_i=\sum_{i=1}^h V_i \omega_i$$
-\end{linenomath}
-as $W_1=\sum_i^h \omega_i=1$, since $\omega_i$ are normalized weights calculated as:
-\begin{linenomath}
-$$\omega_i = \frac{\frac{1}{\sigma^2_{V_i}}}{\sum_i^h\frac{1}{\sigma^2_{V_i}}}$$
-\end{linenomath}
-being $\sigma_{V_i}$ the estimation of the uncertainty in $V_i$ obtained together with $V_i$ from the X-weighted power-law fit described in Section \ref{sec:X-w}, for healthy individuals.
-
-Likewise, the estimation of the standard deviation for the healthy population ($\widehat{\sigma}_V$) is:
-\begin{linenomath}
-$$\widehat{\sigma}_V = \sqrt{\frac{1}{W_1-\frac{W_2}{W_1}}\sum_{i=1}^h\left[\omega_i\left(V_i-\hat{V}\right)^2\right]}$$
-\end{linenomath}
-being $W_2=\sum_i^h \omega_i^2$, which finally yields to:
-\begin{linenomath}
-$$\widehat{\sigma}_V = \sqrt{\frac{1}{1-\sum_i^h \omega_i^2}\sum_{i=1}^h\left[\omega_i\left(V_i-\hat{V}\right)^2\right]}$$
-\end{linenomath}
+%\normalsize

### 2016-09-22 mSystems/mSys_discussion.tex
Some changes in main text and supplementary update leads to RC version
On branch master, changes to be committed:
	deleted:    doc/665380-revised-april.pdf
	new file:   doc/mSys-ITA-8-1-2016.pdf
	new file:   doc/mSys-sub-checklistAugust2016.pdf
	renamed:    mSys_suppl_0.1.pdf -> mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_discussion.tex
	modified:   mSys_intro.tex
	modified:   mSys_results.tex
	new file:   mSys_suppl.pdf
	new file:   mSys_suppl.tex
	deleted:    mSys_suppl_0.1.tex
	modified:   option.tc
	new file:   results/finalPlot11.bkp.pdf
	modifie

@@
 Interestingly, the variability parameter was higher in the healthy subset for the study of the discordant twins suffering from kwashiorkor disease \cite{kwashiorkor}.
-Taking into account that the infant microbiota is evolving toward a definite, adult state \cite{koenig}, it means that the temporal variability would be greater than in a adult who has reached a stability in his gut microbiota, while our results could be directing in the possibility that this variability is necessary in order to reach that adult state  \todo{<-review the wording of the whole sentence}.
+In this regard it has been shown that the infant microbiota needs to develop toward a definite, adult state \cite{koenig}.
+This implies that the temporal variability would be greater in children respect to a healthy adult state, which should be temporally stable.
+Thus, our results could be pointing out the need of this variability in order to reach that adult state.
 Furthermore, as we wanted to see how this variability behaved over time, we calculated the evolution of this parameter for the samples which had enough time sampling.

### 2016-09-22 mSystems/mSys_intro.tex
Several minor changes leading to 0.5.0, another step towards pre-releasing
On branch master changes to be committed:
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_discussion.tex
	modified:   mSys_intro.tex
	modified:   mSys_results.tex
Some changes in main text and supplementary update leads to RC version
On branch master, changes to be committed:
	deleted:    doc/665380-revised-april.pdf
	new file:   doc/mSys-ITA-8-1-2016.pdf
	new file:   doc/mSys-sub-checklistAugust2016.pdf
	renamed:    mSys_suppl_0.1.pdf -> mSys_MAIN.pdf
	modified:   mSy

@@
 Human microbiota has been suggested to be closely related to diseases like type 2 diabetes \cite{diabetes2}, cardiovascular disease (CVD) \cite{CVD}, irritable bowel syndrome \cite{IBS}, Crohn's disease \cite{CD}, some affections as obesity \cite{ob1, ob2}, malnutrition \cite{nutr} among other multiple diseases \cite{Moya_trends}.
-High throughput methods for microbial 16S ribosomal RNA gene and WGS have now begun to reveal the composition of archaeal, bacterial, fungal and viral communities located both, in and on the human body.
+Current studies reveal that gut microbiota also influences brain function and behaviour and is related with neurological disorders like Alzheimer's disease through the brain-gut-microbiome axis\cite{mind,AD}.
+Recently, even a mystifying and elusive diagnosis condition as chronic fatigue syndrome, which has often been suggested to be a psychosomatic disease, has been closely related to reduced diversity and altered composition of the gut microbiome\cite{CFS}.
+
+High throughput methods for microbial 16S ribosomal RNA gene and SMS have now begun to reveal the composition of archaeal, bacterial, fungal and viral communities located both, in and on the human body.
 Modern high-throughput sequencing and bioinformatics tools provide a powerful means of understanding how the human microbiome contributes to health and its potential as a target for therapeutic interventions \cite{microb&health}.
-To define normal microbiota and how it's compositional changes can origin some diseases are important issues still in need for scientific answers \cite{normal1, normal2}.
+To define normal host-gut microbe interactions and how the microbiota compositional changes can origin some diseases are important issues still in need for scientific answers \cite{normal1, normal2, panthropology}.
 
@@
 Microbiota can be approached under the light of ecological theory where we can find, for instance, general principles as the Taylor's law \cite{taylor}, which relates spatial or temporal variability of the population with its mean.
-This law, also known as fluctuation scale law, is ubiquitous in the natural world and can be found in several systems as random walks \cite{randomwalks}, stock markets \cite{economics1, economics2}, animal populations \cite{taylor, animal1, animal2}, gene expression \cite{genexpress}, or in the human genome \cite{genome}.
+This law, also known as fluctuation scale law, is ubiquitous in the natural world and can be found in several systems as random walks \cite{randomwalks}, stock markets \cite{economics1, economics2}, tree \cite{cohen_taylor} and animal populations \cite{taylor, animal1, animal2}, gene expression \cite{genexpress}, or in the human genome \cite{genome}.
 Taylor's law has been applied to microbiota in a spatial way in the work of Zhang {\it et al.}, (2014) \cite{isme1}, where they show that this population tend to be in an aggregated way rather than in a random distribution.

### 2016-09-22 mSystems/mSys_results.tex
Some changes in main text and supplementary update leads to RC version
On branch master, changes to be committed:
	deleted:    doc/665380-revised-april.pdf
	new file:   doc/mSys-ITA-8-1-2016.pdf
	new file:   doc/mSys-sub-checklistAugust2016.pdf
	renamed:    mSys_suppl_0.1.pdf -> mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_discussion.tex
	modified:   mSys_intro.tex
	modified:   mSys_results.tex
	new file:   mSys_suppl.pdf
	new file:   mSys_suppl.tex
	deleted:    mSys_suppl_0.1.tex
	modified:   option.tc
	new file:   results/finalPlot11.bkp.pdf
	modifie

@@
 	\includegraphics[width=0.8\textwidth]{results/fits/IBS_P2_Metatranscriptores_stdVSmean_xWboot_LOG.pdf}
-	\caption{X-weighted power-law fits of the standard deviations versus the mean values for each bacterial genus monitored in time.
+	\caption{X-weighted power-law fits of the standard deviations (std) versus the mean values for each bacterial genus monitored in time.
 We show the fit for samples from a healthy subject (top) and from a subject diagnosed with irritable bowel syndrome (bottom), studied in our lab \cite{IBS}.
@@
 
-Taylor parameters describing the temporal variability of the gut microbiome in our sampled individuals are shown in Tables S1 to S6.
+Taylor parameters describing the temporal variability of the gut microbiome in our sampled individuals are shown in Supplementary Tables S1 to S6.
 Our results hint at an ubiquitous behavior.
@@
 This enables us to define an area in the Taylor parameter space that we called the \emph{healthy zone}.
-% should be changed for any other name: stable zone?
 
-In order to jointly visualize and compare the results of individuals from different studies, their Taylor parameters have been standardized, where standardization means that each parameter is subtracted by the mean value and divided by the standard deviation of the group of healthy individuals for each study (for details of the procedure, please see Standardization subsection in Material and Methods).
+In order to jointly visualize and compare the results of individuals from different studies\cite{IBS,moving,antibiotic,LEA,kwashiorkor,diet}, their Taylor parameters have been standardized, where standardization means that each parameter is subtracted by the mean value and divided by the standard deviation (std) of the group of healthy individuals for each study (for details of the procedure, please see Standardization subsection in Material and Methods).
 The healthy zone and the standardized Taylor parameters for individuals whose gut microbiota is altered (i.e., suffering from kwashiorkor, altered diet, antibiotics or IBS) is shown in Figure \ref{fig:main2}.
@@
 A global property emerges from all worldwide data collected: Taylor parameters characterize the statistical behavior of microbiome changes.
-Furthermore, we have verified that our conclusions are robust to systematic errors due to taxonomic assignment (see Figure Sx in Supplemental Material).
+Furthermore, we have verified that our conclusions are robust to systematic errors due to taxonomic assignment (see Taxa level selection in Material and Methods).
 
@@
 Points with errors place each individual gut microbiome in the Taylor space.
-Note that the parameters have been standardized (stdev units) to the healthy group in each study for demonstrative and comparative purposes.}
+Note that the parameters have been standardized (standard deviation units) to the healthy group in each study for demonstrative and comparative purposes.}
 	\label{fig:main2}

### 2016-09-22 mSystems/mSys_suppl.tex
Some changes in main text and supplementary update leads to RC version
On branch master, changes to be committed:
	deleted:    doc/665380-revised-april.pdf
	new file:   doc/mSys-ITA-8-1-2016.pdf
	new file:   doc/mSys-sub-checklistAugust2016.pdf
	renamed:    mSys_suppl_0.1.pdf -> mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_MandM.tex
	modified:   mSys_discussion.tex
	modified:   mSys_intro.tex
	modified:   mSys_results.tex
	new file:   mSys_suppl.pdf
	new file:   mSys_suppl.tex
	deleted:    mSys_suppl_0.1.tex
	modified:   option.tc
	new file:   results/finalPlot11.bkp.pdf
	modifie

--- 2016-04-15 eLife/elife_suppl_sub.tex
@@
-\documentclass[11pt,oneside,letterpaper]{article}
+% !TEX TS-program = pdflatex
+% !TEX encoding = UTF-8 Unicode
 
-% graphicx package, useful for including eps and pdf graphics
-\usepackage{graphicx}
-\DeclareGraphicsExtensions{.pdf,.png,.jpg}
+\documentclass[12pt,oneside,letterpaper]{article}
 
-% basic packages
-\usepackage{color} 
-\usepackage{parskip}
-\usepackage{float}
+% release codenumber (update it here!)
+\newcommand\releasecode{$\beta$}
+
+%%% PACKAGES
+\usepackage{lmodern} % Activa fuentes y macros de Latin Modern
+\usepackage[T1]{fontenc} % set font encoding
+\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)
+\usepackage[english]{babel} 
+\usepackage[expert, charter]{mathdesign} % Fuentes... disponibles conjuntos {charter|utopia}
+%%\usepackage{calligra}
+\usepackage[printwatermark]{xwatermark}
+%\usepackage{graphicx}
+\usepackage[usenames,dvipsnames,table]{xcolor}
+%\usepackage{parskip}
+%%\usepackage{float}
+%\usepackage[nolists,tablesfirst, nomarkers]{endfloat}
+%\usepackage{moreverb} % for verbatim output
 
 % text layout
-\usepackage{geometry}
-\geometry{textwidth=15cm} % 15.25cm for single-space, 16.25cm for double-space
-\geometry{textheight=22cm} % 22cm for single-space, 22.5cm for double-space
+\usepackage[letterpaper]{geometry}
+\geometry{textwidth=16.25cm} % 15.25cm for single-space, 16.25cm for double-space
+\geometry{textheight=22.5cm} % 22cm for single-space, 22.5cm for double-space
+
+% watermarks (time consuming!)
+\newwatermark[allpages,color=Green!4,angle=45,scale=12,xpos=-60,ypos=0]{RC\releasecode}
+\newwatermark[allpages,color=Green!12,angle=0,scale=1,xpos=0,ypos=137]{
+\emph{mSystems} \LaTeX\ manuscript RC\releasecode}
 
@@
 
-\newcommand{\CC}[0]{\emph{cmplxcruncher}}
-\newcommand{\task}[1]{\texttt{\bfseries\scshape\textcolor{MidnightBlue}{#1}}}
-\newcommand{\un}[1]{\operatorname{#1}}
-\newcommand{\unclassrate}[0]{\un{kpb/s/core}}
+% line numbering
+\usepackage[running,mathlines]{lineno}
+\renewcommand\thelinenumber{\color{red}\arabic{linenumber}}
+\linenumbers
 
@@
 \usepackage[labelfont=bf,labelsep=period,font=small]{caption}
+\captionsetup[table]{name=Supplementary Table S\!\!}
 
@@
 \usepackage{setspace} 
-\doublespacing
 \captionsetup{labelfont=bf,labelsep=period,font=doublespacing}
@@
 
+% Package authblk
 \usepackage{authblk}
@@
 \renewcommand\Affilfont{\small \normalfont}
-\makeatletter
-\renewcommand\AB@affilsepx{, \protect\Affilfont}
-\makeatother
 
@@
 \usepackage{amsmath}
-\usepackage{amssymb}
 
-%%% TITLE %%%
-\title{\vspace{1.0cm} \Large \bf 
-Supplemental material of ``Health and disease imprinted in the time variability of the human microbiome''
+%% Roman footnote numbering style to not collide with bibliographic cites: 
+\renewcommand{\thefootnote}{\Roman{footnote}}
+
+% aux commands
+\newcommand{\CC}[0]{\emph{cmplxcruncher}}
+\newcommand{\task}[1]{\texttt{\bfseries\scshape\textcolor{MidnightBlue}{#1}}}
+\newcommand{\un}[1]{\operatorname{#1}}
+\newcommand{\unclassrate}[0]{\un{kpb/s/core}}
+\newcommand{\todo}[1]{\texttt{\bfseries\textcolor{Orange}{#1}}}
+
+% word count (Need --enable-write18 or --shell-escape)
+\immediate\write18{texcount -opt=option.tc \jobname.tex > wordcount.aux} 
+\newcommand\wordcount{\verbatiminput{wordcount.aux}}
+
+%%% DOCUMENT %%%
+
+\begin{document}
+
+\title{
+	\vspace*{0mm} % May help to include the corresponding author footnote in the title page
+	\singlespacing
+	\begin{flushleft}
+		\texttt{\large Title:} \\
+	\end{flushleft}
+	\vspace*{2mm}
+	Supplemental material of "Health and disease imprinted in the time variability of the human microbiome"\\
+	\vspace*{6mm}
+	\begin{flushleft}
+		\texttt{\large Running title:} \\
+	\end{flushleft}
+	\vspace*{0mm}
+	Supplemental material of "Microbiota, are you sick?"
+	\vspace*{4mm}
 	}
 
-\author[1]{Jose Manuel Mart\' i}
-\author[1,2]{Daniel M. Mart\' inez}
-\author[1]{Manuel Pe\~na}
-\author[1]{C\'esar Gracia}
-\author[2,3,4]{Amparo Latorre}
-\author[2,3,4]{Andr\'es Moya}
-\author[1]{Carlos P. Garay}
+\doublespacing
 
-\affil[1]{Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain.}
-\affil[2]{FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain.}
-\affil[3]{Cavanilles Institute of Biodiversity and Evolutionary Biology, Univ. de Valencia, 46980, Spain. }
-\affil[4]{CIBER en Epidemiolog\' ia y Salud P\'ublica (CIBEResp), Madrid, Spain}
+\author[1,2,$*$]{Jose Manuel Martí}
+\author[1,2,3,$*$]{Daniel Martínez-Martínez}
+\author[2]{Manuel Peña}
+\author[1,2]{César Gracia}
+\author[1,3,4,5]{Amparo Latorre}
+\author[1,3,4,5,\#]{Andrés Moya}
+\author[1,2,\#]{Carlos P. Garay}
+
+\affil[1]{Institute for Integrative Systems Biology (I2SysBio), 46980, Spain.}
+\affil[2]{Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain.}
+\affil[3]{FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain.}
+\affil[4]{Cavanilles Institute of Biodiversity and Evolutionary Biology, UVEG, 46980, Spain.}
+\affil[5]{CIBER en Epidemiología y Salud Pública (CIBEResp), Madrid, Spain}
 
@@
 
-\begin{document}
+\maketitle
+\footnote[0]{$^*$ Equally contributed}
+\footnote[0]{$^\#$ Corresponding authors: andres.moya@uv.es, penagaray@gmail.com}
 
-\maketitle
+\clearpage
+\begin{table}
+  \caption{Taylor parameters.
+Individuals with either animal-based (A) or plant-based (P) diets\cite{diet}.
+Previous to diet,  the population sampled is described by $\bar{V} = 0.09 \pm 0.05, \bar{\beta} = 0.77 \pm 0.04$.}
 
-\section{Model} \label{sec:model}
-
-We model the microbial abundances across time along the lines of Blumm \textit{et al.}\cite{ranking}. The dynamics of taxon relative abundances is described by the Langevin equation: 
-\begin{equation}
-\dot{x_i} = F_i \cdot x_i^\alpha + V \cdot x_i^\beta \xi_i(t) - \phi(t) \cdot x_i,
-\end{equation}
-where F$_i$ captures the fitness of the taxon i, V corresponds to the noise amplitude and $\xi_i$(t) is a Gaussian random noise with zero mean  $<\xi_i(t)>$ = 0 and variance uncorrelated in time, $<\xi_i(t) \xi_i(t')>$ =  $\delta(t'-t)$, .
-The function $\phi(t)$ ensures the normalization at all times, $\sum x_i(t) = 1$, and corresponds to $\phi(t) = \sum F_i x_i^\alpha + \sum V x_i^\beta \xi_i(t)$ .
-The temporal evolution of the probability that a taxon i has a relative abundance $x_i(t)$, P(x$_i$,t), is determined by the Fokker-Planck equation:
-\begin{equation}
-\frac{\partial P}{\partial t} = - \frac{\partial}{\partial x_i}  [(F_i \cdot x_i^\alpha - \phi(t) \cdot x_i ) \cdot P]+ \frac{1}{2} \frac{\partial^2}{\partial x_i^2} (V^2 \cdot x_i^{2\beta}\cdot P).
-\end{equation}
-The microbiota evolves towards a steady-state with a time-independent probability depending on the values of $\alpha$, $\beta$, F$_i$ and V.
-For $\alpha<1$ (otherwise, systems are always unstable), the steady-state probability may be localized in a region around a preferred value or broadly distributed over a wide range, depending on whether the fitness F$_i$  dominates or is overwhelmed by the noise amplitude V.
-The steady-state solution of the Fokker-Planck equation is given by:
-\begin{eqnarray}
-P_0 (x_i) &=& C_{ne}(\alpha,\beta,F_i,V)  \cdot x_i^{-2\beta}  \cdot \exp\Big[\frac{2F_i}{V^2}\frac{x_i^{1+\alpha-2\beta}}{1+\alpha-2\beta}-\frac{\phi_0}{V^2}\frac{x_i^{2-2\beta}}{1-\beta}\Big] \quad \textrm{if} \quad  2\beta \ne 1+\alpha, \notag \\
-P_0 (x_i) &=& C_e(\alpha,\beta,F_i,V)  \cdot x_i^{\frac{2F_i}{V^2} -2\beta}  \cdot \exp\Big[\frac{\phi_0}{V^2}\frac{x_i^{2-2\beta}}{1-\beta}\Big] \quad \textrm{if} \quad  2\beta = 1+\alpha,
-\notag
-\end{eqnarray}
-where $\phi_0 = (\sum_i F_i^{1/(1-\alpha)})^{1-\alpha}$ and C$_{ne}$ and C$_{e}$ are integrals that should be solved numerically for the parameters of interest.
-The ordered phase happens when the solution has a maximum in the physical interval ($0<x_i<1$).
-For larger V, the transition to a disordered phase happens when the maximum shifts to the unphysical region $x_i<0$, which sets the phase transition region V($\alpha,\beta,F_i$).
-The phase transition region can be calculated analytically in particular cases:
-\begin{eqnarray}
-F_i^2 &=& 4 \beta \phi_0 V^2\quad \textrm{if} \quad  \beta = \alpha \neq 1, \notag \\
-F_i &=& \beta V^2\quad \textrm{if} \quad  2\beta = 1+\alpha,
-\notag
-\end{eqnarray}
-where the first case, simplifies to $F= 3 V^2$ if $\beta = 0.75$ and the fitness of this taxon dominates in $\phi_0$. 
-In many physical systems (Brownian motion is the classical example), the two terms of the Langevin equation are related.
-The \emph{fluctuation--dissipation theorem} states a general relationship between the response to an external disturbance and the internal fluctuations of the system\cite{FD}.
-The theorem can be used as the basic formula to derive the fitness from the analysis of fluctuations
-of the microbiota, assuming that it is in equilibrium (the ordered phase).
-
-\section{Selection and Methods}
-
-\subsection{Sample selection}
-We have chosen studies about relevant pathologies containing metagenomic sequencing time data series of bacterial populations from humans in different healthy and non-healthy states.
-We have selected only those individuals who had three or more time points of data available in databases.
-Metadata of each study is provided in Supplementary Tables \ref{tab:diet} to \ref{tab:LEA}.
-All used 16S rRNA gene sequencing except for the study of the discordant kwashiorkor twins\cite{kwashiorkor} (see Supplementary Tables \ref{tab:DH} and \ref{tab:DK}) where shotgun metagenomic sequencing (SMS) and 16S rRNA were used.
-In the latter case we selected to work with SMS data to show that our method is valid regardless of the source of taxonomic information.
-Each one of the datasets was treated as follows:
-
-\subsection{16rRNA sequences processing}
-Reads from the selected studies were first quality filtered using the FastX toolkit\cite{FASTX}, allowing only those reads which had more than 25 of quality along the 75\% of the complete sequence.
-16S rRNA reads were then clustered at 97\% nucleotide sequence identity (97\% ID) into operational taxonomic units (OTUs) using QIIME package software\cite{QIIME} (version 1.8) We followed open reference OTU picking workflow in all cases.
-The clustering method used was uclust, and the OTUs were matched against Silva database\cite{SILVA} (version 111, July 2012) and were assigned to taxonomy with an uclust-based consensus taxonomy assigner.
-The parameters used in this step were: similarity 0.97, prefilter percent id 0.6, max accepts 20, max rejects 500.
-
-\subsection{Metagenomic sequences processing}
-Metagenomic shotgun (and 16S too) sequences were analyzed with LMAT (Livermore Metagenomics Analysis Toolkit) software package\cite{LMAT} (version 1.2.4, with Feb'15 release of data base \emph{LMAT-Grand}).
-LMAT was run using a Bull shared-memory node belonging to the team's HPC (high performance computing) cluster.
-It is equipped with 32 cores (64 threads available using Intel Hyper-threading technology) as it has 2 Haswell-based Xeons, the E5-2698v3@2.3 GHz, sharing half a tebibyte (0.5 TiB, that is, 512 gibibytes) of DRAM memory.
-This node is also provided with a card PCIe SSD as NVRAM, the P420m HHHL, with 1.4 TB, and 750000 reading IOPS, 4 KB, achieving 3.3 GB/s, which Micron kindly issued free of charge, as a sample for testing purposes.
-The computing node was supplied with a RAID-0 (striping) scratch disk area.
-We used the ``Grand'' database\footnote{In this context, ``Grand'' refers to a huge database that contains k-mers from all viral, prokaryote, fungal and protist genomes present in the NCBI database, plus Human reference genome (hg19), plus GenBank Human, plus the 1000 Human Genomes Project (HGP).
-This represent about 31.75 billion k-mers occupying 457.62 GB.}, release Feb'15, provided by the LMAT team.
-Previously to any calculation, the full database was loaded in the NVRAM.
-With this configuration the observed LMAT sustained sequence classification rate was $20\unclassrate$.
-Finally, it is worth mentioning that a complete set of Python scripts have been developed as back-end and front-end of the LMAT pipeline in order to manage the added complexity of time series analysis.
-
-\subsection{Taxa level selection}
-We selected genus as taxonomic level for the subsequent steps of our work.
-In order to ensure that, between adjacent taxonomic levels, there were not crucial differences which could still be of relevance after standardization (see Section \ref{sec:stan}), we tested two different data sets.
-In the former, the antibiotics study\cite{antibiotic} with 16S data, we tested the differences between genus and family levels.
-The latter dataset tested was the kwashiorkor discordant twins study\cite{kwashiorkor} for both genus and species taxonomic levels.
-The Supplementary Figures \ref{fig:taxlev1} (overview) and \ref{fig:taxlev2} (detail) plot the comparison between studies (and so, 16S and SMS) and between adjacent taxonomic levels.
-
-\begin{figure} 
-  \includegraphics[width=0.5\textwidth]{results/taxalevel/sum_raw_16S.pdf}
-  \includegraphics[width=0.5\textwidth]{results/taxalevel/sum_sta_16S.pdf}
-  \includegraphics[width=0.5\textwidth]{results/taxalevel/sum_raw_SMS.pdf} 
-  \includegraphics[width=0.5\textwidth]{results/taxalevel/sum_sta_SMS.pdf}
-\caption{Overview of comparison of different approaches based on adjacent taxonomic levels using plots in the Taylor-parameters space.
-For 16S (former row of subfigures), the levels are family vs.
-genus, whereas for SMS (latter row of subfigures) levels are genus vs.
-species.
-The left column shows the raw results and the right column plots the standardized results (see Section \ref{sec:stan})}
-\label{fig:taxlev1}
-\end{figure}
-
-\begin{figure} 
-  \includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_fam_16S.pdf}
-  \includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_gen_16S.pdf}
-  \includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_gen_SMS.pdf}
-  \includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_spc_SMS.pdf}
-\caption{Detail of comparison of different approaches based on adjacent taxonomic levels using plots of X-weighted power-law fits (see Section \ref{sec:X-w}).
-The former row of subfigures shows examples for 16S, whereas the latter row of subfigures plots examples for SMS.
-The left column shows results for the superior taxonomic level (family for 16S, genus for SMS), while the right column shows results for the inferior level (genus for 16S, specie for SMS).}
-\label{fig:taxlev2}
-\end{figure}
-
-\begin{table} 
   \begin{center}
@@
   \end{center}
-  \caption{Taylor parameters.
-Individuals with either animal-based (A) or plant-based (P) diets\cite{diet}.
-Previous to diet,  the population sampled is described by $\bar{V} = 0.09 \pm 0.05, \bar{\beta} = 0.77 \pm 0.04$.}
   \label{tab:diet}
@@
 
+\clearpage
+
 \begin{table} 
+\caption{Taylor parameters for individuals taking antibiotics\cite{antibiotic}.
+Prior to antibiotics intake, the population sampled is described by $\bar{V} = 0.12 \pm 0.05, \bar{\beta} = 0.75 \pm 0.04$.}
   \begin{center}
@@
   \end{center}
-  \caption{Taylor parameters for individuals taking antibiotics\cite{antibiotic}.
-Prior to antibiotics intake, the population sampled is described by $\bar{V} = 0.12 \pm 0.05, \bar{\beta} = 0.75 \pm 0.04$.}
   \label{tab:antibiotics}
@@
 
+\clearpage
+
 \begin{table}
+\caption{Taylor parameters for persons diagnosed with irritable bowel syndrome (IBS)\cite{IBS}.
+Healthy individuals sampled in this study are characterized by $\bar{V} = 0.134 \pm 0.009, \bar{\beta} = 0.691 \pm 0.025$.}
+
   \begin{center}
@@
   \end{center}
-  \caption{Taylor parameters for persons diagnosed with irritable bowel syndrome (IBS)\cite{IBS}.
-Healthy individuals sampled in this study are characterized by $\bar{V} = 0.134 \pm 0.009, \bar{\beta} = 0.691 \pm 0.025$.}
   \label{tab:IBS}
@@
 
+\clearpage
+
 \begin{table} 
+\caption{Taylor parameters for the healthy subject of the discordant twins\cite{kwashiorkor}.
+This table continues in Supplementary Table \ref{tab:DK}.
+The population of healthy twins is characterized by $\bar{V} = 0.25 \pm 0.10, \bar{\beta} = 0.863 \pm 0.028$.}
   \begin{center}
@@
   \end{center}
-  \caption{Taylor parameters for the healthy subject of the discordant twins\cite{kwashiorkor}.
-This table continues in Supplementary Table \ref{tab:DK}.
-The population of healthy twins is characterized by $\bar{V} = 0.25 \pm 0.10, \bar{\beta} = 0.863 \pm 0.028$.}
   \label{tab:DH}
@@
 
+\clearpage
+
 \begin{table}
+\caption{Taylor parameters for the kwashiorkor part of the discordant twins\cite{kwashiorkor}.
+This is a continuation of Supplementary Table \ref{tab:DH}.
+The population of healthy twins is characterized by $\bar{V} = 0.25 \pm 0.10, \bar{\beta} = 0.863 \pm 0.028$.}
    \begin{center}
@@
   \end{center}
-  \caption{Taylor parameters for the kwashiorkor part of the discordant twins\cite{kwashiorkor}.
-This is a continuation of Supplementary Table \ref{tab:DH}.
-The population of healthy twins is characterized by $\bar{V} = 0.25 \pm 0.10, \bar{\beta} = 0.863 \pm 0.028$.}
     \label{tab:DK}
@@
 
+\clearpage
+
 \begin{table}
+\caption{Taylor parameters for individuals with different degrees of overweight and obesity\cite{LEA}.
+Healthy people in this study, whom were not obese, are characterized by $\bar{V} = 0.19 \pm 0.06, \bar{\beta} = 0.806 \pm 0.034$.}
   \begin{center}
@@
   \end{center}
-  \caption{Taylor parameters for individuals with different degrees of overweight and obesity\cite{LEA}.
-Healthy people in this study, whom were not obese, are characterized by $\bar{V} = 0.19 \pm 0.06, \bar{\beta} = 0.806 \pm 0.034$.}
   \label{tab:LEA}
@@
 
-\section{Temporal evolution of model parameters}
+\clearpage
 
-We have studied the time dependence of the variability $V$ and power law index $\beta$ (see Section \ref{sec:model}) by using a sliding window approach.
-The total number of time points are divided in subsets of five points, where next subset is defined by adding next time sampling and by eliminating the earliest one.
-Both parameters were calculated for each subset against the average time lapse.
-Figure \ref{fig:tempevo1} shows the variability  $V$ as a function of time for the largest sampling: two individuals in the Caporaso's study\cite{moving} corresponding to the gut microbiota of a male (upper plot) and a female (lower plot).
-Figure \ref{fig:tempevo2} shows the time evolution of $V$ for patient P2 of the IBS study\cite{IBS} (upper plot) and patient D in the antibiotics study\cite{antibiotic} (lower plot).
-
-\begin{figure}
-	\includegraphics[width=1.0\textwidth]{results/sliwin/male_mov.pdf}
-	\hspace*{3mm}\includegraphics[width=0.448\textwidth]{results/sliwin/female_mov.pdf}
-\caption{$V$ as a function of time for the two individuals in the Caporaso's study\cite{moving}: samples of gut microbiome of a male (upper plot) and a female (lower plot).
-Both samples show changes in the variability V with quasi--periodic behavior peaked at about 10 days.
-Variability grows more for the gut microbiota of the male and share a minimal value around 0.1 with the gut microbiota of the female.}
-\label{fig:tempevo1}
-\end{figure}
-
-\begin{figure}
-	\centering 
- 	\includegraphics[width=0.8\textwidth]{results/sliwin/patP2_IBS.pdf}
-  	\includegraphics[width=1.0\textwidth]{results/sliwin/patD_antibio.pdf} 
-\caption{$V$ as a function of time for patient P2 of the IBS study\cite{IBS} (upper plot) and patient D in the antibiotics study\cite{antibiotic} (lower plot). The variability 
-of the gut microbiota of P2 decreases from above 0.3 to below 0.2, showing a slow tendency to increase the order of the system.
-Antibiotic intake leaks to a quick increase of variability which lasts for a few days to recover ordering.
-The second antibiotic treatment shows some memory (lower increase of variability) with a slower recovery.
-NOTE: The blue vertical lines in the lower plot are showing the periods of antibiotic treatment.}
-\label{fig:tempevo2}
-\end{figure}
-
-\section{Engineered Software}
-
-A complete software framework, named 'ComplexCruncher', has been engineered to support the analysis of the dynamics of ranking processes in complex systems.
-Although the software was devised with a clear bias towards metagenomics, it is general enough to be able to cope with a ranking process in any complex system.
-Implemented in Python using well-known open-source community software, the software solution is composed of two parts that can be used together or apart: a web-based graphic front-end connected to a database, and a computing kernel.
-Used together, this software enables other users to reproduce our results easily and, furthermore, upload and analyse their own data or experiment with the preloaded metagenomics data sets.
-The next sections of this supplementary material deal with these two software pieces in detail.
-
-\section{Web-based front-end}
-
-`ComplexCruncher WebPortal' (CCWebPortal) is a web platform designed to allow the user to interact with a data repository of selected and well-documented metagenomics data sources.
-Through a few simple steps, the user can perform advanced searches on the complete set of records in the metagenomics repository.
-The web application provides advanced filters that allow the user to reduce the search to a small set of interest.
-After this first step, the user can refine the search and discard those records that do not meet certain requirements.
-
-The web application allows calculations to be done directly by the stable release of the \CC\ computing kernel.
-At the end of the calculations, the results are displayed to the user on the same browser which runs the web application.
-Then, the user can interact over the series of generated graphics thus allowing flexible comparison among them.
-In addition, CCWebPortal enables direct download of generated data (plots, spreadsheets, etc).
-The web application generates a report file summarizing all the results in PDF format.
-If the user has login permissions, CCWebPortal enables the option of insert new database records in addition to editing and deleting existing ones.
-
-CCWebPortal is a web application that runs on current versions of many browsers.
-Additional software is not needed and only requires javaScript to be enabled on the browser to run applications.
-CCWebPortal is, at least, compatible with the following browsers: Internet Explorer 6+, Firefox 3.6+ (PC, Mac), Safari 4+, Chrome 10+ and Opera 11+ (PC, Mac).
-
-CCWebPortal is implemented following the client$-$server distributed programming model.
-The user connects to the website and a set of javaScript files are downloaded and executed on the local machine.
-Subsequently, the client application connects to a remote server that enables the execution of calculations and transactions through a centralized database management system.
-The web application is deployed on the client side using ExtJs libraries.
-ExtJs is a javaScript development framework to implement interactive Web applications using AJAX, DHTML and DOM technologies.
-ExtJS has been widely adopted by large technology projects worldwide.
-Its success lies in its ability to design and implement web interfaces with a clear and flexible functionality for the end user.
-ExtJs is used under the GNU license GPL v3.
-
-The CCWebPortal website communicates with a remote server that hosts a MySQL database.
-The database runs on a MyISAM engine that optimizes the speed of access to records in a transactional web-based environment.
-A set of relational tables allows the structuring of the metagenomics repository to establish relationships between records.
-Thus the search and information threshing is optimized for queries launched from the client interface.
-Access to the database on the server is implemented through Django framework.
-Django is an open-source framework written in Python using the model-view-controller (MVC) architectural pattern for implementing user interfaces, which eases the development and updating of code.
-This framework is used to implement web applications on the server side and allows flexible and modular developments and access to various types of database managers.
-Django runs as a web server, awaiting requests from the client CCWebPortal application.
-Once a user is authorized to perform a given operation, Django interacts with the database serving queries to the client application.
-
-\section{Computing kernel}
-
-The effective data analysis has been performed with a Python tool developed from scratch following the OOP\footnote{Object Oriented Programming, but OOP also stands for Object Oriented Python, which is applicable here too.} paradigm, called \CC.
-This software is the back-end of the website described in the previous subsection.
-However, it could be run as an independent piece of software since it is built as a Python package provided with a command-line front-end (\CC.py).
-Once installed, the tool can be run interactively (using the flag \texttt{-i}/\texttt{--interactive}) but also in automatic mode (flag \texttt{-a}/\texttt{--automatic} followed by the desired graphic format, e.g., \texttt{pdf}, \texttt{png}, \texttt{svg}, \texttt{ps}), which uses parallel computation to speed up the analysis of several data sources.
-
-The external mandatory dependencies of \CC\ are with Python packages \emph{pandas}, \emph{matplotlib}, \emph{numpy}, \emph{scipy} and \emph{xlrd}, while the recommended ones are \emph{pylatex} and \emph{uncertainties}.
-The PEP8 compliant code of \CC\ could be executed under a Python 2.7 interpret, but a 3.X interpret is strongly advised.
-The complete distribution of \CC\ has the code signature shown in the Table \ref{tab:cloc}, obtained through the \emph{CLOC}\footnote{\emph{Count Lines of Code} is a perl script that counts blank lines, comment lines, and physical lines of source code.} tool release 1.62.
-Figure \ref{fig:CC} shows the code layout of the complete distribution of \CC.
-
-\begin{table}
-  \begin{center}
-    \begin{tabular}{lrrrrcrcr}
-    \hline
-	Language & files & blank & comment & code & x & scale & = & 3rd gen. equiv \\
-	\hline
-	Python & 13 & 384 & 666 & 4222 &x& 4.20 &=& 17732.40 \\
-	R & 1 & 11 & 38 & 192 &x& 3.00 &=& 576.00 \\	
-	\hline
-	SUM: & 14 & 395 & 704 & 4414 &x& 4.15 &=& 18308.40 \\
-	\hline
-    \end{tabular}
-  \end{center}
-  \caption{Line counting for the complete distribution of \CC\ by \emph{CLOC}}
-  \label{tab:cloc}
-\end{table}
-
-\begin{figure}
-  \centering
-	\[ \text{\task{ComplexCruncher}}
-	\begin{cases}
-		\text{\task{cmplxcruncher.py}} \\
-		\\ 
-		\text{\task{cmplxcruncher}}
-		\begin{cases}
-			\text{\task{\_\_init\_\_.py}} \\ 
-			\\ 
-			\text{\task{models.py}} \\
-			\\ 
-			\text{\task{plots.py}} \\
-			\\ 
-			\text{\task{sessions.py}} \\ 
-		\end{cases}
-		\\ \\
-		\text{\task{tools}}
-		\begin{cases}
-			\text{\task{power\_analysis.R}} \\
-			\\ 
-			\text{\task{rawlmat2lmat.py}} \\
-			\\ 
-			\text{\task{lmat2cmplx.py}} \\
-			\\ 
-			\text{\task{pyLMAT\_rl.py}} \\
-			\\ 
-			\text{\task{pyLMAT\_cs.py}} \\
-			\\ 
-			\text{\task{pyLMAT\_gl.py}} \\
-			\\
-			\text{\task{LMAT}}
-			\begin{cases}
-	    		\text{\task{pyLMAT\_rescore.py}} \\ 
-	    		\\ 
-	    		\text{\task{HMPspiker.py}} 
-			\end{cases}
-		\end{cases}
-	\end{cases} \]
-	\caption{\CC\ layout}
-	\label{fig:CC}
-\end{figure}
-
-\CC\ performs the power-law fit described in the \emph{Blumm, N.
-et al.} paper, but by fitting the best model, i.e.
-choosing between fitting a power-law using linear regression versus nonlinear regression\cite{ecology}.
-In the power-law fit plots we also show the generalized coefficient of determination computed for continuous models\cite{genR2,disR2}.
-
-\section{Un-weighted power-law fit}\label{sec:unw}
-
-\subsection{Fitting the best model}
-
-As mentioned above, to choose between fitting power laws ($y=Vx^\beta$) using linear regression on log-transformed (LLR) data versus non-linear regression (NLR), we mainly follow \emph{General Guidelines for the Analysis of Biological Power Laws}\cite{ecology}.
-It consists of the following three steps:
-\begin{enumerate}
-    \item Determining the appropriate error structure by likelihood analysis.
-    \begin{enumerate}
-        \item Fit the Non-Linear Regression (NLR) model and obtain $V_\text{NLR}$, $\beta_\text{NLR}$ and $\sigma_\text{NLR}^2$.
-        \item Calculate the loglikelihood that the data ($n$ is sample size) are generated from a normal distribution with additive error:
-        \begin{itemize}
-            \item The likelihood of a normal distribution is:  
-            $$\mathcal{L}_\text{norm} = \prod_{i=1}^n\left[\frac{1}{\sqrt{2\pi\sigma^2_\text{NLR}}}\;\exp{\left(-\frac{\left(y_i-V_\text{NLR}x_i^{\beta_\text{NLR}}\right)^2}{2\sigma^2_\text{NLR}}\right)}\right]$$
-            \item So, the loglikelihood of a normal distribution is:
-            \begin{eqnarray*}
-                \log\mathcal{L}_\text{norm} &=& -\frac{n}{2}\log\left|2\pi\sigma^2_\text{NLR}\right| - \frac{1}{2\sigma^2_\text{NLR}}\underbrace{\sum_{i=1}^n\left(y_i-V_\text{NLR}x_i^{\beta_\text{NLR}}\right)^2}_{\mathrm{RSS
-                }_\text{NLR}}\\
-                &=& -\frac{n}{2}\log\left|2\pi\sigma^2_\text{NLR}\right|-\frac{\mathrm{RSS}_\text{NLR}}{2\sigma^2_\text{NLR}}
-            \end{eqnarray*}
-        \end{itemize}
-        \item Calculate the \emph{corrected Akaike's Information Criterion} for the NLR model:
-        $$\mathrm{AIC_{c_{NLR}}} = 2k - 2\log\mathcal{L}_\text{norm} + \frac{2k(k+1)}{n-k-1}$$
-        \item Fit the Log-transformed Linear Regression (LLR) model and obtain $V_\text{LLR}$, $\beta_\text{LLR}$ and $\sigma_\text{LLR}^2$.
-        \item Calculate the loglikelihood that the data ($n$ is sample size) are generated from a lognormal distribution with multiplicative error:
-        \begin{itemize}
-            \item The likelihood of a lognormal distribution is: 
-            $$\mathcal{L}_\text{logn} = \prod_{i=1}^n\left[\frac{1}{y_i\sqrt{2\pi\sigma^2_\text{LLR}}}\;\exp{\left(-\frac{\left(\log|y_i|-\log|V_\text{LLR}|-\beta_\text{LLR}\log|x_i|\right)^2}{2\sigma^2_\text{LLR}}\right)}\right]$$
-            \item So, the loglikelihood of a lognormal distribution is: 
-            \begin{eqnarray*}
-                \log\mathcal{L}_\text{logn} &=& -\frac{n}{2}\log\left|2\pi\sigma^2_\text{LLR}\right| - \sum_{i=1}^n\log|y_i| -\\
-                &&\qquad-\frac{1}{2\sigma^2_\text{LLR}}\underbrace{\sum_{i=1}^n\left(\log|y_i|-\log|V_\text{LLR}|-\beta_\text{LLR}\log|x_i|\right)^2}_{\mathrm{RSS
-                }_\text{LLR}}\\
-                &=& -\frac{n}{2}\log\left|2\pi\sigma^2_\text{LLR}\right| - \frac{\mathrm{RSS}_\text{LLR}}{2\sigma^2_\text{LLR}} - \sum_{i=1}^n\log|y_i|
-            \end{eqnarray*}
-        \end{itemize}
-        \item Calculate the \emph{corrected Akaike's Information Criterion} for the LR model: 
-        	$$\mathrm{AIC_{c_{LLR}}} = 2k - 2\log\mathcal{L}_\text{logn} + \frac{2k(k+1)}{n-k-1}$$
-    \end{enumerate}
-    \item Compare $\mathrm{AIC_{c_{NLR}}}$ with $\mathrm{AIC_{c_{LLR}}}$:
-    \begin{itemize}
-        \item If $\mathrm{AIC_{c_{NLR}}} - \mathrm{AIC_{c_{LLR}}} < -2$, the assumption of normal error is favoured compared to lognormal error, so proceed with the results obtained from the NLR fit.
-        \item If $\mathrm{AIC_{c_{NLR}}} - \mathrm{AIC_{c_{LLR}}} > 2$, the assumption of lognormal error is favoured compared to normal error, so proceed with the results obtained from the LLR fit.
-        \item If $\left|\mathrm{AIC_{c_{NLR}}} - \mathrm{AIC_{c_{LLR}}}\right| \leq 2$, no model is favoured, so proceed with model averaging:
-        \begin{eqnarray*}
-            B_\text{av} &=& w_\text{NLR}V_\text{NLR} + w_\text{LLR}V_\text{LLR} \\
-            \beta_\text{av} &=& w_\text{NLR}\beta_\text{NLR} + w_\text{LLR}\beta_\text{LLR}
-        \end{eqnarray*}
-        where: 
-        \begin{eqnarray*}
-            w_\text{NLR} &=& \frac{1}{1+\mathrm{e}^{\frac{1}{2}\left(\mathrm{AIC_{c_{NLR}}}-\mathrm{AIC_{c_{LLR}}}\right)}} \\
-            w_\text{LLR} &=& \frac{1}{1+\mathrm{e}^{\frac{1}{2}\left(\mathrm{AIC_{c_{LLR}}}-\mathrm{AIC_{c_{NLR}}}\right)}}
-        \end{eqnarray*}
-        which are obtained to fulfill the next condition: $w_\text{NLR} + w_\text{LLR} = 1$.
-The CIs for $B_\text{av}$ and $\beta_\text{av}$ are to be generated by ordinary bootstrapping\footnote{\CC\ has available the next bootstrapping alternatives\cite{boot}: ordinary, ``Resampling Residuals'' method, ``Wild'' method, and ``Monte-Carlo'' method.}.
-    \end{itemize}
-    \item Assess the validity of the underlying statistical assumptions with diagnostic plots because while it is rare for all the assumptions to be fully satisfied by real-life data sets, major violations indicate the lack of appropriateness of the model and, thus, the potential invalidity of the results.
-\end{enumerate}
-
-\subsection{Calculating the coefficient of determination}
-
-We think the best approach in this situation is to apply the generalized $R^2$ that, for continuous models, was defined as\cite{genR2}:
-$$ R^2 = 1 - \left(\frac{\mathcal{L}(0)}{\mathcal{L}(\hat\theta)}\right)^{\!\!\frac{2}{n}} $$
-where $\mathcal{L}(\hat\theta)$ and $\mathcal{L}(0)$  denote the likelihoods of the fitted and the ``null'' model, respectively, and $n$ is the sample size.
-In terms of the loglikelihoods, the generalized coefficient of determination would be:
-$$ R^2 = 1 - \mathrm{e}^{-\frac{2}{n}\left(\log\mathcal{L}(\hat\theta)-\log\mathcal{L}(0)\right)} $$
-We have the likelihoods calculated from the previous section, but what about the ``null'' models?
-We understand that they are the models with only the intercept.
-So for the Gaussian additive error model:
-$$\mathcal{L}_\text{norm}(0) = \prod_{i=1}^n\left[\frac{1}{\sqrt{2\pi\sigma^2_\text{NLR0}}}\;\exp{\left(-\frac{\left(y_i-\bar y\right)^2}{2\sigma^2_\text{NLR0}}\right)}\right]$$
-So:
-\begin{eqnarray*}
-\log\mathcal{L}_\text{norm}(0) &=& -\frac{n}{2}\log\left|2\pi\sigma^2_\text{NLR0}\right| - \frac{1}{2\sigma^2_\text{NLR0}}\sum_{i=1}^n\left(y_i-\bar y\right)^2\\
- &=& -\frac{n}{2}\left(\log\left|2\pi\sigma^2_\text{NLR0}\right| + 1 \right)
-\end{eqnarray*}
-since $\sigma^2_\text{NLR0}=\frac{1}{n}\sum\left(y_i-\bar y\right)^2=\frac{1}{n}\mathrm{TSS}_\text{NLR}$. Now, coming back to the coefficient of determination, we have:
-\begin{eqnarray*}
-R^2_\text{NLR} &=& 1 - \mathrm{e}^{\frac{2}{n}\left(\log\mathcal{L}_\text{NLR}(0)-\log\mathcal{L}_\text{NLR}(\hat\theta)\right)} = 1 - \exp{\left(\frac{\log(\mathrm{RSS}_\text{NLR})}{\log(\mathrm{TSS}_\text{NLR})}\right)}
- =\\ &=& 1 - \frac{\mathrm{RSS}_\text{NLR}}{\mathrm{TSS}_\text{NLR}} =
- 1 - \frac{\sum_{i=1}^n\left(y_i-V_\text{NLR}x_i^{\beta_\text{NLR}}\right)^2}{\sum_{i=1}^n\left(y_i-\bar y\right)^2}
-\end{eqnarray*}
-recovering the traditional expression for $R^2$. Using the same approach for calculating $R^2_\text{LLR}$, then:
-$$\mathcal{L}_\text{logn}(0) = \prod_{i=1}^n\left[\frac{1}{y_i\sqrt{2\pi\sigma^2_\text{LLR0}}}\;\exp{\left(-\frac{\left(\log|y_i|-\log|B_\text{LLR0}|\right)^2}{2\sigma^2_\text{LLR0}}\right)}\right]$$
-So:
-\begin{eqnarray*}
-\log\mathcal{L}_\text{logn}(0) &=& -\frac{n}{2}\log\left|2\pi\sigma^2_\text{LLR0}\right| - \frac{1}{2\sigma^2_\text{LLR0}}\sum_{i=1}^n\left(\log|y_i|-\overline{\log|y|}\right)^2 - \sum_{i=1}^n\log|y_i|\\
- &=& -\frac{n}{2}\left(\log\left|2\pi\sigma^2_\text{LLR0}\right| + 1 \right) - \sum_{i=1}^n\log|y_i|
-\end{eqnarray*}
-since $\sigma^2_\text{LLR0}=\frac{1}{n}\sum\left(\log|y_i|-\overline{\log|y|}\right)^2=\frac{1}{n}\mathrm{TSS}_\text{logn}$.
-Again, recalling the expression for the generalized coefficient of determination, we have:
-\begin{eqnarray*}
-R^2_\text{LLR} &=& 1 - \mathrm{e}^{\frac{2}{n}\left(\log\mathcal{L}_\text{LLR}(0)-\log\mathcal{L}_\text{LLR}(\hat\theta)\right)} = 1 - \exp{\left(\frac{\log(\mathrm{RSS}_\text{LLR})}{\log(\mathrm{TSS}_\text{LLR})}\right)}
- =\\ &=& 1 - \frac{\mathrm{RSS}_\text{LLR}}{\mathrm{TSS}_\text{LLR}} =
- 1 - \frac{\sum_{i=1}^n\left(\log|y_i|-\log|V_\text{LLR}|-\beta_\text{LLR}\log|x_i|\right)^2}{\sum_{i=1}^n\left(\log|y_i|-\overline{\log|y|}\right)^2}
-\end{eqnarray*}
-
-\section{X-weighted power-law fit}\label{sec:X-w}
-
-When fitting the power-law of std vs. mean, we can take into account that every mean has uncertainty and estimate it for a sample size $n$ by the SEM (\emph{Standard Error of the Mean}):
-$$ \mathrm{SEM} = \frac{s}{\sqrt{n}}$$
-where $s$ is the sample standard deviation. So, the vector of weights is computed with:
-$$ \mathbf{w} = \frac{1}{\overrightarrow{\mathrm{SEM}}} = \frac{\sqrt{\mathbf{n}}}{\mathbf{s}}$$
-
-Here, the uncertainties affect the independent variable, so the fit is not so trivial as a Y-weighted fit, where the uncertainties affect the dependent variable.
-A standard approach to do this fit is: a) invert your variables before applying the weights, b) then perform the weighted fit, and finally, c) revert the inversion.
-This method is deterministic, but the approximate solution worsens with smaller $R^2$.
-For comparison, we develop a stochastic method by using a bootstrapping-like strategy that avoids the inversion and is applicable regardless of $R^2$.
-Both methods, detailed below, are implemented in \CC.
-
-\subsection{Method 1: By inverting the data}
-
-In the case of the log-LR model, we have:
-$$\log y = \log V + \beta\log x \quad\rightarrow\quad \underbrace{\log x}_{\tilde y} = \overbrace{-\frac{1}{\beta}\log V}^{b} + \overbrace{\frac{1}{\beta}}^{m}\underbrace{\log y}_{\tilde x}$$ 
-where $m$ determines the slope or gradient of the fitted line, and $b$ determines the point at which the line crosses the y-axis, otherwise known as the y-intercept.
-Once the model is fitted, the original parameters can be retrieved easily:
-\begin{eqnarray*}
-\beta &=& \frac{1}{m} \\
-V &=& \mathrm{e}^{-\beta b} = \mathrm{e}^{-\frac{b}{m}}
-\end{eqnarray*}
-Their respective uncertainties are to be obtained using \emph{error propagation}:
-\begin{eqnarray*}
-    \sigma_\beta &=& \left|\frac{\mathrm{d}\beta}{\mathrm{d}m}\right|\sigma_m \quad=\quad \frac{1}{m^2}\;\sigma_m \\
-    \sigma_V &=& \sqrt{\left(\frac{\partial V}{\partial b}\right)^{\!\!2}\sigma_b^2 +
-      \left(\frac{\partial V}{\partial m}\right)^{\!\!2}\sigma_m^2} \quad=\quad
-      \frac{1}{m}\;\mathrm{e}^{-\frac{b}{m}}\,\sqrt{\sigma_b^2 + \frac{b^2}{m^2}\;\sigma_m^2} 
-\end{eqnarray*}
-
-\subsection{Method 2: Bootstrapping-like strategy}
-
-The basic idea of bootstrapping is that inference about a population from sample data (sample $\rightarrow$ population) can be modeled by resampling the sample data and performing inference on (resample $\rightarrow$ sample).
-To adapt this general idea to our problem, we resample the x-data array using its errors array.
-That is, for each replicate, a new x-data array is computed based on:
-$$x^*_i = x_i + v_i$$
-where $v_i$ is a Gaussian random variable with mean $\mu_i=0$ and standard deviation $\sigma_i=\mathrm{SEM}_i$, as defined previously in this supplementary material.
-For each replicate a complete un-weighted power-law fit is performed, as described in the previous section.
-It is worth mentioning that each replicate is filtered to avoid values of $x^*_i$ under \emph{eps} (obtained by \texttt{np.finfo(np.double).eps}) in order to keep away from the error of getting log of negatives or zero during the fit.
-
-We devised and implemented a multi-step algorithm to estimate the fit parameters that finishes when a relative error of less than $10^{-4}$ is achieved.
-It also ends if the number of steps reaches $100$ to avoid too much time lapse, to prevent any pathologic numeric case which, in fact, we still have not detected in all the data sets analyzed.
-
-In the previous version of the algorithm, for each step, the method generated $10$ replicates for each x-data point, in other words, it was computing the fit for $10$ times the length of the x-data array replicates, with a maximum of $10000$ fits per step.
-Nevertheless, we found that such an approach depending on the length of the x-data array did not perform better, so we decided to simplify the method and fix the number of fits per step in $100$.
-This latter approach improved the performance.
-
-The parameters of the X-weighted fit are then estimated by averaging through all the replicate fits performed, and their errors are estimated by computing the standard deviation also for all the fits.
-At the end of each step, the relative error is calculated by comparing the fit parameters estimation in the last step with the previous one.
-
-Finally, both the coefficient of determination of the fit and the coefficient of correlation between the fit parameters are estimated by averaging.
-
-\section{Rank Stability Index (RSI)}\label{sec:RSI}
-
-The Rank Stability Index is shown as a percentage in a separate bar on the right of the rank matrix plot provided by \CC.
-The RSI is strictly $1$ for an element whose range never changes over time, and is strictly $0$ for an element whose rank oscillates between the extremes from time to time.
-So, RSI is calculated, per element, as $1$ less the quotient of the number of true rank hops taken between the number of maximum possible rank hops, all powered to $p$:
-$${\rm RSI} = \left(1-\frac{\text{true rank hops}}{\text{possible rank hops}}\right)^p = \left(1-\frac{D}{(N-1)(t-1)}\right)^p$$
-where $D$ is the total of rank hops taken by the studied element, $N$ is the number of elements that have been ranked, and $t$ is the number of time samples.
-The power index $p$ is arbitrarily chosen to increase the resolution in the stable region; the value in the current version of the code is $p=4$.
-
-As an example of this ``zooming'' effect in the stable region, to match a linear ($p=1$) RSI of 0.9 to a powered one of 0.1, we should select $p=21.8543$.
-An alternative way to obtain this effect and exactly map a linear RSI of 0.9 to a non-linear RSI (${\rm RSI'}$) of 0.1, is by applying the following function:
-$${\rm RSI'} = \frac{10^{10\left(1-\frac{D}{(N-1)(t-1)}\right)}-1}{10^{10}-1} \approx 10^{-10\left(\frac{D}{(N-1)(t-1)}\right)} $$
-where the approximation is valid because $10^{10}\gg 1$ but, the small price to pay for it is that, in the worst instability case, the ${\rm RSI'}$ would not be strictly $0$ but $10^{-10}$.
-
-The colour code of the RSI percentage text in the rank plot of \CC\ is chosen following the first condition satisfied from those shown in Table \ref{tab:RSI} (see page \pageref{tab:RSI}).
-
-\begin{table}
-  \begin{center}
-    \begin{tabular}{ccc}
-    \hline
-    Case  &  Condition  &  Colour  \\
-    \hline
-    1  &  $1\ge{\rm RSI}>0.99$  & \textcolor{blue}{blue}  \\ 
-    2  & ${\rm RSI}>0.90$  &  \textcolor{green}{green}  \\
-    3  &  ${\rm RSI}>0.75$  &  \textcolor{orange}{orange} \\
-    4  &  ${\rm RSI}>0.25$  &  \textcolor{red}{red} \\
-    5  &  $0.25\ge{\rm RSI}\ge0$  &  \bfseries{black}  \\
-    \hline
-    \end{tabular}
-  \end{center}
-  \caption{Colour code of the RSI percentage text shown in rank plots, following the first condition satisfied.}
-  \label{tab:RSI}
-\end{table}
-
-\section{\CC\ output}
-In the previous sections, we have discussed details about the methods used in \CC.
-In this section we review the output from the package, which aims at summarizing the as yet undescribed functionality.
-Unless the flag \texttt{-n}/\texttt{--nodirs} is given to the \CC.py front-end script, the output will be spread into subdirectories (\texttt{corrank}, \texttt{fits}, \texttt{hist} and \texttt{other}) under the results directory.
-The path of the results directory is relative to the data directory, and is given by the optional argument \texttt{-r}/\texttt{--results}, which defaults to ``\texttt{results}''.
-A bunch of plots distributed in subdirectories of such path is generated by \CC\ (the flag that is reducing such number to the minimum is \texttt{-l}/\texttt{--less}).
-They are described in the next subsections.
-In addition, the code provides with output in CSV format (one file per input file), in Excel format (one summary spreadsheet with one sheet per input file, plus correlations and rank tables in the \texttt{corrank} subdirectory) and, optionally, in \LaTeX\ and PDF format (thanks to pdf\LaTeX).
-
-\subsection{Fit Plots}
-The subdirectory \texttt{fits} under the results directory contains the fit plots for each data set.
-Both for the unweighted fit (detailed in the section \ref{sec:unw}) and for the X-weighted fit (detailed in the section \ref{sec:X-w}), two plots are generated: the former with logarithmic scale, the latter with lineal scale.
-Figure \ref{fig:unwFit} shows an example of the unweighted fit plots and Figure \ref{fig:X-wFit} does the same with the X-weighted fit plots.
-Additionally, for the unweighted fit, a complete residues analysis is performed, and a 4-in-1 figure is generated as shown in Figure \ref{fig:unwRes}, corresponding to the fit of Figure \ref{fig:unwFit}.
-Among other tests, it allows to check for normality and homoscedasticity of the residues.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_LOG.pdf}
-	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_LIN.pdf}
-	\caption{Example of unweighted fit, shown both in logarithmic and lineal scale plots}
-	\label{fig:unwFit}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_xWboot_LOG.pdf}
-	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_xWboot_LIN.pdf}
-	\caption{The X-weighted fit log and lineal plots corresponding to the fit of Figure \ref{fig:unwFit}}
-	\label{fig:X-wFit}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_LLR_RES.pdf}
-	\caption{Residues analysis plot corresponding to the fit of Figure \ref{fig:unwFit}.
-The top-left subplot is a simple residues plot.
-The top-right subplot is a Normal quantiles plot with linear fitting (value of coefficient of determination is provided).
-The bottom-left subplot shows an accumulated ESS (Explained Sum of Squares) plot.
-Finally, the bottom-right subplot is a residues Normal histogram plot.
-This set of subplots allows to check for normality and homoscedasticity of the residues.}
-	\label{fig:unwRes}
-\end{figure}
-
-\subsection{Histogram Plots} 
-The subdirectory \texttt{hist} under the results directory contains the histogram plots. \CC\ generates three different histogram plots:
-\begin{description}
-	\item[$-$Absolute frequencies plot]: This plot is useful to visually assess the validity of the time points in terms of the accumulated absolute frequency of the elements (taxa), since absolute frequencies far (much higher or much lower) from those typically observed could mean a sampling problem.
-As an example, Figure \ref{fig:histAFP} shows this histogram for the pre-treatment data (first 7 times) of patient ``D'' in the antibiotics study\cite{antibiotic}.
-	\item[$-$2D deviation plot]: The 2D semi-logarithmic histogram representing deviations from the mean versus the mean itself, is a useful tool in the analysis of the stability of ranking processes in complex systems\cite{ranking}.
-Figure \ref{fig:hist2D} shows this plot for the data used in the fit shown in Figure \ref{fig:unwFit}.
-	\item[$-$Zero relative frequency plot]: We could define the ZRF (Zero Relative Frequency, thereby ranging from 0 to 1) of an element (taxon) as the portion of times where it is zero, i.e., it is not found.
-Attending to all the elements (taxa), we can plot the ZRF histogram, which then lies on the horizontal axis of the plot.
-The vertical axis shows the number of elements (taxa), so the height of a bar represents the number of elements that have determinate ZRF.
-In this respect, the bar over $0.0$ counts the number of elements (taxa) that are present at every time point of the data set (aka ``core''), while the bar over $1.0$ would count the number of elements (taxa) that are never found (this bar never appears because all these ``null'' elements are automatically filtered by the code).
-Figure \ref{fig:histZRF} shows an example of this plot.
-There, we can see that $12$ taxa are present at all the time points of the time series while $9$ taxa basically appear only once.
-So, this plot is clearly useful to notice how the ``core'' is distributed.
-\end{description}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.8\textwidth]{results/hist/antibiotics_h_patD_pre_AbsFreqPlot}
-	\caption{Histogram with the absolute frequencies of the pre-treatment data (7 first times) of patient ``D'' in the antibiotics study\cite{antibiotic}}
-	\label{fig:histAFP}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.8\textwidth]{results/hist/IBS_h_A_amplicons_family_hist2D.pdf}
-	\caption{2D histogram deviation plot of the data used in the fit shown in Figure \ref{fig:unwFit}}
-	\label{fig:hist2D}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.8\textwidth]{results/hist/IBS_h_A_amplicons_family_ZRFhist.pdf}
-	\caption{Histogram with the relative frequency of zero for the elements (taxa) in the data used in the fit shown in Figure \ref{fig:unwFit}}
-	\label{fig:histZRF}
-\end{figure}
-
-\subsection{Correlation and Rank Plots} 
-The subdirectory \texttt{corrank} under the results directory contains the correlation matrix and rank\&stability plots, as well as Excel files with the resulting matrices.
-\CC\ generates three different plots falling under this category:
-\begin{description}
-	\item[$-$Elements correlation matrix]: This plot shows a correlation matrix among the elements (taxa), calculated with the time as independent variable.
-For these calculations, the data set is not normalized to avoid entering an additional constraint.
-As an example, Figure \ref{fig:corrElm} shows this matrix for the ``core'' elements (taxa) present in the pre-treatment data (first seven times) of patient ``D'' in the antibiotics study\cite{antibiotic}.
-	\item[$-$Times correlation matrix]: This plot presents a correlation matrix among the time points of the data set, calculated with the elements (taxa) as independent variable.
-Again, the data set is not normalized.
-Figure \ref{fig:corrTim} shows this matrix for the ``core'' elements (taxa) present in pre-treatment data (first seven times) of patient ``D'' in the antibiotics study\cite{antibiotic}.
-	\item[$-$Rank dynamics and stability plot]: This plot shows the variation in the rank with time for the most dominant elements (taxa) and their calculated RSI, as discussed in Section \ref{sec:RSI}.
-Figure \ref{fig:corrank} shows this plot for the elements (taxa) in the data used in the fit shown in Figure \ref{fig:unwFit}.
-\end{description}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\textwidth]{results/corrank/antibiotics_h_patD_pre_CORE_ElementsCorr}
-	\caption{Element correlation plot of the pre-treatment data (first seven times) of patient ``D'' in the antibiotics study\cite{antibiotic} for ``core'' taxa}
-	\label{fig:corrElm}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\textwidth]{results/corrank/antibiotics_h_patD_pre_CORE_TimesCorr}
-	\caption{Times correlation plot of the pre-treatment data (first seven times) of patient ``D'' in the antibiotics study\cite{antibiotic} for ``core'' taxa}
-	\label{fig:corrTim}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\textwidth]{results/corrank/IBS_h_A_amplicons_family_Rank}
-	\caption{Matrix showing the rank variation throughout time for the most dominant elements (taxa) and their calculated RSI (as discussed in Section \ref{sec:RSI}) in the data used in the fit shown in Figure \ref{fig:unwFit}}
-	\label{fig:corrank}
-\end{figure}
-
-\subsection{Other Plots} 
-The subdirectory \texttt{other} under the results directory contains other plots, like the elements (taxa) kurtosis and skewness against the mean.
-These plots are generated when the flag \texttt{-m}/\texttt{--more} is given to the \CC.py front-end script.
-
-\subsection{\CC\ advanced options}
-The advanced option flags are summarized here:
-\begin{description}
-	\item[$-$ZRF filter]: This is activated with \texttt{-c}/\texttt{--core} and allows global and detached results for ``core'', ``tailless'' and ``tail'', which allows comparative analysis depending on the presence/absence rate of the taxa throughout time, which is calculated in practice using the Zero Relative Frequency (ZRF) defined above.
-The criteria is as follows:
-	\begin{itemize}
-		\item In any case, if $\operatorname{ZRF} < 0.2$ taxa are classified as ``core''. In other words, the taxa considered ``core'' are those appearing at over $80\%$ of the time points under analysis.
-		\item Depending on the number of times:
-		\begin{itemize}
-			\item If there are only 3 time points, taxa with $0.2 \le\operatorname{ZRF} < 0.4$ are classified as ``tailless'' and, therefore, taxa with $\operatorname{ZRF} \ge 0.4$ are classified as ``tail''.
-			\item If there are 4 time points, taxa with $0.2 \le\operatorname{ZRF} < 0.5$ are classified as ``tailless'' and, therefore, taxa with $\operatorname{ZRF} \ge 0.5$ are classified as ``tail''.
-			\item If there are 5 or more time points, taxa with $0.2 \le\operatorname{ZRF} < 0.6$ are classified as ``tailless'' and, therefore, taxa with $\operatorname{ZRF} \ge 0.6$ are classified as ``tail''.
-		\end{itemize}
-    \end{itemize}
-	\item[$-$Relative frequency filters]: The minimum frequency filter is activated with \texttt{-z}/\texttt{--fmin} and the maximum one with \texttt{-x}/\texttt{--fmax}, with values between $0$ and $1$.
-If both filters are activated simultaneously, the minimum frequency cannot be equal or higher than the maximum one.
-If no filter is activated, by default, \CC\ is filtering all the taxa that never appear throughout time, in the event there are such taxa among the input data.
-On the one hand, if a minimum frequency value is provided, all the taxa whose relative accumulated frequencies (the normalized sum of the values through time) are under such a value, are eliminated.
-But in addition, once the set of data is normalized (to work with relative frequencies), if there are data at any time points that are under the minimum frequency given value, they are automatically floored to zero (we call this \emph{item@time filtering}), and the set of data is renormalized.
-On the other hand, if a maximum frequency value is provided, all the taxa whose relative accumulated frequencies are over such value, are deleted.
-\end{description}
-
-\section{\CC\ input}
-The input data are provided by two simultaneously available ways:
-\begin{enumerate}
-	\item Text files (\texttt{txt} extension), to be processed in parallel\footnote{Currently, parallelization in \CC\ is only available under GNU/Linux}, with a shared-memory model.
-	\item Excel files (\texttt{xls[x]} extension), whose sheets are related datasets to be processed in parallel\footnotemark[\value{footnote}].
-Optionally, all the sheets of an Excel file are standardized using part of such Excel datasets, following the procedure described in the Supplementary Section \ref{sec:stan}, in which event, additional output is provided by \CC.
-\end{enumerate}
-Typically, the data sets are tables with the taxa lying in different rows and the time points ordered in different columns.
-So, the first column is considered the taxon labels, whereas the first row contains the time labels.
-It is worth mentioning that any Excel sheet or text file beginning with the underscore character is ignored by \CC, as they are supposed to hold metadata.
-
-\section{Standardization} \label{sec:stan}
-In order to show all the studies properly under common axes, we decided to standardize the Taylor parameters using the group of healthy individuals for each study.
-With this approach, all the studies can be visualized in a shared plot with units of Taylor-parameters standard-deviation on their axes.
-
-For a Taylor parameter, e.g. $V$, the estimate of the mean ($\widehat{V}$) for the healthy subpopulation, composed of $h$ individuals, is:
-$$\widehat{V} = \frac{1}{W_1}\sum_{i=1}^h V_i \omega_i=\sum_{i=1}^h V_i \omega_i$$
-as $W_1=\sum_i^h \omega_i=1$, since $\omega_i$ are normalized weights calculated as:
-$$\omega_i = \frac{\frac{1}{\sigma^2_{V_i}}}{\sum_i^h\frac{1}{\sigma^2_{V_i}}}$$
-being $\sigma_{V_i}$ the estimation of the uncertainty in $V_i$ obtained together with $V_i$ from the X-weighted power-law fit described in Section \ref{sec:X-w}, for healthy individuals.
-
-Likewise, the estimation of the standard deviation for the healthy population ($\widehat{\sigma}_V$) is:
-$$\widehat{\sigma}_V = \sqrt{\frac{1}{W_1-\frac{W_2}{W_1}}\sum_{i=1}^h\left[\omega_i\left(V_i-\hat{V}\right)^2\right]}$$
-being $W_2=\sum_i^h \omega_i^2$, which finally yields to:
-$$\widehat{\sigma}_V = \sqrt{\frac{1}{1-\sum_i^h \omega_i^2}\sum_{i=1}^h\left[\omega_i\left(V_i-\hat{V}\right)^2\right]}$$
-
-\section{\CC\ interface with LMAT} \label{sec:lmat}
-With the aim of adapting LMAT (Livermore Metagenomics Analysis Toolkit) software\cite{LMAT} for its easy use in time series data, like the ones we have analyzed in this paper\cite{kwashiorkor}, we developed Python front and back-end scripts.
-They are supplied in the directory \texttt{tools} of the \CC\ package.
-
-\subsection{LMAT front-ends} They are: \texttt{pyLMAT\_rl.py}, the launcher for LMAT \texttt{run\_rl.sh}, that is, the first step in the LMAT pipeline or \emph{read labeling}; \texttt{pyLMAT\_cs.py}, the launcher for LMAT \texttt{run\_cs.sh}, the \emph{content summarization} step in the LMAT pipeline; and, \texttt{pyLMAT\_gl.py}, the launcher for LMAT \texttt{run\_gl.sh}, the \emph{gene labeling/identification} step in the LMAT pipeline.
-Basically, all three scripts permit to manage the execution of parallelized LMAT against several fasta files coming from diferent time points in time series studies.
-
-\subsection{LMAT back-ends} Currently, they are: \texttt{pyLMAT\_rescore.py} and the pipeline composed by \texttt{rawlmat2lmat.py} and \texttt{lmat2cmplx.py}.
-The first one allows not only to quick recalculate the LMAT results for different cutoff levels, but also to pull out reads with specific NCBI taxId (\emph{human} reads by default).
-For example, issuing "\texttt{pyLMAT\_rescore.py -p path/to/results -f file.fna -m -1 0 1 -t=32}" will regenerate the LMAT results for cutoffs $-1$, $0$ and $1$, and will pull human reads for each cutoff level, parallelizing in $32$ threads.
-On the other hand, the mentioned pipeline enables to post-process the several \emph{fastsummary} output files from LMAT first and second steps launched by the LMAT front-ends described above.
-The pipeline transforms the results so that they achieve a format and content ready for \CC\ computations.
-Such data arrangements are done independently for two different taxonomic levels: genus and specie.
-
-\newpage
-\begin{thebibliography}{99}
-	\bibitem{ranking} Blumm, N. \textit{et al.} Dynamics of ranking processes in complex systems. {\it Phys. Rev. Lett.} {\bf 109,} 128701 (2012).
-	\bibitem{FD} Weber, J. \textit{et al.} Fluctuation dissipation theorem. {\it Phys. Rev.} {\bf 101}, 1620-6 (1956).
-	\bibitem{kwashiorkor} Smith M.I.  \textit{et al.} Gut microbiomes of Malawian twin pairs discordant for kwashiorkor. {\it Science} {\bf 339,} 548-54 (2013).
-	\bibitem{FASTX} Gordon, A., Hannon, G.J. FASTX-Toolkit. FASTQ/A shortreads pre-processing tools (2010). http://hannonlab.cshl.edu/fastx\_toolkit/ (accessed 23 Feb 2015).
-	\bibitem{QIIME} Caporaso, J.G. \textit{et al.} QIIME allows analysis of high-throughput community sequencing data. {\it Nature Methods} {\bf 7,} 335-6 (2010).
-	\bibitem{SILVA} Quast C.  \textit{et al.} The SILVA ribosomal RNA gene database project: improved data processing and web-based tools (2013)
-	\bibitem{LMAT} Ames, S.K.  \textit{et al.} Scalable metagenomic taxonomy classification usng a reference genome database. {\it Bioinformatics} {\bf 29,} 2253-2260 (2013).
-	\bibitem{antibiotic} Dethlefsen L., Relman D.
-A.
+\begin{thebibliography}{9}
+\bibitem{diet} {\bf David LA, Maurice CF, Carmody RN, Gootenberg DB, Button JE, Wolfe BE, Ling A V, Devlin AS, Varma Y, Fischbach MA, Biddinger SB, Dutton RJ, Turnbaugh PJ.} 2014.
+Diet rapidly and reproducibly alters the human gut microbiome.
+Nature {\bf 505}:559–63.
+\bibitem{antibiotic} {\bf Dethlefsen L, Relman DA.} 2011.
 Incomplete recovery and individualized responses of the human distal gut microbiota to repeated antibiotic perturbation.
-{\it Proc.
-Nat.
-Acad.
-Sci.
-USA} {\bf 108,} 4554-61 (2011).
-	\bibitem{diet} David, L.A. \textit{et al.} Diet rapidly and reproducibly alters the human gut microbiome. {\it Nature} {\bf 505,} 559-63 (2014).
-	\bibitem{IBS} Durban, A. \textit{et al.} Structural alterations of faecal and mucosa-associated bacterial communities in irritable bowel syndrome. {\it FEMS Microbiol. Ecol.} {\bf 86}, 581-9 (2013).
-	\bibitem{LEA} Faith, J.J.  \textit{et al.} The long-term stability of the human gut microbiota. {\it Science} {\bf 341,} 1237439 (2013).
-	\bibitem{moving} Caporaso, J.G.  \textit{et al.} Moving pictures of the human microbiome. {\it Genome Biol.} {\bf 12,} R50 (2011).
- 	\bibitem{ecology} Xiao Xiao, Ethan P.
-White, Mevin B.
-Hooten, and Susan L.
-Durham.
-On the use of log-transformation vs.
-nonlinear regression for analyzing biological power laws.
-{\it Ecology} {\bf 92}, 10, 1887-1894 (2011).
- 	\bibitem{genR2} Magee L., $R^2$ measures based on wald and likelihood ratio joint significance tests. {\it The American Statistician} {\bf 44}, 3, 250-253 (1990).
-	\bibitem{disR2} Nagelkerke N.J.D., A note on a general definition of the coefficient of determination. {\it Biometrika} {\bf 78}, 3, 691-692 (1991).
-	\bibitem{boot} Wu, C.F.J. Jackknife, bootstrap and other resampling methods in regression analysis. (with discussions) \textit{The Annals of Statistics} {\bf 14}: 1261?1350 (1986)
+Proc Natl Acad Sci {\bf108}:4554–61.
+\bibitem{IBS} {\bf Durbán A, Abellán JJ, Jiménez-Hernández N, Artacho A, Garrigues V, Ortiz V, Ponce J, Latorre A, Moya A.} 2013.
+Instability of the faecal microbiota in diarrhoea-predominant irritable bowel syndrome.
+FEMS Microbiol Ecol {\bf 86}:581–589.
+\bibitem{kwashiorkor} {\bf Smith MI, Yatsunenko T, Manary MJ, Trehan I, Mkakosya R, Cheng J, Kau AL, Rich SS, Concannon P, Mychaleckyj JC, Liu J, Houpt E, Li J V, Holmes E, Nicholson J, Knights D, Ursell LK, Knight R, Gordon JI.} 2013.
+Gut microbiomes of Malawian twin pairs discordant for kwashiorkor.
+Science {\bf 339}:548–54.
+\bibitem{LEA} {\bf Faith JJ, Guruge JL, Charbonneau M, Subramanian S, Seedorf H, Goodman AL, Clemente JC, Knight R, Heath AC, Leibel RL, Rosenbaum M, Gordon JI.} 2013.
+The long-term stability of the human gut microbiota.
+Science {\bf 341}:1237439.
 \end{thebibliography}
-
 \end{document}

### 2016-09-23 mSystems/mSys_MAIN.tex
mSystems: docs of the initial submission
On branch master
	new file:   CoverLetter.DOC
	new file:   CoverLetter.pdf
	deleted:    MSWord_OBSOLETE/WHOLE_draft_0.1.docx
	deleted:    MSWord_OBSOLETE/discussion_draft_0.1.docx
	deleted:    MSWord_OBSOLETE/intro_draft_0.1.docx
	deleted:    MSWord_OBSOLETE/intro_draft_0.2.docx
	deleted:    MSWord_OBSOLETE/mat_met_draft_0.3.docx
	deleted:    MSWord_OBSOLETE/results_draft_0.1.docx
	deleted:    MSWord_OBSOLETE/skeleton_draft_0.3.docx
	new file:   mSys_MAIN.docx
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_suppl.pdf
	modified:

@@
 % release codenumber (update it here!)
-\newcommand\releasecode{$\beta$}
+\newcommand\releasecode{initial}
 
@@
 % watermarks (time consuming!)
-\newwatermark[allpages,color=Green!4,angle=45,scale=12,xpos=-60,ypos=0]{RC\releasecode}
-\newwatermark[allpages,color=Green!12,angle=0,scale=1,xpos=0,ypos=137]{
-\emph{mSystems} \LaTeX\ manuscript RC\releasecode}
+\newwatermark[allpages,color=Purple!4,angle=45,scale=7,xpos=-40,ypos=0]{submitted}
+\newwatermark[allpages,color=Purple!12,angle=0,scale=1,xpos=0,ypos=137]{
+\emph{mSystems} \LaTeX\ manuscript \releasecode\ submission}
 

### 2016-09-25 mSystems/mSys_MAIN.tex
Minor changes to mSys_MAIN to adapt to mSystems' staff requirements
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_suppl.pdf
	modified:   mSys_suppl.tex

@@
 %%\usepackage{float}
-\usepackage[nolists,tablesfirst, nomarkers]{endfloat}
+\usepackage[nolists, nomarkers]{endfloat}
 \usepackage{moreverb} % for verbatim output
@@
 % watermarks (time consuming!)
-\newwatermark[allpages,color=Purple!4,angle=45,scale=7,xpos=-40,ypos=0]{submitted}
-\newwatermark[allpages,color=Purple!12,angle=0,scale=1,xpos=0,ypos=137]{
-\emph{mSystems} \LaTeX\ manuscript \releasecode\ submission}
+%\newwatermark[allpages,color=Purple!4,angle=45,scale=7,xpos=-40,ypos=0]{submitted}
+%\newwatermark[allpages,color=Purple!12,angle=0,scale=1,xpos=0,ypos=137]{
+%\emph{mSystems} \LaTeX\ manuscript \releasecode\ submission}
 
@@
 \usepackage[labelfont=bf,labelsep=period,font=small]{caption}
+\captionsetup[table]{name=Supplementary Table S\!\!} % See EOF
 
@@
 \end{thebibliography}
+
+%% mSystems staff asked to send the captions of the supplementary tables at the end
+%%  of the main text PDF.
+\begin{table}
+  \caption{Taylor parameters.
+Individuals with either animal-based (A) or plant-based (P) diets\cite{diet}.
+Previous to diet,  the population sampled is described by $\bar{V} = 0.09 \pm 0.05, \bar{\beta} = 0.77 \pm 0.04$.}
+
+\caption{Taylor parameters for individuals taking antibiotics\cite{antibiotic}.
+Prior to antibiotics intake, the population sampled is described by $\bar{V} = 0.12 \pm 0.05, \bar{\beta} = 0.75 \pm 0.04$.}
+
+\caption{Taylor parameters for persons diagnosed with irritable bowel syndrome (IBS)\cite{IBS}.
+Healthy individuals sampled in this study are characterized by $\bar{V} = 0.134 \pm 0.009, \bar{\beta} = 0.691 \pm 0.025$.}
+
+\caption{Taylor parameters for the healthy subject of the discordant twins\cite{kwashiorkor}.
+This table continues in Supplementary Table S\ref{tab:DK}.
+The population of healthy twins is characterized by $\bar{V} = 0.25 \pm 0.10, \bar{\beta} = 0.863 \pm 0.028$.}\label{tab:DH}
+
+\caption{Taylor parameters for the kwashiorkor part of the discordant twins\cite{kwashiorkor}.
+This is a continuation of Supplementary Table S\ref{tab:DH}.
+The population of healthy twins is characterized by $\bar{V} = 0.25 \pm 0.10, \bar{\beta} = 0.863 \pm 0.028$.}\label{tab:DK}
+
+\caption{Taylor parameters for individuals with different degrees of overweight and obesity\cite{LEA}.
+Healthy people in this study, whom were not obese, are characterized by $\bar{V} = 0.19 \pm 0.06, \bar{\beta} = 0.806 \pm 0.034$.}
+\end{table}
+
 \end{document}

### 2016-12-07 mSystems/mSys_suppl.tex
Minor changes to mSys_MAIN to adapt to mSystems' staff requirements
On branch master
	modified:   mSys_MAIN.pdf
	modified:   mSys_MAIN.tex
	modified:   mSys_suppl.pdf
	modified:   mSys_suppl.tex
More minor changes
	modified:   ../mSys_MAIN.pdf
	modified:   ../mSys_MAIN.tex
	modified:   ../mSys_suppl.pdf
	modified:   ../mSys_suppl.tex
	modified:   Response to Reviewer Comments.docx
	modified:   mSys_MAIN_rev_marked.docx

@@
 % release codenumber (update it here!)
-\newcommand\releasecode{initial}
+\newcommand\releasecode{final}
 
@@
 % watermarks (time consuming!)
-\newwatermark[allpages,color=Purple!4,angle=45,scale=7,xpos=-40,ypos=0]{submitted}
-\newwatermark[allpages,color=Purple!12,angle=0,scale=1,xpos=0,ypos=137]{
+\newwatermark[allpages,color=PineGreen!4,angle=45,scale=7,xpos=-40,ypos=0]{submitted}
+\newwatermark[allpages,color=PineGreen!12,angle=0,scale=1,xpos=0,ypos=137]{
 \emph{mSystems} \LaTeX\ manuscript \releasecode\ submission}

### 2016-12-21 mSystems/rev/mSys_MAIN_rev.tex
Add LaTeX version for mSystems revised manuscript and improve figs
Changes to be committed:
	modified:   ../.gitignore
	new file:   figs/Fig2/ellipse.m
	new file:   figs/Fig2/ellipsedata.txt
	new file:   figs/Fig2/errorbarxy.m
	modified:   mSys_MAIN.pdf
	modified:   rev/figs/Fig1.eps
	modified:   rev/figs/Fig2.eps
	modified:   rev/figs/_ed_Fig1.pdf
	renamed:    rev/figs/_ed_FigS3.pdf -> rev/figs/_ed_supfig_corrank_IBS_C.pdf
	renamed:    rev/figs/_ed_FigS4.pdf -> rev/figs/_ed_supfig_corrank_IBS_P1.pdf
	renamed:    rev/figs/_ed_FigS1.pdf -> rev/figs/_ed_supfig_taxlev1.pdf
	renamed:    rev/figs/_

--- 2016-09-25 mSystems/mSys_MAIN.tex
@@
 % release codenumber (update it here!)
-\newcommand\releasecode{initial}
+\newcommand\releasecode{REVISION draft A}
 
@@
 \usepackage{parskip}
-%%\usepackage{float}
-\usepackage[nolists, nomarkers]{endfloat}
 \usepackage{moreverb} % for verbatim output
@@
 % watermarks (time consuming!)
-%\newwatermark[allpages,color=Purple!4,angle=45,scale=7,xpos=-40,ypos=0]{submitted}
-%\newwatermark[allpages,color=Purple!12,angle=0,scale=1,xpos=0,ypos=137]{
-%\emph{mSystems} \LaTeX\ manuscript \releasecode\ submission}
+\newwatermark[allpages,color=Mulberry!4,angle=45,scale=7,xpos=-40,ypos=0]{REVISION}
+\newwatermark[allpages,color=Mulberry!12,angle=0,scale=1,xpos=0,ypos=137]{
+\emph{mSystems} \LaTeX\ manuscript \releasecode}
 
@@
 % word count (Need --enable-write18 or --shell-escape)
-\immediate\write18{texcount -opt=option.tc \jobname.tex > wordcount.aux} 
-\newcommand\wordcount{\verbatiminput{wordcount.aux}}
+\immediate\write18{texcount -opt=option_rev.tc \jobname.tex > wordcount_rev.aux} 
+\newcommand\wordcount{\verbatiminput{wordcount_rev.aux}}
+
+% Floating stuff
+\usepackage[nolists, nomarkers]{endfloat}
+%\renewcommand{\processdelayedfloats}{}  % Avoid float processing
+\usepackage{newfloat}
+\DeclareFloatingEnvironment[
+	fileext=lsf,
+	listname={List of Supplementary Figures},
+	name=Supplementary Figure,
+	placement=p
+]{supfig}
+\DeclareDelayedFloat{supfig}{Supplementary Figures}
 
@@
 \author[1,2]{César Gracia}
+\author[1]{Teresa Rubio}
 \author[1,3,4,5]{Amparo Latorre}
@@
 
+%% TeXcount commands (in addition to option_rev.tc)
+%TC:newcounter supfig Number of Supplemental Figures
+%TC:macrocount beginsupfig [supfig]
+%TC:newcounter supfigwords Words Supplemental Figures
+%TC:envir supfig [] supfigwords
 %TC:newcounter abswords Words in abstract
 %TC:macro \abstract [abswords]
+%TC:newcounter impwords Words in importance section
+%TC:macro \importance [impwords]
+
 {\abstract{Animal microbiota (human included) plays an important role keeping healthy the physiological status of the host.
@@
 
-%TC:newcounter impwords Words in importance section
-%TC:macro \importance [impwords]
 \importance{Human microbiota is tightly associated to the health status of a person.
@@
 \section*{Introduction}
-\input{mSys_intro.tex}
+\input{mSys_intro_rev.tex}
 %\clearpage
@@
 \section*{Results}
-\input{mSys_results.tex}
+\input{mSys_results_rev.tex}
 %\clearpage
@@
 \section*{Discussion}
-\input{mSys_discussion.tex}
+\input{mSys_discussion_rev.tex}
 %\clearpage    
@@
 \section*{Materials and Methods}
-\input{mSys_MandM.tex}   
+\input{mSys_MandM_rev.tex}   
 %\clearpage
@@
 \bibitem {FD} {\bf Weber, J.} 1956. Fluctuation Dissipation Theorem. Phys. Rev. {\bf 101}:1620-6
+\bibitem {sislands1} {\bf Ng SC, Lam EFC, Lam TTY, Chan Y, Law W, Tse PCH, Kamm M a, Sung JJY, Chan FKL, Wu JCY.} 2013.
+Effect of probiotic bacteria on the intestinal microbiota in irritable bowel syndrome.
+J Gastroenterol Hepatol {\bf 28}:1624–31.
+\bibitem {sislands2} {\bf Carroll IM, Ringel-Kulka T, Siddle JP, Ringel Y.} 2012.
+Alterations in composition and diversity of the intestinal microbiota in patients with diarrhea-predominant irritable bowel syndrome.
+Neurogastroenterol Motil {\bf 24}:521-30
+\bibitem {sislands3} {\bf Xu D, Gao J, Gillilland M, Wu X, Song I, Kao JY, Owyang C.} 2014.
+Rifaximin alters intestinal bacteria and prevents stress-induced gut inflammation and visceral hyperalgesia in rats.
+Gastroenterology {\bf 146}:484–496.e4.
+\bibitem {sislands4} {\bf Rooks MG, Veiga P, Wardwell-scott LH, Tickle T, Segata N, Michaud M, Gallini CA, Beal C, van Hylckama-Vlieg JET, Ballal SA, Morgan XC, Glickman JN, Gevers D, Huttenhower C, Garrett WS, Hylckama-vlieg JET Van, Ballal SA, Morgan XC, Glickman JN, Gevers D, Huttenhower C, Garrett WS.} 2014.
+Gut microbiome composition and function in experimental colitis during active disease and treatment-induced remission.
+ISME J {\bf 8}:1403–17.
 \bibitem {tikhonov} {\bf  Tikhonov M.} 2016. Community-level cohesion without cooperation. Elife {\bf 5}.

### 2016-12-21 mSystems/rev/mSys_MandM_rev.tex
Add LaTeX version for mSystems revised manuscript and improve figs
Changes to be committed:
	modified:   ../.gitignore
	new file:   figs/Fig2/ellipse.m
	new file:   figs/Fig2/ellipsedata.txt
	new file:   figs/Fig2/errorbarxy.m
	modified:   mSys_MAIN.pdf
	modified:   rev/figs/Fig1.eps
	modified:   rev/figs/Fig2.eps
	modified:   rev/figs/_ed_Fig1.pdf
	renamed:    rev/figs/_ed_FigS3.pdf -> rev/figs/_ed_supfig_corrank_IBS_C.pdf
	renamed:    rev/figs/_ed_FigS4.pdf -> rev/figs/_ed_supfig_corrank_IBS_P1.pdf
	renamed:    rev/figs/_ed_FigS1.pdf -> rev/figs/_ed_supfig_taxlev1.pdf
	renamed:    rev/figs/_

--- 2016-09-22 mSystems/mSys_MandM.tex
@@
 %%
+%TC:ignore
 %\small
@@
 The steady-state solution of the Fokker-Planck equation is given by:
-\begin{eqnarray}
-P_0 (x_i) &=& C_{ne}(\alpha,\beta,F_i,V)  \cdot x_i^{-2\beta}  \cdot \exp\Big[\frac{2F_i}{V^2}\frac{x_i^{1+\alpha-2\beta}}{1+\alpha-2\beta}-\frac{\phi_0}{V^2}\frac{x_i^{2-2\beta}}{1-\beta}\Big] \quad \textrm{if} \quad  2\beta \ne 1+\alpha, \notag \\
+\begin{eqnarray*}
+P_0 (x_i) &=& C_{ne}(\alpha,\beta,F_i,V)  \cdot x_i^{-2\beta}  \cdot \exp\Big[\frac{2F_i}{V^2}\frac{x_i^{1+\alpha-2\beta}}{1+\alpha-2\beta}-\frac{\phi_0}{V^2}\frac{x_i^{2-2\beta}}{1-\beta}\Big] \quad \textrm{if} \quad  2\beta \ne 1+\alpha, \\
 P_0 (x_i) &=& C_e(\alpha,\beta,F_i,V)  \cdot x_i^{\frac{2F_i}{V^2} -2\beta}  \cdot \exp\Big[\frac{\phi_0}{V^2}\frac{x_i^{2-2\beta}}{1-\beta}\Big] \quad \textrm{if} \quad  2\beta = 1+\alpha,
-\notag
-\end{eqnarray}
+\end{eqnarray*}
 where $\phi_0 = (\sum_i F_i^{1/(1-\alpha)})^{1-\alpha}$ and C$_{ne}$ and C$_{e}$ are integrals that should be solved numerically for the parameters of interest.
@@
 The phase transition region can be calculated analytically in particular cases:
-\begin{eqnarray}
-F_i^2 &=& 4 \beta \phi_0 V^2\quad \textrm{if} \quad  \beta = \alpha \neq 1, \notag \\
+\begin{eqnarray*}
+F_i^2 &=& 4 \beta \phi_0 V^2\quad \textrm{if} \quad  \beta = \alpha \neq 1,\\
 F_i &=& \beta V^2\quad \textrm{if} \quad  2\beta = 1+\alpha,
 \notag
-\end{eqnarray}
+\end{eqnarray*}
 where the first case, simplifies to $F= 3 V^2$ if $\beta = 0.75$ and the fitness of this taxon dominates in $\phi_0$. 
@@
 Genus, with best balance between error assignment and number of taxa, was chosen as our reference taxonomic level.
-We have verified that our conclusions are not significantly affected by selecting family or species as the reference taxonomic level (see Figure \ref{fig:taxlev1}).
+We have verified that our conclusions are not significantly affected by selecting family or species as the reference taxonomic level (see Supplementary Figure \ref{supfig:taxlev1}).
 
@@
 The latter dataset tested was the kwashiorkor discordant twins study\cite{kwashiorkor} for both genus and species taxonomic levels.
-The Figures \ref{fig:taxlev1} (overview) and \ref{fig:taxlev2} (detail) plot the comparison between studies (and so, 16S and SMS) and between adjacent taxonomic levels.
+The Supplementary Figures \ref{supfig:taxlev1} (overview) and \ref{supfig:taxlev2} (detail) plot the comparison between studies (and so, 16S and SMS) and between adjacent taxonomic levels.
 
-\begin{figure} 
-  \includegraphics[width=0.5\textwidth]{results/taxalevel/sum_raw_16S.pdf}
-  \includegraphics[width=0.5\textwidth]{results/taxalevel/sum_sta_16S.pdf}
-  \includegraphics[width=0.5\textwidth]{results/taxalevel/sum_raw_SMS.pdf} 
-  \includegraphics[width=0.5\textwidth]{results/taxalevel/sum_sta_SMS.pdf}
+%TC:endignore
+\begin{supfig} 
+  %\includegraphics[width=0.5\textwidth]{results/taxalevel/sum_raw_16S.pdf}
+  %\includegraphics[width=0.5\textwidth]{results/taxalevel/sum_sta_16S.pdf}
+  %\includegraphics[width=0.5\textwidth]{results/taxalevel/sum_raw_SMS.pdf} 
+  %\includegraphics[width=0.5\textwidth]{results/taxalevel/sum_sta_SMS.pdf}
+  \includegraphics[width=0.99\textwidth]{figs/supfig_taxlev1.eps}
 \caption{Overview of comparison of different approaches based on adjacent taxonomic levels using plots in the Taylor-parameters space.
@@
 The left column shows the raw results and the right column plots the standardized results (see Standardization under Material and Methods).}
-\label{fig:taxlev1}
-\end{figure}
+\label{supfig:taxlev1}
+\end{supfig}
 
-\begin{figure} 
-  \includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_fam_16S.pdf}
-  \includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_gen_16S.pdf}
-  \includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_gen_SMS.pdf}
-  \includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_spc_SMS.pdf}
+\begin{supfig} 
+  %\includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_fam_16S.pdf}
+  %\includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_gen_16S.pdf}
+  %\includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_gen_SMS.pdf}
+  %\includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_spc_SMS.pdf}
+  \includegraphics[width=0.99\textwidth]{figs/supfig_taxlev1.eps}
 \caption{Detail of comparison of different approaches based on adjacent taxonomic levels using plots of X-weighted power-law fits (see Material and Methods).
@@
 The left column shows results for the superior taxonomic level (family for 16S, genus for SMS), while the right column shows results for the inferior level (genus for 16S, specie for SMS).}
-\label{fig:taxlev2}
-\end{figure}
+\label{supfig:taxlev2}
+\end{supfig}
+%TC:ignore
 
@@
 %\normalsize
+%TC:endignore

### 2016-12-21 mSystems/rev/mSys_discussion_rev.tex
Add LaTeX version for mSystems revised manuscript and improve figs
Changes to be committed:
	modified:   ../.gitignore
	new file:   figs/Fig2/ellipse.m
	new file:   figs/Fig2/ellipsedata.txt
	new file:   figs/Fig2/errorbarxy.m
	modified:   mSys_MAIN.pdf
	modified:   rev/figs/Fig1.eps
	modified:   rev/figs/Fig2.eps
	modified:   rev/figs/_ed_Fig1.pdf
	renamed:    rev/figs/_ed_FigS3.pdf -> rev/figs/_ed_supfig_corrank_IBS_C.pdf
	renamed:    rev/figs/_ed_FigS4.pdf -> rev/figs/_ed_supfig_corrank_IBS_P1.pdf
	renamed:    rev/figs/_ed_FigS1.pdf -> rev/figs/_ed_supfig_taxlev1.pdf
	renamed:    rev/figs/_

--- 2016-09-22 mSystems/mSys_discussion.tex
@@
 
-The primary hypothesis of this work is that, in adults, having a healthy microbiota means that population is stable in time and does not have huge flips or jumps into another states.
+The primary hypothesis of this work is that, in adults, having a healthy microbiota means that population is stable in time and does not travel into a state where it is highly susceptible to external or internal perturbations, causing in such a case a dysbiotic state in the microbiota.
 In order to use the valuable information which gives us the empirical law of Taylor's work, we propose the use of Langevin equation to model how the ranking stability evolves in time.
@@
 His microbiota is altered enough to lose its stability, and hence be placed in the unstable part.
-So located, it is more sensible to potential perturbations as, for example, opportunist infections.
+So located, it is more sensitive to potential perturbations as, for example, opportunist infections.
 Third and last condition, the subject is already in the unstable phase due to some healthy issue as IBS.
@@
 Interestingly, that stability disappears when we look at the IBS patients.
-One could ask if these taxa are key players in the phase transition of the microbiota, of if they are more susceptible to perturbations than the most abundant.
+From the literature, it seems that some genera from the families Comamonadaceae, Neisseriaceae and Carnobacteriaceae have been reported to lower their abundance in IBS patients against healthy controls \cite{sislands1}.
+In our case, we see that these families that are present in the rank stability island of the healthy patient decrease their rank stability index or even disappear in the IBS patient.
+However, we also see contradictory results in other families as Lactobacillaceae or Fusobacteriaceae, which seems to increase their abundance in IBS patients \cite{sislands2}, while we observe an increase of the former and a decrease of the latter.
+The presence of members of the Lactobacillae family have been reported to have positive effects against gut inflammation and visceral hyperalgesia \cite{sislands3}, usual symptoms of gastrointestinal disorders.
+It could be happening that a disorder in the stability of this particular group may help to arise the onset of symptoms associated to gastrointestinal disorders.
+The Aerococacceae family is also enriched in rats with IBS symptoms who have been treated with immunomodulators \cite{sislands4}.
+Inside this island of stability, we also have families as Fusobacteriaceae o Hallomonadaceae which include pathogenic genera in them, but that are not present in the IBS patient.
+It could be brought into question the rule of these taxa as key players in the phase transition of the microbiota, or whether they are more susceptible to perturbations than the most abundant.
 The types of interactions that could be sustaining this particular behavior are not clear, as these non-abundant taxa are not usually included in dynamical studies in order to get the community matrix.

### 2016-12-21 mSystems/rev/mSys_intro_rev.tex
Add LaTeX version for mSystems revised manuscript and improve figs
Changes to be committed:
	modified:   ../.gitignore
	new file:   figs/Fig2/ellipse.m
	new file:   figs/Fig2/ellipsedata.txt
	new file:   figs/Fig2/errorbarxy.m
	modified:   mSys_MAIN.pdf
	modified:   rev/figs/Fig1.eps
	modified:   rev/figs/Fig2.eps
	modified:   rev/figs/_ed_Fig1.pdf
	renamed:    rev/figs/_ed_FigS3.pdf -> rev/figs/_ed_supfig_corrank_IBS_C.pdf
	renamed:    rev/figs/_ed_FigS4.pdf -> rev/figs/_ed_supfig_corrank_IBS_P1.pdf
	renamed:    rev/figs/_ed_FigS1.pdf -> rev/figs/_ed_supfig_taxlev1.pdf
	renamed:    rev/figs/_

--- 2016-09-22 mSystems/mSys_intro.tex
@@
 
-High throughput methods for microbial 16S ribosomal RNA gene and SMS have now begun to reveal the composition of archaeal, bacterial, fungal and viral communities located both, in and on the human body.
+High throughput methods for microbial 16S ribosomal RNA gene and SMS (shotgun metagenomic sequencing) have now begun to reveal the composition of archaeal, bacterial, fungal and viral communities located both, in and on the human body.
 Modern high-throughput sequencing and bioinformatics tools provide a powerful means of understanding how the human microbiome contributes to health and its potential as a target for therapeutic interventions \cite{microb&health}.

### 2016-12-21 mSystems/rev/mSys_results_rev.tex
Add LaTeX version for mSystems revised manuscript and improve figs
Changes to be committed:
	modified:   ../.gitignore
	new file:   figs/Fig2/ellipse.m
	new file:   figs/Fig2/ellipsedata.txt
	new file:   figs/Fig2/errorbarxy.m
	modified:   mSys_MAIN.pdf
	modified:   rev/figs/Fig1.eps
	modified:   rev/figs/Fig2.eps
	modified:   rev/figs/_ed_Fig1.pdf
	renamed:    rev/figs/_ed_FigS3.pdf -> rev/figs/_ed_supfig_corrank_IBS_C.pdf
	renamed:    rev/figs/_ed_FigS4.pdf -> rev/figs/_ed_supfig_corrank_IBS_P1.pdf
	renamed:    rev/figs/_ed_FigS1.pdf -> rev/figs/_ed_supfig_taxlev1.pdf
	renamed:    rev/figs/_

--- 2016-09-22 mSystems/mSys_results.tex
@@
 The fitting to the power law is always robust ($R^{2}$ > 0.88) and does not depend on the microbiome condition.
-The power law (or scaling) index $\beta$ and the variability $V$ (hereafter Taylor parameters) appear to be correlated with the stability of the community and related with the health status of the host, which we consider the main finding exposed in this article.
+The power law (or scaling) index $\beta$ and the variability $V$ (hereafter Taylor parameters) appear to be correlated with the stability of the community.
+.
+The variability V is a direct estimator of amplitude of fluctuations over time, while the scaling index give us information about the statistic properties of the ecosystem.
+The fact that it is less than the unity (when looking at the standard deviation) tell us that the most abundant taxa are less susceptible to any perturbation than the less abundant elements of the population.
+The Taylor parameters are related with the health status of the host, which we consider the main finding exposed in this article.
 
@@
 	\vspace*{-15mm} % Corrects overbox of the figures
-	\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_xWboot_LOG.pdf}
-	\includegraphics[width=0.8\textwidth]{results/fits/IBS_P2_Metatranscriptores_stdVSmean_xWboot_LOG.pdf}
-	\caption{X-weighted power-law fits of the standard deviations (std) versus the mean values for each bacterial genus monitored in time.
+	%\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_xWboot_LOG.pdf}
+	%\includegraphics[width=0.8\textwidth]{results/fits/IBS_P2_Metatranscriptores_stdVSmean_xWboot_LOG.pdf}
+	\includegraphics[width=0.8\textwidth]{figs/Fig1.eps}
+	\caption{X-weighted power-law fits of the standard deviations versus the mean values for each bacterial genus monitored in time.
 We show the fit for samples from a healthy subject (top) and from a subject diagnosed with irritable bowel syndrome (bottom), studied in our lab \cite{IBS}.
@@
 
-In order to jointly visualize and compare the results of individuals from different studies\cite{IBS,moving,antibiotic,LEA,kwashiorkor,diet}, their Taylor parameters have been standardized, where standardization means that each parameter is subtracted by the mean value and divided by the standard deviation (std) of the group of healthy individuals for each study (for details of the procedure, please see Standardization subsection in Material and Methods).
-The healthy zone and the standardized Taylor parameters for individuals whose gut microbiota is altered (i.e., suffering from kwashiorkor, altered diet, antibiotics or IBS) is shown in Figure \ref{fig:main2}.
+In order to jointly visualize and compare the results of individuals from different studies\cite{IBS,moving,antibiotic,LEA,kwashiorkor,diet}, their Taylor parameters have been standardized, where standardization means that each parameter is subtracted by the mean value and divided by the standard deviation of the group of healthy individuals for each study (for details of the procedure, please see Standardization subsection in Material and Methods).
+The healthy zone and the standardized Taylor parameters for individuals whose gut microbiota is compromised (i.e., suffering from kwashiorkor, altered diet, antibiotics or IBS) is shown in Figure \ref{fig:main2}.
 Children developing kwashiorkor show smaller variability than their healthy twins.
@@
 	\centering
-	\includegraphics[width=0.9\textwidth]{results/finalplot11.pdf}
+	%\includegraphics[width=0.9\textwidth]{results/finalplot11.pdf}
+	\includegraphics[width=0.9\textwidth]{figs/Fig2.eps}
 	\caption{Taylor's law parameter space.
@@
 The coloured circle corresponds to 68\% confidence level (CL) region of healthy individuals in the Taylor parameter space, while dashed line delimites the 98\% CL region.
-Points with errors place each individual gut microbiome in the Taylor space.
+Points with errors place gut microbiome in the Taylor space, for each individual whose microbiota was compromised.
 Note that the parameters have been standardized (standard deviation units) to the healthy group in each study for demonstrative and comparative purposes.}
@@
 	\vspace*{-5mm} % Corrects overbox of the figures
-	\includegraphics[width=0.8\textwidth]{results/finalPlot33_new.pdf}
+	%\includegraphics[width=0.8\textwidth]{results/finalPlot33_new.pdf}
+	\includegraphics[width=0.8\textwidth]{figs/Fig3.eps}
 	\caption{Microbiota states can be placed in the phase space $F$--$V$.
@@
 
-\def \RSItable {
-\scriptsize
-	\begin{tabular}{cccr}
-	\multicolumn{4}{c}{Colour code for the RSI percentage column} \\
-    \hline
-    Case  &  Condition  &  Colour  &  Description \\
-    \hline
-    1  &  $1\ge{\rm RSI}>0.99$  & \textcolor{blue}{blue} & constant rank \\ 
-    2  &  $0.99\ge{\rm RSI}>0.90$  &  \textcolor{green}{green}  & highly stable rank \\
-    3  &  $0.90\ge{\rm RSI}>0.75$  &  \textcolor{orange}{orange} & moderately stable rank \\
-    4  &  $0.75\ge{\rm RSI}>0.25$  &  \textcolor{red}{red} & unstable rank \\
-    5  &  $0.25\ge{\rm RSI}\ge0$  &  \bfseries{black} & very unstable rank \\
-    \hline
-    \end{tabular}	
-}
-
 The rank dynamics and stability plots in Figure \ref{fig:corrank_IBS_A} and \ref{fig:corrank_IBS_P2} show the variation in the rank with time for the most dominant taxa and their calculated Rank Stability Index (RSI, as discussed in Material and Methods) for the taxa of a healthy subject (individual \emph{A}, Figure \ref{fig:corrank_IBS_A}) and from a subject diagnosed with IBS (patient \emph{P2}, Figure \ref{fig:corrank_IBS_P2}) of the IBS study\cite{IBS}.
@@
 	%\vspace*{-10mm} % Corrects overbox of the figures
-	\includegraphics[width=0.99\textwidth]{results/corrank/IBS_h_A_amplicons_family_Rank.pdf}
-	\RSItable	
+	%\includegraphics[width=0.99\textwidth]{results/corrank/IBS_h_A_amplicons_family_Rank.pdf}
+	\includegraphics[width=0.99\textwidth]{figs/Fig4.eps}
 	\caption{Rank variation throughout time for the most dominant elements (taxa) and their calculated Rank Stability Index (as shown in Material and Methods) for samples from a healthy subject studied in our lab\cite{IBS}.}
@@
 	%\vspace*{-10mm} % Corrects overbox of the figures
-	\hspace*{-4.5mm}\includegraphics[width=0.99\textwidth]{results/corrank/IBS_P2_Metatranscriptores_Rank.pdf}
-	\RSItable
+	%\hspace*{-4.5mm}\includegraphics[width=0.99\textwidth]{results/corrank/IBS_P2_Metatranscriptores_Rank.pdf}
+	\includegraphics[width=0.99\textwidth]{figs/Fig5.eps}
 	\caption{Rank variation throughout time for the most dominant elements (taxa) and their calculated Rank Stability Index for samples from a subject diagnosed with irritable bowel syndrome studied in our lab\cite{IBS}.}
@@
 We found the presence of such of \emph{rank stability island} for medium-ranked taxa in the other healthy subjects (\emph{B} and \emph{C}) of the IBS study\cite{IBS} together with its total absence for the other IBS diagnosed patient (patient \emph{P1}), which also presents very high rank instability in its medium-ranked taxa.
+See Supplementary Figure \ref{supfig:corrank_IBS_C} and Supplementary Figure \ref{supfig:corrank_IBS_P1} for further examples of rank plots for a healthy and light IBS diagnosed patients, respectively.
+
+\begin{supfig}
+	\centering
+	%\vspace*{-10mm} % Corrects overbox of the figures
+	\includegraphics[width=0.99\textwidth]{figs/supfig_corrank_IBS_C.eps}
+	\caption{Rank variation throughout time for the most dominant elements (taxa) and their calculated RSI for samples from another healthy subject studied in our lab\cite{IBS}.}
+	\label{supfig:corrank_IBS_C}
+\end{supfig}
+
+\begin{supfig}
+	\centering
+	%\vspace*{-10mm} % Corrects overbox of the figures
+	%\hspace*{-4.5mm}
+	\includegraphics[width=0.99\textwidth]{figs/supfig_corrank_IBS_P1.eps}
+	\caption{Rank variation throughout time for the most dominant elements (taxa) and their calculated RSI for samples from a subject diagnosed with light IBS and studied in our lab\cite{IBS}.}
+	\label{supfig:corrank_IBS_P1}
+\end{supfig}
 
@@
 Figure \ref{fig:tempevo1} shows the variability  $V$ as a function of time for the largest sampling: two individuals in the Caporaso's study\cite{moving} corresponding to the gut microbiota of a male (upper plot) and a female (lower plot).
-Figure \ref{fig:tempevo2} shows the time evolution of $V$ for patient P2 of the IBS study\cite{IBS} (upper plot) and patient D in the antibiotics study\cite{antibiotic} (lower plot).
 Both samples shows changes in the variability V with quasi--periodic behavior peaked at about 10 days.
 Variability grows more for the gut microbiota of the male and share a minimal value around 0.1 with the gut microbiota of the female.
+
+Figure \ref{fig:tempevo2} shows the time evolution of $V$ for patient P2 of the IBS study\cite{IBS} (upper plot) and patient D in the antibiotics study\cite{antibiotic} (lower plot).
 The variability of the gut microbiota of P2 decreases from above 0.3 to below 0.2, showing a slow tendency to increase the order of the system.
@@
 \begin{figure}
-	\includegraphics[width=1.0\textwidth]{results/sliwin/male_mov.pdf}
-	\hspace*{3mm}\includegraphics[width=0.448\textwidth]{results/sliwin/female_mov.pdf}
+	%\includegraphics[width=1.0\textwidth]{results/sliwin/male_mov.pdf}
+	%\hspace*{3mm}\includegraphics[width=0.448\textwidth]{results/sliwin/female_mov.pdf}
+	\includegraphics[width=0.99\textwidth]{figs/Fig6.eps}
 \caption{$V$ as a function of time for the two individuals in the Caporaso's study\cite{moving}: samples of gut microbiome of a male (upper plot) and a female (lower plot).}
@@
 	\centering 
- 	\includegraphics[width=0.8\textwidth]{results/sliwin/patP2_IBS.pdf}
-  	\includegraphics[width=1.0\textwidth]{results/sliwin/patD_antibio.pdf} 
+ 	%\includegraphics[width=0.8\textwidth]{results/sliwin/patP2_IBS.pdf}
+  	%\includegraphics[width=1.0\textwidth]{results/sliwin/patD_antibio.pdf} 
+	\includegraphics[width=0.99\textwidth]{figs/Fig7.eps}	
 \caption{$V$ as a function of time for patient P2 of the IBS study\cite{IBS} (upper plot) and patient D in the antibiotics study\cite{antibiotic} (lower plot).

### 2016-12-21 mSystems/figs/Fig2/ellipse.m

@@ -1,0 +1,158 @@
+fid = fopen('ellipsedata.txt');
+data = textscan(fid, '%f %f %f %f %s');
+X=data{1,1};
+XE=data{1,2};
+Y=data{1,3};
+YE=data{1,4};
+sizeSymbol = 15;
+close all;
+clear title xlabel ylabel;
+%hFigure = figure('Units', 'pixels', 'Position', [100 100 500 375]);
+hFigure = figure('Units', 'pixels', 'Position', [100 100 740 440]);
+daspect([1,1,1]);
+hold on;
+
+%Draw Ellipse
+angle = 43.2/180*pi;
+%a = 0.912317;
+a = 1;
+%b = 0.942536;
+b = 1;
+cx = 0;
+cy = 0;
+%color = [0.48,0.48,0.48];
+color = [176/255,196/255,222/255];
+
+r=0:0.1:2*pi+0.1;
+p=[(a*cos(r))' (b*sin(r))'];
+
+alpha=[cos(angle) -sin(angle)
+       sin(angle) cos(angle)];
+   
+p1=p*alpha;
+ 
+h0 = patch(cx+p1(:,1),cy+p1(:,2),color,'EdgeColor',color);
+
+% Circle
+t = 0 : .1 : 2*pi;
+x = 2 * cos(t) + 0;
+y = 2 * sin(t) + 0;
+ch=plot(x, y,'--');
+set(ch,'Color',color);
+set(ch,'LineWidth',2);
+
+%hh = hatchfill(h0,'cross',45,40,color);
+%set(hh,'color','w','linewidth',1.5)
+
+
+% Twin_Kwashiorkor
+color=[0 0 255/255]; % 9
+initIdx=1;
+endIdx=16;
+%h1 = scatter(X(initIdx:endIdx),Y(initIdx:endIdx),size,color,'filled','MarkerEdgeColor',[0 0 255/255], 'MarkerFaceColor',color);
+h1 = scatter(X(initIdx:endIdx),Y(initIdx:endIdx),sizeSymbol,'d','filled','MarkerEdgeColor',[0 0 255/255], 'MarkerFaceColor',color);
+h1b = errorbarxy(X(initIdx:endIdx),Y(initIdx:endIdx),XE(initIdx:endIdx),YE(initIdx:endIdx),{'.',color,color});
+
+% Twin_healthy
+color=[0/255 255/255 255/255]; % 8
+initIdx=17;
+endIdx=31;
[…]

### 2016-12-21 mSystems/rev/mSys_MandM_rev.tex
new study in the text, several minor modifications

@@
 We considered as the healthy part only those times before any antibiotic treatment, and as perturbations the time of antibiotic intakes and the period after that.
+David et al.
+work \cite{hostlife} was selected due to the great longitudinal data that it provides plus its complete metadata and the interesting events that happened to both subjects (and infection and a travel).
+We considered as a healthy part the previous time points before each event.
 And finally, we also considered a study made in our group carried by Durban et al., \cite{IBS} in which we considered as healthy subjects those who didn't suffer from irritable bowel disease, and as perturbation the patients who had this disease.
 
-Metadata of each study is provided in Supplementary Tables S1 to S4.
+Metadata of each study is provided in Supplementary Tables S1 to S7.
 All used 16S rRNA gene sequencing except for the study of the discordant kwashiorkor twins\cite{kwashiorkor}, where both SMS and 16S rRNA data were used.

### 2016-12-21 mSystems/rev/mSys_results_rev.tex
new study in the text, several minor modifications

@@
 
-Taylor parameters describing the temporal variability of the gut microbiome in our sampled individuals are shown in Supplementary Tables S1 to S6.
+Taylor parameters describing the temporal variability of the gut microbiome in our sampled individuals are shown in Supplementary Tables S1 to S7.
 Our results hint at an ubiquitous behavior.
@@
 
-In order to jointly visualize and compare the results of individuals from different studies\cite{IBS,moving,antibiotic,LEA,kwashiorkor,diet}, their Taylor parameters have been standardized, where standardization means that each parameter is subtracted by the mean value and divided by the standard deviation of the group of healthy individuals for each study (for details of the procedure, please see Standardization subsection in Material and Methods).
-The healthy zone and the standardized Taylor parameters for individuals whose gut microbiota is compromised (i.e., suffering from kwashiorkor, altered diet, antibiotics or IBS) is shown in Figure \ref{fig:main2}.
+In order to jointly visualize and compare the results of individuals from different studies\cite{IBS,moving,antibiotic,LEA,kwashiorkor,diet,hostlife}, their Taylor parameters have been standardized, where standardization means that each parameter is subtracted by the mean value and divided by the standard deviation of the group of healthy individuals for each study (for details of the procedure, please see Standardization subsection in Material and Methods).
+The healthy zone and the standardized Taylor parameters for individuals whose gut microbiota is compromised (i.e., suffering from kwashiorkor, a Salmonella infection, travel abroad, altered diet, antibiotics or IBS) is shown in Figure \ref{fig:main2}.
 Children developing kwashiorkor show smaller variability than their healthy twins.
 A meat/fish-based diet increases the variability significantly when compared to a plant-based diet.
-All other cases presented increased variability, which is particularly severe, and statistically significant at more than 95\% CL, for obese patients grade III on a diet, individuals taking antibiotics or IBS--diagnosed patients.
+All other cases presented increased variability, which is particularly severe, and statistically significant at more than 95\% CL, for obese patients grade III on a diet, individuals taking antibiotics, the subject who had a Salmonella infection, the subject who did a travel abroad or the IBS--diagnosed patients.
 A global property emerges from all worldwide data collected: Taylor parameters characterize the statistical behavior of microbiome changes.

### 2016-12-21 mSystems/figs/Fig2/ellipse.m
Minor modifications and update Fig2 source scripts&data
Changes to be committed:
	modified:   ../README.md
	modified:   figs/Fig2/ellipse.m
	modified:   figs/Fig2/ellipsedata.txt
	modified:   rev/mSys_MAIN_rev.pdf
	modified:   rev/mSys_MAIN_rev.tex
	modified:   rev/mSys_discussion_rev.tex

@@ -4,7 +4,7 @@ X=data{1,1};
 XE=data{1,2};
 Y=data{1,3};
 YE=data{1,4};
-sizeSymbol = 15;
+sizeSymbol = 20;
 close all;
 clear title xlabel ylabel;
 %hFigure = figure('Units', 'pixels', 'Position', [100 100 500 375]);
@@ -112,23 +112,28 @@ h8b = errorbarxy(X(initIdx:endIdx),Y(initIdx:endIdx),XE(initIdx:endIdx),YE(initI
 color=[0/255 0/255 0/255]; % 1
 initIdx= 65;
 endIdx= 66;
-h9 = scatter(X(initIdx:endIdx),Y(initIdx:endIdx),sizeSymbol,color,'filled','MarkerEdgeColor',[0/255 0/255 0/255], 'MarkerFaceColor',color);
+h9 = scatter(X(initIdx:endIdx),Y(initIdx:endIdx),sizeSymbol,'p','filled','MarkerEdgeColor',[0/255 0/255 0/255], 'MarkerFaceColor',color);
 h9b = errorbarxy(X(initIdx:endIdx),Y(initIdx:endIdx),XE(initIdx:endIdx),YE(initIdx:endIdx),{'w.',color,color});
 
-% TERESA
-color=[255/255 0/255 255/255]; % 1
-%initIdx= 67;
-%endIdx= 66;
-%h9 = scatter(X(initIdx:endIdx),Y(initIdx:endIdx),sizeSymbol,color,'filled','MarkerEdgeColor',[0/255 0/255 0/255], 'MarkerFaceColor',color);
-%h9b = errorbarxy(X(initIdx:endIdx),Y(initIdx:endIdx),XE(initIdx:endIdx),YE(initIdx:endIdx),{'w.',color,color});
-h10 = scatter([-2,-1,0,1,2,3,4,5,6],[-2,-1,0,1,2,3,4,5,6],sizeSymbol,'d','filled','MarkerEdgeColor',color, 'MarkerFaceColor',color);
+% HLS_TRIP
+color=[255/255 0/255 255/255]; % 
+initIdx= 67;
+endIdx= 67;
+h10 = scatter(X(initIdx:endIdx),Y(initIdx:endIdx),sizeSymbol,'^','filled','MarkerEdgeColor',color, 'MarkerFaceColor',color);
[…]

### 2016-12-23 mSystems/rev/mSys_MAIN_rev.tex
new study in the text, several minor modifications
Minor modifications and update Fig2 source scripts&data
Changes to be committed:
	modified:   ../README.md
	modified:   figs/Fig2/ellipse.m
	modified:   figs/Fig2/ellipsedata.txt
	modified:   rev/mSys_MAIN_rev.pdf
	modified:   rev/mSys_MAIN_rev.tex
	modified:   rev/mSys_discussion_rev.tex
Update supplementary, Figure 2 and various minor enhancements
Changes to be committed:
	modified:   rev/figs/Fig2.eps
	new file:   rev/figs/Fig2/Fig2.pdf
	new file:   rev/figs/Fig2/Fig2.svg
	new file:   rev/figs/Fig2/README
	new file:   rev/figs/Fig2/ellipse.

@@
 % release codenumber (update it here!)
-\newcommand\releasecode{REVISION draft A}
+\newcommand\releasecode{REVISION draft C}
 
@@
 Nature {\bf 505}:559–63.
+\bibitem{hostlife} {\bf David LA, Materna AC, Friedman J, Campos-Baptista MI, Blackburn MC, Perrotta A, Erdman SE, Alm EJ.} 2014.
+Host lifestyle affects human microbiota on daily timescales.
+Genome Biol {\bf 15}:R89.
 % <- END main studies
@@
 \caption{Taylor parameters for persons diagnosed with irritable bowel syndrome (IBS)\cite{IBS}.
-Healthy individuals sampled in this study are characterized by $\bar{V} = 0.134 \pm 0.009, \bar{\beta} = 0.691 \pm 0.025$.}
+Healthy individuals sampled in this study are characterized by $\bar{V} = 0.135 \pm 0.010, \bar{\beta} = 0.692 \pm 0.024$.}
 
@@
 Healthy people in this study, whom were not obese, are characterized by $\bar{V} = 0.19 \pm 0.06, \bar{\beta} = 0.806 \pm 0.034$.}
+
+\caption{Taylor parameters for special intervals concerning gut microbiota in the host lifestyle study\cite{hostlife}.
+The healthy and quotidian periods are characterized by $\bar{V} = 0.25 \pm 0.09, \bar{\beta} = 0.777 \pm 0.025$.}
 \end{table}

### 2016-12-23 mSystems/rev/mSys_discussion_rev.tex
new study in the text, several minor modifications
Minor modifications and update Fig2 source scripts&data
Changes to be committed:
	modified:   ../README.md
	modified:   figs/Fig2/ellipse.m
	modified:   figs/Fig2/ellipsedata.txt
	modified:   rev/mSys_MAIN_rev.pdf
	modified:   rev/mSys_MAIN_rev.tex
	modified:   rev/mSys_discussion_rev.tex
Update supplementary, Figure 2 and various minor enhancements
Changes to be committed:
	modified:   rev/figs/Fig2.eps
	new file:   rev/figs/Fig2/Fig2.pdf
	new file:   rev/figs/Fig2/Fig2.svg
	new file:   rev/figs/Fig2/README
	new file:   rev/figs/Fig2/ellipse.

@@
 In addition, it was shown that this subject improved its healthy status in the time when the experiment was done, implying that his microbiota also recovered the lost stability.
+It is interesting to notice that in the subject from the study of David et al.
+\cite{hostlife}, who suffered a Salmonella infection during the experiment, we can observe a huge shift in the variability and a final recovery from the perturbed state (see \ref{supfig:HLS_xWSummary}).
+
+\begin{supfig}
+	\centering
+	%\vspace*{-10mm} % Corrects overbox of the figures
+	\includegraphics[width=0.99\textwidth]{figs/supfig_HLS_xWSummary.eps}
+	\caption{Taylor's law parameter space for intervals concerning gut microbiota in the host lifestyle study\cite{hostlife}.}
+	\label{supfig:HLS_xWSummary}
+\end{supfig}
 

### 2016-12-23 mSystems/rev/mSys_suppl_rev.tex
Update supplementary, Figure 2 and various minor enhancements
Changes to be committed:
	modified:   rev/figs/Fig2.eps
	new file:   rev/figs/Fig2/Fig2.pdf
	new file:   rev/figs/Fig2/Fig2.svg
	new file:   rev/figs/Fig2/README
	new file:   rev/figs/Fig2/ellipse.m
	new file:   rev/figs/Fig2/ellipsedata.txt
	new file:   rev/figs/Fig2/errorbarxy.m
	new file:   rev/figs/supfig_HLS_xWSummary.eps
	new file:   rev/figs/supfig_HLS_xWSummary.pdf
	modified:   rev/mSys_MAIN_rev.pdf
	modified:   rev/mSys_MAIN_rev.tex
	modified:   rev/mSys_discussion_rev.tex
	new file:   rev/mSys_suppl_rev.pdf
	modified:   re

--- 2016-09-22 mSystems/mSys_suppl.tex
@@
 % release codenumber (update it here!)
-\newcommand\releasecode{$\beta$}
+\newcommand\releasecode{REVISION draft C}
 
@@
 % watermarks (time consuming!)
-\newwatermark[allpages,color=Green!4,angle=45,scale=12,xpos=-60,ypos=0]{RC\releasecode}
-\newwatermark[allpages,color=Green!12,angle=0,scale=1,xpos=0,ypos=137]{
-\emph{mSystems} \LaTeX\ manuscript RC\releasecode}
+\newwatermark[allpages,color=Mulberry!4,angle=45,scale=7,xpos=-40,ypos=0]{REVISION}
+\newwatermark[allpages,color=Mulberry!12,angle=0,scale=1,xpos=0,ypos=137]{
+\emph{mSystems} \LaTeX\ manuscript \releasecode}
 
@@
 		\hline
-		IBS&$0.204 \pm 0.034$&$0.739 \pm 0.029$&$0.916$&&$7.6 \pm 3.7$&$1.9 \pm 1.2$\\
-		IBS&$0.35 \pm 0.05$&$0.793 \pm 0.023$&$0.935$&&$23.1 \pm 5.9$&$4.0 \pm 0.9$\\
+		IBS (minor)&$0.205 \pm 0.034$&$0.740 \pm 0.029$&$0.917$&&$6.9 \pm 3.3$&$2.0 \pm 1.2$\\
+		IBS (severe)&$0.35 \pm 0.06$&$0.793 \pm 0.025$&$0.934$&&$21 \pm 6$&$4.2 \pm 1.0$\\
 	     \hline
@@
 
+\begin{table}
+\caption{Taylor parameters for special intervals concerning gut microbiota in the host lifestyle study\cite{hostlife}.
+The healthy and quotidian periods are characterized by $\bar{V} = 0.25 \pm 0.09, \bar{\beta} = 0.777 \pm 0.025$.}
+
+  \begin{center}
+    \begin{tabular}{ccccccc}
+	    \hline
+		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
+		\hline
+		HLS (abroad)&$0.51 \pm 0.06$&$0.820 \pm 0.012$&$0.928$&&$2.8 \pm 0.6$&$1.7 \pm 0.5$\\
+		HLS (infection)&$0.49 \pm 0.08$&$0.828 \pm 0.018$&$0.923$&&$2.6 \pm 0.9$&$2.0 \pm 0.7$\\
+		HLS (after infection)&$0.36 \pm 0.05$&$0.776 \pm 0.015$&$0.922$&&$1.1 \pm 0.6$&$-0.0 \pm 0.6$\\
+	     \hline
+	     \hline
+    \end{tabular}
+  \end{center}
+  \label{tab:IBS}
+\end{table}
+
+\clearpage
+
 \begin{thebibliography}{9}
@@
 Science {\bf 341}:1237439.
+\bibitem{hostlife} {\bf David LA, Materna AC, Friedman J, Campos-Baptista MI, Blackburn MC, Perrotta A, Erdman SE, Alm EJ.} 2014.
+Host lifestyle affects human microbiota on daily timescales.
+Genome Biol {\bf 15}:R89.
 \end{thebibliography}

### 2016-12-23 mSystems/rev/figs/Fig2/ellipse.m

@@ -1,0 +1,167 @@
+fid = fopen('ellipsedata.txt');
+data = textscan(fid, '%f %f %f %f %s');
+X=data{1,1};
+XE=data{1,2};
+Y=data{1,3};
+YE=data{1,4};
+sizeSymbol = 20;
+close all;
+clear title xlabel ylabel;
+%hFigure = figure('Units', 'pixels', 'Position', [100 100 740 440]);
+hFigure = figure('Units', 'pixels', 'Position', [100 100 800 400]);
+daspect([1,1,1]);
+hold on;
+
+%Draw Ellipse
+angle = 43.2/180*pi;
+%a = 0.912317;
+a = 1;
+%b = 0.942536;
+b = 1;
+cx = 0;
+cy = 0;
+color = [176/255,196/255,222/255];
+r=0:0.1:2*pi+0.1;
+p=[(a*cos(r))' (b*sin(r))'];
+alpha=[cos(angle) -sin(angle)
+       sin(angle) cos(angle)];   
+p1=p*alpha; 
+h0 = patch(cx+p1(:,1),cy+p1(:,2),color,'EdgeColor',color);
+
+% Circle
+t = 0 : .1 : 2*pi;
+x = 2 * cos(t) + 0;
+y = 2 * sin(t) + 0;
+ch=plot(x, y,'--');
+set(ch,'Color',color);
+set(ch,'LineWidth',2);
+
+%hh = hatchfill(h0,'cross',45,40,color);
+%set(hh,'color','w','linewidth',1.5)
+
+% Twin_Kwashiorkor
+color=[0 0 255/255]; % 9
+initIdx=1;
+endIdx=16;
+%h1 = scatter(X(initIdx:endIdx),Y(initIdx:endIdx),size,color,'filled','MarkerEdgeColor',[0 0 255/255], 'MarkerFaceColor',color);
+h1 = scatter(X(initIdx:endIdx),Y(initIdx:endIdx),sizeSymbol,'d','filled','MarkerEdgeColor',[0 0 255/255], 'MarkerFaceColor',color);
+h1b = errorbarxy(X(initIdx:endIdx),Y(initIdx:endIdx),XE(initIdx:endIdx),YE(initIdx:endIdx),{'.',color,color});
+
+% Twin_healthy
+color=[0/255 255/255 255/255]; % 8
+initIdx=17;
+endIdx=31;
[…]

### 2016-12-23 mSystems/rev/mSys_discussion_rev.tex
Minor enhancements
Changes to be committed:
	modified:   rev/figs/Fig2.eps
	modified:   rev/figs/Fig2/Fig2.pdf
	modified:   rev/mSys_MAIN_rev.pdf
	modified:   rev/mSys_MAIN_rev.tex
	modified:   rev/mSys_discussion_rev.tex
	modified:   rev/mSys_results_rev.tex

--- 2016-12-21 mSystems/rev/mSys_discussion_rev.tex
@@
 In addition, it was shown that this subject improved its healthy status in the time when the experiment was done, implying that his microbiota also recovered the lost stability.
+It is interesting to notice that in the subject from the study of David et al.
+\cite{hostlife} who suffered a Salmonella infection during the experiment, we can observe a significative shift in the variability and a final recovery from the perturbed state (see \ref{supfig:HLS_xWSummary}).
+
+\begin{supfig}
+	\centering
+	%\vspace*{-10mm} % Corrects overbox of the figures
+	\includegraphics[width=0.99\textwidth]{figs/supfig_HLS_xWSummary.eps}
+	\caption{Taylor's law parameter space for intervals concerning gut microbiota in the host lifestyle study\cite{hostlife}.
+We observe that the subject B, who suffered a Salmonella infection during the experiment, had a relevant shift in the parameters from \emph{\_before} to \emph{\_infection} and a final recovery from the perturbed state to \emph{\_afterinfec}.
+The subject A also had a shift in the variability from \emph{\_before} to \emph{\_abroad} and back to \emph{\_returned}.}
+	\label{supfig:HLS_xWSummary}
+\end{supfig}
 

### 2016-12-23 mSystems/rev/mSys_discussion_rev.tex
Minor changes which lead to REVISION draft E
Changes to be committed:
	modified:   rev/mSys_MAIN_rev.pdf
	modified:   rev/mSys_MAIN_rev.tex
	modified:   rev/mSys_discussion_rev.tex
	modified:   rev/mSys_suppl_rev.pdf
	modified:   rev/mSys_suppl_rev.tex

@@
 	\caption{Taylor's law parameter space for intervals concerning gut microbiota in the host lifestyle study\cite{hostlife}.
-We observe that the subject B, who suffered a Salmonella infection during the experiment, had a relevant shift in the parameters from \emph{\_before} to \emph{\_infection} and a final recovery from the perturbed state to \emph{\_afterinfec}.
-The subject A also had a shift in the variability from \emph{\_before} to \emph{\_abroad} and back to \emph{\_returned}.}
+We observe that the subject B, who suffered a Salmonella infection during the experiment, had a relevant shift in the parameters from \emph{\_before} to \emph{\_infection} and a final recovery from the perturbed state to \emph{\_afterinfec}, which lays in the parameters area compatible with the healthy and stable intervals (see Supplementary Table S7).
+The subject A also had a shift in the variability from \emph{\_before} to \emph{\_abroad} and back to \emph{\_returned}, also in the proximity zone of healthy and stable periods.}
 	\label{supfig:HLS_xWSummary}

### 2017-01-10 mSystems/rev/mSys_MAIN_rev.tex
Minor enhancements
Changes to be committed:
	modified:   rev/figs/Fig2.eps
	modified:   rev/figs/Fig2/Fig2.pdf
	modified:   rev/mSys_MAIN_rev.pdf
	modified:   rev/mSys_MAIN_rev.tex
	modified:   rev/mSys_discussion_rev.tex
	modified:   rev/mSys_results_rev.tex
Minor changes which lead to REVISION draft E
Changes to be committed:
	modified:   rev/mSys_MAIN_rev.pdf
	modified:   rev/mSys_MAIN_rev.tex
	modified:   rev/mSys_discussion_rev.tex
	modified:   rev/mSys_suppl_rev.pdf
	modified:   rev/mSys_suppl_rev.tex
English corrections applied, which leads to REVISION draft F
On branch master, changes

@@
 % release codenumber (update it here!)
-\newcommand\releasecode{REVISION draft C}
+\newcommand\releasecode{REVISION draft F}
 
@@
 \affil[1]{Institute for Integrative Systems Biology (I2SysBio), 46980, Spain.}
-\affil[2]{Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain.}
-\affil[3]{FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain.}
+\affil[2]{Instituto de Física Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain.}
+\affil[3]{FISABIO, Avda. de Catalunya, 21, 46020, Valencia, Spain.}
 \affil[4]{Cavanilles Institute of Biodiversity and Evolutionary Biology, UVEG, 46980, Spain.}
@@
 
-{\abstract{Animal microbiota (human included) plays an important role keeping healthy the physiological status of the host.
-Increasing research activity is dedicated to understand how changes in composition and function of the microbiota are associated to disease or not.
-We analyze 16S rRNA and shotgun metagenomic sequencing (SMS) published data from the gut microbiota of 97 individuals monitored in time.
-Temporal fluctuations in the microbial composition reveal significant differences due to factors such us dietary changes, antibiotic intake, age or disease.
-Here we show that a fluctuation scaling law describes the temporal changes in the gut microbiota.
-This law allows to estimate the temporal variability of the microbial population and quantitatively characterizes the path toward disease by a noise-induced phase transition.
-The estimation of the systemic parameters for follow-up studies may have clinical use and, more generally, applications in other fields where it is important to know if a given community is stable or not.}}
+{\abstract{Animal microbiota (including human microbiota) plays an important role in keeping the physiological status of the host healthy.
+Research activity into understanding whether changes in the composition and function of the microbiota are associated with disease is increasing.
+We analyzed 16S rRNA and shotgun metagenomic sequencing (SMS) published data from the gut microbiota of 97 individuals monitored over time.
+Temporal fluctuations in the microbial composition revealed significant differences due to factors such us dietary changes, antibiotic intake, age and disease.
+This article shows that a fluctuation scaling law describes the temporal changes in the gut microbiota.
+This law enables the temporal variability of the microbial population to be estimated and quantitatively characterizes the path toward disease via a noise-induced phase transition.
+The estimation of the systemic parameters for follow-up studies may have clinical use and, more generally, may also have applications in other fields where it is important to know whether a given community is stable or not.}}
 
-\importance{Human microbiota is tightly associated to the health status of a person.
-Here we analyse the microbial composition of several subjects under different conditions, over a time span that ranges from days to months.
-Using the Langevin equation as the basis of our mathematical framework in order to evaluate microbial temporal stability, we prove that we are capable to distinguish stable from unstable microbiotas.
-This first step will help us to determine how microbiota temporal stability is related to the healthiness of the people, and it will allow the development of a more complete framework in order to deepen the knowledge of this complex system.}
+\importance{Human microbiota is closedly linked to the health status of a person.
+This article analyzes the microbial composition of several subjects under different conditions, over a time span that ranged from days to months.
+Using the Langevin equation as the basis of our mathematical framework in order to evaluate microbial temporal stability, we proved that stable microbiotas can be distinguised from unstable microbiotas.
+This first step will help us to determine how microbiota temporal stability is related to the healthiness of people, and it will enable the development of a more comprehensive framework in order to obtain more in-depth knowledge of this complex system.}
 
@@
 \section*{Acknowledgments}
-Authors declare that there are no competing financial interests in relation to the work described here.
-We thereby express our acknowledgement to Bull/Atos and Micron Technology for providing us with the PCIe SSD card Micron P420m HHHL as a free-of-charge sample for high performance throughput database testing purposes.
+The authors declare that there are no competing financial interests in relation to the work described here.
+We thereby express our acknowledgement to Bull/Atos and Micron Technology for providing us with the PCIe SSD card Micron P420m HHHL as a free-of-charge sample for high performance throughput for database testing purposes.
 
 \section*{Funding Information}
-This work was supported by grants to AM from the Spanish Ministry of Science and Competitivity (projects SAF2012-31187, SAF2013-49788-EXP, SAF2015-65878-R), Carlos III Institute of Health (projects PIE14/00045 and AC15/00022), Generalitat Valenciana (project PrometeoII/2014/065) and co-financed by FEDER, and grants to CPG from the Generalitat Valenciana Prometeo Grants II/2014/050, II/2014/065, 419 by the Spanish Grants FPA2011-29678, BFU2012-39816-C02-01 of MINECO and by PITN-GA-420 2011-289442-INVISIBLES.
+This work was supported by grants to AM from the Spanish Ministry of Science and Competitiveness (projects SAF2012-31187, SAF2013-49788-EXP, SAF2015-65878-R), Carlos III Institute of Health (projects PIE14/00045 and AC15/00022), Generalitat Valenciana (project PrometeoII/2014/065) and co-financed by ERDF, and grants to CPG from the Generalitat Valenciana Prometeo Grants II/2014/050, II/2014/065, 419 by the Spanish Grants FPA2011-29678, BFU2012-39816-C02-01 of MINECO and by PITN-GA-420 2011-289442-INVISIBLES.
 From the Ministry of Economy and Competitiveness (grants FPI BES-2012-052900 and FPI BES-2013-062767).
@@
 \bibitem{scfa2} {\bf Smith PM, Howitt MR, Panikov N, Michaud M, Gallini CA, Bohlooly-Y M, Glickman JN, Garrett WS.} 2013.
-The Microbial Metabolites, Short-Chain Fatty Acids, Regulate Colonic Treg Cell Homeostasis.
-Science (80- ) {\bf 341}:569–573.
+The Microbial Metabolites, Short--Chain Fatty Acids, Regulate Colonic Treg Cell Homeostasis.
+Science {\bf 341}:569–573.
 \bibitem {scfa3} {\bf Kimura I, Ozawa K, Inoue D, Imamura T, Kimura K, Maeda T, Terasawa K, Kashihara D, Hirano K, Tani T, Takahashi T, Miyauchi S, Shioi G, Inoue H, Tsujimoto G.} 2013.
@@
 The gut microbiota and host health: a new clinical frontier.
-Gut 1–10.
+Gut {\bf 65}:330–9.
 \bibitem{normal1} {\bf Falony G, Joossens M, Vieira-Silva S, Wang J, Darzi Y, Faust K, Kurilshikov A, Bonder MJ, Valles-Colomer M, Vandeputte D, Tito RY, Chaffron S, Rymenans L, Verspecht C, De Sutter L, Lima-Mendez G, Dhoe K, Jonckheere K, Homola D, Garcia R, Tigchelaar EF, Eeckhaudt L, Fu J, Henckaerts L, Zhernakova A, Wijmenga C, Raes J.} 2016.
 Population-level analysis of gut microbiome variation.
-Science (80- ) {\bf 352}:560–564.
+Science {\bf 352}:560–564.
 \bibitem{normal2} {\bf Zhernakova A, Kurilshikov A, Bonder MJ, Tigchelaar EF, Schirmer M, Vatanen T, Mujagic Z, Vila AV, Falony G, Vieira-Silva S, Wang J, Imhann F, Brandsma E, Jankipersadsing SA, Joossens M, Cenit MC, Deelen P, Swertz MA, Weersma RK, Feskens EJM, Netea MG, Gevers D, Jonkers D, Franke L, Aulchenko YS, Huttenhower C, Raes J, Hofker MH, Xavier RJ, Wijmenga C, Fu J.} 2016.
 Population-based metagenomics analysis reveals markers for gut microbiome composition and diversity.
-Science (80- ) {\bf 352}:565–569.
+Science {\bf 352}:565–569.
 \bibitem{panthropology} {\bf Amato KR} 2016. Incorporating the Gut Microbiota Into Models of Human and Non-Human Primate Ecology and Evolution. Yearbook Of Physical Anthropology {\bf 157}:S196–S215.
@@
 \bibitem {uni_dynam} {\bf Bashan A, Gibson TE, Friedman J, Carey VJ, Weiss ST, Hohmann EL, Liu Y-Y.} 2016. Universality of human microbial dynamics. Nature {\bf 534}:259–262.
-\bibitem{taylor} {\bf Taylor, L.R.} 1961. Aggregation, Variance and the mean. Nature {\bf 189,} 732-35.
+\bibitem{taylor} {\bf Taylor, L.R.} 1961. Aggregation, Variance and the mean. Nature {\bf 189}:732-35.
 \bibitem{randomwalks} {\bf de Menezes MA, Barabási A-L.} 2004. Fluctuations in network dynamics. Phys Rev Lett {\bf 92}:1–4.
@@
 \bibitem{convergence1} {\bf Fronczak A, Fronczak P.} 2010. Origins of Taylor’s power law for fluctuation scaling in complex systems. Phys Rev E {\bf 81}:066112.
-\bibitem{convergence2} {\bf Kendal, W.S., Jorgensen,B.} Taylor's power law and fluctuation scaling explained by a central-limit-like convergence.  Phys. Rev. E {\bf 83} 066115.
+\bibitem{convergence2} {\bf Kendal, W.S., Jorgensen,B.} Taylor's power law and fluctuation scaling explained by a central-limit-like convergence.  Phys. Rev. E {\bf 83}:066115.
 \bibitem{convergence3} {\bf Kendal WS, Jørgensen B.} 2011. Tweedie convergence: A mathematical basis for Taylor’s power law, 1/f noise, and multifractality. Phys Rev E {\bf 84}:066120.
@@
 \begin{table}
-  \caption{Taylor parameters.
+  \caption{Taylor's parameters.
 Individuals with either animal-based (A) or plant-based (P) diets\cite{diet}.
-Previous to diet,  the population sampled is described by $\bar{V} = 0.09 \pm 0.05, \bar{\beta} = 0.77 \pm 0.04$.}
+Previous to the diet, the population sampled is described by $\bar{V} = 0.09 \pm 0.05, \bar{\beta} = 0.77 \pm 0.04$.}
 
-\caption{Taylor parameters for individuals taking antibiotics\cite{antibiotic}.
-Prior to antibiotics intake, the population sampled is described by $\bar{V} = 0.12 \pm 0.05, \bar{\beta} = 0.75 \pm 0.04$.}
+\caption{Taylor's parameters for individuals taking antibiotics\cite{antibiotic}.
+Prior to the antibiotics intake, the population sampled is described by $\bar{V} = 0.12 \pm 0.05, \bar{\beta} = 0.75 \pm 0.04$.}
 
-\caption{Taylor parameters for persons diagnosed with irritable bowel syndrome (IBS)\cite{IBS}.
+\caption{Taylor's parameters for persons diagnosed with irritable bowel syndrome (IBS)\cite{IBS}.
 Healthy individuals sampled in this study are characterized by $\bar{V} = 0.135 \pm 0.010, \bar{\beta} = 0.692 \pm 0.024$.}
 
-\caption{Taylor parameters for the healthy subject of the discordant twins\cite{kwashiorkor}.
+\caption{Taylor's parameters for the healthy subject of the discordant twins\cite{kwashiorkor}.
 This table continues in Supplementary Table S\ref{tab:DK}.
@@
 
-\caption{Taylor parameters for the kwashiorkor part of the discordant twins\cite{kwashiorkor}.
+\caption{Taylor's parameters for the kwashiorkor part of the discordant twins\cite{kwashiorkor}.
 This is a continuation of Supplementary Table S\ref{tab:DH}.
@@
 
-\caption{Taylor parameters for individuals with different degrees of overweight and obesity\cite{LEA}.
-Healthy people in this study, whom were not obese, are characterized by $\bar{V} = 0.19 \pm 0.06, \bar{\beta} = 0.806 \pm 0.034$.}
+\caption{Taylor's parameters for individuals with different degrees of excess weight and obesity\cite{LEA}.
+The healthy people in this study, who were not obese, are characterized by $\bar{V} = 0.19 \pm 0.06, \bar{\beta} = 0.806 \pm 0.034$.}
 
-\caption{Taylor parameters for special intervals concerning gut microbiota in the host lifestyle study\cite{hostlife}.
+\caption{Taylor's parameters for special intervals concerning gut microbiota in the host lifestyle study\cite{hostlife}.
 The healthy and quotidian periods are characterized by $\bar{V} = 0.25 \pm 0.09, \bar{\beta} = 0.777 \pm 0.025$.}

### 2017-01-10 mSystems/rev/mSys_MandM_rev.tex
English corrections applied, which leads to REVISION draft F
On branch master, changes to be committed:
	new file:   ../documentation/mSys_MAIN_revEnglish.doc
	modified:   mSys_MAIN_rev.pdf
	modified:   mSys_MAIN_rev.tex
	modified:   mSys_MandM_rev.tex
	modified:   mSys_discussion_rev.tex
	modified:   mSys_intro_rev.tex
	modified:   mSys_results_rev.tex
	modified:   mSys_suppl_rev.tex

--- 2016-12-21 mSystems/rev/mSys_MandM_rev.tex
@@
 
-We model the microbial abundances across time along the lines of Blumm \textit{et al.}\cite{ranking}. The dynamics of taxon relative abundances is described by the Langevin equation:
+We modeled microbial abundances over time along the lines of Blumm \textit{et al.}\cite{ranking}. The dynamics of taxon relative abundances was described by the Langevin equation:
 \begin{linenomath}
@@
 \end{linenomath}
-where F$_i$ captures the fitness of the taxon i, V corresponds to the noise amplitude and $\xi_i$(t) is a Gaussian random noise with zero mean  $<\xi_i(t)>$ = 0 and variance uncorrelated in time, $<\xi_i(t) \xi_i(t')>$ =  $\delta(t'-t)$, .
-The function $\phi(t)$ ensures the normalization at all times, $\sum x_i(t) = 1$, and corresponds to $\phi(t) = \sum F_i x_i^\alpha + \sum V x_i^\beta \xi_i(t)$.
-The temporal evolution of the probability that a taxon i has a relative abundance $x_i(t)$, P(x$_i$,t), is determined by the Fokker-Planck equation:
+where $F_i$ captured the fitness of the taxon $i$, $V$ corresponded to the noise amplitude and $\xi_i$(t) was a Gaussian random noise with zero mean  $<\xi_i(t)> = 0$, and variance which was uncorrelated over time, $<\xi_i(t) \xi_i(t')> = \delta(t'-t)$.
+The function $\phi(t)$ ensured the normalization at all times, $\sum x_i(t) = 1$, and corresponded to $\phi(t) = \sum F_i x_i^\alpha + \sum V x_i^\beta \xi_i(t)$.
+The temporal evolution of the probability that taxon $i$ had a relative abundance $x_i(t)$, $P(x_i,t)$, was determined by the Fokker-Planck equation:
 \begin{linenomath}
@@
 \end{linenomath}
-The microbiota evolves towards a steady-state with a time-independent probability depending on the values of $\alpha$, $\beta$, F$_i$ and V.
-For $\alpha<1$ (otherwise, systems are always unstable), the steady-state probability may be localized in a region around a preferred value or broadly distributed over a wide range, depending on whether the fitness F$_i$  dominates or is overwhelmed by the noise amplitude V.
-The steady-state solution of the Fokker-Planck equation is given by:
+The microbiota evolved towards a steady state with a time-independent probability dependent on the values of $\alpha$, $\beta$, $F_i$ and $V$.
+For $\alpha<1$ (otherwise, systems are always unstable), the steady state probability was localized in a region around a preferred value or broadly distributed over a wide range, depending on whether the fitness $F_i$ dominated or was overwhelmed by the noise amplitude $V$.
+The steady-state solution of the Fokker--Planck equation was given by:
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-where $\phi_0 = (\sum_i F_i^{1/(1-\alpha)})^{1-\alpha}$ and C$_{ne}$ and C$_{e}$ are integrals that should be solved numerically for the parameters of interest.
-The ordered phase happens when the solution has a maximum in the physical interval ($0<x_i<1$).
-For larger V, the transition to a disordered phase happens when the maximum shifts to the unphysical region $x_i<0$, which sets the phase transition region V($\alpha,\beta,F_i$).
-The phase transition region can be calculated analytically in particular cases:
+where $\phi_0 = (\sum_i F_i^{1/(1-\alpha)})^{1-\alpha}$ and $C_{ne}$ and $C_{e}$ were integrals that were solved numerically for the parameters of interest.
+The ordered phase occurred when the solution had a maximum in the physical interval ($0<x_i<1$).
+For a larger V, the transition to a disordered phase happened when the maximum shifted to the unphysical region $x_i<0$, which sets the phase transition region $V(\alpha,\beta,F_i)$.
+The phase transition region was calculated analytically in specific cases:
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-where the first case, simplifies to $F= 3 V^2$ if $\beta = 0.75$ and the fitness of this taxon dominates in $\phi_0$. 
-In many physical systems (Brownian motion is the classical example \cite{Einstein}), the two terms of the Langevin equation are related.
-The \emph{Fluctuation Dissipation Theorem} states a general relationship between the response to an external disturbance and the internal fluctuations of the system\cite{FD}.
+where the first case, simplified to $F = 3 V^2$ if $\beta = 0.75$ and the fitness of this taxon dominated in $\phi_0$. 
+In many physical systems (Brownian motion is the classic example \cite{Einstein}), the two terms of the Langevin's equation are related.
+The \emph{Fluctuation Dissipation Theorem} states out a general relationship between the response to an external disturbance and the internal fluctuations of the system\cite{FD}.
 The theorem can be used as the basic formula to derive the fitness from the analysis of fluctuations
@@
 \subsection*{Standardization} \label{sec:stan}
-In order to properly show all the studies under common axes, we decided to standardize the Taylor parameters using the group of healthy individuals for each study.
-With this approach, all the studies can be visualized in a shared plot with units of Taylor-parameters standard-deviation on their axes.
+In order to properly show all the studies under common axes, we decided to standardize the Taylor parameters using the group of healthy individuals for each study independently.
+With this approach, all the studies can be visualized in a shared plot with units of Taylor--parameter standard--deviation on their axes.
 
-For a Taylor parameter, e.g. $V$, the estimate of the mean ($\widehat{V}$) for the healthy subpopulation, composed of $h$ individuals, is:
+For a Taylor's parameter, e.g. $V$, the estimate of the mean ($\widehat{V}$) for the healthy subpopulation, composed of $h$ individuals, is:
 \begin{linenomath}
@@
 \end{linenomath}
-being $\sigma_{V_i}$ the estimation of the uncertainty in $V_i$ obtained together with $V_i$ from the X-weighted power-law fit described in Section \ref{sec:X-w}, for healthy individuals.
+$\sigma_{V_i}$ being the estimation of the uncertainty in $V_i$ obtained together with $V_i$ from the X-weighted power-law fit described in Section \ref{sec:X-w}, for healthy individuals.
 
@@
 \end{linenomath}
-being $W_2=\sum_i^h \omega_i^2$, which finally yields to:
+with $W_2=\sum_i^h \omega_i^2$, which finally yields to:
 \begin{linenomath}
@@
 
-The bacteria and archaea taxonomic assignations were obtained by analysing 16S rRNA sequences, which were clustered into operational taxonomic units (OTUs) sharing 97 \% sequence identity using QIIME\cite{QIIME}.
-Shotgun metagenomic sequencing (SMS) data\cite{kwashiorkor} were analysed and assigned at strain level by the Livermore Metagenomic Analysis Toolkit (LMAT)\cite{LMAT}, according to their default quality threshold.
-Genus, with best balance between error assignment and number of taxa, was chosen as our reference taxonomic level.
-We have verified that our conclusions are not significantly affected by selecting family or species as the reference taxonomic level (see Supplementary Figure \ref{supfig:taxlev1}).
+The bacteria and archaea taxonomic assignations were obtained by analyzing 16S rRNA sequences, which were clustered into operational taxonomic units (OTUs) sharing 97\% of their sequence identity using QIIME \cite{QIIME}.
+Shotgun metagenomic sequencing (SMS) data\cite{kwashiorkor} were analyzed and assigned at strain level by the Livermore Metagenomic Analysis Toolkit (LMAT)\cite{LMAT}, according to their default quality threshold.
+Genus, with the best balance between error assignment and number of taxa, was chosen as our reference taxonomic level.
+We verified that our conclusions were not significantly affected by selecting family or species as the reference taxonomic level (see Supplementary Figure \ref{supfig:taxlev1}).
 
@@
 \subsubsection*{Sample selection}
-We have chosen studies about relevant pathologies containing metagenomic sequencing time data series of bacterial populations from humans in different healthy and non-healthy states.
-We have selected only those individuals who had three or more time points of data available in databases.
-Caporaso et al.
-study \cite{moving} was selected as it has two healthy individuals measured over a very large timespan, with almost daily sampling.
-Faith et al.
-study \cite{LEA} was selected due to the BMI differences between subjects, moreover some of them had diets which can be treated as system perturbations.
-We considered healthy only those individuals who had normal or overweight BMI.
-Smith et al., study \cite{kwashiorkor} was selected for both the age of the patients and the rare disease.
-We only worked with the discordant twins, and considered healthy those who were not affected by kwashiorkor in each pair of patients.
-David et al., study \cite{diet} was selected for its differential diets.
-We considered as the healthy part the first time samples of each individual before the diet, and the rest of time points as perturbations.
-Dethlefsen and Relman work \cite{antibiotic} was selected due to the interesting treatment of two antibiotic intakes of the same antibiotic by three different subjects.
-We considered as the healthy part only those times before any antibiotic treatment, and as perturbations the time of antibiotic intakes and the period after that.
-And finally, we also considered a study made in our group carried by Durban et al., \cite{IBS} in which we considered as healthy subjects those who didn't suffer from irritable bowel disease, and as perturbation the patients who had this disease.
+We chose studies about relevant pathologies containing metagenomic sequencing time data series of bacterial populations from humans in different healthy and non-healthy states.
+Only those individuals who had three or more time points of data available in databases were selected.
+The study by Caporaso \textit{et al.} study \cite{moving} was selected as it featured two healthy individuals measured over a very long timespan, with almost daily sampling.
+The study of Faith \textit{et al.} \cite{LEA} was selected given the BMI differences between subjects.
+Moreover, some of them followed diets which could be treated as system perturbations.
+Only those individuals who had normal or overweight BMI were considered as healthy.
+The study by Smith \textit{et al.} \cite{kwashiorkor} was selected for both the age of the patients and the rare disease.
+We only worked with the discordant twins, and considered those who were not affected by kwashiorkor in each pair of patients as being healthy.
+The study by David \textit{et al.} \cite{diet}was selected for its differential diets.
+The healthy part was considered to be the first time samples of each individual before the diet, while the rest of the time points were considered as perturbations.
+Dethlefsen and Relman's work \cite{antibiotic} was selected due to the interesting treatment of two antibiotic intakes by three different subjects of the same antibiotic.
+The healthy part was considered to be only those times before any antibiotic treatment, and the time of antibiotic intakes and the period after that as perturbations.
+The work by David \textit{et al.} \cite{hostlife} was selected due to the comprehensive longitudinal data that it provides plus its complete metadata and the interesting events that happened to both subjects (an infection and a trip abroad).
+The healthy part was taken as the previous time points before each event.
+Finally, we also considered a study from our group carried out by Durban \textit{et al.} \cite{IBS} in which the healthy subjects were considered as those who did not suffer from irritable bowel syndrome, while the patients who had this disease were taken as perturbations.
 
-Metadata of each study is provided in Supplementary Tables S1 to S4.
-All used 16S rRNA gene sequencing except for the study of the discordant kwashiorkor twins\cite{kwashiorkor}, where both SMS and 16S rRNA data were used.
-In the latter case we selected to work with SMS data to show that our method is valid regardless of the source of taxonomic information.
-Each one of the datasets was treated as follows:
+The metadata for each study is provided in Supplementary Tables S1 to S7.
+They all used 16S rRNA gene sequencing, except for the study of the discordant kwashiorkor twins\cite{kwashiorkor}, in which both SMS and 16S rRNA data were used.
+In the latter case, we chose to work with SMS data to show that our method was valid, regardless of the source of taxonomic information.
+Each of the datasets was treated as follows:
 
@@
 \subsubsection*{16rRNA sequences processing}
-Reads from the selected studies were first quality filtered using the FastX toolkit\cite{FASTX}, allowing only those reads which had more than 25 of quality along the 75\% of the complete sequence.
-16S rRNA reads were then clustered at 97\% nucleotide sequence identity (97\% ID) into operational taxonomic units (OTUs) using QIIME package software\cite{QIIME} (version 1.8) We followed open reference OTU picking workflow in all cases.
-The clustering method used was uclust, and the OTUs were matched against Silva database\cite{SILVA} (version 111, July 2012) and were assigned to taxonomy with an uclust-based consensus taxonomy assigner.
-The parameters used in this step were: similarity 0.97, prefilter percent id 0.6, max accepts 20, max rejects 500.
+Reads from the selected studies were first quality filtered using the FastX toolkit\cite{FASTX}, allowing only those reads which had more than a quality score of over 25 in 75\% of the complete sequence.
+16S rRNA reads were then clustered at a 97\% nucleotide sequence identity (97\% ID) into operational taxonomic units (OTUs), using the QIIME software package\cite{QIIME} (version 1.8).
+We followed an open reference OTU picking workflow in all cases.
+The clustering method used was UCLUST, and the OTUs were matched against the Silva database\cite{SILVA} (version 111, July 2012) and were assigned to a taxonomy with a UCLUST-based consensus taxonomy assigner.
+The parameters used in this step were: similarity 0.97, prefilter percent id 0.6, maximum accepts 20, maximum rejects 500.
 
@@
 \subsubsection*{Metagenomic sequences processing}
-Shotgun metagenomic (and 16S too) sequences were analyzed with LMAT (Livermore Metagenomics Analysis Toolkit) software package\cite{LMAT} (version 1.2.4, with Feb'15 release of data base \emph{LMAT-Grand}).
-LMAT was run using a Bull shared-memory node belonging to the team's HPC (high performance computing) cluster.
-It is equipped with 32 cores (64 threads available using Intel Hyper-threading technology) as it has two Haswell-based Xeons ($22\un{nm}$ technology), the E5-2698v3@2.3 GHz, sharing half a tebibyte (0.5 TiB, that is, 512 gibibytes) of DRAM memory.
+Shotgun metagenomic sequences were analyzed with LMAT (Livermore Metagenomics Analysis Toolkit) software package\cite{LMAT} (version 1.2.4, with Feb'15 release of the LMAT-Grand database).
+LMAT was run using a Bull shared-memory node belonging to the team?s HPC (high performance computing) cluster.
+It was equipped with 32 cores (64 threads available using Intel Hyper-Threading Technology) as it has two Haswell-based Xeons (22 nm technology), the E5-2698v3@2.3 GHz, sharing half a tebibyte of DRAM memory.
 This node is also provided with a PCIe SSD card as NVRAM, the Micron P420m HHHL, with 1.4 TB, and 750000 reading IOPS, 4 KB, achieving 3.3 GB/s.
 The computing node was supplied with a RAID-0 (striping) scratch disk area.
-We used the ``Grand'' database\cite{LMAT2}, release Feb'15, provided by the LMAT team, where ``Grand'' refers to a huge database that contains k-mers from all viral, prokaryote, fungal and protist genomes present in the NCBI database, plus Human reference genome (hg19), plus GenBank Human, plus the 1000 Human Genomes Project (HGP) (this represent about 31.75 billion k-mers occupying 457.62 GB)\cite{LMAT2}.
-Previously to any calculation, the full database was loaded in the NVRAM.
-With this configuration the observed LMAT sustained sequence classification rate was $20\unclassrate$.
-Finally, it is worth mentioning that a complete set of Python scripts have been developed as back-end and front-end of the LMAT pipeline in order to manage the added complexity of time series analysis.
+We used the ``Grand'' database\cite{LMAT2}, released in Feb'15, provided by the LMAT team, where ``Grand'' refers to a huge database that contains k-mers from all the viral, prokaryote, fungal and protist genomes present in the NCBI database, plus the Human reference genome (hg19), plus GenBank Human, plus the 1000 Human Genomes Project (HGP) (this represents about 31.75 billion k-mers occupying 457.62 GB)\cite{LMAT2}.
+Before any calculations were made, the entire database was loaded into the NVRAM.
+With this configuration, the observed LMAT sustained sequence classification rate was $20\unclassrate$.
+Finally, it is worth mentioning that a complete set of Python scripts was developed as back-end and front-end of the LMAT pipeline in order to manage the added complexity of time series analysis.
 
@@
 \subsubsection*{Taxa level robustness}
-We selected genus as taxonomic level for the subsequent steps of our work.
-In order to ensure that, between adjacent taxonomic levels, there were not crucial differences which could still be of relevance after standardization (see last subsection of Material and Methods), we tested two different data sets.
+We selected genus as the taxonomic level for the subsequent steps of our work.
+In order to ensure that there were no crucial differences between adjacent taxonomic levels which could still be of relevance after standardization (see the last subsection of Material and Methods), we tested two different data sets.
 In the former, the antibiotics study\cite{antibiotic} with 16S data, we tested the differences between genus and family levels.
@@
   \includegraphics[width=0.99\textwidth]{figs/supfig_taxlev1.eps}
-\caption{Overview of comparison of different approaches based on adjacent taxonomic levels using plots in the Taylor-parameters space.
+\caption{Overview of the comparison of different approaches based on adjacent taxonomic levels using plots in the Taylor-parameters space.
 The former row of subfigures is for 16S, where levels are family (blue circles) vs.
-genus (purple triangles).
-Whereas the latter row of subfigures is for SMS, where levels are genus (blue circles) vs.
+genus (purple triangles), whereas the latter row of subfigures is for SMS, where levels are genus (blue circles) vs.
 species (purple triangles).
-The left column shows the raw results and the right column plots the standardized results (see Standardization under Material and Methods).}
+The left column shows the raw results and the right column plots the standardized results (see Standardization in Material and Methods).}
 \label{supfig:taxlev1}
@@
 When fitting the power-law of std vs.
-mean, we can take into account that every mean has uncertainty and estimate it for a sample size $n$ by the SEM (\emph{Standard Error of the Mean}).
-Here, the uncertainties affect the independent variable, so the fit is not so trivial as a Y-weighted fit, where the uncertainties affect the dependent variable.
-A standard approach to do this fit is: a) invert your variables before applying the weights, b) then perform the weighted fit, and finally, c) revert the inversion.
+mean, we took into account that every mean has uncertainty and can be estimated for a sample size $n$ by the SEM (\emph{Standard Error of the Mean}).
+Here, the uncertainties affected the independent variable, so the fit was not as trivial as a Y-weighted fit, where the uncertainties affect the dependent variable.
+A standard approach for this fit is: a) to invert the variables before applying the weights, b) then perform the weighted fit, and finally, c) revert the inversion.
 This method is deterministic, but the approximate solution worsens with smaller coefficients of determination.
-To overcome this limitation, we developed a stochastic method by using a bootstrapping-like strategy that avoids the inversion and is applicable regardless of the coefficient of determination.
+To overcome this limitation, we developed a stochastic method by using a bootstrapping-like strategy that avoided inversion and could be applied regardless of the coefficient of determination.
 
 The basic idea of bootstrapping is that inference about a population from sample data (sample $\rightarrow$ population) can be modeled by resampling the sample data and performing inference on (resample $\rightarrow$ sample)\cite{boot}.
-To adapt this general idea to our problem, we resample the x-data array using its errors array.
-That is, for each replicate, a new x-data array is computed based on:
+To adapt this general idea to our problem, we resampled the x-data array using its errors array.
+That is, for each replicate, a new x-data array was computed based on:
 \begin{linenomath}
@@
 where $v_i$ is a Gaussian random variable with mean $\mu_i=0$ and standard deviation $\sigma_i=\mathrm{SEM}_i$, as defined previously.
-For each replicate a complete un-weighted power-law fit is performed, where to choose between fitting power laws ($y=Vx^\beta$) using linear regression on log-transformed (LLR) data versus non-linear regression (NLR) we mainly follow \emph{General Guidelines for the Analysis of Biological Power Laws}\cite{biopwrlaw}.
-The parameters of the X-weighted fit are then estimated by averaging through all the replicate fits performed, and their errors are estimated by computing the standard deviation also for all the fits.
-At the end of each step, the relative error is calculated by comparing the fit parameters estimation in the last step with the previous one.
-Finally, both the coefficient of determination of the fit and the coefficient of correlation between the fit parameters are estimated by averaging.
+For each replicate, a complete unweighted power-law fit was performed, where in order to choose between fitting power laws ($y=Vx^\beta$) using linear regression on log-transformed (LLR) data versus non-linear regression (NLR) we mainly followed the \emph{General Guidelines for the Analysis of Biological Power Laws}\cite{biopwrlaw}.
+The parameters of the X-weighted fit were then estimated by averaging through all the replicate fits performed, and their errors were estimated by computing the standard deviation for all the fits.
+At the end of each step, the relative error was calculated by comparing the fit parameter estimation in the last step with the previous one.
+Finally, both the coefficient of determination of the fit and the coefficient of correlation between the fit parameters were estimated by averaging.
 
@@
 
-The Rank Stability Index (RSI) is shown as a percentage in a separate bar on the right of the rank matrix plot shown in Figures \ref{fig:corrank_IBS_A} and \ref{fig:corrank_IBS_P2}.
-The RSI is strictly $1$ for an element whose range never changes over time, and is strictly $0$ for an element whose rank oscillates between the extremes from time to time.
-So, RSI is calculated, per element, as $1$ less the quotient of the number of true rank hops taken between the number of maximum possible rank hops, all powered to $p$:
+The Rank Stability Index (RSI) is shown as a percentage in a separate bar on the right of the rank matrix plot in Figures \ref{fig:corrank_IBS_A} and \ref{fig:corrank_IBS_P2}.
+The RSI is strictly $1$ for an element whose range never changes over time, and is strictly $0$ for an element whose rank oscillates between the extremes over time.
+So, the RSI is calculated, per element, as $1$ less the quotient of the number of true rank hops taken between the number of maximum possible rank hops, all powered to $p$:
 \begin{linenomath}
@@
 where $D$ is the total of rank hops taken by the studied element, $N$ is the number of elements that have been ranked, and $t$ is the number of time samples.
-The power index $p=4$ is arbitrarily chosen to increase the resolution in the stable region.
+The power index $p=4$ was arbitrarily chosen to increase the resolution in the stable region.
 

### 2017-01-10 mSystems/rev/mSys_discussion_rev.tex
English corrections applied, which leads to REVISION draft F
On branch master, changes to be committed:
	new file:   ../documentation/mSys_MAIN_revEnglish.doc
	modified:   mSys_MAIN_rev.pdf
	modified:   mSys_MAIN_rev.tex
	modified:   mSys_MandM_rev.tex
	modified:   mSys_discussion_rev.tex
	modified:   mSys_intro_rev.tex
	modified:   mSys_results_rev.tex
	modified:   mSys_suppl_rev.tex

--- 2016-12-23 mSystems/rev/mSys_discussion_rev.tex
@@
-One of the mains features of this work is to have shown that, independently of its condition, the microbiota follows the Taylor's law.
-We have seen that the value of the scaling index in each case is always less than the unity (using standard deviation as the measurement for dispersion), which is informing us about the community structure.
+One of the mains features of this work is to have demonstrated that, independently of its condition, microbiota follows Taylor's law.
+We have seen that the value of the scaling index in each case is always less than the unity (using standard deviation as the measurement for dispersion), which provides us with information about the community structure.
 This means that, in relative terms, the most abundant elements in the population are less volatile to perturbations than the less abundant ones.
-The explanation for this universal pattern is not clear although some hypothesis have been tested in other studies, as the presence of negative interactions in the population \cite{kilpatrick}, or the demonstration that it may depend on reproductive correlation \cite{ballantyne}.
-Nevertheless, none of these explanations are enough when we are talking about microbiota as the reproduction term is diffuse, the interactions between its components are not only based on competition \cite{joao, mehta, bucci} and that even that kind of negative interaction may not effectively yield in values less than the unity when referring to a bacterial species \cite{ramslayer}.
-Anyhow, the values obtained in all cases are very similar among them, which could be suggesting that the community structure is preserved throughout the different scenarios that we have studied.
+The explanation for this universal pattern is not clear although some hypotheses have been tested in other studies, such as the presence of negative interactions in the population \cite{kilpatrick}, and a demonstration that this may depend on reproductive correlation \cite{ballantyne}.
+Nevertheless, none of these explanations are sufficient when we are talking about microbiota, as the reproduction term is diffuse, the interactions between its components are not only based on competition \cite{joao, mehta, bucci}, and that even that kind of negative interaction may not effectively yield values less than the unity when referring to a bacterial species \cite{ramslayer}.
+Anyhow, the values obtained in all cases were very similar from one to another, which could suggest that the community structure is preserved throughout the different scenarios we studied.
 
-The second parameter is informing about the noise and can be directly related with the variability or the fluctuation amplitude of the population over time.
+The second parameter provides information about noise and can be directly linked to the variability or fluctuation amplitude of the population over time.
 It is a direct estimator of the stability of the system under study.
-As we showed above, the healthy subset of each study have lower variability than the non-healthy subset when dealing with adult individuals.
-Interestingly, the variability parameter was higher in the healthy subset for the study of the discordant twins suffering from kwashiorkor disease \cite{kwashiorkor}.
-In this regard it has been shown that the infant microbiota needs to develop toward a definite, adult state \cite{koenig}.
-This implies that the temporal variability would be greater in children respect to a healthy adult state, which should be temporally stable.
-Thus, our results could be pointing out the need of this variability in order to reach that adult state.
+As we have shown above, the healthy subset of each study has lower variability than the non--healthy subset, when dealing with adult individuals.
+Interestingly, the variability parameter was higher in the healthy subset in the study of the discordant twins suffering from kwashiorkor disease \cite{kwashiorkor}.
+In this regard, it has been shown that infant microbiota needs to develop toward a definite, adult state \cite{koenig}.
+This implies that temporal variability is greater in children compared to a healthy adult state, which should be temporally stable.
+Thus, our results could point to the need for this variability in order to reach that adult state.
 Furthermore, as we wanted to see how this variability behaved over time, we calculated the evolution of this parameter for the samples which had enough time sampling.
-As can be seen in Figure \ref{fig:tempevo1}, the variability of the microbiota has some fluctuations over time.
-It is interesting to note in Figure \ref{fig:tempevo2} how this parameter can capture the two antibiotic intakes in one of the patients from the study of Dethlefsen and Relman \cite{antibiotic}, especially that it seems to be some resilience process in the microbiota due to the lower variability increase in the second antibiotic intake.
+As shown in Figure \ref{fig:tempevo1}, the variability of microbiota fluctuated over time.
+It is interesting to note in Figure \ref{fig:tempevo2} how this parameter captured the two antibiotic intakes in one of the patients from the study by Dethlefsen and Relman \cite{antibiotic}, especially in that there seems to be some kind of a resilience process in the microbiota due to the lower variability increase in the second antibiotic intake.
 
-The primary hypothesis of this work is that, in adults, having a healthy microbiota means that population is stable in time and does not travel into a state where it is highly susceptible to external or internal perturbations, causing in such a case a dysbiotic state in the microbiota.
-In order to use the valuable information which gives us the empirical law of Taylor's work, we propose the use of Langevin equation to model how the ranking stability evolves in time.
-While we can measure directly the component of the noise of the system as their variability, the other main term needs to be inferred from the model.
-This term, which we have named as 'fitness', is the one that gives the ability to the system to be stable to potential perturbations.
-In ecological terms, this could mean the nature of interactions that are present among the bacteria, between bacteria and other minority populations as fungi or archaea, between bacteria and the viral component in the microbiota, and the interactions between host and the whole microbiota.
-Being this a first step to model the temporal stability of the microbiota and due to its complicated nature, we have calculated the fitness term using the Fluctuation Dissipation Theorem as a first approximation\cite{FD}.
-Thus, the fitness of the microbiota still remains to be modelled in future works in order to make the model more accurate and with a higher predictive power.
+The primary hypothesis of this work is that, in adults, having a healthy microbiota means that the population is stable over time and does not move into a state where it is highly susceptible to external or internal perturbations, causing a dysbiotic state in the microbiota.
+In order to use the valuable information provided by the empirical law of Taylor's work, we proposed the use of Langevin's equation to model how ranking stability evolved over time.
+While the system noise component can be directly measured as its variability, the other main term needs to be inferred from the model.
+This term, which we named "fitness", is the one that enables the system to be stable in the face of potential perturbations.
+In ecological terms, this could represent the nature of interactions that are present among bacteria, between bacteria and other minority populations, such as fungi or archaea, between bacteria and the viral component in microbiota, and interactions between the host and the whole microbiota.
+As this is a first step to model the temporal stability of microbiota, and given its complex nature, we calculated fitness using the Fluctuation Dissipation Theorem as a first approximation\cite{FD}.
+Thus, the fitness of microbiota will still need to be modeled in future works in order to make the model more accurate and give it a higher predictive power.
 
-By solving the Langevin differential equation, we can obtain a phase diagram where each microbiota sample can be placed according to its fitness and variability in one of two phases according to the ranking stability of the system.
-As we can see in the phase-space in Figure \ref{fig:main3}, we are showing three different conditions that could happen.
-First, we can have a healthy microbiota which could have some fluctuations as showed by one of the subjects of Caporasso et al study \cite{moving}.
-Because the fitness of this cases will be high enough, the temporal variability will not place the microbiota in the unstable phase of the diagram.
-Second, we have a subject from the study of Dethlefsen and Relman \cite{antibiotic} which is perturbed twice by an antibiotic intake.
-His microbiota is altered enough to lose its stability, and hence be placed in the unstable part.
-So located, it is more sensitive to potential perturbations as, for example, opportunist infections.
-Third and last condition, the subject is already in the unstable phase due to some healthy issue as IBS.
-This can be observed in one of the patients from Durban et al study \cite{IBS}.
-In addition, it was shown that this subject improved its healthy status in the time when the experiment was done, implying that his microbiota also recovered the lost stability.
-It is interesting to notice that in the subject from the study of David et al.
-\cite{hostlife}, who suffered a Salmonella infection during the experiment, we can observe a huge shift in the variability and a final recovery from the perturbed state (see \ref{supfig:HLS_xWSummary}).
+By solving Langevin's differential equation, we can obtain a phase diagram where each microbiota sample can be placed according to its fitness and variability into one of two phases, according to the ranking stability of the system.
+As we can see in the phase--space in Figure \ref{fig:main3}, three different conditions that could occur are shown.
+First, we could have a healthy microbiota with some fluctuations, as shown by one of the subjects of Caporaso \emph{et al.}'s study \cite{moving}.
+Because this case would have good fitness, its temporal variability would not place the microbiota in the unstable phase of the diagram.
+Second, we have a subject from the study by Dethlefsen and Relman \cite{antibiotic} who was perturbed twice by an antibiotic intake.
+His microbiota underwent sufficient change so as to lose its stability, and hence be placed in the unstable part.
+In this location, it is more sensitive to potential perturbations such as, for example, opportunist infections.
+In the third and last condition, the subject was already in the unstable phase due to a health issue, i.e.
+IBS.
+This can be observed in one of the patients from Durban \emph{et al.}'s study \cite{IBS}.
+In addition, it was shown that this subject's health status improved during the time the experiment was carried out, implying that his microbiota also recovered the stability it had lost.
+It is interesting to note that in the subject's health from the study made by David \emph{et al.} \cite{hostlife} who suffered a Salmonella infection during the experiment, there was a significant shift in variability and a final recovery from the perturbed state (see \ref{supfig:HLS_xWSummary}).
 
@@
 	\includegraphics[width=0.99\textwidth]{figs/supfig_HLS_xWSummary.eps}
-	\caption{Taylor's law parameter space for intervals concerning gut microbiota in the host lifestyle study\cite{hostlife}.}
+	\caption{Taylor's law parameter space for intervals concerning gut microbiota in the host lifestyle study\cite{hostlife}.
+We observe that subject \emph{B}, who suffered a Salmonella infection during the experiment, had a relevant shift in the parameters from \emph{\_before} to \emph{\_infection} and a final recovery from the perturbed state to \emph{\_afterinfec}, which lies in the parameter area compatible with the healthy and stable intervals (see Supplementary Table S7).
+Subject \emph{A} also had a shift in variability from \emph{\_before} to \emph{\_abroad} and back to \emph{\_returned}, also in the proximity zone of healthy and stable periods.}
 	\label{supfig:HLS_xWSummary}
@@
 
-Specifically, the analysis of the rank stability of the samples of healthy and IBS diagnosed patients studied in our lab\cite{IBS}, suggests that the presence of \emph{rank stability islands} among medium-ranked taxa is a interesting feature.
+Specifically, the analysis of the rank stability of the samples of healthy and IBS--diagnosed patients studied in our lab\cite{IBS}, suggests that the presence of \emph{rank stability islands} among medium-ranked taxa is a interesting feature.
 The higher stability of these taxa goes against the global meaning of the scaling index.
-Interestingly, that stability disappears when we look at the IBS patients.
+Interestingly, this stability disappears when we look at IBS patients.
 From the literature, it seems that some genera from the families Comamonadaceae, Neisseriaceae and Carnobacteriaceae have been reported to lower their abundance in IBS patients against healthy controls \cite{sislands1}.
@@
 It could be brought into question the rule of these taxa as key players in the phase transition of the microbiota, or whether they are more susceptible to perturbations than the most abundant.
-The types of interactions that could be sustaining this particular behavior are not clear, as these non-abundant taxa are not usually included in dynamical studies in order to get the community matrix.
-Further experiments and data analysis is needed to clarify if this is not a unique event, or it is a widespread feature of stable microbiotas.
+The types of interactions that could sustain this particular behavior are not clear, as these non-abundant taxa are not usually included in dynamic studies in order to obtain a community matrix.
+Further experiments and data analysis are needed to clarify whether this was a unique event, or whether it is a widespread feature of stable microbiotas.
 
-However, we have to be aware that the hypothesis above is too simplistic to be directly related with reality.
-It has been demonstrated that the situation is more complex than the outlook provided separating healthy people from non-healthy people just by compositional terms, as Moya and Ferrer underline in their recent review \cite{Moya_trends}.
-There are several different feasible scenarios in which we can consider the microbiota as stable independently of their compositional evolution over time.
-For example, depending on their ability to recover the initial composition (resilience), or whether it can recover the original function despite the composition (functional redundancy).
-What we have shown in this work could be explained as the transitions of a stable microbiota into a dysbiosis state.
+However, we have to be aware that the hypothesis above is too simplistic to be directly related to reality.
+It has been demonstrated that the situation is more complex than the outlook provided which separate healthy people from non-healthy people just by compositional terms, as Moya and Ferrer underlined in their recent review \cite{Moya_trends}.
+There are several different feasible scenarios in which we can consider microbiota as being stablem irrespective of their compositional evolution over time.
+For example, depending on its ability to recover its initial composition (resilience), or whether it can recover its original function despite its composition (functional redundancy).
+What we have shown in this work could be explained as the transition of stable microbiota into a state of dysbiosis.
 
-As a first step toward understanding the microbiota stability, the model presents some limitations and there is still work to do.
-From the biological perspective, many questions arise from this work.
-We have observed the same pattern in Taylor's parameters in all the different conditions we have studied, but a pertinent question is whether it is really a universal feature in the huge diversity of microbial niches.
-Furthermore, another relevant question is which mechanisms are involved in maintaining the population structure.
-The nature of the interactions among the elements of the community is surely of great importance in this matter, and it is related to the fitness of the community as has been commented above.
-How we should address the community fitness is not clear, but works as Tikhonov's \cite{tikhonov} could help us to aim at the correct direction toward unraveling the complexity of the microbiota.
+As a first step towards understanding microbiota stability, the model presents some limitations and there is still work to do.
+From a biological perspective, many questions arise from this work.
+We have observed the same pattern in Taylor's parameters in all the different conditions we studied, but a pertinent question is whether this is really a universal feature in the huge diversity of microbial niches.
+Furthermore, another relevant question centers on which mechanisms are involved in maintaining the population structure.
+The nature of the interactions between the elements of community is surely of great importance in this matter, and this is related to the fitness of the community, as mentioned above.
+How we should address community fitness is not clear, but works such as the one by Tikhonov \cite{tikhonov} could point us in the right direction to unravel the complexity of microbiota.

### 2017-01-10 mSystems/rev/mSys_intro_rev.tex
English corrections applied, which leads to REVISION draft F
On branch master, changes to be committed:
	new file:   ../documentation/mSys_MAIN_revEnglish.doc
	modified:   mSys_MAIN_rev.pdf
	modified:   mSys_MAIN_rev.tex
	modified:   mSys_MandM_rev.tex
	modified:   mSys_discussion_rev.tex
	modified:   mSys_intro_rev.tex
	modified:   mSys_results_rev.tex
	modified:   mSys_suppl_rev.tex

--- 2016-06-23 mSystems/mSys_intro.tex
@@
-The desire to understand the factors that influence human health and cause diseases has always been one of the major driving forces of biological research.
-As evidence of new concepts 'holobiont' and 'hologenome' is increasing each day \cite{holo1, holo2}, research not only focus on the human physiology but also on the microbial population that surround ourselves.
-However, these concepts are still in debate \cite{holo3}.
-We are populated by a myriad of microorganisms that are interacting with us in several physiological processes such as metabolism of the bile acids \cite{bileacids}, of the choline \cite{choline} or key-route metabolites as short-chain fatty acids \cite{scfa1, scfa2} which are also involved in immune system maturation \cite{scfa3, scfa4}.
-Human microbiota has been suggested to be closely related to diseases like type 2 diabetes \cite{diabetes2}, cardiovascular disease (CVD) \cite{CVD}, irritable bowel syndrome \cite{IBS}, Crohn's disease \cite{CD}, some affections as obesity \cite{ob1, ob2}, malnutrition \cite{nutr} among other multiple diseases \cite{Moya_trends}.
-High throughput methods for microbial 16S ribosomal RNA gene and WGS have now begun to reveal the composition of archaeal, bacterial, fungal and viral communities located both, in and on the human body.
+The desire to understand the factors that influence human health and cause disease has always been one of the major driving forces of biological research.
+As evidence of the new "holobiont"  and "hologenome" concepts is increasing each day \cite{holo1, holo2}, research not only focuses on the human physiology but also on the microbial population that surrounds us.
+However, these concepts are still under debate \cite{holo3}.
+We are populated by a myriad of microorganisms that interact with us in several physiological processes such as the metabolism of bile acids \cite{bileacids}, of choline \cite{choline} and key-route metabolites, such as short-chain fatty acids \cite{scfa1, scfa2} which are also involved in immune system maturation \cite{scfa3, scfa4}.
+Human microbiota has been suggested to be closely related to diseases like type 2 diabetes \cite{diabetes2}, cardiovascular disease (CVD) \cite{CVD}, irritable bowel syndrome \cite{IBS}, Crohn's disease \cite{CD}, some affections like obesity \cite{ob1, ob2} and malnutrition \cite{nutr} as well as other multiple diseases \cite{Moya_trends}.
+Current studies reveal that gut microbiota also influences brain function and behaviour and is related to neurological disorders like Alzheimer's disease through the brain-gut-microbiome axis\cite{mind,AD}.
+Recently, even a mystifying and elusive condition which is hard to diagnose like chronic fatigue syndrome, which has often been suggested to be a psychosomatic disease, has been closely related to reduced diversity and altered composition of the gut microbiome\cite{CFS}.
+
+High throughput methods for microbial 16S ribosomal RNA gene and SMS (shotgun metagenomic sequencing) have now begun to reveal the composition of archaeal, bacterial, fungal and viral communities located both, in and on the human body.
 Modern high-throughput sequencing and bioinformatics tools provide a powerful means of understanding how the human microbiome contributes to health and its potential as a target for therapeutic interventions \cite{microb&health}.
-To define normal microbiota and how it's compositional changes can origin some diseases are important issues still in need for scientific answers \cite{normal1, normal2}.
+To define normal host-gut microbe interactions and how microbiota compositional changes can cause some diseases are important issues that still require scientific answers \cite{normal1, normal2, panthropology}.
 
-Biology has recently acquired new technological and conceptual tools to investigate, model and understand living organisms at the system level, thanks to the spectacular progress in quantitative techniques, large-scale measurement methods and the integration of experimental and computational approaches.
-In particular, Systems Biology has placed a great effort to unveil the general laws governing the complex behaviour of microbial communities \cite{sysbio&microb, msys1, metasysbio}, even proposing that they have universal dynamics \cite{uni_dynam}.
-Microbiota can be approached under the light of ecological theory where we can find, for instance, general principles as the Taylor's law \cite{taylor}, which relates spatial or temporal variability of the population with its mean.
-This law, also known as fluctuation scale law, is ubiquitous in the natural world and can be found in several systems as random walks \cite{randomwalks}, stock markets \cite{economics1, economics2}, animal populations \cite{taylor, animal1, animal2}, gene expression \cite{genexpress}, or in the human genome \cite{genome}.
-Taylor's law has been applied to microbiota in a spatial way in the work of Zhang {\it et al.}, (2014) \cite{isme1}, where they show that this population tend to be in an aggregated way rather than in a random distribution.
-Despite its ubiquity, it has been studied only in experimental settings \cite{cohen_bac, ramslayer} but never been applied in follow-up studies from microbiota even that a great effort has been made to infer the community structure from a dynamical point of view \cite{cobas, schloss, ravel}
+Biology has recently acquired new technological and conceptual tools to investigate, model and understand living organisms at system level, thanks to spectacular progress in quantitative techniques, large-scale measurement methods and the integration of experimental and computational approaches.
+In particular, Systems Biology has made great efforts to reveal the general laws governing the complex behaviour of microbial communities \cite{sysbio&microb, msys1, metasysbio}, including a proposal suggesting they have universal dynamics \cite{uni_dynam}.
+Microbiota can be approached under the light of ecological theory which includes general principles like Taylor's law \cite{taylor} that relates the spatial or temporal variability of the population with its mean.
+This law, also known as fluctuation scale law, is ubiquitous in the natural world and can be found in several systems like random walks \cite{randomwalks}, stock markets \cite{economics1, economics2}, tree \cite{cohen_taylor} and animal populations \cite{taylor, animal1, animal2}, gene expression \cite{genexpress}, and the human genome \cite{genome}.
+Taylor's law has been applied to microbiota in a spatial way in the work of Zhang {\it et al.}, (2014) \cite{isme1}, where they show that this population tends to be an aggregated one rather than having a random distribution.
+Despite its ubiquity, it has only been studied in experimental settings \cite{cohen_bac, ramslayer} but has never been applied in follow-up studies on microbiota, even though major efforts have been made to infer the community structure from a dynamic point of view \cite{cobas, schloss, ravel}
 
-Here we present the imprints of health status (healthy or disease) in macroscopic properties of microbiota, by studying its temporal variability.
-We have analyzed more than 35000 time series of taxa from the gut microbiome of 97 individuals obtained from publicly available high throughput sequencing data on different conditions: diseases, diets, obese status, antibiotic therapy and healthy individuals.
-Having seen that all cases follows Taylor's law, we use this empirical fact to model how the relative abundances of taxa evolves toward time thanks to the Langevin equation, in a similar way as it was applied recently by Blumm {\it et al.} \cite{ranking}.
-We use this mathematical framework to explore the temporal stability of the microbiota in different conditions in order to understand how this affects the healthy status of the subjects.
+This paper presents the imprints of health status (healthy or disease) in the macroscopic properties of microbiota, by studying its temporal variability.
+We analyzed more than 35000 time series of taxa from the gut microbiome of 99 individuals obtained from publicly available high throughput sequencing data about different conditions: diseases, diets, obese status, antibiotic therapy and healthy individuals.
+Having seen that all the cases followed Taylor's law, we used this empirical fact to model how the relative abundances of taxa evolved over time thanks to the Langevin equation, in a similar way as the approach applied recently by Blumm {\it et al.} \cite{ranking}.
+We used this mathematical framework to explore the temporal stability of microbiota under different conditions in order to understand how this affected the healthy status of the subjects.

### 2017-01-10 mSystems/rev/mSys_results_rev.tex
Minor enhancements
Changes to be committed:
	modified:   rev/figs/Fig2.eps
	modified:   rev/figs/Fig2/Fig2.pdf
	modified:   rev/mSys_MAIN_rev.pdf
	modified:   rev/mSys_MAIN_rev.tex
	modified:   rev/mSys_discussion_rev.tex
	modified:   rev/mSys_results_rev.tex
English corrections applied, which leads to REVISION draft F
On branch master, changes to be committed:
	new file:   ../documentation/mSys_MAIN_revEnglish.doc
	modified:   mSys_MAIN_rev.pdf
	modified:   mSys_MAIN_rev.tex
	modified:   mSys_MandM_rev.tex
	modified:   mSys_discussion_rev.tex
	modified:   mSys_intro_rev.tex
	modified:   mSys_

--- 2016-12-21 mSystems/rev/mSys_results_rev.tex
@@
-We have analyzed the microbiome temporal variability to extract global properties of the system.
-As fluctuations in total counts are plagued by systematic errors we worked on temporal variability of relative abundances for each taxon.
-Our first finding was that, in all cases, changes in relative abundances of taxa follow a ubiquitous pattern known as the fluctuation scaling law\cite{fs} or Taylor's power law\cite{taylor}, i.e., microbiota of all detected taxa follows $\sigma_i  = V\cdot x_i^{\beta}$, a power law dependence between mean relative abundance $x_i$ and dispersion $\sigma_i$.
+The microbiome temporal variability was analyzed to extract the global properties of the system.
+As fluctuations in total counts are plagued by systematic errors we worked on the temporal variability of relative abundances for each taxon.
+Our first finding was that, in all cases, changes in the relative abundances of taxa followed a ubiquitous pattern, known as the fluctuation scaling law\cite{fs} or Taylor's power law\cite{taylor}, i.e., the microbiota of all detected taxa followed $\sigma_i  = V\cdot x_i^{\beta}$, a power law dependence between the mean relative abundance $x_i$ and the dispersion $\sigma_i$.
 The law seem to be ubiquitous, spanning even to six orders of magnitude in the observed relative abundances.
-As can be seen in Figure \ref{fig:main1}, the most abundant species are less volatile in relative terms than the less abundant.
-The fitting to the power law is always robust ($R^{2}$ > 0.88) and does not depend on the microbiome condition.
-The power law (or scaling) index $\beta$ and the variability $V$ (hereafter Taylor parameters) appear to be correlated with the stability of the community.
-.
-The variability V is a direct estimator of amplitude of fluctuations over time, while the scaling index give us information about the statistic properties of the ecosystem.
-The fact that it is less than the unity (when looking at the standard deviation) tell us that the most abundant taxa are less susceptible to any perturbation than the less abundant elements of the population.
-The Taylor parameters are related with the health status of the host, which we consider the main finding exposed in this article.
+As shown in Figure \ref{fig:main1}, the most abundant species were less volatile in relative terms than those which were less abundant.
+The fitting to the power law was always robust ($R^{2}$ > 0.88) and did not depend on the microbiome condition.
+The power law (or scaling) index $\beta$ and the variability $V$ (hereafter Taylor's parameters) appear to be correlated with the stability of the community.
+The variability V was a direct estimator of the amplitude of fluctuations over time, while the scaling index gave us information about the statistical properties of the ecosystem.
+The fact that it was less than the unity (when looking at the standard deviation) tells us that the most abundant taxa were less susceptible to any perturbation than the less abundant elements of the population.
+The Taylor parameters were related to the health status of the host, which we consider as constituting the main finding contributed by this article.
 
@@
 	\includegraphics[width=0.8\textwidth]{figs/Fig1.eps}
-	\caption{X-weighted power-law fits of the standard deviations versus the mean values for each bacterial genus monitored in time.
-We show the fit for samples from a healthy subject (top) and from a subject diagnosed with irritable bowel syndrome (bottom), studied in our lab \cite{IBS}.
+	\caption{X-weighted power-law fits of the standard deviations versus the mean values for each bacterial genus monitored over time.
+The fit is shown for samples from a healthy subject (top) and from a subject diagnosed with irritable bowel syndrome (bottom), studied in our lab \cite{IBS}.
 Taylor's power law seems to be ubiquitous, spanning to six orders of magnitude.}
@@
 
-Taylor parameters describing the temporal variability of the gut microbiome in our sampled individuals are shown in Supplementary Tables S1 to S6.
-Our results hint at an ubiquitous behavior.
-On the first hand, the variability (which corresponds to the maximum amplitude of fluctuations) is large, which suggests resilient capacity of the microbiota.
-On the other hand, the scaling index is always smaller than one, which means that more abundant taxa are less volatile than less abundant ones.
-In addition, Taylor parameters for the microbiome of healthy individuals in different studies are compatible within estimated errors.
-This enables us to define an area in the Taylor parameter space that we called the \emph{healthy zone}.
+Taylor's parameters describing the temporal variability of the gut microbiome in our sampled individuals are shown in Supplementary Tables S1 to S7.
+Our results hint at ubiquitous behavior.
+Firtly, the variability (which corresponds to the maximum amplitude of fluctuations) was large, which suggests the resilient capacity of the microbiota.
+Secondly, the scaling index was always smaller than one, which means that more abundant taxa were less volatile than less abundant ones.
+In addition, Taylor's parameters for the microbiome of healthy individuals in different studies were compatible with estimated errors.
+This enabled us to define an area in the Taylor parameter space that we called the \emph{healthy zone}.
 
-In order to jointly visualize and compare the results of individuals from different studies\cite{IBS,moving,antibiotic,LEA,kwashiorkor,diet}, their Taylor parameters have been standardized, where standardization means that each parameter is subtracted by the mean value and divided by the standard deviation of the group of healthy individuals for each study (for details of the procedure, please see Standardization subsection in Material and Methods).
-The healthy zone and the standardized Taylor parameters for individuals whose gut microbiota is compromised (i.e., suffering from kwashiorkor, altered diet, antibiotics or IBS) is shown in Figure \ref{fig:main2}.
-Children developing kwashiorkor show smaller variability than their healthy twins.
-A meat/fish-based diet increases the variability significantly when compared to a plant-based diet.
-All other cases presented increased variability, which is particularly severe, and statistically significant at more than 95\% CL, for obese patients grade III on a diet, individuals taking antibiotics or IBS--diagnosed patients.
-A global property emerges from all worldwide data collected: Taylor parameters characterize the statistical behavior of microbiome changes.
-Furthermore, we have verified that our conclusions are robust to systematic errors due to taxonomic assignment (see Taxa level selection in Material and Methods).
+In order to jointly visualize and compare the results of individuals from different studies\cite{IBS,moving,antibiotic,LEA,kwashiorkor,diet,hostlife}, their Taylor parameters were standardized, with standardization meaning that each parameter was subtracted by the mean value and divided by the standard deviation of the group of healthy individuals for each study independently (for details of the procedure, please see Standardization subsection in Material and Methods).
+The healthy zone and the standardized Taylor parameters for individuals whose gut microbiota was compromised (i.e., they were suffering from IBS, kwashiorkor, altered diet, intake of antibiotics, a Salmonella infection, or had gone on a trip abroad) are shown in Figure \ref{fig:main2}.
+The variability in children developing kwashiorkor was smaller than that of their healthy twins.
+A meat/fish-based diet significantly increased variability when compared to a plant-based diet.
+All other cases presented increased variability, which was particularly severe and statistically significant at over 95\% confidence level (CL), for grade III obese patients on a diet, individuals taking antibiotics, the subject who had a Salmonella infection, the subject who did a travel abroad or the IBS-diagnosed patients.
+One global property emerged from all the worldwide data collected: the Taylor's parameters characterized the statistical behavior of microbiome changes.
+Furthermore, we verified that our conclusions were robust to systematic errors as a result of taxonomic assignment (see Taxa level selection in Material and Methods).
 
@@
 	%\includegraphics[width=0.9\textwidth]{results/finalplot11.pdf}
-	\includegraphics[width=0.9\textwidth]{figs/Fig2.eps}
+	\includegraphics[width=1.0\textwidth]{figs/Fig2.eps}
 	\caption{Taylor's law parameter space.
-We have compiled here all the data studied in this work.
-The coloured circle corresponds to 68\% confidence level (CL) region of healthy individuals in the Taylor parameter space, while dashed line delimites the 98\% CL region.
-Points with errors place gut microbiome in the Taylor space, for each individual whose microbiota was compromised.
-Note that the parameters have been standardized (standard deviation units) to the healthy group in each study for demonstrative and comparative purposes.}
+All the data studied in this work were compiled here.
+The colored circle corresponds to a 68\% confidence level (CL) region of healthy individuals in the Taylor's parameter space, while the dashed line delimites the 98\% CL region.
+Points with errors place gut microbiome in the Taylor's parameter space, for each individual whose microbiota was compromised.
+It should be noted that the parameters were standardized (standard deviation units) to the healthy group in each study for demonstrative and comparative purposes.}
 	\label{fig:main2}
@@
 
-Taylor's power law has been explained in terms of various effects, all without general consensus.
-It can be shown to have its origin in a mathematical convergence similar to the central limit theorem, so virtually any statistical model designed to produce a Taylor law converge to a Tweedie distribution\cite{stat}, providing a mechanistic explanation based on the statistical theory of errors\cite{convergence1,convergence2,convergence3}.
-To unveil the generic mechanisms that drive different scenarios in the $\beta$--$V$ space, we model the system by assuming that taxon relative abundance follows a Langevin equation with, on the one hand, a deterministic term that captures the fitness of each taxon and, on the other hand, a randomness term associated with Gaussian random noise\cite{ranking}.
-Both terms are modeled by power laws, with coefficients that can be interpreted as the taxon fitness $F_i$ and the variability $V$ (see Model under Material and Methods).
+Taylor's power law has been explained in terms of various effects, though none have brought a general consensus.
+It can be shown to have its origin in mathematical convergence which is similar to the central limit theorem, and thus virtually any statistical model designed to produce a Taylor law converges to a Tweedie distribution\cite{stat}, providing a mechanistic explanation based on the statistical theory of errors\cite{convergence1,convergence2,convergence3}.
+To reveal the generic mechanisms that drive different scenarios in the $\beta-V$ space, we modeled the system by assuming that taxon relative abundance followed a Langevin equation with, on the one hand, a deterministic term that captured the fitness of each taxon and, on the other hand, a randomness term associated with Gaussian random noise\cite{ranking}.
+Both terms were modeled by power laws, with coefficients that can be interpreted as the taxon fitness $F_i$ and the variability $V$ (see Model in Material and Methods).
 In this model, when $V$ is sufficiently low, abundances are stable in time.
-Differences in variability $V$ can induce a noise-induced phase transition in relative abundances of taxa.
-The temporal evolution of the probability of a taxon having abundance $x_i$ given its fitness is governed by the Fokker--Planck equation.
-The results of solving this equation show that the stability is best captured by a phase space determined by fitness $F$ and amplitude of fluctuations $V$ (see Figure \ref{fig:main3}).
+Differences in the variability $V$ can induce a noise-induced phase transition in the relative abundances of taxa.
+The temporal evolution of the probability of a taxon having the abundance $x_i$, given its fitness, is governed by the Fokker--Planck equation.
+The results of solving this equation show that stability is best captured by a phase space determined by the fitness $F$ and the amplitude of fluctuations $V$ (see Figure \ref{fig:main3}).
 
@@
 	\includegraphics[width=0.8\textwidth]{figs/Fig3.eps}
-	\caption{Microbiota states can be placed in the phase space $F$--$V$.
-The light blue shaded region corresponds to the stable phase, while the grey shaded region is the unstable phase (the phase transition line is calculated for  $\alpha$ = $\beta$ = 0.75).
-We place healthy individuals (green) and individuals whose gut microbiota is threatened (antibiotics, IBS) in the phase space fitness--variability.
-Gut microbiota of healthy individuals over a long term span show a quasi--periodical variability (central period is ten days).
-We show that taking antibiotics (AB1 and AB2 correspond to first and second treatment respectively) induces a phase transition in the gut microbiota, which impacts its future changes.
+	\caption{Microbiota states can be placed in the phase space $F-V$.
+The light-blue shaded region corresponds to the stable phase, while the grey shaded region is the unstable phase (the phase transition line is calculated for  $\alpha$ = $\beta$ = 0.75).
+We placed healthy individuals (green) and individuals whose gut microbiota is threatened (antibiotics, IBS) in the phase space fitness--variability.
+The gut microbiota of healthy individuals over a long term span show a quasi--periodical variability (central period is ten days).
+We show that taking antibiotics (AB1 and AB2 correspond to the first and second treatment, respectively) induces a phase transition in gut microbiota, which impacts on future changes.
 We also show an IBS--diagnosed patient transiting from the unstable to the stable phase.}
@@
 
-The model predicts two phases for the gut microbiome: a stable phase with large variability that permits some changes in the relative abundances of taxa; and an unstable phase with larger variability, above the phase transition, where the order of abundant taxa varies significantly with time.
+The model predicted two phases for the gut microbiome: a stable phase with large variability that enabled some changes in the relative abundances of taxa; and an unstable phase with larger variability, above the phase transition, where the order of abundant taxa varies significantly over time.
 The microbiome of all healthy individuals was found to be in the stable phase, while the microbiome of several other individuals was shown to be in the unstable phase.
-In particular, individuals taking antibiotics and IBS--diagnosed patient P2 had the most severe symptoms.
+In particular, individuals taking antibiotics and the IBS--diagnosed patient P2 had the most severe symptoms.
 In this phase diagram, each microbiota state is represented by a point at its measured variability $V$ and inferred fitness $F$.
-The model predicts high average fitness for all taxa, i.e., taxa are narrowly distributed in F.
-The fitness parameter has been chosen with different values for demonstrative purposes.
-Fitness is larger for the healthiest subjects and smaller for the IBS--diagnosed patients.
+The model predicted high average fitness for all taxa, i.e., taxa were narrowly distributed in F.
+The fitness parameter was chosen with different values for demonstrative purposes.
+Fitness was larger for the healthiest subjects and smaller for the IBS--diagnosed patients.
 
@@
 
-The rank dynamics and stability plots in Figure \ref{fig:corrank_IBS_A} and \ref{fig:corrank_IBS_P2} show the variation in the rank with time for the most dominant taxa and their calculated Rank Stability Index (RSI, as discussed in Material and Methods) for the taxa of a healthy subject (individual \emph{A}, Figure \ref{fig:corrank_IBS_A}) and from a subject diagnosed with IBS (patient \emph{P2}, Figure \ref{fig:corrank_IBS_P2}) of the IBS study\cite{IBS}.
-The taxa are listed ordered by the accumulated frequency along the time series, so y-axis is an overall dominance axis for each sample set.
-Generally speaking, we observe that the most dominant taxa are the most rank stable.
+The rank dynamics and stability plots in Figure \ref{fig:corrank_IBS_A} and \ref{fig:corrank_IBS_P2} show the variations in rank over time for the most dominant taxa and their calculated Rank Stability Index (RSI, as discussed in Material and Methods) for the taxa of a healthy subject (individual \emph{A}, Figure \ref{fig:corrank_IBS_A}) and for a subject diagnosed with IBS (patient \emph{P2}, Figure \ref{fig:corrank_IBS_P2}) of the IBS study\cite{IBS}.
+The taxa were listed ordered by their accumulated frequency over the time series, with the y-axis being the overall dominance axis for each sample set.
+Generally speaking, we observed that the most dominant taxa had the highest rank stability.
 
@@
 	\includegraphics[width=0.99\textwidth]{figs/Fig4.eps}
-	\caption{Rank variation throughout time for the most dominant elements (taxa) and their calculated Rank Stability Index (as shown in Material and Methods) for samples from a healthy subject studied in our lab\cite{IBS}.}
+	\caption{Rank variation over time for the most dominant elements (taxa) and their calculated Rank Stability Index (as shown in Material and Methods) for samples from a healthy subject (individual \emph{A}) studied in our lab\cite{IBS}.}
 	\label{fig:corrank_IBS_A}
@@
 	\includegraphics[width=0.99\textwidth]{figs/Fig5.eps}
-	\caption{Rank variation throughout time for the most dominant elements (taxa) and their calculated Rank Stability Index for samples from a subject diagnosed with irritable bowel syndrome studied in our lab\cite{IBS}.}
+	\caption{Rank variation over time for the most dominant elements (taxa) and their calculated Rank Stability Index for samples from a subject diagnosed with irritable bowel syndrome (patient \emph{P2}) studied in our lab\cite{IBS}.}
 	\label{fig:corrank_IBS_P2}
@@
 
-Nevertheless, in the particular case of the healthy individual in Figure \ref{fig:corrank_IBS_A}, \emph{Burkholderiales} and \emph{Betaproteobacteria} (taxa ordered as 18th and 25th in the dominance axis) show comparatively very low rank stability regarding similar dominant taxa while, on the other hand, \emph{Comamonadaceae}, \emph{Lactobacillaceae}, \emph{Fusobacteriaceae}, \emph{Aerococcaceae} and \emph{Carnobacteriaceae} show higher stability than other more dominant taxa, forming a kind of \emph{rank stability island} for medium-ranked taxa around position 40 in the dominance axis, and thus colored in orange, following the color criteria shown in the table included in Figure \ref{fig:corrank_IBS_A}, since they show a moderately stable RSI.
+Nevertheless, in the particular case of the healthy individual in Figure \ref{fig:corrank_IBS_A}, \emph{Burkholderiales} and \emph{Betaproteobacteria} (the 18th and 25th taxa in the dominance axis) comparatively showed very low rank stability regarding similar dominant taxa while, on the other hand, \emph{Comamonadaceae}, \emph{Lactobacillaceae}, \emph{Fusobacteriaceae}, \emph{Aerococcaceae} and \emph{Carnobacteriaceae} showed higher stability than other more dominant taxa, forming a kind of \emph{rank stability island} for medium--ranked taxa around position 40 in the dominance axis, and thus colored brigher, since they show a moderately stable (RSI over 75\%).
 
-In the IBS diagnosed patient of Figure \ref{fig:corrank_IBS_P2}, beyond the differences in dominance for the particular taxa, we still observe that the most dominant are the most rank stable.
-However, as opposed to the healthy individual results, far from presenting a \emph{rank stability island}, the medium-ranked taxa are very rank unstable, mostly due to transient (often one or two consecutive samples) but deep drops in their relative abundance, which are usually happening more than twice along the time series.
-That is, for instance, the case of \emph{Sphingobacteriales} with two non-consecutive samples dropping to 111th rank position.
-In other cases, the high rank instability comes from a rank fluctuation over all the time series, as for \emph{Streptococcaceae} and \emph{Burkholderiales}, which are ranking 26th and 29th respectively in the overall dominance axis but show very low RSI, and thus colored in black attending to the color criteria shown in the table included in Figure \ref{fig:corrank_IBS_P2}.
+In the IBS--diagnosed patient in Figure \ref{fig:corrank_IBS_P2}, beyond the differences in dominance for the particular taxa, we still observed that the most dominant were the most rank stable.
+However, as opposed to the healthy individual results, far from presenting a \emph{rank stability island}, the medium-ranked taxa were very rank unstable, mostly due to transient (often one or two consecutive samples) yet dramatic drops in their relative abundance, which usually occurred more than twice during their time series.
+For example, this was the case of \emph{Sphingobacteriales} with two non-consecutive samples dropping in the ranking to 111th.
+In other cases, the high rank instability came from a rank fluctuation throughout the time series, such as for \emph{Streptococcaceae} and \emph{Burkholderiales}, which ranked 26th and 29th, respectively, in the overall dominance axis but had a very low RSI.
 
-We found the presence of such of \emph{rank stability island} for medium-ranked taxa in the other healthy subjects (\emph{B} and \emph{C}) of the IBS study\cite{IBS} together with its total absence for the other IBS diagnosed patient (patient \emph{P1}), which also presents very high rank instability in its medium-ranked taxa.
-See Supplementary Figure \ref{supfig:corrank_IBS_C} and Supplementary Figure \ref{supfig:corrank_IBS_P1} for further examples of rank plots for a healthy and light IBS diagnosed patients, respectively.
+We also found a \emph{rank stability island} for medium-ranked taxa in the other healthy subjects (\emph{B} and \emph{C}) in the IBS study\cite{IBS} together with its total absence for the other IBS--diagnosed patient (patient \emph{P1}), which also showed very high rank instability in its medium-ranked taxa.
+See Supplementary Figure \ref{supfig:corrank_IBS_C} and Supplementary Figure \ref{supfig:corrank_IBS_P1} for further examples of rank plots for a healthy and mild IBS--diagnosed patients, respectively.
 
@@
 	\includegraphics[width=0.99\textwidth]{figs/supfig_corrank_IBS_C.eps}
-	\caption{Rank variation throughout time for the most dominant elements (taxa) and their calculated RSI for samples from another healthy subject studied in our lab\cite{IBS}.}
+	\caption{Rank variation over time for the most dominant elements (taxa) and their calculated RSI for samples from another healthy subject studied in our lab\cite{IBS}.}
 	\label{supfig:corrank_IBS_C}
@@
 	\includegraphics[width=0.99\textwidth]{figs/supfig_corrank_IBS_P1.eps}
-	\caption{Rank variation throughout time for the most dominant elements (taxa) and their calculated RSI for samples from a subject diagnosed with light IBS and studied in our lab\cite{IBS}.}
+	\caption{Rank variation over time for the most dominant elements (taxa) and their calculated RSI for samples from a subject diagnosed with mild IBS and studied in our lab\cite{IBS}.}
 	\label{supfig:corrank_IBS_P1}
@@
 
-Finally, we have studied the time dependence of the variability $V$ and power law index $\beta$ (see Model under Material and Methods) by using a sliding window approach.
-The total number of time points are divided in subsets of five points, where the next subset is defined by adding the next time sampling and by eliminating the earliest one.
+Finally, we studied the time dependence of the variability $V$ and power law index $\beta$ (see Model in Material and Methods) by using a sliding window approach.
+The total number of time points was divided into subsets of five points, where the following subset was defined by adding the next time sampling and eliminating the earliest one.
 Both parameters were calculated for each subset against the average time lapse.
-Figure \ref{fig:tempevo1} shows the variability  $V$ as a function of time for the largest sampling: two individuals in the Caporaso's study\cite{moving} corresponding to the gut microbiota of a male (upper plot) and a female (lower plot).
-Both samples shows changes in the variability V with quasi--periodic behavior peaked at about 10 days.
-Variability grows more for the gut microbiota of the male and share a minimal value around 0.1 with the gut microbiota of the female.
+Figure \ref{fig:tempevo1} shows the variability $V$ as a function of time for the largest sampling: two individuals in Caporaso's study\cite{moving} corresponding to the gut microbiota of a male (upper plot) and a female (lower plot).
+Both samples showed changes in the variability V with quasi--periodic behavior peaking at about 10 days.
+Variability grew more for the gut microbiota of the male and shared a minimal value of around 0.1 with the gut microbiota of the female.
 
-Figure \ref{fig:tempevo2} shows the time evolution of $V$ for patient P2 of the IBS study\cite{IBS} (upper plot) and patient D in the antibiotics study\cite{antibiotic} (lower plot).
-The variability of the gut microbiota of P2 decreases from above 0.3 to below 0.2, showing a slow tendency to increase the order of the system.
-Antibiotic intake leaks to a quick increase of variability which lasts for a few days to recover ordering.
-The second antibiotic treatment shows some memory (lower increase of variability) with a slower recovery.
+Figure \ref{fig:tempevo2} shows the time evolution of $V$ for patient \emph{P2} in the IBS study\cite{IBS} (upper plot) and patient \emph{D} in the antibiotics study\cite{antibiotic} (lower plot).
+The variability of the gut microbiota of \emph{P2} decreased from over 0.3 to below 0.2, showing a slow tendency to increase the order of the system.
+Antibiotic intake led to a quick increase in variability which lasted for a few days to recover ordering.
+The second antibiotic treatment showed some memory traits (lower increase of variability) with a slower recovery.
 
@@
 	\includegraphics[width=0.99\textwidth]{figs/Fig6.eps}
-\caption{$V$ as a function of time for the two individuals in the Caporaso's study\cite{moving}: samples of gut microbiome of a male (upper plot) and a female (lower plot).}
+\caption{$V$ as a function of time for the two individuals in Caporaso's study\cite{moving}: samples of gut microbiome of a male (upper plot) and a female (lower plot).}
 \label{fig:tempevo1}
@@
 	\includegraphics[width=0.99\textwidth]{figs/Fig7.eps}	
-\caption{$V$ as a function of time for patient P2 of the IBS study\cite{IBS} (upper plot) and patient D in the antibiotics study\cite{antibiotic} (lower plot).
-The blue vertical lines in the lower plot are showing the periods of antibiotic treatment.}
+\caption{$V$ as a function of time for patient \emph{P2} in the IBS study\cite{IBS} (upper plot) and patient \emph{D} in the antibiotics study\cite{antibiotic} (lower plot).
+The blue vertical lines in the lower plot show the periods of antibiotic treatment.}
 \label{fig:tempevo2}

### 2017-01-11 mSystems/rev/mSys_MAIN_rev.tex
New draft approach to rank stability islands, which leads to REVISION draft G
Changes to be committed:
	modified:   ../documentation/MediaInspector_macOSX_mSystems.jnlp
	modified:   figs/Fig4.eps
	modified:   figs/Fig5.eps
	deleted:    figs/_ed_supfig_corrank_IBS_C.pdf
	deleted:    figs/_ed_supfig_corrank_IBS_P1.pdf
	new file:   figs/supfig_corrank_HLS_StoolA_after.eps
	new file:   figs/supfig_corrank_HLS_StoolA_before.eps
	deleted:    figs/supfig_corrank_IBS_C.eps
	deleted:    figs/supfig_corrank_IBS_P1.eps
	modified:   mSys_MAIN_rev.pdf
	modified:   mSys_MAIN_rev.tex
	modified:   mSys_MandM_

@@
 % release codenumber (update it here!)
-\newcommand\releasecode{REVISION draft F}
+\newcommand\releasecode{REVISION draft G}
 
@@
 \bibitem {FD} {\bf Weber, J.} 1956. Fluctuation Dissipation Theorem. Phys. Rev. {\bf 101}:1620-6
-\bibitem {sislands1} {\bf Ng SC, Lam EFC, Lam TTY, Chan Y, Law W, Tse PCH, Kamm M a, Sung JJY, Chan FKL, Wu JCY.} 2013.
+%\bibitem {sislands1} {\bf Ng SC, Lam EFC, Lam TTY, Chan Y, Law W, Tse PCH, Kamm M a, Sung JJY, Chan FKL, Wu JCY.} 2013.
 Effect of probiotic bacteria on the intestinal microbiota in irritable bowel syndrome.
 J Gastroenterol Hepatol {\bf 28}:1624–31.
-\bibitem {sislands2} {\bf Carroll IM, Ringel-Kulka T, Siddle JP, Ringel Y.} 2012.
+%\bibitem {sislands2} {\bf Carroll IM, Ringel-Kulka T, Siddle JP, Ringel Y.} 2012.
 Alterations in composition and diversity of the intestinal microbiota in patients with diarrhea-predominant irritable bowel syndrome.
 Neurogastroenterol Motil {\bf 24}:521-30
-\bibitem {sislands3} {\bf Xu D, Gao J, Gillilland M, Wu X, Song I, Kao JY, Owyang C.} 2014.
+%\bibitem {sislands3} {\bf Xu D, Gao J, Gillilland M, Wu X, Song I, Kao JY, Owyang C.} 2014.
 Rifaximin alters intestinal bacteria and prevents stress-induced gut inflammation and visceral hyperalgesia in rats.
 Gastroenterology {\bf 146}:484–496.e4.
-\bibitem {sislands4} {\bf Rooks MG, Veiga P, Wardwell-scott LH, Tickle T, Segata N, Michaud M, Gallini CA, Beal C, van Hylckama-Vlieg JET, Ballal SA, Morgan XC, Glickman JN, Gevers D, Huttenhower C, Garrett WS, Hylckama-vlieg JET Van, Ballal SA, Morgan XC, Glickman JN, Gevers D, Huttenhower C, Garrett WS.} 2014.
+%\bibitem {sislands4} {\bf Rooks MG, Veiga P, Wardwell-scott LH, Tickle T, Segata N, Michaud M, Gallini CA, Beal C, van Hylckama-Vlieg JET, Ballal SA, Morgan XC, Glickman JN, Gevers D, Huttenhower C, Garrett WS, Hylckama-vlieg JET Van, Ballal SA, Morgan XC, Glickman JN, Gevers D, Huttenhower C, Garrett WS.} 2014.
 Gut microbiome composition and function in experimental colitis during active disease and treatment-induced remission.

### 2017-01-11 mSystems/rev/mSys_MandM_rev.tex
New draft approach to rank stability islands, which leads to REVISION draft G
Changes to be committed:
	modified:   ../documentation/MediaInspector_macOSX_mSystems.jnlp
	modified:   figs/Fig4.eps
	modified:   figs/Fig5.eps
	deleted:    figs/_ed_supfig_corrank_IBS_C.pdf
	deleted:    figs/_ed_supfig_corrank_IBS_P1.pdf
	new file:   figs/supfig_corrank_HLS_StoolA_after.eps
	new file:   figs/supfig_corrank_HLS_StoolA_before.eps
	deleted:    figs/supfig_corrank_IBS_C.eps
	deleted:    figs/supfig_corrank_IBS_P1.eps
	modified:   mSys_MAIN_rev.pdf
	modified:   mSys_MAIN_rev.tex
	modified:   mSys_MandM_

@@
 
-The Rank Stability Index (RSI) is shown as a percentage in a separate bar on the right of the rank matrix plot in Figures \ref{fig:corrank_IBS_A} and \ref{fig:corrank_IBS_P2}.
+The Rank Stability Index (RSI) is shown as a percentage in a separate bar on the right of the rank matrix plot in Figures \ref{fig:corrank_HLS_abroad} and \ref{fig:corrank_HLS_returned} and Supplementary Figures \ref{supfig:corrank_HLS_before} and \ref{supfig:corrank_HLS_after}.
 The RSI is strictly $1$ for an element whose range never changes over time, and is strictly $0$ for an element whose rank oscillates between the extremes over time.

### 2017-01-11 mSystems/rev/mSys_discussion_rev.tex
New draft approach to rank stability islands, which leads to REVISION draft G
Changes to be committed:
	modified:   ../documentation/MediaInspector_macOSX_mSystems.jnlp
	modified:   figs/Fig4.eps
	modified:   figs/Fig5.eps
	deleted:    figs/_ed_supfig_corrank_IBS_C.pdf
	deleted:    figs/_ed_supfig_corrank_IBS_P1.pdf
	new file:   figs/supfig_corrank_HLS_StoolA_after.eps
	new file:   figs/supfig_corrank_HLS_StoolA_before.eps
	deleted:    figs/supfig_corrank_IBS_C.eps
	deleted:    figs/supfig_corrank_IBS_P1.eps
	modified:   mSys_MAIN_rev.pdf
	modified:   mSys_MAIN_rev.tex
	modified:   mSys_MandM_

@@
 
-Specifically, the analysis of the rank stability of the samples of healthy and IBS--diagnosed patients studied in our lab\cite{IBS}, suggests that the presence of \emph{rank stability islands} among medium-ranked taxa is a interesting feature.
-The higher stability of these taxa goes against the global meaning of the scaling index.
-Interestingly, this stability disappears when we look at IBS patients.
-From the literature, it seems that some genera from the families Comamonadaceae, Neisseriaceae and Carnobacteriaceae have been reported to lower their abundance in IBS patients against healthy controls \cite{sislands1}.
-In our case, we see that these families that are present in the rank stability island of the healthy patient decrease their rank stability index or even disappear in the IBS patient.
-However, we also see contradictory results in other families as Lactobacillaceae or Fusobacteriaceae, which seems to increase their abundance in IBS patients \cite{sislands2}, while we observe an increase of the former and a decrease of the latter.
-The presence of members of the Lactobacillae family have been reported to have positive effects against gut inflammation and visceral hyperalgesia \cite{sislands3}, usual symptoms of gastrointestinal disorders.
-It could be happening that a disorder in the stability of this particular group may help to arise the onset of symptoms associated to gastrointestinal disorders.
-The Aerococacceae family is also enriched in rats with IBS symptoms who have been treated with immunomodulators \cite{sislands4}.
-Inside this island of stability, we also have families as Fusobacteriaceae o Hallomonadaceae which include pathogenic genera in them, but that are not present in the IBS patient.
+Specifically, the analysis of the rank stability of different periods of time belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife}, suggests that the presence of \emph{rank stability islands} among medium-ranked taxa is a interesting feature.
+Interestingly, this stability was compromised when the period was not an ordinary one, suggesting that those taxa was even sensitive to changes in the lifestyle.
 It could be brought into question the rule of these taxa as key players in the phase transition of the microbiota, or whether they are more susceptible to perturbations than the most abundant.

### 2017-01-11 mSystems/rev/mSys_results_rev.tex
New draft approach to rank stability islands, which leads to REVISION draft G
Changes to be committed:
	modified:   ../documentation/MediaInspector_macOSX_mSystems.jnlp
	modified:   figs/Fig4.eps
	modified:   figs/Fig5.eps
	deleted:    figs/_ed_supfig_corrank_IBS_C.pdf
	deleted:    figs/_ed_supfig_corrank_IBS_P1.pdf
	new file:   figs/supfig_corrank_HLS_StoolA_after.eps
	new file:   figs/supfig_corrank_HLS_StoolA_before.eps
	deleted:    figs/supfig_corrank_IBS_C.eps
	deleted:    figs/supfig_corrank_IBS_P1.eps
	modified:   mSys_MAIN_rev.pdf
	modified:   mSys_MAIN_rev.tex
	modified:   mSys_MandM_

@@
 
-The rank dynamics and stability plots in Figure \ref{fig:corrank_IBS_A} and \ref{fig:corrank_IBS_P2} show the variations in rank over time for the most dominant taxa and their calculated Rank Stability Index (RSI, as discussed in Material and Methods) for the taxa of a healthy subject (individual \emph{A}, Figure \ref{fig:corrank_IBS_A}) and for a subject diagnosed with IBS (patient \emph{P2}, Figure \ref{fig:corrank_IBS_P2}) of the IBS study\cite{IBS}.
+The rank dynamics and stability plots in Figure \ref{fig:corrank_HLS_abroad} and \ref{fig:corrank_HLS_returned} show the variations in rank over time for the most dominant taxa and their calculated Rank Stability Index (RSI, as discussed in Material and Methods) for the gut microbiome taxa of a healthy subject, the individual \emph{A} in the host lifestyle study \cite{hostlife}.
+The Figure \ref{fig:corrank_HLS_abroad} covers the period when the individual is travelling abroad and the Figure \ref{fig:corrank_HLS_returned} covers the subsequent period.
 The taxa were listed ordered by their accumulated frequency over the time series, with the y-axis being the overall dominance axis for each sample set.
@@
 	%\vspace*{-10mm} % Corrects overbox of the figures
-	%\includegraphics[width=0.99\textwidth]{results/corrank/IBS_h_A_amplicons_family_Rank.pdf}
-	\includegraphics[width=0.99\textwidth]{figs/Fig4.eps}
-	\caption{Rank variation over time for the most dominant elements (taxa) and their calculated Rank Stability Index (as shown in Material and Methods) for samples from a healthy subject (individual \emph{A}) studied in our lab\cite{IBS}.}
-	\label{fig:corrank_IBS_A}
+	\includegraphics[width=1.0\textwidth]{figs/Fig4.eps}
+	\caption{Rank variation over time for the 50 most dominant elements (taxa) and their calculated Rank Stability Index (as shown in Material and Methods) for a special period (days 72 to 122, travelling abroad) belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife}.}
+	\label{fig:corrank_HLS_abroad}
 \end{figure}
@@
 	\centering
-	%\vspace*{-10mm} % Corrects overbox of the figures
-	%\hspace*{-4.5mm}\includegraphics[width=0.99\textwidth]{results/corrank/IBS_P2_Metatranscriptores_Rank.pdf}
-	\includegraphics[width=0.99\textwidth]{figs/Fig5.eps}
-	\caption{Rank variation over time for the most dominant elements (taxa) and their calculated Rank Stability Index for samples from a subject diagnosed with irritable bowel syndrome (patient \emph{P2}) studied in our lab\cite{IBS}.}
-	\label{fig:corrank_IBS_P2}
+	%\vspace*{-10mm} % Corrects overbox of the figures \hspace*{-4.5mm}
+	\includegraphics[width=1.0\textwidth]{figs/Fig5.eps}
+	\caption{Rank variation over time for the 50 most dominant elements (taxa) and their calculated Rank Stability Index (as shown in Material and Methods) for an ordinary period (days 123 to 256, after the trip) belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife}.}
+	\label{fig:corrank_HLS_returned}
 \end{figure}
 
-Nevertheless, in the particular case of the healthy individual in Figure \ref{fig:corrank_IBS_A}, \emph{Burkholderiales} and \emph{Betaproteobacteria} (the 18th and 25th taxa in the dominance axis) comparatively showed very low rank stability regarding similar dominant taxa while, on the other hand, \emph{Comamonadaceae}, \emph{Lactobacillaceae}, \emph{Fusobacteriaceae}, \emph{Aerococcaceae} and \emph{Carnobacteriaceae} showed higher stability than other more dominant taxa, forming a kind of \emph{rank stability island} for medium--ranked taxa around position 40 in the dominance axis, and thus colored brigher, since they show a moderately stable (RSI over 75\%).
+For the trip abroad in Figure \ref{fig:corrank_HLS_abroad}, beyond the differences in dominance for the particular taxa, we still observed that the most dominant were the most rank stable.
+Moreover, the medium-ranked taxa were quite rank unstable, mostly due to transient (often one or two consecutive samples) yet dramatic drops in their relative abundance, which usually occurred more than twice during their time series.
 
-In the IBS--diagnosed patient in Figure \ref{fig:corrank_IBS_P2}, beyond the differences in dominance for the particular taxa, we still observed that the most dominant were the most rank stable.
-However, as opposed to the healthy individual results, far from presenting a \emph{rank stability island}, the medium-ranked taxa were very rank unstable, mostly due to transient (often one or two consecutive samples) yet dramatic drops in their relative abundance, which usually occurred more than twice during their time series.
-For example, this was the case of \emph{Sphingobacteriales} with two non-consecutive samples dropping in the ranking to 111th.
-In other cases, the high rank instability came from a rank fluctuation throughout the time series, such as for \emph{Streptococcaceae} and \emph{Burkholderiales}, which ranked 26th and 29th, respectively, in the overall dominance axis but had a very low RSI.
+Nevertheless, in the particular case of the next period, the one subsequent to the trip, in Figure \ref{fig:corrank_HLS_returned}, some taxa showed higher stability than other more dominant taxa, forming a kind of \emph{rank stability islands} for medium--ranked taxa, since they show a moderately stable index (RSI roughly over 70\%).
+In particular, this is the case for the genera \emph{Actinomyces}, \emph{Leuconostoc}, \emph{Lachnobacterium}, \emph{Eggerthella}, \emph{Clostridium} and	\emph{Collinsella}.
+For those genera, both the overall rank and the RSI were clearly lower during the trip (RSI under 70\%).
+\emph{Actinomyces} and \emph{Lachnobacterium} are even not shown in Figure \ref{fig:corrank_HLS_abroad} because they sank to positions 56 and 77, respectively.
+On the other hand, \emph{Leuconostoc} was the less sensitive to the change of lifestyle.
 
-We also found a \emph{rank stability island} for medium-ranked taxa in the other healthy subjects (\emph{B} and \emph{C}) in the IBS study\cite{IBS} together with its total absence for the other IBS--diagnosed patient (patient \emph{P1}), which also showed very high rank instability in its medium-ranked taxa.
-See Supplementary Figure \ref{supfig:corrank_IBS_C} and Supplementary Figure \ref{supfig:corrank_IBS_P1} for further examples of rank plots for a healthy and mild IBS--diagnosed patients, respectively.
+We also found those \emph{rank stability islands} for medium-ranked taxa in the other periods belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife} (see Supplementary Figure \ref{supfig:corrank_HLS_before} and Supplementary Figure \ref{supfig:corrank_HLS_after} for the corresponding rank plots).
 
@@
 	\centering
-	%\vspace*{-10mm} % Corrects overbox of the figures
-	\includegraphics[width=0.99\textwidth]{figs/supfig_corrank_IBS_C.eps}
-	\caption{Rank variation over time for the most dominant elements (taxa) and their calculated RSI for samples from another healthy subject studied in our lab\cite{IBS}.}
-	\label{supfig:corrank_IBS_C}
+	\includegraphics[width=1.0\textwidth]{figs/supfig_corrank_HLS_StoolA_before.eps}
+	\caption{Rank variation over time for the 50 most dominant elements (taxa) and their calculated Rank Stability Index (as shown in Material and Methods) for an ordinary period (days 0 to 70, before the trip) belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife}.}
+	\label{supfig:corrank_HLS_before}
 \end{supfig}
@@
 	\centering
-	%\vspace*{-10mm} % Corrects overbox of the figures
-	%\hspace*{-4.5mm}
-	\includegraphics[width=0.99\textwidth]{figs/supfig_corrank_IBS_P1.eps}
-	\caption{Rank variation over time for the most dominant elements (taxa) and their calculated RSI for samples from a subject diagnosed with mild IBS and studied in our lab\cite{IBS}.}
-	\label{supfig:corrank_IBS_P1}
+	\includegraphics[width=1.0\textwidth]{figs/supfig_corrank_HLS_StoolA_after.eps}
+	\caption{Rank variation over time for the 50 most dominant elements (taxa) and their calculated Rank Stability Index (as shown in Material and Methods) for an ordinary period (days 257 to 364, further after the trip) belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife}.}
+	\label{supfig:corrank_HLS_after}
 \end{supfig}

### 2017-01-11 mSystems/rev/mSys_MAIN_rev.tex
Improve approach to rank stability islands, which leads to REVISION draft H
Changes to be committed:
	modified:   mSys_MAIN_rev.pdf
	modified:   mSys_MAIN_rev.tex
	modified:   mSys_MandM_rev.tex
	modified:   mSys_discussion_rev.tex
	modified:   mSys_results_rev.tex

@@
 \author[1,2,3,$*$]{Daniel Martínez-Martínez}
+\author[1]{Teresa Rubio}
+\author[1,2]{César Gracia}
 \author[2]{Manuel Peña}
-\author[1,2]{César Gracia}
-\author[1]{Teresa Rubio}
 \author[1,3,4,5]{Amparo Latorre}
@@
 \bibitem {FD} {\bf Weber, J.} 1956. Fluctuation Dissipation Theorem. Phys. Rev. {\bf 101}:1620-6
-%\bibitem {sislands1} {\bf Ng SC, Lam EFC, Lam TTY, Chan Y, Law W, Tse PCH, Kamm M a, Sung JJY, Chan FKL, Wu JCY.} 2013.
-Effect of probiotic bacteria on the intestinal microbiota in irritable bowel syndrome.
-J Gastroenterol Hepatol {\bf 28}:1624–31.
-%\bibitem {sislands2} {\bf Carroll IM, Ringel-Kulka T, Siddle JP, Ringel Y.} 2012.
-Alterations in composition and diversity of the intestinal microbiota in patients with diarrhea-predominant irritable bowel syndrome.
-Neurogastroenterol Motil {\bf 24}:521-30
-%\bibitem {sislands3} {\bf Xu D, Gao J, Gillilland M, Wu X, Song I, Kao JY, Owyang C.} 2014.
-Rifaximin alters intestinal bacteria and prevents stress-induced gut inflammation and visceral hyperalgesia in rats.
-Gastroenterology {\bf 146}:484–496.e4.
-%\bibitem {sislands4} {\bf Rooks MG, Veiga P, Wardwell-scott LH, Tickle T, Segata N, Michaud M, Gallini CA, Beal C, van Hylckama-Vlieg JET, Ballal SA, Morgan XC, Glickman JN, Gevers D, Huttenhower C, Garrett WS, Hylckama-vlieg JET Van, Ballal SA, Morgan XC, Glickman JN, Gevers D, Huttenhower C, Garrett WS.} 2014.
-Gut microbiome composition and function in experimental colitis during active disease and treatment-induced remission.
-ISME J {\bf 8}:1403–17.
+\bibitem {rsi_dysbiosis} {\bf Larsen PE, Dai Y.} 2015. Metabolome of human gut microbiome is predictive of host dysbiosis. Gigascience {\bf 4}:42.
+\bibitem {rsi_actino} {\bf Peters BA, Dominianni C, Shapiro JA, Church TR, Wu J, Miller G, Yuen E, Freiman H, Lustbader I, Salik J, Friedlander C, Hayes RB, Ahn J.} 2016.
+The gut microbiota in conventional and serrated precursors of colorectal cancer.
+Microbiome {\bf 4}:69.
+\bibitem {rsi_egg} {\bf Gardiner BJ, Tai AY, Kotsanas D, Francis MJ, Roberts SA, Ballard SA, Junckerstorff RK, Kormana TM.} 2015.
+Clinical and microbiological characteristics of eggerthella lenta bacteremia.
+J Clin Microbiol {\bf 53}:626–635.
 \bibitem {tikhonov} {\bf  Tikhonov M.} 2016. Community-level cohesion without cooperation. Elife {\bf 5}.
@@
 %%  of the main text PDF.
+%TC:ignore
 \begin{table}
@@
 The healthy and quotidian periods are characterized by $\bar{V} = 0.25 \pm 0.09, \bar{\beta} = 0.777 \pm 0.025$.}
+
+\caption{Rank and Rank Stability Index (RSI, as discussed in Material and Methods) over different periods for the taxa listed as \emph{rank stability islands} regarding the gut microbiome of the individual \emph{A} in the host lifestyle study \cite{hostlife}.}
 \end{table}
+%TC:endignore
 

### 2017-01-11 mSystems/rev/mSys_MandM_rev.tex
Improve approach to rank stability islands, which leads to REVISION draft H
Changes to be committed:
	modified:   mSys_MAIN_rev.pdf
	modified:   mSys_MAIN_rev.tex
	modified:   mSys_MandM_rev.tex
	modified:   mSys_discussion_rev.tex
	modified:   mSys_results_rev.tex

@@
 With this configuration, the observed LMAT sustained sequence classification rate was $20\unclassrate$.
-Finally, it is worth mentioning that a complete set of Python scripts was developed as back-end and front-end of the LMAT pipeline in order to manage the added complexity of time series analysis.
+Finally, it is worth mentioning that a complete set of Python scripts  was developed as back-end and front-end of the LMAT pipeline in order to manage the added complexity of time series analysis (\href{https://github.com/DLSteam/MAUS_scripts}{[https://github.com/DLSteam/MAUS\_scripts]}).
 

### 2017-01-11 mSystems/rev/mSys_discussion_rev.tex
Improve approach to rank stability islands, which leads to REVISION draft H
Changes to be committed:
	modified:   mSys_MAIN_rev.pdf
	modified:   mSys_MAIN_rev.tex
	modified:   mSys_MandM_rev.tex
	modified:   mSys_discussion_rev.tex
	modified:   mSys_results_rev.tex

@@
 Specifically, the analysis of the rank stability of different periods of time belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife}, suggests that the presence of \emph{rank stability islands} among medium-ranked taxa is a interesting feature.
-Interestingly, this stability was compromised when the period was not an ordinary one, suggesting that those taxa was even sensitive to changes in the lifestyle.
-It could be brought into question the rule of these taxa as key players in the phase transition of the microbiota, or whether they are more susceptible to perturbations than the most abundant.
+Interestingly, this stability was compromised when the period was not an ordinary one, suggesting that those taxa were sensitive to changes in the lifestyle.
+Among the genera identified as \emph{rank stability islands}, \emph{Lachnobacterium} and \emph{Clostridium} were catalogued as genera predictive of dysbiosis in the work of Larsen and Dai \cite{rsi_dysbiosis}, which analysed the same dataset \cite{hostlife}.
+Furthermore, very recently, it has been confirmed a clear relationship between \emph{Actinomyces} and conventional adenoma \cite{rsi_actino}, one of the two main precursors of the colorectal cancer.
+Finally,\emph{Eggerthella} is an opportunistic pathogen that is often associated with serious gastrointestinal pathology \cite{rsi_egg}.
+
+It could be brought into question the role of these taxa as key players in the phase transition of the microbiota, or whether they are more susceptible to perturbations than the most abundant.
 The types of interactions that could sustain this particular behavior are not clear, as these non-abundant taxa are not usually included in dynamic studies in order to obtain a community matrix.
-Further experiments and data analysis are needed to clarify whether this was a unique event, or whether it is a widespread feature of stable microbiotas.
+Further experiments and data analysis are needed to clarify whether \emph{rank stability island} is a widespread feature of microbiotas and whether it also appears at lower taxonomic levels.
 

### 2017-01-11 mSystems/rev/mSys_results_rev.tex
Improve approach to rank stability islands, which leads to REVISION draft H
Changes to be committed:
	modified:   mSys_MAIN_rev.pdf
	modified:   mSys_MAIN_rev.tex
	modified:   mSys_MandM_rev.tex
	modified:   mSys_discussion_rev.tex
	modified:   mSys_results_rev.tex

@@
 \emph{Actinomyces} and \emph{Lachnobacterium} are even not shown in Figure \ref{fig:corrank_HLS_abroad} because they sank to positions 56 and 77, respectively.
-On the other hand, \emph{Leuconostoc} was the less sensitive to the change of lifestyle.
+On the contrary, \emph{Leuconostoc} was the least sensitive to the change of lifestyle.
+In addition, it is worth mentioning that \emph{Lachnobacterium} showed anti-correlation over time against the vast majority of the taxa classified in this study.
 
 We also found those \emph{rank stability islands} for medium-ranked taxa in the other periods belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife} (see Supplementary Figure \ref{supfig:corrank_HLS_before} and Supplementary Figure \ref{supfig:corrank_HLS_after} for the corresponding rank plots).
+See Supplementary Table S8 for details about the rank and RSI for the above-mentioned taxa over the different periods considered.
 

### 2017-01-12 mSystems/rev/mSys_suppl_rev.tex
Minor changes which lead to REVISION draft E
Changes to be committed:
	modified:   rev/mSys_MAIN_rev.pdf
	modified:   rev/mSys_MAIN_rev.tex
	modified:   rev/mSys_discussion_rev.tex
	modified:   rev/mSys_suppl_rev.pdf
	modified:   rev/mSys_suppl_rev.tex
English corrections applied, which leads to REVISION draft F
On branch master, changes to be committed:
	new file:   ../documentation/mSys_MAIN_revEnglish.doc
	modified:   mSys_MAIN_rev.pdf
	modified:   mSys_MAIN_rev.tex
	modified:   mSys_MandM_rev.tex
	modified:   mSys_discussion_rev.tex
	modified:   mSys_intro_rev.tex
	modified:   mSys_results

@@
 \documentclass[12pt,oneside,letterpaper]{article}
+%\pagestyle{empty}
 
 % release codenumber (update it here!)
-\newcommand\releasecode{REVISION draft C}
+\newcommand\releasecode{REVISION draft I}
 
@@
 \usepackage[usenames,dvipsnames,table]{xcolor}
+\usepackage{rotating}
 %\usepackage{parskip}
@@
 
+% Disable page numbering
+%\pagenumbering{gobble}
+
 \title{
@@
 \begin{table}
-  \caption{Taylor parameters.
+  \caption{Taylor's parameters.
 Individuals with either animal-based (A) or plant-based (P) diets\cite{diet}.
-Previous to diet,  the population sampled is described by $\bar{V} = 0.09 \pm 0.05, \bar{\beta} = 0.77 \pm 0.04$.}
+Previous to the diet, the population sampled is described by $\bar{V} = 0.09 \pm 0.05, \bar{\beta} = 0.77 \pm 0.04$.}
 
   \begin{center}
-    \begin{tabular}{ccccccc}
+    \begin{tabular}{*{7}{c}}
 	    \hline
@@
 \begin{table} 
-\caption{Taylor parameters for individuals taking antibiotics\cite{antibiotic}.
-Prior to antibiotics intake, the population sampled is described by $\bar{V} = 0.12 \pm 0.05, \bar{\beta} = 0.75 \pm 0.04$.}
+\caption{Taylor's parameters for individuals taking antibiotics\cite{antibiotic}.
+Prior to the antibiotics intake, the population sampled is described by $\bar{V} = 0.12 \pm 0.05, \bar{\beta} = 0.75 \pm 0.04$.}
   \begin{center}
-    \begin{tabular}{ccccccc}
+    \begin{tabular}{*{7}{c}}
 	    \hline
@@
 \begin{table}
-\caption{Taylor parameters for persons diagnosed with irritable bowel syndrome (IBS)\cite{IBS}.
+\caption{Taylor's parameters for persons diagnosed with irritable bowel syndrome (IBS)\cite{IBS}.
 Healthy individuals sampled in this study are characterized by $\bar{V} = 0.135 \pm 0.010, \bar{\beta} = 0.692 \pm 0.024$.}
@@
   \begin{center}
-    \begin{tabular}{ccccccc}
+    \begin{tabular}{*{7}{c}}
 	    \hline
@@
 \begin{table} 
-\caption{Taylor parameters for the healthy subject of the discordant twins\cite{kwashiorkor}.
+\caption{Taylor's parameters for the healthy subject of the discordant twins\cite{kwashiorkor}.
 This table continues in Supplementary Table \ref{tab:DK}.
@@
 \begin{table}
-\caption{Taylor parameters for the kwashiorkor part of the discordant twins\cite{kwashiorkor}.
+\caption{Taylor's parameters for the kwashiorkor part of the discordant twins\cite{kwashiorkor}.
 This is a continuation of Supplementary Table \ref{tab:DH}.
@@
    \begin{center}
-    \begin{tabular}{ccccccc}
+    \begin{tabular}{*{7}{c}}
 	    \hline
@@
 \begin{table}
-\caption{Taylor parameters for individuals with different degrees of overweight and obesity\cite{LEA}.
-Healthy people in this study, whom were not obese, are characterized by $\bar{V} = 0.19 \pm 0.06, \bar{\beta} = 0.806 \pm 0.034$.}
+\caption{Taylor's parameters for individuals with different degrees of overweight and obesity\cite{LEA}.
+Healthy people in this study, who were not obese, are characterized by $\bar{V} = 0.19 \pm 0.06, \bar{\beta} = 0.806 \pm 0.034$.}
   \begin{center}
-    \begin{tabular}{ccccccc}
+    \begin{tabular}{*{7}{c}}
 	    \hline
@@
 \begin{table}
-\caption{Taylor parameters for special intervals concerning gut microbiota in the host lifestyle study\cite{hostlife}.
+\caption{Taylor's parameters for special intervals concerning gut microbiota in the host lifestyle study\cite{hostlife}.
 The healthy and quotidian periods are characterized by $\bar{V} = 0.25 \pm 0.09, \bar{\beta} = 0.777 \pm 0.025$.}
@@
   \end{center}
-  \label{tab:IBS}
+  \label{tab:HLS}
 \end{table}
+
+\clearpage
+
+% Rank and RSI for the HLS study (StoolA samples)
+\begin{sidewaystable}
+\caption{Rank and Rank Stability Index (RSI, as discussed in Material and Methods) over different periods for the taxa listed as \emph{rank stability islands} regarding the gut microbiome of the individual \emph{A} in the host lifestyle study \cite{hostlife}.}
+
+  \begin{center}
+  \small
+    \begin{tabular}{*{21}{c}}
+\cline{2-3}\cline{5-21}
+	\multicolumn{1}{c}{\rule[-4mm]{0mm}{10mm}}\vline&\multicolumn{2}{c}{Period}\vline&
+	\multicolumn{1}{c}{}\vline&\multicolumn{17}{c}{Genera}\vline\\
+\cline{2-3}\cline{5-21}
+\rule[-3mm]{0mm}{9mm}&&&&
+	\multicolumn{2}{c}{\bf\emph{Actinomyces}}&&\multicolumn{2}{c}{\bf\emph{Leuconostoc}}&&
+	\multicolumn{2}{c}{\bf\emph{Lachnobacterium}}&&\multicolumn{2}{c}{\bf\emph{Eggerthella}}&&
+	\multicolumn{2}{c}{\bf\emph{Clostridium}}&&\multicolumn{2}{c}{\bf\emph{Collinsella}}\\
+\cline{5-6}\cline{8-9}\cline{11-12}\cline{14-15}\cline{17-18}\cline{20-21}
+\rule[-3mm]{0mm}{9mm}&name&days&&					rank&RSI&&	rank&RSI&&	rank&RSI&&	rank&RSI&&	rank&RSI&&	rank&RSI\\
+\cline{2-3}\cline{5-6}\cline{8-9}\cline{11-12}\cline{14-15}\cline{17-18}\cline{20-21}
+\rule[-2mm]{0mm}{7mm}&\emph{before}&0 to 70&&		46&72.5&&	44&76.3&&	45&70.2&&	35&73.3&&	28&77.2&&	25&84.2\\
+\rule[-2mm]{0mm}{7mm}&\emph{abroad}&72 to 122&&		56&67.1&&	46&66.2&&	77&53.3&&	48&53.4&&	36&49.9&&	41&63.5\\
+\rule[-2mm]{0mm}{7mm}&\emph{returned}&123 to 256&&	44&79.3&&	41&69.5&&	31&74.2&&	33&77.5&&	34&71.6&&	27&81.0\\
+\rule[-2mm]{0mm}{7mm}&\emph{after}&257 to 364&&		43&79.0&&	39&72.2&&	33&68.4&&	30&78.5&&	34&76.7&&	26&80.4\\
+\cline{2-3}\cline{5-6}\cline{8-9}\cline{11-12}\cline{14-15}\cline{17-18}\cline{20-21}
+\rule[-3mm]{0mm}{8mm}&\multicolumn{2}{c}{Overall}&&	47&76.4&&	43&71.0&&	36&69.2&&	35&74.1&&	34&70.7&&	28&79.5\\
+\cline{2-3}\cline{5-21}
+    \end{tabular}
+  \end{center}
+  \label{tab:HLS_RSI}
+\end{sidewaystable}
 

### 2017-01-12 mSystems/rev/mSys_discussion_rev.tex
Corrections and pre-submission, which leads to REVISION draft I
Changes to be committed:
	modified:   figs/supfig_taxlev1.eps
	modified:   figs/supfig_taxlev2.eps
	modified:   mSys_MAIN_rev.pdf
	modified:   mSys_MAIN_rev.tex
	modified:   mSys_MAIN_rev_marked.docx
	modified:   mSys_MandM_rev.tex
	modified:   mSys_discussion_rev.tex
	modified:   mSys_suppl_rev.pdf
	modified:   mSys_suppl_rev.tex
	new file:   tab/mSys_suppl_tab1.pdf
	new file:   tab/mSys_suppl_tab2.pdf
	new file:   tab/mSys_suppl_tab3.pdf
	new file:   tab/mSys_suppl_tab4.pdf
	new file:   tab/mSys_suppl_tab5.pdf
	new file:   tab/m

@@
 In addition, it was shown that this subject's health status improved during the time the experiment was carried out, implying that his microbiota also recovered the stability it had lost.
-It is interesting to note that in the subject's health from the study made by David \emph{et al.} \cite{hostlife} who suffered a Salmonella infection during the experiment, there was a significant shift in variability and a final recovery from the perturbed state (see \ref{supfig:HLS_xWSummary}).
+It is interesting to note that in the subject's health from the study made by David \emph{et al.} \cite{hostlife} who suffered a Salmonella infection during the experiment, there was a significant shift in variability and a final recovery from the perturbed state (see Supplementary Figure  \ref{supfig:HLS_xWSummary}).
 

### 2017-01-13 mSystems/rev/mSys_suppl_rev.tex
Corrections and pre-submission, which leads to REVISION draft I
Changes to be committed:
	modified:   figs/supfig_taxlev1.eps
	modified:   figs/supfig_taxlev2.eps
	modified:   mSys_MAIN_rev.pdf
	modified:   mSys_MAIN_rev.tex
	modified:   mSys_MAIN_rev_marked.docx
	modified:   mSys_MandM_rev.tex
	modified:   mSys_discussion_rev.tex
	modified:   mSys_suppl_rev.pdf
	modified:   mSys_suppl_rev.tex
	new file:   tab/mSys_suppl_tab1.pdf
	new file:   tab/mSys_suppl_tab2.pdf
	new file:   tab/mSys_suppl_tab3.pdf
	new file:   tab/mSys_suppl_tab4.pdf
	new file:   tab/mSys_suppl_tab5.pdf
	new file:   tab/m

@@
 % watermarks (time consuming!)
-\newwatermark[allpages,color=Mulberry!4,angle=45,scale=7,xpos=-40,ypos=0]{REVISION}
-\newwatermark[allpages,color=Mulberry!12,angle=0,scale=1,xpos=0,ypos=137]{
-\emph{mSystems} \LaTeX\ manuscript \releasecode}
+%\newwatermark[allpages,color=Mulberry!4,angle=45,scale=7,xpos=-40,ypos=0]{REVISION}
+%\newwatermark[allpages,color=Mulberry!12,angle=0,scale=1,xpos=0,ypos=137]{
+%\emph{mSystems} \LaTeX\ manuscript \releasecode}
 
@@
 \renewcommand\thelinenumber{\color{red}\arabic{linenumber}}
-\linenumbers
+%\linenumbers
 
@@
 \captionsetup[table]{name=Supplementary Table S\!\!}
+%\DeclareCaptionFormat{empty}{}  %%%%% COMMENT THIS LINE AND THE NEXT FOR REACTIVATING CAPTIONS %%%%%
+%\captionsetup{format=empty,aboveskip=0pt,belowskip=0pt} %%%%%
 
@@
 % Disable page numbering
-%\pagenumbering{gobble}
+\pagenumbering{gobble}
 
@@
 	\vspace*{2mm}
-	Supplemental material of "Health and disease imprinted in the time variability of the human microbiome"\\
+	Supplementary tables of "Health and disease imprinted in the time variability of the human microbiome"\\
 	\vspace*{6mm}
@@
 	\vspace*{0mm}
-	Supplemental material of "Microbiota, are you sick?"
+	Supplementary tables of "Microbiota, are you sick?"
 	\vspace*{4mm}
@@
 \end{document}
+
+Rank and RSI for the taxa listed as rank stability island in the main document.

### 2017-01-14 mSystems/rev/mSys_MAIN_rev.tex
Corrections and pre-submission, which leads to REVISION draft I
Changes to be committed:
	modified:   figs/supfig_taxlev1.eps
	modified:   figs/supfig_taxlev2.eps
	modified:   mSys_MAIN_rev.pdf
	modified:   mSys_MAIN_rev.tex
	modified:   mSys_MAIN_rev_marked.docx
	modified:   mSys_MandM_rev.tex
	modified:   mSys_discussion_rev.tex
	modified:   mSys_suppl_rev.pdf
	modified:   mSys_suppl_rev.tex
	new file:   tab/mSys_suppl_tab1.pdf
	new file:   tab/mSys_suppl_tab2.pdf
	new file:   tab/mSys_suppl_tab3.pdf
	new file:   tab/mSys_suppl_tab4.pdf
	new file:   tab/mSys_suppl_tab5.pdf
	new file:   tab/m

@@
 % release codenumber (update it here!)
-\newcommand\releasecode{REVISION draft G}
+\newcommand\releasecode{REVISION R1 FINAL}
 
@@
 	listname={List of Supplementary Figures},
-	name=Supplementary Figure,
+	name=Supplementary Figure S\!\!,
 	placement=p
@@
 \bibitem {uni_dynam} {\bf Bashan A, Gibson TE, Friedman J, Carey VJ, Weiss ST, Hohmann EL, Liu Y-Y.} 2016. Universality of human microbial dynamics. Nature {\bf 534}:259–262.
+\bibitem{pretaylor} {\bf Smith H. F.} 1938. An empirical law describing heterogeneity in the yields of agricultural crops. J. Agric. Sci. {\bf 28}:1-23
 \bibitem{taylor} {\bf Taylor, L.R.} 1961. Aggregation, Variance and the mean. Nature {\bf 189}:732-35.
@@
 \begin{table}
-  \caption{Taylor's parameters.
-Individuals with either animal-based (A) or plant-based (P) diets\cite{diet}.
-Previous to the diet, the population sampled is described by $\bar{V} = 0.09 \pm 0.05, \bar{\beta} = 0.77 \pm 0.04$.}
 
-\caption{Taylor's parameters for individuals taking antibiotics\cite{antibiotic}.
-Prior to the antibiotics intake, the population sampled is described by $\bar{V} = 0.12 \pm 0.05, \bar{\beta} = 0.75 \pm 0.04$.}
+	\caption{Taylor's parameters of individuals with either animal-based (A) or plant-based (P) diets\cite{diet}.
+Previous to the diet, the population sampled was described by $\bar{V} = 0.09 \pm 0.05, \bar{\beta} = 0.77 \pm 0.04$.}
 
-\caption{Taylor's parameters for persons diagnosed with irritable bowel syndrome (IBS)\cite{IBS}.
-Healthy individuals sampled in this study are characterized by $\bar{V} = 0.135 \pm 0.010, \bar{\beta} = 0.692 \pm 0.024$.}
+	\caption{Taylor's parameters for individuals taking antibiotics (Ab) in the antibiotics study \cite{antibiotic}, persons diagnosed with irritable bowel syndrome (IBS) in the IBS study \cite{IBS} and for special intervals concerning gut microbiota (HLS) in the host lifestyle study\cite{hostlife}.
+Prior to the antibiotics intake, the population sampled in the antibiotics study \cite{antibiotic} was described by $\bar{V} = 0.12 \pm 0.05, \bar{\beta} = 0.75 \pm 0.04$.
+Healthy individuals sampled in the IBS study \cite{IBS} were characterized by $\bar{V} = 0.135 \pm 0.010, \bar{\beta} = 0.692 \pm 0.024$.
+The healthy and quotidian periods in the host lifestyle study \cite{hostlife} are characterized by $\bar{V} = 0.25 \pm 0.09, \bar{\beta} = 0.777 \pm 0.025$.}
 
-\caption{Taylor's parameters for the healthy subject of the discordant twins\cite{kwashiorkor}.
-This table continues in Supplementary Table S\ref{tab:DK}.
-The population of healthy twins is characterized by $\bar{V} = 0.25 \pm 0.10, \bar{\beta} = 0.863 \pm 0.028$.}\label{tab:DH}
+	\caption{Taylor's parameters for the healthy subject (DH) and kwashiorkor part (DK) of the discordant twins\cite{kwashiorkor}.
+The population of healthy twins is characterized by $\bar{V} = 0.25 \pm 0.10, \bar{\beta} = 0.863 \pm 0.028$.}
 
-\caption{Taylor's parameters for the kwashiorkor part of the discordant twins\cite{kwashiorkor}.
-This is a continuation of Supplementary Table S\ref{tab:DH}.
-The population of healthy twins is characterized by $\bar{V} = 0.25 \pm 0.10, \bar{\beta} = 0.863 \pm 0.028$.}\label{tab:DK}
-
-\caption{Taylor's parameters for individuals with different degrees of excess weight and obesity\cite{LEA}.
-The healthy people in this study, who were not obese, are characterized by $\bar{V} = 0.19 \pm 0.06, \bar{\beta} = 0.806 \pm 0.034$.}
-
-\caption{Taylor's parameters for special intervals concerning gut microbiota in the host lifestyle study\cite{hostlife}.
-The healthy and quotidian periods are characterized by $\bar{V} = 0.25 \pm 0.09, \bar{\beta} = 0.777 \pm 0.025$.}
+	\caption{Taylor's parameters for individuals with different degrees of overweight and obesity\cite{LEA}.
+Healthy people in this study, who were not obese, are characterized by $\bar{V} = 0.19 \pm 0.06, \bar{\beta} = 0.806 \pm 0.034$.}
 
 	\caption{Rank and Rank Stability Index (RSI, as discussed in Material and Methods) over different periods for the taxa listed as \emph{rank stability islands} regarding the gut microbiome of the individual \emph{A} in the host lifestyle study \cite{hostlife}.}
+
 \end{table}

### 2017-01-14 mSystems/rev/mSys_discussion_rev.tex
mSystems00144-16R1: solved excess of supplementary files detected at QC step
Changes to be committed:
	modified:   mSys_MAIN_rev.docx
	modified:   mSys_MAIN_rev.pdf
	modified:   mSys_MAIN_rev.tex
	modified:   mSys_MAIN_rev_marked.docx
	modified:   mSys_discussion_rev.tex
	modified:   mSys_suppl_rev.pdf
	modified:   mSys_suppl_rev.tex
	deleted:    tab/mSys_suppl_tab5.pdf
	deleted:    tab/mSys_suppl_tab6.pdf
	deleted:    tab/mSys_suppl_tab7.pdf
	renamed:    tab/mSys_suppl_tab1.pdf -> tab/mSys_suppl_tabS1.pdf
	renamed:    tab/mSys_suppl_tab2.pdf -> tab/mSys_suppl_tabS2.pdf
	renamed:    tab/mSys_s

@@
 	\caption{Taylor's law parameter space for intervals concerning gut microbiota in the host lifestyle study\cite{hostlife}.
-We observe that subject \emph{B}, who suffered a Salmonella infection during the experiment, had a relevant shift in the parameters from \emph{\_before} to \emph{\_infection} and a final recovery from the perturbed state to \emph{\_afterinfec}, which lies in the parameter area compatible with the healthy and stable intervals (see Supplementary Table S7).
+We observe that subject \emph{B}, who suffered a Salmonella infection during the experiment, had a relevant shift in the parameters from \emph{\_before} to \emph{\_infection} and a final recovery from the perturbed state to \emph{\_afterinfec}, which lies in the parameter area compatible with the healthy and stable intervals (see Supplementary Table S2).
 Subject \emph{A} also had a shift in variability from \emph{\_before} to \emph{\_abroad} and back to \emph{\_returned}, also in the proximity zone of healthy and stable periods.}

### 2017-01-14 mSystems/rev/mSys_suppl_rev.tex
mSystems00144-16R1: solved excess of supplementary files detected at QC step
Changes to be committed:
	modified:   mSys_MAIN_rev.docx
	modified:   mSys_MAIN_rev.pdf
	modified:   mSys_MAIN_rev.tex
	modified:   mSys_MAIN_rev_marked.docx
	modified:   mSys_discussion_rev.tex
	modified:   mSys_suppl_rev.pdf
	modified:   mSys_suppl_rev.tex
	deleted:    tab/mSys_suppl_tab5.pdf
	deleted:    tab/mSys_suppl_tab6.pdf
	deleted:    tab/mSys_suppl_tab7.pdf
	renamed:    tab/mSys_suppl_tab1.pdf -> tab/mSys_suppl_tabS1.pdf
	renamed:    tab/mSys_suppl_tab2.pdf -> tab/mSys_suppl_tabS2.pdf
	renamed:    tab/mSys_s

@@
 \begin{table}
-  \caption{Taylor's parameters.
-Individuals with either animal-based (A) or plant-based (P) diets\cite{diet}.
-Previous to the diet, the population sampled is described by $\bar{V} = 0.09 \pm 0.05, \bar{\beta} = 0.77 \pm 0.04$.}
+  \caption{Taylor's parameters of individuals with either animal-based (A) or plant-based (P) diets\cite{diet}.
+Previous to the diet, the population sampled was described by $\bar{V} = 0.09 \pm 0.05, \bar{\beta} = 0.77 \pm 0.04$.}
 
@@
 \begin{table} 
-\caption{Taylor's parameters for individuals taking antibiotics\cite{antibiotic}.
-Prior to the antibiotics intake, the population sampled is described by $\bar{V} = 0.12 \pm 0.05, \bar{\beta} = 0.75 \pm 0.04$.}
+\caption{Taylor's parameters for individuals taking antibiotics (Ab) in the antibiotics study \cite{antibiotic}, persons diagnosed with irritable bowel syndrome (IBS) in the IBS study \cite{IBS} and for special intervals concerning gut microbiota (HLS) in the host lifestyle study\cite{hostlife}.
+Prior to the antibiotics intake, the population sampled in the antibiotics study \cite{antibiotic} was described by $\bar{V} = 0.12 \pm 0.05, \bar{\beta} = 0.75 \pm 0.04$.
+Healthy individuals sampled in the IBS study \cite{IBS} were characterized by $\bar{V} = 0.135 \pm 0.010, \bar{\beta} = 0.692 \pm 0.024$.
+The healthy and quotidian periods in the host lifestyle study \cite{hostlife} are characterized by $\bar{V} = 0.25 \pm 0.09, \bar{\beta} = 0.777 \pm 0.025$.}
+
   \begin{center}
@@
 	 	\hline
-    \end{tabular}
-  \end{center}
-  \label{tab:antibiotics}
-\end{table}
-
-\clearpage
-
-\begin{table}
-\caption{Taylor's parameters for persons diagnosed with irritable bowel syndrome (IBS)\cite{IBS}.
-Healthy individuals sampled in this study are characterized by $\bar{V} = 0.135 \pm 0.010, \bar{\beta} = 0.692 \pm 0.024$.}
-
-  \begin{center}
-    \begin{tabular}{*{7}{c}}
-	    \hline
-		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
-		\hline
 		IBS (minor)&$0.205 \pm 0.034$&$0.740 \pm 0.029$&$0.917$&&$6.9 \pm 3.3$&$2.0 \pm 1.2$\\
 		IBS (severe)&$0.35 \pm 0.06$&$0.793 \pm 0.025$&$0.934$&&$21 \pm 6$&$4.2 \pm 1.0$\\
+        \hline
+    	\hline
+		HLS (abroad)&$0.51 \pm 0.06$&$0.820 \pm 0.012$&$0.928$&&$2.8 \pm 0.6$&$1.7 \pm 0.5$\\
+		HLS (infection)&$0.49 \pm 0.08$&$0.828 \pm 0.018$&$0.923$&&$2.6 \pm 0.9$&$2.0 \pm 0.7$\\
+		HLS (after infection)&$0.36 \pm 0.05$&$0.776 \pm 0.015$&$0.922$&&$1.1 \pm 0.6$&$-0.0 \pm 0.6$\\
 	    \hline
@@
 \begin{table} 
-\caption{Taylor's parameters for the healthy subject of the discordant twins\cite{kwashiorkor}.
-This table continues in Supplementary Table \ref{tab:DK}.
+\caption{Taylor's parameters for the healthy subject (DH) and kwashiorkor part (DK) of the discordant twins\cite{kwashiorkor}.
 The population of healthy twins is characterized by $\bar{V} = 0.25 \pm 0.10, \bar{\beta} = 0.863 \pm 0.028$.}
@@
 		\hline
-    \end{tabular}
-  \end{center}
-  \label{tab:DH}
-\end{table}
-
-\clearpage
-
-\begin{table}
-\caption{Taylor's parameters for the kwashiorkor part of the discordant twins\cite{kwashiorkor}.
-This is a continuation of Supplementary Table \ref{tab:DH}.
-The population of healthy twins is characterized by $\bar{V} = 0.25 \pm 0.10, \bar{\beta} = 0.863 \pm 0.028$.}
-   \begin{center}
-    \begin{tabular}{*{7}{c}}
-	    \hline
-		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
-		\hline
 		DK&$0.40 \pm 0.07$&$0.859 \pm 0.017$&$0.926$&&$1.5 \pm 0.7$&$-0.1 \pm 0.6$\\
@@
   \end{center}
-    \label{tab:DK}
+    \label{tab:Kwash}
 \end{table}
@@
   \label{tab:LEA}
-\end{table}
-
-\clearpage
-
-\begin{table}
-\caption{Taylor's parameters for special intervals concerning gut microbiota in the host lifestyle study\cite{hostlife}.
-The healthy and quotidian periods are characterized by $\bar{V} = 0.25 \pm 0.09, \bar{\beta} = 0.777 \pm 0.025$.}
-
-  \begin{center}
-    \begin{tabular}{*{7}{c}}
-	    \hline
-		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
-		\hline
-		HLS (abroad)&$0.51 \pm 0.06$&$0.820 \pm 0.012$&$0.928$&&$2.8 \pm 0.6$&$1.7 \pm 0.5$\\
-		HLS (infection)&$0.49 \pm 0.08$&$0.828 \pm 0.018$&$0.923$&&$2.6 \pm 0.9$&$2.0 \pm 0.7$\\
-		HLS (after infection)&$0.36 \pm 0.05$&$0.776 \pm 0.015$&$0.922$&&$1.1 \pm 0.6$&$-0.0 \pm 0.6$\\
-	     \hline
-	     \hline
-    \end{tabular}
-  \end{center}
-  \label{tab:HLS}
 \end{table}

### 2017-01-20 mSystems/rev/mSys_MAIN_rev.tex
mSys: Sync of Word and LaTeX versions after sending REV1
Changes to be committed:
	modified:   mSys_MAIN_rev.docx
	modified:   mSys_MAIN_rev.pdf
	modified:   mSys_MAIN_rev.tex
	modified:   mSys_MAIN_rev_marked.docx
	modified:   mSys_MandM_rev.tex
	modified:   mSys_discussion_rev.tex
	modified:   mSys_intro_rev.tex
	modified:   mSys_results_rev.tex
	modified:   mSys_suppl_rev.pdf
	modified:   mSys_suppl_rev.tex

@@
 % release codenumber (update it here!)
-\newcommand\releasecode{REVISION R1 FINAL}
+\newcommand\releasecode{REV2 DRAFT$\alpha$}
 
@@
 % watermarks (time consuming!)
-%\newwatermark[allpages,color=Mulberry!4,angle=45,scale=7,xpos=-40,ypos=0]{REVISION}
-%\newwatermark[allpages,color=Mulberry!12,angle=0,scale=1,xpos=0,ypos=137]{
+%\newwatermark[allpages,color=BurntOrange!4,angle=45,scale=7,xpos=-40,ypos=0]{REVISION}
+%\newwatermark[allpages,color=BurntOrange!12,angle=0,scale=1,xpos=0,ypos=137]{
 %\emph{mSystems} \LaTeX\ manuscript \releasecode}
@@
 \captionsetup[table]{name=Supplementary Table S\!\!} % See EOF
+%\DeclareCaptionFormat{empty}{}  %%%%% COMMENT THIS LINE AND THE NEXT FOR REACTIVATING CAPTIONS %%%%%
+%\captionsetup{format=empty,aboveskip=0pt,belowskip=0pt} %%%%%
 
@@
 % Floating stuff
+\usepackage{rotating}
 \usepackage[nolists, nomarkers]{endfloat}
@@
 \DeclareDelayedFloat{supfig}{Supplementary Figures}
+\DeclareDelayedFloatFlavor{sidewaysfigure}{figure}
+\DeclareDelayedFloatFlavor{sidewaystable}{table}
 
@@
 Research activity into understanding whether changes in the composition and function of the microbiota are associated with disease is increasing.
-We analyzed 16S rRNA and shotgun metagenomic sequencing (SMS) published data from the gut microbiota of 97 individuals monitored over time.
+We analyzed 16S rRNA and shotgun metagenomic sequencing (SMS) published data from the gut microbiota of 99f individuals monitored over time.
 Temporal fluctuations in the microbial composition revealed significant differences due to factors such us dietary changes, antibiotic intake, age and disease.
@@
 
-%\clearpage
+%===================================== BIBLIOGRAPHY =======================================
+
 \begin{thebibliography}{99}
@@
 \bibitem{economics1} {\bf Mantegna RN, Stanley HE.} 1995. Scaling behaviour in the dynamics of an economic index. Nature {\bf 376}:46–49.
-\bibitem{economics2} {\bf Eisler Z, Kertesz J, Yook SH, Barabasi AL.} 2005. Multiscaling and non-universality in fluctuations of driven complex systems. Europhys Lett {\bf 69}:664–670.
+\bibitem{economics2} {\bf Eisler Z, Kertesz J, Yook SH, Barabási AL.} 2005. Multiscaling and non-universality in fluctuations of driven complex systems. Europhys Lett {\bf 69}:664–670.
 \bibitem {cohen_taylor} {\bf Cohen JE, Xu M, Schuster WSF}.
@@
 %%  of the main text PDF.
+%============================== SUPPLEMENTARY TABLES ================================
+
 %TC:ignore
 \begin{table}
-
 \caption{Taylor's parameters of individuals with either animal-based (A) or plant-based (P) diets\cite{diet}.
@@
 
+  \begin{center}
+    \begin{tabular}{*{7}{c}}
+	    \hline
+		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
+		\hline
+		A&$0.26 \pm 0.05$&$0.826 \pm 0.025$&$0.918$&&$3.1 \pm 0.9$&$1.2 \pm 0.6$\\
+		A&$0.32 \pm 0.06$&$0.857 \pm 0.025$&$0.924$&&$4.4 \pm 1.1$&$2.0 \pm 0.6$\\
+		A&$0.194 \pm 0.033$&$0.813 \pm 0.024$&$0.918$&&$1.9 \pm 0.6$&$0.9 \pm 0.6$\\
+		A&$0.24 \pm 0.04$&$0.824 \pm 0.020$&$0.924$&&$2.7 \pm 0.7$&$1.2 \pm 0.5$\\
+		A&$0.34 \pm 0.06$&$0.855 \pm 0.024$&$0.931$&&$4.7 \pm 1.1$&$1.9 \pm 0.6$\\
+		A&$0.30 \pm 0.05$&$0.847 \pm 0.022$&$0.921$&&$3.9 \pm 1.0$&$1.7 \pm 0.5$\\
+		A&$0.133 \pm 0.021$&$0.784 \pm 0.023$&$0.916$&&$0.7 \pm 0.4$&$0.2 \pm 0.6$\\
+		A&$0.25 \pm 0.04$&$0.831 \pm 0.024$&$0.929$&&$3.0 \pm 0.8$&$1.4 \pm 0.6$\\
+		\hline
+		P&$0.23 \pm 0.05$&$0.804 \pm 0.035$&$0.885$&&$2.6 \pm 0.9$&$0.7 \pm 0.8$\\
+		P&$0.097 \pm 0.018$&$0.705 \pm 0.031$&$0.891$&&$0.03 \pm 0.34$&$-1.6 \pm 0.7$\\
+		P&$0.037 \pm 0.006$&$0.642 \pm 0.025$&$0.881$&&$-1.12 \pm 0.11$&$-3.1 \pm 0.6$\\
+		P&$0.118 \pm 0.019$&$0.723 \pm 0.025$&$0.895$&&$0.4 \pm 0.4$&$-1.2 \pm 0.6$\\
+		P&$0.17 \pm 0.04$&$0.78 \pm 0.04$&$0.842$&&$1.5 \pm 0.7$&$0.1 \pm 0.9$\\
+		P&$0.123 \pm 0.020$&$0.757 \pm 0.026$&$0.914$&&$0.5 \pm 0.4$&$-0.4 \pm 0.6$\\
+		P&$0.19 \pm 0.05$&$0.77 \pm 0.04$&$0.871$&&$1.8 \pm 0.9$&$-0.0 \pm 0.9$\\
+		P&$0.121 \pm 0.020$&$0.736 \pm 0.027$&$0.921$&&$0.5 \pm 0.4$&$-0.9 \pm 0.6$\\
+		P&$0.187 \pm 0.034$&$0.771 \pm 0.030$&$0.908$&&$1.8 \pm 0.7$&$-0.1 \pm 0.7$\\
+		P&$0.097 \pm 0.015$&$0.735 \pm 0.025$&$0.922$&&$0.05 \pm 0.28$&$-0.9 \pm 0.6$\\
+	   	\hline
+    	\hline
+    \end{tabular}
+  \end{center}
+  \label{tab:diet}
+\end{table}
+
+\clearpage
+
+\begin{table} 
 \caption{Taylor's parameters for individuals taking antibiotics (Ab) in the antibiotics study \cite{antibiotic}, persons diagnosed with irritable bowel syndrome (IBS) in the IBS study \cite{IBS} and for special intervals concerning gut microbiota (HLS) in the host lifestyle study\cite{hostlife}.
@@
 
+  \begin{center}
+    \begin{tabular}{*{7}{c}}
+	    \hline
+		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
+		\hline
+		Ab&$0.35 \pm 0.07$&$0.81 \pm 0.04$&$0.925$&&$4.3 \pm 1.4$&$1.3 \pm 0.9$\\
+		Ab&$0.41 \pm 0.09$&$0.82 \pm 0.04$&$0.908$&&$5.6 \pm 1.8$&$1.6 \pm 0.9$\\
+		Ab&$0.23 \pm 0.04$&$0.770 \pm 0.031$&$0.920$&&$2.1 \pm 0.8$&$0.5 \pm 0.7$\\
+		Ab&$0.165 \pm 0.029$&$0.738 \pm 0.031$&$0.928$&&$0.9 \pm 0.6$&$-0.3 \pm 0.7$\\
+		Ab&$0.34 \pm 0.06$&$0.812 \pm 0.032$&$0.936$&&$4.1 \pm 1.2$&$1.5 \pm 0.7$\\
+		Ab&$0.26 \pm 0.05$&$0.798 \pm 0.033$&$0.931$&&$2.8 \pm 0.9$&$1.1 \pm 0.8$\\
+	    \hline
+	 	\hline
+		IBS (minor)&$0.205 \pm 0.034$&$0.740 \pm 0.029$&$0.917$&&$6.9 \pm 3.3$&$2.0 \pm 1.2$\\
+		IBS (severe)&$0.35 \pm 0.06$&$0.793 \pm 0.025$&$0.934$&&$21 \pm 6$&$4.2 \pm 1.0$\\
+        \hline
+    	\hline
+		HLS (abroad)&$0.51 \pm 0.06$&$0.820 \pm 0.012$&$0.928$&&$2.8 \pm 0.6$&$1.7 \pm 0.5$\\
+		HLS (infection)&$0.49 \pm 0.08$&$0.828 \pm 0.018$&$0.923$&&$2.6 \pm 0.9$&$2.0 \pm 0.7$\\
+		HLS (after infection)&$0.36 \pm 0.05$&$0.776 \pm 0.015$&$0.922$&&$1.1 \pm 0.6$&$-0.0 \pm 0.6$\\
+	    \hline
+    	\hline  
+    \end{tabular}
+  \end{center}
+  \label{tab:Ab-IBS-HLS}
+\end{table}
+
+\clearpage
+
+\begin{table} 
 \caption{Taylor's parameters for the healthy subject (DH) and kwashiorkor part (DK) of the discordant twins\cite{kwashiorkor}.
 The population of healthy twins is characterized by $\bar{V} = 0.25 \pm 0.10, \bar{\beta} = 0.863 \pm 0.028$.}
+  \begin{center}
+    \begin{tabular}{*{7}{c}}
+	    \hline
+		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
+		\hline
+		DH&$0.27 \pm 0.04$&$0.835 \pm 0.016$&$0.925$&&$0.2 \pm 0.4$&$-1.0 \pm 0.6$\\
+		DH&$0.36 \pm 0.06$&$0.858 \pm 0.015$&$0.929$&&$1.1 \pm 0.6$&$-0.2 \pm 0.5$\\
+		DH&$0.35 \pm 0.06$&$0.859 \pm 0.014$&$0.926$&&$1.0 \pm 0.5$&$-0.1 \pm 0.5$\\
+		DH&$0.25 \pm 0.04$&$0.829 \pm 0.014$&$0.911$&&$0.0 \pm 0.4$&$-1.2 \pm 0.5$\\
+		DH&$0.30 \pm 0.05$&$0.844 \pm 0.014$&$0.920$&&$0.5 \pm 0.4$&$-0.7 \pm 0.5$\\
+		DH&$0.29 \pm 0.05$&$0.850 \pm 0.016$&$0.915$&&$0.4 \pm 0.5$&$-0.5 \pm 0.5$\\
+		DH&$0.28 \pm 0.05$&$0.848 \pm 0.016$&$0.921$&&$0.3 \pm 0.5$&$-0.5 \pm 0.6$\\
+		DH&$0.35 \pm 0.07$&$0.861 \pm 0.017$&$0.918$&&$0.9 \pm 0.6$&$-0.0 \pm 0.6$\\
+		DH&$0.31 \pm 0.04$&$0.833 \pm 0.012$&$0.916$&&$0.6 \pm 0.4$&$-1.1 \pm 0.4$\\
+		DH&$0.33 \pm 0.05$&$0.843 \pm 0.013$&$0.925$&&$0.8 \pm 0.5$&$-0.7 \pm 0.5$\\
+		DH&$0.31 \pm 0.05$&$0.852 \pm 0.014$&$0.925$&&$0.6 \pm 0.5$&$-0.4 \pm 0.5$\\
+		DH&$0.31 \pm 0.05$&$0.853 \pm 0.015$&$0.930$&&$0.6 \pm 0.5$&$-0.4 \pm 0.5$\\
+		DH&$0.203 \pm 0.033$&$0.815 \pm 0.015$&$0.907$&&$-0.44 \pm 0.32$&$-1.7 \pm 0.5$\\
+		\hline
+		DK&$0.40 \pm 0.07$&$0.859 \pm 0.017$&$0.926$&&$1.5 \pm 0.7$&$-0.1 \pm 0.6$\\
+		DK&$0.44 \pm 0.08$&$0.868 \pm 0.016$&$0.919$&&$1.8 \pm 0.8$&$0.2 \pm 0.6$\\
+		DK&$0.196 \pm 0.031$&$0.819 \pm 0.014$&$0.916$&&$-0.50 \pm 0.30$&$-1.5 \pm 0.5$\\
+		DK&$0.160 \pm 0.026$&$0.798 \pm 0.015$&$0.904$&&$-0.85 \pm 0.25$&$-2.3 \pm 0.5$\\
+		DK&$0.30 \pm 0.05$&$0.845 \pm 0.014$&$0.924$&&$0.5 \pm 0.4$&$-0.6 \pm 0.5$\\
+		DK&$0.23 \pm 0.04$&$0.834 \pm 0.014$&$0.908$&&$-0.1 \pm 0.4$&$-1.0 \pm 0.5$\\
+		DK&$0.27 \pm 0.05$&$0.848 \pm 0.015$&$0.930$&&$0.2 \pm 0.4$&$-0.5 \pm 0.5$\\
+		DK&$0.35 \pm 0.07$&$0.860 \pm 0.019$&$0.916$&&$1.0 \pm 0.7$&$-0.1 \pm 0.7$\\
+		DK&$0.34 \pm 0.05$&$0.835 \pm 0.012$&$0.917$&&$0.9 \pm 0.5$&$-1.0 \pm 0.4$\\
+		DK&$0.25 \pm 0.04$&$0.831 \pm 0.012$&$0.912$&&$0.0 \pm 0.4$&$-1.1 \pm 0.4$\\
+		DK&$0.36 \pm 0.06$&$0.858 \pm 0.013$&$0.918$&&$1.1 \pm 0.5$&$-0.2 \pm 0.5$\\
+		DK&$0.31 \pm 0.06$&$0.851 \pm 0.016$&$0.924$&&$0.6 \pm 0.6$&$-0.4 \pm 0.6$\\
+		DK&$0.149 \pm 0.022$&$0.799 \pm 0.013$&$0.905$&&$-0.96 \pm 0.22$&$-2.2 \pm 0.5$\\
+	    \hline
+    	\hline
+    \end{tabular}
+  \end{center}
+    \label{tab:Kwash}
+\end{table}
 
+\clearpage
+
+\begin{table}
 \caption{Taylor's parameters for individuals with different degrees of overweight and obesity\cite{LEA}.
 Healthy people in this study, who were not obese, are characterized by $\bar{V} = 0.19 \pm 0.06, \bar{\beta} = 0.806 \pm 0.034$.}
+  \begin{center}
+    \begin{tabular}{*{7}{c}}
+	    \hline
+		Metadata&V&$\beta$&$\bar{R}^2$&&V$_{st}$&$\beta_{st}$\\
+		\hline
+		OW&$0.59 \pm 0.12$&$0.894 \pm 0.034$&$0.920$&&$6.6 \pm 2.0$&$2.6 \pm 1.0$\\
+		OW&$0.22 \pm 0.04$&$0.830 \pm 0.030$&$0.904$&&$0.5 \pm 0.6$&$0.7 \pm 0.9$\\
+		\hline
+		OBI&$0.28 \pm 0.04$&$0.855 \pm 0.022$&$0.958$&&$1.5 \pm 0.6$&$1.4 \pm 0.6$\\
+		OBI&$0.33 \pm 0.07$&$0.870 \pm 0.031$&$0.916$&&$2.4 \pm 1.1$&$1.9 \pm 0.9$\\
+		\hline
+		OBII&$0.223 \pm 0.032$&$0.823 \pm 0.023$&$0.938$&&$0.6 \pm 0.5$&$0.5 \pm 0.7$\\
+		OBII&$0.208 \pm 0.029$&$0.844 \pm 0.022$&$0.935$&&$0.4 \pm 0.5$&$1.1 \pm 0.7$\\
+		\hline
+		OBIII&$0.34 \pm 0.05$&$0.855 \pm 0.025$&$0.943$&&$2.5 \pm 0.9$&$1.4 \pm 0.7$\\
+		OBIII&$0.26 \pm 0.04$&$0.845 \pm 0.026$&$0.954$&&$1.1 \pm 0.7$&$1.2 \pm 0.8$\\
+		OBIII&$0.33 \pm 0.06$&$0.870 \pm 0.027$&$0.908$&&$2.4 \pm 1.0$&$1.9 \pm 0.8$\\
+		OBIII&$0.200 \pm 0.026$&$0.843 \pm 0.020$&$0.949$&&$0.2 \pm 0.4$&$1.1 \pm 0.6$\\
+		OBIII&$0.30 \pm 0.05$&$0.846 \pm 0.026$&$0.929$&&$1.9 \pm 0.8$&$1.2 \pm 0.7$\\
+		OBIII&$0.176 \pm 0.029$&$0.826 \pm 0.026$&$0.894$&&$-0.2 \pm 0.5$&$0.6 \pm 0.8$\\
+		OBIII&$0.30 \pm 0.06$&$0.841 \pm 0.031$&$0.896$&&$1.8 \pm 0.9$&$1.0 \pm 0.9$\\
+		OBIII&$0.28 \pm 0.04$&$0.857 \pm 0.025$&$0.941$&&$1.5 \pm 0.7$&$1.5 \pm 0.7$\\
+		OBIII&$0.122 \pm 0.018$&$0.822 \pm 0.024$&$0.930$&&$-1.05 \pm 0.30$&$0.5 \pm 0.7$\\
+		\hline
+		OBIIId&$0.47 \pm 0.08$&$0.872 \pm 0.023$&$0.945$&&$4.7 \pm 1.3$&$1.9 \pm 0.7$\\
+		OBIIId&$0.38 \pm 0.06$&$0.846 \pm 0.023$&$0.951$&&$3.2 \pm 1.0$&$1.2 \pm 0.7$\\
+		OBIIId&$0.36 \pm 0.06$&$0.842 \pm 0.022$&$0.954$&&$2.9 \pm 0.9$&$1.1 \pm 0.6$\\
+	    \hline
+    	\hline
+    \end{tabular}
+  \end{center}
+  \label{tab:LEA}
+\end{table}
 
+\clearpage
+
+% Rank and RSI for the HLS study (StoolA samples)
+\begin{sidewaystable}
 \caption{Rank and Rank Stability Index (RSI, as discussed in Material and Methods) over different periods for the taxa listed as \emph{rank stability islands} regarding the gut microbiome of the individual \emph{A} in the host lifestyle study \cite{hostlife}.}
 
-\end{table}
+  \begin{center}
+  \small
+    \begin{tabular}{*{21}{c}}
+\cline{2-3}\cline{5-21}
+	\multicolumn{1}{c}{\rule[-4mm]{0mm}{10mm}}\vline&\multicolumn{2}{c}{Period}\vline&
+	\multicolumn{1}{c}{}\vline&\multicolumn{17}{c}{Genera}\vline\\
+\cline{2-3}\cline{5-21}
+\rule[-3mm]{0mm}{9mm}&&&&
+	\multicolumn{2}{c}{\bf\emph{Actinomyces}}&&\multicolumn{2}{c}{\bf\emph{Leuconostoc}}&&
+	\multicolumn{2}{c}{\bf\emph{Lachnobacterium}}&&\multicolumn{2}{c}{\bf\emph{Eggerthella}}&&
+	\multicolumn{2}{c}{\bf\emph{Clostridium}}&&\multicolumn{2}{c}{\bf\emph{Collinsella}}\\
+\cline{5-6}\cline{8-9}\cline{11-12}\cline{14-15}\cline{17-18}\cline{20-21}
+\rule[-3mm]{0mm}{9mm}&name&days&&					rank&RSI&&	rank&RSI&&	rank&RSI&&	rank&RSI&&	rank&RSI&&	rank&RSI\\
+\cline{2-3}\cline{5-6}\cline{8-9}\cline{11-12}\cline{14-15}\cline{17-18}\cline{20-21}
+\rule[-2mm]{0mm}{7mm}&\emph{before}&0 to 70&&		46&72.5&&	44&76.3&&	45&70.2&&	35&73.3&&	28&77.2&&	25&84.2\\
+\rule[-2mm]{0mm}{7mm}&\emph{abroad}&72 to 122&&		56&67.1&&	46&66.2&&	77&53.3&&	48&53.4&&	36&49.9&&	41&63.5\\
+\rule[-2mm]{0mm}{7mm}&\emph{returned}&123 to 256&&	44&79.3&&	41&69.5&&	31&74.2&&	33&77.5&&	34&71.6&&	27&81.0\\
+\rule[-2mm]{0mm}{7mm}&\emph{after}&257 to 364&&		43&79.0&&	39&72.2&&	33&68.4&&	30&78.5&&	34&76.7&&	26&80.4\\
+\cline{2-3}\cline{5-6}\cline{8-9}\cline{11-12}\cline{14-15}\cline{17-18}\cline{20-21}
+\rule[-3mm]{0mm}{8mm}&\multicolumn{2}{c}{Overall}&&	47&76.4&&	43&71.0&&	36&69.2&&	35&74.1&&	34&70.7&&	28&79.5\\
+\cline{2-3}\cline{5-21}
+    \end{tabular}
+  \end{center}
+  \label{tab:HLS_RSI}
+\end{sidewaystable}
+
 %TC:endignore

### 2017-01-20 mSystems/rev/mSys_MandM_rev.tex
Corrections and pre-submission, which leads to REVISION draft I
Changes to be committed:
	modified:   figs/supfig_taxlev1.eps
	modified:   figs/supfig_taxlev2.eps
	modified:   mSys_MAIN_rev.pdf
	modified:   mSys_MAIN_rev.tex
	modified:   mSys_MAIN_rev_marked.docx
	modified:   mSys_MandM_rev.tex
	modified:   mSys_discussion_rev.tex
	modified:   mSys_suppl_rev.pdf
	modified:   mSys_suppl_rev.tex
	new file:   tab/mSys_suppl_tab1.pdf
	new file:   tab/mSys_suppl_tab2.pdf
	new file:   tab/mSys_suppl_tab3.pdf
	new file:   tab/mSys_suppl_tab4.pdf
	new file:   tab/mSys_suppl_tab5.pdf
	new file:   tab/m

@@
 \subsection*{Standardization} \label{sec:stan}
-In order to properly show all the studies under common axes, we decided to standardize the Taylor parameters using the group of healthy individuals for each study independently.
+In order to properly show all the studies under common axes, we decided to standardize the Taylor parameters using the group of healthy individuals for every single study independently.
 With this approach, all the studies can be visualized in a shared plot with units of Taylor--parameter standard--deviation on their axes.
@@
 Genus, with the best balance between error assignment and number of taxa, was chosen as our reference taxonomic level.
-We verified that our conclusions were not significantly affected by selecting family or species as the reference taxonomic level (see Supplementary Figure \ref{supfig:taxlev1}).
+We verified that our conclusions were not significantly affected by selecting family or species as the reference taxonomic level (see Supplementary Figure S\ref{supfig:taxlev1}).
 
@@
 We only worked with the discordant twins, and considered those who were not affected by kwashiorkor in each pair of patients as being healthy.
-The study by David \textit{et al.} \cite{diet}was selected for its differential diets.
+The study by David \textit{et al.} \cite{diet} was selected for its differential diets.
 The healthy part was considered to be the first time samples of each individual before the diet, while the rest of the time points were considered as perturbations.
@@
 The work by David \textit{et al.} \cite{hostlife} was selected due to the comprehensive longitudinal data that it provides plus its complete metadata and the interesting events that happened to both subjects (an infection and a trip abroad).
-The healthy part was taken as the previous time points before each event.
+The healthy part was taken from time points before or after each event.
 Finally, we also considered a study from our group carried out by Durban \textit{et al.} \cite{IBS} in which the healthy subjects were considered as those who did not suffer from irritable bowel syndrome, while the patients who had this disease were taken as perturbations.
 
-The metadata for each study is provided in Supplementary Tables S1 to S7.
+The metadata for each study is provided in Supplementary Tables S\ref{tab:diet} to S\ref{tab:LEA}.
 They all used 16S rRNA gene sequencing, except for the study of the discordant kwashiorkor twins\cite{kwashiorkor}, in which both SMS and 16S rRNA data were used.
@@
 The latter dataset tested was the kwashiorkor discordant twins study\cite{kwashiorkor} for both genus and species taxonomic levels.
-The Supplementary Figures \ref{supfig:taxlev1} (overview) and \ref{supfig:taxlev2} (detail) plot the comparison between studies (and so, 16S and SMS) and between adjacent taxonomic levels.
+The Supplementary Figures S\ref{supfig:taxlev1} (overview) and S\ref{supfig:taxlev2} (detail) plot the comparison between studies (and so, 16S and SMS) and between adjacent taxonomic levels.
 
@@
   %\includegraphics[width=0.5\textwidth]{results/taxalevel/sum_sta_SMS.pdf}
-  \includegraphics[width=0.99\textwidth]{figs/supfig_taxlev1.eps}
+  \includegraphics[width=1.0\textwidth]{figs/supfig_taxlev1.eps}
 \caption{Overview of the comparison of different approaches based on adjacent taxonomic levels using plots in the Taylor-parameters space.
@@
   %\includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_spc_SMS.pdf}
-  \includegraphics[width=0.99\textwidth]{figs/supfig_taxlev1.eps}
+  \includegraphics[width=1.0\textwidth]{figs/supfig_taxlev2.eps}
 \caption{Detail of comparison of different approaches based on adjacent taxonomic levels using plots of X-weighted power-law fits (see Material and Methods).
@@
 
-The Rank Stability Index (RSI) is shown as a percentage in a separate bar on the right of the rank matrix plot in Figures \ref{fig:corrank_HLS_abroad} and \ref{fig:corrank_HLS_returned} and Supplementary Figures \ref{supfig:corrank_HLS_before} and \ref{supfig:corrank_HLS_after}.
+The Rank Stability Index (RSI) is shown as a percentage in a separate bar on the right of the rank matrix plot in Figures \ref{fig:corrank_HLS_abroad} and \ref{fig:corrank_HLS_returned} and Supplementary Figures S\ref{supfig:corrank_HLS_before} and S\ref{supfig:corrank_HLS_after}.
 The RSI is strictly $1$ for an element whose range never changes over time, and is strictly $0$ for an element whose rank oscillates between the extremes over time.
@@
 
+Finally, under the rank matrix of the aforementioned Figures there are plots relevant to the variability of the rank over time.
+On the one hand, the RV (Rank Variability) for a sampling point shows the absolute difference between every taxon?s rank and its accumulated abundance rank (the overall rank), averaged for all the taxa shown.
+On the other hand, the DV (Differences Variability) for a sampling point shows the absolute difference between every taxon?s rank at that time and the value that it had at the previous sampling point, averaged for all the taxa shown.
+
 %\normalsize

### 2017-01-20 mSystems/rev/mSys_results_rev.tex
mSys: Sync of Word and LaTeX versions after sending REV1
Changes to be committed:
	modified:   mSys_MAIN_rev.docx
	modified:   mSys_MAIN_rev.pdf
	modified:   mSys_MAIN_rev.tex
	modified:   mSys_MAIN_rev_marked.docx
	modified:   mSys_MandM_rev.tex
	modified:   mSys_discussion_rev.tex
	modified:   mSys_intro_rev.tex
	modified:   mSys_results_rev.tex
	modified:   mSys_suppl_rev.pdf
	modified:   mSys_suppl_rev.tex

@@
 The power law (or scaling) index $\beta$ and the variability $V$ (hereafter Taylor's parameters) appear to be correlated with the stability of the community.
-The variability V was a direct estimator of the amplitude of fluctuations over time, while the scaling index gave us information about the statistical properties of the ecosystem.
-The fact that it was less than the unity (when looking at the standard deviation) tells us that the most abundant taxa were less susceptible to any perturbation than the less abundant elements of the population.
+On the one hand, $\beta$ is a scaling index that gave us information about the statistical properties of the ecosystem.
+If it is $1/2$, the system behaves like a Poisson distribution.
+If $\beta$ is 1, the system behaves as an exponential distribution.
+Generally speaking, metagenomes vary with time with $\beta$ between these two universal classes.
+In our case, the fact that $\beta$ was less than 1 tells us that the most abundant taxa in the microbial community were less susceptible to any perturbation than the other taxa.
+On the other hand, the variability $V$ was a direct estimator of the amplitude of fluctuations over time.
+$V$ represents the maximum variability attainable by a hypothetical dominant genus (with relative abundance close to 1).
+It is an important parameter that characterizes the type of system.
+If $V$ is small the ranking is stable such as, for example, the number of diagnoses of a particular disease recorded in Medicare during a month.
+If $V$ is large, as it seems to be the case of metagenomic samples, the ranking might be unstable, like the number of hourly page views of articles in Wikipedia \cite{ranking,fs}.
 The Taylor parameters were related to the health status of the host, which we consider as constituting the main finding contributed by this article.
@@
 The fit is shown for samples from a healthy subject (top) and from a subject diagnosed with irritable bowel syndrome (bottom), studied in our lab \cite{IBS}.
-Taylor's power law seems to be ubiquitous, spanning to six orders of magnitude.}
+Taylor's power law seems to be ubiquitous, spanning to six orders of magnitude.
+The error bars (\emph{mean-axis}) are the SEM.}
 	\label{fig:main1}
@@
 
-Taylor's parameters describing the temporal variability of the gut microbiome in our sampled individuals are shown in Supplementary Tables S1 to S7.
+Taylor's parameters describing the temporal variability of the gut microbiome in our sampled individuals are shown in Supplementary Tables S\ref{tab:diet} to S\ref{tab:LEA}.
 Our results hint at ubiquitous behavior.
-Firtly, the variability (which corresponds to the maximum amplitude of fluctuations) was large, which suggests the resilient capacity of the microbiota.
+Firstly, the variability (which corresponds to the maximum amplitude of fluctuations) was large, which suggests the resilient capacity of the microbiota.
 Secondly, the scaling index was always smaller than one, which means that more abundant taxa were less volatile than less abundant ones.
@@
 
-In order to jointly visualize and compare the results of individuals from different studies\cite{IBS,moving,antibiotic,LEA,kwashiorkor,diet,hostlife}, their Taylor parameters were standardized, with standardization meaning that each parameter was subtracted by the mean value and divided by the standard deviation of the group of healthy individuals for each study independently (for details of the procedure, please see Standardization subsection in Material and Methods).
+In order to jointly visualize and compare the results of individuals from different studies\cite{IBS,moving,antibiotic,LEA,kwashiorkor,diet,hostlife}, their Taylor parameters were standardized, with standardization meaning that each parameter was subtracted by the mean value and divided by the standard deviation of the group of healthy individuals for every single study independently.
+Due to the different systematics in each study, we defined a healthy region for each one of them, standardized to mean zero and variance one and computed mean and variance of ?unhealthy?
+with this standardization (for details of the procedure, please see Standardization subsection in Material and Methods).
+Therefore, different studies were isolated so that individuals from a given study did not affect the results on the ?unhealthy?
+individuals of the other studies.
+We think this statistical approach was safer, as we avoided to combine data with very different systematic errors.
 The healthy zone and the standardized Taylor parameters for individuals whose gut microbiota was compromised (i.e., they were suffering from IBS, kwashiorkor, altered diet, intake of antibiotics, a Salmonella infection, or had gone on a trip abroad) are shown in Figure \ref{fig:main2}.
@@
 Points with errors place gut microbiome in the Taylor's parameter space, for each individual whose microbiota was compromised.
-It should be noted that the parameters were standardized (standard deviation units) to the healthy group in each study for demonstrative and comparative purposes.}
+It should be noted that the parameters were standardized (standard deviation units) to the healthy group in each study for every single study independently, for demonstrative and comparative purposes.}
 	\label{fig:main2}
@@
 It can be shown to have its origin in mathematical convergence which is similar to the central limit theorem, and thus virtually any statistical model designed to produce a Taylor law converges to a Tweedie distribution\cite{stat}, providing a mechanistic explanation based on the statistical theory of errors\cite{convergence1,convergence2,convergence3}.
-To reveal the generic mechanisms that drive different scenarios in the $\beta-V$ space, we modeled the system by assuming that taxon relative abundance followed a Langevin equation with, on the one hand, a deterministic term that captured the fitness of each taxon and, on the other hand, a randomness term associated with Gaussian random noise\cite{ranking}.
-Both terms were modeled by power laws, with coefficients that can be interpreted as the taxon fitness $F_i$ and the variability $V$ (see Model in Material and Methods).
+To reveal the generic mechanisms that drive different scenarios in the $\beta-V$ space, we modeled the system by assuming that taxon relative abundance followed a Langevin equation with, on the one hand, a deterministic term that captured the fitness of each taxon and, on the other hand, a randomness term associated with Gaussian random noise\cite{ranking}.Both terms were modeled by power laws, with coefficients that can be interpreted as the taxon fitness $F_i$ and the variability $V$ (see Model in Material and Methods).
+Fitness $F_i$ captures the time scale that the system needs to reach equilibrium (the size of variability $V$ may or may not allow to reach it).
+$F_i$ has dimensions of 1/time and roughly corresponds to the half-life of the system when decaying to the stable state.
+In fact, it is exactly the half-life if $\beta$ is one and $V$ is negligible.
 In this model, when $V$ is sufficiently low, abundances are stable in time.
@@
 The microbiome of all healthy individuals was found to be in the stable phase, while the microbiome of several other individuals was shown to be in the unstable phase.
-In particular, individuals taking antibiotics and the IBS--diagnosed patient P2 had the most severe symptoms.
+In particular, individuals taking antibiotics and the IBS--diagnosed patient \emph{P2} had the most severe symptoms.
 In this phase diagram, each microbiota state is represented by a point at its measured variability $V$ and inferred fitness $F$.
-The model predicted high average fitness for all taxa, i.e., taxa were narrowly distributed in F.
+The model predicted high average fitness for all taxa, i.e., taxa were narrowly distributed in $F$.
 The fitness parameter was chosen with different values for demonstrative purposes.
@@
 
-We also found those \emph{rank stability islands} for medium-ranked taxa in the other periods belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife} (see Supplementary Figure \ref{supfig:corrank_HLS_before} and Supplementary Figure \ref{supfig:corrank_HLS_after} for the corresponding rank plots).
-See Supplementary Table S8 for details about the rank and RSI for the above-mentioned taxa over the different periods considered.
+We also found those \emph{rank stability islands} for medium-ranked taxa in the other periods belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife} (see Supplementary Figure S\ref{supfig:corrank_HLS_before} and Supplementary Figure S\ref{supfig:corrank_HLS_after} for the corresponding rank plots).
+See Supplementary Table S\ref{tab:HLS_RSI} for details about the rank and RSI for the above-mentioned taxa over the different periods considered.
 
@@
 Figure \ref{fig:tempevo1} shows the variability $V$ as a function of time for the largest sampling: two individuals in Caporaso's study\cite{moving} corresponding to the gut microbiota of a male (upper plot) and a female (lower plot).
-Both samples showed changes in the variability V with quasi--periodic behavior peaking at about 10 days.
+Both samples showed changes in the variability $V$ with quasi--periodic behavior peaking at about 10 days.
 Variability grew more for the gut microbiota of the male and shared a minimal value of around 0.1 with the gut microbiota of the female.

### 2017-01-20 mSystems/rev2/mSys_discussion_rev2.tex
mSys: Kick off for REV2 (DRAFT rel alpha)
Changes to be committed:
	new file:   ../documentation/DetailedEditorAnswer-R1.pdf
	renamed:    ../rev/Response to Reviewer Comments.docx -> ../rev1/Response to Reviewer Comments.docx
	renamed:    ../rev/figs/Fig1.eps -> ../rev1/figs/Fig1.eps
	renamed:    ../rev/figs/Fig2.eps -> ../rev1/figs/Fig2.eps
	renamed:    ../rev/figs/Fig2/Fig2.pdf -> ../rev1/figs/Fig2/Fig2.pdf
	renamed:    ../rev/figs/Fig2/Fig2.svg -> ../rev1/figs/Fig2/Fig2.svg
	renamed:    ../rev/figs/Fig2/README -> ../rev1/figs/Fig2/README
	renamed:    ../rev/figs/Fig2/ellipse.m -> ../rev1/fi

--- 2017-01-11 mSystems/rev/mSys_discussion_rev.tex
@@
 In addition, it was shown that this subject's health status improved during the time the experiment was carried out, implying that his microbiota also recovered the stability it had lost.
-It is interesting to note that in the subject's health from the study made by David \emph{et al.} \cite{hostlife} who suffered a Salmonella infection during the experiment, there was a significant shift in variability and a final recovery from the perturbed state (see \ref{supfig:HLS_xWSummary}).
+It is interesting to note that in the subject's health from the study made by David \emph{et al.} \cite{hostlife} who suffered a Salmonella infection during the experiment, there was a significant shift in variability and a final recovery from the perturbed state (see Supplementary Figure S\ref{supfig:HLS_xWSummary}).
 
@@
 	\caption{Taylor's law parameter space for intervals concerning gut microbiota in the host lifestyle study\cite{hostlife}.
-We observe that subject \emph{B}, who suffered a Salmonella infection during the experiment, had a relevant shift in the parameters from \emph{\_before} to \emph{\_infection} and a final recovery from the perturbed state to \emph{\_afterinfec}, which lies in the parameter area compatible with the healthy and stable intervals (see Supplementary Table S7).
+We observe that subject \emph{B}, who suffered a Salmonella infection during the experiment, had a relevant shift in the parameters from \emph{\_before} to \emph{\_infection} and a final recovery from the perturbed state to \emph{\_afterinfec}, which lies in the parameter area compatible with the healthy and stable intervals (see Supplementary Table S\ref{tab:Ab-IBS-HLS}).
 Subject \emph{A} also had a shift in variability from \emph{\_before} to \emph{\_abroad} and back to \emph{\_returned}, also in the proximity zone of healthy and stable periods.}
@@
 
-Specifically, the analysis of the rank stability of different periods of time belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife}, suggests that the presence of \emph{rank stability islands} among medium-ranked taxa is a interesting feature.
+Specifically, the analysis of the rank stability of different periods of time belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife}, suggests that the presence of \emph{rank stability islands} among medium-ranked taxa is an interesting feature.
 Interestingly, this stability was compromised when the period was not an ordinary one, suggesting that those taxa were sensitive to changes in the lifestyle.
@@
 The types of interactions that could sustain this particular behavior are not clear, as these non-abundant taxa are not usually included in dynamic studies in order to obtain a community matrix.
-Further experiments and data analysis are needed to clarify whether \emph{rank stability island} is a widespread feature of microbiotas and whether it also appears at lower taxonomic levels.
+Further experiments and data analysis are needed to clarify whether \emph{rank stability islands} are a widespread feature of microbiotas and whether they appear at lower taxonomic levels too.
 

### 2017-01-20 mSystems/rev2/mSys_intro_rev2.tex
mSys: Kick off for REV2 (DRAFT rel alpha)
Changes to be committed:
	new file:   ../documentation/DetailedEditorAnswer-R1.pdf
	renamed:    ../rev/Response to Reviewer Comments.docx -> ../rev1/Response to Reviewer Comments.docx
	renamed:    ../rev/figs/Fig1.eps -> ../rev1/figs/Fig1.eps
	renamed:    ../rev/figs/Fig2.eps -> ../rev1/figs/Fig2.eps
	renamed:    ../rev/figs/Fig2/Fig2.pdf -> ../rev1/figs/Fig2/Fig2.pdf
	renamed:    ../rev/figs/Fig2/Fig2.svg -> ../rev1/figs/Fig2/Fig2.svg
	renamed:    ../rev/figs/Fig2/README -> ../rev1/figs/Fig2/README
	renamed:    ../rev/figs/Fig2/ellipse.m -> ../rev1/fi

--- 2017-01-10 mSystems/rev/mSys_intro_rev.tex
@@
 In particular, Systems Biology has made great efforts to reveal the general laws governing the complex behaviour of microbial communities \cite{sysbio&microb, msys1, metasysbio}, including a proposal suggesting they have universal dynamics \cite{uni_dynam}.
-Microbiota can be approached under the light of ecological theory which includes general principles like Taylor's law \cite{taylor} that relates the spatial or temporal variability of the population with its mean.
+Microbiota can be approached under the light of ecological theory which includes general principles like Taylor's law \cite{pretaylor,taylor} that relates the spatial or temporal variability of the population with its mean.
 This law, also known as fluctuation scale law, is ubiquitous in the natural world and can be found in several systems like random walks \cite{randomwalks}, stock markets \cite{economics1, economics2}, tree \cite{cohen_taylor} and animal populations \cite{taylor, animal1, animal2}, gene expression \cite{genexpress}, and the human genome \cite{genome}.
@@
 This paper presents the imprints of health status (healthy or disease) in the macroscopic properties of microbiota, by studying its temporal variability.
-We analyzed more than 35000 time series of taxa from the gut microbiome of 99 individuals obtained from publicly available high throughput sequencing data about different conditions: diseases, diets, obese status, antibiotic therapy and healthy individuals.
+We analyzed more than 40000 time series of taxa from the gut microbiome of 99 individuals obtained from publicly available high throughput sequencing data about different conditions: diseases, diets, trips, obese status, antibiotic therapy and healthy individuals.
 Having seen that all the cases followed Taylor's law, we used this empirical fact to model how the relative abundances of taxa evolved over time thanks to the Langevin equation, in a similar way as the approach applied recently by Blumm {\it et al.} \cite{ranking}.

### 2017-01-20 mSystems/rev2/mSys_MAIN_rev2.tex
mSys: Kick off for REV2 (DRAFT rel alpha)
Changes to be committed:
	new file:   ../documentation/DetailedEditorAnswer-R1.pdf
	renamed:    ../rev/Response to Reviewer Comments.docx -> ../rev1/Response to Reviewer Comments.docx
	renamed:    ../rev/figs/Fig1.eps -> ../rev1/figs/Fig1.eps
	renamed:    ../rev/figs/Fig2.eps -> ../rev1/figs/Fig2.eps
	renamed:    ../rev/figs/Fig2/Fig2.pdf -> ../rev1/figs/Fig2/Fig2.pdf
	renamed:    ../rev/figs/Fig2/Fig2.svg -> ../rev1/figs/Fig2/Fig2.svg
	renamed:    ../rev/figs/Fig2/README -> ../rev1/figs/Fig2/README
	renamed:    ../rev/figs/Fig2/ellipse.m -> ../rev1/fi

--- 2017-01-20 mSystems/rev/mSys_MAIN_rev.tex
@@
 % watermarks (time consuming!)
-%\newwatermark[allpages,color=BurntOrange!4,angle=45,scale=7,xpos=-40,ypos=0]{REVISION}
-%\newwatermark[allpages,color=BurntOrange!12,angle=0,scale=1,xpos=0,ypos=137]{
-%\emph{mSystems} \LaTeX\ manuscript \releasecode}
+\newwatermark[allpages,color=BurntOrange!4,angle=45,scale=7,xpos=-40,ypos=0]{REVISION}
+\newwatermark[allpages,color=BurntOrange!12,angle=0,scale=1,xpos=0,ypos=137]{
+\emph{mSystems} \LaTeX\ manuscript \releasecode}
 
@@
 \section*{Introduction}
-\input{mSys_intro_rev.tex}
+\input{mSys_intro_rev2.tex}
 %\clearpage
@@
 \section*{Results}
-\input{mSys_results_rev.tex}
+\input{mSys_results_rev2.tex}
 %\clearpage
@@
 \section*{Discussion}
-\input{mSys_discussion_rev.tex}
+\input{mSys_discussion_rev2.tex}
 %\clearpage    
@@
 \section*{Materials and Methods}
-\input{mSys_MandM_rev.tex}   
+\input{mSys_MandM_rev2.tex}   
 %\clearpage
@@
 \bibitem{fs} {\bf Eisler Z, Bartos I, Kertész J.} 2008. Fluctuation scaling in complex systems: Taylor’s law and beyond1. Adv Phys {\bf 57}:89–142.
+\bibitem{medicare} {\bf Hidalgo CA, Blumm N, Barabási AL, Christakis NA.} 2009. A Dynamic Network Approach for the Study of Human Phenotypes. PLoS Comput Biol {5}.
 % -> main studies

### 2017-01-20 mSystems/rev2/mSys_intro_rev2.tex
causality and medicare factoid points fixed in the text

@@
 Having seen that all the cases followed Taylor's law, we used this empirical fact to model how the relative abundances of taxa evolved over time thanks to the Langevin equation, in a similar way as the approach applied recently by Blumm {\it et al.} \cite{ranking}.
-We used this mathematical framework to explore the temporal stability of microbiota under different conditions in order to understand how this affected the healthy status of the subjects.
+We used this mathematical framework to explore the temporal stability of microbiota under different conditions in order to understand how this is related to the healthy status of the subjects.

### 2017-01-26 mSystems/rev2/mSys_results_rev2.tex
causality and medicare factoid points fixed in the text
figure 1 caption changed, V time dependence better explained

--- 2017-01-20 mSystems/rev/mSys_results_rev.tex
@@
 The law seem to be ubiquitous, spanning even to six orders of magnitude in the observed relative abundances.
-As shown in Figure \ref{fig:main1}, the most abundant species were less volatile in relative terms than those which were less abundant.
+As shown in Figure \ref{fig:main1}, where $V$ corresponds to the y-intercept and $\beta$ is the slope of the fit, the most abundant species were less volatile in relative terms than those which were less abundant.
 The fitting to the power law was always robust ($R^{2}$ > 0.88) and did not depend on the microbiome condition.
@@
 It is an important parameter that characterizes the type of system.
-If $V$ is small the ranking is stable such as, for example, the number of diagnoses of a particular disease recorded in Medicare during a month.
+If $V$ is small the ranking is stable, similar to the number of diagnoses of a particular disease recorded in Medicare  during a month \cite{medicare}.
 If $V$ is large, as it seems to be the case of metagenomic samples, the ranking might be unstable, like the number of hourly page views of articles in Wikipedia \cite{ranking,fs}.
@@
 Taylor's power law seems to be ubiquitous, spanning to six orders of magnitude.
+$V$ corresponds to the y-intercept and $\beta$ to the slope of the line.
 The error bars (\emph{mean-axis}) are the SEM.}
@@
 Both parameters were calculated for each subset against the average time lapse.
-Figure \ref{fig:tempevo1} shows the variability $V$ as a function of time for the largest sampling: two individuals in Caporaso's study\cite{moving} corresponding to the gut microbiota of a male (upper plot) and a female (lower plot).
+Figure \ref{fig:tempevo1} shows the variability $V$ as a function of time for the two individuals in Caporaso's study\cite{moving} corresponding to the gut microbiota of a male (upper plot) and a female (lower plot).
 Both samples showed changes in the variability $V$ with quasi--periodic behavior peaking at about 10 days.

### 2017-01-27 mSystems/rev2/figs/Fig2/ellipse.m
mSys: Figure 2 (MATLAB) regenerated with labels improved
Changes to be committed:
	modified:   figs/Fig2.eps
	new file:   figs/Fig2/Fig2.eps
	modified:   figs/Fig2/Fig2.pdf
	modified:   figs/Fig2/Fig2.svg
	modified:   figs/Fig2/ellipse.m
	modified:   mSys_MAIN_rev2.pdf
	modified:   mSys_MAIN_rev2.tex

@@ -127,8 +127,8 @@ h11 = scatter(X(initIdx:endIdx),Y(initIdx:endIdx),sizeSymbol,'^','filled','Marke
 h11b = errorbarxy(X(initIdx:endIdx),Y(initIdx:endIdx),XE(initIdx:endIdx),YE(initIdx:endIdx),{'w.',color,color});
 
 %hTitle  = title ('Diagram of regions for health and sick members.');
-hXLabel = xlabel('Variability (SD units)');
-hYLabel = ylabel('Scaling Index (SD units)');
+hXLabel = xlabel('V  (SD units)');
+hYLabel = ylabel('\beta  (SD units)');
 lbl = {'Zone of Health','Kwashiorkor','Plant Diet','Animal Diet','Ob III','Ob III Diet','Antibiotics','IBS','Lifestyle Trip','Lifestyle Sick'};
 hLegend = legend([h0 h1 h3 h4 h6 h7 h8 h9 h10 h11],lbl,'Location','southeast');
 

### 2017-01-30 mSystems/rev2/mSys_intro_rev2.tex
response to reviewer is more complete intro modified accoding to the rev2

@@
 However, these concepts are still under debate \cite{holo3}.
-We are populated by a myriad of microorganisms that interact with us in several physiological processes such as the metabolism of bile acids \cite{bileacids}, of choline \cite{choline} and key-route metabolites, such as short-chain fatty acids \cite{scfa1, scfa2} which are also involved in immune system maturation \cite{scfa3, scfa4}.
-Human microbiota has been suggested to be closely related to diseases like type 2 diabetes \cite{diabetes2}, cardiovascular disease (CVD) \cite{CVD}, irritable bowel syndrome \cite{IBS}, Crohn's disease \cite{CD}, some affections like obesity \cite{ob1, ob2} and malnutrition \cite{nutr} as well as other multiple diseases \cite{Moya_trends}.
-Current studies reveal that gut microbiota also influences brain function and behaviour and is related to neurological disorders like Alzheimer's disease through the brain-gut-microbiome axis\cite{mind,AD}.
-Recently, even a mystifying and elusive condition which is hard to diagnose like chronic fatigue syndrome, which has often been suggested to be a psychosomatic disease, has been closely related to reduced diversity and altered composition of the gut microbiome\cite{CFS}.
+The human microbiome is intimately linked to our physiology with the metabolism of bile acids \cite{bileacids}, of choline \cite{choline} and key metabolites like short-chain fatty acids \cite{scfa1, scfa2} which are also involved in immune system maturation \cite{scfa3, scfa4}.
+Human microbiota has been suggested to be closely related to diseases like type 2 diabetes \cite{diabetes2}, cardiovascular disease (CVD) \cite{CVD}, irritable bowel syndrome \cite{IBS}, Crohn's disease \cite{CD}, some afflictions like obesity \cite{ob1, ob2} and malnutrition \cite{nutr} as well as many other diseases \cite{Moya_trends}.
+Recent studies have revealed that microbes also influence brain function and behaviour and are related to neurological disorders like Alzheimer's disease through the gut-brain axis\cite{mind,AD}.
+Recently, even a subtle condition like chronic fatigue syndrome, which has often been suggested to be a psychosomatic disease, has been associated with reduced diversity and altered composition of the gut microbiome\cite{CFS}.
 

### 2017-01-30 mSystems/rev2/mSys_results_rev2.tex
minor modifications in the text and the response

@@
 The power law (or scaling) index $\beta$ and the variability $V$ (hereafter Taylor's parameters) appear to be correlated with the stability of the community.
+For this, we assume that Taylor's parameters behave as proxies for the stability.
 On the one hand, $\beta$ is a scaling index that gave us information about the statistical properties of the ecosystem.
@@
 
-The model predicted two phases for the gut microbiome: a stable phase with large variability that enabled some changes in the relative abundances of taxa; and an unstable phase with larger variability, above the phase transition, where the order of abundant taxa varies significantly over time.
-The microbiome of all healthy individuals was found to be in the stable phase, while the microbiome of several other individuals was shown to be in the unstable phase.
+The model predicted two phases for the gut microbiome: a stable phase with large variability that enabled some changes in the relative abundances of taxa; and an unstable phase with even larger variability, above the phase transition, where the order of abundant taxa varies significantly over time.
+The phase transition is continuous (of second order), as is the crossing of the boundary.
+The state variable is the composition.
+Any disturbance modifies the composition of the microbiota, with different compositions encoding different $F$ and $V$.
+Our model can be solved analytically, which allows to a simple understanding of the different regimes, and in particular, to calculate the formula of the transition region.
+If $V$ is sufficiently small compared to $F$, there is a maximum in the likelihood in the physical region (relative compositions larger than zero and smaller than one), i.e., there is a best composition solution of the differential equation, which is the ordered solution.
+On the contrary, if $V$ is sufficiently large compared to $F$ the maximum in the likelihood is outside the physical region, i.e., the best composition solution of the differential equation is at the boundaries (either zero or one) and all physical solutions have comparable likelihoods, which is the noisy phase.
+The microbiome of healthy individuals was found to be in the stable phase, while the microbiome of several other individuals was shown to be in the unstable phase.
 In particular, individuals taking antibiotics and the IBS--diagnosed patient \emph{P2} had the most severe symptoms.

### 2017-01-31 mSystems/rev2/mSys_MandM_rev2.tex
mSys: several figs&text changes which lead to REV2 DRAFT delta
On branch master changes to be committed:
	modified:   Response to Reviewer Comments.docx
	modified:   figs/Fig1.eps
	modified:   figs/Fig4.eps
	modified:   figs/Fig5.eps
	new file:   figs/HostLifeStyle_StoolA_day72to122_abroad_TAILLESS_Rank.pdf
	new file:   figs/HostLifeStyle_h_StoolA_day0to70_before_TAILLESS_Rank.pdf
	new file:   figs/HostLifeStyle_h_StoolA_day123to256_returned_TAILLESS_Rank.pdf
	new file:   figs/HostLifeStyle_h_StoolA_day257to364_after_TAILLESS_Rank.pdf
	modified:   figs/_ed_Fig1.pdf
	modified:   figs/_ed_supfig

--- 2017-01-20 mSystems/rev/mSys_MandM_rev.tex
@@
 %%
-\subsection*{Rank Stability Index}\label{sec:RSI}
+\subsection*{Rank stability and variability}\label{sec:RSI}
 
@@
 Finally, under the rank matrix of the aforementioned Figures there are plots relevant to the variability of the rank over time.
-On the one hand, the RV (Rank Variability) for a sampling point shows the absolute difference between every taxon?s rank and its accumulated abundance rank (the overall rank), averaged for all the taxa shown.
-On the other hand, the DV (Differences Variability) for a sampling point shows the absolute difference between every taxon?s rank at that time and the value that it had at the previous sampling point, averaged for all the taxa shown.
+On the one hand, the RV (Rank Variability) for a sampling point shows the absolute difference between every taxon's rank and its accumulated abundance rank (the overall rank), averaged for all the taxa shown.
+On the other hand, the DV (Differences Variability) for a sampling point shows the absolute difference between every taxon's rank at that time and the value that it had at the previous sampling point, averaged for all the taxa shown.
 

### 2017-01-31 mSystems/rev2/mSys_intro_rev2.tex
mSys: several figs&text changes which lead to REV2 DRAFT delta
On branch master changes to be committed:
	modified:   Response to Reviewer Comments.docx
	modified:   figs/Fig1.eps
	modified:   figs/Fig4.eps
	modified:   figs/Fig5.eps
	new file:   figs/HostLifeStyle_StoolA_day72to122_abroad_TAILLESS_Rank.pdf
	new file:   figs/HostLifeStyle_h_StoolA_day0to70_before_TAILLESS_Rank.pdf
	new file:   figs/HostLifeStyle_h_StoolA_day123to256_returned_TAILLESS_Rank.pdf
	new file:   figs/HostLifeStyle_h_StoolA_day257to364_after_TAILLESS_Rank.pdf
	modified:   figs/_ed_Fig1.pdf
	modified:   figs/_ed_supfig

@@
 Recent studies have revealed that microbes also influence brain function and behaviour and are related to neurological disorders like Alzheimer's disease through the gut-brain axis\cite{mind,AD}.
-Recently, even a subtle condition like chronic fatigue syndrome, which has often been suggested to be a psychosomatic disease, has been associated with reduced diversity and altered composition of the gut microbiome\cite{CFS}.
+Even a difficult to diagnose condition like chronic fatigue syndrome, which in the past had often been suggested to be a psychosomatic disease, has recently been associated with reduced diversity and altered composition of the gut microbiome\cite{CFS}.
 

### 2017-01-31 mSystems/rev2/mSys_results_rev2.tex
mSys: several figs&text changes which lead to REV2 DRAFT delta
On branch master changes to be committed:
	modified:   Response to Reviewer Comments.docx
	modified:   figs/Fig1.eps
	modified:   figs/Fig4.eps
	modified:   figs/Fig5.eps
	new file:   figs/HostLifeStyle_StoolA_day72to122_abroad_TAILLESS_Rank.pdf
	new file:   figs/HostLifeStyle_h_StoolA_day0to70_before_TAILLESS_Rank.pdf
	new file:   figs/HostLifeStyle_h_StoolA_day123to256_returned_TAILLESS_Rank.pdf
	new file:   figs/HostLifeStyle_h_StoolA_day257to364_after_TAILLESS_Rank.pdf
	modified:   figs/_ed_Fig1.pdf
	modified:   figs/_ed_supfig

@@
 	\centering
-	\vspace*{-15mm} % Corrects overbox of the figures
+	%\vspace*{-15mm} % Corrects overbox of the figures
 	%\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_xWboot_LOG.pdf}
 	%\includegraphics[width=0.8\textwidth]{results/fits/IBS_P2_Metatranscriptores_stdVSmean_xWboot_LOG.pdf}
-	\includegraphics[width=0.8\textwidth]{figs/Fig1.eps}
+	\includegraphics[width=0.7\textwidth]{figs/Fig1.eps}
 	\caption{X-weighted power-law fits of the standard deviations versus the mean values for each bacterial genus monitored over time.
@@
 	\includegraphics[width=1.0\textwidth]{figs/Fig4.eps}
-	\caption{Rank variation over time for the 50 most dominant elements (taxa) and their calculated Rank Stability Index (as shown in Material and Methods) for a special period (days 72 to 122, travelling abroad) belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife}.}
+	\caption{Rank variation over time for the 50 most dominant elements (taxa) and their calculated RSI (Rank Stability Index), Rank Variability (RV) and Differences Variability (DV), as detailed in Rank stability and variability in Material and Methods, for a special period (days 72 to 122, travelling abroad) belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife}.}
 	\label{fig:corrank_HLS_abroad}
@@
 	\includegraphics[width=1.0\textwidth]{figs/Fig5.eps}
-	\caption{Rank variation over time for the 50 most dominant elements (taxa) and their calculated Rank Stability Index (as shown in Material and Methods) for an ordinary period (days 123 to 256, after the trip) belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife}.}
+	\caption{Rank variation over time for the 50 most dominant elements (taxa) and their calculated RSI (Rank Stability Index), Rank Variability (RV) and Differences Variability (DV), as detailed in Rank stability and variability in Material and Methods, for an ordinary period (days 123 to 256, after the trip) belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife}.}
 	\label{fig:corrank_HLS_returned}
@@
 	\includegraphics[width=1.0\textwidth]{figs/supfig_corrank_HLS_StoolA_before.eps}
-	\caption{Rank variation over time for the 50 most dominant elements (taxa) and their calculated Rank Stability Index (as shown in Material and Methods) for an ordinary period (days 0 to 70, before the trip) belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife}.}
+	\caption{Rank variation over time for the 50 most dominant elements (taxa) and their calculated RSI (Rank Stability Index), Rank Variability (RV) and Differences Variability (DV), as detailed in Rank stability and variability in Material and Methods, for an ordinary period (days 0 to 70, before the trip) belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife}.}
 	\label{supfig:corrank_HLS_before}
@@
 	\includegraphics[width=1.0\textwidth]{figs/supfig_corrank_HLS_StoolA_after.eps}
-	\caption{Rank variation over time for the 50 most dominant elements (taxa) and their calculated Rank Stability Index (as shown in Material and Methods) for an ordinary period (days 257 to 364, further after the trip) belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife}.}
+	\caption{Rank variation over time for the 50 most dominant elements (taxa) and their calculated RSI (Rank Stability Index), Rank Variability (RV) and Differences Variability (DV), as detailed in Rank stability and variability in Material and Methods, for an ordinary period (days 257 to 364, further after the trip) belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife}.}
 	\label{supfig:corrank_HLS_after}

### 2017-02-21 mSystems/rev2/mSys_MAIN_rev2.tex
figure 1 caption changed, V time dependence better explained
mSys: Figure 2 (MATLAB) regenerated with labels improved
Changes to be committed:
	modified:   figs/Fig2.eps
	new file:   figs/Fig2/Fig2.eps
	modified:   figs/Fig2/Fig2.pdf
	modified:   figs/Fig2/Fig2.svg
	modified:   figs/Fig2/ellipse.m
	modified:   mSys_MAIN_rev2.pdf
	modified:   mSys_MAIN_rev2.tex
mSys: several figs&text changes which lead to REV2 DRAFT delta
On branch master changes to be committed:
	modified:   Response to Reviewer Comments.docx
	modified:   figs/Fig1.eps
	modified:   figs/Fig4.eps
	modified:   figs/Fig5.eps
	ne

@@
-% !TEX TS-program = pdflatex
-% !TEX encoding = UTF-8 Unicode
-
+% !TeX TS-program = pdflatex
+% !TeX encoding = UTF-8 Unicode
 \documentclass[12pt,oneside,letterpaper]{article}
 
-% release codenumber (update it here!)
-\newcommand\releasecode{REV2 DRAFT$\alpha$}
+% release codenumber for mSystems 00144-16R2 (update it here!)
+\newcommand\releasecode{ACCEPTED}
 
@@
 \usepackage[english]{babel} 
-\usepackage[expert, charter]{mathdesign} % Fuentes... disponibles conjuntos {charter|utopia}
+\usepackage[expert,charter]{mathdesign} % Fuentes... disponibles conjuntos {charter|utopia}
 %%\usepackage{calligra}
@@
 % watermarks (time consuming!)
-\newwatermark[allpages,color=BurntOrange!4,angle=45,scale=7,xpos=-40,ypos=0]{REVISION}
-\newwatermark[allpages,color=BurntOrange!12,angle=0,scale=1,xpos=0,ypos=137]{
-\emph{mSystems} \LaTeX\ manuscript \releasecode}
+\newwatermark[allpages,color=Aquamarine!5,angle=45,scale=6,xpos=-30,ypos=0]{ACCEPTED}
+\newwatermark[allpages,color=Aquamarine!14,angle=0,scale=1,xpos=0,ypos=137]{
+\textbf{ASM} \emph{mSystems} \LaTeX\ manuscript \releasecode}
 
@@
 \renewcommand{\textfraction}{0.1}
+
 %% Roman footnote numbering style to not collide with bibliographic cites: 
@@
 \usepackage{amsmath}
-
-%% Roman footnote numbering style to not collide with bibliographic cites: 
-\renewcommand{\thefootnote}{\Roman{footnote}}
 
@@
 {\abstract{Animal microbiota (including human microbiota) plays an important role in keeping the physiological status of the host healthy.
-Research activity into understanding whether changes in the composition and function of the microbiota are associated with disease is increasing.
-We analyzed 16S rRNA and shotgun metagenomic sequencing (SMS) published data from the gut microbiota of 99f individuals monitored over time.
-Temporal fluctuations in the microbial composition revealed significant differences due to factors such us dietary changes, antibiotic intake, age and disease.
-This article shows that a fluctuation scaling law describes the temporal changes in the gut microbiota.
-This law enables the temporal variability of the microbial population to be estimated and quantitatively characterizes the path toward disease via a noise-induced phase transition.
-The estimation of the systemic parameters for follow-up studies may have clinical use and, more generally, may also have applications in other fields where it is important to know whether a given community is stable or not.}}
+Research seeks greater insight into whether changes in the composition and function of the microbiota are associated with disease.
+We analyzed 16S rRNA and shotgun metagenomic sequencing (SMS) published data pertaining to the gut microbiota of 99 subjects monitored over time.
+Temporal fluctuations in the microbial composition revealed significant differences due to factors such as dietary changes, antibiotic intake, age and disease.
+This article shows that a fluctuation scaling law can describe the temporal changes in the gut microbiota.
+This law estimates the temporal variability of the microbial population and quantitatively characterizes the path toward disease via a noise-induced phase transition.
+Estimation of the systemic parameters may be of clinical utility in follow-up studies and have more general applications in fields where it is important to know whether a given community is stable or not.}}
 
-\importance{Human microbiota is closedly linked to the health status of a person.
+\importance{The human microbiota correlates closely with the health status of its host.
 This article analyzes the microbial composition of several subjects under different conditions, over a time span that ranged from days to months.
-Using the Langevin equation as the basis of our mathematical framework in order to evaluate microbial temporal stability, we proved that stable microbiotas can be distinguised from unstable microbiotas.
-This first step will help us to determine how microbiota temporal stability is related to the healthiness of people, and it will enable the development of a more comprehensive framework in order to obtain more in-depth knowledge of this complex system.}
+Using the Langevin equation as the basis of our mathematical framework to evaluate microbial temporal stability, we proved that stable microbiotas can be distinguished from unstable microbiotas.
+This initial step will help us to determine how microbiota temporal stability is related to a subject’s health status, and to develop a more comprehensive framework which will provide greater insight into this complex system.}
 
@@
 The authors declare that there are no competing financial interests in relation to the work described here.
-We thereby express our acknowledgement to Bull/Atos and Micron Technology for providing us with the PCIe SSD card Micron P420m HHHL as a free-of-charge sample for high performance throughput for database testing purposes.
+We thereby express our acknowledgment to Bull/Atos and Micron Technology for providing us with the PCIe SSD card Micron P420m HHHL as a free-of-charge sample for high performance throughput for database testing purposes.
 
@@
 \begin{table}
-\caption{Taylor's parameters of individuals with either animal-based (A) or plant-based (P) diets\cite{diet}.
+\caption{Taylor's parameters of subjects with either animal-based (A) or plant-based (P) diets\cite{diet}.
 Previous to the diet, the population sampled was described by $\bar{V} = 0.09 \pm 0.05, \bar{\beta} = 0.77 \pm 0.04$.}
@@
 \begin{table} 
-\caption{Taylor's parameters for individuals taking antibiotics (Ab) in the antibiotics study \cite{antibiotic}, persons diagnosed with irritable bowel syndrome (IBS) in the IBS study \cite{IBS} and for special intervals concerning gut microbiota (HLS) in the host lifestyle study\cite{hostlife}.
+\caption{Taylor's parameters for subjects taking antibiotics (Ab) in the antibiotics study \cite{antibiotic}, persons diagnosed with irritable bowel syndrome (IBS) in the IBS study \cite{IBS} and for special intervals concerning gut microbiota (HLS) in the host lifestyle study\cite{hostlife}.
 Prior to the antibiotics intake, the population sampled in the antibiotics study \cite{antibiotic} was described by $\bar{V} = 0.12 \pm 0.05, \bar{\beta} = 0.75 \pm 0.04$.
-Healthy individuals sampled in the IBS study \cite{IBS} were characterized by $\bar{V} = 0.135 \pm 0.010, \bar{\beta} = 0.692 \pm 0.024$.
+Healthy subjects sampled in the IBS study \cite{IBS} were characterized by $\bar{V} = 0.135 \pm 0.010, \bar{\beta} = 0.692 \pm 0.024$.
 The healthy and quotidian periods in the host lifestyle study \cite{hostlife} are characterized by $\bar{V} = 0.25 \pm 0.09, \bar{\beta} = 0.777 \pm 0.025$.}
@@
 \begin{table}
-\caption{Taylor's parameters for individuals with different degrees of overweight and obesity\cite{LEA}.
+\caption{Taylor's parameters for subjects with varying degrees of overweight and obesity\cite{LEA}.
 Healthy people in this study, who were not obese, are characterized by $\bar{V} = 0.19 \pm 0.06, \bar{\beta} = 0.806 \pm 0.034$.}
@@
 \begin{sidewaystable}
-\caption{Rank and Rank Stability Index (RSI, as discussed in Material and Methods) over different periods for the taxa listed as \emph{rank stability islands} regarding the gut microbiome of the individual \emph{A} in the host lifestyle study \cite{hostlife}.}
+\caption{Rank and Rank Stability Index (RSI, as discussed in Material and Methods) over different periods for the taxa listed as \emph{rank stability islands} regarding the gut microbiome of the subject \emph{A} in the host lifestyle study \cite{hostlife}.}
 

### 2017-02-21 mSystems/rev2/mSys_MandM_rev2.tex
mSys LaTeX files updated to ACCEPTED version (and beyond...)
"Beyond" means that formulae have been improved, typos corrected
and LaTeX magic comments added.

On branch master, changes committed:
	new file:   rev2/mSys_MAIN_rev2.dvi
	modified:   rev2/mSys_MAIN_rev2.pdf
	modified:   rev2/mSys_MAIN_rev2.tex
	modified:   rev2/mSys_MandM_rev2.tex
	modified:   rev2/mSys_discussion_rev2.tex
	modified:   rev2/mSys_intro_rev2.tex
	modified:   rev2/mSys_results_rev2.tex

@@
-%%
+% !TeX root = ./mSys_MAIN_rev2.tex
 %TC:ignore
@@
 \end{linenomath}
-where $F_i$ captured the fitness of the taxon $i$, $V$ corresponded to the noise amplitude and $\xi_i$(t) was a Gaussian random noise with zero mean  $<\xi_i(t)> = 0$, and variance which was uncorrelated over time, $<\xi_i(t) \xi_i(t')> = \delta(t'-t)$.
+where $F_i$ captured the fitness of the taxon $i$, $V$ corresponded to the noise amplitude and $\xi_i$(t) was a Gaussian random noise with zero mean  $\left<\xi_i(t)\right> = 0$, and variance which was uncorrelated over time, $\left<\xi_i(t) \xi_i(t')\right> = \delta(t'-t)$.
 The function $\phi(t)$ ensured the normalization at all times, $\sum x_i(t) = 1$, and corresponded to $\phi(t) = \sum F_i x_i^\alpha + \sum V x_i^\beta \xi_i(t)$.
-The temporal evolution of the probability that taxon $i$ had a relative abundance $x_i(t)$, $P(x_i,t)$, was determined by the Fokker-Planck equation:
+The time--course changes in the probability that taxon $i$ had a relative abundance $x_i(t)$, $P(x_i,t)$, was determined by the Fokker--Planck equation:
 \begin{linenomath}
 \begin{equation*}
-\frac{\partial P}{\partial t} = - \frac{\partial}{\partial x_i}  [(F_i \cdot x_i^\alpha - \phi(t) \cdot x_i ) \cdot P]+ \frac{1}{2} \frac{\partial^2}{\partial x_i^2} (V^2 \cdot x_i^{2\beta}\cdot P).
+\frac{\partial P}{\partial t} = - \frac{\partial}{\partial x_i}  \left[(F_i \; x_i^\alpha - \phi(t) \; x_i ) \; P\right]+ \frac{1}{2} \frac{\partial^2}{\partial x_i^2} \left(V^2 \; x_i^{2\beta}\; P\right).
 \end{equation*}
 \end{linenomath}
-The microbiota evolved towards a steady state with a time-independent probability dependent on the values of $\alpha$, $\beta$, $F_i$ and $V$.
-For $\alpha<1$ (otherwise, systems are always unstable), the steady state probability was localized in a region around a preferred value or broadly distributed over a wide range, depending on whether the fitness $F_i$ dominated or was overwhelmed by the noise amplitude $V$.
-The steady-state solution of the Fokker--Planck equation was given by:
+The microbiota evolved towards a steady state with a time--independent probability dependent on the values of $\alpha$, $\beta$, $F_i$ and $V$.
+For $\alpha<1$ (otherwise, systems are always unstable), the steady--state probability was localized in a region around a preferred value or broadly distributed over a wide range, depending on whether the fitness $F_i$ dominated or was overwhelmed by the noise amplitude $V$.
+The steady--state solution of the Fokker--Planck equation was given by:
 \begin{eqnarray*}
-P_0 (x_i) &=& C_{ne}(\alpha,\beta,F_i,V)  \cdot x_i^{-2\beta}  \cdot \exp\Big[\frac{2F_i}{V^2}\frac{x_i^{1+\alpha-2\beta}}{1+\alpha-2\beta}-\frac{\phi_0}{V^2}\frac{x_i^{2-2\beta}}{1-\beta}\Big] \quad \textrm{if} \quad  2\beta \ne 1+\alpha, \\
-P_0 (x_i) &=& C_e(\alpha,\beta,F_i,V)  \cdot x_i^{\frac{2F_i}{V^2} -2\beta}  \cdot \exp\Big[\frac{\phi_0}{V^2}\frac{x_i^{2-2\beta}}{1-\beta}\Big] \quad \textrm{if} \quad  2\beta = 1+\alpha,
+P_0 (x_i) &=& C_{ne}(\alpha,\beta,F_i,V)  \; x_i^{-2\beta}  \; \exp\left[\frac{2F_i}{V^2}\frac{x_i^{1+\alpha-2\beta}}{1+\alpha-2\beta}-\frac{\phi_0}{V^2}\frac{x_i^{2-2\beta}}{1-\beta}\right] \quad \textrm{if} \quad  2\beta \ne 1+\alpha, \\
+P_0 (x_i) &=& C_e(\alpha,\beta,F_i,V)  \; x_i^{\frac{2F_i}{V^2} -2\beta}  \; \exp\left[\frac{\phi_0}{V^2}\frac{x_i^{2-2\beta}}{1-\beta}\right] \quad \textrm{if} \quad  2\beta = 1+\alpha,
 \end{eqnarray*}
-where $\phi_0 = (\sum_i F_i^{1/(1-\alpha)})^{1-\alpha}$ and $C_{ne}$ and $C_{e}$ were integrals that were solved numerically for the parameters of interest.
-The ordered phase occurred when the solution had a maximum in the physical interval ($0<x_i<1$).
+where $\phi_0 = \left(\sum_i F_i^{\frac{1}{1-\alpha}}\right)^{1-\alpha}$ and $C_{ne}$ together with $C_{e}$ were integrals solved numerically for the parameters of interest.
+The ordered phase occurred when the solution had a maximum in the physical interval $(0<x_i<1)$.
 For a larger V, the transition to a disordered phase happened when the maximum shifted to the unphysical region $x_i<0$, which sets the phase transition region $V(\alpha,\beta,F_i)$.
@@
 In many physical systems (Brownian motion is the classic example \cite{Einstein}), the two terms of the Langevin's equation are related.
-The \emph{Fluctuation Dissipation Theorem} states out a general relationship between the response to an external disturbance and the internal fluctuations of the system\cite{FD}.
-The theorem can be used as the basic formula to derive the fitness from the analysis of fluctuations
-of the microbiota, assuming that it is in equilibrium (the ordered phase).
+The \emph{Fluctuation Dissipation Theorem} states a general relationship between the response to an external disturbance and the internal fluctuations of the system\cite{FD}.
+The theorem can be used as the basic formula to derive the fitness from the analysis of fluctuations in the microbiota, assuming that it is in equilibrium (the ordered phase).
 
@@
 \subsection*{Standardization} \label{sec:stan}
-In order to properly show all the studies under common axes, we decided to standardize the Taylor parameters using the group of healthy individuals for every single study independently.
+In order to properly show all the studies under common axes, we decided to standardize the Taylor parameters using the group of healthy subjects for every single study independently.
 With this approach, all the studies can be visualized in a shared plot with units of Taylor--parameter standard--deviation on their axes.
 
-For a Taylor's parameter, e.g. $V$, the estimate of the mean ($\widehat{V}$) for the healthy subpopulation, composed of $h$ individuals, is:
+For a Taylor's parameter, e.g. $V$, the estimate of the mean ($\widehat{V}$) for the healthy subpopulation, composed of $h$ subjects, is:
 \begin{linenomath}
@@
 \end{linenomath}
-$\sigma_{V_i}$ being the estimation of the uncertainty in $V_i$ obtained together with $V_i$ from the X-weighted power-law fit described in Section \ref{sec:X-w}, for healthy individuals.
+$\sigma_{V_i}$ being the estimation of the uncertainty in $V_i$ obtained together with $V_i$ from the X-weighted power-law fit described in Section \ref{sec:X-w}, for healthy subjects.
 
@@
 \subsubsection*{Sample selection}
-We chose studies about relevant pathologies containing metagenomic sequencing time data series of bacterial populations from humans in different healthy and non-healthy states.
-Only those individuals who had three or more time points of data available in databases were selected.
-The study by Caporaso \textit{et al.} study \cite{moving} was selected as it featured two healthy individuals measured over a very long timespan, with almost daily sampling.
+We chose studies about relevant pathologies, containing metagenomic sequencing time data series of bacterial populations from humans in various healthy and non-healthy states.
+Only those subjects who had three or more time points of data available in databases were selected.
+The study by Caporaso \textit{et al.} study \cite{moving} was selected as it featured two healthy subjects measured over a very long time-span, with almost daily sampling.
 The study of Faith \textit{et al.} \cite{LEA} was selected given the BMI differences between subjects.
 Moreover, some of them followed diets which could be treated as system perturbations.
-Only those individuals who had normal or overweight BMI were considered as healthy.
+Only those subjects whose BMI was normal or overweight were considered healthy.
 The study by Smith \textit{et al.} \cite{kwashiorkor} was selected for both the age of the patients and the rare disease.
-We only worked with the discordant twins, and considered those who were not affected by kwashiorkor in each pair of patients as being healthy.
+Regarding kwashiorkor, we considered only the discordant twins, and deemed subjects unaffected by kwashiorkor as being healthy in each pair of patients.
 The study by David \textit{et al.} \cite{diet} was selected for its differential diets.
-The healthy part was considered to be the first time samples of each individual before the diet, while the rest of the time points were considered as perturbations.
-Dethlefsen and Relman's work \cite{antibiotic} was selected due to the interesting treatment of two antibiotic intakes by three different subjects of the same antibiotic.
-The healthy part was considered to be only those times before any antibiotic treatment, and the time of antibiotic intakes and the period after that as perturbations.
-The work by David \textit{et al.} \cite{hostlife} was selected due to the comprehensive longitudinal data that it provides plus its complete metadata and the interesting events that happened to both subjects (an infection and a trip abroad).
-The healthy part was taken from time points before or after each event.
+The healthy period  was considered to be the initial samples of each subject before starting the diet, while the remaining time points were considered as perturbations.
+Dethlefsen and Relman’s work \cite{antibiotic} was selected due to the interesting treatment of two antibiotic intakes of the same antibiotic by three different subjects.
+The healthy period was considered to correspond only to those times before any antibiotic treatment, whereas the periods during and after antibiotic intake were considered as perturbations.
+The work by David \textit{et al.} \cite{hostlife} was selected due to the comprehensive longitudinal data that it provides, plus its complete metadata and the interesting events experienced by both subjects (an infection and a trip abroad).
+The healthy period was taken from time points before or after each event.
 Finally, we also considered a study from our group carried out by Durban \textit{et al.} \cite{IBS} in which the healthy subjects were considered as those who did not suffer from irritable bowel syndrome, while the patients who had this disease were taken as perturbations.
@@
 \subsubsection*{16rRNA sequences processing}
-Reads from the selected studies were first quality filtered using the FastX toolkit\cite{FASTX}, allowing only those reads which had more than a quality score of over 25 in 75\% of the complete sequence.
+Reads from the selected studies were first quality filtered using the FastX toolkit\cite{FASTX}, allowing only those reads scoring over 25 for quality in  75\% of the complete sequence.
 16S rRNA reads were then clustered at a 97\% nucleotide sequence identity (97\% ID) into operational taxonomic units (OTUs), using the QIIME software package\cite{QIIME} (version 1.8).
@@
 Shotgun metagenomic sequences were analyzed with LMAT (Livermore Metagenomics Analysis Toolkit) software package\cite{LMAT} (version 1.2.4, with Feb'15 release of the LMAT-Grand database).
-LMAT was run using a Bull shared-memory node belonging to the team?s HPC (high performance computing) cluster.
+LMAT was run using a Bull shared-memory node belonging to the team's HPC (high performance computing) cluster.
 It was equipped with 32 cores (64 threads available using Intel Hyper-Threading Technology) as it has two Haswell-based Xeons (22 nm technology), the E5-2698v3@2.3 GHz, sharing half a tebibyte of DRAM memory.
@@
 Here, the uncertainties affected the independent variable, so the fit was not as trivial as a Y-weighted fit, where the uncertainties affect the dependent variable.
-A standard approach for this fit is: a) to invert the variables before applying the weights, b) then perform the weighted fit, and finally, c) revert the inversion.
+A standard approach to perform this fit is: a) to invert the variables before applying the weights, b) then perform the weighted fit and, finally, c) revert the inversion.
 This method is deterministic, but the approximate solution worsens with smaller coefficients of determination.
-To overcome this limitation, we developed a stochastic method by using a bootstrapping-like strategy that avoided inversion and could be applied regardless of the coefficient of determination.
+To overcome this limitation, we developed a stochastic method using a bootstrapping-like strategy that avoided inversion and was applicable regardless of the coefficient of determination.
 

### 2017-02-21 mSystems/rev2/mSys_discussion_rev2.tex
mSys: several figs&text changes which lead to REV2 DRAFT delta
On branch master changes to be committed:
	modified:   Response to Reviewer Comments.docx
	modified:   figs/Fig1.eps
	modified:   figs/Fig4.eps
	modified:   figs/Fig5.eps
	new file:   figs/HostLifeStyle_StoolA_day72to122_abroad_TAILLESS_Rank.pdf
	new file:   figs/HostLifeStyle_h_StoolA_day0to70_before_TAILLESS_Rank.pdf
	new file:   figs/HostLifeStyle_h_StoolA_day123to256_returned_TAILLESS_Rank.pdf
	new file:   figs/HostLifeStyle_h_StoolA_day257to364_after_TAILLESS_Rank.pdf
	modified:   figs/_ed_Fig1.pdf
	modified:   figs/_ed_supfig

@@
-One of the mains features of this work is to have demonstrated that, independently of its condition, microbiota follows Taylor's law.
-We have seen that the value of the scaling index in each case is always less than the unity (using standard deviation as the measurement for dispersion), which provides us with information about the community structure.
+% !TeX root = ./mSys_MAIN_rev2.tex
+One of the highlights of this study is that is shows, independently of its condition, that microbiota follows Taylor’s law.
+We have seen that in each case the value of the scaling index is always less than the unity (using standard deviation as the dispersion measurement), which provides us with information about the community structure.
 This means that, in relative terms, the most abundant elements in the population are less volatile to perturbations than the less abundant ones.
 The explanation for this universal pattern is not clear although some hypotheses have been tested in other studies, such as the presence of negative interactions in the population \cite{kilpatrick}, and a demonstration that this may depend on reproductive correlation \cite{ballantyne}.
-Nevertheless, none of these explanations are sufficient when we are talking about microbiota, as the reproduction term is diffuse, the interactions between its components are not only based on competition \cite{joao, mehta, bucci}, and that even that kind of negative interaction may not effectively yield values less than the unity when referring to a bacterial species \cite{ramslayer}.
-Anyhow, the values obtained in all cases were very similar from one to another, which could suggest that the community structure is preserved throughout the different scenarios we studied.
+Nevertheless, none of these explanations are sufficient when it comes to microbiota, as the term reproduction is diffuse the interactions between its components are not only based on competition \cite{joao, mehta, bucci}.
+Moreover, even such negative interaction may not effectively yield values less than the unity when referring to a bacterial species (41).
+Nonetheless, the values obtained in all cases were very similar to each other, which may suggest that the community structure is preserved throughout the different scenarios studied herein.
 
@@
 It is a direct estimator of the stability of the system under study.
-As we have shown above, the healthy subset of each study has lower variability than the non--healthy subset, when dealing with adult individuals.
-Interestingly, the variability parameter was higher in the healthy subset in the study of the discordant twins suffering from kwashiorkor disease \cite{kwashiorkor}.
-In this regard, it has been shown that infant microbiota needs to develop toward a definite, adult state \cite{koenig}.
-This implies that temporal variability is greater in children compared to a healthy adult state, which should be temporally stable.
-Thus, our results could point to the need for this variability in order to reach that adult state.
-Furthermore, as we wanted to see how this variability behaved over time, we calculated the evolution of this parameter for the samples which had enough time sampling.
+As we have shown above, the healthy subset of each study has lower variability than the non-healthy subset, when dealing with adult subjects.
+Interestingly, the variability parameter was higher in the healthy subset in the study of discordant twins suffering from kwashiorkor disease \cite{kwashiorkor}.
+In this respect, research has shown that infant microbiota needs to develop toward a definite, adult state \cite{koenig}.
+This implies that temporal variability is greater in children than in a healthy adult, which should be temporally stable.
+Thus, our results could indicate this variability is necessary in order to reach that adult state.
+Furthermore, as we wanted to see how this variability shifted over time, we calculated the changes in this parameter for the samples which had enough time sampling.
 As shown in Figure \ref{fig:tempevo1}, the variability of microbiota fluctuated over time.
-It is interesting to note in Figure \ref{fig:tempevo2} how this parameter captured the two antibiotic intakes in one of the patients from the study by Dethlefsen and Relman \cite{antibiotic}, especially in that there seems to be some kind of a resilience process in the microbiota due to the lower variability increase in the second antibiotic intake.
+Interestingly, Figure \ref{fig:tempevo2} shows how this parameter reflected the two antibiotic intakes in one of the patients in the study by Dethlefsen and Relman \cite{antibiotic} particularly the apparent resilience of the microbiota due to the reduced increase in variability during the second antibiotic intake.
 
-The primary hypothesis of this work is that, in adults, having a healthy microbiota means that the population is stable over time and does not move into a state where it is highly susceptible to external or internal perturbations, causing a dysbiotic state in the microbiota.
-In order to use the valuable information provided by the empirical law of Taylor's work, we proposed the use of Langevin's equation to model how ranking stability evolved over time.
+The primary hypothesis of this work is that, in adults, having a healthy microbiota means that the microbial population is stable over time.
+This stability means the microbiota does not shift and become susceptible to external or internal perturbations causing dysbiosis.
+In order to use the valuable information provided by the empirical law of Taylor’s work, herein we have proposed the use of Langevin’s equation to model how stability ranking changed  over time.
 While the system noise component can be directly measured as its variability, the other main term needs to be inferred from the model.
-This term, which we named "fitness", is the one that enables the system to be stable in the face of potential perturbations.
-In ecological terms, this could represent the nature of interactions that are present among bacteria, between bacteria and other minority populations, such as fungi or archaea, between bacteria and the viral component in microbiota, and interactions between the host and the whole microbiota.
-As this is a first step to model the temporal stability of microbiota, and given its complex nature, we calculated fitness using the Fluctuation Dissipation Theorem as a first approximation\cite{FD}.
-Thus, the fitness of microbiota will still need to be modeled in future works in order to make the model more accurate and give it a higher predictive power.
+This term, which we have named "fitness", enables the system to remain stable when confronted with potential perturbations.
+In ecological terms, this could represent the nature of interactions present among bacteria, between bacteria and other minority populations, such as fungi or archaea, between bacteria and the viral component in microbiota, and interactions between the host and the whole microbiota.
+As this is a first step to model the temporal stability of microbiota, and given its complex nature, we calculated fitness using the Fluctuation Dissipation Theorem as a first approximation \cite{FD}.
+Thus, future works are required to model the fitness of microbiota in order to provide a more accurate model with higher predictive power.
 
-By solving Langevin's differential equation, we can obtain a phase diagram where each microbiota sample can be placed according to its fitness and variability into one of two phases, according to the ranking stability of the system.
-As we can see in the phase--space in Figure \ref{fig:main3}, three different conditions that could occur are shown.
-First, we could have a healthy microbiota with some fluctuations, as shown by one of the subjects of Caporaso \emph{et al.}'s study \cite{moving}.
+By solving Langevin’s differential equation, we obtain a phase diagram where each microbiota sample can be placed by its fitness and variability into one of two phases, according to the stability ranking of the system.
+As shown by the phase--space in Figure \ref{fig:main3}, three different conditions can occur.
+The first is a healthy microbiota with some fluctuations, as shown by one of the subjects in Caporaso \emph{et al.}'s study \cite{moving}.
 Because this case would have good fitness, its temporal variability would not place the microbiota in the unstable phase of the diagram.
-Second, we have a subject from the study by Dethlefsen and Relman \cite{antibiotic} who was perturbed twice by an antibiotic intake.
-His microbiota underwent sufficient change so as to lose its stability, and hence be placed in the unstable part.
-In this location, it is more sensitive to potential perturbations such as, for example, opportunist infections.
+Secondly, we have a subject from the study by Dethlefsen and Relman \cite{antibiotic} whose microbiota was perturbed twice by an antibiotic intake, undergoing sufficient change so as to lose its stability, and hence be placed in the unstable part.
+In this location, it is more sensitive to potential perturbations such as, for example, opportunistic infections.
 In the third and last condition, the subject was already in the unstable phase due to a health issue, i.e.
 IBS.
-This can be observed in one of the patients from Durban \emph{et al.}'s study \cite{IBS}.
-In addition, it was shown that this subject's health status improved during the time the experiment was carried out, implying that his microbiota also recovered the stability it had lost.
-It is interesting to note that in the subject's health from the study made by David \emph{et al.} \cite{hostlife} who suffered a Salmonella infection during the experiment, there was a significant shift in variability and a final recovery from the perturbed state (see Supplementary Figure S\ref{supfig:HLS_xWSummary}).
+This can be observed in one of the patients in Durban \emph{et al.} \cite{IBS}.
+In addition, it was shown that this subject’s health status improved during the experiment, implying that his/her microbiota also recovered stability.
+Interestingly, in the study made by David \emph{et al.} \cite{hostlife} the subject who had a Salmonella infection during the experiment underwent a significant shift in variability with eventual recovery from the perturbed state (see Supplementary Figure S\ref{supfig:HLS_xWSummary}).
 
@@
 
-Specifically, the analysis of the rank stability of different periods of time belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife}, suggests that the presence of \emph{rank stability islands} among medium-ranked taxa is an interesting feature.
-Interestingly, this stability was compromised when the period was not an ordinary one, suggesting that those taxa were sensitive to changes in the lifestyle.
+Specifically, in the host lifestyle study \cite{hostlife}, the presence of \emph{rank stability islands} among medium--ranked taxa is an interesting feature revealed by the analysis of rank stability at different time periods in subject \emph{A}.
+Interestingly, this stability was compromised when the period was not an ordinary one, suggesting that those taxa were sensitive to changes in lifestyle.
 Among the genera identified as \emph{rank stability islands}, \emph{Lachnobacterium} and \emph{Clostridium} were catalogued as genera predictive of dysbiosis in the work of Larsen and Dai \cite{rsi_dysbiosis}, which analysed the same dataset \cite{hostlife}.
-Furthermore, very recently, it has been confirmed a clear relationship between \emph{Actinomyces} and conventional adenoma \cite{rsi_actino}, one of the two main precursors of the colorectal cancer.
-Finally,\emph{Eggerthella} is an opportunistic pathogen that is often associated with serious gastrointestinal pathology \cite{rsi_egg}.
+Furthermore, research has recently confirmed a clear relationship between \emph{Actinomyces} and conventional adenoma \cite{rsi_actino}, one of the two main precursors of the colorectal cancer.
+Finally, \emph{Eggerthella} is an opportunistic pathogen that is often associated with serious gastrointestinal pathology \cite{rsi_egg}.
 
-It could be brought into question the role of these taxa as key players in the phase transition of the microbiota, or whether they are more susceptible to perturbations than the most abundant.
-The types of interactions that could sustain this particular behavior are not clear, as these non-abundant taxa are not usually included in dynamic studies in order to obtain a community matrix.
+One might  question the role of these taxa as key players in the phase transition of the microbiota and wonder whether they are more susceptible to perturbations than the most abundant taxa.
+The types of interactions that could sustain this particular behavior are unclear, as these non-abundant taxa are usually excluded from dynamic studies to obtain a community matrix.
 Further experiments and data analysis are needed to clarify whether \emph{rank stability islands} are a widespread feature of microbiotas and whether they appear at lower taxonomic levels too.
 
-However, we have to be aware that the hypothesis above is too simplistic to be directly related to reality.
-It has been demonstrated that the situation is more complex than the outlook provided which separate healthy people from non-healthy people just by compositional terms, as Moya and Ferrer underlined in their recent review \cite{Moya_trends}.
-There are several different feasible scenarios in which we can consider microbiota as being stablem irrespective of their compositional evolution over time.
-For example, depending on its ability to recover its initial composition (resilience), or whether it can recover its original function despite its composition (functional redundancy).
+Notwithstanding the above, we should be aware that the above hypothesis is too simplistic to directly apply to reality.
+Indeed, the situation is more complex than the idea that healthy people can be distinguished from non-healthy people in solely compositional terms, as highlighted by Moya and Ferrer in their recent review \cite{Moya_trends}.
+There are several feasible scenarios in which we can consider microbiota to be stable, irrespective of its compositional shifts over time.
+For example, it may depend on the ability of the microbiota to recover its initial composition (resilience), or its ability to recover its original function despite its composition (functional redundancy).
 What we have shown in this work could be explained as the transition of stable microbiota into a state of dysbiosis.
 
-As a first step towards understanding microbiota stability, the model presents some limitations and there is still work to do.
+This is a first step towards understanding microbiota stability, although the model presents some limitations and thus further research is required.
 From a biological perspective, many questions arise from this work.
-We have observed the same pattern in Taylor's parameters in all the different conditions we studied, but a pertinent question is whether this is really a universal feature in the huge diversity of microbial niches.
-Furthermore, another relevant question centers on which mechanisms are involved in maintaining the population structure.
-The nature of the interactions between the elements of community is surely of great importance in this matter, and this is related to the fitness of the community, as mentioned above.
-How we should address community fitness is not clear, but works such as the one by Tikhonov \cite{tikhonov} could point us in the right direction to unravel the complexity of microbiota.
+We have observed the same pattern in Taylor’s parameters under all the conditions studied, but a pertinent question is whether this is really a universal feature in the hugely diverse microbial niches.
+Furthermore, another relevant question relates to mechanisms involved in maintaining the population structure.
+Undoubtedly, the nature of the interactions between community components has a great bearing on this issue, and this is related to community fitness, as mentioned above.
+How we should address community fitness remains unclear, but studies like the one by Tikhonov  \cite{tikhonov} could point us in the right direction and help us to unravel the complexity of microbiota and its relationship to host health.

### 2017-02-21 mSystems/rev2/mSys_intro_rev2.tex
mSys LaTeX files updated to ACCEPTED version (and beyond...)
"Beyond" means that formulae have been improved, typos corrected
and LaTeX magic comments added.

On branch master, changes committed:
	new file:   rev2/mSys_MAIN_rev2.dvi
	modified:   rev2/mSys_MAIN_rev2.pdf
	modified:   rev2/mSys_MAIN_rev2.tex
	modified:   rev2/mSys_MandM_rev2.tex
	modified:   rev2/mSys_discussion_rev2.tex
	modified:   rev2/mSys_intro_rev2.tex
	modified:   rev2/mSys_results_rev2.tex

--- 2017-01-30 mSystems/rev2/mSys_intro_rev2.tex
@@
-The desire to understand the factors that influence human health and cause disease has always been one of the major driving forces of biological research.
-As evidence of the new "holobiont"  and "hologenome" concepts is increasing each day \cite{holo1, holo2}, research not only focuses on the human physiology but also on the microbial population that surrounds us.
-However, these concepts are still under debate \cite{holo3}.
-The human microbiome is intimately linked to our physiology with the metabolism of bile acids \cite{bileacids}, of choline \cite{choline} and key metabolites like short-chain fatty acids \cite{scfa1, scfa2} which are also involved in immune system maturation \cite{scfa3, scfa4}.
-Human microbiota has been suggested to be closely related to diseases like type 2 diabetes \cite{diabetes2}, cardiovascular disease (CVD) \cite{CVD}, irritable bowel syndrome \cite{IBS}, Crohn's disease \cite{CD}, some afflictions like obesity \cite{ob1, ob2} and malnutrition \cite{nutr} as well as many other diseases \cite{Moya_trends}.
+% !TeX root = ./mSys_MAIN_rev2.tex
+The quest to understand the factors that influence human health and cause disease has always been one of the major driving forces of biological research.
+With growing evidence of the new "holobiont"  and "hologenome" concepts \cite{holo1, holo2}, research not only focuses on the human physiology but also on the associated microbial population, although these concepts are still under debate \cite{holo3}.
+Research has revealed that the human microbiome is intimately linked to our physiology through the metabolism of bile acids \cite{bileacids}, of choline \cite{choline} and key metabolites like short-chain fatty acids \cite{scfa1, scfa2}, which are also involved in immune system maturation \cite{scfa3, scfa4}.
+Human microbiota is plausibly related to diseases such as type 2 diabetes \cite{diabetes2}, cardiovascular disease (CVD) \cite{CVD}, irritable bowel syndrome \cite{IBS}, Crohn's disease \cite{CD}, some afflictions like obesity \cite{ob1, ob2} and malnutrition \cite{nutr}, as well as many other diseases \cite{Moya_trends}.
 Recent studies have revealed that microbes also influence brain function and behaviour and are related to neurological disorders like Alzheimer's disease through the gut-brain axis\cite{mind,AD}.
-Recently, even a subtle condition like chronic fatigue syndrome, which has often been suggested to be a psychosomatic disease, has been associated with reduced diversity and altered composition of the gut microbiome\cite{CFS}.
+Recently, chronic fatigue syndrome, a subtle condition often cited as a psychosomatic disease, has been associated with a reduced diversity and altered composition of the gut microbiome\cite{CFS}.
 
-High throughput methods for microbial 16S ribosomal RNA gene and SMS (shotgun metagenomic sequencing) have now begun to reveal the composition of archaeal, bacterial, fungal and viral communities located both, in and on the human body.
+This research area has progressed greatly thanks to high-throughput methods for microbial 16S ribosomal RNA gene and SMS (shotgun metagenomic sequencing), which reveal the composition of archaeal, bacterial, fungal and viral communities located in and on the human body.
 Modern high-throughput sequencing and bioinformatics tools provide a powerful means of understanding how the human microbiome contributes to health and its potential as a target for therapeutic interventions \cite{microb&health}.
-To define normal host-gut microbe interactions and how microbiota compositional changes can cause some diseases are important issues that still require scientific answers \cite{normal1, normal2, panthropology}.
+Research is underway to establish normal host-gut microbe interactions and understand how microbiota compositional changes can cause certain diseases  \cite{normal1, normal2, panthropology}.
 
-Biology has recently acquired new technological and conceptual tools to investigate, model and understand living organisms at system level, thanks to spectacular progress in quantitative techniques, large-scale measurement methods and the integration of experimental and computational approaches.
-In particular, Systems Biology has made great efforts to reveal the general laws governing the complex behaviour of microbial communities \cite{sysbio&microb, msys1, metasysbio}, including a proposal suggesting they have universal dynamics \cite{uni_dynam}.
-Microbiota can be approached under the light of ecological theory which includes general principles like Taylor's law \cite{pretaylor,taylor} that relates the spatial or temporal variability of the population with its mean.
-This law, also known as fluctuation scale law, is ubiquitous in the natural world and can be found in several systems like random walks \cite{randomwalks}, stock markets \cite{economics1, economics2}, tree \cite{cohen_taylor} and animal populations \cite{taylor, animal1, animal2}, gene expression \cite{genexpress}, and the human genome \cite{genome}.
-Taylor's law has been applied to microbiota in a spatial way in the work of Zhang {\it et al.}, (2014) \cite{isme1}, where they show that this population tends to be an aggregated one rather than having a random distribution.
-Despite its ubiquity, it has only been studied in experimental settings \cite{cohen_bac, ramslayer} but has never been applied in follow-up studies on microbiota, even though major efforts have been made to infer the community structure from a dynamic point of view \cite{cobas, schloss, ravel}
+Biology has recently acquired new technological and conceptual tools to investigate, model and understand living organisms at a systems level, thanks to progress in quantitative techniques, large-scale measurement methods and joint experimental and computational approaches.
+In particular, Systems Biology strives to reveal the general laws governing the complex behavior of microbial communities \cite{sysbio&microb, msys1, metasysbio}, including a proposal for universal dynamics \cite{uni_dynam}.
+Microbiota can be approached in the light of ecological theory, which includes general principles like Taylor’s law \cite{pretaylor,taylor} relating the spatial or temporal variability of the population with its mean.
+This law, also known as fluctuation scale law, is ubiquitous in the natural world and can be found in several systems such as random walks \cite{randomwalks}, stock markets \cite{economics1, economics2}, tree \cite{cohen_taylor} and animal populations \cite{taylor, animal1, animal2}, gene expression \cite{genexpress}, and the human genome \cite{genome}.
+Taylor's law has been applied to microbiota spatially by Zhang {\it et al.}, (2014) \cite{isme1}, with results showing that this population tends to be an aggregated one rather than having a random distribution.
+Despite its ubiquity, this law has only been tested in experimental settings \cite{cohen_bac, ramslayer} but has never been applied in follow-up studies on microbiota, despite major efforts to infer the community structure from a dynamic point of view \cite{cobas, schloss, ravel}
 
-This paper presents the imprints of health status (healthy or disease) in the macroscopic properties of microbiota, by studying its temporal variability.
-We analyzed more than 40000 time series of taxa from the gut microbiome of 99 individuals obtained from publicly available high throughput sequencing data about different conditions: diseases, diets, trips, obese status, antibiotic therapy and healthy individuals.
-Having seen that all the cases followed Taylor's law, we used this empirical fact to model how the relative abundances of taxa evolved over time thanks to the Langevin equation, in a similar way as the approach applied recently by Blumm {\it et al.} \cite{ranking}.
-We used this mathematical framework to explore the temporal stability of microbiota under different conditions in order to understand how this is related to the healthy status of the subjects.
+This paper presents the hallmarks of health status (healthy or diseased) in the macroscopic properties of microbiota, by studying its temporal variability.
+We analyzed over 40,000 time series of taxa from the gut microbiome of 99 subjects obtained from publicly available high-throughput sequencing data related to different conditions: diseases, diets, trips, obese status, antibiotic therapy and healthy subjects.
+On finding that all the cases followed Taylor’s law, we used this empirical fact to model how the relative abundances of taxa evolved over time using the Langevin equation, similarly to the approach applied by Blumm {\it et al.} \cite{ranking}.
+We used this mathematical framework to explore the temporal stability of microbiota under different conditions in order to understand how this is related with the health status of the subjects.

### 2017-02-21 mSystems/rev2/mSys_results_rev2.tex
mSys LaTeX files updated to ACCEPTED version (and beyond...)
"Beyond" means that formulae have been improved, typos corrected
and LaTeX magic comments added.

On branch master, changes committed:
	new file:   rev2/mSys_MAIN_rev2.dvi
	modified:   rev2/mSys_MAIN_rev2.pdf
	modified:   rev2/mSys_MAIN_rev2.tex
	modified:   rev2/mSys_MandM_rev2.tex
	modified:   rev2/mSys_discussion_rev2.tex
	modified:   rev2/mSys_intro_rev2.tex
	modified:   rev2/mSys_results_rev2.tex

--- 2017-01-30 mSystems/rev2/mSys_results_rev2.tex
@@
-The microbiome temporal variability was analyzed to extract the global properties of the system.
+% !TeX root = ./mSys_MAIN_rev2.tex
+Microbiome temporal variability was analyzed to extract the global properties of the system.
 As fluctuations in total counts are plagued by systematic errors we worked on the temporal variability of relative abundances for each taxon.
-Our first finding was that, in all cases, changes in the relative abundances of taxa followed a ubiquitous pattern, known as the fluctuation scaling law\cite{fs} or Taylor's power law\cite{taylor}, i.e., the microbiota of all detected taxa followed $\sigma_i  = V\cdot x_i^{\beta}$, a power law dependence between the mean relative abundance $x_i$ and the dispersion $\sigma_i$.
+Our first finding was, without exception, that changes in the relative abundances of taxa followed a ubiquitous pattern, known as the fluctuation scaling law \cite{fs} or Taylor's power law \cite{taylor}.
+In other words, the microbiota of all detected taxa followed $\sigma_i  = V\cdot x_i^{\beta}$, a power law dependence between the mean relative abundance $x_i$ and the dispersion $\sigma_i$.
 The law seem to be ubiquitous, spanning even to six orders of magnitude in the observed relative abundances.
-As shown in Figure \ref{fig:main1}, where $V$ corresponds to the y-intercept and $\beta$ is the slope of the fit, the most abundant species were less volatile in relative terms than those which were less abundant.
-The fitting to the power law was always robust ($R^{2}$ > 0.88) and did not depend on the microbiome condition.
-The power law (or scaling) index $\beta$ and the variability $V$ (hereafter Taylor's parameters) appear to be correlated with the stability of the community.
-For this, we assume that Taylor's parameters behave as proxies for the stability.
-On the one hand, $\beta$ is a scaling index that gave us information about the statistical properties of the ecosystem.
+As shown in Figure \ref{fig:main1}, where $V$ corresponds to the y-intercept and $\beta$ is the slope of the fit, the most abundant species were less volatile in relative terms than the less abundant ones.
+The fit to the power law was always robust ($R^{2}$ > 0.88) and did not depend on microbiome condition.
+The power law (or scaling) index $\beta$ and the variability $V$ (hereafter Taylor's parameters) appear to be correlated with community stability.
+Accordingly, we assume that Taylor's parameters behave as proxies for stability.
+On the one hand, $\beta$ is a scaling index that provides information about the statistical properties of the ecosystem.
 If it is $1/2$, the system behaves like a Poisson distribution.
-If $\beta$ is 1, the system behaves as an exponential distribution.
-Generally speaking, metagenomes vary with time with $\beta$ between these two universal classes.
-In our case, the fact that $\beta$ was less than 1 tells us that the most abundant taxa in the microbial community were less susceptible to any perturbation than the other taxa.
-On the other hand, the variability $V$ was a direct estimator of the amplitude of fluctuations over time.
-$V$ represents the maximum variability attainable by a hypothetical dominant genus (with relative abundance close to 1).
+If $\beta$ is 1, the system behaves like an exponential distribution.
+Generally speaking, metagenomes undergo time-course variations with $\beta$ between these two universal classes.
+In our study, the fact that $\beta$ was less than 1 indicates that the most abundant taxa in the microbial community were less susceptible to perturbations than the other taxa.
+On the other hand, the variability $V$ is a direct estimator of the amplitude of fluctuations over time.
+$V$ represents the maximum variability attainable by a hypothetically dominant genus (with relative abundance close to 1).
 It is an important parameter that characterizes the type of system.
-If $V$ is small the ranking is stable, similar to the number of diagnoses of a particular disease recorded in Medicare  during a month \cite{medicare}.
-If $V$ is large, as it seems to be the case of metagenomic samples, the ranking might be unstable, like the number of hourly page views of articles in Wikipedia \cite{ranking,fs}.
-The Taylor parameters were related to the health status of the host, which we consider as constituting the main finding contributed by this article.
+If $V$ is small the ranking is stable.
+As way of example, this would be the case for the number of diagnoses of a particular disease recorded in Medicare during a month \cite{medicare}.
+If $V$ is large, as occurs for metagenomic samples, the ranking might be unstable, as it would be for the number of hourly page views of articles in Wikipedia \cite{ranking,fs}.
+Interestingly, the Taylor parameters were related to the health status of the host, which is the main finding of this study.
 
@@
 
-Taylor's parameters describing the temporal variability of the gut microbiome in our sampled individuals are shown in Supplementary Tables S\ref{tab:diet} to S\ref{tab:LEA}.
-Our results hint at ubiquitous behavior.
+Taylor's parameters describing the temporal variability of the gut microbiome in our sampled subjects are shown in Supplementary Tables S\ref{tab:diet} to S\ref{tab:LEA}.
+Our results are indicative of ubiquitous behavior.
 Firstly, the variability (which corresponds to the maximum amplitude of fluctuations) was large, which suggests the resilient capacity of the microbiota.
 Secondly, the scaling index was always smaller than one, which means that more abundant taxa were less volatile than less abundant ones.
-In addition, Taylor's parameters for the microbiome of healthy individuals in different studies were compatible with estimated errors.
+In addition, Taylor's parameters for the microbiome of healthy individuals in different studies  \cite{IBS,moving,antibiotic,LEA,kwashiorkor,diet,hostlife} were compatible with estimated errors.
 This enabled us to define an area in the Taylor parameter space that we called the \emph{healthy zone}.
 
-In order to jointly visualize and compare the results of individuals from different studies\cite{IBS,moving,antibiotic,LEA,kwashiorkor,diet,hostlife}, their Taylor parameters were standardized, with standardization meaning that each parameter was subtracted by the mean value and divided by the standard deviation of the group of healthy individuals for every single study independently.
-Due to the different systematics in each study, we defined a healthy region for each one of them, standardized to mean zero and variance one and computed mean and variance of ?unhealthy?
-with this standardization (for details of the procedure, please see Standardization subsection in Material and Methods).
-Therefore, different studies were isolated so that individuals from a given study did not affect the results on the ?unhealthy?
-individuals of the other studies.
-We think this statistical approach was safer, as we avoided to combine data with very different systematic errors.
-The healthy zone and the standardized Taylor parameters for individuals whose gut microbiota was compromised (i.e., they were suffering from IBS, kwashiorkor, altered diet, intake of antibiotics, a Salmonella infection, or had gone on a trip abroad) are shown in Figure \ref{fig:main2}.
-The variability in children developing kwashiorkor was smaller than that of their healthy twins.
+In order to jointly visualize and compare the results of subjects from the above-mentioned studies \cite{IBS,moving,antibiotic,LEA,kwashiorkor,diet,hostlife}, their Taylor parameters were standardized, with standardization meaning that each parameter was subtracted by the mean value and divided by the standard deviation of the group of healthy subjects for every study independently.
+Due to the different systematics in each study, we defined a healthy region for each of them, standardized to mean zero and variance one, then we computed mean and variance of “unhealthy” with this standardization (for details of the procedure, please see Standardization subsection in Material and Methods).
+Therefore, different studies were isolated so that subjects from a given study did not affect the results for the “unhealthy” subjects of the other studies.
+We think this statistical approach was safer, as we avoided combining data with very different systematic errors.
+The healthy zone and the standardized Taylor parameters for the temporal variability of the gut microbiome in subjects whose gut microbiota was compromised (i.e., they were suffering from IBS, kwashiorkor, altered diet, antibiotic intake, a Salmonella infection, or had gone on a trip abroad) are shown in Figure \ref{fig:main2}.
+The variability in children with kwashiorkor was smaller than that of their healthy twins.
 A meat/fish-based diet significantly increased variability when compared to a plant-based diet.
-All other cases presented increased variability, which was particularly severe and statistically significant at over 95\% confidence level (CL), for grade III obese patients on a diet, individuals taking antibiotics, the subject who had a Salmonella infection, the subject who did a travel abroad or the IBS-diagnosed patients.
-One global property emerged from all the worldwide data collected: the Taylor's parameters characterized the statistical behavior of microbiome changes.
-Furthermore, we verified that our conclusions were robust to systematic errors as a result of taxonomic assignment (see Taxa level selection in Material and Methods).
+All other cases presented increased variability, which was particularly severe and statistically significant at over 95\% confidence level (CL), for grade III obese patients on a diet, subjects taking antibiotics, the subject who had a Salmonella infection, the subject who had traveled abroad and the IBS--diagnosed patients.
+One global property emerged from these comprehensive data: the Taylor’s parameters characterized the statistical behavior of microbiome changes.
+Furthermore, we verified that our conclusions were robust to systematic errors resulting from taxonomic assignment (see Taxa-level selection in Material and Methods).
 
@@
 All the data studied in this work were compiled here.
-The colored circle corresponds to a 68\% confidence level (CL) region of healthy individuals in the Taylor's parameter space, while the dashed line delimites the 98\% CL region.
-Points with errors place gut microbiome in the Taylor's parameter space, for each individual whose microbiota was compromised.
+The colored circle corresponds to a 68\% confidence level (CL) region of healthy subjects in the Taylor's parameter space, while the dashed line delimits the 98\% CL region.
+Points with errors place gut microbiome in the Taylor's parameter space, for each subject whose microbiota was compromised.
 It should be noted that the parameters were standardized (standard deviation units) to the healthy group in each study for every single study independently, for demonstrative and comparative purposes.}
@@
 
-Taylor's power law has been explained in terms of various effects, though none have brought a general consensus.
-It can be shown to have its origin in mathematical convergence which is similar to the central limit theorem, and thus virtually any statistical model designed to produce a Taylor law converges to a Tweedie distribution\cite{stat}, providing a mechanistic explanation based on the statistical theory of errors\cite{convergence1,convergence2,convergence3}.
+Taylor’s power law has been explained in terms of various effects, though without a general consensus.
+It has its origin in mathematical convergence, which is similar to the central limit theorem, and thus virtually any statistical model designed to produce a Taylor law converges to a Tweedie distribution \cite{stat}, providing a mechanistic explanation based on the statistical theory of errors\cite{convergence1,convergence2,convergence3}.
 To reveal the generic mechanisms that drive different scenarios in the $\beta-V$ space, we modeled the system by assuming that taxon relative abundance followed a Langevin equation with, on the one hand, a deterministic term that captured the fitness of each taxon and, on the other hand, a randomness term associated with Gaussian random noise\cite{ranking}.Both terms were modeled by power laws, with coefficients that can be interpreted as the taxon fitness $F_i$ and the variability $V$ (see Model in Material and Methods).
-Fitness $F_i$ captures the time scale that the system needs to reach equilibrium (the size of variability $V$ may or may not allow to reach it).
+Fitness $F_i$ captures the time scale that the system needs to reach equilibrium (the size of variability $V$ may or may not allow equilibrium to be reached).
 $F_i$ has dimensions of 1/time and roughly corresponds to the half-life of the system when decaying to the stable state.
@@
 Differences in the variability $V$ can induce a noise-induced phase transition in the relative abundances of taxa.
-The temporal evolution of the probability of a taxon having the abundance $x_i$, given its fitness, is governed by the Fokker--Planck equation.
+The Fokker-Planck equation governs the temporal changes in the likelihood that a given taxon has the abundance $x_i$, given its fitness.
 The results of solving this equation show that stability is best captured by a phase space determined by the fitness $F$ and the amplitude of fluctuations $V$ (see Figure \ref{fig:main3}).
@@
 The light-blue shaded region corresponds to the stable phase, while the grey shaded region is the unstable phase (the phase transition line is calculated for  $\alpha$ = $\beta$ = 0.75).
-We placed healthy individuals (green) and individuals whose gut microbiota is threatened (antibiotics, IBS) in the phase space fitness--variability.
-The gut microbiota of healthy individuals over a long term span show a quasi--periodical variability (central period is ten days).
+We placed healthy subjects (green) and subjects whose gut microbiota is threatened (antibiotics, IBS) in the phase space fitness--variability.
+The gut microbiota of healthy subjects over a long term span show a quasi--periodical variability (central period is ten days).
 We show that taking antibiotics (AB1 and AB2 correspond to the first and second treatment, respectively) induces a phase transition in gut microbiota, which impacts on future changes.
@@
 Any disturbance modifies the composition of the microbiota, with different compositions encoding different $F$ and $V$.
-Our model can be solved analytically, which allows to a simple understanding of the different regimes, and in particular, to calculate the formula of the transition region.
-If $V$ is sufficiently small compared to $F$, there is a maximum in the likelihood in the physical region (relative compositions larger than zero and smaller than one), i.e., there is a best composition solution of the differential equation, which is the ordered solution.
-On the contrary, if $V$ is sufficiently large compared to $F$ the maximum in the likelihood is outside the physical region, i.e., the best composition solution of the differential equation is at the boundaries (either zero or one) and all physical solutions have comparable likelihoods, which is the noisy phase.
-The microbiome of healthy individuals was found to be in the stable phase, while the microbiome of several other individuals was shown to be in the unstable phase.
-In particular, individuals taking antibiotics and the IBS--diagnosed patient \emph{P2} had the most severe symptoms.
+We have shown that effective perturbations significantly change V and lead the microbiota to a transition from the ordered phase to the noise--induced one.
+Our model can be solved analytically, which allows for a simple understanding of the different regimes and, in particular, to calculate the formula of the transition region.
+The order parameter is the composition $x_M$ that maximizes the probability distribution: $0<x_M<1$ defines the ordered phase, while $x_M>1$ defines the disordered phase.
+If $V$ is sufficiently small compared to $F$, the likelihood peaks in the physical region (relative compositions larger than zero and smaller than one), i.e., there is a best composition solution of the differential equation, which is the ordered solution.
+Conversely, if $V$ is sufficiently large compared to $F$, the likelihood peaks outside the physical region, i.e., the best composition solution of the differential equation is at the boundaries (either zero or one) and all physical solutions have comparable likelihoods, i.e., the noise--induced phase.
+The microbiome of healthy subjects was found to be in the stable phase, while the microbiome of several other subjects was in the unstable phase.
+In particular, subjects taking antibiotics and the IBS–-diagnosed patient \emph{P2} had the most severe symptoms.
 In this phase diagram, each microbiota state is represented by a point at its measured variability $V$ and inferred fitness $F$.
@@
 
-The rank dynamics and stability plots in Figure \ref{fig:corrank_HLS_abroad} and \ref{fig:corrank_HLS_returned} show the variations in rank over time for the most dominant taxa and their calculated Rank Stability Index (RSI, as discussed in Material and Methods) for the gut microbiome taxa of a healthy subject, the individual \emph{A} in the host lifestyle study \cite{hostlife}.
-The Figure \ref{fig:corrank_HLS_abroad} covers the period when the individual is travelling abroad and the Figure \ref{fig:corrank_HLS_returned} covers the subsequent period.
-The taxa were listed ordered by their accumulated frequency over the time series, with the y-axis being the overall dominance axis for each sample set.
+The rank dynamics and stability plots in Figures \ref{fig:corrank_HLS_abroad} and \ref{fig:corrank_HLS_returned} show the variations in rank over time for the most dominant taxa and their calculated Rank Stability Index (RSI, as discussed in Material and Methods) for the gut microbiome taxa of a healthy subject, namely subject \emph{A} in the host lifestyle study \cite{hostlife}.
+The Figure \ref{fig:corrank_HLS_abroad} covers the period when the subject is travelling abroad and Figure \ref{fig:corrank_HLS_returned} covers the subsequent period.
+The taxa are listed according to their accumulated frequency over the time series, with the y-axis being the overall dominant axis for each sample set.
 Generally speaking, we observed that the most dominant taxa had the highest rank stability.
@@
 	\includegraphics[width=1.0\textwidth]{figs/Fig4.eps}
-	\caption{Rank variation over time for the 50 most dominant elements (taxa) and their calculated Rank Stability Index (as shown in Material and Methods) for a special period (days 72 to 122, travelling abroad) belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife}.}
+	\caption{Rank variation over time for the 50 most dominant elements (taxa) and their calculated RSI (Rank Stability Index), Rank Variability (RV) and Differences Variability (DV), as detailed in Rank stability and variability in Material and Methods, for a special period (days 72 to 122, travelling abroad) belonging to subject \emph{A} in the host lifestyle study \cite{hostlife}.}
 	\label{fig:corrank_HLS_abroad}
@@
 	\includegraphics[width=1.0\textwidth]{figs/Fig5.eps}
-	\caption{Rank variation over time for the 50 most dominant elements (taxa) and their calculated Rank Stability Index (as shown in Material and Methods) for an ordinary period (days 123 to 256, after the trip) belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife}.}
+	\caption{Rank variation over time for the 50 most dominant elements (taxa) and their calculated RSI (Rank Stability Index), Rank Variability (RV) and Differences Variability (DV), as detailed in Rank stability and variability in Material and Methods, for an ordinary period (days 123 to 256, after the trip) belonging to subject \emph{A} in the host lifestyle study \cite{hostlife}.}
 	\label{fig:corrank_HLS_returned}
@@
 
-For the trip abroad in Figure \ref{fig:corrank_HLS_abroad}, beyond the differences in dominance for the particular taxa, we still observed that the most dominant were the most rank stable.
-Moreover, the medium-ranked taxa were quite rank unstable, mostly due to transient (often one or two consecutive samples) yet dramatic drops in their relative abundance, which usually occurred more than twice during their time series.
+For the trip abroad in Figure \ref{fig:corrank_HLS_abroad}, beyond the differences in dominance of particular taxa, we observed that the most dominant were the also most rank--stable.
+Moreover, the medium--ranked taxa were quite rank unstable, mostly due to transient (often one or two consecutive samples) albeit dramatic declines in their relative abundance, which usually occurred more than twice during their time series.
 
-Nevertheless, in the particular case of the next period, the one subsequent to the trip, in Figure \ref{fig:corrank_HLS_returned}, some taxa showed higher stability than other more dominant taxa, forming a kind of \emph{rank stability islands} for medium--ranked taxa, since they show a moderately stable index (RSI roughly over 70\%).
-In particular, this is the case for the genera \emph{Actinomyces}, \emph{Leuconostoc}, \emph{Lachnobacterium}, \emph{Eggerthella}, \emph{Clostridium} and	\emph{Collinsella}.
-For those genera, both the overall rank and the RSI were clearly lower during the trip (RSI under 70\%).
-\emph{Actinomyces} and \emph{Lachnobacterium} are even not shown in Figure \ref{fig:corrank_HLS_abroad} because they sank to positions 56 and 77, respectively.
-On the contrary, \emph{Leuconostoc} was the least sensitive to the change of lifestyle.
-In addition, it is worth mentioning that \emph{Lachnobacterium} showed anti-correlation over time against the vast majority of the taxa classified in this study.
+Nevertheless, in the particular case of the next period (Figure \ref{fig:corrank_HLS_returned}), the one subsequent to the trip, some taxa showed higher stability than other more dominant taxa, forming \emph{rank stability islands} for medium--ranked taxadisplaying a moderately stable index (RSI roughly over 70\%).
+In particular, this was the case for the genera \emph{Actinomyces}, \emph{Leuconostoc}, \emph{Lachnobacterium}, \emph{Eggerthella}, \emph{Clostridium} and	\emph{Collinsella}.
+For the aforementioned genera, both the overall rank and the RSI were clearly lower during the trip (RSI under 70\%).
+\emph{Actinomyces} and \emph{Lachnobacterium} are not shown in Figure \ref{fig:corrank_HLS_abroad} because they sank to positions 56 and 77, respectively.
+By contrast, \emph{Leuconostoc} was the least sensitive to the lifestyle change.
+Interestingly, \emph{Lachnobacterium} showed anti-correlation over time compared to the vast majority of the taxa classified in this study.
 
-We also found those \emph{rank stability islands} for medium-ranked taxa in the other periods belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife} (see Supplementary Figure S\ref{supfig:corrank_HLS_before} and Supplementary Figure S\ref{supfig:corrank_HLS_after} for the corresponding rank plots).
+We also found those \emph{rank stability islands} for medium-ranked taxa in the other periods belonging to subject \emph{A} in the host lifestyle study \cite{hostlife} (see Supplementary Figure S\ref{supfig:corrank_HLS_before} and Supplementary Figure S\ref{supfig:corrank_HLS_after} for the corresponding rank plots).
 See Supplementary Table S\ref{tab:HLS_RSI} for details about the rank and RSI for the above-mentioned taxa over the different periods considered.
@@
 	\includegraphics[width=1.0\textwidth]{figs/supfig_corrank_HLS_StoolA_before.eps}
-	\caption{Rank variation over time for the 50 most dominant elements (taxa) and their calculated Rank Stability Index (as shown in Material and Methods) for an ordinary period (days 0 to 70, before the trip) belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife}.}
+	\caption{Rank variation over time for the 50 most dominant elements (taxa) and their calculated RSI (Rank Stability Index), Rank Variability (RV) and Differences Variability (DV), as detailed in Rank stability and variability in Material and Methods, for an ordinary period (days 0 to 70, before the trip) belonging to subject \emph{A} in the host lifestyle study \cite{hostlife}.}
 	\label{supfig:corrank_HLS_before}
@@
 	\includegraphics[width=1.0\textwidth]{figs/supfig_corrank_HLS_StoolA_after.eps}
-	\caption{Rank variation over time for the 50 most dominant elements (taxa) and their calculated Rank Stability Index (as shown in Material and Methods) for an ordinary period (days 257 to 364, further after the trip) belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife}.}
+	\caption{Rank variation over time for the 50 most dominant elements (taxa) and their calculated RSI (Rank Stability Index), Rank Variability (RV) and Differences Variability (DV), as detailed in Rank stability and variability in Material and Methods, for an ordinary period (days 257 to 364, further after the trip) belonging to subject \emph{A} in the host lifestyle study \cite{hostlife}.}
 	\label{supfig:corrank_HLS_after}
@@
 Finally, we studied the time dependence of the variability $V$ and power law index $\beta$ (see Model in Material and Methods) by using a sliding window approach.
-The total number of time points was divided into subsets of five points, where the following subset was defined by adding the next time sampling and eliminating the earliest one.
+The total number of time points was divided into subsets of five points, where the following subset was defined by adding the next sampling time and eliminating the earliest one.
 Both parameters were calculated for each subset against the average time lapse.
-Figure \ref{fig:tempevo1} shows the variability $V$ as a function of time for the two individuals in Caporaso's study\cite{moving} corresponding to the gut microbiota of a male (upper plot) and a female (lower plot).
+Figure \ref{fig:tempevo1} shows the variability $V$ as a function of time for the two subjects in Caporaso's study\cite{moving} corresponding to the gut microbiota of a male (upper plot) and a female (lower plot).
 Both samples showed changes in the variability $V$ with quasi--periodic behavior peaking at about 10 days.
@@
 
-Figure \ref{fig:tempevo2} shows the time evolution of $V$ for patient \emph{P2} in the IBS study\cite{IBS} (upper plot) and patient \emph{D} in the antibiotics study\cite{antibiotic} (lower plot).
+Figure \ref{fig:tempevo2} shows time--course changes in $V$ for patient \emph{P2} in the IBS study\cite{IBS} (upper plot) and patient \emph{D} in the antibiotics study\cite{antibiotic} (lower plot).
 The variability of the gut microbiota of \emph{P2} decreased from over 0.3 to below 0.2, showing a slow tendency to increase the order of the system.
@@
 	\includegraphics[width=0.99\textwidth]{figs/Fig6.eps}
-\caption{$V$ as a function of time for the two individuals in Caporaso's study\cite{moving}: samples of gut microbiome of a male (upper plot) and a female (lower plot).}
+\caption{$V$ as a function of time for the two subjects in Caporaso's study \cite{moving}: samples of gut microbiome of a male (upper plot) and a female (lower plot).}
 \label{fig:tempevo1}
